### 2018-12-21 thesis_sebastian_bek/content/0-Abstract.tex
Initial commit

\chapter*{Abstract}
\label{chap:Abstract}
%Sebastian
In den letzten Jahren erreichen uns immer häufiger Nachrichten über große Veranstaltungen, bei denen sich aufgrund zu hoher Besucherzahlen in Kombination mit unzureichender Planung Tumulte und Paniken, häufig mit Verletzten, ausbilden.
Solche Großveranstaltungen werden immer beliebter und sind fester Bestandteil der menschlichen Freizeitaktivitäten geworden, weswegen Einschränkungen Dieser unwahrscheinlich sind. Die Tendenz zeigt, das mit steigenden Besucherzahlen auch die Gefahr für Paniken und Unruhen steigt, weswegen videogestütztes Monitoring hier sinnvoll sein kann. Derzeit erfolgt das Monitoring von Veranstaltungen noch durch menschliche Betrachter, die oft unaufmerksam sind und meist nur subjektiv beurteilen können, wie hoch die Personenströme und -dichten wirklich sind. Zukünftig wird sich aber das automatisierte Monitoring durchsetzen, mit dem es möglich ist, ohne menschliches Eingreifen, Besucherzahlen und Personenströme quantitativ und relativ genau zu messen. Die automatische Erfassung und Analyse von Menschenströmen stellt jedoch zum heutigen Zeitpunkt noch eine technische Herausforderung dar.

In den ersten Jahren seines Lebens lernt der Mensch Formen, Strukturen und Kontouren richtig zu klassifizieren und aufgrund von Erfahrungswerten richtig einzuordnen. Mit diesen Erfahrungswerten, die sich über eine lange Entwicklung hin gebildet haben, kann der Mensch, mit einer sehr hohen Trefferquote, Personen als solche von anderen Strukturen in Bildern unterscheiden. Rechner und entsprechende Algorithmen sind hinsichtlich ihrer Erkennungsleistung noch lange nicht mit dem Menschen vergleichbar. Insbesondere bei niedriger Auflösung und starken Verdeckungsartefakten in der Szene, ist eine automatische Detektion von Personen mit den bekannten Verfahren bisher kaum möglich.

Im Gegenzug besitzt der Rechner, im Falle einer möglichen Detektion von Objekten, signifikante Vorteile bei der Analyse großer Datenmengen und bei der quantitativen Abschätzung von Messgrößen gegenüber der visuellen Auswertung. Um die Erkennungsleistung eines automatischen Systems, trotz der schwierigen Rahmenbedigungen bei der Beobachtung von Menschenmassen (niedrige Auflösung und Verdeckungen) auf ein einsetzbares Niveau zu steigern, werden im Rahmen dieser Arbeit Methoden basierend auf "`indirekten Merkmalen'', wie beispielsweise Bewegungsmerkmalen, untersucht, um auf Basis Dieser Informationen über die lokale Dichte und den lokalen Personenfluss innerhalb der Menschenmenge zu erschließen. Diese Dichte- und Flussinformationen sollen klassifiziert, quantifiziert und veranschaulicht werden, um etwaige Folgemaßnahmen bei kritischen Situationen zu erschließen. 
\newpage
Kurzfristige Folgemaßnahmen könnten \zb Evakuierung und Absperrung bestimmter Gebiete oder Eingriffe in Unruhen sein. Langfristige Folgemaßnahmen können \zb eine verbesserte Planung der Infrastruktur der Veranstaltung oder die Wahl eines größeren Veranstaltungsgeländes sein. Es soll ein Verfahren zur Erkennung und Messung der Bewegungs- und Gruppierungsmuster entwickelt und hinsichtlich seiner Performanz bewertet werden, um zu zeigen, dass es, auch unter Verwendung von niedrig aufgelösten Videodaten, für die Analyse von Dichte- und Flussinformationen in dicht gedrängten Personengruppen geeignet ist.

### 2018-12-21 thesis_sebastian_bek/content/1-Einleitung.tex
Initial commit

\chapter{Einleitung}
\label{chap:ein}

%Sebastian
Diese Arbeit befasst sich mit der algorithmischen Detektion und Messung von Bewegungs- und Gruppierungsmustern von Personen, wobei Trajektorien (Pfade) als Grundlage für die darauf aufsetzende Dichte- und Flussschätzung verwendet werden. Besonders intensiv werden dabei die Personenzählung (in Personen/Bildausschnitt), die Berechnung einer mittleren Strömung (in Personen/Frame) und die Ableitung eines Dichtefaktors (in \% einer Maximaldichte) in einem bestimmten Bildausschnitt behandelt. Das Verfahren soll in Kombination mit Anderen in Form einer integrierten Software zur komfortablen Analyse von Gruppierungs- und Bewegungsmustern und zur Einrichtung von Schnellwarnsystemen eingesetzt werden.

Der nachfolgende Abschnitt (\ref{sec:motiv}) befasst sich mit der Motivation für die automatisierte Videoauswertung. Er basiert auf der Dissertation "`Automatische Erfassung präziser Trajektorien in Personenströmen hoher Dichte'' von Maik Boltes \cite{boltes} und dem Artikel "`A Method for Counting Moving People in Video Surveillance Videos'' von Donatello Conte et al. \cite{conte}. Anschließend werden die Ziele dieser Arbeit (s. Abschnitt \ref{sec:ziele}) und im folgenden Abschnitt \ref{sec:std} der aktuelle Stand der Technik behandelt.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
\label{sec:motiv}

Personenströme und -anordnungen im Alltag sind für uns gewöhnliche Abläufe, die aber dennoch interressante Erkenntnisse über die Selbstorganisation und die Gewohnheiten der Menschen liefern können. Zudem wäre es mit Sicherheit von Vorteil ein tiefgreifendes Verständnis der bisher nur grob erforschten Dynamik von Personenströmen zu gewinnen, um Beobachtungen zu machen, die in sicherheitskritischen und ökonomischen Anwendungsgebieten hilfreich sein können. 
\newpage
Die Erfassung dicht gedrängter Personengruppen mit Kameras und Bildauswertungssoftware erweist sich dabei als besonders zielführend, da analysierte Szenen vom Computer direkt ausgewertet und die angezeigten Informationen aufbereitet werden können. An das Personal werden dabei geringere Anforderungen an Aufmerksamkeit und Organisation gestellt. Es können außerdem Messdaten aufgenommen und quantitative Aussagen über das Verhalten von Personengruppen gemacht werden.

Gestützte Monitoringssysteme sind vor allem an Orten, die von vielen Menschen gleichzeitig besucht werden (\zb Großveranstaltungen oder Fußgängerzonen) sinnvoll. Denn mit höheren Besucherzahlen nimmt möglicherweise auch das Risiko für Schadensereignisse oder Kriminalität zu, weil die Zahl an Personen, die bereit sind Verbrechen zu begehen und Tumulte zu verursachen, mit der Besucherzahl steigt. Solche Großveranstaltungen bergen zudem die Gefahr für Massenpaniken, weswegen als Grundsatz für eine Veranstaltungsplanung Flucht- und Rettungswege ausnahmslos freigehalten werden sollten. Langfristig kann mit flächigem videogestütztem Monitoring die Planung verbessert werden, indem \zb kritische Gebiete entlastet werden (breitere Tore/Türen, Fluchtwege) oder die Zeitplanung für die öffentlichen Verkehrsmittel überdacht wird. Für kurzfristige Reaktionen kann mit der integrierten Software ein Schnellwarnsystem, in Form eines Kameranetzwerkes, am Veranstaltungsort installiert werden, das schnell ortsbezogene Notrufe tätigen und exakte Positionen von besonders kritischen Gebieten des Veranstaltungsgeländes übermitteln kann. Kritische Gebiete sind zum Beispiel Staugebiete, also Orte, an denen die Personenzahl hoch ist und eine geringe Dynamik herrscht. Ein solches Schnellwarnsystem sollte kritische Situationen schnell erkennen und melden können, weswegen als Vorraussetzung die Extraktion der Laufwege auch bei hohen Personendichten (, die kritischen Situationen entsprechen,) verlässlich anwendbar sein muss.
Außerdem ist es möglich, durch eine Kalibrierung der Kameras, georegistrierte Messwerte zu erhalten, die beispielsweise auf einer Karte oder Aufnahme des Veranstaltungsgeländes dargestellt werden können. Dies ermöglicht eine einfache visuelle Auswertung der Daten, die geringe Anforderungen an die Aufmerksamkeit der Nutzer stellt. Die Daten können entweder in Echtzeit oder im Nachhinein zur verbesserten Planung analysiert werden. Zusätzlich kann das aufgenommene Beweismaterial zu Rate gezogen werden, wenn Unklarheiten über die Ursache für den sicherheitskritischen Fall bestehen. Dies sollte aber ohne Identifizierung von Einzelnen erfolgen.

%Neue Seite?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Aufgabenstellung und Ziele}
\label{sec:ziele}
In dieser Arbeit wird die Untersuchung und Entwicklung eines Videoanalyseverfahrens zur automatischen Extraktion von Trajektorien (Pfaden) bewegter Personen behandelt. Das Verfahren soll im Rahmen eines Videoanalysesystems, das für die Realisierung eines Schnellwarnsystems und für die komfortable Analyse von Bewegungs- und Gruppierungsmustern (s. Abschnitt \ref{sec:motiv}) zur Langzeitplanung von Veranstaltungen geeignet ist, entwickelt und dokumentiert werden. 
\newpage

Es sollen Laufwege bewegter Personen, deren Ecken zuvor algorithmisch detektiert wurden, als Bildkoordinaten in Listen eingetragen und verwaltet werden. Solche Laufwege werden als Trajektorien bezeichnet. Dabei sollen nur bewegte Eckpunkte verwendet werden, weswegen gänzlich stillstehende Personen (zunächst) nicht erfasst werden. Weiterhin können aus den gespeicherten Positionen der Trajektorien, innerhalb von bestimmten Bildausschnitten, Messgrößen wie Personenfluss, Personenzahl und Dichtefaktoren abgeleitet werden. Das Verfahren sollte zuverlässig sein und keine auffälligen, kritischen Situationen übersehen. Es soll erprobt und mithilfe von Grundwahrheiten bewertet werden, um eine Aussage über die Wahrheitstreue der Messgrößen treffen zu können.

%\\ \\
%Nachfolgend werden Vor- und Nachteile der Nutzung von Bildauswertungssoftware gegenüber manueller Videoüberwachung erfasst:

%\begin{center}
%\fbox{
%\begin{minipage}[t]{0.47\textwidth}
%\begin{center}
%\bigskip
%\textbf{Vorteile}:
%\bigskip
%\hline
%\end{center}
%\begin{itemize}
%\item[+] System benötigt geringeren Personaleinsatz
%\item[+] komfortable Videoüberwachung
%\item[+] quantitative Aussagen(ohne subjektiven Einfluss) möglich
%\item[+] System hat keinen Konzentrationsverlust
%\item[+] System übersieht in der Regel keine Gefahren
%\end{itemize}
%\bigskip
%\bigskip
%\bigskip
%\end{minipage}
%\hskip 5pt
%\vline
%\hskip 7pt
%\begin{minipage}[t]{0.47%\textwidth}
%\begin{center}
%\bigskip
%\textbf{Nachteile}:
%\bigskip
%\hline
%\end{center}
%\begin{itemize}
%\item[-] unflexible Reaktion des Systems auf Szenenwechsel
%\item[-] System löst Falschalarme aus
%\item[-] System besitzt keine menschlichen Erfahrungswerte
%\item[-] Leistung des Systems ist hardwareabhängig
%\item[-] aufwendige Kalibrierung des Systems
%\end{itemize}
%\bigskip
%\end{minipage}
%\bigskip
%\bigskip %\%\
%}
%\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Stand der Technik}
%Bei Anne geht es mehr um Segmentierung - hier mehr um Erstellung von Trajektorien(Tracking) der Personen
\label{sec:std}

Die Idee, automatische Videoanalyseverfahren zur Erkennung von Mustern und Objekten einzusetzen, ist keine neue Vorstellung. Wegen der zunächst nur wenig fortgeschrittenen Technik, setzte sich diese Idee jedoch erst in den 1980er Jahren durch. Heute gibt es vielseitige Ansätze, um diese Aufgaben zuverlässig durchzuführen. Dennoch haben diese Verfahren ihre Grenzen und es existiert ein weitreichendes Verbesserungspotenzial.

Wie in Abschnitt \ref{sec:ziele} beschrieben, wird in dieser Arbeit die algorithmische Detektion und Messung von Dichte- und Flussinformationen in Personengruppen behandelt. Um solche Gruppierungs- und Bewegungsmuster erfassen und messen zu können, müssen die Personen zunächst quantitativ erfasst werden. Zur Zählung von Personen/Objekten existieren, nach aktuellem Stand, zwei grundlegende Ansätze. Zunächst muss, in beiden Fällen, eine Extraktion einer ROI (region of interest - \zb Punkte, Flächen) zur groben Ausfilterung von uninteressanten Bereichen, an denen sich gerade keine Personen befinden können, vorgenommen werden. Dabei werden uninteressante, statische Bereiche im Bild ausgefiltert, während die interessanten Bereiche hinsichtlich Personendetektion, Dichte- und Flussschätzung weiterverarbeitet werden. Nachfolgend werden die beiden Ansätze beschrieben und mit Beispielen ausgeführt.

%\subsection{Extraktion einer ROI}
% Zur Extraktion einer ROI gibt es derzeit einige Möglichkeiten, wobei die Wichtigsten nachfolgend aufgezählt werden:

%\subsubsection{Segmentierungsverfahren:}

%\begin{itemize}

%\item Pixelorientierte Verfahren: Diverse Schwellwertverfahren eignen sich zur Segmentierung, wenn die Dichte der Grauwerte im Histogramm ausreichend bimodal ist(s. Otsu-Schwellwertverfahren %\cite{segcourse}).

%\item Kantenorientierte Verfahren: Diverse Kantenfilter, wie Scharr- und Sobel-Operator eignen sich zur Segmentierung, wenn anschließend ein "`Flood-Fill-Algorithmus'' eingesetzt wird, um den Hintergrund zu fluten. %\cite{segcourse2}

%\item Modellbasierte Verfahren: Diverse geometrische Modelle wie Kreise oder Geraden eignen sich zur Segmentierung, wenn beispielsweise die Form der Kanten mit ihnen verglichen wird. %\cite{segcourse2}

%\item Regionenorientierte Verfahren: Diese Verfahren wählen Pixelpositionen als Ausgangspunkte und lassen Vordergrundregionen nach verschiedenen Kriterien wie Ähnlichkeit ausgehend von dieser Position wachsen oder sich vereinen. %\cite{segcourse2}

%\item Texturorientierte Verfahren: Texturen in Vordergrund-Regionen werden erkannt und anhand ihrer Beschaffenheit interpretiert. Unruhige Strukturen deuten %\zb auf Personengruppen und Glatte auf eine leere Oberfläche hin(s. "`Haar-Wavelets'' oder "`gray-level-co-occurence-matrix'' %\cite{segcourse2}).

%\item Selektion des Hintergrunds: Der Hintergrund wird erfasst und subtrahiert(Subtraktion des Hintergrunds). Der Hintergrund wird abgezogen und erhält somit idealerweise die Farbe Schwarz, während der Vordergrund erhalten bleibt. Wie in %\cite{zhao} beschrieben, kann Hintergrund zum Beispiel detektiert werden, indem statische Regionen im Bild markiert werden.

%\item Detektion mit Sensor-Array/3D-Kameras: Eine Szene wird mit einer 3D-Kamera gefilmt, um die räumliche Struktur zu dokumentieren und 3D-Objekte zu segmentieren, die als 3D-Modelle gespeichert werden.

%\end{itemize}

%\subsubsection{Merkmalsdetektionsverfahren}

%\begin{itemize}
    %\item Extraktion von Vordergrund-Merkmalen("`Features'' - \zb Punkten), die beispielsweise durch extrahierte Eckpunkte(s. "`Harris Corner Detector'') gegeben sind und in Kombination mit einer Bewegungsschätzung (optischer Fluss) als lokale Bewegungsmerkmale verwendet werden können.
%\end{itemize}

\subsection{direkter Ansatz}
\begin{enumerate}
\item Segmentierung von VG-Bereichen: nach Möglichkeit mit beinhalteten Personen
\item Detektion und individuelle Separation der Personen
\item Zählung der separierten Vordergrund-Regionen
\end{enumerate}

Im Fall des direkten Ansatzes, werden in Kombination zur Extraktion einer ROI in VG-Regionen Muster und Merkmale gesucht, die Menschen zugeordnet werden sollen, um diese einzeln zu separieren. 
\newpage
Nachfolgend werden einige Beispiele für den direkten Ansatz aufgeführt:
\vskip 10pt
\emph{Objektklassifizierung:}
\begin{itemize}

\item Vergleich der Form: Kanten von Vordergrund-Regionen werden mit Modellen (Rechtecke, Ellipsen o.ä.), die entweder ganze Menschen oder Körperteile modellieren, verglichen (s.\cite{rittscher}).

\item Vergleich der Kantenstatistik: Anzahl, Ausrichtung und Form der Kanten können Menschen beschreiben, wobei Statistiken/Histogramme über die Häufigkeit von Kantenrichtungen erstellt und gespeichert werden, falls sich ein/kein Mensch im Bildausschnitt befindet (s. HOG = histogram of oriented gradients). Anschließend kann, über einen Vergleich der Kantenstatistik des analysierten Bildausschnitts mit der gespeicherten Kantenstatistik, entschieden werden, ob ein Mensch in einem Bildausschnitt befindlich ist. Dies wird von einem lernenden Algorithmus (SVM: support vector machine) durchgeführt. Zusätzlich kann hier, wie in \cite{loy2013crowd} beschrieben, bestimmt werden, ob sich gerade mehr (komplexere Kanten) oder weniger Menschen (einfache Kanten) in einem größeren Bildausschnitt aufhalten.

\item Vergleich von aufgenommenen 3D-Modellen mit Standardmodellen: Nach \cite{zhao} können Menschen erkannt werden, indem ein 3D-Modell, das mit einer 3D-Kamera oder einem Sensor-Array erstellt wurde, mit Standardmodellen für Menschen verglichen wird.

\item Gruppieren von extrahierten Merkmalen ("`Features'' - \zb Punkten) nach ihren Bewegungscharakteristika: Punkte, die sich nahezu gleich schnell in eine Richtung bewegen, gehören mit hoher Wahrscheinlichkeit nur zu einer Person (s. \cite{brostow}).

\end{itemize}
\vskip 5pt
Anschließend wird, über ein Labeling-Verfahren (\zb ein sog. "`Point Clustering''), eine Vergabe von Identifikationsnummern durchgeführt, wobei die einzeln (als Vordergrund-Regionen) separierten Personen mit IDs versehen und gezählt werden können.

Beim nachfolgenden indirekten Ansatz wird eine solche Objektklassifizierung nicht durchgeführt, sondern direkt mit der extrahierten ROI gearbeitet.

\subsection{indirekter Ansatz}

\begin{enumerate}
\item Detektion und Extraktion von lokalen Merkmalen (segmentierungslos - \zb Punkte/Pixelpositionen), die beispielsweise durch extrahierte Eckpunkte (s. "`Harris Corner Detector'' \cite{albiol}) gegeben sind und in Kombination mit einer Bewegungsschätzung (optischer Fluss) als lokale Bewegungsmerkmale verwendet werden können.
\item Ableitung der Personenzahl aus der Zahl der lokalen, extrahierten Merkmale, unabhängig von einer separaten Detektion von Personen
\end{enumerate}

Dieser Ansatz gilt als robuster, weil die vereinzelte Separation der Personen, besonders in überfüllten Gebieten, ein zu großes Problem darstellt und bisher nicht zuverlässig gelöst wurde.
\vskip 5pt
Nachfolgend werden einige Beispiele für den indirekten Ansatz aufgelistet:

\begin{itemize}

\item Zählen der Menge an nicht-statischen Pixeln (s. \cite{cho})

\item Zählen der Menge an nicht-statischen Eckpunkten, extrahiert von einem Corner Detector (\zb Harris Corner Detector, wie in \cite{albiol} beschrieben)

\item Erfassen der Größen und Häufigkeiten von VG-Bereichen (vgl. \cite{kong})

\item Erfassen der fraktalen Dimension von VG-Bereichen (vgl. \cite{marana}):\\
\begin{center}
$D = \frac{\text{log(Zahl der selbstähnlichen Bestandteile)}}{\text{log(Abbildungsmaßstab)}}$
\end{center}
\end{itemize}

\vskip 5pt
In dieser Arbeit wird eine Ausführung des indirekten Ansatzes behandelt. Dabei kommt der Harris Corner Detector, wie in \cite{albiol} beschrieben, zum Einsatz, um Eckpunkte/Merkmale zu extrahieren und vom Hintergrund zu trennen. Ein Corner Tracker erfasst die relative Bewegung (den optischen Fluss) dieser Merkmale entlang der Bildzeilen und -spalten. Die Anzahl an bewegten Personen wird in ein Verhältnis mit der Anzahl an sich bewegenden, extrahierten Eckpunkten gesetzt und daraus werden weitere Größen wie Personenfluss und ein Dichtefaktor abgeleitet.

Wichtige Verfahren, in Bezug auf diese Arbeit, sind somit zunächst der Harris Corner Detector, gekoppelt an einen Corner Tracker, der die Bewegung (den optischen Fluss) erfasst, sowie das Verfahren zur Erfassung von Trajektorien (Pfaden) der extrahierten Eckpunkte. Diese Grundlagen werden im Kapitel \ref{chap:grund} näher erläutert. Diese Arbeit soll zeigen, dass das entwickelte Verfahren, auch mit den niedrig aufgelösten Daten \zb von CCTV-Überwachungskameras, in Personengruppen hoher Dichte zuverlässig angewandt werden kann.

### 2018-12-21 thesis_sebastian_bek/content/2-Grundlagen.tex
Initial commit

\chapter{Grundlagen}
\label{chap:grund}

Dieses Kapitel befasst sich mit den Basismethoden, welche in den untersuchten Verfahren eingesetzt werden. Zunächst wird in Abschnitt \ref{sec:harris} das zugrundeliegende Verfahren zur Extraktion lokaler Bildmerkmale beschrieben, das auf dem Harris Corner Detector \cite{collinscourse} basiert. Anschließend wird im Abschnitt \ref{sec:corner} der verwendete Corner Tracker behandelt, welcher zu den lokalen Bildmerkmalen Bewegungsinformationen (vgl. Abschnitt \ref{sec:planes}) hinzufügt (motion vectors - Bewegungsvektoren). Im letzten Abschnitt \ref{sec:trajektorien} wird das verwendete Verfahren zur Erfassung von Trajektorien (Pfaden) der lokalen Bildmerkmale, und dadurch indirekt der Personen (-gruppen), näher beschrieben.

\section{Harris Corner Detector}
\label{sec:harris}
Der "`Harris Corner Detector'' \cite{collinscourse} ist ein Algorithmus, der verschiedene Bildausschnitte (Patches) einer Bildmatrix $I$ nach Ecken (engl.: Corners) durchsucht und diese in Form der Position des Center Pixels extrahiert, falls sich eine Ecke im Bildausschnitt befindet. Eine grafische Beschreibung der Idee des Harris Corner Detector wird auf der folgenden Seite dargestellt. Dabei werden unterschiedliche Patches (s. Abbildung \ref{gradient} - blau) mit unterschiedlichen Situationen betrachtet:

\begin{figure}[H]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    \label{c)}
  \end{minipage}
\caption{Betrachtung der Gradientenbeiträge verschiedener Patches}
\label{gradient}
\end{figure}

%BILD Patch im flachen und auf einer Ecke
Diese anschauliche Beschreibung der Idee des Harris Corner Detectors zeigt, dass in der Umgebung einer Ecke immer viele hohe Intensitätsunterschiede (Gradientenbeiträge) in verschiedene Richtungen vorliegen. In Abbildung \ref{harris} werden Verteilungen der Gradientenbeiträge in x- und y-Richtung für die verschiedenen Fälle in einem Bildausschnitt betrachtet:
\bigskip
\bigskip
\begin{figure}
  \centering
  \fbox{
    \includegraphics[width=0.8\textwidth]
    {images/dummy.png}
  }
  \caption{Verteilungen von Gradientenbeiträgen $I_x,I_y$ \cite{collinscourse}}
  \label{harris}
\end{figure}
\newpage
Dabei wird klar, dass je nach Situation eine andere Verteilung der Daten vorliegt und zwischen allen drei Fällen klar unterschieden werden kann. Dafür muss aber unterschieden werden können, ob die Varianz der Verteilung mit Bezug zum Ursprung generell klein (im Fall einer Fläche), groß in eine Richtung (im Fall einer Kante) oder groß in mindestens 2 Richtungen (im Fall einer Ecke) ist. Dazu wird eine Kovarianzmatrix (auch bekannt als Strukturtensor) $M$ für die Verteilung der Gradientenbeiträge $I_x$,$I_y$ hergeleitet (Herleitung siehe \cite{cscourse}), die die Varianz der Daten mit Bezug zum Ursprungswert ($I_x=0$, $I_y=0$) im aktuellen Bildausschnitt mit Center Pixel $x,y$ beschreibt:

\begin{align}
M = \sum_{u,v}{w(u,v)
\begin{bmatrix}
I_x^2 & I_xI_y\\
I_xI_y & I_y^2\\
\end{bmatrix}
}
&=
\begin{bmatrix}
a = \sum_{u,v}{w(u,v)I_x^2} & b = \sum_{u,v}{w(u,v)I_xI_y}\\
b = \sum_{u,v}{w(u,v)I_xI_y} & c = \sum_{u,v}{w(u,v)I_y^2}\\
\end{bmatrix}
\notag \\ \bigskip \bigskip \bigskip
\begin{bmatrix}
a & b\\
b & c\\
\end{bmatrix}
&=
\begin{bmatrix}
\text{var}(I_x) & \text{covar}(I_x,I_y)\\
\text{covar}(I_x,I_y) & \text{var}(I_y)\\
\end{bmatrix}
\end{align}
\vskip 5pt
\begin{flushleft}
mit  $u,v \in W(u,v)$, $I_x=I_x(x+u,y+v)$ und $I_y=I_y(x+u,y+v)$
\end{flushleft}
\vskip 5pt
$I_x$ und $I_y$ werden wegen der kompakteren Schreibweise abgekürzt.
$w(u,v)$ ist hier eine Gewichtungsfunktion wie die Rechteck- oder Gauss-Funktion. $W(u,v)$ beschreibt den zulässigen Wertebereich, innerhalb dem $u$ und $v$ liegen müssen, um den definierten Bildausschnitt nicht zu überschreiten. Für die Eigenwerte der Matrix $M$ gilt:
\begin{equation}
\lambda_{1/2} = \frac{a+c \pm \sqrt{(a-c)^2 + 4b^2}}{2}
\end{equation}
Im Fall einer Kovarianzmatrix geben die Eigenvektoren Dieser die 2 orthogonalen Hauptrichtungskomponenten der Varianz der Daten an. Die Eigenwerte $\lambda_1$ und $\lambda_2$ beschreiben die Länge der größten Eigenvektoren und damit die Varianz der Verteilung von $I_x$, $I_y$ in Richtung der Eigenvektoren. Damit ergeben sich im Falle einer:
\begin{itemize}
\item flachen Region (s. Abbildung \ref{harris} - flat): 2 kleine Eigenwerte der Matrix B (keine Varianz) $\Rightarrow \lambda_1\sim \lambda_2\approx 0$ 
\item Kantenregion (s. Abbildung \ref{harris} - edge): 1 großer, 1 kleiner Eigenwert der Matrix B (große Varianz in eine Richtung) $\Rightarrow \lambda_1 >> \lambda_2$ \hskip 5pt $v$ \hskip 5pt $\lambda_2 >> \lambda_1$
\item Eckregion (s. Abbildung \ref{harris} - corner): 2 große Eigenwerte der Matrix B (große Varianz in mindestens 2 Richtungen) $\Rightarrow \lambda_1\sim \lambda_2$ 
\end{itemize}

\newpage

Um in allen Fällen verschiedene Ergebnisse zu erhalten, wird eine sogenannte Corner Response an jedem Pixel definiert, die große, positive Werte für Ecken, große Negative für Kanten und kleine Werte für flache Regionen annimmt:
\begin{equation}
R = \lambda_1\lambda_2 - k(\lambda_1 + \lambda_2)^2
\end{equation}
$k$ ist ein empirisch festgelegter Parameter, dessen Wertebereich zwischen 0,04 und 0,06 liegt.
Um den Rechenaufwand, der durch die Berechnung der Eigenwerte entsteht (Quadratwurzeln sind rechenintensiv) zu verringern, wird die Corner Response oft so definiert:
\begin{equation}
R = \det{B} - k(\text{spur\ } B)^2
\end{equation}
mit: $\det{B} = \lambda_1\lambda_2$ und $\text{spur\ } B = \lambda_1+\lambda_2$\\

\section{Corner Tracker}
\label{sec:corner}
Als Grundlage der Arbeit liegt eine Bibliothek "`OCV\_CornerTracker'' vor, in der im Rahmen eines sogenannten Corner Trackers der Harris Corner Detector bereits implementiert ist. Das Blockschaltbild dieser Bibliothek wird in Abbildung \ref{ocv} dargestellt.
\bigskip
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {images/dummy.png}
  }
  \caption{Blockschaltbild: OCV\_CornerTracker}
  \label{ocv}
\end{figure}
\begin{flushleft}
\underline{Funktion:}
\end{flushleft}
Das hier eingesetzte Verfahren zum Tracking von Harris Corner Features (Merkmalen) berechnet zunächst für jeden Pixel im vorherigen (n-1.) Eingangsbild den sogenannten optischen Fluss \cite{Baker07adatabase}. Dabei wird ein Intensitätsprofil aus der Umgebung eines Pixels erstellt. Dieses Profil wird in der Umgebung des Pixels im aktuellen (n.) Eingangsbild gesucht. Die relative Verschiebung des Mittelpunkts im Profil im Vektorraum, zwischen beiden Eingangsbildern, gibt den optischen Fluss des Pixels in Zeilen- und Spaltenrichtung an. 
\newpage
In Abbildung \ref{gaussians} werden beispielhaft solche Verschiebungsvektoren in einem Videobild in blau eingezeichnet. Anschließend wird zu den jeweiligen Pixelpositionen, die im vorherigen (n-1.) Eingangsbild als Harris Corners klassifiziert wurden, der zugehörige optische Fluss als Bewegungsschätzung hinzugefügt. An Positionen, an denen die Merkmale keine Bewegungen aufweisen oder keine Merkmale extrahiert wurden, werden die Verschiebungsvektoren gleich 0 gesetzt. Zusammenfassend enthalten die generierten Verschiebungsvektoren ausschließlich pixelbezogene Bewegungen der extrahierten Merkmale des Harris Corner Detectors vom letzten (n-1.) zum aktuellen (n.) Frame.
\vskip 5pt
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.75\textwidth]
    {images/dummy.png}
  }
  \caption{optischer Fluss von detektierten Ecken an Marathonläufern \cite{AliS07}}
  \label{gaussians}
\end{figure}

\section{Verschiebungsvektoren}
\label{sec:planes}

Die Verschiebungsvektoren, die vom Corner Tracker (s. Abschnitt \ref{sec:corner}) ausgegeben werden, beinhalten die relativen Bewegungen der extrahierten Merkmale in Zeilen- und Spaltenrichtung am Ort des Ursprungs. Aus den beiden Vektorkomponenten können für jede detektierte Position im Bild Vektorbeträge berechnet werden:\\
\begin{equation}
\abs{\vec{a}} = \sqrt{DX^2 + DY^2} \hskip 30pt [\text{px}]
\end{equation}
\vskip 5pt
Weil diese Vektorlänge sich im Verfahren auf die Bewegung eines Merkmals über einen Frame hinweg bezieht, ist diese Angabe eine Merkmalsgeschwindigkeit:\\
\begin{equation}
v = \frac{\abs{\vec{a}}}{1\text{frame}} = \frac{\sqrt{DX^2 + DY^2}}{1\text{frame}} \hskip 30pt [\frac{\text{px}}{\text{frame}}]
\end{equation}
\newpage
Kennt man den genäherten Abstand $d(x,y)$ zwischen den Pixeln im betreffenden Bildausschnitt als abbildende Funktion(\zb in Form einer Karte mit Einträgen in $[\frac{m}{\text{px}}]$) und zusätzlich die Framerate $\text{FR}$ (in $[\text{fps}]$) erhält man die reale Merkmalsgeschwindigkeit genähert in SI-Einheiten:\vskip 3pt
\begin{equation}
v_{SI}(x,y) \approx v(x,y)\cdot d(x,y)\cdot\text{FR} \hskip 30pt [\frac{m}{s}]
\end{equation}

%evtl abbildende Funktion f(v(x,y)) hinzunehmen

\section{Erfassung von Trajektorien}
\label{sec:trajektorien}

In diesem Abschnitt wird das Verfahren zur Erfassung von Trajektorien (Pfaden) der Personen beschrieben. Die Funktion dieses Verfahrens wird in Abbildung \ref{idMaps} schematisch dargestellt:
\bigskip
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.8\textwidth]
    {images/dummy.png}
  }
  \caption{Funktionsweise des Verfahrens zur Erfassung von Trajektorien}
  \label{idMaps}
\end{figure}

Die Trajektorien werden im verwendeten Verfahren durch eine ID (Identifikationsnummer) und eine Trajektorienadresse (s. Abb. \ref{idMaps} - trackID, adress) beschrieben, an der die bereits passierten Positionen gespeichert werden.

Die Verschiebungsvektoren des optischen Flusses werden zunächst für jeden Eintrag nach Positionen durchsucht, an denen die Vektorlängen (Merkmalsgeschwindigkeiten) oberhalb einer festgelegten Geschwindigkeitsschwelle liegen. Wird eine solche Position gefunden, liegt dort ein sich deutlich bewegendes, interessantes Merkmal vor, dessen nähere Umgebung in der ID Map des Vorgängerframes nach bereits vorhandenen IDs durchsucht wird (vgl. Abb. \ref{searchRadius}). Wird keine andere ID in der Umgebung in der ID Map gefunden, wird eine neue ID für diese, neu erstellte, Trajektorie vergeben (vgl. Abbildung \ref{idMaps} ID Map - schwarz). Zum Erweitern der vorhandenen Trajektorien wird für jede Pixelposition zusätzlich die ID Map des Vorgängerframes nach bereits vorhandenen IDs durchsucht.
\newpage
Wird eine ID gefunden, wird umgekehrt die Umgebung Dieser in den Verschiebungsvektorlisten nach Positionen durchsucht, an denen die Merkmalsgeschwindigkeit oberhalb der Geschwindigkeitsschwelle liegt. Wird ein solches Merkmal gefunden, wird die ID für dieses Merkmal übernommen (vgl. Abbildung \ref{idMaps} ID Map - blau). Um die Positionen der Trajektorien zu aktualisieren, werden die erstellten/übernommenen IDs mit den Versätzen DX und DY des Verschiebungsvektors in eine neue ID Map eingetragen, die nach der Evaluierung eines Frames gespeichert wird, um im nächsten Frame zum Verknüpfen der Trajektorien (als ID Map des Vorgängerframes) zu dienen. Wird eine neue Trajektorie erstellt, wird dieser Versatz auf die Merkmalsposition addiert. Beim Erweitern einer bereits vorhandenen Trajektorie wird der Versatz auf die Position der ID addiert, um möglichst gerade Trajektorien zu erhalten. So beugt man instabilen Merkmalen vor, die ihre Detektionsposition an der Person ändern. Zugleich werden die IDs in einer Liste von Trajektorien (vgl. Abb. \ref{idMaps} - TrackList) zusammen mit allen Positionen eingetragen, die eine bestimmte Trajektorie bisher passiert hat. Verhält sich eine Trajektorie länger als die maximale Verweilzeit statisch, wird sie entfernt. 
\bigskip
\bigskip
\begin{figure}[H]
  \begin{minipage}{0.6\textwidth}
  In nebenstehender Grafik erkennt man beispielhaft die Suche nach IDs von einem gefundenen Merkmal in der ID Map des Vorgängerframes in einer sog. 8er-Nachbarschaft. Hier ist keine ID vorhanden, weswegen eine neue ID für dieses Merkmal erstellt wird. Diese wird mit dem DX/DY-Versatz des Verschiebungsvektors in die neue ID Map eingetragen. Das rot markierte Pixel ist der Center Pixel, an dem ein sich bewegendes Merkmal vorliegt. Dieser Ablauf entspricht der Situation in Frame 1 in Abbildung \ref{idMaps} (schwarz).
  \end{minipage}
\hfill
  \begin{minipage}{0.2\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    \caption{Suche nach IDs}
    \label{searchRadius}
  \end{minipage}
\end{figure}

### 2018-12-21 thesis_sebastian_bek/content/3-Implementierung.tex
Initial commit

\chapter{Systemarchitektur}
\label{chap:imp}
In folgendem Kapitel wird die Implementierung des, in Kapitel \ref{chap:ein} beschriebenen, Videoanalysesystems behandelt. Dabei wird zunächst in Abschnitt \ref{sec:mess} der Aufbau des Systems und die Funktion der Bildauswertungssoftware als Teil der gesamten Systemarchitektur erklärt. Im Abschnitt \ref{sec:sig} wird die Umsetzung der Eingangsgrößen (Bilder als Sensorsignale) in Messgrößen wie Personenzahl, -fluss oder -dichte in bestimmten Bildausschnitten beschrieben. Die Umsetzung erfolgt auf Basis der, in Kapitel \ref{chap:grund} beschriebenen, Grundlagen. Weiterhin werden in Abschnitt \ref{sec:konfig} geplante Ablaufschritte zur Konfigurierung der Software beschrieben.

\section{Beschreibung des Messsystems}
\label{sec:mess}
Großveranstaltungen sind meist nicht mehr mit nur einer Kamera zu überblicken. Daher werden meist ganze Kameranetzwerke als Verbund mehrerer Kameras als Messsysteme eingesetzt. Ein möglicher Messaufbau wird in Abbildung \ref{messsystem} dargestellt.

\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {./images/dummy.png}
  }
  \caption{Aufbau eines möglichen Videoanalysesystems}
  \label{messsystem}
\end{figure}

Die Kameras sollten in ihrer Ausrichtung und ihren Abbildungseigenschaften statisch sein. Setzt man dies voraus, besitzen die Kameras fest definierte Anordnungen zueinander. Die Auflösung des Kamerabilds muss ausreichend sein, um einzelne Personen aufzulösen. Nachdem das System aufgestellt und eingerichtet ist, müssen Grenzwerte der Messgrößen für jede Kamera erlernt und die Konfigurierung, wie in Abschnitt \ref{sec:konfig} beschrieben, durchgeführt werden. Das Eingangsbild der Kameras wird in rechteckige Bildausschnitte eingeteilt, für die jeweils Statistiken und Messwerte generiert werden. Diese werden als Listen oder Tabellen übertragen. In der Überwachungszentrale (siehe Abbildung \ref{messsystem}), in der die Datenströme der Kameras zusammengeführt werden, können, durch eine vorherige Kalibrierung, die Patch Positionen der Kamerabilder georegistriert auf Karten des Überwachungsgeländes übertragen werden. 
\newpage
Damit können zudem die Messungen aus Bildausschnitten verschiedener Kameras zusammengeführt und auf eine einzelne Ortsposition bezogen werden. Geben die Kameras als Netzwerk an einer Ortsposition eine durchschnittliche Messgröße in einem kritischen Bereich aus, können automatisch Notrufe und Warnungen ausgesendet werden. Zusätzlich kann das Personal in der Überwachungszentrale eine generierte Karte des Überwachungsgeländes, die georegistrierte und gemittelte Messungen der versch. Kameras darstellt, analysieren. Hier werden die durchschnittlichen Messwerte der verschiedenen Kameras als Heat Map wahlweise mit angezeigten Messwerten dargestellt. Zusätzlich kann im Nachhinein nach der Veranstaltung das Bewegungsverhalten der Personengruppen im Detail analysiert werden, indem Überwachungsvideos gespeichert werden und in Ruhe das Verhalten der Messgrößen betrachtet wird.
\newpage
Nachfolgend werden die Anforderungen des Systems an die Nutzer beschrieben:

\begin{flushleft}
\underline{Anforderungen des Systems:}
\end{flushleft}
\begin{itemize}
\item statische Rahmenbedingungen: statische Kameraperspektive
\item Kalibrierung der Positionen/Perspektiven der Kameras (wird nicht im Rahmen dieser Arbeit behandelt)
\item Konfigurierung mit, für die Szene sinnvollen, Evaluations-Parametern wie Such- oder Abstands-Radius
\item Konfiguration mit intrinsischen und extrinsischen Kameraparametern zur Skalierung von Evaluations-/Filtereinstellungen und Extremwerten (\zb in Form einer Patch-Karte)
\item Trainieren von Grenzwerten, um relative Messwerte zu erhalten
\end{itemize}

\section{Signalverarbeitung}
\label{sec:sig}
Um innerhalb eines Bildes lokale, statistische Messergebnisse zu erhalten, wird das Eingangsbild zunächst in rechteckige Bildausschnitte (Patches) eingeteilt, deren Höhe und Breite im Konfigurationsdialog bestimmt werden können. Die Trajektorien werden dabei ausgewertet, um sie in die Bildausschnitte entsprechend einzugliedern und patchbezogene Messwerte zu erhalten. Dieses Verfahren stellt in dieser Arbeit die Grundlage für die Signalverarbeitung, also die Umsetzung der Eingangsgrößen in Messgrößen, dar. Die Laufvariable $j$ ist in diesem Kapitel konsistent ein Index für die Frames einer Eingangsbildfolge und indiziert dabei den aktuellen Frame. Die Laufvariablen $x,y$ indizieren eine Position im Eingangsbild. Diese müssen immer innerhalb des Wertebereichs $W(w,h)$ des Bildausschnitts liegen. Die nachfolgend beschriebenen Messgrößen werden in jedem neuen Frame $j$ aktualisiert, d.h. neu berechnet und in sogenannte Patch Maps eingetragen.

\subsection{Trajektorien}
Die Umsetzung der Eingangsgrößen (-bilder) in Messgrößen erfolgt in dieser Arbeit auf Basis von erstellten Trajektorien (wie in Kapitel \ref{chap:grund} Abschnitt \ref{sec:trajektorien} beschrieben). 
\newpage
Die Trajektorien können mit dem erstellten Softwaremodul als Linien gezeichnet werden. Die Erfassung solcher Trajektorien anhand einer künstlich generierten Bildsequenz erkennt man beispielhaft in Abbildung \ref{trajektorien}. 
\bigskip
\begin{figure}[H]
\centering
  \begin{minipage}{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    \label{b)}
  \end{minipage}
\caption{Erfassung und Darstellung von Trajektorien aus versch. Perspektiven \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
\label{trajektorien}
\end{figure}

Hier erkennt man die automatische Erfassung von Trajektorien des entwickelten Verfahrens. Dabei werden 27 Frames einer künstlich generierten Bildsequenz analysiert, in der Personen, die anfangs in gleichem Abstand stehen, in alle Richtungen gegen die Wände drängen. Alle erfassten Koordinaten einer Trajektorie werden als zusammenhängende Linie in blau eingezeichnet. Nachfolgend erkennt man in Abbildung \ref{trajektorien_scenes} die Erfassung und Darstellung von Trajektorien in weiteren Szenen des verwendeten Datensatzes "`AGORASET''(siehe \cite{CourtyPRL2014} \cite{Allain2012ICPR}):
\bigskip
\begin{figure}[H]
\centering
  \begin{minipage}{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    \label{b)}
  \end{minipage}
\caption{Erfassung und Darstellung von Trajektorien in diversen Szenarien \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
\label{trajektorien_scenes}
\end{figure}

\newpage

\subsection{Personenzahl}
Die Zahl der bewegten Personen $N$ wird in ein Verhältnis mit der Summe an Trajektorienendpunkten (-spitzen) im Patch gesetzt. Um die Summe Dieser zu erhalten, wird die Kardinalität (Mächtigkeit) der Menge $S$ an Trajektorienspitzen (Trajektorien an ihrer Position im aktuellen Frame $j$), die sich innerhalb des betrachteten Bildausschnitts befinden, berechnet:

\begin{equation}
N = c_K\cdot |S| \hskip 30pt [\text{personen}]
\end{equation}
\vskip 5pt
\begin{flushleft}
für alle verwendeten Positionen $P(x,y)\in S$ gilt: $x,y\in W(w,h)$
\end{flushleft}
Hier bezeichnet $W(w,h)$ den zulässigen Wertebereich für die Bildkoordinaten $x,y$, falls diese sich im Bildausschnitt befinden. Dieser ist abhängig von der Breite $w$ und Höhe $h$ des Bildausschnitts.
%Damit werden nur die aktuellen Positionen der erstellten Pfade ausgewertet. Dazu wird die aktuelle (ggf. gefilterte) Trajektorienliste betrachtet und jede gefundene ID in einer neuen, mit 0 initialisierten Bildmatrix $S[x,y]$(Trajektorienspitzen) an ihrer aktuellen Position mit 1 markiert. Über Bewegungen $u,v$ vom Center Pixel an $x,y$ werden alle markierten Trajektorienspitzen, die sich im Bildausschnitt befinden, summiert.
%\begin{equation}
%N = C\cdot\sum_{u,v}{S[x+u,y+v]} \hskip 30pt [\text{personen}]
%\end{equation}
%mit $u,v \in W(u,v)$
\vskip 5pt

Wendet man auf $N$ einen zeitlichen gleitenden Mittelwert an, um Rauschen zu unterdrücken, erhält man die mittlere Personenzahl $\bar{N}_j$:

\begin{equation}
\bar{N}_j = \frac{1}{m}\sum_{i=0}^{m-1}{N_{j-i}} \hskip 30pt [\text{personen}]
\end{equation}
\vskip 10pt
Um relative Messgrößen zu erhalten, kann, im Laufe einer Testsequenz, ein Maximalwert $N_{max}$ für die (gemittelte) Personenzählung trainiert werden, der anschließend gespeichert wird. Die relative Personenzahl $N_{rel}$ ergibt sich somit zu:

\begin{equation}
    N_{rel}=\frac{\bar{N}_j}{N_{max}} \hskip 30pt [\%]
\end{equation}
mit $N_{rel}\in \{0..1\}$
\vskip 10pt
Ist dieser Maximalwert von Hand in einem vollen Patch gezählt worden oder anderweitig genau bekannt, kann dieser absolut als $N_{abs}$ im Konfigurationsdialog zusätzlich eingetragen werden. Dann kann aus der relativen Personenzahl im Bildausschnitt näherungsweise ein manuell definierbarer Normierungsterm (Kalibrierung des Zählsystems) $N_{scal}$ in $[\text{personen}]$ berechnet werden:

\begin{equation}
    N_{scal}=N_{rel}\cdot N_{abs} \hskip 30pt [\text{personen}]
\end{equation}
\vskip 10pt
Ist weiterhin die projezierte Fläche A des Bildausschnitts bekannt kann außerdem ein manuell definierbarer Normierungsterm für die Personendichte $D_{scal}$ in $[\frac{\text{personen}}{m^2}]$ abgeleitet werden:

\begin{equation}
    D_{scal}=\frac{N_{rel}\cdot N_{abs}}{A} \hskip 30pt [\frac{\text{personen}}{m^2}]
\end{equation}
\newpage
\subsection{Personenfluss}
Der Personenfluss wird als der mittlere Nettozufluss an Trajektorien in einem Patch über die letzten $p$ (Schrittweite) Frames definiert. Um in jedem Frame zu- oder abfließende Trajektorien zu erfassen, wird ein Fluss-Akkumulator $F_{acc}$ eingerichtet, der den Beitrag +1 für eine in den Patch Eintretende und den Beitrag -1 für eine aus dem Patch austretende Trajektorie erhält. Dazu wird in jedem Frame die Differenz der Kardinalität der Menge S an Trajektorienspitzen und der Kardinalität der Menge $K$ an Trajektorienknoten (Position der Trajektorien im vorherigen Frame $j-1$) im Patch gebildet und über $p$ Frames gemittelt.

%Dazu werden die Trajektorien der aktuellen (ggf. gefilterten) Trajektorienliste in einer neuen, mit 0 initialisierten Bildmatrix $K[x,y]$ an ihrer Position im vorherigen Frame mit 1 markiert(Trajektorienknoten). Trajektorienspitzen S inkrementieren den Fluss im Patch, die Trajektorienknoten K dekrementieren ihn. Damit ergibt sich im Fall einer Grenzüberschreitung einer Trajektorie ein Beitrag +1 im Zielpatch und ein Beitrag -1 im Ursprungspatch.

\begin{equation}
F_{acc} = c_K\cdot (|S| - |K|) \hskip 30pt [\frac{\text{personen}}{\text{frame}}]
\end{equation}
\vskip 5pt
\begin{flushleft}
für alle verwendeten Positionen $P(x,y)\in S\cup K$ gilt: $x,y\in W(w,h)$
\end{flushleft}
\vskip 3pt
Der mittlere Nettozufluss über die letzten $p$ Frames beträgt somit:

\begin{equation}
    F = \frac{1}{p}\sum_{i=0}^{p-1}F_{acc,j-i} \hskip 30pt [\frac{\text{personen}}{\text{frame}}]
\end{equation}
\vskip 10pt
Wendet man auf $F$ einen zeitlichen gleitenden Mittelwert über M Frames an, um Rauschen zu unterdrücken, erhält man den mittleren Personenfluss $\bar{F_j}$.

\begin{equation}
\bar{F}_j = \frac{1}{m}\sum_{i=0}^{m-1}F_{j-i} \hskip 30pt [\frac{\text{personen}}{\text{frame}}]
\end{equation}
\vskip 10pt
Wird ein betragliches Maximum $F_{max}$ dieses mittleren Personenflusses algorithmisch gelernt, kann ein relativer Fluss angegeben werden:

\begin{equation}
F_{rel} = \frac{\bar{F}_j}{F_{max}} \hskip 30pt [\%]
\end{equation}
mit $F_{rel}\in \{0..1\}$
\vskip 10pt
Analog zur Personenzahl kann bei Bekanntheit von $N_{abs}$ ein manuell definierbarer Normierungsterm für den Personenfluss erhalten werden:
\vskip 5pt
\begin{equation}
    \bar{F}_{scal} = \bar{F}_j \cdot \frac{N_{abs}}{N_{max}} \hskip 30pt [\frac{\text{personen}}{\text{frame}}]
\end{equation}

\newpage
\subsection{Dynamik}
Zur Ableitung eines aussagekräftigen Dichtefaktors wird zusätzlich zur Personenzahl eine Kennzahl benötigt, die die vorherrschende Dynamik im Patch beschreibt. Dazu wird eine mittlere Merkmalsgeschwindigkeit im Patch eingeführt. Es wird erneut die Menge $S$ an Trajektorien betrachtet, deren Spitzen (aktuelle Positionen) sich innerhalb des Bildausschnitts befinden. Es werden Vektorbeträge für jede dieser Trajektorien nach folgendem Prinzip berechnet, um die Bewegungen über die letzten $q$ (Schrittweite) Frames zu erfassen. Im Fall, dass eine betrachtete Trajektorie noch keine $q$ Frames existiert, wird die Trajektorie aussortiert:
\begin{align}
\text{für} j-q \geq 0&: \hskip 10pt \abs{\vec{b_j}}(x,y) = \sqrt{[P_{j}(x) - P_{j-q}(x)]^2 + [P_{j}(y) - P_{j-q}(y)]^2} \notag \\
\text{sonst}:
\end{align}
\begin{flushleft}
für alle verwendeten Positionen $P(x,y)\in S$ gilt: $x,y\in W(w,h)$\vskip 5pt
Hier bezeichnet $P_j(x)$ die x-Position der Trajektorienspitze im aktuellen Frame j. $P_{j-q}(x)$ bezeichnet die x-Position der Trajektorienspitze vor $q$ Frames.
\end{flushleft} \vskip 5pt
Die mittlere Merkmalsgeschwindigkeit $B_j$ eines Bildausschnitts erhält man nun durch Mittelung über alle Vektorlängen $\abs{\vec{b}}_j(x,y)$ innerhalb des Patchs:

%\begin{equation}
%B_j = \frac{1}{|Z|}\sum_{u,v}\abs{\vec{b}}_j(x+u,y+v) \hskip 30pt [\frac{\text{px}}{q\cdot \text{frame}}]
%end{equation}
\begin{equation}
B_j = \frac{1}{|Z|}\sum_{k=0}Z_k \hskip 30pt [\frac{\text{px}}{q\cdot \text{frame}}]
\end{equation}

\begin{flushleft}
für alle verwendeten Vektorlängen $\abs{\vec{b}}_j(x,y)\in Z$ gilt: $x,y\in W(w,h)$.\vskip 5pt
Hier beschreibt $Z$ die Menge der berechneten Vektorlängen $\abs{\vec{b_j}}(x,y)$.\vskip 5pt
Um Rauschen zu unterdrücken, wird diese Kennzahl noch zeitlich gemittelt:
\end{flushleft}
\begin{equation}
\bar{B}_j = \frac{1}{m}\sum_{i=0}^{m-1}B_{j-i} \hskip 30pt [\frac{\text{px}}{q\cdot \text{frame}}]
\end{equation}
\vskip 5pt
Bei längerer Beobachtung der mittleren Merkmalsgeschwindigkeit in einer Testsequenz wird deutlich, dass sie in einem nicht vollgestauten (also unkritischen) Bildausschnitt immer oberhalb eines Schwellwertes liegt (nach Einschwingen des Systems). Dieser Schwellwert kann festgelegt werden, indem \zb ein Bildausschnitt betrachtet wird, der stetig vollgestaut wird und damit eine abnehmende mittlere Merkmalsgeschwindigkeit aufweist. Der Benutzer liest diese Messgröße, kurz nachdem er die Situation als kritisch einstuft, ab und trägt sie als Schwellwert $B_S$ ein. Der Schwellwert sollte, sobald kein Zu-/Abfluss in/aus dem Patch mehr existiert, schnell erreicht werden. Mit der Festlegung des Schwellwerts kann nun eine relative mittlere Merkmalsgeschwindigkeit $B_{rel}$ berechnet werden:
\newpage
\begin{equation}
    B_{rel}=\frac{B_S}{\bar{B}_j} \hskip 30pt [\%]
\end{equation}
mit: $B_{rel}\in \{0..1\}$
\vskip 10pt
Der Trägheitsfaktor $B_{rel}$ ist eine Art Staufaktor, der hohe Werte für geringe Dynamiken im Patch annimmt. Eine kritische Situation, wie beispielsweise ein Stau, liegt aber nur dann wirklich vor, wenn sich viele Personen im Bildausschnitt befinden und eine geringe Dynamik in Kombination vorliegt.

\subsection{Dichtefaktor}
Schlussendlich kann aus den zuvor abgeleiteten Größen ein Dichtefaktor definiert werden, der nur auf hohe Personenzahlen und hohe Trägheit in Kombination reagiert und hohe Werte annimmt.\vskip 5pt
\begin{equation}
\text{dFactor} = N_{rel}\cdot B_{rel} \hskip 30pt [\%]
\end{equation}
mit: $\text{dFactor}\in \{0..1\}$
\vskip 10pt

Der definierte Dichtefaktor funktioniert nur dann zuverlässig, wenn die Auflösung hoch genug ist, um Trajektorien von vereinzelten Personen zu erstellen. Treten Personengruppen hingegen nur noch als Anhäufung gleich heller Pixel auf, sind die Personen nicht separat erkennbar und die Personenzählung wird immer zu niedrig ausfallen, wobei gleiches für den Dichtefaktor gilt. Dieser Dichtefaktor macht weniger eine Aussage über eine Dichte in $[\frac{\text{personen}}{m^2}]$ als über den aktuellen Grad der Gefährdung der Besucher im betrachteten Bildausschnitt. Denn für hohe Dichtefaktoren, die kritischen Situationen entsprechen, ist sowohl die Personenzahl als auch die Trägheit im Patch hoch.

\subsection{Darstellung der Messgrößen - Aufbereitung}
Das Eingangsbild wird vom implementierten Verfahren wahlweise mit angezeigten Messwerten und eingezeichneten Bildausschnitten manipuliert. Außerdem können Messgrößen für einen genaueren Überblick durch Heat Maps dargestellt werden. Heat Maps sind farbkodierte Karten (ähnlich einer Wärmeverteilungskarte). Dabei wird das Bild zunächst in ein Grauwertbild konvertiert, damit es selbst keine Farbinformation mehr enthält. Dann wird für jeden Pixel eine neue Farbe im HSV-Farbraum bestimmt. Im HSV (hue, saturation, value)-Farbmodell kann der berechntete Grauwert an der Pixelposition als Helligkeitswert value eingesetzt werden, während die Farbsättigung (saturation) auf den Maximalwert gesetzt wird. Der Farbton (hue) wird dann auf den Wertebereich einer Messgröße skaliert und je nach Betrag färbt sich der Bildausschnitt mit zunehmender Messgröße von cyanblau bis dunkelrot. Zur Skalierung der Messgröße ist ein Training mit einer Testsequenz erforderlich. Dabei werden Maximal-/Minimalwerte der Messgröße vom Verfahren erlernt.
\newpage
Diese können in Konfigurationsdateien gespeichert und wiederverwendet werden. Hier wurde der gemessene Dichtefaktor zunächst auf einen Wertebereich von ${0..1}$ und anschließend auf dem Farbenkreis auf die Farben von Cyanblau bis Dunkelrot skaliert. Im HSV-Farbmodell entspricht dies einem zahlenmäßigen Farbton von 180-359. Formelmäßig kann man den Farbton hue so ausdrücken, falls $X_{rel}$ allgemein eine relative Messgröße beschreibt:

\begin{equation}
    \text{hue} = X_{rel}\cdot(359-180) + 180
\end{equation}
\vskip 10pt
Ein Beispiel für die Darstellung einer Heat Map erkennt man in Abbildung \ref{heatmap}:
\vskip 10pt
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.8\textwidth]
    {images/dummy.png}
  }
  \caption{Darstellung des Dichtefaktors als Heat Map \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
  \label{heatmap}
\end{figure}

Weil der Dichtefaktor aus der relativen Personenzählung und dem relativen Trägheitsfaktor abgeleitet wird, kann er nur für hohe Personenzahlen hohe Werte annehmen. Die maximale Personenzahl wird vom Verfahren in einer Lernphase statistisch geschätzt und tritt somit nur in einem voll befüllten Patch auf. Der abgeleitete Dichtefaktor beschreibt die Dichte der Personen im Bezug zur Fläche des Bildausschnitts. Personen, die zwar dicht aneinander stehen (siehe Abbildung \ref{heatmap} - lila), jedoch den Bildausschnitt nicht ausfüllen, erhalten keinen hohen Dichtefaktor. In einem solchen Fall würden aber in einem realen Szenario die Menschen sich instinktiv innerhalb des Bildausschnitts verteilen, weswegen eine Erkennung einer solchen Situation auch unerwünscht wäre.
\newpage
Nur im Fall einer hohen Personenzahl und zusätzlich hoher Trägheit im Patch nimmt der Dichtefaktor hohe Werte an, wie in Abbildung \ref{heatmap} im dunkelrot gefärbten Bildausschnitt zu erkennen ist. Ist die Erkennung von kleineren, dicht gedrängten Personengruppen erwünscht, können die Bildausschnitte auch verkleinert werden, damit weniger Personen den Bildausschnitt bereits ausfüllen. Dies birgt allerdings die Gefahr für häufige Fehlalarme.

\section{Konfigurierung}
\label{sec:konfig}
Der folgende Abschnitt befasst sich mit geplanten Ablaufschritten zur Konfigurierung des erstellten Softwaremoduls. Für optimale Ausnutzung des Funktionsumfangs müssen Grenzwerte geschätzt, Kameraparameter geladen, sowie Evaluations-, Filter- und Darstellungseinstellungen vorgenommen werden. Die Konfigurierung muss für jede Kamera/Sequenz einmal vorgenommen werden und kann dann gespeichert und wiederverwendet werden. Der Konfigurationsdialog des Softwaremoduls ist in Abb. \ref{konfig} dargestellt
\vskip 10pt
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.65\textwidth]
    {images/dummy.png}
  }
  \caption{Konfigurationsdialog des Softwaremoduls}
  \label{konfig}
\end{figure}

%\subsection{Plugin: OCV\_Corner Tracker}
%Der vorgeschaltene Corner Tracker, wie in Kapitel \ref{chap:grund} Abschnitt \ref{sec:trajektorien} beschrieben, muss zunächst konfiguriert werden. Dazu wird das Ausgangsbild dieses Plugins betrachtet, in dem die Detektionen des Harris Corner Detectors als Punkte eingezeichnet werden können. Die Einstellung "`maxCorners'', die die maximale Anzahl an detektierten Ecken angibt, wird idealerweise auf einen hohen Wert gesetzt(\zb $10000$). Nun wird das "`QualityLevel'', das die minimale Qualität der detektierten Ecken angibt, stetig vermindert, bis die im Bild zu sehenden Personen im Durchschnitt mindestens eine Eckendetektion/Person erhalten. Dieses Quality Level wird eingestellt und nicht mehr verändert, solange keine neue Szene verwendet wird.

\subsection{Patch-Map Dimension}
Um die Patch Maps den Wünschen des Anwenders anzupassen, kann Höhe und Breite eines Bildausschnitts im Konfigurationsdialog gewählt werden. Anschließend muss das aktuelle Bild nach dem Betätigen eines Reset-Button neu geladen werden, um die Änderungen zu aktualisieren. Passen die gewählten Patches nicht gradzahlig ins Ausgangsbild, werden über den Rand überlappende Bildausschnitte am Rand erstellt. Wie bereits in Abschnitt \ref{sec:sig} erwähnt, können die Patches wahlweise aufbereitet mit Messwerten oder Heat Maps eingezeichnet werden. Dieses aufbereitete Ausgangsbild wird vom Nutzer schließlich vereinfacht analysiert.

\subsection{Geschwindigkeitsschwelle}
Zur angepassten Unterscheidung von statischen Merkmalen zu Bewegten, wird eine konfigurierbare Geschwindigkeitsschwelle in $[\frac{\text{px}}{\text{frame}}]$ festgelegt. Zur Verwaltung (d.h. Erstellung und Fortführung) von Trajektorien werden nur Merkmale betrachtet, die oberhalb dieser festgelegten Geschwindigkeitsschwelle liegen. Statische Merkmale wie Ecken von Häusern oder Gegenständen werden somit weitgehend gefiltert, jedoch auch statische Personen. Eine geeignete Geschwindigkeitsschwelle erhält man empirisch, indem man zunächst den Suchradius vorübergehend auf $0$ setzt, um ein Verknüpfen der Trajektorien weitgehend zu verhindern. So entstehen ständig neue Trajektorien an detektierten Ecken von Personen ohne unberechenbare Verknüpfungen. Anschließend vermindert man die Geschwindigkeitsschwelle stetig und lässt sich die Spitzen (aktuelle Positionen) der Trajektorien in der laufenden Bildsequenz einzeichnen. Die bewegten Personen im Bild sollten im Durchschnitt mindestens eine Trajektorie pro Person erhalten. Die höchste Geschwindigkeitsschwelle für die dies zutrifft, wird eingestellt.

\subsection{Maximale Verweilzeit}
Um zu erlauben, dass Trajektorien für eine gewisse Frame-Dauer statisch auf ihrer Position verharren, wird eine maximale Verweilzeit in $[\text{frame}]$ gewählt. Dazu wird bestimmt, ob in der Umgebung einer Trajektorien-ID Merkmale oberhalb der Geschwindigkeitsschwelle liegen. Wenn ja, ist diese Trajektorie redundant, weil die Trajektorie fortgeführt wird. Wird kein solches Merkmal gefunden, kann die Trajektorie nicht fortgeführt werden und die Position der ID wird in einer Matrix (Stop-and-Go-Map) mit der Bildgröße inkrementiert, sofern der Wert die maximale Verweilzeit noch nicht erreicht hat. Sollte die Toleranz bereits erreicht sein, wird die Position in der Stop-And-Go-Map zurück auf 0 gesetzt und die Trajektorie entfernt.

\newpage

\subsection{Mittelungsdistanz}
Wie in Abschnitt \ref{sec:sig} behandelt, werden zeitliche Mittelungen im entwickelten Verfahren als gleitender Mittelwertfilter ausgeführt. Dazu kann im Konfigurationsdialog eine Mittelungsdistanz $m$ festgelegt werden, die die Ordnung des Filters angibt, also über wieviele Frames der Mittelwertfilter ausgeführt wird. Meist eignet sich eine Mittelungsdistanz von $10-20$ Frames, um Rauschen weitgehend zu unterdrücken.

\subsection{Suchradius}
Wie bereits in Kapitel \ref{chap:grund} Abschnitt \ref{sec:trajektorien} behandelt, werden nur gefundene Merkmale, die oberhalb der Geschwindigkeitsschwelle liegen, behandelt. Wird ein Merkmal in der Umgebung einer ID in der ID Map gefunden, wird diese, bereits vorhandene, Trajektorie mit dem gefundenen Merkmal verknüpft und fortgeführt. Zur Suche nach Merkmalen in der Nachbarschaft ist die Definition eines Suchradius in $[\text{px}]$ nötig, innerhalb dessen eine Suche nach Merkmalen/IDs vorgenommen wird. Dieser kann global im Konfigurationsdialog festgelegt werden. Einen geeigneten Suchradius für die Merkmale erhält man empirisch, indem man zunächst beim kleinsten Wert (1) beginnt und sich die Trajektorien, nach gewisser Dauer, in einem bestimmten Frame der Testsequenz einzeichnen lässt. Dies wiederholt man mit steigenden Radien, bis man durchgängige, lange Pfade erkennen kann. Der niedrigste Radius für den dies zutrifft, wird als Suchradius gewählt. Mit weiter steigenden Radien erscheinen die Pfade als eckiger/zackiger, weil die Merkmale großzügiger mit den ID's verknüpft werden. Damit ergibt sich eine höhere Toleranz für weit von der ID entfernte Merkmale, die möglicherweise schon zu einer anderen Person gehören und damit ein individuelles Tracking maßgeblich erschweren. Wechselt eine Trajektorie zwischen verschiedenen Personen, entstehen dabei Zacken. Gezeichnete Trajektorien werden in Abbildung \ref{trajektorien_drawn} für verschiedene Suchradien dargestellt.

\newpage
\begin{figure}[h]
\centering
    \includegraphics[width=0.49\textwidth]{images/dummy.png}
    a) $r=1$

    \includegraphics[width=0.49\textwidth]{images/dummy.png}
    b) $r=2$

    \includegraphics[width=0.49\textwidth]{images/dummy.png}
    c) $r=5$

\caption{Gezeichnete Trajektorien für versch. Suchradien}
\label{trajektorien_drawn}
\end{figure}
\newpage

Man erkennt hier, dass für kleinere Radien (siehe a)) die Trajektorien fragmentiert wirken. Erhöht man den Radius (vgl. b)), funktioniert das Verknüpfen der Pfade zuverlässiger und man erkennt durchgängige Trajektorien. Wird der Radius wie in c) noch weiter erhöht, entstehen Zacken in den Pfaden, weil die Trajektorien großzügiger verknüpft werden und eine höhere Toleranz für Merkmale gesetzt wird, die von der ID weit entfernt liegen. Hier wurde der Radius 2 als Suchradius gewählt und gespeichert.

\subsection{Abstands-Radius}
Ebenso wie es zu vermeiden gilt, dass eine Trajektorie zwischen verschiedenen Personen wechselt, gilt es zu vermeiden, dass im zeitlichen Verlauf der Eingangssequenz im Durchschnitt unterschiedlich viele Trajektorien pro Person geführt werden. Dazu wird ein Abstands-Radius eingeführt, innerhalb dessen, nach der Evaluation eines Frames, die Nachbarschaft in der ID Map nach anderen IDs abgesucht wird. Wird eine ID gefunden, wird die länger bereits vorhandene Trajektorie behalten und die Kürzere entfernt, sofern ein Unterschied besteht. Damit unterdrückt man neue, kurze Trajektorien, die durch Rauschen entstehen und sorgt zudem dafür, dass die Trajektorien und damit auch die detektierten Personen zusammen eine große Fläche einnehmen müssen, um eine hohe Zahl an Trajektorien zu ermöglichen. Dies ist erwünscht, weil sonst beliebig viele Trajektorien an einer Person entstehen können, auch wenn diese alleine keine große Fläche einnimmt, was zu einer zu hohen Personenzählung führt. Eine große Zahl an Personen nimmt eine große Bildfläche ein und infolgedessen sollten auch nur in diesem Fall viele Trajektorien erlaubt sein. Durch Einführung des Abstands-Radius vermindert man Probleme, die durch gegenseitige Verdeckung der Personen entstehen. In weniger dichten Gebieten nimmt eine Person in der Regel eine größere Fläche ein, weil sie nicht durch andere Personen verdeckt wird. So können allerdings auch mehr Trajektorien an dieser Person entstehen, weil mehr Kanten dieser Person sichtbar sind. Nimmt die Personendichte an diesem Ort im Verlauf der Bildsequenz zu, wird diese Person mehr und mehr von Anderen verdeckt und es entstehen im Durchschnitt weniger Trajektorien pro Person, weswegen trotz der deutlich erhöhten Personenzahl, etwa die gleiche Zahl an Trajektorien und damit die gleiche Personenzählung wie zuvor vorliegt. Führt man einen Abstands-Radius ein, können nur bedingt viele Trajektorien an der unbedeckten Person entstehen und die Zahl an Trajektorien pro Person ist sowohl im Zustand geringer als auch hoher Personendichte in etwa gleich, wobei sich im Zustand hoher Dichte folglich, wie gewünscht, eine deutlich erhöhte Personenzählung ergibt.

\subsection{Ortsbezogener Ausgleich}
Bei Einführung dieses Abstands-Radius muss bedacht werden, dass die Ursprungsorte der Pixel im Eingangsbild, je nach Perspektive, unterschiedliche Abstände zum Kameraobjektiv aufweisen. Dieser Effekt wirkt sich umso stärker aus, je weiter man von der Vogelperspektive (Draufsichtperspektive) abweicht. 
\newpage
Demnach ist der Pixelabstand $d(x,y)$ in $[\frac{m}{\text{px}}]$ in entfernteren Bildregionen höher, weil solche Bildregionen kleiner skaliert abgebildet sind. Dies wird in Abblidung \ref{Skalierung} verdeutlicht:
\vskip 10pt
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.6\textwidth]
    {images/dummy.png}
  }
  \caption{versch. Bildregionen mit unterschdl. Abständen zum Kameraobjektiv}
  \label{Skalierung}
\end{figure}

Deshalb sind Menschen in weiter entfernten Bildregionen auch weniger Pixel hoch und voneinander entfernt. Sie nehmen damit, bei gleichbleibender Personenzahl, eine geringere Bildfläche ein, wobei weniger Trajektorien pro Person erfasst werden, weil die Kantenlänge in $[\text{px}]$ verringert ist. Der Abstands-Radius sollte demnach in entfernteren Bildregionen sehr klein sein, um die ohnehin schon geringe Zahl an Trajektorien pro Person nicht weiter zu verringern. Dieser kleinste Abstands-Radius kann im Konfigurationsdialog festgelegt werden. Je näher sich das Objekt an der Kamera befindet, desto größer wird es abgebildet und der Abstands-Radius sollte mit einem entsprechenden (nichtlinearen) Skalierungfaktor vergrößert werden. Infolgedessen sollte idealerweise eine Skalierungsfunktion in Form einer Bildmatrix vorliegen, in der der Pixelabstand (die Skalierung) in $\frac{m}{\text{px}}$ relativ zum größten Pixelabstand (zur kleinsten Abbildung) in $\frac{m}{\text{px}}$ im Eingangsbild für alle Bildkoordinaten $x,y$ eingetragen ist. Mit Skalierungsfunktionen können Abstands-Radius oder andere Evaluations- und Filtereinstellungen sowie Grenzwerte skaliert werden. Eine solche Skalierungsfunktion kann leicht aus einer vollständig dokumentierten Kamerakalibrierung generiert werden. Alternativ kann auch eine Abstands-Radius Karte erlernt werden. Dazu muss der Abstands-Radius in $\text{px}$ am unteren und oberen Bildrand als größter und kleinster Radius eingesetzt werden. Die Pixel zwischen oberer und unterer Bildzeile erhalten interpolierte Abstands-Radien zwischen den beiden Vorgegebenen. Dabei wird davon ausgegangen, dass die Entfernung vom Kameraobjektiv vom Unteren zum oberen Bildrand hin zunimmt, was oft in guter Näherung zutrifft. Wird keine Abstands-Radius-Karte vorgegeben wird der kleinste Abstands-Radius global verwendet. Für den kleinsten Abstands-Radius kann die Personenhöhe (in $\text{px}$) für den maximal zulässigen Abstand von Person zu Kamera eingesetzt werden (minimale Auflösung einer Person).
\newpage
Alternativ kann auch eine geometrische Kalibrierung vorgenommen werden, um den Einfluss der unterschiedlichen Abstände der Bildregionen vom Kameraobjektiv zu verringern. Dazu werden bestimmte Bildregionen neu skaliert, so dass ursprünglich größer skalierte Gebiete, in denen also eine Person mit deutlich mehr Pixeln aufgelöst wird, eine geringere Auflösung erhalten, als Gebiete in denen die Personen ohnehin schon klein skaliert sind. Dabei verzerrt sich das Bild geometrisch. Da nun in allen Bildregionen eine Person mit näherungsweise gleich vielen Pixeln aufgelöst wird, werden nun Inhomogenitäten in der durchschnittlichen Zahl an Trajektorien pro Person vermieden. Dies wird in Abbildung \ref{geokal} dargestellt:

\begin{figure}[H]
\centering
  \begin{minipage}{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
  \end{minipage}
  \begin{minipage}{0.45\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
  \end{minipage}
\caption{Verfahren: geometrische Kalibrierung}
\label{geokal}
\end{figure}

\subsection{Distanzschätzung}
Sind Aperturwinkel $\omega$, Tilt-Winkel $T$, sowie die Höhe $h$ der Kamera bekannt, lassen sich näherungsweise die Abstände der Ursprungsorte von oberer und unterer Bildzeile vom Kameraobjektiv bestimmen. Dazu betrachtet man folgende Skizze in Abbildung \ref{distanz}:

\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.6\textwidth]
    {images/dummy.png}
  }
  \caption{Distanzschätzung der Bildzeilen}
  \label{distanz}
\end{figure}

Wendet man die trigonometrischen Gesetze an, wird hierbei deutlich:
\begin{align}
    d_1 &\approx \frac{h}{\sin{(T + \frac{\omega}{2})}} \notag \\
    d_2 &\approx \frac{h}{\sin{(T - \frac{\omega}{2}})}
\end{align}

Die Abstände der Ursrpungsorte von oberer und unterer Bildzeile vom Objektiv können mit diesen Formeln näherungsweise berechnet werden. Um eine Patch Map zu erhalten, die genäherte Distanzen der Bildauschnitte enthält, wird die Distanz zeilenweise zwischen oberer und unterer Bildzeile interpoliert und innerhalb der Patches gemittelt.

Liegt eine vollständig dokumentierte Kamerakalibrierung vor, ist dieses Verfahren redundant, weil aus der Kalibrierung der Abstand jedes Ursprungsorts im Eingangsbild berechnet werden kann. Dies ist jedoch für die Bilddaten, die für diese Arbeit vorliegen, nicht gegeben.

%\subsection{Trainieren von Extremwerten}
%Um relative Messgrößen zu erhalten, müssen Extremwerte von bestimmten Messgrößen trainiert werden. Relative Messungen sind nötig, um Angaben in $\%$ zu erhalten und damit Skalierungen der Messgrößen beispielsweise auf den Farbbereich von Heat Maps vornehmen zu können. Nach aktuellem Stand können diese Extremwerte automatisch trainiert werden:

%\begin{itemize}
    %\item Maximum der Personenzählung
    %\item Maximum/Minimum des Personenflusses
    %\item Maximum des Dichtefaktors
%\end{itemize}

%Zum Trainieren der Extremwerte können Frame Intervalle(\zb $0-50$) festgelegt werden, innerhalb derer das Training vorgenommen wird.

%Alternativ können diese Extremwerte auch im Konfigurationsdialog eingetragen werden, falls diese schon bekannt sind. Der Schwellwert der mittleren Merkmalsgeschwindigkeit über die letzten 10 Frames muss manuell eingetragen werden und wird meist aus dem Verhalten der Messgröße erschlossen. 
%\newpage
%Idealerweise sollten Patch Maps vorliegen oder generiert werden, in denen Extremwerte für jede Messgröße skaliert auf jeden Bildausschnitt eingetragen sind. Die Extremwerte für Personenzahl und andere Messgrößen sowie Evaluations- und Filtereinstellungen sollten über das Eingangsbild hinweg örtlich variieren. Eine klein skalierte Bildregion enthält beispielsweise klein skalierte Personen, weswegen sich auch viele Personen gleichzeitig in ihr befinden können. In einer größer skalierten Bildregion sind auch die Personen größer skaliert und es passen insgesamt weniger Personen in diesen Bildausschnitt. Die Skalierung der Extremwerte auf die verschieden weit entfernten Bildausschnitte kann mit der zuvor beschriebenen Karte, in der an jedem Ort $x,y$ der Pixelabstand relativ zum größten Pixelabstand im Eingangsbild eingetragen ist, durchgeführt werden.

%Nach dem Training/der Eingabe der Extremwerte können diese in Konfigurationsdateien gespeichert und wiederverwendet werden.

\subsection{Filtereinstellungen}
Leichte Kamerabewegungen lassen Trajektorien auch an Ecken von Gegenständen entstehen. Zusätzlich nehmen Wasser- und Windbewegungen (\zb Blätter) Einfluss. Diese Bewegungen sind meist klein, können aber dennoch die gewählte Geschwindigkeitsschwelle betragsmäßig überschreiten. Erwünscht sind jedoch nur Trajektorien an Personen. Um solche fälschlicherweise entstandenen Trajektorien zu entfernen, können, vor der Umsetzung in Messgrößen, Filter auf die Trajektorien angewandt werden.

Nachfolgend werden die implementierten Filter aufgezählt:

\begin{itemize}
    \item minimale Vektorlänge der Trajektorie über die letzten $m$ Frames$\Rightarrow$kleine Kreisbewegungen werden gefiltert, exzentrische, gestreckte Bewegungen werden behalten
    \item minimale Eintragslänge der Trajektorie$\Rightarrow$kurze Trajektorien werden gefiltert, robuste, lange Trajektorien werden behalten
    \item maximale Eintragslänge der Trajektorie$\Rightarrow$besonders lange Trajektorien werden gefiltert 
\end{itemize}

### 2018-12-21 thesis_sebastian_bek/content/4-Ergebnis.tex
Initial commit

\addtocontents{toc}{\protect\newpage}
\chapter{Evaluation}
\label{chap:erg}

Das folgende Kapitel befasst sich mit der Auswertung des entwickelten Verfahrens. Dabei soll der erstellte Algorithmus erprobt und hinsichtlich seiner Zuverlässigkeit und Leistung untersucht werden. Um eine Erprobung vorzunehmen, müssen zunächst Referenzdaten durch manuelle Annotationen generiert werden. Eine solche manuelle Annotation, wird fachsprachlich als "`Ground Truth'' (dt.: Grundwahrheit) bezeichnet. Das, im Rahmen dieser Arbeit entwickelte, Verfahren erhält die Bezeichnung "`MFFT'' (multi frame feature tracking).

\section{Testdaten}
\label{test:data}

\subsection{3D-Animation: AGORASET/Dispersion \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
Zum Testen der generierten Messungen des entwickelten Verfahrens wird eine vordefinierte, bekannte Szene verwendet, die in Form einer künstlich generierten Bildsequenz als Datensatz vorliegt. Die Szene zeigt initial gleichmäßig verteilte Personen in einer Halle aus der Vogelperspektive. Im Verlauf der Bildsequenz stäuben diese Personen von einem scheinbaren Ausgangspunkt in alle Richtungen auseinander. Nach einiger Zeit stauen sich die Personen an den Rändern der Halle. Die Sequenz besteht aus 110 Einzelbildern. In Abbildung \ref{crowd_burst} werden beispielhaft Ausschnitte der beschriebenen Sequenz gezeigt:

\begin{figure}[H]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    a) Bild Nr. $01$
    \label{frame01}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    b) Bild Nr. $30$
    \label{frame30}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    c) Bild Nr. $87$
    \label{frame87}
  \end{minipage}
\caption{Ausschnitte der Testsequenz Dispersion \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
\label{crowd_burst}
\end{figure}

Das AGORASET ist ein Datensatz, der entwickelt wurde, um Algorithmen auf dem Gebiet der Analyse von Menschenmengen zu testen. Er besteht aus insgesamt 8 Szenen, die alle künstlich generiert wurden, um Szenarien mit interessanten Personengruppenformationen/\\-bewegungen zu erhalten.

\subsection{Videosequenz: UCF\_CrowdsDataset/Business-Getümmel \cite{AliS07}}
Zusätzlich zur Evaluation der Messgrößen mittels einer künstlich generierten Bildsequenz soll die Evaluation mit realen Daten durchgeführt werden. Dazu wird eine Szene des UCF\_CrowdsDataset (siehe \cite{AliS07}) verwendet, in der sich Personen ungerichtet im Alltagsstress auf einem Platz bewegen. Die Sequenz besteht aus 255 Einzelbildern. In Abbildung \ref{business} werden Ausschnitte dieser Testsequenz gezeigt:

\begin{figure}[H]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    a) Bild Nr. $01$
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    b) Bild Nr. $131$
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    c) Bild Nr. $255$
  \end{minipage}
\caption{Ausschnitte der Testsequenz Business-Getümmel \cite{AliS07}}
\label{business}
\end{figure}

Anhand dieser realen Szene sollen Messgrößen wie Dichtefaktor, Personenfluss-/zahl zusätzlich evaluiert werden. Wegen stark varriierender Personenzahl und Dichte in verschiedenen Bildausschnitten ist diese Sequenz geeignet, um die Messgrößen hinsichtlich ihrer Sensitivität und ihrem Verhalten zu beurteilen. Die Szene liegt in einer maximalen Auflösung von 480x360 vor, weswegen diese Szene in besonders kurzer Laufzeit ausgewertet werden kann.

Das UCF\_CrowdsDataset ist ein Datensatz mit aufgenommenen Bilddaten, der bewegte Menschenmengen in realen Szenarien zeigt. Er ist für die Erprobung von Algorithmen auf dem Gebiet der Analyse von Menschenmengen geeignet. Der Datensatz besteht insgesamt aus 30 Szenen, die aus öffentlichen Quellen (\zb BBC) zusammengetragen wurden.

\subsection{Videosequenz: UCF\_CrowdsDataset/Marathon \cite{AliS07}}
Um eine Testsequenz einzubeziehen, die extrem viele sowohl Bewegte als auch statische Personen enthält wird die Aufnahme eines Marathonlaufs verwendet. Die Sequenz liegt in Form einer Bildfolge mit einer Auflösung von 720x404 vor, weshalb dabei unter Ausführung des Verfahrens (MFFT) nur geringe Anforderungen an die Rechenleistung gestellt werden. In folgender Abbildung \ref{run} werden beispielhaft Ausschnitte der Videosequenz des Marathonlaufs dargestellt.

\begin{figure}[H]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    a) Bild Nr. $01$
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    b) Bild Nr. $31$
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    c) Bild Nr. $62$
  \end{minipage}
\caption{Ausschnitte der Testsequenz Marathonlauf \cite{AliS07}}
\label{run}
\end{figure}

Man erkennt hier deutlich, dass sich am Rande der Marathonstrecke stehende (statische) Personen befinden, während die Läufer sich bewegen. Daher wird diese Testsequenz dazu verwendet, die Trennung von statischen (HG) und nicht statischen (VG) Merkmalen zu evaluieren.

\subsection{Videosequenz: Hamburger Hafenfest 2014}
Um die Erfassung präziser Trajektorien in Personengruppen hoher Dichte zu evaluieren, wird das Szenario einer Massenveranstaltung verwendet. Dazu liegen Testsequenzen des Hamburger Hafenfests 2014 vor. Die Testdaten des Hamburger Hafenfests 2014 umfassen die Aufnahmen von einer in 80m Höhe montierten Kamera, die nach schräg links geneigt ist. Im Videobild ist hauptsächlich Wasser in Kanälen des Hamburger Hafens zu sehen. Während des Videos werden diese sporadisch von Schiffen befahren. 2 aufgehängte Brücken führen zur meist belebteren Hafenpromenade, die im Videobild teilweise sichtbar ist. Hier sind Buden und weiße Zelte zu erkennen, sowie Treppen und eine benachbarte Straße. Es liegen Aufnahmen von jeglichen Wetterlagen und Tageszeiten von dieser Kamera vor, um verschiedene Szenarien überprüfen zu können. In Abbildung \ref{hamburg:regen} - \ref{hamburg:sonne} erkennt man beispielhaft Aufnahmen der verschiedenen Wetterlagen:

\newpage
\begin{figure}[h]
  \centering  
    \begin{minipage}{0.9\textwidth} 
      \includegraphics[width=\textwidth]{images/dummy.png}\\
	\end{minipage}
    \put(0,-100){\textbf{{ \color{black}(a)}}}\\
    \begin{minipage}{0.9\textwidth} 
      \includegraphics[width=\textwidth]{images/dummy.png}
	\end{minipage}
    \put(0,-100){\textbf{{ \color{black}(b)}}}\\
  \caption{Testsequenz des Hamburger Hafenfests bei\\
  (a) Nebel; (b) Regen}
  \label{hamburg:regen}
\end{figure} 
\newpage
\begin{figure}[h]
  \centering     
    \begin{minipage}{0.9\textwidth} 
      \includegraphics[width=\textwidth]{images/dummy.png}\\
	\end{minipage}
    \put(0,-100){\textbf{{ \color{black}(c)}}}\\
    \vfill
    \begin{minipage}{0.9\textwidth} 
      \includegraphics[width=\textwidth]{images/dummy.png}
	\end{minipage}
    \put(0,-100){\textbf{{ \color{black}(d)}}}    
  \caption{Testsequenz des Hamburger Hafenfests bei\\
  (c) Sonne; (d) Sonnenuntergang}
  \label{hamburg:sonne}
\end{figure}
\newpage

Die Aufnahmen liegen in einer maximalen Auflösung von 1920x1082 vor. Wegen hoher Rechenzeit beziehungsweise nicht verfügbarer Rechenleistung werden jedoch zum Testen skalierte Aufnahmen mit einer Auflösung von 960x540 verwendet. Moderne Computer können ein Eingangsbild einer solchen Bildfolge unter Ausführung des entwickelten Verfahrens MFFT in weniger als 50ms verarbeiten.

\section{Testergebnisse}
\label{test:erg}

\subsection{3D-Animation: AGORASET/Dispersion \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
\label{dispersion_test}
Die Personen im Video sind mit einer Auflösung von 640x480 für eine Zählung ausreichend hoch aufgelöst. Zum Testen werden große Bildausschnitte mit einer Größe von 80x96 eingestellt und der Bildausschnitt gewählt, der den größten Wertebereich im Verlauf der Sequenz für die Personenzahl aufweist. Dieser Bildausschnitt ist am aussagekräftigsten für die Sensitivität und das Verhalten der Messgrößen (Personenzahl, -fluss, Dichtefaktor). Beim Testen werden nur relative Messgrößen betrachtet, weil, wie in Kapitel \ref{chap:imp} Abschnitt \ref{sec:sig} beschrieben, die Ableitung einer richtig skalierten absoluten Messgröße aus einer Relativen, durch Definition einer absoluten maximalen Personenzahl $N_{abs}$, jederzeit möglich ist. Diese wird entweder manuell oder durch theoretische Betrachtungen erhalten. Das entwickelte Verfahren wird, wie in Kapitel \ref{chap:imp} Abschnitt \ref{sec:konfig} beschrieben, mit einem geeigneten Suchradius konfiguriert. Die maximale Verweilzeit für die Trajektorien wird auf eine längere Verweilzeit als die Dauer der Bildsequenz gesetzt, um stehenbleibende Personen mitzuzählen. Außerdem wird ein globaler Abstands-Radius eingestellt, der Probleme abmindert, die durch Verdeckungen entstehen. Die Schrittweiten $p$ und $q$ werden in diesem Abschnitt auf den Wert 10 festgelegt.

\subsubsection{Personenzählung}
Die Personen innerhalb des Bildausschnitts wurden für die 110 Frames der Testsequenz manuell gezählt und die Ergebnisse notiert. Um die Analogie der manuellen Zählung zum entwickelten Verfahren zu bewahren, werden die gezählten Personenzahlen mit einer Mittelungsdistanz von $m$ Frames mit einem gleitenden Mittelwertfilter gemittelt. Die Mittelungsdistanz $m$ wird hier auf 10 Frames festgelegt. Anschließend werden die Messungen des Verfahrens MFFT für die mittlere Personenzahl für den Verlauf der Testsequenz exportiert und gespeichert. Beide Ergebnisse werden über den zeitlichen Verlauf der Sequenz aufgetragen und gegenübergestellt, wie in Abbildung \ref{zahl} dargestellt wird:
\newpage
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {images/dummy.png}
  }
  \caption{Graph: Verlauf der gemessenen/gezählten Personenzahl \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
  \label{zahl}
\end{figure}

Man erkennt hier, dass das Verhalten des erstellten Verfahrens (MFFT) mit dem Verhalten der Ground Truth korreliert. Im 1. Viertel der Sequenz erkennt man allerdings eine leicht erhöhte Zählung gegenüber der Ground Truth. Diese erhöhte Zählung tritt aufgrund von Verdeckungen auf, was auf dem Gebiet der Bildverarbeitung ein häufiges Problem darstellt. Dabei erhalten die Personen im Zustand geringer Personendichte (Beginn der Sequenz) deutlich mehr Trajektorien pro Person, weil eine frei stehende Person schlicht mehr Bildfläche einnimmt als eine Bedeckte, wodurch mehr Ecken (Harris Corners) sichtbar sind. Da das verwendete Verfahren die Spitzen der Trajektorien zählt, erhält man im Zustand geringer Personendichte eine zu hohe Personenzählung. Dieses Problem wird bereits vermindert, indem ein Abstands-Radius, wie in Kapitel \ref{chap:imp} Abschnitt \ref{sec:konfig} beschrieben, eingestellt wird. Um es jedoch gänzlich aufzuheben, muss Aufwand betrieben werden, der den Umfang der Arbeit weit überschreiten würde. Es könnten beispielsweise Merkmalspunkte mit gleicher Bewegungsrichtung und -geschwindigkeit zusammengefasst werden (siehe Clustering), um pro Person jeweils nur eine Detektion zu erhalten.

\subsubsection{Personenfluss}
\label{dispersion_fluss}
Das verwendete Verfahren zur Erfassung des Personenflusses stützt sich auf die Zählung von Trajektorienspitzen und -knoten (siehe Kapitel \ref{chap:imp} Abschnitt \ref{sec:sig}) über die letzten $p$ (Schrittweite) Frames. Die Schrittweite $p$ wird hier auf 10 Frames gestgelegt. Dabei stellt man fest, dass der so erhaltene Personenfluss aufgrund der hohen Dynamik in der Szene stark fluktuiert und erneut gemittelt werden muss. Ab einer Mittelungsdistanz $m$ von etwa 30 Frames erhält man eine robuste Schätzung. Dieses Verfahren kann in dieser Form nicht manuell annotiert werden, da der zeitliche Rahmen dieser Arbeit die Annotation ganzer Trajektorien, und dies wiederum für eine hohe Anzahl an Personen (Menschenmassen) nicht ermöglicht.
\newpage
Als Alternative wird die Differenz der manuell gezählten aktuellen Personenzahl und der Personenzahl vor $m$ (hier 30) Frames als absoluter Personenfluss über die letzten $m$ Frames verwendet. Wird dieser durch die Mittelungsdistanz 30 geteilt, erhält man den mittleren Personenfluss der letzten $m$ Frames, wie in Gleichung \ref{fluss_gt} dargestellt wird:

\begin{equation}
    \bar{F}=\frac{N_j-N_{j-m}}{m}
    \label{fluss_gt}
\end{equation}
\vskip 5pt
 Die so erhaltenen Messwerte werden über die Zahl der Frames aufgetragen und verglichen, wie in Abbildung \ref{fluss} dargestellt wird:
\vskip 10pt
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {images/dummy.png}
  }
  \caption{Graph: Verlauf des gemessenen/gezählten Personenflusses \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
  \label{fluss}
\end{figure}

Man erkennt, dass der vom entwickelten Verfahren abgeleitete Messwert für den Personenfluss zwar träge ist und eine Verzögerung aufweist, dem prinzipiellen Verhalten der Ground Truth aber folgt. Die so erhaltene Messgröße Personenfluss kann also als aussagekräftig betrachet werden. Die Verzögerung tritt auf, weil sich das Verfahren zum Generieren der Ground Truth vom entwickelten Verfahren unterscheidet. Die Verzögerung um 10 Frames entsteht, weil die Berechnung des Personenfluss im entwickelten Verfahren (MFFT) zunächst über eine Schrittweite von $p$ (hier 10) Frames berechnet wird. In manueller Form kann dieses Verfahren nicht ausgeführt werden, weswegen im Fall der Ground Truth diese Verzögerung ausbleibt.

\subsubsection{Dichtefaktor}
\label{dispersion_dichte}
Der, in dieser Arbeit, abgeleitete Dichtefaktor stützt sich auf eine Kombination aus der Messung eines relativen Trägheitsfaktors (in Form einer relativen Merkmalsgeschwindigkeit im Patch) und einer Personenzahl (siehe Kapitel \ref{chap:imp} Abschnitt \ref{sec:sig}). 
\newpage
Diese Messgrößen werden multipliziert um den Dichtefaktor zu erhalten. Da die manuelle Bestimmung der mittleren Merkmalsgeschwindigkeit der Personenbewegungen der letzten $q$ (hier 10) Frames den Umfang der Arbeit überschreiten würde, wird für diese Messgröße keine Ground Truth durch Annotationen generiert. Der Verlauf des ausgegeben Dichtefaktors wird hier mit dem Verlauf der rel. Personenzahl verglichen, um zu zeigen, wie der abgeleitete Dichtefaktor funktioniert und sich von der reinen Messung einer Personenzahl unterscheidet. Die vom Verfahren (MFFT) generierten Messwerte werden exportiert und gespeichert. Die so erhaltenen Ergebnisse werden über den zeitlichen Verlauf der Testsequenz aufgetragen und betrachtet (siehe Abbildung \ref{dichte}):
\vskip 10pt
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {images/dummy.png}
  }
  \caption{Graph: Verlauf des Dichtefaktors und der Personenzahl (MFFT) \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
  (mit eingezeichneter Farbcodierung der verwendeten Heat Map)
  \label{dichte}
\end{figure}
\bigskip

Man erkennt hier, dass die Verläufe korrelieren, jedoch zu Beginn starke Abweichungen auftreten. Zu Beginn der Bildsequenz (Testbereich 1) hat die rel. Personenzahl bereits die Hälfte ihres Maximalwerts erreicht, weil sich bereits einige Personen im Bildausschnitt befinden. In diesem Bereich existiert jedoch noch eine hohe Dynamik (keine Staus) im Patch, weswegen kleine Messwerte für den relativen Trägheitsfaktor und damit auch für den Dichtefaktor generiert werden. Gegen Ende der Sequenz (Testbereich 2) beginnen die Personen im Bildausschnitt sich zu stauen, weswegen die Dynamik im Patch stark abnimmt. Damit werden immer größere Werte für den relativen Trägheitsfaktor und damit auch für den Dichtefaktor generiert, weswegen in diesem Bereich der Dichtefaktor stark ansteigt und mit der relativen Personenzahl in Übereinstimmung kommt. Ab hier nimmt der relative Trägheitsfaktor den Maximalwert $1,0$ an, denn hier stimmen die Werte für Dichtefaktor und Personenzahl überein. Das Maximum des Dichtefaktors zu Beginn der Sequenz (Frame 10) entsteht durch den Einschwingvorgang der Messgröße relativer Trägheitsfaktor (vgl. Kapitel \ref{chap:imp} Abschnitt \ref{sec:sig}).
\newpage
Weil die mittlere Vektorlänge ($B_j$) über $q$ Frames im Patch erst dann Messwerte generiert, wenn bereits $q$ (hier 10) Frames vergangen sind und Diese noch zeitlich über $m$ Frames gemittelt wird, dauert es weitere $m$ (hier 10) Frames bis die mittlere Merkmalsgeschwindigkeit ($\bar{B}_j$) ihren hohen Betriebswert für die Dynamik annehmen kann (vgl. Fr. 20). Deswegen ist ihr Wert zu Beginn im Einschwingvorgang immer klein und erhöht sich dann auf den Betriebswert. Weil der relative Trägheitsfaktor ($B_{rel}$) als Schwellwert/Messwert ($\frac{\bar{B}_S}{\bar{B}_j}$) berechnet wird, wird man für kleine mittlere Merkmalsgeschwindigkeiten immer große relative Trägheitsfaktoren erhalten, was den Überschwinger zu Beginn der Sequenz erklärt.

\subsubsection{Darstellung als Heat Map}
Die Darstellung als Heat Map wird anhand des relativen Dichtefaktors in Abbildung \ref{heatmap_series} gezeigt. Dabei wird erneut die künstlich generierte Bildsequenz betrachtet, mit der die Messgrößen bereits evaluiert wurden. Der Verlauf des vom Verfahren (MFFT) ausgegebenen Dichtefaktors im Patch mit den Koordinaten (1|1) wird in Abbildung \ref{dichte} - blau dargestellt. Zusätzlich wird auf der Ordinate die Farbcodierung der verwendeten Heat Map angegeben. Dieser Verlauf wird auf den Wertebereich, den die Heat Map für den Farbton (hue) vorsieht (${180..359}$), skaliert. Hier werden exemplarisch Aufnahmen im Verlauf der Testsequenz Dispersion gezeigt, die mit Heat Maps aufbereitet werden. Die Heat Map stellt den Dichtefaktor dar. Kleine Werte für den Dichtefaktor ergeben die Farbtöne Cyan/Blau, hohe Werte ergeben die Farbtöne Lila/Rot. Wird dieser Dichtefaktor als Heat Map dargestellt, müssen, wie im Diagramm in Abbildung \ref{dichte} dargestellt, zu Beginn der Testsequenz Cyan-/Blautöne ($\text{hue}=180-250$) überwiegen. Gegen Ende (etwa ab Frame 70) der Testsequenz muss der Farbton in diesem Bildausschnitt ins Lila/Rote ($\text{hue}=300-359$) wechseln. 
\vskip 10pt
\begin{figure}[H]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    a) Bild Nr. 10
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    b) Bild Nr. 70
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{images/dummy.png}
    c) Bild Nr. 100
  \end{minipage}
\caption{Darstellung der Heat Map in der Testsequenz Dispersion  \cite{CourtyPRL2014} \cite{Allain2012ICPR}}
\label{heatmap_series}
\end{figure}

Zum Vergleich wird der Bildausschnitt mit den Koordinaten (1|1) betrachtet, da dieser Verlauf bereits in Abbildung \ref{dichte} dargestellt ist. Man erkennt, dass in Frame 10 (vgl. a)) Cyan-/Blautöne überwiegen, sich die Messgröße zunächst einschwingt und dann im Arbeitsbetrieb befindet. Im Diagramm steigt der Dichtefaktor nun bis zum Ende der Testsequenz über die Farbtöne blau und lila (vgl. b)) auf seinen Maximalwert an, was ebenfalls zu erkennen ist. Wird, wie in c), der Maximalwert erreicht, wird der Bildausschnitt dunkelrot eingefärbt.

\subsection{Videosequenz: UCF\_CrowdsDataset/Business-Getümmel \cite{AliS07}}
Die Testsequenz "`Business-Getümmel'' wird verwendet, um erneut die Messgrößen Dichtefaktor, Personenfluss/-zahl mit realen Daten zu evaluieren. Die Personen sind mit einer Auflösung von 480x360 wegen der nahen Perspektive ausreichend aufgelöst. Für die Evaluation wird das Eingangsbild in nur einen Bildausschnitt mit einer Auflösung von 480x360 eingeteilt, weil ohnehin nur wenige Personen (maximal 60 auf einmal) in der Testsequenz zu sehen sind und ein möglichst hoher Wertebereich für die Messgrößen erzeugt werden soll. In der zuvor verwendeten künstlich generierten Bildsequenz waren bis zu 80 Personen in einem einzelnen Bildausschnitt zu sehen. Dieser Test mit realen Daten soll, in Analogie zum zuvor durchgeführten Test in Abschnitt \ref{dispersion_test}, zur Evaluation des Verhaltens der Messgrößen ausgeführt werden. Erneut werden nur relative Messgrößen zum Testen verwendet, weil relative Messgrößen bei Bekanntheit eines Skalierungsfaktors jederzeit in richtig skalierte, absolute Messgrößen umgerechnet werden können (vgl. Abschnitt \ref{dispersion_test} und Kapitel \ref{chap:imp} Abschnitt \ref{sec:konfig}). Das Verfahren (MFFT) wird, wie in Kapitel \ref{chap:imp} Abschnitt \ref{sec:konfig} beschrieben, konfiguriert und ein geeigneter Suchradius sowie die Geschwindigkeitsschwelle eingestellt. Die maximale Verweilzeit wird nun auf eine kurze Dauer gesetzt, da keine lang stehenden Personen in der Szene zu sehen sind. Die Schrittweiten $p$ und $q$ werden in diesem Abschnitt auf den Wert 10 festgelegt.

\subsubsection{Personenzählung}
Zur Durchführung der Tests müssen zunächst wieder Referenzdaten durch manuelle Annotationen generiert werden, die erneut durch manuelles Zählen der Personen in jedem Eingangsbild erhalten werden. Die Personen innerhalb des Bildausschnitts werden in 187 Frames manuell gezählt und die Ergebnisse notiert. Um erneut die Analogie zum verwendeten Verfahren (MFFT) zu bewahren, werden die gezählten Personenzahlen mit einem gleitenden Mittelwertfilter mit einer Mittelungsdistanz von $m$ Frames gemittelt. Die Mittelungsdistanz $m$ wird auf 10 Frames festgelgt. Anschließend werden die vom Verfahren ausgegebenen Messwerte für den Verlauf der Testsequenz exportiert und gespeichert. Die Ergebnisse werden in Abb. \ref{business:zahl} dargestellt und betrachtet.
\newpage
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {images/dummy.png}
  }
  \caption{Graph: Verlauf der gemessenen/gezählten Personenzahl \cite{AliS07}}
  \label{business:zahl}
\end{figure}

Man erkennt hier die Auftragung der manuell Annotierten und der algorithmisch geschätzten Messwerte des entwickelten Verfahrens (MFFT). Die geschätzte Personenzahl folgt dem Trend der erstellten Ground Truth, ist jedoch rauschbehaftet. Vergleicht man die Graphen mit denen der Personenzählung aus Abschnitt \ref{test:erg} Abbildung \ref{zahl}, stellt man fest, dass die, zu Beginn auftretende, überhöhte Personenzählung hier ausbleibt. Die überhöhte Personenzählung tritt im Fall von Personenverdeckungen auf, die sich im Laufe einer Sequenz, in der die Personenzahl stetig zunimmt, ausbilden können. In der hier verwendeten Sequenz beträgt die Personenzählung inital nach dem Einschwingen über 60\%. Hier nimmt also die Dichte nicht so drastisch (60\%-100\%) wie in der in Abbildung \ref{zahl} verwendeten Sequenz zu, in der sie von 10\%-100\% zunimmt, weswegen eine Zunahme der Verdeckungen im Verlauf der Testsequenz hier sehr viel schwächer ausgeprägt ist. Die überhöhte Personenzählung zu Beginn der Sequenz bleibt also aus, weil bereits zu Beginn mehr als 60\% der maximalen Personenzahl erreicht wird. Die Personen verdecken sich also demnach schon zu Beginn der Sequenz und verdecken sich im Verlauf der Sequenz nicht merklich zunehmend weiter.

\subsubsection{Personenfluss}
Wie bereits in Abschnitt \ref{dispersion_test} behandelt, kann das entwickelte Verfahren (MFFT) zur Erfassung des Personenflusses nicht manuell annotiert werden. Dieses Verfahren stützt sich auf das Zählen von Trajektorienspitzen und -knoten und es wäre sehr aufwendig für Menschen solche Trajektorien in einer Testsequenz zu erstellen, verwalten und schlussendlich auszuwerten. Daher wird erneut, wie bereits in Abschnitt \ref{dispersion_fluss}, der Personenfluss über die letzten $m$ Frames nach Gleichung \ref{fluss_gt} bestimmt. Die Schrittweite $p$ wird auf 10 Frames und die Mittelungsdistanz $m$ auf 30 Frames festgelegt. Der mittlere, relative Personenfluss wird über den Verlauf der Sequenz aufgetragen und in Abb. \ref{business:fluss} dargestellt.
\vskip 10pt
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {images/dummy.png}
  }
  \caption{Graph: Verlauf des gemessenen/gezählten Personenflusses \cite{AliS07}}
  \label{business:fluss}
\end{figure}
\newpage
Man erkennt, dass der Verlauf des mittleren Personenflusses erneut, wie schon mit den computergenerierten Daten, mit dem Verhalten der Ground Truth korreliert. Der Verlauf scheint jedoch um etwa 10 Frames verzögert. Die Verzögerung resultiert aus den unterschiedlichen verwendeten Verfahren. Die Verfahren erfassen beide einen Personenfluss, sind sich jedoch in ihrem Ansatz Einzelschätzungen zu glätten unterschiedlich. Diese Verzögerung tritt auf, weil die Ground Truth, wie in Gleichung \ref{fluss_gt} dargestellt, generiert wird. Im Verfahren jedoch, wird der Personenfluss zunächst über eine Schrittweite von $p$ Frames berechnet und anschließend über $m$ Frames gemittelt. Da dieses Verfahren in manueller Form nicht ausgeführt werden kann, wird der Personenfluss als Ground Truth direkt über $m$ Frames erhalten, was die Verzögerung um $p$ (hier 10) Frames erklärt.

\subsubsection{Dichtefaktor}
Wie bereits in Abschnitt \ref{dispersion_dichte} erwähnt, wird der abgeleitete Dichtefaktor aus der Multiplikation der Messgröße relativer Trägheitsfaktor $B_{rel}$ mit der Messgröße relative Personenzählung erhalten. Da sich erneut, wie bereits in Abschnitt \ref{dispersion_dichte}, das manuelle Vorgehen zur Bestimmung einer mittleren Merkmalsgeschwindigkeit den Umfang der Arbeit überschreiten würde, werden für den Dichtefaktor keine Referenzdaten durch manuelle Annotationen generiert. Der Dichtefaktor wird mit der, bereits evaluierten, rel. Personenzahl verglichen, um seine Funktion und Unterscheidung von der reinen Messung einer Personenzahl zu zeigen. Die geschätzten Messungen des Verfahrens für den Dichtefaktor werden mit denen für die relative Personenzahl in Abbildung \ref{business:dichte} verglichen.

%Da sich erneut das Vorgehen des Menschen zur Berechnung einer mittleren Mermalsgeschwindigkeit kaum von der algorithmischen Vorgehensweise unterscheidet, können die vom Plugin generierten Messergebnisse für den relativen Trägheitsfaktor als Ground Truth angenommen werden. Die manuell erhaltenen Werte für die relative Personenzählung werden mit den Generierten für den relativen Trägheitsfaktor multipliziert, um eine Ground Truth für den relativen Dichtefaktor zu erhalten. Anschließend werden die intern im Plugin generierten Messwerte für den relativen Dichtefaktor exportiert und gespeichert. 
\newpage
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {images/dummy.png}
  }
  \caption{Graph: Verlauf des Dichtefaktors und der Personenzahl (MFFT) \cite{AliS07}}
  \label{business:dichte}
\end{figure}
Hier erkennt man, dass die Verläufe der beiden Messgrößen stark korrelieren. Die Abweichung zu Beginn der Sequenz, wie in Abschnitt \ref{dispersion_dichte} beschrieben, bleibt hier aus. Dies resultiert aus der hohen Dynamik, die im Verlauf dieser Testsequenz durchgängig vorliegt. Im Fall einer hohen Dynamik werden immer kleine Messwerte für den relativen Trägheitsfaktor generiert, wobei durch die Multiplikation gleiches für den Dichtefaktor gilt. Der Verlauf für den Dichtefaktor ist nach unten versetzt, weil die kleinen Werte des relativen Trägheitsfaktors multipliziert mit den hohen Werten für die relativen Personenzahl eine nach unten versetzte Kurve ergeben.

Außerdem erkennt man den, aus Abbildung \ref{dichte} bekannten, Überschwinger zu Beginn der Testsequenz. Dieser ergibt sich erneut durch den Einschwingvorgang der Messgröße relativer Trägheitsfaktor (vgl. Kapitel \ref{chap:imp} Abschnitt \ref{sec:sig}). Weil zunächst $q$ (hier 10) Frames vergehen müssen, bis die mittlere Vektorlänge der Bewegung über die letzten $q$ Frames ($B_j$) Messwerte generiert und diese anschließend noch über $m$ Frames gemittelt wird, dauert es weitere $m$ (hier 10) Frames, bis die mittlere Merkmalsgeschwindigkeit ($\bar{B_j}$) ihren hohen Betriebswert annimmt. Diese Größe nimmt also zu Beginn im Einschwingvorgang kleine Werte an, weswegen der rel. Trägheitsfaktor ($B_{rel}$), der als $(\frac{B_S}{\bar{B}_j})$ berechnet wird, zu Beginn immer hohe Werte annimmt. Weil aus dieser Größe der Dichtefaktor abgeleitet wird, nimmt dieser ebenfalls zu Beginn immer hohe Werte an und schwingt über. Nach dem Einschwingvorgang (etwa ab Frame 20) funktioniert diese Messgröße allerdings passabel und nimmt sinnvolle Werte an. Generell korreliert der Verlauf des abgeleiteten Dichtefaktors (MFFT) mit dem Verlauf der Ground Truth und kann als aussagekräftig betrachtet werden.
\newpage
\subsection{Videosequenz: UCF\_CrowdsDataset/Marathon \cite{AliS07}}
Die bereits beschriebene Videosequenz Marathon wird dazu verwendet, die Trennung von Bewegten und statischen Merkmalen zu evaluieren. Dafür wird die maximale Verweilzeit auf 0 Frame gestellt, um die Erfassung statischer Merkmale nicht zu erlauben. Das Verfahren wird wie gewohnt für die Szene konfiguriert. Anschließend werden über 30 Frames Trajektorien mit dem entwickelten Verfahren (MFFT) erfasst und gezeichnet. Die erfassten Trajektorien werden in Abbildung \ref{run_trajectories} dargestellt:
\vskip 10pt
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {images/dummy.png}
  }
  \caption{Darstellung von erfassten Trajektorien von Marathonläufern \cite{AliS07}}
  \label{run_trajectories}
\end{figure}

Man erkennt hier, dass nahezu ausschließlich in der bewegten Masse an Läufern Trajektorien erfasst werden, die als Gruppe oberhalb der eingestellten Geschwindigkeitsschwelle liegt. Statische Punkte wie Ecken an Häusern oder unbewegten Personen werden nicht erfasst. Die Trennung der Merkmale in Statische und nicht-Statische funktioniert also zureichend und kann als plausibel angenommen werden.

\subsection{Videosequenz: Hamburger Hafenfest 2014}
Die zuvor beschriebene Videosequenz des Hamburger Hafenfests wird verwendet, um die Erfassung präziser Trajektorien in Personengruppen hoher Dichte zu evaluieren. Dabei werden niedrig aufgelöste Aufnahmen mit 960x540 verwendet, um zu zeigen, dass die Erfassung solcher Trajektorien auch mit relativ niedrig aufgelösten Daten zuverlässig funktioniert. Die Trajektorien werden erfasst und als blaue Linien im Eingangsbild in Abbildung \ref{hafen} dargestellt:
\newpage
\begin{figure}[h]
  \centering
  \fbox{
    \includegraphics[width=0.7\textwidth]
    {images/dummy.png}
  }
  \caption{Erfasste Trajektorien in der Testsequenz Hamburger Hafenfest}
  \label{hafen}
\end{figure}

Es ist zu erkennen, dass mit dem erstellten Algorithmus selbst in Personengruppen hoher Dichte individuelle Trajektorien geführt werden können und somit die Abbleitung von aussagekräftigen Messgrößen in derartigen Szenarien auf Massenveranstaltungen möglich ist. Außerdem erkennt man hier, dass die Fahnenbewegungen im Wind kleine Trajektorien entstehen lassen. Diese Trajektorien können gefiltert werden, indem man eine minimale Vektorlänge über die letzten M Frames vorgibt (vgl. Kapitel \ref{chap:imp} Abschnitt \ref{sec:konfig}). Da die Fahnenbewegungen Kreisbewegungen sind und keine exzentrische Orientierung haben, kann sich die Fahne nie über große Teile des Bildes bewegen, wie dies etwa Menschen tun. Damit wird sie nie eine hohe Vektorlänge über die letzten M Frames erreichen, weil sie immer wieder in etwa zu ihrem Ausgangspunkt zurückkehrt. Dabei werden zwar in Ausnahmefällen auch gewünschte Trajektorien gefiltert, jedoch können so Falschdetektionen (sog. False Positives) weitgehend verhindert werden.

### 2018-12-21 thesis_sebastian_bek/content/5-Fazit.tex
Initial commit

\chapter{Fazit}
\label{chap:faz}
Das Ziel dieser Arbeit war es, ein algorithmisches Verfahren zur Analyse von Gruppierungs- und Bewegungsmustern zu entwickeln. Dieses sollte zunächst Trajektorien von Personen erfassen und analysieren. Aus den Trajektorien sollten weiterhin Messgrößen abgeleitet und dargestellt werden, mit denen eine möglichst genaue Analyse von Bewegungs- und Gruppierungsmustern möglich ist. Das entworfene Verfahren basiert auf dem Harris Corner Detector und dem sog. optischen Fluss (vgl. Kapitel \ref{chap:grund}). Es wurde im Rahmen eines Videoanalysesystems entwickelt, dessen Struktur in Kapitel \ref{chap:imp} näher beschrieben wurde. Im Kapitel \ref{chap:erg} wurde das entwickelte Verfahren mithilfe von Grundwahrheiten quantitativ bewertet. Dabei wurde festgestellt, dass die generierten relativen Messwerte mit dem Verhalten der manuell generierten Wahrheit (Ground Truth) korrelieren. Die generierten Messwerte können also als aussagekräftig betrachtet werden.

Probleme die zeitweilige Abweichungen verursachen sind \zb Verdeckungen, die auftreten wenn im Verlauf der Sequenz die Personendichte stark variiert. Dieses Problem ist auf dem Gebiet der Bildverarbeitung bekannt und es existieren teilweise sehr aufwändige aber wirksame Lösungsansätze. Außerdem tritt oft zu Beginn der Sequenz ein Einschwingen bestimmter Messgrößen auf. Dies ist aber für die vorliegende Aufgabenstellung unkritisch, da das System für ein Langzeitmonitoring ausgelegt ist. Das vorliegende System wurde primär für die Erfassung bewegter Personen entwickelt und es konnte gezeigt werden, dass die Erfassung von Trajektorien in Personengruppen hoher Dichte, auch unter Verwendung von niedrig aufgelösten Bilddaten, zuverlässig funktioniert.

Als Ausblick könnte das Verfahren durch die Nutzung von intrinsischen und extrinsischen Kameraparametern, welche in der Regel durch eine Kalibrierung während der Einrichtphase gewonnen werden, hinsichtlich Systemparametrisierung und Robustifizierung verbessert werden. So könnten Evaluations-/Filtereinstellungen und Grenzwerte über den Bildort $x,y$ skaliert werden, um Einflüsse der unterschiedlich hohen Skalierung von Personen einzuschränken. Außerdem könnte versucht werden, dass Tracking der Personen durch die Wahl eines anderen Corner Detectors zu verbessern. 
\newpage
Möglicherweise existieren andere Eckendetektoren (\zb SIFT, SURF, ORB uvm.), die in diesem Kontext eine höhere Erkennungsleistung liefern. Um stehenbleibende Personen mitzuzählen, ohne zugleich statische Objekte mit zu erfassen, könnte die maximale Verweilzeit nur auf Trajektorien angewandt werden, die bereits eine große Strecke zurückgelegt haben. So hätten spontan an Objektkanten entstandene Trajektorien keine Berechtigung weiter zu existieren. Zusätzlich könnte das Verfahren mit einem vereinfachten Konfigurationsablauf sowie einer Konfigurationsdokumentation implementiert werden.

Abschließend kann festgestellt werden, dass das entwickelte Verfahren vielversprechende Ergebnisse liefert und in einer weiteren Entwicklungsstufe unter realen Bedingungen evaluiert werden sollte. Insbesondere sollten Crowd Manager(Sicherheitspersonal bei Großveranstaltungen) die Praxistauglichkeit des Verfahrens bewerten und dadurch die Einsetzbarkeit in der Praxis bestätigen.

%So könnte sie verbreitet genutzt werden, ohne dass ein zu großer Einrichtungsaufwand besteht. Solche Überwachungssysteme müssten, am Veranstaltungsort einmal installiert, vor jeder Veranstaltung einmal kalibriert werden und könnten im Verlauf der Veranstaltung für eine erhöhte Sicherheit sorgen. 
%Zusätzlich könnten die Veranstaltungsplaner im Nachhinein besonders kritische Gebiete des Veranstaltungsgeländes markieren, die entlastet werden sollten. Das sogenannte Tracking von Personen/Objekten kann auch in anderen Forschungsgebieten eingesetzt werden. Beispielsweise kann das Verhalten von Konsumenten in Supermärkten analysiert werden, um die Anordnung der Produkte zu optimieren oder auf kleinerer Skala könnten Proben im Labor unter dem Mikroskop hinsichtlich ihres Bewegungs- und Gruppierungsverhaltens analysiert werden. Diese und viele weitere Möglichkeiten bietet die algorithmische Erfassung von Trajektorien. Zum Einsatz einer solchen Software muss aber noch viel Forschung betrieben und gegebenenfalls müssen umfangreiche Praxistests durchgeführt werden. Die Arbeit hat jedoch gezeigt, dass die Erfassung präziser Trajektorien, nach ausgewähltem Verfahren, in Gebieten hoher Personendichte, auch mit niedrig aufgelösten Daten, grundsätzlich möglich ist. Außerdem wurde bestätigt, dass relevante, aussagekräftige Messgrößen aus den erstellten Trajektorien abgeleitet und dargestellt werden können. Demnach ist das Verfahren zur verbesserten Analyse von Gruppierungs- und Bewegungsmustern geeignet, weswegen eine weitere Behandlung des Themas sinnvoll ist.
%

%mnr: Du bist hier etwas ausgeschweift, und hast über mögliche Anwendungsgebiete gesprochen. Das sollte nicht sein, denn dann hättest du zunächst zeigen müssen, dass es für diese Anwendungen nicht bereits bessere Verfahren gibt. Wir bleiben also mit dieser Arbeit beim Thema Großveranstaltung!
%mnr: Hier wäre es besser wenn du noch Überlegungen anstellen könntest, wie man vielleicht einzelne Komponenten deines Verfahrens verbessern könnte. z.B. das Tracking, die Problematik "Person läuft irgendwo hin, und dann bleibt sie lange stehen"...

### 2018-12-21 thesis_sebastian_bek/content/nomenclatures.tex
Initial commit

\nomenclature{HG}{Hintergrund}
\nomenclature{VG}{Vordergrund}
\nomenclature{$\abs{\vec{a}}$}{Betrag eines Vektors}
\nomenclature{$\abs{\vec{b}}$}{Betrag eines Vektors der Bewegung über $q$ Frames}
\nomenclature{$v(x,y)$}{Merkmalsgeschwindigkeit}
\nomenclature{$v_{SI}(x,y)$}{Merkmalsgeschwindigkeit in SI-Einheiten}
\nomenclature{$d(x,y)$}{Abstand zwischen 2 Pixeln in $\frac{m}{px}$}
\nomenclature{$D$}{fraktale Dimension}
\nomenclature{$FR$}{Framerate - Bildfrequenz in $[\text{fps}] = [\frac{\text{frame}}{s}]$}
\nomenclature{c}{Center Pixel/Hotspot, auf den sich eine Nachbarschaft/Umgebung oder ein Patch/Bildausschnitt bezieht}
\nomenclature{ROI}{region of interest - Vordergrundregion(\zb Positionen oder Polygonzüge)}
\nomenclature{HSV}{hue, saturation, value - Farbmodell, für intuitivere Definition von Farben}
\nomenclature{$w(u,v)$}{Fenster-/Wichtungsfunktion}
\nomenclature{$W(u,v)$}{Wertebereich für die Laufvariablen $u$ und $v$ innerhalb eines Bildausschnitts konstanter Größe in $[\text{px}]$}
\nomenclature{$W(w,h)$}{Wertebereich für die Bildkoordinaten $x$ und $y$ innerhalb eines Bildausschnitts festgelegter Größe in $[\text{px}]$}
\nomenclature{$w$}{Breite(width) eines Bildausschnitts}
\nomenclature{$h$}{Höhe(height) eines Bildausschnitts}
\nomenclature{$I(x,y)$}{Intensitäts-/Bildmatrix}
\nomenclature{$I_x$}{diskrete Ableitung der Intensitätsmatrix in x-Richtung in $[\frac{1}{\text{px}}]$}
\nomenclature{$I_y$}{diskrete Ableitung der Intensitätsmatrix in y-Richtung in $[\frac{1}{\text{px}}]$}
\nomenclature{$M$}{Kovarianzmatrix der Verteilung von Gradientenbeiträgen $I_x$,$I_y$}
\nomenclature{$\lambda_{1/2}$}{Eigenwerte der Matrix A}
\nomenclature{$DX$}{DX-Wert eines Verschiebungsvektors in $[\text{px}]$}
\nomenclature{$DY$}{DY-Wert einer Verschiebungsvektors in $[\text{px}]$}
\nomenclature{$c_K$}{Konstante}
\nomenclature{$j$}{Index für den aktuellen Frame}
\nomenclature{$m$}{Mittelungsdistanz - Anzahl der rückwärtigen Messwerte $m$, die in die Mittelung einbezogen werden in $[\text{frame}]$}
\nomenclature{$x,y$}{Bildkoordinaten - Breite, Höhe}
\nomenclature{$P(x,y)$}{Zeilen-/Spaltenposition im Eingangsbild}
\nomenclature{$P_j(x)$}{x-Position im Eingangsbild im aktuellen Frame j}
\nomenclature{$\text{var}()$}{Varianz}
\nomenclature{$\text{covar}()$}{Kovarianz}
\nomenclature{$u,v$}{Laufvariablen für die Verschiebung in Zeilen- und Spaltenrichtung}
\nomenclature{$A$}{projezierte Fläche eines Bildausschnitts in $[m^2]$}
\nomenclature{$S$}{Menge an Trajektorienspitzen(aktuelle Position) im Patch}
\nomenclature{$K$}{Menge an Trajektorienknoten(vorherige Position) im Patch}
\nomenclature{$Z$}{Menge an berechneten Vektorlängen $\abs{\vec{b}}_j(x,y)$ im Patch}
\nomenclature{$N$}{Personenzahl in $[\text{personen}]$}
\nomenclature{$\bar{N}_j$}{mittlere Personenzahl im aktuellen Frame j in $[\text{personen}]$}
\nomenclature{$N_{max}$}{trainierte maximale Personenzahl in $[\text{personen}]$}
\nomenclature{$N_{abs}$}{absolute maximale Personenzahl in $[\text{personen}]$}
\nomenclature{$N_{rel}$}{relative mittlere Personenzahl in $[\%]$}
\nomenclature{$D_{scal}$}{skalierte Personendichte in $[\frac{\text{personen}}{m^2}]$}
\nomenclature{$N_{scal}$}{skalierte Personenzahl in $[\text{personen}]$}
\nomenclature{$F_{acc}$}{Fluss-Akkumulator in $[\frac{\text{personen}}{\text{frame}}]$}
\nomenclature{$F$}{Personenfluss in $[\frac{\text{personen}}{\text{frame}}]$}
\nomenclature{$\bar{F}_j$}{mittlerer Personenfluss in $[\frac{\text{personen}}{\text{frame}}]$}
\nomenclature{$F_{rel}$}{relativer mittlerer Personenfluss $[\%]$}
\nomenclature{$F_{max}$}{trainierter maximaler mittlerer Personenfluss $[\%]$}
\nomenclature{$F_{scal}$}{skalierter Personenfluss in $[\frac{\text{personen}}{\text{frame}}]$}
\nomenclature{$B_j$}{mittlere Merkmalsgeschwindigkeit innerhalb eines Bildausschnitts in $[\frac{px}{q\cdot \text{frame}}]$}
\nomenclature{$B_S$}{niedrigster Schwellwert für die mittlere Merkmalsgeschwindigkeit in $[\frac{px}{q\cdot \text{frame}}]$}
\nomenclature{$\bar{B}_j$}{zeitlich gemittelte mittlere Merkmalsgeschwindigkeit innerhalb eines Bildausschnitts}
\nomenclature{$\bar{B}_{rel}$}{relative mittlere Merkmalsgeschwindigkeit innerhalb eines Bildausschnitts in $[\%]$}
\nomenclature{$\text{dFactor}$}{Dichtefaktor in $[\%]$}
\nomenclature{$X_{rel}$}{relative Messgröße allgemein in $[\%]$}
\nomenclature{$\text{hue}$}{hue: zahlenmäßiger Farbton im HSV-Farbmodell in  $[\ ]$}

\nomenclature{$r$}{Suchradius in $[\text{px}]$}
\nomenclature{$p$}{Schrittweite in $[\text{frame}]$}
\nomenclature{$q$}{Schrittweite in $[\text{frame}]$}

\nomenclature{$d_1$}{Abstand der unteren Bildzeile zum Kameraobjektiv in $[m]$}
\nomenclature{$d_2$}{Abstand der oberen Bildzeile zum Kameraobjektiv in $[m]$}

### 2018-12-21 C++ Projects/Eigenframe.cpp

@@ -1,0 +1,129 @@
+#include <iostream>
+#include <vector>
+#include <stdexcept>
+#include <Eigen/Eigen>
+using namespace Eigen;
+
+//using namespace std;
+
+
+class my_matrix {
+	std::vector<std::vector<double>> m;
+	int width;
+	int height;
+public:
+	my_matrix(unsigned int x, unsigned int y)
+	{
+		m.resize(x, std::vector<double>(y, 0.0));
+		width = x;
+		height = y;
+	}
+	int getWidth() {return width;}
+	int getHeight() {return height;}
+	class matrix_row {
+		std::vector<double>& row;
+	public:
+		matrix_row(std::vector<double>& r) : row(r) // row gets assigned with adress r
+		{
+		}
+		double& operator[](unsigned int y) // inner operator
+		{
+			return row.at(y); // This is where the doubles come from !!!
+		}
+	};
+	matrix_row operator[](unsigned int x) // outer operator
+	{
+		return matrix_row(m.at(x));
+	}
+};
+
+
+struct complex { double r, i; bool real; };
+template <typename T>
+struct pair { T p1, p2; };
+
+pair<complex> midnightFormula(double a, double b, double c)
+{
+	pair<complex> result = { 0 };
+
+	if (a < 0.000001)    // ==0
+	{
+		if (b > 0.000001)  // !=0
+			result.p1.r = result.p2.r = -c / b;
+		else
+			if (c > 0.00001) throw std::exception("no solutions");
+		return result;
+	}
+
+	double delta = b * b - 4 * a*c;
+	if (delta >= 0)
+	{
+		result.p1.r = (-b - sqrt(delta)) / 2 / a;
+		result.p2.r = (-b + sqrt(delta)) / 2 / a;
+		result.p1.real = result.p2.real = true;
+	}
+	else
+	{
+		result.p1.r = result.p2.r = -b / 2 / a;
+		result.p1.i = sqrt(-delta) / 2 / a;
+		result.p2.i = -sqrt(-delta) / 2 / a;
+	}
+
+	return result;
+}
+
+pair<complex> Eigenvalues(my_matrix matrix)
+{
+	double lambda1 = 0.0;
+	double lambda2 = 0.0;
+
+	double a = 1;
+	double b = -1 * (matrix[0][0] + matrix[1][1]);
+	double c = matrix[0][0] * matrix[1][1] - matrix[0][1] * matrix[1][0];
+
[…]

### 2018-12-21 C++ Projects/sfml2DGrid7x7.cpp

@@ -1,0 +1,278 @@
+#define _USE_MATH_DEFINES
+#include <SFML/Graphics.hpp>
+#include <iostream>
+#include "muParser.h"
+#include <cmath>
+#include <sstream>
+#include <fstream>
+#include <vector>
+#include <unistd.h>
+#include <math.h>
+#include "exprtk.hpp"
+#include <Eigen/Dense>
+#include <Eigen/Eigenvalues>
+
+
+#ifdef WINDOWS
+#include <direct.h>
+#define GetCurrentDir _getcwd
+#else
+#include <unistd.h>
+#define GetCurrentDir getcwd
+#endif
+
+#define MAXBUFSIZE  ((int) 1e3)
+
+using namespace std;
+using namespace Eigen;
+
+//length and width of window
+int wSize = 700;
+sf::Vector2i windowsize;
+
+//recording frameskip 
+int record_frameskip = -1;
+
+//directory to write images to, must exist
+std::string record_folder = "frames";
+
+//fullscreen flag
+bool fullscreen = false;
+
+//definition of pi
+const double pi = M_PI;
+
+// define function string
+std::string functionStr = "cos(theta) + sin(3*theta)";
+
+// get current working directory to assign matrix.txt path
+std::string GetCurrentWorkingDir(void) {
+	char buff[FILENAME_MAX];
+	GetCurrentDir(buff, FILENAME_MAX);
+	std::string current_working_dir(buff);
+	return current_working_dir;
+}
+
+void defineConvexEllipse(sf::ConvexShape* ellipse, double radius_x, double radius_y, unsigned short quality)
+{
+	for (unsigned short i = 0; i < quality; ++i)
+	{
+		float rad = (360 / quality * i) / (360 / M_PI / 2);
+		float x = cos(rad)*radius_x;
+		float y = sin(rad)*radius_y;
+
+		ellipse[0].setPoint(i, sf::Vector2f(x, y));
+	}
+}
+
+MatrixXd readMatrix(std::string filepath, int* colsCount, int* rowsCount)
+{
+	int cols = 0, rows = 0;
+	double buff[MAXBUFSIZE];
+	ifstream infile(filepath);
+
+	while (!infile.eof())
+	{
+		string line;
+		getline(infile, line);
+
+		int temp_cols = 0;
+		stringstream stream(line);
+		while (!stream.eof())
+			stream >> buff[cols*rows + temp_cols++];
+
+		if (temp_cols == 0)
+			continue;
+
[…]

### 2018-12-21 C++ Projects/sfml2DTensorVisEllipses.cpp

@@ -1,0 +1,1052 @@
+#define _USE_MATH_DEFINES
+#include <SFML/Graphics.hpp>
+#include <iostream>
+#include "muParser.h"
+#include <cmath>
+#include <sstream>
+#include <fstream>
+#include <vector>
+#include <unistd.h>
+#include <math.h>
+#include "exprtk.hpp"
+#include <Eigen/Dense>
+#include <Eigen/Eigenvalues>
+#include <algorithm>
+#include <functional>
+
+#define MAXBUFSIZE  ((int) 1e3)
+
+using namespace std;
+using namespace Eigen;
+
+//length and width of window
+int wSize = 700;
+sf::Vector2i windowsize;
+
+//recording frameskip 
+int record_frameskip = -1;
+
+//directory to write images to, must exist
+std::string record_folder = "frames";
+
+//fullscreen flag
+bool fullscreen = false;
+
+//definition of pi
+const double pi = M_PI;
+
+std::string functionString; // Prototyping
+
+template <typename T>
+T muPlus(T& a, T& b)
+{
+	T res = a + b;
+	return res;
+}
+
+// OPERATOR DEFINITIONS //
+template <typename T>
+std::vector<T> operator+(const std::vector<T>& a, const std::vector<T>& b)
+{
+	assert(a.size() == b.size());
+	
+	std::vector<T> result(a.size());
+
+	for(int i = 0; i < a.size(); i++)
+		result[i] = a[i] + b[i];
+	
+	return result;
+}
+
+
+// CLASS DEFINITIONS //
+class neighborhood
+{
+	bool neighborR = true;
+	bool neighborT = true;
+	bool neighborL = true;
+	bool neighborB = true;
+
+public:
+
+	//constructor
+	neighborhood(int j, int i, const int length, std::vector<bool>& processMap) // check the neighbourhood of (j,i) for missing neighbors..
+	{
+		// check if frame exceeded, outside FOV, offscreen..
+		if (i == 0)
+			neighborL = false; // left
+		else if (i == length - 1)
+			neighborR = false; // right
+		if (j == 0)
+			neighborT = false; // top
+		else if (j == length - 1)
+			neighborB = false; // bottom
+		// leave order.. functional!
+
+		// check neighbors in process(ed) map, for neighbors already processed ..
+		if (neighborR) // if neighbor existent..
[…]

### 2018-12-21 C++ Projects/sfml2DTensorVisV1Strings.cpp

@@ -1,0 +1,1209 @@
+#define _USE_MATH_DEFINES
+#include <SFML/Graphics.hpp>
+#include <iostream>
+#include "muParser.h"
+#include <cmath>
+#include <sstream>
+#include <fstream>
+#include <vector>
+#include <unistd.h>
+#include <math.h>
+#include "exprtk.hpp"
+#include <Eigen/Dense>
+#include <Eigen/Eigenvalues>
+#include <algorithm>
+#include <functional>
+#include <array>
+#include <filesystem>
+#include <sstream>
+#include <string>
+
+#ifdef WINDOWS
+#include <direct.h>
+#define GetCurrentDir _getcwd
+#else
+#include <unistd.h>
+#define GetCurrentDir getcwd
+#endif
+
+#define MAXBUFSIZE  ((int) 1e3)
+
+using namespace std;
+using namespace Eigen;
+using namespace mu;
+
+// PROTOTYPES
+std::string functionString; // Prototyping of functionString for strFunction!
+int width;
+int height;
+
+//length and width of window
+int wSize = 700;
+sf::Vector2i windowsize;
+
+//recording frameskip 
+int record_frameskip = 0; // --> disables recording
+
+//directory to write images to, must exist
+std::string record_folder = "frames";
+
+//fullscreen flag
+bool fullscreen = false;
+
+//definition of pi
+const double pi = M_PI;
+
+// get current working directory to assign matrix.txt path
+std::string GetCurrentWorkingDir(void) {
+	char buff[FILENAME_MAX];
+	GetCurrentDir(buff, FILENAME_MAX);
+	std::string current_working_dir(buff);
+	return current_working_dir;
+}
+
+void defineConvexEllipse(sf::ConvexShape* ellipse, double radius_x, double radius_y, unsigned short quality)
+{
+	for (unsigned short i = 0; i < quality; ++i)
+	{
+		float rad = (360 / quality * i) / (360 / M_PI / 2);
+		float x = cos(rad)*radius_x;
+		float y = sin(rad)*radius_y;
+
+		ellipse[0].setPoint(i, sf::Vector2f(x, y));
+	}
+}
+
+
+// OPERATORS
+template <typename T>
+T clip(const T& n, const T& lower, const T& upper) // template clip function
+{
+	return std::max(lower, std::min(n, upper));
+}
+
[…]

### 2018-12-21 C++ Projects/sfmlTensorVisV1Array.cpp

@@ -1,0 +1,1209 @@
+#define _USE_MATH_DEFINES
+#include <SFML/Graphics.hpp>
+#include <iostream>
+#include "muParser.h"
+#include <cmath>
+#include <sstream>
+#include <fstream>
+#include <vector>
+#include <unistd.h>
+#include <math.h>
+#include "exprtk.hpp"
+#include <Eigen/Dense>
+#include <Eigen/Eigenvalues>
+#include <algorithm>
+#include <functional>
+#include <array>
+#include <filesystem>
+#include <sstream>
+#include <string>
+
+#ifdef WINDOWS
+#include <direct.h>
+#define GetCurrentDir _getcwd
+#else
+#include <unistd.h>
+#define GetCurrentDir getcwd
+#endif
+
+#define MAXBUFSIZE  ((int) 1e3)
+
+using namespace std;
+using namespace Eigen;
+using namespace mu;
+
+// PROTOTYPES
+std::string functionString; // Prototyping of functionString for strFunction!
+int width;
+int height;
+
+//length and width of window
+int wSize = 700;
+sf::Vector2i windowsize;
+
+//recording frameskip 
+int record_frameskip = 0; // --> disables recording
+
+//directory to write images to, must exist
+std::string record_folder = "frames";
+
+//fullscreen flag
+bool fullscreen = false;
+
+//definition of pi
+const double pi = M_PI;
+
+// get current working directory to assign matrix.txt path
+std::string GetCurrentWorkingDir(void) {
+	char buff[FILENAME_MAX];
+	GetCurrentDir(buff, FILENAME_MAX);
+	std::string current_working_dir(buff);
+	return current_working_dir;
+}
+
+void defineConvexEllipse(sf::ConvexShape* ellipse, double radius_x, double radius_y, unsigned short quality)
+{
+	for (unsigned short i = 0; i < quality; ++i)
+	{
+		float rad = (360 / quality * i) / (360 / M_PI / 2);
+		float x = cos(rad)*radius_x;
+		float y = sin(rad)*radius_y;
+
+		ellipse[0].setPoint(i, sf::Vector2f(x, y));
+	}
+}
+
+
+// OPERATORS
+template <typename T>
+T clip(const T& n, const T& lower, const T& upper) // template clip function
+{
+	return std::max(lower, std::min(n, upper));
+}
+
[…]

### 2018-12-21 C++ Projects/sfmlTensorVisV1Array_dep.cpp

@@ -1,0 +1,1008 @@
+#define _USE_MATH_DEFINES
+#include <SFML/Graphics.hpp>
+#include <iostream>
+#include "muParser.h"
+#include <cmath>
+#include <sstream>
+#include <fstream>
+#include <vector>
+#include <unistd.h>
+#include <math.h>
+#include "exprtk.hpp"
+#include <Eigen/Dense>
+#include <Eigen/Eigenvalues>
+#include <algorithm>
+#include <functional>
+#include <array>
+
+#define MAXBUFSIZE  ((int) 1e3)
+
+using namespace std;
+using namespace Eigen;
+using namespace mu;
+
+//length and width of window
+int wSize = 700;
+sf::Vector2i windowsize;
+
+//recording frameskip 
+int record_frameskip = -1;
+
+//directory to write images to, must exist
+std::string record_folder = "frames";
+
+//fullscreen flag
+bool fullscreen = false;
+
+//definition of pi
+const double pi = M_PI;
+
+std::string functionString; // Prototyping
+
+// OPERATORS
+template <typename T>
+T clip(const T& n, const T& lower, const T& upper) // template clip function
+{
+	return std::max(lower, std::min(n, upper));
+}
+
+// -> muparser custom clip function
+static value_type clipNeg(value_type v) { return clip(v, 0.0, 1.0); }
+
+// OPERATOR DEFINITIONS //
+template <typename T> // element-wise plus for std::vector
+std::array<T, 21> operator+(const std::array<T, 21>& a, const std::array<T, 21>& b)
+{
+	assert(a.size() == b.size());
+
+	std::array<T, 21> result;
+
+	for (int i = 0; i < a.size(); i++)
+		result.at(i) = a.at(i) + b.at(i);
+
+	return result;
+}
+
+// OPERATOR DEFINITIONS //
+template <typename T> // element-wise plus for std::vector
+std::vector<std::array<T, 21>> operator+(const std::vector<std::array<T, 21>>& a, const std::vector<std::array<T, 21>>& b)
+{
+	assert(a.size() == b.size());
+
+	std::vector<std::array<T, 21>> result(a.size());
+
+	for (int i = 0; i < a.size(); i++)
+		result.at(i) = a.at(i) + b.at(i);
+
+	return result;
+}
+
+// CLASS DEFINITIONS //
+class neighborhood
+{
+	bool neighborR = true;
+	bool neighborT = true;
[…]

### 2018-12-21 PV-TF-generators/double_gyre_org.py

@@ -1,0 +1,115 @@
+import numpy as np
+from vtk.util.vtkAlgorithm import * # use (standalone) vtk module for sake of completeness
+from vtk.vtkCommonDataModel import vtkImageData
+from vtk.numpy_interface import dataset_adapter as dsa
+from timeit import default_timer as timer
+from paraview.pythonalgorithm import smproxy, smhint, smdomain, smproperty # import custom pythonalgorithm to generate XMLs via decorators in deprecated versions (<5.6.0)
+
+
+@smproxy.source(label="Double Gyre")
+@smhint.xml('''<ShowInMenu category="My Sources" />''')
+class DoubleGyre(VTKPythonAlgorithmBase):
+    def __init__(self):
+        self.Dimensions = [101, 101, 101]
+        self.T = 10.0
+        self.A = 0.1
+        self.eps = 0.1
+        self.Vectorize = True
+        VTKPythonAlgorithmBase.__init__(self, nInputPorts=0, nOutputPorts=1, outputType='vtkImageData')
+
+    @smproperty.intvector(name='Dimensions', default_values=[101, 101, 101])
+    def SetDimensions(self, dx, dy, dz):
+        self.Dimensions = [dx, dy, dz]
+        self.Modified()
+
+    @smproperty.doublevector(name='T', default_values=10.0)
+    def SetT(self, T):
+        self.T = T
+        self.Modified()
+
+    @smproperty.doublevector(name='A', default_values=0.1)
+    def SetA(self, A):
+        self.A = A
+        self.Modified()
+
+    @smproperty.doublevector(name='eps', default_values=0.1)
+    def Seteps(self, eps):
+        self.eps = eps
+        self.Modified()
+
+    @smproperty.intvector(name='Vectorize', default_values=1)
+    @smdomain.xml('''<BooleanDomain name="bool" />''')
+    def SetVectorize(self, vectorize):
+        self.Vectorize = vectorize
+        self.Modified()
+
+    def RequestInformation(self, request, inInfo, outInfo):
+        executive = self.GetExecutive()
+        info = outInfo.GetInformationObject(0)
+        info.Set(executive.WHOLE_EXTENT(),
+                 [0, self.Dimensions[0] - 1, 0, self.Dimensions[1] - 1, 0, self.Dimensions[2] - 1], 6)
+        return 1
+
+    def double_gyre_vectorized(self, x, y, t):
[…]

### 2018-12-21 PV-TF-generators/double_gyre_psrc2.py

@@ -1,0 +1,59 @@
+import vtk
+import numpy as np
+from vtk.util import numpy_support
+from vtk.numpy_interface import dataset_adapter as dsa
+import numpy as np
+import vtk
+import vtk.util.numpy_support as ns
+
+# set user parameters
+Dimensions = [101,101,101]
+T = 10.0
+A = 0.1
+eps = 0.1
+
+def double_gyre_vectorized( x, y, t):
+	a = eps * np.sin(2.0 * np.pi / T * t)
+	b = 1.0 - 2.0 * eps * np.sin(2.0 * np.pi / T * t)
+	f = a * x**2 + b * x
+	df = 2.0 * a * x + b
+
+	dx = -np.pi * A * np.sin(np.pi * f) * np.cos(np.pi * y)
+	dy = np.pi * A * np.cos(np.pi * f) * np.sin(np.pi * y) * df
+	dt = np.full_like(t, 1.0)
+
+	return np.dstack([dx, dy, dt])
+
+# create meshgrid and spacing
+xs = np.linspace(0.0, 2.0, Dimensions[0])
+ys = np.linspace(0.0, 1.0, Dimensions[1])
+ts = np.linspace(0.0, 10.0, Dimensions[2])
+x, y, t = np.meshgrid(xs, ys, ts, indexing='ij')
+
+# set data sink and image attributes
+img = self.GetImageDataOutput()
+img.SetDimensions(Dimensions)
+dims = Dimensions
+img.SetSpacing(abs(xs[1] - xs[0]), abs(ys[1] - ys[0]), abs(ts[1] - ts[0]))
+img.SetOrigin(xs[0], ys[0], ts[0])
+img = dsa.WrapDataObject(output)
+
+v = double_gyre_vectorized(x, y, t)
+
+img.PointData.append(v.reshape((-1, 3), order='F'), 'v')
+img.VTKObject.GetPointData().SetActiveScalars('v')
+
+# Request Information Script 1:
+# from paraview import util
+# Dimensions = [101,101,101]
+# op = self.GetOutput()
+# util.SetOutputWholeExtent(self, [0, Dimensions[0] - 1, 0, Dimensions[1] - 1, 0, Dimensions[2] - 1])
+
+# //////
+
+# Request Information Script 2
+# Dimensions = [101,101,101]
+# executive = self.GetExecutive()
+# outInfo = executive.GetOutputInformation(0)
[…]

### 2018-12-21 PV-TF-generators/import_script.py

@@ -1,0 +1,7 @@
+import vtk
+import numpy as np
+from vtk.util.vtkAlgorithm import * # use (standalone) vtk module for sake of completeness
+from vtk.vtkCommonDataModel import vtkImageData
+from vtk.numpy_interface import dataset_adapter as dsa
+from timeit import default_timer as timer
+from vtk.util import numpy_support

### 2018-12-21 PV-TF-generators/tensor_field_psrc.py

@@ -1,0 +1,65 @@
+import vtk
+import numpy as np
+from vtk.util import numpy_support
+from vtk.numpy_interface import dataset_adapter as dsa
+import numpy as np
+import vtk
+import vtk.util.numpy_support as ns
+
+# set user parameters
+Dimensions = [11,11,11]
+T = 10.0
+A = 0.1
+eps = 0.1
+
+def double_gyre_vectorized( x, y, t):
+	a = eps * np.sin(2.0 * np.pi / T * t)
+	b = 1.0 - 2.0 * eps * np.sin(2.0 * np.pi / T * t)
+	f = a * x**2 + b * x
+	df = 2.0 * a * x + b
+
+	xx = -np.pi * A * np.sin(np.pi * f) * np.cos(np.pi * y)
+	yx = np.pi * A * np.cos(np.pi * f) * np.sin(np.pi * y) * df
+	zx = np.pi * A * np.cos(np.pi * f)
+	xy = -np.pi * A * np.sin(np.pi * f) * np.cos(np.pi * y)
+	yy = np.pi * A * np.cos(np.pi * f) * np.sin(np.pi * y) * df
+	zy = np.pi * A * np.sin(np.pi * f)
+	xz = -np.pi * A * np.sin(np.pi * f) * np.cos(np.pi * y)
+	yz = np.pi * A * np.cos(np.pi * f) * np.sin(np.pi * y) * df
+	zz = np.pi * A * np.sin(np.pi * f) * np.cos(np.pi * f)
+
+	return np.dstack([xx, yx, zx, xy, yy, zy, xz, yz, zz])
+
+# create meshgrid and spacing
+xs = np.linspace(0.0, 2.0, Dimensions[0])
+ys = np.linspace(0.0, 1.0, Dimensions[1])
+ts = np.linspace(0.0, 10.0, Dimensions[2])
+x, y, t = np.meshgrid(xs, ys, ts, indexing='ij')
+
+# set data sink and image attributes
+img = self.GetImageDataOutput()
+img.SetDimensions(Dimensions)
+dims = Dimensions
+img.SetSpacing(abs(xs[1] - xs[0]), abs(ys[1] - ys[0]), abs(ts[1] - ts[0]))
+img.SetOrigin(xs[0], ys[0], ts[0])
+img = dsa.WrapDataObject(output)
+
+v = double_gyre_vectorized(x, y, t)
+
+img.PointData.append(v.reshape((-1, 9), order='F'), 'v')
[…]

### 2018-12-21 PV-TF-generators/test_env.py

@@ -1,0 +1,14 @@
+import numpy as np
+xs = np.linspace(0.0, 2.0, 10)
+ys = np.linspace(0.0, 1.0, 10)
+ts = np.linspace(0.0, 10.0, 10)
+x, y, t = np.meshgrid(xs, ys, ts, indexing='ij')
+for k in range(10):
+    for j in range(10):
+        for i in range(10):
+            px = x[i, j, k]
+            py = y[i, j, k]
+            pt = t[i, j, k]
+            print(px + i)
+            print(py + j)
+            print(pt + k)

### 2018-12-21 TF-5x5/X.py

@@ -1,0 +1,98 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1#/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 5
+
+deg = 0
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. zeros
[…]

### 2018-12-21 TF-5x5/Y.py

@@ -1,0 +1,98 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 5
+
+deg = 90
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. zeros
[…]

### 2018-12-21 TF-5x5/cone.py

@@ -1,0 +1,93 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 5
+
+deg = -45
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
+            if i==j: # if diagonal hit..
+                scaled = np.matmul(scale(4-i*(3/(length-1)),1),identity) # chronological transformation order: right->left
+                rotated = rotate(scaled, rad)
[…]

### 2018-12-21 TF-5x5/cross-X.py

@@ -1,0 +1,111 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 5
+
+deg1 = -45
+rad1 = deg1*(np.pi/180.0)
+deg2 = 45
+rad2 = deg2*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
[…]

### 2018-12-21 TF-5x5/drain.py

@@ -1,0 +1,94 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 5
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            y = length - 1 + round(j-(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
+            x = round(i-(length-1)/2) # --> center around midpoint
+            matrixArray[j][i] = normalize(np.array([[1, 0], [0, 1]])) # initialization w. normed identity
[…]

### 2018-12-21 TF-5x5/isotropic.py

@@ -1,0 +1,98 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 5
+
+deg = 0
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. zeros
[…]

### 2018-12-21 TF-5x5/rings.py

@@ -1,0 +1,139 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 5
+
+deg = -45
+rad = deg*(np.pi/180.0)
+
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations (polar form)
+radrange = 2*m.pi
+radres = m.pi/180 # 1deg/step
+steps = radrange/radres
+radarr = [float(i)*radrange/(steps-1) for i in range(round(steps))]
+curDeltaX = [[1.0]*length]*length
+curDeltaY = [[1.0]*length]*length
+curDelta = np.ones((length,length))
+
[…]

### 2018-12-21 TF-5x5/spiral.py

@@ -1,0 +1,116 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+# create identity
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 5
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations (polar form)
+radrange = 4*m.pi # .. rad range
+radres = m.pi/180 # 1deg/step
+steps = radrange/radres # of steps :)
+radarr = [float(i)*radrange/(steps-1) for i in range(round(steps))] # create array w. steps in rad
+curDelta = np.ones((length,length)) # create current delta (error) array initialized w. delta=1
+
+for j in range(length): # rows
+    for i in range(length): # cols
[…]

### 2018-12-21 TF-7x7/X.py

@@ -1,0 +1,98 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 7
+
+deg = 0
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. zeros
[…]

### 2018-12-21 TF-7x7/Y.py

@@ -1,0 +1,98 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 7
+
+deg = 90
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. zeros
[…]

### 2018-12-21 TF-7x7/cone.py

@@ -1,0 +1,93 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 7
+
+deg = -45
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
+            if i==j: # if diagonal hit..
+                scaled = np.matmul(scale(4-i*(3/(length-1)),1),identity) # chronological transformation order: right->left
+                rotated = rotate(scaled, rad)
[…]

### 2018-12-21 TF-7x7/corner(base).py

@@ -1,0 +1,94 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 7
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
+            rad = m.atan2(j,i) # use vector atan2 to determine angle and quadrant of vector [x|y]
+            if i==0 and j==0:
+                scaled = np.matmul(scale(4, 1), identity)  # chronological transformation order: right->left
+                rotated = rotate(scaled, m.pi/4)
[…]

### 2018-12-21 TF-7x7/cross-X.py

@@ -1,0 +1,109 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 7
+
+deg1 = -45
+rad1 = deg1*(np.pi/180.0)
+deg2 = 45
+rad2 = deg2*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
[…]

### 2018-12-21 TF-7x7/drain.py

@@ -1,0 +1,94 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 7
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            y = length - 1 - round(j-(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
+            x = round(i-(length-1)/2) # --> center around midpoint
+            matrixArray[j][i] = normalize(np.array([[1, 0], [0, 1]])) # initialization w. normed identity
[…]

### 2018-12-21 TF-7x7/isotropic.py

@@ -1,0 +1,98 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 7
+
+deg = 0
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. zeros
[…]

### 2018-12-21 TF-7x7/rings.py

@@ -1,0 +1,166 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 7
+
+deg = -45
+rad = deg*(np.pi/180.0)
+
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations (polar form)
+radrange = 2*m.pi
+radres = m.pi/180 # 1deg/step
+steps = radrange/radres
+radarr = [float(i)*radrange/(steps-1) for i in range(round(steps))]
+curDelta = np.ones((length,length))
+
+for j in range(length): # rows
+    for i in range(length): # cols
[…]

### 2018-12-21 TF-7x7/spiral.py

@@ -1,0 +1,149 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+# create identity
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 7
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations (polar form)
+radrange = 12.3*m.pi # .. rad range WARNING: DO NOT TOUCH - experimentally derived parameter!
+radres = m.pi/180 # 1deg/step
+steps = radrange/radres # of steps :)
+radarr = [float(i)*radrange/(steps-1) for i in range(round(steps))] # create array w. steps in rad
+curDelta = np.ones((length,length)) # create current delta (error) array initialized w. delta=1
+
+for j in range(length): # rows
[…]

### 2018-12-21 TF-7x7/star(src-sink).py

@@ -1,0 +1,100 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 7
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            y = round(length - 1 - j -(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
+            x = round(i-(length-1)/2) # --> center around midpoint
+            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
[…]

### 2018-12-21 pv_examples/example.py

@@ -1,0 +1,39 @@
+from vtk.util.vtkAlgorithm import *
+from vtk.vtkCommonDataModel import vtkPolyData
+from vtk.numpy_interface import dataset_adapter as dsa
+from paraview.pythonalgorithm import smproxy, smhint, smdomain, smproperty
+
+
+@smhint.xml('''<ShowInMenu category="My Submenu" />''')
+@smproperty.input(name='Input', port_index=0)
+@smproxy.filter(label='My Example Filter')
+@smdomain.datatype(dataTypes=['vtkPolyData'])
+class MyExampleFilter(VTKPythonAlgorithmBase):
+    def __init__(self):
+        VTKPythonAlgorithmBase.__init__(self, nInputPorts=1, nOutputPorts=1, outputType='vtkPolyData')
+ 
+    def FillInputPortInformation(self, port, info):
+        info.Set(self.INPUT_REQUIRED_DATA_TYPE(), 'vtkPolyData')
+        return 1
+ 
+    def RequestData(self, request, inInfo, outInfo):
+        inp = vtkPolyData.GetData(inInfo[0], 0)
+        output = vtkPolyData.GetData(outInfo, 0)
+        if inp is None or output is None:
+            return 0
+        # copy input to output
+        output.ShallowCopy(inp)
+        # wrap input for numpy interface
+        inp = dsa.WrapDataObject(inp)
+        my_array = inp.PointData['Array']
+        print(my_array)
+        return 1
+
+ 
+if __name__ == '__main__':
+    from paraview.pythonalgorithm import get_plugin_xmls
+    from xml.dom.minidom import parseString
+    for xml in get_plugin_xmls(globals()):
+        dom = parseString(xml)
+    print(dom.toprettyxml(' ', '\n'))
+	

### 2018-12-21 symmetric_tensor_field_src.py

@@ -1,0 +1,128 @@
+from paraview.util.vtkAlgorithm import *
+from vtkmodules.vtkCommonDataModel import vtkImageData
+from vtkmodules.numpy_interface import dataset_adapter as dsa
+import numpy as np
+from timeit import default_timer as timer
+ 
+ 
+@smproxy.source(label="Double Gyre Tensor - symmetric")
+@smhint.xml('''<ShowInMenu category="My Sources" />''')
+class DoubleGyreTensor(VTKPythonAlgorithmBase):
+    def __init__(self):
+        self.Dimensions = [10, 10, 10]
+        self.T = 10.0
+        self.A = 0.1
+        self.eps = 0.1
+        self.Vectorize = True
+        VTKPythonAlgorithmBase.__init__(self, nInputPorts=0, nOutputPorts=1, outputType='vtkImageData')
+ 
+    @smproperty.intvector(name='Dimensions', default_values=[101, 101, 101])
+    def SetDimensions(self, dx, dy, dz):
+        self.Dimensions = [dx, dy, dz]
+        self.Modified()
+ 
+    @smproperty.doublevector(name='T', default_values=10.0)
+    def SetT(self, T):
+        self.T = T
+        self.Modified()
+ 
+    @smproperty.doublevector(name='A', default_values=0.1)
+    def SetA(self, A):
+        self.A = A
+        self.Modified()
+ 
+    @smproperty.doublevector(name='eps', default_values=0.1)
+    def Seteps(self, eps):
+        self.eps = eps
+        self.Modified()
+ 
+    @smproperty.intvector(name='Vectorize', default_values=1)
+    @smdomain.xml('''<BooleanDomain name="bool" />''')
+    def SetVectorize(self, vectorize):
+        self.Vectorize = vectorize
+        self.Modified()
+ 
+    def RequestInformation(self, request, inInfo, outInfo):
+        executive = self.GetExecutive()
+        info = outInfo.GetInformationObject(0)
+        info.Set(executive.WHOLE_EXTENT(),
+                 [0, self.Dimensions[0] - 1, 0, self.Dimensions[1] - 1, 0, self.Dimensions[2] - 1], 6)
+        return 1
+ 
+    def double_gyre_vectorized(self, x, y, t):
+		a = self.eps * np.sin(2.0 * np.pi / self.T * t)
[…]

### 2018-12-21 tensor_field_generator.py

@@ -1,0 +1,140 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 5
+
+deg = -45
+rad = deg*(np.pi/180.0)
+
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations (cartesian form)
+# with open('temp.txt', 'wb') as f:
+#     for j in range(length): # rows
+#         for i in range(length): # cols
+#             matrixArray[j][i] = normalize(np.array([[1, 0],[0, 1]])) # initialization w. normed identity
[…]

### 2018-12-21 tensor_field_src.py

@@ -1,0 +1,128 @@
+from paraview.util.vtkAlgorithm import *
+from vtkmodules.vtkCommonDataModel import vtkImageData
+from vtkmodules.numpy_interface import dataset_adapter as dsa
+import numpy as np
+from timeit import default_timer as timer
+ 
+ 
+@smproxy.source(label="Double Gyre Tensor")
+@smhint.xml('''<ShowInMenu category="My Sources" />''')
+class DoubleGyreTensor(VTKPythonAlgorithmBase):
+    def __init__(self):
+        self.Dimensions = [101, 101, 101]
+        self.T = 10.0
+        self.A = 0.1
+        self.eps = 0.1
+        self.Vectorize = True
+        VTKPythonAlgorithmBase.__init__(self, nInputPorts=0, nOutputPorts=1, outputType='vtkImageData')
+ 
+    @smproperty.intvector(name='Dimensions', default_values=[101, 101, 101])
+    def SetDimensions(self, dx, dy, dz):
+        self.Dimensions = [dx, dy, dz]
+        self.Modified()
+ 
+    @smproperty.doublevector(name='T', default_values=10.0)
+    def SetT(self, T):
+        self.T = T
+        self.Modified()
+ 
+    @smproperty.doublevector(name='A', default_values=0.1)
+    def SetA(self, A):
+        self.A = A
+        self.Modified()
+ 
+    @smproperty.doublevector(name='eps', default_values=0.1)
+    def Seteps(self, eps):
+        self.eps = eps
+        self.Modified()
+ 
+    @smproperty.intvector(name='Vectorize', default_values=1)
+    @smdomain.xml('''<BooleanDomain name="bool" />''')
+    def SetVectorize(self, vectorize):
+        self.Vectorize = vectorize
+        self.Modified()
+ 
+    def RequestInformation(self, request, inInfo, outInfo):
+        executive = self.GetExecutive()
+        info = outInfo.GetInformationObject(0)
+        info.Set(executive.WHOLE_EXTENT(),
+                 [0, self.Dimensions[0] - 1, 0, self.Dimensions[1] - 1, 0, self.Dimensions[2] - 1], 6)
+        return 1
+ 
+    def double_gyre_vectorized(self, x, y, t):
+        a = self.eps * np.sin(2.0 * np.pi / self.T * t)
[…]

### 2019-01-30 C++ Projects/sfml2DTensorVisV1Node-BasedHybridLobes.cpp
add doc

@@ -104,7 +104,7 @@ class neighborhood
 public:
 
 	//constructor
-	neighborhood(int j, int i, const int length, std::vector<bool>& processMap) // check the neighbourhood of (j,i) for missing neighbors..
+	neighborhood(int j, int i, const int length) // check the neighbourhood of (j,i) for missing neighbors..
 	{
 		// check if frame exceeded, outside FOV, offscreen..
 		if (i == 0)
@@ -116,21 +116,6 @@ public:
 		else if (j == height - 1)
 			neighborB = false; // bottom
 		// leave order.. functional!
-
-		// check neighbors in process(ed) map, for neighbors already processed ..
-		if (neighborR) // if neighbor existent..
-			if (processMap.at(i + 1 + j * width) == true)
-				neighborR = false; // right
-		if (neighborT)
-			if (processMap.at(i + (j - 1) * width) == true)
-				neighborT = false; // top
-		if (neighborL)
-			if (processMap.at(i - 1 + j * width) == true)
-				neighborL = false; // left
-		if (neighborB)
-			if (processMap.at(i + (j + 1) * width) == true)
-				neighborB = false; // bottom
-		// ordered again..
 	}
 
 	bool getR() { return neighborR; }
@@ -765,58 +750,32 @@ class propagator
 	// define principal central directions array (12 central directions in 2D --> 30 in 3D, ufff..)
 	std::array<double, 12> centralDirections{ 3.0 / 2 * M_PI + 1.0172232666228471, 0, 0.5535748055013016, 1.0172232666228471, M_PI / 2, M_PI / 2 + 0.5535748055013016, M_PI / 2 + 1.0172232666228471, M_PI, M_PI + 0.5535748055013016, M_PI + 1.0172232666228471, 3.0 / 2 * M_PI, 3.0 / 2 * M_PI + 0.5535748055013016 };
 	std::array<double, 12> apertureAngles{ beta, alpha, beta, beta, alpha, beta, beta, alpha, beta, beta, alpha, beta }; // define aperture angles array in order
-	std::array<int, 12> coneDirections{ 3,0,1,0,1,2,1,2,3,2,3,0 };
-
-	int jIndex = 0;
-	int iIndex = 0;
-
-	std::array<int, 4> centralEdgeIndex{ 0,1,2,3 };
 
 	// define function parser (muparser)
-	double tinc = 2 * pi / 72; // set theta increment tinc
[…]

### 2019-01-30 C++ Projects/sfmlTensorVisV1Array.cpp

@@ -924,12 +924,12 @@ void computeGlyphs(std::vector<std::string>& functionStrEllipse, std::vector<std
 		}
 		else // if mirroring at one single axis.. swap!
 		{
-			if (abs(yy) < abs(xx)) // if anisotropic x-scaling, scale x
+			if (abs(yy) < abs(xx)) // if anisotropic x-scaling, scale y
 			{
 				a = sv2;
 				b = sv1;
 			}
-			else // if anisotropic y-scaling, scale y
+			else // if anisotropic y-scaling, scale x
 			{
 				a = sv1;
 				b = sv2;

### 2019-03-29 C++ Projects/sfml2DTensorVisV1FinalShort.cpp
update

@@ -780,14 +780,15 @@ void computeGlyphs(std::vector<std::vector<double>>& glyphBuffer, std::vector<st
 	// define parameters
 	double radres = (2 * pi)/steps;
 	
+	std::vector<std::complex<double>> result(2, std::complex<double>(0.0));
 	std::vector<bool> signs(3, false);
 	// iterate through the matrixList/svdList (grid) and construct (scaled) ellipses in polar form (function) from the repsective singular values/vectors
 	for (int i = 0; i < matrixList.size(); i++)
 	{
-		double y1 = svdList.at(i).matrixU().col(0)[1]; // use x - coordinate of both semi-axes 
-		double x1 = svdList.at(i).matrixU().col(0)[0]; // use x - coordinate of both semi-axes "sigma_xx"
-		double y2 = svdList.at(i).matrixU().col(1)[1]; // use x - coordinate of both semi-axes 
-		double x2 = svdList.at(i).matrixU().col(1)[0]; // use x - coordinate of both semi-axes "sigma_xx"
+		double y1 = svdList.at(i).matrixU().col(0)[1]; // use x - coordinate of both semi-axes -- Get LEFT U-vector
+		double x1 = svdList.at(i).matrixU().col(0)[0]; // use x - coordinate of both semi-axes
+		double y2 = svdList.at(i).matrixU().col(1)[1]; // use x - coordinate of both semi-axes -- Get RIGHT U-vector
+		double x2 = svdList.at(i).matrixU().col(1)[0]; // use x - coordinate of both semi-axes
 		double xx = matrixList.at(i).row(0)[0]; // "sigma_xx"
 		double xy = matrixList.at(i).row(0)[1]; // "sigma_xy"
 		double yx = matrixList.at(i).row(1)[1]; // "sigma_yx"
@@ -796,33 +797,42 @@ void computeGlyphs(std::vector<std::vector<double>>& glyphBuffer, std::vector<st
 		double deg2 = atan2(y2, x2) * 180.0 / M_PI; // use vector atan2 to get rotational angle (phase) of both basis vectors [-180°,180°]
 
 		glyphParameters.at(i).at(2) = deg1;
+		
+		// calculate principal stresses w. formula.. https://vergleichsspannung.de/vergleichsspannungen/normalspannungshypothese-nh/herleitung-der-hauptspannungen/
[…]

### 2019-04-02 C++ Projects/sfml2DTensorVisEllipses.cpp
add

@@ -15,6 +15,14 @@
 #include <functional>
 
 #define MAXBUFSIZE  ((int) 1e3)
+#ifdef WINDOWS
+#include <direct.h>
+#define GetCurrentDir _getcwd
+#else
+#include <unistd.h>
+#define GetCurrentDir getcwd
+#endif
+std::string workDir;
 
 using namespace std;
 using namespace Eigen;
@@ -22,6 +30,9 @@ using namespace Eigen;
 //length and width of window
 int wSize = 700;
 sf::Vector2i windowsize;
+int width;
+int height;
+int steps;
 
 //recording frameskip 
 int record_frameskip = -1;
@@ -58,7 +69,13 @@ std::vector<T> operator+(const std::vector<T>& a, const std::vector<T>& b)
 	return result;
 }
 
-
+// get current working directory to assign matrix.txt path
+std::string GetCurrentWorkingDir(void) {
+	char buff[FILENAME_MAX];
+	GetCurrentDir(buff, FILENAME_MAX);
+	std::string current_working_dir(buff);
+	return current_working_dir;
+}
 // CLASS DEFINITIONS //
 class neighborhood
 {
@@ -117,6 +134,8 @@ class polarPlot
 	// draw speed
 	int speed;
 	int line_width;
+	int steps;
+	int ctr = 0;
 	// theta
 	double t;
 	// radius
@@ -124,11 +143,14 @@ class polarPlot
 	double tmax;
 	double tinc;
 	double rotation;
+	double radres;
 	sf::Color color;
+	std::vector<double> sample;
+	bool discretePlot = false;
 
 public:
 
-	//constructor
+	//constructor #1 string parser - muParser: tinc_in passed as tinc used for parsing
 	polarPlot(std::string tag,
 		int speed_in,
 		int line_width_in,
@@ -149,6 +171,53 @@ public:
 		p.SetExpr(tag);
 	}
 
+	//constructor #2 array parser w. array-inherited resolution radres
+	polarPlot(std::vector<std::vector<double>> SampleBuffer,
+		int speed_in,
+		int line_width_in,
+		int steps_in,
+		double tmax_in,
+		double res,
+		double rotation_in,
+		sf::Color color_in)
+
+		: speed(speed_in),
+		line_width(line_width_in),
+		steps(steps_in),
+		t(0),
+		tmax(tmax_in),
+		radres(res),
+		rotation(rotation_in),
+		color(color_in)
+	{
+		discretePlot = true;
[…]

### 2019-04-04 C++ Projects/sfml2DTensorVisV1FinalShort(GIGradient).cpp
update thesis

@@ -402,18 +402,45 @@ void parse_options(int argc, char* argv[], std::vector<std::string>& funcs, std:
 }
 
 // OPERATOR DEFINITIONS //
-template <typename T> // element-wise plus for std::vector
+//template <typename T> // element-wise plus for std::vector
+//std::vector<T> operator+(const std::vector<T>& a, const std::vector<T>& b)
+//{
+//	std::vector<T> result(b.size());
+//
+//	for (int i = 0; i < b.size(); i++)
+//		result.at(i) = a.at(i) + b.at(i);
+//
+//	return result;
+//}
+
+// vector PLUS(+) and MINUS(-)
+template <typename T>
 std::vector<T> operator+(const std::vector<T>& a, const std::vector<T>& b)
 {
-	std::vector<T> result(b.size());
+	assert(a.size() == b.size());
 
-	for (int i = 0; i < b.size(); i++)
-		result.at(i) = a.at(i) + b.at(i);
+	std::vector<T> result;
+	result.reserve(a.size());
 
+	std::transform(a.begin(), a.end(), b.begin(),
+		std::back_inserter(result), std::plus<T>());
 	return result;
 }
 
-template <typename T> // element-wise plus for std::vector
+template <typename T>
+std::vector<T> operator-(const std::vector<T>& a, const std::vector<T>& b)
+{
+	assert(a.size() == b.size());
+
+	std::vector<T> result;
+	result.reserve(a.size());
+
+	std::transform(a.begin(), a.end(), b.begin(),
+		std::back_inserter(result), std::minus<T>());
+	return result;
+}
+
+template <typename T> // element-wise multiplication for number, std::vector
 std::vector<T> operator*(const T a, const std::vector<T>& b)
 {
 	std::vector<T> result(b.size());
@@ -424,16 +451,18 @@ std::vector<T> operator*(const T a, const std::vector<T>& b)
 	return result;
 }
 
-template <typename T> // element-wise plus for std::vector
+template <typename T> // element-wise minus for vector of vectors
 std::vector<std::vector<T>> operator-(const std::vector<std::vector<T>>& a, const std::vector<std::vector<T>>& b)
 {
 	std::vector<std::vector<T>> result(b.size());
 
 	for (int i = 0; i < b.size(); i++)
-		result.at(i) = a.at(i) + (-1.0)*b.at(i);
[…]

### 2019-04-05 C++ Projects/sfml2DTensorVisV1FinalShort(GIGradient).cpp
update code

@@ -441,7 +441,7 @@ std::vector<T> operator-(const std::vector<T>& a, const std::vector<T>& b)
 }
 
 template <typename T> // element-wise multiplication for number, std::vector
-std::vector<T> operator*(const T a, const std::vector<T>& b)
+std::vector<T> operator*(const T& a, const std::vector<T>& b)
 {
 	std::vector<T> result(b.size());
 
@@ -645,6 +645,13 @@ std::vector<double> sample(double(*f)(double x), std::vector<std::vector<double>
 	return sample;
 }
 
+int fast_mod(const int input, const int ceil) {
+	// apply the modulo operator only when needed
+	// (i.e. when the input is greater than the ceiling)
+	return input >= ceil ? input % ceil : input;
+	// NB: the assumption here is that the numbers are positive
+}
+
 class propagator
 {
 	// set principal arture angles in rads
@@ -678,7 +685,7 @@ class propagator
 	bool flag = false;
 
 	std::vector<double> initArray;
-	std::vector<int> deltaIndex;
+	std::vector<int> deltaIndex{ 1, 1 - width, -width, -1 - width, -1, -1 + width, width, 1 + width };
 public:
 	propagator(const int dim, double* mean, std::vector<std::vector<double>>* sampleBuffA, std::vector<std::vector<double>>* sampleBuffB, std::vector<std::vector<double>>* ellipseArray)
 	{
@@ -694,7 +701,6 @@ public:
 		out = std::vector<double>(steps, 0.0);
 		initArray = std::vector<double>(steps, 0.0);
 
-		deltaIndex = { 1, 1 - width, -width, -1 - width, -1, -1 + width, width, 1 + width };
 		double energy_sum = 0.0;
 		for (int k = 0; k < 8; k++) // for each node..
 		{
@@ -742,13 +748,13 @@ public:
 			for (int j = 0; j < steps; j++)
 			{
 				sum1 += read.at(j); // evaluate for iMean (sum1)
-				sum2 += glyphBuffer->at(i).at(j);
-				sum3 += read.at(j)*glyphBuffer->at(i).at(j);
+				sum2 += glyph.at(j);
+				sum3 += read.at(j)*glyph.at(j);
 
-				if (read.at(j) < 0)
+				/*if (read.at(j) < 0)
[…]

### 2019-04-10 C++ Projects/sfml2DTensorVisV1FinalShort(energy-tests).cpp
update

@@ -22,17 +22,18 @@
 #include <math.h>
 #include <Eigen/Dense>
 #include <Eigen/Eigenvalues>
-#include <array>
+//#include <array>
 #include <filesystem>
 #include <string>
 #include <algorithm> 
-#include <cctype>
-#include <locale>
-#include <functional>
+//#include <cctype>
+//#include <locale>
+//#include <functional>
 #include <numeric>
 #include <boost/iostreams/stream.hpp>
 #include <boost/iostreams/tee.hpp>
 #include <limits>
+#include <ctime>
 
 // NAMESPACE IMPORTS
 using namespace std;
@@ -47,6 +48,9 @@ typedef std::numeric_limits< double > dbl;
 std::string functionString = "2.1"; // Prototyping of functionString for strFunction (symbolic string arg parsed by muparser in cpp functional ptr rep)!
 int width;
 int height;
+int steps; // use n steps for angular resolution - radres
+double radres;
+
 template <typename T>
 T clip(const T& n, const T& lower, const T& upper); // template clip function
 
@@ -61,7 +65,14 @@ const double pi = M_PI;
 
 // -> muparser custom clip function: template functions need to be static callback (call after) functions, which can be passed as arguments
 static value_type clipNeg(value_type v) { return clip(v, 0.0, 1.0); }
-
+static value_type deltaFunction(value_type v)
+{
+	double radres = 2 * pi / steps;
+	if (abs(v) < radres / 2.0 || v == radres / 2.0)
+		return steps;
+	else
+		return 0.0;
+}
 //writes frames to file 
 //directory must already exist -- parse file methods
 void record(sf::RenderTexture& in, int frameskip, std::string directory)
@@ -111,9 +122,30 @@ class neighborhood
 
 public:
 
+	neighborhood() // standard constructor
+	{}
 	//constructor
-	neighborhood(int j, int i, const int length) // check the neighbourhood of (j,i) for missing neighbors..
+	neighborhood(int j, int i) // check the neighbourhood of (j,i) for missing neighbors..
 	{
+		// check if frame exceeded, outside FOV, offscreen..
+		if (i == 0)
+			neighborL = false; // left
+		else if (i == width - 1)
+			neighborR = false; // right
[…]

### 2019-04-10 C++ Projects/sfml2DTensorVisV1FinalShort.cpp

@@ -1072,9 +1072,9 @@ public:
 					//}
 				}
 
-				out = cosines.at
-				// multiply respective c(k); // assign cosine cone w.r.t cone direction (index) k
-osine cone by valsum*=radres, because of energy normalization to cosine_sum (pre-computed in constructor)
+				out = cosines.at(k); // assign cosine cone w.r.t cone direction (index) k
+
+				// multiply respective cosine cone by valsum*=radres, because of energy normalization to cosine_sum (pre-computed in constructor)
 				std::transform(out.begin(), out.end(), out.begin(), std::bind(std::multiplies<double>(), std::placeholders::_1, val_sum *= radres));
 
 				*meanA += val_sum;

### 2019-04-10 TF-13x13/isotropic.py

@@ -51,7 +51,7 @@ def file_len(fname):
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
 zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
-length = 29
+length = 69
 
 deg = 0
 rad = deg*(np.pi/180.0)

### 2019-05-22 WissArbIEEE/shortpaper_sebastian_bek.tex
add old shortpaper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

%\documentclass[letterpaper, 10 pt, conference]{orbieeeconfpre}  % Comment this line out if you need a4paper

\documentclass[a4paper, 10pt, journal]{wissarbIEEE}      % Use this line for a4 paper
%\conference{IEEE Conference for Awesome ORB Research}

\bibliographystyle{orbref-num}

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{lipsum}

\title{\LARGE \bf
The crowd congestion level - a new measure for risk assessment in video-based crowd monitoring
}

\author{Sebastian Bek$^{1}$ and Dr.-Ing. Eduardo Monari$^{2}$% <-this % stops a space
\thanks{*Based on the guidelines published on the \href{http://conf.papercept.net/conferences/support/tex.php}{PaperCept conference manuscript management website}}% <-this % stops a space
\thanks{$^{1}$Sebastian Bek is with the master program for Applied Computer Science at the Institute for Computer Science, Heidelberg University
        {\tt\small sebibek@gmail.com}}%
\thanks{$^{2}$Dr.-Ing. Eduardo Monari, is with the Department of Video Exploitation Systems (VID), Fraunhofer Institute for Optronics System Technology and Image Exploitation (IOSB), Fraunhoferstraße 1, 76131 Karlsruhe
        {\tt\small eduardo.monari@iosb.fraunhofer.de, \href{https://www.iosb.fraunhofer.de}{}}}%
}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

In this paper, we propose a new characteristic measure relative crowd density and motion dynamics for the purpose of long-term crowd monitoring. Furthermore, we will discuss the derivation of a so-called Congestion Level of local areas in the crowd, which takes the current dynamics and density within a certain image region into account.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
The visitor rates at public events are increasing steadily since several years. In particular, besides large events in closed-area environments (e.g. stadiums, festival areas, etc.) crowds have also become a common phenomenon in public urban environment. In such environments video-based monitoring might be useful for both, safety-related and security-related applications.

\section{State-of-the-art approaches}
\label{sec:sota}
For automated counting of people in images, there exist two main approaches, the so-called direct, and the indirect approach. In direct approaches people are first detected, segmented from background and then counted. Examples for direct approaches can be found in e.g. \cite{rittscher}. The indirect approach however, is considered to be more robustly, since the individual segmentation of persons as foreground regions is itself a challenging problem, that is not solved reliably yet. Yet a promising approach has been proposed by Albiol et al. \cite{albiol}, whereas the number of moving corner features is used to estimate the number of persons in the image region.

\section{Proposed Method}
\label{sec:method}
Since the estimation of dynamics is useful for indication of congestion and flow jams in a crowd, it would be interesting to gather information about them as well. In our approach we use motion (tracks) in the scene as basic features, as well as relative changes in track velocities (inertia) for generation of a so-called Congestion Level. We observed that even in scenarios with high people density, the situation can be regarded as non-critical, in case people can still move freely and smoothly through the crowd. As a consequence, we believe that information on the flow dynamics should be taken into account for risk assessment. Our approach assumes, that a local spot in the crowd might be critical, if the density is continuously increasing (relative density) over time, and simultaneously a signiﬁcant reduction of motion dynamics (increasing inertia) is observed. The proposed approach is an indirect one, since it is considered to be more robust in case of small objects in the scene. The approach used is similar to the one proposed by Albiol at Al. in \cite{albiol}. However, in addition to just clustering and counting moving features (people counting), we extend the motion vector extraction by multi-frame feature tracking and by estimation of densities and dynamics. The approach used to obtain trajectories from motion features is as follows: (1) Detection of a set of Harris Corners \cite{harris} as pixel coordinates $\mathcal{F} = \{\mathbf{f}_1, \mathbf{f}_2, ..., \mathbf{f}_n\}$, (2) estimation of LK optical flow \cite{LK} for detected features and assignment of resulting motion vectors $\mathcal{V} = \{\mathbf{v}_1, \mathbf{v}_2, ..., \mathbf{v}_n\}$ to each feature, and (3) classiﬁcation of features/motion vectors above a minimum motion threshold β as moving features:

%% Figure
\begin{figure}[h]
   \centering
   \includegraphics[width=0.35\textwidth]{fig/trajektorien.png}
   \caption{Acquisition of Trajectories (tracklets).}
   \label{fig:trajectories}
\end{figure}

\begin{equation}
	\mathcal{F'} = \{f_i\in \mathcal{F} | \ |v_i| \ge \beta\} \hskip 5pt {with} \ i = \{1..n\}
\end{equation}

Here $f_i=(x,y)^T$ is a feature point located at pixel position $(x,y)$ while $v_i=(\Delta x,\Delta y)$ is the corresponding motion vector, indicating the absolute movement of the feature between the preceding and the current frame. Finally in step (4) motion feature tracking is performed by association of detected moving features to trajectory candidates. In most scenes, this simple classiﬁcation will sort out almost any detected feature point not corresponding to a moving object (e.g. person). Additionally to obtain information about dynamics and flow, the feature points are used to generate tracks, by concatenation of motion vectors. Hereby, we first try to extend previously created trajectories evaluating the distance between new detected features in the last video frame, and the estimated position of features in the previous video frames, shifted by the optical flow motion vector, which is basically a motion prediction. If no previously created trajectories are found in a defined neighborhood, a new trajectory with the corresponding motion vector is created. Finally, we obtain a set of tracks $\mathcal{T}^k = \{\mathcal{T}_1^k, \mathcal{T}_2^k, ..., \mathcal{T}_m^k\}$, summarizing all available $m = |\mathcal{F'}|$ tracks at time $k$. Each track in turn is basically a set of features $\mathcal{T}_j^k = \{ \mathbf{f}^k, \mathbf{f}^{k-1}, ..., \mathbf{f}^{k-s_j}\}, j\in\{1, ..., m\}$, and $s_j$ the length of the track $j$. 
 The trajectories are assigned with IDs for proper identification and managed in a track list. Fig. \ref{fig:trajectories} shows example results of our local feature tracking approach. Now, since we have track information available, we use them to create statistics on track density, dynamics and flow behavior. To generate local, statistics, the image is split into $R$ smaller image patches $\mathcal{P}$ first, whereas $\mathcal{P}_r, r = \{1, ..., R\}$ represents the set of pixels of each patch. The number of estimated persons (local density) is defined as the amount of trajectory (tracklet) tips in a set of feature points $\mathcal{G}_r$ within a patch $\mathcal{P}_r$ in the very last video frame:

\begin{equation}
	d_r = \kappa\cdot|\mathcal{G}_r| \ \ \ \ \ \ \ \ \ [\frac{tracklets}{patch}]
\end{equation}
with $\mathcal{G}_r \supseteq \mathcal{F′}\cap \mathcal{P}_r$.\\

Here $\kappa$ is some heuristically deﬁned scale factor, indicating that the number of feature points (tracklets) can differ from the actual number of persons.

To efficiently sense dynamics as well, we propose a hypothesis about congestions: \emph{Congestions can be interpreted as a discontinuity in track flow, which equals low excentric dynamics.} Accordingly, we want to measure the excentric dynamics of a track. Therefore, we measure the Euclidean length of each track position vector within the last q frames. This length is then averaged over all tracklets within an image patch. We call this measure local flow inertia:

\begin{equation}
	i_r^k = \frac{1}{|\mathcal{G}_r'|}\sum_{\forall f\in\mathcal{G}_r'} ||f^k, f^{k-q}||_2 \ \ \ \ \ \ \ \ \ [\frac{{px}}{q\cdot {frame}}] 
\end{equation}

with $\mathcal{G}_r′ = \{T\in\mathcal{G}_r | \ |T| \ge q\}$.\\

For higher robustness of the approach, only tracks with a minimum length of q are taken into account. These tracks are summarized by $\mathcal{G}_r′$, as a subset of $\mathcal{G}_r$.

To measure the risk level for the people in the crowd, we propose a combined coefficient, we call congestion level (${cl}$), which incorporates the estimated inertia and dynamics in form of relative (${rel}$) estimates. To generate relative estimates, it is required to set proper thresholds, obtained by theoretical or experimental analysis. Our derived congestion level is defined as the product of the previously decribed characteristic figures $i_{rel}$ and $d_{rel}$:

\begin{equation}
	{cl} = d_{rel}\cdot i_{rel} \ \ \ \ \ \ \ \ \ [\%] 
\end{equation}

\section{Evaluation}
\label{sec:eval}

To demonstrate the plausibility of the derived measures, we first compared the obtained track density with manually generated ground truth data. Second, we evaluate the difference between the proposed congestion level and the popular density metric. We used both artiﬁcial and real video data publicly available. In the following table we compare the estimated density with manually generated ground-truth data in a certain image patch. Also, we show the impact of the included dynamics measure in the derived congestion level. All measures have been averaged by a temporal mean filter over $10$ frames:

%% Table
\begin{table}[h]
%\tiny
\centering
\caption{Evaluation of Estimates (averaged over $10$ frames).}
   \begin{tabularx}{0.5\textwidth}{p{0.03\textwidth}p{0.02\textwidth}p{0.02\textwidth}p{0.02\textwidth}p{0.02\textwidth}p{0.02\textwidth}p{0.02\textwidth}p{0.02\textwidth}p{0.02\textwidth}p{0.02\textwidth}p{0.02\textwidth}}
   		 \toprule 
   		 fr. & $10$ & $20$ & $30$ & $40$ & $50$ & $60$ & $70$ & $80$ & $90$ & $100$ \\
   		 \midrule
		 GT. & $0.34$ & $0.37$ & $0.39$ & $0.46$ & $0.53$ & $0.62$ & $0.67$ & $0.81$ & $0.89$ & $0.95$ \\
		 \midrule
		 rD. & $0.35$ & $0.45$ & $0.49$ & $0.54$ & $0.59$ & $0.65$ & $0.72$ & $0.81$ & $0.88$ & $0.93$ \\
		 \midrule
		 cl. & $0.35$ & $0.07$ & $0.09$ & $0.01$ & $0.19$ & $0.34$ & $0.49$ & $0.81$ & $0.88$ & $0.93$ \\ 
		 \midrule
   \end{tabularx}  \label{tab:eval}
   \vskip 1pt
   \begin{tiny}
	\textbf{index:} \hskip 5pt fr.: frame; \hskip 5pt GT.: ground-truth relative density; \hskip 5pt rD.: relative density; \hskip 5pt cl: congestion level;
	\end{tiny}
\end{table}
It can be observed, that the estimated relative density follows the ground-truth density and is therefore a reasonable measure for risk assessment. Additionally, dynamics are taken into account with the congestion level. The congestion level rises with a little delay, which represents the growing process of congestions/jams. At first, people are congested from high densities. As a consequence, dynamics decrease rapidly. Thus, the congestion level increases a little delayed because the congestions themselves need high densities as a condition to evolve over time.

\section{Conclusion}
\label{sec:conclusion}
In this paper, we proposed a new characteristic measure for density-related risk assessment in crowd analysis, we call 
Congestion Level. This measure indicates the endangering of local areas in a crowd, due to increasing people density by simultaneous reduction of motion dynamics (stop-and- go or slackening of crowd motion). It has been shown that the proposed Congestion Level provides a suitable measure for dynamics and density, which might be of interest for
(semi-)automated risk assessment systems. Future works might include evaluation of applicability for risk assessment in practice.

\bibliography{mybibfile}

\end{document}

### 2019-05-22 TF-13x13/rings.py
add

@@ -49,7 +49,8 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 79# use width+50 because of subsequent cropping for 13x13 fields --> hack to prevent circular function errors
+width = 117
+length = width+50# use width+50 because of subsequent cropping for 13x13 fields --> hack to prevent circular function errors
 
 deg = -45
 rad = deg*(np.pi/180.0)
@@ -133,5 +134,5 @@ with open('temp.txt', 'r+') as txtfile:
                 str1 = str2 = "" # ..reset line strings
 
             i += 1 #increment line index
-
+print("line_index: " + str(i)) # print len of matrixArray: should equal length
 os.remove('temp.txt')

### 2019-05-22 TF-13x13/spiral.py

@@ -47,7 +47,7 @@ def file_len(fname):
 
 # create identity
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 29+50
+length = 117+50
 matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
 
 # generate normalized tensors from linear transformations (polar form)

### 2019-05-22 TF-13x13/star(src-sink).py

@@ -49,7 +49,7 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 13
+length = 117
 matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
 
 # generate normalized tensors from linear transformations

### 2019-05-24 TF-13x13/spiral.py
update

@@ -51,7 +51,7 @@ length = 117+50
 matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
 
 # generate normalized tensors from linear transformations (polar form)
-radrange = 200*m.pi # .. rad range WARNING: DO NOT TOUCH - experimentally derived parameter!
+radrange = 260*m.pi # .. rad range WARNING: DO NOT TOUCH - experimentally derived parameter!
 radres = m.pi/180 # 1deg/step
 steps = radrange/radres # of steps :)
 radarr = [float(i)*radres for i in range(round(steps))] # create array w. steps in rad

### 2019-05-27 TF-13x13/corner(base).py
add

@@ -46,7 +46,7 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 29
+length = 13
 matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
 
 # generate normalized tensors from linear transformations

### 2019-05-29 TF-13x13/drain_old.py
update

@@ -46,14 +46,14 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 117
+length = 13
 matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
 
 # generate normalized tensors from linear transformations
 with open('temp.txt', 'wb') as f:
     for j in range(length): # rows
         for i in range(length): # cols
-            y = length - 1 - round((length-1)/2-j) # use inverted y coordinates (mirroring at x-axis in array reference frame)
+            y = length - 1 - round(j-(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
             x = round(i-(length-1)/2) # --> center around midpoint
             matrixArray[j][i] = normalize(np.array([[1, 0], [0, 1]])) # initialization w. normed identity
             rad = m.atan2(y,x) # use vector atan2 to determine angle and quadrant of vector [x|y]

### 2019-06-10 C++ Projects/GIGradient(on-the-fly)-1D-opt.cpp
update

@@ -1,1100 +1,0 @@
-// INLINE DEFINES
-
-#ifdef WINDOWS
-#include <direct.h>
-#define GetCurrentDir _getcwd
-#else
-#include <unistd.h>
-#define GetCurrentDir getcwd
-#endif
-
-#define MAXBUFSIZE  ((int) 1e8)
-#define _USE_MATH_DEFINES
-
-
-// INCLUDES (IMPORTS)
-//#include <SFML/Graphics.hpp>
-#include <iostream>
-//#include "muParser.h"
-#include <sstream>
-#include <fstream>
-#include <vector>
-//#include <unistd.h>
-#include <math.h>
-#include <eigen3/Eigen/Dense>
-#include <eigen3/Eigen/Eigenvalues>
-//#include <array>
-#include <experimental/filesystem>
-#include <string>
-//#include <algorithm> 
-//#include <cctype>
-//#include <locale>
-//#include <functional>
-#include <numeric>
-#include <boost/iostreams/stream.hpp>
-#include <boost/iostreams/tee.hpp>
-#include <limits>
-#include <ctime>
-
-// VTK Includes
-//#include <vtkVersion.h>
-#include <vtkSmartPointer.h>
-#include <vtkXMLImageDataWriter.h>
-#include <vtkImageData.h>
-#include <vtkPointData.h>
-#include <vtkDoubleArray.h>
-
-//thrust
-#include <thrust/device_vector.h>
-#include <thrust/transform.h>
-#include <thrust/copy.h>
-#include <thrust/fill.h>
-#include <thrust/functional.h>
-
-// NAMESPACE IMPORTS
-using namespace std;
-using namespace Eigen;
-
-typedef boost::iostreams::tee_device<std::ostream, std::ofstream> Tee;
-typedef boost::iostreams::stream<Tee> TeeStream;
-typedef std::numeric_limits< double > dbl;
-
-double buffer[MAXBUFSIZE];
-
-//definition of pi
-const double pi = M_PI;
-
-// PROTOTYPES
-int width;
-int height;
-int steps; // use n steps for angular resolution - radres
-double radres;
-
-template <typename T>
-T clip(const T& n, const T& lower, const T& upper); // template clip function
-
-//length and width of window
-int wSize = 701;
-
-// GENERIC FUNCTION DEFINITIONS
-
-// get current working directory to assign matrix.txt path
-std::string GetCurrentWorkingDir(void) {
[…]

### 2019-06-10 C++ Projects/GIGradient(on-the-fly)-1D-simple.cpp

@@ -1,1083 +1,0 @@
-// INLINE DEFINES
-
-#ifdef WINDOWS
-#include <direct.h>
-#define GetCurrentDir _getcwd
-#else
-#include <unistd.h>
-#define GetCurrentDir getcwd
-#endif
-
-#define MAXBUFSIZE  ((int) 1e8)
-#define _USE_MATH_DEFINES
-
-
-// INCLUDES (IMPORTS)
-//#include <SFML/Graphics.hpp>
-#include <iostream>
-//#include "muParser.h"
-#include <sstream>
-#include <fstream>
-#include <vector>
-//#include <unistd.h>
-#include <math.h>
-#include <eigen3/Eigen/Dense>
-#include <eigen3/Eigen/Eigenvalues>
-//#include <array>
-#include <experimental/filesystem>
-#include <string>
-//#include <algorithm> 
-//#include <cctype>
-//#include <locale>
-//#include <functional>
-#include <numeric>
-#include <boost/iostreams/stream.hpp>
-#include <boost/iostreams/tee.hpp>
-#include <limits>
-#include <ctime>
-
-// VTK Includes
-//#include <vtkVersion.h>
-#include <vtkSmartPointer.h>
-#include <vtkXMLImageDataWriter.h>
-#include <vtkImageData.h>
-#include <vtkPointData.h>
-#include <vtkDoubleArray.h>
-
-//thrust
-#include <thrust/device_vector.h>
-#include <thrust/transform.h>
-#include <thrust/copy.h>
-#include <thrust/fill.h>
-#include <thrust/functional.h>
-
-// NAMESPACE IMPORTS
-using namespace std;
-using namespace Eigen;
-
-typedef boost::iostreams::tee_device<std::ostream, std::ofstream> Tee;
-typedef boost::iostreams::stream<Tee> TeeStream;
-typedef std::numeric_limits< double > dbl;
-
-double buffer[MAXBUFSIZE];
-
-//definition of pi
-const double pi = M_PI;
-
-// PROTOTYPES
-int width;
-int height;
-int steps; // use n steps for angular resolution - radres
-double radres;
-
-template <typename T>
-T clip(const T& n, const T& lower, const T& upper); // template clip function
-
-//length and width of window
-int wSize = 701;
-
-// GENERIC FUNCTION DEFINITIONS
-
-// get current working directory to assign matrix.txt path
-std::string GetCurrentWorkingDir(void) {
[…]

### 2019-06-10 C++ Projects/GIGradientUbuntu(on-the-fly).cu

@@ -1,1254 +1,0 @@
-// INLINE DEFINES
-
-#ifdef WINDOWS
-#include <direct.h>
-#define GetCurrentDir _getcwd
-#else
-#include <unistd.h>
-#define GetCurrentDir getcwd
-#endif
-
-#define MAXBUFSIZE  ((int) 1e8)
-#define _USE_MATH_DEFINES
-
-
-// INCLUDES (IMPORTS)
-//#include <SFML/Graphics.hpp>
-#include <iostream>
-//#include "muParser.h"
-#include <sstream>
-#include <fstream>
-#include <vector>
-//#include <unistd.h>
-#include <math.h>
-#include <Eigen/Dense>
-#include <Eigen/Eigenvalues>
-//#include <array>
-#include <experimental/filesystem>
-#include <string>
-//#include <algorithm> 
-//#include <cctype>
-//#include <locale>
-//#include <functional>
-#include <numeric>
-#include <boost/iostreams/stream.hpp>
-#include <boost/iostreams/tee.hpp>
-#include <limits>
-#include <ctime>
-
-// VTK Includes
-//#include <vtkVersion.h>
-#include <vtkSmartPointer.h>
-#include <vtkXMLImageDataWriter.h>
-#include <vtkImageData.h>
-#include <vtkPointData.h>
-#include <vtkDoubleArray.h>
-
-//thrust
-#include <thrust/device_vector.h>
-#include <thrust/transform.h>
-#include <thrust/copy.h>
-#include <thrust/fill.h>
-#include <thrust/functional.h>
-
-// NAMESPACE IMPORTS
-using namespace std;
-using namespace Eigen;
-
-typedef boost::iostreams::tee_device<std::ostream, std::ofstream> Tee;
-typedef boost::iostreams::stream<Tee> TeeStream;
-typedef std::numeric_limits< double > dbl;
-
-double buffer[MAXBUFSIZE];
-//definition of pi
-const double pi = M_PI;
-
-// PROTOTYPES
-int width;
-int height;
-int steps; // use n steps for angular resolution - radres
-double radres;
-
-template <typename T>
-T clip(const T& n, const T& lower, const T& upper); // template clip function
-
-//length and width of window
-int wSize = 701;
-
-// GENERIC FUNCTION DEFINITIONS
-
-// THRUST's parallelized GPU methods
-struct saxpy_functor
-{
-	const double a;
-
-	saxpy_functor(double _a) : a(_a) {}
-
[…]

### 2019-06-10 C++ Projects/GIGradientUbuntu.cpp

@@ -1,1162 +1,0 @@
-// INLINE DEFINES
-
-#ifdef WINDOWS
-#include <direct.h>
-#define GetCurrentDir _getcwd
-#else
-#include <unistd.h>
-#define GetCurrentDir getcwd
-#endif
-
-#define MAXBUFSIZE  ((int) 1e5)
-#define _USE_MATH_DEFINES
-
-// INCLUDES (IMPORTS)
-//#include <SFML/Graphics.hpp>
-#include <iostream>
-//#include "muParser.h"
-#include <sstream>
-#include <fstream>
-#include <vector>
-#include <unistd.h>
-#include <math.h>
-#include <eigen3/Eigen/Dense>
-#include <eigen3/Eigen/Eigenvalues>
-//#include <array>
-#include <experimental/filesystem>
-#include <string>
-//#include <algorithm> 
-//#include <cctype>
-//#include <locale>
-//#include <functional>
-#include <numeric>
-#include <boost/iostreams/stream.hpp>
-#include <boost/iostreams/tee.hpp>
-#include <limits>
-#include <ctime>
-// VTK Includes
-//#include <vtkVersion.h>
-#include <vtkSmartPointer.h>
-#include <vtkXMLImageDataWriter.h>
-#include <vtkImageData.h>
-#include <vtkPointData.h>
-#include <vtkDoubleArray.h>
-
-// NAMESPACE IMPORTS
-using namespace std;
-using namespace Eigen;
-//using namespace mu;
-
-//typedef boost::iostreams::tee_device<std::ostream, std::ofstream> Tee;
-//typedef boost::iostreams::stream<Tee> TeeStream;
-typedef std::numeric_limits< double > dbl;
-
-//definition of pi
-const double pi = M_PI;
-
-// PROTOTYPES
-int width;
-int height;
-int steps; // use n steps for angular resolution - radres
-double radres;
-
-template <typename T>
-T clip(const T& n, const T& lower, const T& upper); // template clip function
-
-//length and width of window
-int wSize = 701;
-//sf::Vector2i windowsize;
-
-// GENERIC FUNCTION DEFINITIONS
-
-// -> muparser custom clip function: template functions need to be static callback (call after) functions, which can be passed as arguments
-//static mu::value_type clipNeg(mu::value_type v) { return clip(v, 0.0, 1.0); }
-//static mu::value_type deltaFunction(mu::value_type v)
-//{
-//	if (abs(v) < radres/2.0 || v == radres/2.0) // use margin related to angular resolution radres..
[…]

### 2019-06-10 C++ Projects/depr/GIGradientUbuntu(on-the-fly).cpp

@@ -1,0 +1,1159 @@
+// INLINE DEFINES
+
+#ifdef WINDOWS
+#include <direct.h>
+#define GetCurrentDir _getcwd
+#else
+#include <unistd.h>
+#define GetCurrentDir getcwd
+#endif
+
+#define MAXBUFSIZE  ((int) 1e5)
+#define _USE_MATH_DEFINES
+
+// INCLUDES (IMPORTS)
+//#include <SFML/Graphics.hpp>
+#include <iostream>
+//#include "muParser.h"
+#include <sstream>
+#include <fstream>
+#include <vector>
+#include <unistd.h>
+#include <math.h>
+#include <eigen3/Eigen/Dense>
+#include <eigen3/Eigen/Eigenvalues>
+//#include <array>
+#include <experimental/filesystem>
+#include <string>
+//#include <algorithm> 
+//#include <cctype>
+//#include <locale>
+//#include <functional>
+#include <numeric>
+#include <boost/iostreams/stream.hpp>
+#include <boost/iostreams/tee.hpp>
+#include <limits>
+#include <ctime>
+// VTK Includes
+#include <vtkVersion.h>
+#include <vtkSmartPointer.h>
+#include <vtkXMLImageDataWriter.h>
+#include <vtkImageData.h>
+#include <vtkPointData.h>
+#include <vtkDoubleArray.h>
+
+// NAMESPACE IMPORTS
+using namespace std;
+using namespace Eigen;
+
+typedef boost::iostreams::tee_device<std::ostream, std::ofstream> Tee;
+typedef boost::iostreams::stream<Tee> TeeStream;
+typedef std::numeric_limits< double > dbl;
+
+//definition of pi
+const double pi = M_PI;
+
+// PROTOTYPES
+int width;
+int height;
+int steps; // use n steps for angular resolution - radres
+double radres;
+
+template <typename T>
+T clip(const T& n, const T& lower, const T& upper); // template clip function
+
+//length and width of window
+int wSize = 701;
+
+// GENERIC FUNCTION DEFINITIONS
+
+// get current working directory to assign matrix.txt path
+std::string GetCurrentWorkingDir(void) {
+	char buff[FILENAME_MAX];
+	GetCurrentDir(buff, FILENAME_MAX);
+	std::string current_working_dir(buff);
+	return current_working_dir;
+}
+
+// CLASS DEFINITIONS //
+
+class neighborhood
+{
+	bool neighborR = true;
[…]

### 2019-06-10 C++ Projects/depr/GIGradientUbuntu(on-the-fly).cu

@@ -1,0 +1,1254 @@
+// INLINE DEFINES
+
+#ifdef WINDOWS
+#include <direct.h>
+#define GetCurrentDir _getcwd
+#else
+#include <unistd.h>
+#define GetCurrentDir getcwd
+#endif
+
+#define MAXBUFSIZE  ((int) 1e8)
+#define _USE_MATH_DEFINES
+
+
+// INCLUDES (IMPORTS)
+//#include <SFML/Graphics.hpp>
+#include <iostream>
+//#include "muParser.h"
+#include <sstream>
+#include <fstream>
+#include <vector>
+//#include <unistd.h>
+#include <math.h>
+#include <Eigen/Dense>
+#include <Eigen/Eigenvalues>
+//#include <array>
+#include <experimental/filesystem>
+#include <string>
+//#include <algorithm> 
+//#include <cctype>
+//#include <locale>
+//#include <functional>
+#include <numeric>
+#include <boost/iostreams/stream.hpp>
+#include <boost/iostreams/tee.hpp>
+#include <limits>
+#include <ctime>
+
+// VTK Includes
+//#include <vtkVersion.h>
+#include <vtkSmartPointer.h>
+#include <vtkXMLImageDataWriter.h>
+#include <vtkImageData.h>
+#include <vtkPointData.h>
+#include <vtkDoubleArray.h>
+
+//thrust
+#include <thrust/device_vector.h>
+#include <thrust/transform.h>
+#include <thrust/copy.h>
+#include <thrust/fill.h>
+#include <thrust/functional.h>
+
+// NAMESPACE IMPORTS
+using namespace std;
+using namespace Eigen;
+
+typedef boost::iostreams::tee_device<std::ostream, std::ofstream> Tee;
+typedef boost::iostreams::stream<Tee> TeeStream;
+typedef std::numeric_limits< double > dbl;
+
+double buffer[MAXBUFSIZE];
+//definition of pi
+const double pi = M_PI;
+
+// PROTOTYPES
+int width;
+int height;
+int steps; // use n steps for angular resolution - radres
+double radres;
+
+template <typename T>
+T clip(const T& n, const T& lower, const T& upper); // template clip function
+
+//length and width of window
+int wSize = 701;
+
+// GENERIC FUNCTION DEFINITIONS
+
+// THRUST's parallelized GPU methods
+struct saxpy_functor
+{
+	const double a;
+
+	saxpy_functor(double _a) : a(_a) {}
+
[…]

### 2019-06-10 C++ Projects/final/sfml2DTensorVisV1FinalShort.cpp

@@ -1,0 +1,1342 @@
+// INLINE DEFINES
+
+#ifdef WINDOWS
+#include <direct.h>
+#define GetCurrentDir _getcwd
+#else
+#include <unistd.h>
+#define GetCurrentDir getcwd
+#endif
+
+#define MAXBUFSIZE  ((int) 1e8)
+#define _USE_MATH_DEFINES
+
+// INCLUDES (IMPORTS)
+#include <SFML/Graphics.hpp>
+#include <iostream>
+#include "muParser.h"
+#include <sstream>
+#include <fstream>
+#include <vector>
+#include <unistd.h>
+#include <math.h>
+#include <Eigen/Dense>
+#include <Eigen/Eigenvalues>
+//#include <array>
+#include <filesystem>
+#include <string>
+#include <algorithm> 
+//#include <cctype>
+//#include <locale>
+//#include <functional>
+#include <numeric>
+#include <boost/iostreams/stream.hpp>
+#include <boost/iostreams/tee.hpp>
+#include <limits>
+#include <ctime>
+
+// NAMESPACE IMPORTS
+using namespace std;
+using namespace Eigen;
+using namespace mu;
+
+typedef boost::iostreams::tee_device<std::ostream, std::ofstream> Tee;
+typedef boost::iostreams::stream<Tee> TeeStream;
+typedef std::numeric_limits< double > dbl;
+
+// PROTOTYPES
+std::string functionString = "2.1"; // Prototyping of functionString for strFunction (symbolic string arg parsed by muparser in cpp functional ptr rep)!
+int width;
+int height;
+int steps; // use n steps for angular resolution - radres
+double radres;
+double buff[MAXBUFSIZE];
+
+template <typename T>
+T clip(const T& n, const T& lower, const T& upper); // template clip function
+
+//length and width of window
+int wSize = 701;
+sf::Vector2i windowsize;
+
+//definition of pi
+const double pi = M_PI;
+
+// GENERIC FUNCTION DEFINITIONS
+
+// -> muparser custom clip function: template functions need to be static callback (call after) functions, which can be passed as arguments
+static value_type clipNeg(value_type v) { return clip(v, 0.0, 1.0); }
+static value_type deltaFunction(value_type v)
+{
+	double radres = 2 * pi / steps;
+	if (abs(v) < radres/2.0 || v == radres/2.0)
+		return steps;
+	else
[…]

### 2019-06-13 TF-13x13/gyre.py
update

@@ -46,8 +46,8 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101
-
+length = 13
+s = 10
 #deg = -45
 
 matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
@@ -62,10 +62,25 @@ with open('temp.txt', 'wb') as f:
             dy = np.pi * np.cos(np.pi * (i/(length-1))*2.0) * np.sin(np.pi * ((length-1-j)/(length-1))*2.0)
 
             rad = m.atan2(dy, dx)
-            if i==(length-1)/2 and j==(length-1)/2 or i==(length-1)/2/2 and j==(length-1)/2/2 or i==(length-1)/2+(length-1)/2/2 and j==(length-1)/2/2 or i==(length-1)/2/2 and j==(length-1)/2+(length-1)/2/2 or i==(length-1)/2+(length-1)/2/2 and j==(length-1)/2+(length-1)/2/2 :
+            if i==0:
+                if j==0 or j==(length-1)/2 or j==(length-1):
+                    scaled = np.matmul(scale(1, 1), identity)  # use isotropic scaling
+                else:
+                    scaled = np.matmul(scale(s, 1), identity)  # chronological transformation order: right->left
+            elif i == (length-1)/2:
+                if j==0 or j==(length-1)/2 or j==(length-1):
+                    scaled = np.matmul(scale(1, 1), identity)  # use isotropic scaling
+                else:
+                    scaled = np.matmul(scale(s, 1), identity)  # chronological transformation order: right->left
+            elif i == (length-1):
+                if j==0 or j==(length-1)/2 or j==(length-1):
+                    scaled = np.matmul(scale(1, 1), identity)  # use isotropic scaling
+                else:
+                    scaled = np.matmul(scale(s, 1), identity)  # chronological transformation order: right->left
+            elif i==(length-1)/2 and j==(length-1)/2 or i==(length-1)/2/2 and j==(length-1)/2/2 or i==(length-1)/2+(length-1)/2/2 and j==(length-1)/2/2 or i==(length-1)/2/2 and j==(length-1)/2+(length-1)/2/2 or i==(length-1)/2+(length-1)/2/2 and j==(length-1)/2+(length-1)/2/2 :
                 scaled = np.matmul(scale(1, 1), identity)  # use isotropic scaling
             else:
[…]

### 2019-06-17 C++ Projects/final/vtkReader.cpp
update

@@ -1,0 +1,154 @@
+// INLINE DEFINES
+
+#ifdef WINDOWS
+#include <direct.h>
+#define GetCurrentDir _getcwd
+#else
+#include <unistd.h>
+#define GetCurrentDir getcwd
+#endif
+
+#define MAXBUFSIZE  ((int) 1e8)
+#define _USE_MATH_DEFINES
+
+
+// INCLUDES (IMPORTS)
+//#include <SFML/Graphics.hpp>
+#include <iostream>
+//#include "muParser.h"
+#include <sstream>
+#include <fstream>
+#include <vector>
+//#include <unistd.h>
+#include <math.h>
+#include <Eigen/Dense>
+#include <Eigen/Eigenvalues>
+//#include <array>
+#include <experimental/filesystem>
+#include <string>
+//#include<sys/io.h>
+//#include <algorithm> 
+//#include <cctype>
+//#include <locale>
+//#include <functional>
+#include <numeric>
+#include <boost/iostreams/stream.hpp>
+#include <boost/iostreams/tee.hpp>
+#include <limits>
+#include <ctime>
+#include <omp.h>
+#include <chrono>
+typedef std::chrono::high_resolution_clock Clock;
+
+#include <vtkSmartPointer.h>
+#include <vtkProperty.h>
+//#include <vtkDataSetMapper.h>
+//#include <vtkImageActor.h>
+//#include <vtkImageViewer2.h>
+#include <vtkXMLImageDataReader.h>
+//#include <vtkRenderWindow.h>
+//#include <vtkRenderWindowInteractor.h>
+//#include <vtkRenderer.h>
+//#include <vtkXMLImageDataWriter.h>
+#include <vtkXMLPolyDataWriter.h>
+#include <vtkImageData.h>
+#include <vtkPointData.h>
+//#include <vtkDoubleArray.h>
+#include <vtkCutter.h>
+#include <vtkPlane.h>
+
+
+// NAMESPACE IMPORTS
+using namespace std;
+using namespace Eigen;
+
+typedef boost::iostreams::tee_device<std::ostream, std::ofstream> Tee;
+typedef boost::iostreams::stream<Tee> TeeStream;
+typedef std::numeric_limits< double > dbl;
+
+double buffer[MAXBUFSIZE];
+
+//definition of pi
+const double pi = M_PI;
+
+// PROTOTYPES
+int width;
+int height;
+int steps; // use n steps for angular resolution - radres
[…]

### 2019-06-19 C++ Projects/final/GIGradient(on-the-fly)-1D-vtkReader.cpp
update

@@ -761,6 +761,7 @@ void computeGlyphsFromVTK(std::vector<double>& glyphBuffer, std::vector<std::vec
 	std::vector<double>::iterator glyphEnd;
 	//std::advance(glyphEnd, steps);
 	// iterate through the matrixList/svdList (grid) and construct (scaled) ellipses in polar form (function) from the repsective singular values/vectors
+	double rMeanMax = 0.0;
 	for (int i = 0; i < matrixList.size(); i++)
 	{
 		glyphStart = glyphBuffer.begin() + i * steps;
@@ -821,17 +822,23 @@ void computeGlyphsFromVTK(std::vector<double>& glyphBuffer, std::vector<std::vec
 			}
 		}
 		double rMean = sum / steps; // compute rMean from cartesian (rectangular) energy-based integral as opposed to the polar integral relevant to the geometrical (triangular/circular) area
+		if (rMean > rMeanMax)
+			rMeanMax = rMean;
 		// write glyphParameters
 		glyphParameters.at(i).at(0) = 1.0 / rMean * sv1;
 		glyphParameters.at(i).at(1) = 1.0 / rMean * sv2;
 
 		// multiply respective cosine cone by valsum*=radres, because of energy normalization to cosine_sum (pre-computed in constructor)
-		std::transform(glyphStart, glyphEnd, glyphStart, std::bind(std::multiplies<double>(), std::placeholders::_1, 1.0 / rMean));
+		//std::transform(glyphStart, glyphEnd, glyphStart, std::bind(std::multiplies<double>(), std::placeholders::_1, 1.0 / rMean));
 		//glyphBuffer.at(i) = 1.0 / rMean * glyphBuffer.at(i);
 
 		//std::advance(glyphStart, steps);
 		//std::advance(glyphEnd, steps);
 	}
+	if (rMeanMax > 0)
+		std::transform(glyphBuffer.begin(), glyphBuffer.end(), glyphBuffer.begin(), std::bind(std::multiplies<double>(), std::placeholders::_1, 1.0 / rMeanMax));
+	else
+		cout << "NULL field encountered: no normalization possible" << endl;
 }
 
 int fast_mod(const int input, const int ceil) {

### 2019-06-21 C++ Projects/final/GIGradient(on-the-fly)-1D-vtkReader.cpp
add

@@ -1236,7 +1236,7 @@ int main(int argc, char* argv[])
 		dim = getVTKdim();
 		width = height = sqrt(dim);
 	}
-	ctrLimit = 2*width;
+	//ctrLimit = 2*width;
 	
 	cout << "width|height|steps: " << width << "|" << height << "|" << steps << endl;
 

### 2019-06-21 TF-13x13/rings.py

@@ -49,7 +49,7 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-width = 101
+width = 29
 length = width+50# use width+50 because of subsequent cropping for 13x13 fields --> hack to prevent circular function errors
 
 deg = -45

### 2019-06-23 TF-13x13/isotropic.py
ûpdate

@@ -51,7 +51,7 @@ def file_len(fname):
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
 zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
-length = 69
+length = 51
 
 deg = 0
 rad = deg*(np.pi/180.0)

### 2019-06-23 TF-13x13/rings.py

@@ -49,7 +49,7 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-width = 29
+width = 51
 length = width+50# use width+50 because of subsequent cropping for 13x13 fields --> hack to prevent circular function errors
 
 deg = -45

### 2019-06-23 C++ Projects/final/GIGradient(on-the-fly)-1D-vtkReader.cpp
update

@@ -660,7 +660,7 @@ void computeGlyphs(std::vector<double>& glyphBuffer, std::vector<std::vector<boo
 	}
 }
 
-int getVTKdim()
+int getVTKdim(int& width, int& height)
 {
 	// set file name "brain.vti"
 	std::string inputFilename = "brain.vtp";
@@ -668,8 +668,7 @@ int getVTKdim()
 	vtkSmartPointer<vtkXMLPolyDataReader> reader = vtkSmartPointer<vtkXMLPolyDataReader>::New();
 	reader->SetFileName(inputFilename.c_str());
 	reader->Update();
-	cout << "size: " << reader->GetOutput()->GetNumberOfPolys() << endl;
-
+	
 	// Create a plane to cut,here it cuts in the XY direction (xz normal=(1,0,0);XY =(0,0,1),YZ =(0,1,0)
 	//vtkSmartPointer<vtkPlane> plane = vtkSmartPointer<vtkPlane>::New();
 	//plane->SetOrigin(0, 0, slice);
@@ -680,14 +679,15 @@ int getVTKdim()
 	//cutter->SetCutFunction(plane);
 	//cutter->SetInputConnection(reader->GetOutputPort());
 	//cutter->Update();
-
+	
 	// get polyData of SLICE
 	vtkSmartPointer<vtkPolyData> polyData = reader->GetOutput(); //vtkSmartPointer<vtkImageData>::New();
-
+	
 	// create tensor array of slice and crop down to 2x2 matrices
 	vtkDataArray* tensors = polyData->GetPointData()->GetArray("tensors");// cutter->get()->GetScalars();
+	cout << "size: " << tensors->GetNumberOfComponents() << endl;
 	cout << "array size: " << tensors->GetSize() / 9 << endl;
-	
+
 	return tensors->GetSize() / 9;
 }
 void computeGlyphsFromVTK(std::vector<double>& glyphBuffer, std::vector<std::vector<bool>>& signMap, std::vector<std::vector<double>>& glyphParameters)
@@ -699,10 +699,9 @@ void computeGlyphsFromVTK(std::vector<double>& glyphBuffer, std::vector<std::vec
 	std::string inputFilename = "brain.vtp";
 
 	// Read the file
-	vtkSmartPointer<vtkXMLPolyDataReader> reader =	vtkSmartPointer<vtkXMLPolyDataReader>::New();
[…]

### 2019-06-23 TF-13x13/Y.py

@@ -1,98 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    if det != 0:
-        c = 1/np.sqrt(det)#*np.pi)
-    else:
-        c = 0
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
-length = 13
-
-deg = 90
-rad = deg*(np.pi/180.0)
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
-            matrixArray[j][i] = normalize(identity) # initialization w. zeros
[…]

### 2019-06-23 TF-13x13/cone.py

@@ -1,93 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    c = 1/np.sqrt(det)#*np.pi)
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 13
-
-deg = -45
-rad = deg*(np.pi/180.0)
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
-            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
-            if i==j: # if diagonal hit..
-                scaled = np.matmul(scale(4-i*(3/(length-1)),1),identity) # chronological transformation order: right->left
-                rotated = rotate(scaled, rad)
[…]

### 2019-06-23 TF-13x13/corner(base).py

@@ -1,91 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    c = 1/np.sqrt(det)#*np.pi)
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 117
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
-            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
-            rad = m.atan2(j+1,i+1) # use vector atan2 to determine angle and quadrant of vector [x|y]
-            if i==0 and j==0:
[…]

### 2019-06-23 TF-13x13/cross-X.py

@@ -1,105 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    if det != 0:
-        c = 1/np.sqrt(det)#*np.pi)
-    else:
-        c = 0
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
-length = 29
-
-deg1 = -45
-rad1 = deg1*(np.pi/180.0)
-deg2 = 45
-rad2 = deg2*(np.pi/180.0)
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
[…]

### 2019-06-23 TF-13x13/diagonalDrain.py

@@ -1,101 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    c = 1/np.sqrt(det)#*np.pi)
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101
-
-#deg = -45
-
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
-            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
-
-            dx = ((length-1-j)/(length-1)) * np.cos((i/(length-1))*((length-1-j)/(length-1)))
-            dy = (i/(length-1)) * np.cos((i/(length-1))*((length-1-j)/(length-1)))
-
[…]

### 2019-06-23 TF-13x13/drain.py

@@ -1,91 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(matrix): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return np.matmul(mirror, matrix)
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return np.matmul(mirror, matrix)
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    c = 1/np.sqrt(det)#*np.pi)
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
-            y = length - 1 - ((length-1)/2-j) #j # use inverted y coordinates (mirroring at x-axis in array reference frame)
-            x = (length - 1 - j)/(length-1)*round(i-(length-1)/2) # --> center around midpoint
[…]

### 2019-06-23 TF-13x13/drain_alt.py

@@ -1,91 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(matrix): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return np.matmul(mirror, matrix)
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return np.matmul(mirror, matrix)
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    c = 1/np.sqrt(det)#*np.pi)
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
-            y = j #length - 1 - ((length-1)/2-j) #j # use inverted y coordinates (mirroring at x-axis in array reference frame)
[…]

### 2019-06-23 TF-13x13/drain_old.py

@@ -1,90 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    c = 1/np.sqrt(det)#*np.pi)
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 117
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
-            y = length - 1 - round(j-(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
-            x = round(i-(length-1)/2) # --> center around midpoint
-            matrixArray[j][i] = normalize(np.array([[1, 0], [0, 1]])) # initialization w. normed identity
[…]

### 2019-06-23 TF-13x13/inverse.py

@@ -1,90 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    c = 1/np.sqrt(det)#*np.pi)
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
-            y = round(j-(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
-            x = round(i-(length-1)/2) # --> center around midpoint
-            matrixArray[j][i] = normalize(np.array([[1, 0], [0, 1]])) # initialization w. normed identity
[…]

### 2019-06-23 TF-13x13/isotropic.py

@@ -1,97 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    if det != 0:
-        c = 1/np.sqrt(det)#*np.pi)
-    else:
-        c = 0
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
-length = 51
-
-deg = 0
-rad = deg*(np.pi/180.0)
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
[…]

### 2019-06-23 TF-13x13/rings.py

@@ -1,142 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    if det == 0:
-        c = 1.0
-    else:
-        c = 1/np.sqrt(det)#*np.pi)
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-width = 51
-length = width+50# use width+50 because of subsequent cropping for 13x13 fields --> hack to prevent circular function errors
-
-deg = -45
-rad = deg*(np.pi/180.0)
-
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations (polar form)
-radrange = 2*m.pi
-radres = m.pi/180 # 1deg/step
-steps = radrange/radres
-radarr = [float(i)*radres for i in range(round(steps))]
[…]

### 2019-06-23 TF-13x13/spiral.py

@@ -1,151 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    c = 1/np.sqrt(det)#*np.pi)
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-# create identity
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101+50
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations (polar form)
-radrange = 260*m.pi # .. rad range WARNING: DO NOT TOUCH - experimentally derived parameter!
-radres = m.pi/180 # 1deg/step
-steps = radrange/radres # of steps :)
-radarr = [float(i)*radres for i in range(round(steps))] # create array w. steps in rad
-curDelta = np.ones((length,length)) # create current delta (error) array initialized w. delta=1
-
-for j in range(length): # rows
-    for i in range(length): # cols
[…]

### 2019-06-23 TF-13x13/star(src-sink).py

@@ -1,96 +1,0 @@
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    if det != 0:
-        c = 1/np.sqrt(det)#*np.pi)
-    else:
-        c = 0
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
-    return i + 1
-
-identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 51
-matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
-
-# generate normalized tensors from linear transformations
-with open('temp.txt', 'wb') as f:
-    for j in range(length): # rows
-        for i in range(length): # cols
-            y = round(length - 1 - j -(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
-            x = round(i-(length-1)/2) # --> center around midpoint
-            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
[…]

### 2019-06-23 TensorFields/Y.py

@@ -1,0 +1,98 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 13
+
+deg = 90
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. zeros
[…]

### 2019-06-23 TensorFields/corner(base).py

@@ -1,0 +1,91 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 117
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
+            rad = m.atan2(j+1,i+1) # use vector atan2 to determine angle and quadrant of vector [x|y]
+            if i==0 and j==0:
[…]

### 2019-06-23 TensorFields/diagonalDrain.py

@@ -1,0 +1,101 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 101
+
+#deg = -45
+
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
+
+            dx = ((length-1-j)/(length-1)) * np.cos((i/(length-1))*((length-1-j)/(length-1)))
+            dy = (i/(length-1)) * np.cos((i/(length-1))*((length-1-j)/(length-1)))
+
[…]

### 2019-06-23 TensorFields/drain.py

@@ -1,0 +1,91 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(matrix): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return np.matmul(mirror, matrix)
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return np.matmul(mirror, matrix)
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 101
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            y = length - 1 - ((length-1)/2-j) #j # use inverted y coordinates (mirroring at x-axis in array reference frame)
+            x = (length - 1 - j)/(length-1)*round(i-(length-1)/2) # --> center around midpoint
[…]

### 2019-06-23 TensorFields/drain_alt.py

@@ -1,0 +1,91 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(matrix): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return np.matmul(mirror, matrix)
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return np.matmul(mirror, matrix)
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 101
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            y = j #length - 1 - ((length-1)/2-j) #j # use inverted y coordinates (mirroring at x-axis in array reference frame)
[…]

### 2019-06-23 TensorFields/drain_old.py

@@ -1,0 +1,90 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 117
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            y = length - 1 - round(j-(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
+            x = round(i-(length-1)/2) # --> center around midpoint
+            matrixArray[j][i] = normalize(np.array([[1, 0], [0, 1]])) # initialization w. normed identity
[…]

### 2019-06-23 TensorFields/gyre.py

@@ -1,0 +1,112 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 101
+s = 4
+#deg = -45
+
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            matrixArray[j][i] = normalize(identity) # initialization w. normed identity
+
[…]

### 2019-06-23 TensorFields/inverse.py

@@ -1,0 +1,90 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 29
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            y = round(j-(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
+            x = round(i-(length-1)/2) # --> center around midpoint
+            matrixArray[j][i] = normalize(np.array([[1, 0], [0, 1]])) # initialization w. normed identity
[…]

### 2019-06-23 TensorFields/isotropic.py

@@ -1,0 +1,97 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 51
+
+deg = 0
+rad = deg*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
[…]

### 2019-06-23 TensorFields/spiral.py

@@ -1,0 +1,151 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+# create identity
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 101+50
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations (polar form)
+radrange = 260*m.pi # .. rad range WARNING: DO NOT TOUCH - experimentally derived parameter!
+radres = m.pi/180 # 1deg/step
+steps = radrange/radres # of steps :)
+radarr = [float(i)*radres for i in range(round(steps))] # create array w. steps in rad
+curDelta = np.ones((length,length)) # create current delta (error) array initialized w. delta=1
+
+for j in range(length): # rows
+    for i in range(length): # cols
[…]

### 2019-06-24 TensorFields/absorb.py
update tests

@@ -1,0 +1,100 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 101
+
+deg1 = 90
+rad1 = deg1*(np.pi/180.0)
+deg2 = 45
+rad2 = deg2*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
[…]

### 2019-06-24 TensorFields/bow.py

@@ -1,0 +1,90 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    c = 1/np.sqrt(det)#*np.pi)
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+length = 101
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
+            y = round((length-1)*m.cos(2*i/(length-1))-(length-1)/2) # use inverted y coordinates (mirroring at x-axis in array reference frame)
+            x = round(j - (length-1)/2) # --> center around midpoint
+            matrixArray[j][i] = normalize(np.array([[1, 0], [0, 1]])) # initialization w. normed identity
[…]

### 2019-06-24 TensorFields/step.py

@@ -1,0 +1,100 @@
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
+    return i + 1
+
+identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
+zero = np.array([[0, 0],[0, 0]]) ## row-major ordering: row-by-row
+length = 101
+
+deg1 = 90
+rad1 = deg1*(np.pi/180.0)
+deg2 = 45
+rad2 = deg2*(np.pi/180.0)
+matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
+
+
+# generate normalized tensors from linear transformations
+with open('temp.txt', 'wb') as f:
+    for j in range(length): # rows
+        for i in range(length): # cols
[…]

### 2019-06-25 TensorFields/step.py
add

@@ -67,7 +67,7 @@ with open('temp.txt', 'wb') as f:
             rotated = normalize(identity) # initialization w. zeros
             iInverse = length-1-i  # use inverted x coordinates (mirroring at y-axis in array reference frame)
             if j >= round((length-1)/2):
-                scaled = np.matmul(scale(4, 1), identity)  # use isotropic scaling (1,1) - src; (0,0) - sink
+                scaled = np.matmul(scale(10, 1), identity)  # use isotropic scaling (1,1) - src; (0,0) - sink
             else:
                 scaled = np.matmul(scale(1, 1), identity)  # chronological transformation order: right->left
 

### 2019-06-25 thesis-template/dummy.tex
add

\documentclass{scrartcl}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{csquotes}
% Include Graphic-files:
\usepackage{graphicx}
\usepackage{caption}

\newcommand{\source}[1]{\vspace{-3pt} \caption*{ Source: {#1}} }

\begin{document}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_drain_alt.png}
    \label{grada)}
    \caption*{a) LTG($90^\circ$) }
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/drain_alt_gradient.png}
    \label{gradb)}
    \caption*{b) gradient for a)}
  \end{minipage}
  \caption{``drain'' test field}
  \label{grad)}
\end{figure}

\end{document}

### 2019-06-27 C++ Projects/final/GIGradient(on-the-fly)-1D-vtkReader.cpp
update

@@ -682,12 +682,17 @@ int getVTKdim(int& width, int& height)
 	
 	// get polyData of SLICE
 	vtkSmartPointer<vtkPolyData> polyData = reader->GetOutput(); //vtkSmartPointer<vtkImageData>::New();
+	 //= imgData->GetDimensions();
 	
 	// create tensor array of slice and crop down to 2x2 matrices
-	vtkDataArray* tensors = polyData->GetPointData()->GetArray("tensors");// cutter->get()->GetScalars();
+	vtkDataArray* tensors = polyData->GetPointData()->GetArray("nrrd70723");// cutter->get()->GetScalars();
+	
 	cout << "size: " << tensors->GetNumberOfComponents() << endl;
 	cout << "array size: " << tensors->GetSize() / 9 << endl;
-
+	//vtkImageData* imgData = vtkImageData::SafeDownCast(polyData-> GetPointData()->getD);
+	//double* dimL = polyData->GetBounds();
+	//cout << "bounds: " << dimL[1] << endl;
+	//cout << "dim1: " << dim[1] << endl;
 	return tensors->GetSize() / 9;
 }
 void computeGlyphsFromVTK(std::vector<double>& glyphBuffer, std::vector<std::vector<bool>>& signMap, std::vector<std::vector<double>>& glyphParameters)
@@ -718,12 +723,12 @@ void computeGlyphsFromVTK(std::vector<double>& glyphBuffer, std::vector<std::vec
 	vtkSmartPointer<vtkPolyData> polyData = reader->GetOutput(); //vtkSmartPointer<vtkImageData>::New();
 
 	// create tensor array of slice and crop down to 2x2 matrices
-	vtkSmartPointer < vtkDataArray> tensors = vtkDataArray::SafeDownCast(polyData->GetPointData()->GetArray("tensors"));// cutter->get()->GetScalars();
+	vtkSmartPointer < vtkDataArray> tensors = vtkDataArray::SafeDownCast(polyData->GetPointData()->GetArray("nrrd70723"));// cutter->get()->GetScalars();
 	cout << "size: " << tensors->GetNumberOfComponents() << endl;
 	cout << "array size: " << tensors->GetSize() / 9 << endl;
 	cout << "array width: " << sqrt(tensors->GetSize() / 9) << endl;
 	double tensor[9];
-	int dim = 128 * 128;
+	int dim = width*height;
[…]

### 2019-07-02 C++ Projects/final/sfml2DTensorVisFTLE.cpp
update

@@ -1140,47 +1140,46 @@ int main(int argc, char* argv[])
 	computeGlyphs(glyphBuffer, signMap, glyphParameters);
 
 	// POLAR GRAPHER START //
-		//vector to store functions
 
-		// set options
-	int Alpha = 255; // set opacity ("Deckung")
-	int drawSpeed = 0; // set instant drawing
-	int lineWidth = 0; // set line width = 1px
-	double thetaMax = 2 * pi;
-	double thetaInc = (2 * pi) / steps; // set theta increment
-	double rotSpeed = 0.0; // set rotation speed
-	sf::Color red = sf::Color(255, 0, 0);
-	sf::Color green = sf::Color(0, 255, 0, Alpha);
+	// set options
+	//int Alpha = 255; // set opacity ("Deckung")
+	//int drawSpeed = 0; // set instant drawing
+	//int lineWidth = 0; // set line width = 1px
+	//double thetaMax = 2 * pi;
+	//double thetaInc = (2 * pi) / steps; // set theta increment
+	//double rotSpeed = 0.0; // set rotation speed
+	//sf::Color red = sf::Color(255, 0, 0);
+	//sf::Color green = sf::Color(0, 255, 0, Alpha);
 
-	polarPlot glyphs(glyphBuffer, drawSpeed, lineWidth, steps, thetaMax, radres, rotSpeed, green);
+	//polarPlot glyphs(glyphBuffer, drawSpeed, lineWidth, steps, thetaMax, radres, rotSpeed, green);
 
-	//declare renderwindow
-	sf::RenderWindow window;
-	if (fullscreen == false)
-		window.create(sf::VideoMode(wSize, wSize, 32), "Polar Grapher", sf::Style::Default);
-	else
-		window.create(sf::VideoMode::getDesktopMode(), "Polar Grapher", sf::Style::Fullscreen);
-	window.setFramerateLimit(60);
-	windowsize.x = window.getSize().x;
-	windowsize.y = window.getSize().y;
+	////declare renderwindow
+	//sf::RenderWindow window;
+	//if (fullscreen == false)
+	//	window.create(sf::VideoMode(wSize, wSize, 32), "Polar Grapher", sf::Style::Default);
+	//else
+	//	window.create(sf::VideoMode::getDesktopMode(), "Polar Grapher", sf::Style::Fullscreen);
+	//window.setFramerateLimit(60);
+	//windowsize.x = window.getSize().x;
+	//windowsize.y = window.getSize().y;
 
-	//initialize function graphics
[…]

### 2019-07-03 thesis-template/dummy.tex
update

@@
 
+\chapter{Fundamentals}
+\label{chap:grundlagen}
+
+In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and we will capture tensor fields, polar coordinates, PCA, FTLE, GT datasets and related work in that sense.
+
+%Die ersten paar Abschnitte in diesem Kapitel führen in die Grundlagen
+%zur Arbeit ein. Das können beispielsweise Grundlagen zu Netzwerken
+%oder zur Informationsextraktion sein.
+
+\section{Tensor Fields}
+\label{sec:tensors}
+Now what is a tensor? A tensor is a $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
+\begin{table}[!b]
+\centering
+\caption{Tensor Shapes}
+ \begin{tabular}{c|c|c|c|c|c|c}
+\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
+\hline 
+\textbf{shape} & scalar & vector & matrix & ``3D matrix'' & ... & $\mathop{o}$D matrix\\ 
+\end{tabular}
+\label{tensor}
+ \end{table} 
+In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ in space (the dimension of the row and column vectors), which is equal to the rank of the tensor for full rank (in other words the number of column elements n for a $n{\times}n$-matrix).
+A tensor is considered non-degenerate if all eigenvalues $\lambda_i>\lambda_{i+1}$ manifesting in significant anisotropy.
+Tensors follow certain transformation rules which are defined for covariant/contravariant tensors, which indicate the change characteristics under a basis change.
+Contravariant tensors do not change under a change of basis, whereas covariant tensors are not invariant w.r.t.
+basis changes.
+Tensors are most commonly obtained and defined from mathematical transformation equations.
+Hence, any $oD$ number array could be a tensor, but the definition holds only if the transformation rules apply to these $oD$ number arrays (in a sense of mathematical/formal interpretation).
+This is mostly true for component indexed $oD$ number arrays following matrix multiplication and transformation rules found in physics and math manifesting in calculation rules like matrix multiplication.
+Thus, a tensor in general is a $o$D number array arranged in meaningful (transformation) order, to put it simple.
+In fact, tensors themselves are interpreted as transformations which reflect an incoming direction vector (measuring, e.g., the stress in that direction) to an outgoing (stress) vector.
+To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities are not simple numbers or vectors.
+They occur in form of a tensors, and thus as entities or quantities in physics.
+That is, at each point (typically a Euclidean space or manifold) in space (indexed by the grid) there is a whole distribution of directions, e.g., stresses (elastic and viscous), which needs to be characterized and represented by a single tensor each.
+Collectively, defined at each point in space, these tensors form a tensor field.
+\paragraph{stress tensor}
+\begin{figure}[!t]
+  \centering
+ {
+    \includegraphics[width=0.6\textwidth]
+    {img/stress_tensor.png}
+  }
+  \caption{Cauchy stress Tensor}
+  \captionsetup{font={footnotesize,bf,it}}
+  \caption*{src: \url{https://en.wikipedia.org/wiki/Cauchy_stress_tensor\#/media/File:Components_stress_tensor_cartesian.svg}}
+  \label{cauchy}
+\end{figure}
+
+The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig.
+\ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$, arranged in row-major order.
+It appears in similar form as viscous stress tensor in fluid mechanics:
+\begin{align*}
+\boldsymbol{\sigma}
+ = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
+\boldsymbol{T}^{(\mathbf{e}_2)} \\
+\boldsymbol{T}^{(\mathbf{e}_3)} \\
+\end{matrix}}\right] =
+\left[{\begin{matrix}
+\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
+\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
+\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
+\end{matrix}}\right]  = \left[{\begin{matrix}
+\sigma _x & \tau _{xy} & \tau _{xz} \\
+\tau _{yx} & \sigma _y & \tau _{yz} \\
+\tau _{zx} & \tau _{zy} & \sigma _z \\
+\end{matrix}}\right] =
+\begin{bmatrix}
+\sigma _{11} & \sigma _{12} & \sigma _{13} \\
+\sigma _{21} & \sigma _{22} & \sigma _{23} \\
+\sigma _{31} & \sigma _{32} & \sigma _{33} 
+\end{bmatrix}.
+\end{align*}
+These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in plane normal directions ${x,y,z}$ ($3^2=9$ individual numbers).
+Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space.
+The stress tensor itself can be interpreted as a transformation, which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$.
+The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent (kind of a Euclidean distance of the stress components).
+The tensor itself is a complicated representation and needs to be projected onto its principal axes to lower its dimensionality, as it is the case for principal component analysis.
+It is possible to find $2$ principal stress directions in $2D$ through PCA (see sect.
+\ref{sec:pca} below), for which there are no shear stresses existent in equilibrium.
+These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour.
+The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''.
+So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig.
+\ref{hat} in sect.
+\ref{sec:pca}.
+Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle.
+It is therefore somehow representing the distortion of an absolutely isotropic object (unit circle) by whatever the underlying affine tensor transformation effects (e.g.
+stretching, shearing, mirroring, scaling, rotating).
+Let us consider it as some lower dimensional, more intuitive representation of the tensor which we were left with in the first place.
+To put it all together, a tensor is a mathematical transformation which is encoded inside of an abstract $oD$ number array, which comes in need, whenever we have a whole distribution of directions manifesting itself inside of a tensor (matrix) transformation.
+
+\section{Polar Coordinates}
+\label{sec:polar}
+\begin{figure}[!t]
+  \centering
+ {
+    \includegraphics[width=0.8\textwidth]
+    {img/polarkoordinaten.png}
+  }
+  \caption{polar space}
+  \captionsetup{font={footnotesize,bf,it}}
+   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:Ebene_polarkoordinaten.PNG}}
+  \label{polar}
+\end{figure}\vskip 3pt
+Mathematically speaking, a polar space is a system, which maps each angle $\omega$ a magnitude $r(\omega)$ through a symbolic function.
+The conversion formula from polar (given a radius $r$ and angle $\omega$) to Cartesian coordinates is depicted in Fig.
+\ref{polar} and is given by:
+\begin{align*}
+	x &= r \cos \omega, \\
+ 	y &= r \sin \omega.
+\end{align*}
+Note, that we switched angle $\phi$ to $\omega$ for reasonable discrimination to the radiant intensity. The inverse direction (given coordinates $x,y$) is a bit more complicated:
+\begin{align*}
+	r &= \sqrt{x^2 + y^2} \\
+\omega &= \operatorname{atan2}(y, x)
+\end{align*}
+The polar domain is a circular axis version of the domain $[0..2\pi]$, with negative magnitudes in $r$ shifted by 180 degrees (point reflected) as this is the formal behaviour defined and required by the conversion formulas.
+
+\section{Principal Component Analysis}
+\label{sec:pca}
+Note, that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following.
+Also keep in mind, that we interpret a matrix or tensor as affine transformation.
+For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent.
+Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted.
+The anisotropy can be represented by, e.g., ellipsoid glyphs.
+
+\begin{figure}[!t]
+\hskip 22pt
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/HAT.png}
+   \caption{2D principal component ellipsoid\/ glyph (eigenframe)}
+   \label{hat}
+   \captionsetup{font={footnotesize,bf,it}}
+   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:HAT-ellipse0.svg}}
+  \end{minipage} \hskip 40pt
+  %\raggedright
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/glyphs.png}
+    \caption{3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
+    \label{3Dglyphs}
+    \captionsetup{font={footnotesize,bf,it}}
+   \caption*{src: \url{https://slideplayer.com/slide/5291764/}}
+  \end{minipage}
+\end{figure}
+
+Principal Component Analysis is an algorithm, which is able to capture the principal components (directions) in stochastic data.
+But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling.
+Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices.
+For (skew-) symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf.
+\cite{kieburg}).
+Note, that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases (cf.
+\cite{moler}).
+Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling (classically: $\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$).
+We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive).
+Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig.
+\ref{HAT} for any kind of transformation matrix.
+The singular value decomposition is defined as follows:
+\begin{align*}
+	\mathbf{A} = \mathbf{U} \boldsymbol{\Sigma} \mathbf{V}
+\end{align*}
+where
+\begin{itemize}
+	\item $\mathbf{U}$ is an $m × m$ unitary matrix
+	\item $\boldsymbol{\Sigma}$ is a diagonal matrix {$m × n$} with positive real numbers on the diagonal
+	\item $\mathbf{V}$ is an {$n × n$} unitary matrix, $\mathbf{V}^*$ is its conjugate transpose
+\end{itemize}
+
+In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues (\cite{mroz}) ordered decreasingly by absolute value (corresponding to singular value order).
+The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses.
+We do this decomposition analysis once for each matrix and then store the precomputed results in a grid.
+The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step, which makes it a basic and central concept within the scope of this work.
+Within this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition.
+This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor.
+We will then use global-illumination techniques on top to generate a light transport flow map, which permits us somehow to forward the intensities which are given as predefined initialization values in form of symbolic functions defined by the user.
+This kind of light transport propagation scheme will also allow us to compute an FTLE-like field on the tensor fields.
+
+\section{Finite-Time Lyapunov Exponent}
+\label{sec:ftle}
+The Finite-Time Lyapunov Exponent is a measure, developed to analyze time-dependent dynamical systems for their separation abilities concerning massless tracer particles.
+We imagine a flow field, i.e., a vector field which in turn represents the effective amount or concentration shifted into a particular direction at each point in space.
+It could induced by e.g.
+concentration or height gradients.
+This flow field is considered a time-dependent dynamical system, since it usually holds a time-component $t$ and thus maps each initial position an aged (altered) position: $\mathbf{\upvarphi}(\mathbf{x}) : \mathbf{x}(t_0) \longmapsto \mathbf{x}(t_0+T)$.
+We can now imagine placing two close by particles in distance $\delta$ into a flow field and run a time dependent simulation.
+Remark that this corresponds to the Lagrangian view in vector fields, which travels with the particles (instead of taking the position of a resting observer - Eulerian view):
+
+\begin{figure}[!t]
+  \centering
+ {
+    \includegraphics[width=0.6\textwidth]
+    {img/ftle_sep.pdf}
+  }
+  \caption{FTLE: separation of massless tracer particles}
+  \captionsetup{font={footnotesize,bf,it}}
+  \caption*{src: Skript Prof. Sadlo, Scientific Visualization SS2017, Heidelberg University}
+  \label{ftle_sep}
+\end{figure}
+
+We observe particle trajectories diverging w.r.t. the central axis ending up in distance $\Delta$. The FTLE measure for this single example (point in space) is then defined as:
+
+\begin{align*}
+	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
+\end{align*}
+
+Alternatively, an FTLE can also be computed from a flow map's gradient, which is more convenient to compute in general:
+\begin{align*}
+	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \lvert\lvert\nabla\mathbf{\upvarphi(x)}\rvert\rvert _2,
+\end{align*}
+whereas $\lvert\lvert A\rvert\rvert _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$.
+This will yield a map, which responds most where path lines diverge w.r.t.
+to a central axis.
+We consider an example which detects FTLE ridges inside of the cross-section of a double gyre.
+Note, that this example considers single trajectories diverging implicitly:
+\begin{figure}[!t]
+\hskip 22pt
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/diverge.png}
+   \caption{diverging streamlines for double gyre}
+   \label{diverge}
+   \captionsetup{font={footnotesize,bf,it}}
+   \caption*{src: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dg_3traj.png}}
+  \end{minipage} \hskip 40pt
+  %\raggedright
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/ftle_gyre.png}
+    \caption{FTLE field for double gyre}
+    \label{ftle_gyre}
+    \captionsetup{font={footnotesize,bf,it}}
+   \caption*{src: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dle_incT.png}}
+  \end{minipage}
+\end{figure}
+Fig \ref{diverge} and \ref{ftle_gyre} depict a scenario of diverging streamlines for a double gyre snapshot in time (yielding a time-independent dynamical system).
+Fig.
+\ref{ftle_gyre} features a high ridge on the margins (borders) and in particular at the center of the domain.
+This region of interest (ROI) is considered to indicate the maximum relative local divergence of tensor field lines, by the approach.
+Ridges also represent Lagrangian Coherent Structures, which are found in nature in vortices of correlating flow behavior.
+We could imagine to introduce this process for a whole distribution of trajectories, which would implement an FTLE for uncertain vector fields in a sense, that whole propability distributions of trajectories could be heeded.
+An overview about existing visualization techniques for uncertain data is given by Griete \& Schuhmann \cite{griete}.
+We require that the trajectories can cross each other unaffectedly, i.e., the superposition principle is supposed to be respected for the magnitudes.
+But behold, is this behavior not somehow familiar?
+If we elaborately think this through, we can come up with the idea that this is the behavior of light propagation in vacuum.
+Light paths are physically allowed to cross each other with no influence exerted on each other at all.
+Consequently, we choose to model the distributions of trajectories as light intensity profiles.
+That is, we interpret the polar profiles as a propability density function (PDF) for particle trajectories heading into this direction.
+To put it into context, our main contribution method LTG uses the FTLE as an inspiration or template to design a similar approach for tensor field visualization.
+The main difference to the template FTLE is that we follow eigenvectors instead of flow vectors and that we use whole distributions of trajectories, as a tensor field suggests whole distributions of directions, accordingly.
+
+\section{GT Datasets / Methods}
+\label{sec:gt}
+\paragraph{Glyphs and Tensor Field Lines}
+\begin{figure}[!t]
+  \centering
+ {
+    \includegraphics[width=0.4\textwidth]
+    {img/tensorfieldlines.png}
+  }
+  \caption{tensor field lines: bold lines; ellipsoids: ellipsoid glyphs}
+  \label{scheme}
+\end{figure}\vskip 3pt
+Glyphs are derived as a basis for the weighting profiles of the tensors in the following propagation scheme.
+They were chosen because it is an easy-to-use (understand) concept, which can also be interpreted nicely.
+In fact, these representations are used for the principal component analysis of direction vectors for the tensor field line extrapolation (integration) and also as transmission profiles for the light propagation scheme in a subsequent step.
+We can again emphasize glyphs as a crucial, central concept within this work.
+Tensor field lines (the type of tensor streamline analog that does not introduce artificial intertia) were implemented for ground truth (GT) data generation purposes.
+For example, when designing a FTLE-like field computation technique, we will need to know when tensor field lines diverge/converge.
+
+\paragraph{Test Data (Generation)}
+We used python's numpy library to generate predefined, synthetic test cases and caught two real data example from \url{TensorVis.org} and \url{https://www.cs.purdue.edu/homes/cs530/projects/project5.html}, which needed to be pre-processed through slicing and subsequent cropping to obtain $2D$ samples ($2\times2$) matrices).
+These test fields can be grouped into symmetric and asymmetric ones:\\
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+    \includegraphics[height=\textwidth]{img/corner-TFL.png}
+    \caption{``corner'' test field}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+    \includegraphics[height=\textwidth]{img/spiral-TFL.png}
+    \caption{``spiral'' test field}
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+    \includegraphics[height=\textwidth]{img/drain_old.png}
+    \caption{``V'' test field    }
+    \label{b)}
+  \end{minipage}
+%\caption{a) , b) , c) }
+\label{test-fields}
+\end{figure}
+
+\textbf{Symmteric Tensor Fields}
+We also aim to use a real data example of the $3D$ DTI-MRI diffusion tensor of the brain for a male subject at the age of 27 and another one featuring a human heart.
+As we visualize 2D symmetric second order tensor fields, we need to slice the dataset and crop the $2{\times}2$ matrices omitting the $z$-component of the original $3{\times}3$ matrices.
+Fig.
+\ref{brainglyphs} depicts the glyph representation of the test dataset for slice index $8$:
 \begin{figure}[!t]
@@
   \begin{minipage}{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_drain_alt.png}
-    \label{grada)}
-    \caption*{a) LTG($90^\circ$) }
+    \includegraphics[width=\textwidth]{img/brain_cyl_glyphs.png}
+    \label{a)}
+    \caption*{cylinder glyphs}
   \end{minipage}
   \begin{minipage}{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/drain_alt_gradient.png}
-    \label{gradb)}
-    \caption*{b) gradient for a)}
+    \includegraphics[width=\textwidth]{img/brain_ellipsoid_glyphs.png}
+    \label{b)}
+    \caption*{ellipsoid glyphs}
   \end{minipage}
-  \caption{``drain'' test field}
-  \label{grad)}
+\caption{``brain''-test field}
+\label{brainglyphs}
 \end{figure}
 
+\textbf{Asymmetric Tensor Fields}
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/drain(alt).png}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/drain(alt)-TFL.png}
+    \label{b)}
+  \end{minipage}
+\caption{a) ``drain''-testfield, b) tensor field lines for a)}
+\label{drain}
+\end{figure}
+
+The ``drain''-Testfield, depicted in Fig.
+\ref{drain} is generated to have a kind of proof-of-concept (POC) for the FTLE-related approach.
+It poses raw, bare diverging tensor field lines without any other feature modulated on top.
+Thus, it can be expected to induce a high FTLE response, since it is an interpretable ground truth test and is considered to be a basic trigger/stimulus for the approach with no side effects introduced.
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/rings.png}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/tensorfieldlines.png}
+    \label{b)}
+  \end{minipage}
+\caption{a) ``rings'' test field, b) tensor field lines for a)}
+\label{rings}
+\end{figure}
+The ``rings'' test field, depicted in Fig. \ref{rings} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/spiral.png}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/spiral-TFL.png}
+    \label{b)}
+  \end{minipage}
+\caption{a) ``spiral''-test field, b) tensor field lines for a)}
+\label{spiral}
+\end{figure}
+The ``spiral'' test field, depicted in Fig. \ref{spiral} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/inverse.png}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/inverse-TFL.png}
+    \label{b)}
+  \end{minipage}
+\caption{a) ``inverse''-test field, b) tensor field lines for a)}
+\label{inverse}
+\end{figure}
+With the \enquote{inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/corner.png}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/corner-TFL.png}
+    \label{b)}
+  \end{minipage}
+\caption{a) ``corner''-test field, b) tensor field lines for a)}
+\label{corner}
+\end{figure}
+The test field ``corner'', depicted in Fig \ref{corner}, constitutes a sustained turn over the grid representing the redirection or forwarding along domain corners.
+To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{brain} and \enquote{heart} scans, which can be observed in Fig.
+\ref{real1}.
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/brain_org.png}
+    \label{a)}
+    \caption*{tensor magnitude}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/brain_cyl_glyphs.png}
+    \label{b)}
+    \caption*{cyclinder glyphs}
+  \end{minipage}
+\caption{Real data examples: \enquote{brain} dataset}
+\label{real1}
+\end{figure}
+Observing the \enquote{brain} dataset in Fig \ref{real1}, one may notice that it introduces a lot of variances to tensor magnitude and linear anisotropy.
+We choose this DTI-MRI scan as an example-par-excellence for real application data given in a reasonable resolution.
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \includegraphics[width=0.8\textwidth]{img/heart_org.png}
+    \label{a)}
+    \caption*{tensor magnitude}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+      \includegraphics[width=0.8\textwidth]{img/heart_cyl_glyphs.png}
+    \label{b)}
+    \caption*{cyclinder glyphs}
+  \end{minipage}
+\caption{Real data examples \enquote{heart} dataset}
+\label{real2}
+\end{figure}
+When we look at this DT-MRI scan of a human heart we can notice the elongated shape, which we consider to represent the aorta and a part of the rest of the heart.
+If we consider the glyph representation, we again infer, that we incorporate lots of anisotropic regions with varying tensor magnitudes here.
+
+\chapter{Related Work}
+\label{chap:related}
+
+There has been extensive recent work on symmetric tensors but in comparison little on asymmetric ones.
+Since we are not able to cover the whole scope in this work, we will focus on the most relevant works on tensor field visualization.
+We will talk about the publications from Hlawatsch et al.
+\cite{hlawatsch} and Zirr et al.
+\cite{zirr} in particular, since they constitute thematically close related works.
+But first, we will cover related works on global illumination methods, as these form a basis or prerequisite for our work.
+Please note, that they do not focus on tensor field visualization at all.
+Remember, that we will inventively employ global illumination methods for tensor field visualization as an innovative step and that we want to cover previous works on our base approach, as well.
+
+\paragraph{Global Illumination Methods}
+Since we use a specific category of global illumination methods as a basis, we decide to only discuss lattice based methods which work, corresponding to their name, on a grid.
+Remark, that Global Illumination methods classically focus on realistic indirect lighting.
+Thus, they aim on producing a most realistic image, what in our case turns out to be a most informative image in correspondence.
+The most early work in this field is propably the Discrete Ordinates Method (DOM) by Chandrasekhar \cite{Chandrasekhar}, which is used specifically for computing radiative transfer in participating media (volume rendering).
+The radiative transfer equation is discretized spatially and in the angular domain.
+The radiance distribution (PDF) is stored in a grid and light is distributed and exchanged between neighboring volume elements which reduces the overall effort to local and independent computations.
+Furthermore, there has been more recent work on describing the light propagation as a diffusion process within a simple photon transport model based on the Lattice-Boltzmann method \cite{geist}.
+DOMs were recently revisited by Fattal \cite{fattal}.
+This work adresses artifacts stemming from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}.
+These methods are considered to be too slow to \enquote{run at interactive framerates} by Dachsbacher et al.
+\cite{dachsbacher}, though \enquote{efficient and potentially highly parallel}.
+Our work is inspired by Lattice-Boltzmann methods (volume grid radiosity) and the lattice-based discrete ordinate methods concerning the light propagation scheme.
+
+\paragraph{Symmetric Tensor Field Visualization}
+Zheng \& and Pang \cite{pang&zheng} et al.
+proposed a texture-based visualization approach called Hyper LIC, which extends the concept of LIC to symmetric tensor fields by using an anisotropic 2D-filter kernel oriented along the major/minor eigenvectors.
+The concept of following the major/minor eigenvector along tensor field lines was invented by Vilanova et al.
+\cite{vilanova}.
+Tensorlines \cite{weinstein} introduce a kind of artificial inertia (running average), making them resistent to noise.
+The notion of tensor field topology (critical, degenerate points, separatrices) and the concept of hyperstreamlines was first introduced by Delmarcelle and Hesselink \cite{delmarcelle&hesselink}.
+The topology in $3D$ tensor fields was first analyzed by Zheng et al.
+\cite{zheng}.
+The placement of tensor field lines as hyperstreamlines for glyph packing was recently improved by Spencer et al.
+\cite{spencer}.
+A good overview of seeding strategies concerning glyphs is given by McLoughlin et al.
+\cite{mcloughlin}.
+Feng et al.
+\cite{feng} used Voronoi Tesselation for placing the glyphs.
+Superquadric glyphs have been proposed by Schultz and Kindlmann \cite{schultz&kindlmann} for general (non-positive definite) tensors.
+De Leeuw and van Wijk \cite{deleeuw&vanwijk} visualize the partial derivative gradient, the Jacobian of the tensor field, to sense the local properties of the field.
+A diffusion tensor can also be visualized by box \cite{makris}, ellipsoid \cite{pierpaoli}, composite \cite{westin2} or superquadric \cite{kindlmann} glyphs.
+There is also a set of scalar measures available in symmetric tensor field visualization such as: mean diffusivity, fractional anisotropy \cite{basser} and anisotropy coefficients \cite{westin}.
+These coefficients can also be visualized by direct volume rendering \cite{kindlmann2}.
+Hlawatsch et al.
+\cite{hlawatsch} obtain a coordinate (flow) map from tensor field lines following major/minor eigenvectors (fibre trajectories) using the maximum eigenvalue of the Cauchy-Green deformation tensor to compute the FTLE-field in homogeneous regions.
+In a similar manner, we aim to use light transport techniques to obtain a final light distribution (flow map) through successive light propagation cycles directed (in form of transmission profiles obtained from ellipsoid glyphs) by the tensor field.
+DT-MRI diffusion tensor imaging can help to reveal the functional \cite{denis} or topological \cite{wakana} structure of the brain and muscles by analysing celebral and muscle tissue tractography, which yields a tensor field representing the anisotropy characteristcs of the brain's white matter.
+An overview about DT-MRI visualization techniques is given by \cite{vilanova}.
+Tricoche et al.
+\cite{xavier} use invariant manifolds to extract topological features from tensor fields.
+Other feature extraction-based approaches extract general features as: edges \cite{hancock}, tensor shape\/ orientation \cite{gordon}, interfaces \cite{donnel}, creases \cite{tricochet}.
+\\
+\paragraph{Asymmetric Tensor Field Visualization}
+Zheng and Pang \cite{pang&zheng} proposed the concept of dual eigenvectors and Zhang et al.
+\cite{zhang} extended it to pseudo-eigenvectors and introduced the eigenvector/eigenvalue manifolds to visualize eigenvectors in the complex domain.
+Laramee et al.
+\cite{laramee} focused on the efficient implementation and visualization of these structures and provided an interactive visualization system for asymmetric tensors applicable in fluid and solid dynamics.
+The concept of tensor magnitude has been introduced by \cite{laramee} for means of physical interpretation.
+They also proposed an efficient glyph and hyperstreamline hybrid approach, which made dynamic interaction in real-time in $2D$ tensor fields feasible.
+Palacios et al.
+\cite{palacios} extracted isosurfaces of tensor magnitude, mode and isotropy index.
+
+Last, we will discuss our work in relation to other publications.
+While these previously mentioned works use classical approaches like tensor field lines and glyphs, we will also introduce the concept of Lagrangian coherent structures into our framework contribution to tensor field analysis.
+LCS have been defined as time-dependent analog of separatrices \cite{haller} which are concerned to be robust under noisy conditions \cite{haller2}, which we expect from our technique as well by implication.
+Hlawatsch et al.
+\cite{hlawatsch} obtain similar results (a FTLE-like field on tensor fields), but obtain them by sampling of fiber trajectories instead of global illumination light distributions, as it is the case for this work.
+Our work shares the same basic idea, deriving and computing a FTLE-like field for tensor field visualization generalizing FTLE from vectors to tensors, yet using a light propagation scheme and hence using whole distributions of trajectories incorporating polar profiles.
+Concerning the tensor field visualization method (LTG), our approach is based on FTLE, and principal component analysis (PCA).
+We will detect LCS with an FTLE-derived method considering differences in final light transport distributions through computing the gradient of the resulting flow maps.
+The FTLE-related approach named ``light transport gradient'', to be designed, also states a generalization of the light transport visualization method FTPD proposed by Zirr et al.
+\cite{zirr}, since it allows a similar measure to be computed through defining a geometrical scene and setting the transmission (defined by the tensor field) to an isotropic and constant value of $100\%$.
+In this operating mode, the approach operates as a light transport visualization technique neglecting any tensor field transmission profiles and is thus capable of displaying bounding separatrices of a light transport scenario perturbated by obstacles \cite{zirr}.
+%Typischerweise im letzten Abschnitt dieses Kapitels wird dann auf
+%verwandte Arbeiten eingegangen. Entsprechende Arbeiten sind geeignet
+%zu zitieren. Beispiel: Die wurde erstmalig in den Arbeiten von Spitz
+%und Gertz \cite{Spitz2016a} gezeigt \ldots Details dazu werden in
+%dem Buch von Newman zu Netzwerken \cite{Newman2010} erläutert \ldots.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Alternative: put content in separate files
+% Check the difference between including these files using \input{filename} and \include{filename} and see which one you like better
+%\chapter{Einleitung}\label{intro}
+%\input{introduction}
+%
+%\chapter{Voraussetzungen}\label{bg}
+%\input{background}
+
 \end{document}

### 2019-07-03 C++ Projects/final/sfml2DTensorVisFTLE.cpp

@@ -250,7 +250,7 @@ public:
 	//{
 	//	discretePlot = true;
 	//}
-	
+
 	//initialize muparser and graphics 
 	void init() {
 		p.DefineVar("theta", &t);
@@ -647,7 +647,7 @@ void parse_options(int argc, char* argv[]) {
 // mathematical functions
 double circleFunction(double x)
 {
-	double os = 1.0/sqrt(M_PI);
+	double os = 1.0 / sqrt(M_PI);
 	return os;
 }
 
@@ -670,13 +670,13 @@ double circleFunction(double x)
 //}
 
 
-double integral(double(*f)(double x), double a, double b, int r) 
+double integral(double(*f)(double x), double a, double b, int r)
 {
 	double step = (b - a) / r;  // width of each small rectangle
 	double area = 0.0;  // signed area
 	for (int i = 0; i < r; i++)
 		area += f(a + (i + 0.5) * step) * step; // sum up each small rectangle
-	
+
 	return area;
 }
 
@@ -684,10 +684,10 @@ double aIntegral(double(*f)(double x), double a, double b, int r, int n)
 {
 	double step = (b - a) / r;  // width of each small rectangle
 	double area = 0.0;  // signed area
-	
-	for (int i = 0; i < r; i++) 
+
+	for (int i = 0; i < r; i++)
 		area += f(a + (i + 0.5) * step)*cos(n*(a + (i + 0.5) * step)) * step; // sum up each small rectangle
-	
+
 	return area;
 }
 
@@ -697,13 +697,13 @@ double bIntegral(double(*f)(double x), double a, double b, int r, int n)
 	double area = 0.0;  // signed area
 	for (int i = 0; i < r; i++)
 		area += f(a + (i + 0.5) * step)*sin(n*(a + (i + 0.5) * step)) * step; // sum up each small rectangle
-	
+
 	return area;
 }
 
 std::vector<std::vector<double>> calcCoeff(double(*f)(double x)) // for 2Pi periodic functions
 {
-	double a0 = 1 /(2*M_PI) * integral(f, 0, 2*M_PI, 50);
+	double a0 = 1 / (2 * M_PI) * integral(f, 0, 2 * M_PI, 50);
 	cout << "a0: " << a0 << endl;
 	bool finish = false;
 	std::vector<double> an(21, 0.0);
[…]

### 2019-07-03 TensorFields/diagonalDrain.py

@@ -46,7 +46,7 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101
+length = 13
 
 #deg = -45
 
@@ -63,9 +63,9 @@ with open('temp.txt', 'wb') as f:
 
             rad = m.atan2(dy, dx)
             if i==0 and j==length-1:
-                rotated = rotate(scale(10,1), m.pi/4)
+                rotated = rotate(scale(4,1), m.pi/4)
             else:
-                rotated = rotate(scale(10,1), rad)
+                rotated = rotate(scale(4,1), rad)
             # if i==(length-1)/2 and j==(length-1)/2 or i==(length-1)/2/2 and j==(length-1)/2/2 or i==(length-1)/2+(length-1)/2/2 and j==(length-1)/2/2 or i==(length-1)/2/2 and j==(length-1)/2+(length-1)/2/2 or i==(length-1)/2+(length-1)/2/2 and j==(length-1)/2+(length-1)/2/2 :
             #     scaled = np.matmul(scale(1, 1), identity)  # use isotropic scaling
             # else:

### 2019-07-06 thesis-template/msc-thesis.tex
update

	
% This template was initially provided by Dulip Withanage.
% Modifications for the database systems research group
% were made by Conny Junghans,  Jannik Strötgen and Michael Gertz

\documentclass[%enabledeprecatedfontcommands,
     12pt,         % font size
     a4paper,      % paper format
     BCOR10mm,     % binding correction
     DIV14,        % stripe size for margin calculation
%     liststotoc,   % table listing in toc
%     bibtotoc,     % bibliography in toc
%     idxtotoc,     % index in toc
%     parskip       % paragraph skip instad of paragraph indent
     ]{scrreprt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% PACKAGES:

% fix bitmap fonts
\usepackage{lmodern}
\usepackage{tabularx}
\usepackage{booktabs}
% Use German :
\usepackage[ngerman,english]{babel}
\usepackage{amsmath}
\usepackage{upgreek}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
% Input and font encoding
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
% Index-generation
\usepackage{makeidx}
% Einbinden von URLs:
\usepackage{url}
% Special \LaTex symbols (e.g. \BibTeX):
%\usepackage{doc}
% Include Graphic-files:
\usepackage{graphicx}
\usepackage{csquotes}
% Include doc++ generated tex-files:
%\usepackage{docxx}
% Include PDF links
%\usepackage[pdftex, bookmarks=true]{hyperref}

% Fuer anderthalbzeiligen Textsatz
\usepackage{setspace}

\usepackage{caption}

% hyperrefs in the documents
\usepackage[bookmarks=true,colorlinks,pdfpagelabels,pdfstartview = FitH,bookmarksopen = true,bookmarksnumbered = true,linkcolor = black,plainpages = false,hypertexnames = false,citecolor = black,urlcolor=black]{hyperref} 
%\usepackage{hyperref}
\usepackage{letterspace}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% OTHER SETTINGS:

% Pagestyle:
\pagestyle{headings}

% Choose language
\newcommand{\setlang}[1]{\selectlanguage{#1}\nonfrenchspacing}
\newcommand*\mean[1]{\bar{#1}}
\newcommand{\source}[1]{\vspace{-3pt} \caption*{ Source: {#1}} }
\setlength{\emergencystretch}{3em}
\begin{document}

% TITLE:
\pagenumbering{roman} 
\begin{titlepage}

\vspace*{1cm}
\begin{center}
\vspace*{3cm}
\textbf{ 
\Large Heidelberg University\\
\smallskip
\Large Institute of Computer Science\\
\smallskip
\Large Visual Computing Group (VCG)\\
\smallskip
}

\vspace{3cm}

\textbf{\large Master-Arbeit} % Bachelor-Arbeit

\vspace{0.5\baselineskip}
{\huge
\textbf{Light Transport Techniques for Tensor Field Visualization}
}
\end{center}

\vfill

{\large
\begin{tabular}[l]{ll}
Sebastian Bek\\
Student ID: & 3481802\\
Supervisors: & Prof. Filip Sadlo, Dr. Susanne Kr?mker\\
Submission date: & 07.07.2019
\end{tabular}
}

\end{titlepage}

\onehalfspacing

\thispagestyle{empty}

\vspace*{100pt}
\noindent
\emph{dt.:}
Ich versichere, dass ich diese Master-Arbeit selbstst?ndig verfasst und nur die angegebenen
Quellen und Hilfsmittel verwendet habe und die Grunds?tze und
,Empfehlungen ``Verantwortung in der Wissenschaft'' der Universit?t Heidelberg beachtet wurden. 
\bigskip
\\
\emph{eng.:}
I hereby assure, that I drafted this thesis independently and only used the sources and materials labeled as references and that the conventions/principles and recommendations ``Verantwortung in der Wissenschaft'' of the Heidelberg University have been regarded. 
\vspace*{50pt}
\noindent

\underline{\phantom{mmmmmmmmmmmmmmmmmmmm}}

\medskip
\noindent 
Abgabedatum / Due Date: 07.07.2019
\newpage

% Add a brief summary of your topic and contributions (Zusammenfassung) in German *and* in English:
\chapter*{Zusammenfassung}
%\input{zusammenfassung}

Tensorfelder werden meistens in Verbindung mit mechanischen Spanunngsverteilungen in 2D/3D-Gittern gebracht (vgl. Cauchy-Stress Tensor), haben aber auch andere praktische Bedeutungen in der Physik.
Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren, die aber zus?tzlich auch verwendet werden kann um die Lichtausbreitung in einer topologischen Szene zu beschreiben. Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema f?r kartesische Gitter ab, dass die Prinzipien der Ausbreitungsd?mpfung und Energieerhaltung beachtet und Licht-verteilung/en f?r gegebene Lichtquellenrichtung/en und -positionen bis zur Konvergenz approximieren kann. Als unterliegendes Modell werden die Transmissionsprofile innerhalb diesem Gitter als Kristallfiberstruktur angenommen (vgl. Edelsteine: Katzenaugeneffekt beim Tigerauge). Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiema?en des unterliegenden Tensorfelds zu modellieren. Daf?r erstellen wir per Hauptachsentransformation (PCA) ein Eigensystem f?r jede Zelle des Tensorfelds und modellieren eine zugeh?rige Ellipsengleichung als (Transfer-/Wichtungs-) Transmissionsprofil. Folglich kann die resultierende Lichtverteilung als 2D Skalarfeld oder Polarplots f?r Richtungsinformation visualisiert/repr?sentiert werden. Wir messen Impulsantworten des Tensorfelds mit Delta-Pulsen an jeder Position und in jede Richtung um eine uniform gesamplete Map der globalen Lichtverteilungen zu erhalten. Diese Map wird als ein globales Energieflussfeld betrachtet. Wir lassen uns von der Technik FTLE aus dem Bereich Particle Tracing der Vector Field Visualization inspirieren, indem wir den Gradient des Flussfelds der resultierenden Lichtverteilungen analysieren und visualisieren. ?hnlich wie FTLE in Vektorfeldern Gr?te/K?mme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, anziehende/absto?ende/sattelnde Tensorfeldlinien die h?ufig Schl?sselstrukturen in der Natur repr?sentieren, in Tensorfeldern. Wir f?hren diese Gr??e als Global Illumination Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtflusses, zu visualisieren.
\newpage

\chapter*{Abstract}
%\input{abstract}
Many problems in science and engineering require
tensor field representations. Tensor fields, are most commonly associated with stress distributions (cf. Cauchy-stress tensor) in 2D/3D grids, but also have some other meanings in practice in physics. By means of global illumination techniques, we motivate a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in a topologically defined scene geometry. As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence. The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (, e.g., gemstones: tiger's eye's cat's eye effect). The consequent task is to model anisotropic fiber structures with the orientation and anisotropy measures of the underlying tensor field in the grid. For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer) function for the transmitted light profiles. Hence, the resulting light distribution can be visualized via 2D scalar field color coding or polar plots for directional information. On this basis, we measure impulse responses of the tensor field with Dirac-pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions. This map is considered as a global illumination energy flow field. We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field, formed by the resulting light distributions. Similar to how the FTLE detects ridges/LCS (Lagrangian Coherent Structures) in vector fields (flow fields), our approach captures these structures likewise, namely regions revealing attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields. We denote this entity as light transport gradient (LTG) which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global light transport in the grid, directed by imprints in crystelline fiber structures modeled by the tensor field's ellipsoid glyhps. We evaluate our approaches for plausibility and comparability within an extensive test campaign and demonstrate its performance on several examples of both synthetic and real datasets.
\newpage

% MAIN PART:
% Table of contents (Inhaltsverzeichnis)
\tableofcontents
\cleardoublepage
\pagenumbering{arabic}

% List of figures (Abbildungsverzeichnis):
%\listoffigures
% List of tables (Tabellenverzeichnis):
%\listoftables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Here, the actual content of your thesis begins
% You can either put all the text here or use individual files to store the chapters of your thesis.
% Below are templates for both alternatives.

\chapter{Introduction}\label{intro}

In this work, we will discuss the advantages of tensor field analysis and motivate a new method for tensor field visualization through a coherent light transport visualization. At first, we will shortly state the problem setting, objectives and give a quick overview about the structure of this work. The main contributions of this work include:

\paragraph{Contributions of this Work}
\begin{itemize}
	\item light transport model (propagation scheme) following basic but crucial physical principles for tensor field visualization using tensors as transmission profiles,
	\item visualization of the propagation results in $2D$ polar profiles and scalar fields,
	\item a FTLE-related approach called light transport gradient (LTG) for visualizing key structures (ROIs), namely LCS in $2D$ second-order tensor fields, and
	\item application of our approach to real data involving brain and heart datasets.
\end{itemize}
%Dieses Kapitel gibt einen ?berblick ?ber die Arbeit. Gerade der
%Abschnitt zur Motivation soll allgemein verst?ndlich geschrieben
%werden. Die Einleitung sollte auch wichtige Referenzen enthalten.

\section{Motivation}
\label{sec:motiv}
Tensor field visualization is gaining importance as a relevant tool for the analysis of fluid and solid mechanics, where tensor representations are crucially needed. This is always the case, when a directional distribution is needed for each single point in space. Tensors occur most commonly in medical, scientific and engineering applications, making them a frequently used mathematical entity in science. In general, a tensor is a $o$-D generalization of scalar, vector and matrix and appears, e.g., as Jacobian matrix (spatial gradient) in vector (flow) field visualization or as stress tensor in solid and fluid continuum mechanics. Basically, tensors describe compressions, tensions, rotations and volume changes in both solid and fluid material, as illustrated in Fig. \ref{ToS}. While most techniques focus on symmetric tensor field visualization (like glyphs and tensor field lines), some other recent works adress the problem of asymmetric tensor field visualization. The issue with asymmetric (non-symmetric) or therein included antisymmetric (skew-symmetric) tensors is, that they do not always yield real eigenvalues, which is needed to determine an eigenbasis to set up glyphs, hyperstreamlines or tensor field lines (classically). Nevertheless, there has been extensive work in this domain as well. Classically, the issue with asymmetric tensor fields is, that they do not yield real eigenvalues in general, which is needed to set up glyph representations. In this work, we will use the $\mathbf{U}$-matrix obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA principal ellipsoid representing the tensor transformation as suggested by Moler \cite{moler} to state another alternative of processing asymmetric tensors. 2D tensor field visualization is useful for cases where only two-dimensional data is available, such as satellite remote sensing, particle imagery velocimetry (PIV) and earthquake data \cite{laramee}. In their field of work, geologists and medical scientists are struggling with complicated number arrays which seem almost impossible to interpret in their raw, unprocessed form. Nevertheless, comprehensive tensor field visualization is aiding scientists in their everyday lifes understanding complex data loads, such as tensor fields. To cope with the arising problem of tensor field visualization, we can only try to invent new methods, continuously evaluating them for suitability for certain domains of science, to provide our techniques at best, while considering them for the right applications.\\

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/DIFFERENT_TYPES_OF_STRESS.png}
  }
  \caption{Different types of stress}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{https://upload.wikimedia.org/wikipedia/commons/e/eb/DIFFERENT_TYPES_OF_STRESS.png}}	
  \label{ToS}
\end{figure}

\section{Objectives}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.5\textwidth]
    {img/hyperstreamlines.png}
  }
  \caption{Hyperstreamlines in tensor field visualization}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{http://www2.cs.uh.edu/~chengu/Teaching/Fall2018/Lecs/Lec17.pdf}}	
  \label{streamlines}
\end{figure}
The main purpose of visualization in general, is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies, e.g., clustering, structuring or projection techniques. When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor - medical resonance imaging) \cite{hlawatsch}, when measuring water molecule diffusion characteristics within tissue or in simulated (or measured) stress tensor fields from solid and fluid mechanics. In these named domains of science the problem arises, that tensor fields can be generated (simulated) or measured with physical/mathematical background, but not be regarded intuitively. Necessity begets ingenuity as always, which lets visualization methods and systems emerge in this context. Concurrently, the problem setting of this work encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids, commonly consisting of second-order tensors (cf. Cauchy stress tensor). In general, these grids span a spatial domain with sample positions indicated by vertices or cell centers in the dual grid, depending on the mode of interpretation. Tensor fields feature a sampling grid, consistent of tensors, which is spaced uniformly in the commonly used regular grid. A Cartesian grid, which we will utilize in this context, is a grid normalized to a certain unit length, as a special case of a regular grid. That is, each cell is formed as a unit cube in relation to the normalizing unit length. We aim to use global illumination methods to model light transport, directed by tensor fields, within such type of grids, which we consider as light propagation volumes (LPVs) \cite{dachsbacher}. Congruently, the most basic objective of this work is to design a simple light propagation scheme following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation, for tensor field visualization. Of course, the propagation scheme will need some kind of stimulus which can be defined through symbolic functions (, e.g., $r(\omega)=1$) by the user which we sample (discretize) as a primary step. This way, the approach will operate as a light transport visualization technique for light distributions given in polar coordinates, provided the user set an initial profile into the grid. Furthermore, this propagation scheme should either be used to visualize the light transport paths, defined by the anisotropy characteristics of tensor fields and tensor magnitudes of a given tensor field in a $2D$ grid. Or, as a secondary purpose, be used to visualize the light transport in a defined, given geometrical scene in a $2D$ grid. However, the tensor field needs to be preprocessed somehow to generate a footprint map of transmission profiles defined by the tensor field defined by the tensor field's anisotropy characteristics, which we will handle through matrix decompositions. For this purpose, we will let the transmission profiles act on our simple light propagation scheme to have a kind of tensor controlled directive effect on the light transport. The transmission profiles should modulate the outgoing intensities in a sense of redirection or more appropriately, redistribution. This will yield an approach, capable of computing approximated final light distributions for a given light source configuration inside of a $2D$ lattice. As this process requires a lot of user-interaction and is not to be concerned as generalizing to all possible scenarios. In the sense of automization and optimzation, another consequent aim is to implement an approach borrowed from vector-field visualization to reveal structures, which are hard to grasp by ordinary methods. We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges representing LCS (Lagrangian Coherent Structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature, such as moraines in glaciers or vortices in flow fields. Commonly, in vector field visualization, LCS separate regions of different flow behavior, but in our case they separate regions/subdomains of varying directional distributions (manifesting in diverging/converging tensor field lines $\rightarrow$ compressions, tensions, roations or varying tensor magnitude $\rightarrow$ volume changes) represented by tensors, as we use our tensor-controlled light transport visualization (propagation scheme) as a basis or framework to compute the FTLE field. Consequently, we propose another method to compute an FTLE-like field on tensor fields. The FTLE field responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions and since we effectively measure the finite-time Lyapunov exponent (FTLE) with both forward and backward integration time (stimulus directions, in our case), we will detect both cases. Both invented methods, will give new insights into the topology and anisotropy of tensor fields and provide new tools for the structural analysis. The greater goal of this work is therefore the encoding of a tensor field into certain abstracted representations and the interpretation of meaningful results evaluating the proposed techniques for possible uses. The designed approach could ,e.g. , be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations. A classical example for tensor field visualization is plotting hyperstreamlines following major/minor eigenvectors, as depicted in Fig. \ref{streamlines}, which features a degenerate point (purple), a wedge (red) and a trisector (green).

%In diesem Abschnitt sollen neben den Herausforderungen und der
%Problemstellung insbesondere die Ziele der Arbeit beschrieben werden. 
\section{Structure of this Work}

For the rest of this work, we will give a quick overview on the structure and outline. Next, in chap. \ref{chap:grundlagen}, we will capture important fundamentals like tensors, polar coordinates, PCA in the context of singular value decomposition (generalization of eigenvalue decomposition), FTLE and GT datasets. Also we will discuss the most relevant related work in the domain of tensor field visualization and also in the domain of global illumination (indirect illumination rendering). Subsequently, in chap. \ref{chap:main}, we will propose the main contributions and methods of this work and document the results and evaluation in chap. \ref{chap:eval}. At last, a compact summary, conclusion and outlook will be given in chap. \ref{chap:concl}.\\

%Dieser Abschnitt wird meist recht kurz gehalten und beschreibt im
%Prinzip nur den Aufbau des Rests der Arbeit. Zum Beispiel: In Kapitel
%\ref{chap:grundlagen} geben wir einen ?berblick ?ber die  Grundlagen
%zu der Arbeit sowie ?ber verwandte Arbeiten. In Kapitel
%\ref{chap:main} stellen wir dann \ldots vor. \ldots etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Fundamentals}
\label{chap:grundlagen}

In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and we will capture tensor fields, polar coordinates, PCA, FTLE, GT datasets and related work in that sense.

%Die ersten paar Abschnitte in diesem Kapitel f?hren in die Grundlagen
%zur Arbeit ein. Das k?nnen beispielsweise Grundlagen zu Netzwerken
%oder zur Informationsextraktion sein.

\section{Tensor Fields}
\label{sec:tensors}
Now what is a tensor? A tensor is a $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
\begin{table}[!b]
\centering
\caption{Tensor Shapes}
 \begin{tabular}{c|c|c|c|c|c|c}
\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
\hline 
\textbf{shape} & scalar & vector & matrix & ``3D matrix'' & ... & $\mathop{o}$D matrix\\ 
\end{tabular}
\label{tensor}
 \end{table} 
In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ in space (the dimension of the row and column vectors), which is equal to the rank of the tensor for full rank (in other words the number of column elements n for a $n{\times}n$-matrix). A tensor is considered non-degenerate if all eigenvalues $\lambda_i>\lambda_{i+1}$ manifesting in significant anisotropy. Tensors follow certain transformation rules which are defined for covariant/contravariant tensors, which indicate the change characteristics under a basis change. Contravariant tensors do not change under a change of basis, whereas covariant tensors are not invariant w.r.t. basis changes. Tensors are most commonly obtained and defined from mathematical transformation equations. Hence, any $oD$ number array could be a tensor, but the definition holds only if the transformation rules apply to these $oD$ number arrays (in a sense of mathematical/formal interpretation). This is mostly true for component indexed $oD$ number arrays following matrix multiplication and transformation rules found in physics and math manifesting in calculation rules like matrix multiplication. Thus, a tensor in general is a $o$D number array arranged in meaningful (transformation) order, to put it simple. In fact, tensors themselves are interpreted as transformations which reflect an incoming direction vector (measuring, e.g., the stress in that direction) to an outgoing (stress) vector. To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities are not simple numbers or vectors. They occur in form of a tensors, and thus as entities or quantities in physics. That is, at each point (typically a Euclidean space or manifold) in space (indexed by the grid) there is a whole distribution of directions, e.g., stresses (elastic and viscous), which needs to be characterized and represented by a single tensor each. Collectively, defined at each point in space, these tensors form a tensor field.
\paragraph{stress tensor}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/stress_tensor.png}
  }
  \caption{Cauchy stress Tensor}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{https://en.wikipedia.org/wiki/Cauchy_stress_tensor\#/media/File:Components_stress_tensor_cartesian.svg}}
  \label{cauchy}
\end{figure}

The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig. \ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$, arranged in row-major order. It appears in similar form as viscous stress tensor in fluid mechanics:
\begin{align*}
\boldsymbol{\sigma}
 = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
\boldsymbol{T}^{(\mathbf{e}_2)} \\
\boldsymbol{T}^{(\mathbf{e}_3)} \\
\end{matrix}}\right] =
\left[{\begin{matrix}
\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
\end{matrix}}\right]  = \left[{\begin{matrix}
\sigma _x & \tau _{xy} & \tau _{xz} \\
\tau _{yx} & \sigma _y & \tau _{yz} \\
\tau _{zx} & \tau _{zy} & \sigma _z \\
\end{matrix}}\right] =
\begin{bmatrix}
\sigma _{11} & \sigma _{12} & \sigma _{13} \\
\sigma _{21} & \sigma _{22} & \sigma _{23} \\
\sigma _{31} & \sigma _{32} & \sigma _{33} 
\end{bmatrix}.
\end{align*}
These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in plane normal directions ${x,y,z}$ ($3^2=9$ individual numbers). Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space. The stress tensor itself can be interpreted as a transformation, which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$. The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent (kind of a Euclidean distance of the stress components). The tensor itself is a complicated representation and needs to be projected onto its principal axes to lower its dimensionality, as it is the case for principal component analysis. It is possible to find $2$ principal stress directions in $2D$ through PCA (see sect. \ref{sec:pca} below), for which there are no shear stresses existent in equilibrium. These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour. The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''. So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig. \ref{hat} in sect. \ref{sec:pca}. Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle. It is therefore somehow representing the distortion of an absolutely isotropic object (unit circle) by whatever the underlying affine tensor transformation effects (e.g. stretching, shearing, mirroring, scaling, rotating). Let us consider it as some lower dimensional, more intuitive representation of the tensor which we were left with in the first place. To put it all together, a tensor is a mathematical transformation which is encoded inside of an abstract $oD$ number array, which comes in need, whenever we have a whole distribution of directions manifesting itself inside of a tensor (matrix) transformation.

\section{Polar Coordinates}
\label{sec:polar}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.8\textwidth]
    {img/polarkoordinaten.png}
  }
  \caption{polar space}
  \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:Ebene_polarkoordinaten.PNG}}
  \label{polar}
\end{figure}\vskip 3pt
Mathematically speaking, a polar space is a system, which maps each angle $\omega$ a magnitude $r(\omega)$ through a symbolic function. The conversion formula from polar (given a radius $r$ and angle $\omega$) to Cartesian coordinates is depicted in Fig. \ref{polar} and is given by:
\begin{align*}
	x &= r \cos \omega, \\
 	y &= r \sin \omega.
\end{align*}
Note, that we switched angle $\phi$ to $\omega$ for reasonable discrimination to the radiant intensity. The inverse direction (given coordinates $x,y$) is a bit more complicated:
\begin{align*}
	r &= \sqrt{x^2 + y^2} \\
\omega &= \operatorname{atan2}(y, x)
\end{align*}
The polar domain is a circular axis version of the domain $[0..2\pi]$, with negative magnitudes in $r$ shifted by 180 degrees (point reflected) as this is the formal behaviour defined and required by the conversion formulas.

\section{Principal Component Analysis}
\label{sec:pca}
Note, that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following. Also keep in mind, that we interpret a matrix or tensor as affine transformation. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by, e.g., ellipsoid glyphs.

\begin{figure}[!t]
\hskip 22pt
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/HAT.png}
   \caption{2D principal component ellipsoid\/ glyph (eigenframe)}
   \label{hat}
   \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:HAT-ellipse0.svg}}
  \end{minipage} \hskip 40pt
  %\raggedright
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/glyphs.png}
    \caption{3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
    \label{3Dglyphs}
    \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://slideplayer.com/slide/5291764/}}
  \end{minipage}
\end{figure}

Principal Component Analysis is an algorithm, which is able to capture the principal components (directions) in stochastic data. But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling. Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices. For (skew-) symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf. \cite{kieburg}). Note, that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases (cf. \cite{moler}). Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling (classically: $\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of transformation matrix. The singular value decomposition is defined as follows:
\begin{align*}
	\mathbf{A} = \mathbf{U} \boldsymbol{\Sigma} \mathbf{V}
\end{align*}
where
\begin{itemize}
	\item $\mathbf{U}$ is an $m ? m$ unitary matrix
	\item $\boldsymbol{\Sigma}$ is a diagonal matrix {$m ? n$} with positive real numbers on the diagonal
	\item $\mathbf{V}$ is an {$n ? n$} unitary matrix, $\mathbf{V}^*$ is its conjugate transpose
\end{itemize}

In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues (\cite{mroz}) ordered decreasingly by absolute value (corresponding to singular value order). The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses. We do this decomposition analysis once for each matrix and then store the precomputed results in a grid. The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step, which makes it a basic and central concept within the scope of this work. Within this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition. This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor. We will then use global-illumination techniques on top to generate a light transport flow map, which permits us somehow to forward the intensities which are given as predefined initialization values in form of symbolic functions defined by the user. This kind of light transport propagation scheme will also allow us to compute an FTLE-like field on the tensor fields.

\section{Finite-Time Lyapunov Exponent}
\label{sec:ftle}
The Finite-Time Lyapunov Exponent is a measure, developed to analyze time-dependent dynamical systems for their separation abilities concerning massless tracer particles. We imagine a flow field, i.e., a vector field which in turn represents the effective amount or concentration shifted into a particular direction at each point in space. It could induced by e.g. concentration or height gradients. This flow field is considered a time-dependent dynamical system, since it usually holds a time-component $t$ and thus maps each initial position an aged (altered) position: $\mathbf{\upvarphi}(\mathbf{x}) : \mathbf{x}(t_0) \longmapsto \mathbf{x}(t_0+T)$. We can now imagine placing two close by particles in distance $\delta$ into a flow field and run a time dependent simulation. Remark that this corresponds to the Lagrangian view in vector fields, which travels with the particles (instead of taking the position of a resting observer - Eulerian view):

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/ftle_sep.pdf}
  }
  \caption{FTLE: separation of massless tracer particles}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: Skript Prof. Sadlo, Scientific Visualization SS2017, Heidelberg University}
  \label{ftle_sep}
\end{figure}

We observe particle trajectories diverging w.r.t. the central axis ending up in distance $\Delta$. The FTLE measure for this single example (point in space) is then defined as:

\begin{align*}
	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
\end{align*}

Alternatively, an FTLE can also be computed from a flow map's gradient, which is more convenient to compute in general:
\begin{align*}
	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \lvert\lvert\nabla\mathbf{\upvarphi(x)}\rvert\rvert _2,
\end{align*}
whereas $\lvert\lvert A\rvert\rvert _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$. This will yield a map, which responds most where path lines diverge w.r.t. to a central axis. We consider an example which detects FTLE ridges inside of the cross-section of a double gyre. Note, that this example considers single trajectories diverging implicitly:
\begin{figure}[!t]
\hskip 22pt
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/diverge.png}
   \caption{diverging streamlines for double gyre}
   \label{diverge}
   \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dg_3traj.png}}
  \end{minipage} \hskip 40pt
  %\raggedright
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_gyre.png}
    \caption{FTLE field for double gyre}
    \label{ftle_gyre}
    \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dle_incT.png}}
  \end{minipage}
\end{figure}
Fig \ref{diverge} and \ref{ftle_gyre} depict a scenario of diverging streamlines for a double gyre snapshot in time (yielding a time-independent dynamical system). Fig. \ref{ftle_gyre} features a high ridge on the margins (borders) and in particular at the center of the domain. This region of interest (ROI) is considered to indicate the maximum relative local divergence of tensor field lines, by the approach. Ridges also represent Lagrangian Coherent Structures, which are found in nature in vortices of correlating flow behavior.  We could imagine to introduce this process for a whole distribution of trajectories, which would implement an FTLE for uncertain vector fields in a sense, that whole propability distributions of trajectories could be heeded. An overview about existing visualization techniques for uncertain data is given by Griete \& Schuhmann \cite{griete}. We require that the trajectories can cross each other unaffectedly, i.e., the superposition principle is supposed to be respected for the magnitudes. But behold, is this behavior not somehow familiar? If we elaborately think this through, we can come up with the idea that this is the behavior of light propagation in vacuum. Light paths are physically allowed to cross each other with no influence exerted on each other at all. Consequently, we choose to model the distributions of trajectories as light intensity profiles. That is, we interpret the polar profiles as a propability density function (PDF) for particle trajectories heading into this direction. To put it into context, our main contribution method LTG uses the FTLE as an inspiration or template to design a similar approach for tensor field visualization. The main difference to the template FTLE is that we follow eigenvectors instead of flow vectors and that we use whole distributions of trajectories, as a tensor field suggests whole distributions of directions, accordingly.

\section{GT Datasets / Methods}
\label{sec:gt}
\paragraph{Glyphs and Tensor Field Lines}
%\begin{figure}[!t]
%  \centering
% {
%    \includegraphics[width=0.4\textwidth]
%    {img/tensorfieldlines.png}
%  }
%  \caption{tensor field lines: bold lines; ellipsoids: ellipsoid glyphs}
%  \label{scheme}
%\end{figure}
Glyphs are derived as a basis for the weighting profiles of the tensors in the following propagation scheme. They were chosen because it is an easy-to-use (understand) concept, which can also be interpreted nicely. In fact, these representations are used for the principal component analysis of direction vectors for the tensor field line extrapolation (integration) and also as transmission profiles for the light propagation scheme in a subsequent step. We can again emphasize glyphs as a crucial, central concept within this work. Tensor field lines (the type of tensor streamline analog that does not introduce artificial intertia) were implemented for ground truth (GT) data generation purposes. For example, when designing a FTLE-like field computation technique, we will need to know when tensor field lines diverge/converge. Examples for glyphs (ellipsoids) and tensor field lines (thin lines) are depicted in Fig. \ref{test-fields}.

\paragraph{Test Data (Generation)}
We used python's numpy library to generate predefined, synthetic test cases and caught two real data example from \url{TensorVis.org} and \url{https://www.cs.purdue.edu/homes/cs530/projects/project5.html}, which needed to be pre-processed through slicing and subsequent cropping to obtain $2D$ samples ($2\times2$) matrices). These test fields can be grouped into symmetric and asymmetric ones:\\

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/corner-TFL.png}
    \caption*{``corner''}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/spiral-TFL.png}
    \caption*{``spiral''}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/drain_old.png}
    \caption*{``V''}
    \label{b)}
  \end{minipage}
\caption{Named test fields}
\label{test-fields}
\end{figure}

\textbf{Asymmetric Tensor Fields}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.8\textwidth]{img/drain(alt).png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/drain(alt)-TFL.png}
    \caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{``drain'' test field}
\label{drain}
\end{figure}

The ``drain''-Testfield, depicted in Fig. \ref{drain} is generated to have a kind of proof-of-concept (POC) for the FTLE-related approach. It poses raw, bare diverging tensor field lines without any other feature modulated on top. Thus, it can be expected to induce a high FTLE response, since it is an interpretable ground truth test and is considered to be a basic trigger/stimulus for the approach with no side effects introduced.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.8\textwidth]{img/rings.png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
   \centering
    \includegraphics[width=0.8\textwidth]{img/tensorfieldlines.png}
	\caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{``rings'' test field}
\label{rings}
\end{figure}
The ``rings'' test field, depicted in Fig. \ref{rings} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/spiral.png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/spiral-TFL.png}
    \caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{``spiral''-test field}
\label{spiral}
\end{figure}
The ``spiral'' test field, depicted in Fig. \ref{spiral} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/inverse.png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/inverse-TFL.png}
    \caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{``inverse'' test field}
\label{inverse}
\end{figure}
With the \enquote{inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/gyre.png}
    \caption*{a) glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/gyre-TFL.png}
    \caption*{b) tensor field lines for a)}
    \label{b)}
  \end{minipage}
\caption{a) ``gyre''-test field, b) tensor field lines for a)}
\label{gyre}
\end{figure}
The test field ``gyre'', depicted in Fig \ref{gyre}, constitutes a sustained turn over the grid representing the redirection or forwarding along domain corners.

\textbf{Symmteric Tensor Fields}
We also aim to use a real data example of the $3D$ DTI-MRI diffusion tensor of the brain for a male subject at the age of 27 and another one featuring a human heart. As we visualize 2D symmetric second order tensor fields, we need to slice the dataset and crop the $2{\times}2$ matrices omitting the $z$-component of the original $3{\times}3$ matrices. Fig. \ref{brainglyphs} depicts the glyph representation of the test dataset for slice index $15$. To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{brain} and \enquote{heart} scans, which can be observed in Fig. \ref{real1} and \ref{real2}.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.8\textwidth]{img/brain_org.png}
    \label{a)}
    \caption*{tensor magnitude}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.8\textwidth]{img/brain_cyl_glyphs.png}
    \label{b)}
    \caption*{cyclinder glyphs}
  \end{minipage}
\caption{Real data examples: \enquote{brain} dataset}
\label{real1}
\end{figure}
Observing the \enquote{brain} dataset in Fig \ref{real1}, one may notice that it introduces a lot of variances to tensor magnitude and linear anisotropy. We choose this DTI-MRI scan as an example-par-excellence for real application data given in a reasonable resolution.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/heart_org.png}
    \label{a)}
    \caption*{tensor magnitude}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
      \includegraphics[width=0.8\textwidth]{img/heart_cyl_glyphs.png}
    \label{b)}
    \caption*{cyclinder glyphs}
  \end{minipage}
\caption{Real data examples \enquote{heart} dataset}
\label{real2}
\end{figure}
When we look at this DT-MRI scan of a human heart we can notice the elongated shape, which we consider to represent the aorta and a part of the rest of the heart. If we consider the glyph representation, we again infer, that we incorporate lots of anisotropic regions with varying tensor magnitudes here.

\chapter{Related Work}
\label{chap:related}

There has been extensive recent work on symmetric tensors but in comparison little on asymmetric ones. Since we are not able to cover the whole scope in this work, we will focus on the most relevant works on tensor field visualization. We will talk about the publications from Hlawatsch et al. \cite{hlawatsch} and Zirr et al. \cite{zirr} in particular, since they constitute thematically close related works. But first, we will cover related works on global illumination methods, as these form a basis or prerequisite for our work. Please note, that they do not focus on tensor field visualization at all. Remember, that we will inventively employ global illumination methods for tensor field visualization as an innovative step and that we want to cover previous works on our base approach, as well.

\paragraph{Global Illumination Methods}
Since we use a specific category of global illumination methods as a basis, we decide to only discuss lattice based methods which work, corresponding to their name, on a grid. Remark, that Global Illumination methods classically focus on realistic indirect lighting. Thus, they aim on producing a most realistic image, what in our case turns out to be a most informative image in correspondence. The most early work in this field is propably the Discrete Ordinates Method (DOM) by Chandrasekhar \cite{Chandrasekhar}, which is used specifically for computing radiative transfer in participating media (volume rendering). The radiative transfer equation is discretized spatially and in the angular domain. The radiance distribution (PDF) is stored in a grid and light is distributed and exchanged between neighboring volume elements which reduces the overall effort to local and independent computations. Furthermore, there has been more recent work on describing the light propagation as a diffusion process within a simple photon transport model based on the Lattice-Boltzmann method \cite{geist}. DOMs were recently revisited by Fattal \cite{fattal}. This work adresses artifacts stemming from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}. These methods are considered to be too slow to \enquote{run at interactive framerates} by Dachsbacher et al. \cite{dachsbacher}, though \enquote{efficient and potentially highly parallel}. Our work is inspired by Lattice-Boltzmann methods (volume grid radiosity) and the lattice-based discrete ordinate methods concerning the light propagation scheme.

\paragraph{Symmetric Tensor Field Visualization}
Zheng \& and Pang \cite{pang&zheng} et al. proposed a texture-based visualization approach called Hyper LIC, which extends the concept of LIC to symmetric tensor fields by using an anisotropic 2D-filter kernel oriented along the major/minor eigenvectors. The concept of following the major/minor eigenvector along tensor field lines was invented by Vilanova et al. \cite{vilanova}. Tensorlines \cite{weinstein} introduce a kind of artificial inertia (running average), making them resistent to noise. The notion of tensor field topology (critical, degenerate points, separatrices) and the concept of hyperstreamlines was first introduced by Delmarcelle and Hesselink \cite{delmarcelle&hesselink}. The topology in $3D$ tensor fields was first analyzed by Zheng et al. \cite{zheng}. The placement of tensor field lines as hyperstreamlines for glyph packing was recently improved by Spencer et al. \cite{spencer}. A good overview of seeding strategies concerning glyphs is given by McLoughlin et al. \cite{mcloughlin}. Feng et al. \cite{feng} used Voronoi Tesselation for placing the glyphs. Superquadric glyphs have been proposed by Schultz and Kindlmann \cite{schultz&kindlmann} for general (non-positive definite) tensors. De Leeuw and van Wijk \cite{deleeuw&vanwijk} visualize the partial derivative gradient, the Jacobian of the tensor field, to sense the local properties of the field. A diffusion tensor can also be visualized by box \cite{makris}, ellipsoid \cite{pierpaoli}, composite \cite{westin2} or superquadric \cite{kindlmann} glyphs. There is also a set of scalar measures available in symmetric tensor field visualization such as: mean diffusivity, fractional anisotropy \cite{basser} and anisotropy coefficients \cite{westin}. These coefficients can also be visualized by direct volume rendering \cite{kindlmann2}. Hlawatsch et al. \cite{hlawatsch} obtain a coordinate (flow) map from tensor field lines following major/minor eigenvectors (fibre trajectories) using the maximum eigenvalue of the Cauchy-Green deformation tensor to compute the FTLE-field in homogeneous regions. In a similar manner, we aim to use light transport techniques to obtain a final light distribution (flow map) through successive light propagation cycles directed (in form of transmission profiles obtained from ellipsoid glyphs) by the tensor field. DT-MRI diffusion tensor imaging can help to reveal the functional \cite{denis} or topological \cite{wakana} structure of the brain and muscles by analysing celebral and muscle tissue tractography, which yields a tensor field representing the anisotropy characteristcs of the brain's white matter. An overview about DT-MRI visualization techniques is given by \cite{vilanova}. Tricoche et al. \cite{xavier} use invariant manifolds to extract topological features from tensor fields. Other feature extraction-based approaches extract general features as: edges \cite{hancock}, tensor shape\/ orientation \cite{gordon}, interfaces \cite{donnel}, creases \cite{tricochet}.
\\
\paragraph{Asymmetric Tensor Field Visualization}
Zheng and Pang \cite{pang&zheng} proposed the concept of dual eigenvectors and Zhang et al. \cite{zhang} extended it to pseudo-eigenvectors and introduced the eigenvector/eigenvalue manifolds to visualize eigenvectors in the complex domain. Laramee et al. \cite{laramee} focused on the efficient implementation and visualization of these structures and provided an interactive visualization system for asymmetric tensors applicable in fluid and solid dynamics. The concept of tensor magnitude has been introduced by \cite{laramee} for means of physical interpretation. They also proposed an efficient glyph and hyperstreamline hybrid approach, which made dynamic interaction in real-time in $2D$ tensor fields feasible. Palacios et al. \cite{palacios} extracted isosurfaces of tensor magnitude, mode and isotropy index.

Last, we will discuss our work in relation to other publications. While these previously mentioned works use classical approaches like tensor field lines and glyphs, we will also introduce the concept of Lagrangian coherent structures into our framework contribution to tensor field analysis. LCS have been defined as time-dependent analog of separatrices \cite{haller} which are concerned to be robust under noisy conditions \cite{haller2}, which we expect from our technique as well by implication. Hlawatsch et al. \cite{hlawatsch} obtain similar results (a FTLE-like field on tensor fields), but obtain them by sampling of fiber trajectories instead of global illumination light distributions, as it is the case for this work. Our work shares the same basic idea, deriving and computing a FTLE-like field for tensor field visualization generalizing FTLE from vectors to tensors, yet using a light propagation scheme and hence using whole distributions of trajectories incorporating polar profiles. Concerning the tensor field visualization method (LTG), our approach is based on FTLE, and principal component analysis (PCA). We will detect LCS with an FTLE-derived method considering differences in final light transport distributions through computing the gradient of the resulting flow maps. The FTLE-related approach named ``light transport gradient'', to be designed, also states a generalization of the light transport visualization method FTPD proposed by Zirr et al. \cite{zirr}, since it allows a similar measure to be computed through defining a geometrical scene and setting the transmission (defined by the tensor field) to an isotropic and constant value of $100\%$. In this operating mode, the approach operates as a light transport visualization technique neglecting any tensor field transmission profiles and is thus capable of displaying bounding separatrices of a light transport scenario perturbated by obstacles \cite{zirr}. 
%Typischerweise im letzten Abschnitt dieses Kapitels wird dann auf
%verwandte Arbeiten eingegangen. Entsprechende Arbeiten sind geeignet
%zu zitieren. Beispiel: Die wurde erstmalig in den Arbeiten von Spitz
%und Gertz \cite{Spitz2016a} gezeigt \ldots Details dazu werden in
%dem Buch von Newman zu Netzwerken \cite{Newman2010} erl?utert \ldots.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Alternative: put content in separate files
% Check the difference between including these files using \input{filename} and \include{filename} and see which one you like better
%\chapter{Einleitung}\label{intro}
%\input{introduction}
%
%\chapter{Voraussetzungen}\label{bg}
%\input{background}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

\chapter{Method}
\label{chap:meth}

This chapter will summarize the main contributions of this work studied and state targeted aims and goals as a first step in sect. \ref{sec:motiv}. Then we will have a short introduction to the implemented global illumination scheme, derived and simplified from a global illumination approach, invented for the Crytek GmbH in the context of global indirect illumination rendering, in sect. \ref{sec:scheme}. Last, we will state the effectively implemented tensor field visualization methods: light transport visualization (directed propagation scheme yielding $2D$ scalar fields) in sects. \ref{sec:scheme} \& \ref{sec:trans} and light transport gradient (LTG: yielding $3D$ scalar fields) in sect. \ref{sec:ltg}, which are proposed as the invented tensor visualization techniques and hence as conceptual contributions to the domain of tensor field visualization.

%Dieses Kapitel stellt meist den Hauptteil der Arbeit dar. Vor dem
%ersten Abschnitt sollte ein kurzer ?berblick (ein paar wenige S?tze
%mit Verweise auf nachfolgende Abschnitte) gegeben werden. Beispiel: Im
%nachfolgenden Abschnitt \ref{sec:overview} wir ein ?berblick ?ber die Anforderungen an
%das Modell gegeben.

\section{Motivation}
\label{sec:spec} 
Our ambition is, to motivate a simple and efficient global illumination propagation scheme, able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t. energy conservation and propagation attenuation principles. We then generate transmission profiles, which we sample in a preliminary step, from the eigensystems (principal component ellipsoids) of the tensors for each cell in the domain (field). This step is done, to provide an orientation for a lattice of crystal fibre structures (cf. optical fibres) to redirect the intensities in analogy to the anisotropy characteristics of the tensor field within a glyph representation, obtained from principal component analysis (PCA). Our approach should be implicitly adaptive to any kind of 2D dataset or slices of 3D datasets. Thus, we require it to be compatible to handle any kind of matrix (symm./anti-symm.) and any kind of resolution of the field (fully-adaptive). At last, our ambition is to segment key locations (LCS/ridges) in the field with tensor field lines (hyperstreamlines) converging/diverging, which is necessarily the same (for bidirectional eigenvectors/singular vectors of tensors). Thus, we aim to generate a new FTLE-related method for tensor field visualization by applying a light transport gradient to the resulting light distributions for every possible position and direction of the light source in the grid (lattice) generated by stimulation with Dirac-pulses. Our approach should efficiently handle (relatively) large datasets with reasonable resolutions, which turns out to be possible with additional computing power empowered by a cluster. 
%Knapp zwei Seiten, in dem die Anforderungen, die Zielsetzung und die
%Methoden ?berblicksartig beschrieben werden. Hier sollte die
%Beschreibung ``technischer'' bzw.~``formaler'' sein als in der
%Einleitung, da der Leser nun mit den Grundlagen und verwandten
%Arbeiten vertraut ist.

\section{Light Transport (Propagation) Scheme}
\label{sec:scheme}
As already discussed, we will design a new method for visualizing tensor fields, which is based on the physical nature of light propagation. For this purpose, we motivate a light transport propagation scheme designed to efficiently propagate initial intensity profiles stored in polar coordinates given as symbolic impulses (stimuli) by the user defined in the configuration file. This polar profiles are first sampled into a vector with elements ranging over the interval $[0,2\pi]$. Polar profiles for different light sources are shown in Fig. \ref{polar}.
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.8\textwidth]
    {img/strahler.png}
  }
  \caption{Polar profiles for different types of emitters ($m=0$: spherical emitter, $m=1$: Lambert emitter, $m=2$: cone emitter}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{http://www.optik.uni-erlangen.de/odem/vorlesung/ws1213/Photometrie_Geometrische_Optik.pdf}}
  \label{polar}
\end{figure}\vskip 3pt
As a reminder, polar profiles map each angle $\omega$ a magnitude $r$ forming a function $r(\omega)$. They are circular axis versions of the domain $[0..2\pi]$, with negative magnitudes point reflected by 180 degrees. Thus, they are functions with an angular domain making them directionally dependent, which is a well suited representation for point lights (the type of light source with theoretically no finite extend), as we use them in a style of Dachsbacher et al. \cite{dachsbacher}. For the sake of interpretability, we exclude negative values (energies) from the scope of our calculations which we express through mathematical clip functions, limiting negative values to $0.0$, which we denote by subscript $+$ in the following. We will describe the emission charactersitics of point lights exchanging energies in the lattice by parsing such polar functions. Our grid consists of $\mathop{dim}=\mathop{width}\cdot \mathop{height}$ virtual point lights (VPLs), which are in turn light source with (theoretically) no finite extend represented through polar functions. As a start-up thinker for the propagation scheme, we will think of a grid cell in 2D as a center pixel within an 8-neighborhood with a previously injected point light (imagine a spherical emitter: $r(\omega)=1$ here for simplicity). Inside of this cell, we imagine a circular polar profile of radiant intensities for every angle $\omega$ given in $[\frac{W}{\mathop{rad}}]$. Remember, that all considerations from now on are given concerning a single grid cell respectively. Each cell has 8 (4 faces/edges and 4 diagonals) unique neighbors, which are adjacent point lights in the grid (initialized to $I(\omega)=0$). We consider one face neighbor (top) to shortly explain the steps of the scheme. The diagonals and the others are obtained from derivation and symmetry. Naively, one would consider to just transmit the intensity in the direction its pointing into. But this would omit every propagation attenuation principles (any principles at all specifically), if one takes a closer look:
\paragraph{Accumulation stage}
Each cone type (yellow and green) is an individual neighbor-dependent angular area. The diagonals take in a greater angular domain while they are located at greater distance and thus need to be attenuated more strongly (consider an isotropic stimulus). We aim for distributing contributions for overlapping cones to transmit attenuated (because of the larger factor $\sqrt{2}$ distance) intensity in form of a shared part to the diagonals in a single step (which also prevents from zigzagging diagonal intensities in case we would transmit diagonal intensities around the corner). Therefore, the intensity gathered in polar coordinates inside of angular ranges marked in green, needs to be partitioned between 2 of the 8 cone index neighbors with following considerations:

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.8\textwidth]
    {img/geogebra-export_large.pdf}
  }
  \caption{Propagation Scheme}
  \label{scheme}
\end{figure}\vskip 3pt

The face neighbors take in the angular domain of $d_1=\alpha+2\beta$ (yellow cone), whereas the diagonal neighbors (green cones) span a domain of $d_2=2\beta$.
The linear combination weights for the overlapping (green) parts ($\beta$) are obtained from the relative overlapping angular area of the angular span $\beta$ w.r.t. diagonal cones ($2\beta$) and yellow face cones ($\alpha+2\beta = 90^\circ$) (yellow cone is overlapped by green cones at full angle). The linear combination weights are subsequently normalized to $\sum = 1.0$ (normalization condition):
\begin{align}
	\varepsilon_\alpha &= \frac{\frac{\beta}{\alpha+2\beta}}{\frac{\beta}{\alpha+2\beta} + \frac{\beta}{2\beta}} \approx 0.362291\\
	\varepsilon_\beta &= \frac{\frac{\beta}{2\beta}}{\frac{\beta}{\alpha+2\beta} + \frac{\beta}{2\beta}} \approx 0.63771.
\end{align}

Remark, that we can precompute these linear combination weights and reutilize them in each iteration. We define the energetic contributions (presuming the cone is given in local coordinates - i.e. centered around the origin) as follows:
\begin{align}
	\Phi_{\alpha} &= \varepsilon_\alpha\int_\alpha {I}(\omega)\mathop{d\omega} =\varepsilon_\alpha(\int_{-\pi/4}^{-\alpha/2}{I}(\omega)\mathop{d\omega}+\int_{\alpha/2}^{\pi/4}{I}(\omega)\mathop{d\omega})+\int_{-\alpha/2}^{\alpha/2}{I}(\omega)\mathop{d\omega}\\
	\Phi_{\beta} &= \varepsilon_\beta\int_\beta {I}(\omega)\mathop{d\omega} = \varepsilon_\beta\int_{-\beta}^{\beta}{I}(\omega)\mathop{d\omega}
\end{align}

\paragraph{Injection stage}
 We place a cosine lobe scaled with the integrated (accumulated) energies in each of the 8 cone directions (angular domains), whereas outgoing integrated and summed energies are weighted by the linear combination weights (factors) in overlapping (green) areas to split up the contributions accordingly. The contribution areas are depicted in Fig. \ref{scheme}, the cosine lobes can be observed in Fig. \ref{iter} respectively. This manner of propagation is inspired by the propagation scheme of Dachsbacher et al. \cite{dachsbacher} (see the paper on how to ``project the flux into a point light''). Basically, we sum up the weighted intensities corresponding to the respective neighbor index and scale an oriented cosine lobe (in direction of the neighbor) with the total flux $\Phi_t$ accumulated inside the neighbor range to transmit all of the parsed intensity for each neighbor distributed into a cosine distribution. We repeat this process once for each neighbor and each cell respectively. When we are done with the whole grid, we just recursively use the result and keep on running propagation cycles. The implementation is done in a dual buffer approach which pushes the energies back and forth (buffer A to buffer B) for each propagation cyvle until a state of convergence is reached, which is the case when the energy is spreaded throughout the grid and enters a stationary state of equilibrium, characterized by equal out (at grid borders) to in (at light source positions) energy-flow with no more nettings going on inside the domain. We measure the convergence by setting up an overall distribution error w.r.t to the previous iteration. Note that even though we propagate the energy directly through the diagonal cones (which forms a square initially), the 8 degrees of freedom introduce a circular (spherical) wave front after few iterations (for greater resolutions) as depicted in Fig. \ref{iter}. This occurs, because of a faster propagation attenuation which counteracts the larger angle of diagonal cones. There is a redistribution of intensities in the long term resulting in a, since the shared part effects a uniformly spaced isotropic propagation profile resulting in a circular torus-shaped distribution of intensities, as also observed in Fig. \ref{iter}. Also note, that this approach satisfies energy conservation and propagation attenuation principles and applies to the principles of light propagation in vacuum (by default). We infer, that we can assume to follow physical models of spherical waves. The verification of these principal requirements and specifications can be observed in the evaluation in chap. \ref{chap:eval}. We first show the approach for a few iterations with a manual stop criterion to have a kind of imagination for the propagation process:

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=1.0\textwidth]
    {img/steps_gallery_format_arrow_alt.png}
  }
  \caption{Impulse response for isotropic stimulus (few iteration snaphots: $t$)}
  \label{iter}
\end{figure}\vskip 3pt
Fig \ref{iter} depicts the named iteration indices $0,1,2,3,7,11$ (cycles) for the propagation scheme for a circular pulse stimulus which gets turned off after 1 iteration to demonstrate raw energy spreading in the field (impulse response of the algorithm). In step 1 one can observe the placing of cosine lobes scaled with the respective accumulated and weighted energy for the summed area in each of the 8 neighbor directions. In the subsequent steps, remark the evolution from an 8-neighborhood in step 1 to a circular shaped torus.

As already mentioned, a dual-buffer approach is chosen here, as we can efficiently propagate from one into the other and just swap pointers to 
start all over (after reinitialization) until convergence. Another reason is it yields a reasonable source-target or read-write segmentation. We introduce a new operator called \textit{propagate}, permitting us to pre-define the propagation in a symbolic manner. The propagation definition denotes the transformation from one grid (A-buffer) into another (B-buffer) through proper application of the \textit{propagate}-operator to each cell. We will extend this operator with a normalization correction in sect. \ref{sec:trans} in form of a precise mathematical formulation respectively. For now, just imagine the approach in raw operating mode, with no tensor influence incorporated. That is, evaluating and accumulating the polar profiles, weighting with the linear combination weights and scaling and placing of a cosine lobe to the respective neighbor. For now, we want to define our operator $\mathop{propagate}$: Each Grid or Buffer is basically a set of cells defined as $\mathcal{C} = \{\mathbf{c}_1, \mathbf{c}_2, ..., \mathbf{c}_i\}$.  The dual buffer result $\mathcal{C'}$ is denoted symbolically as follows for all cells with a mean intensity greater than zero (those which do not contain trivial null samples):
\begin{align}
	\mathcal{C'} = \mathop{propagate}\{c_i\in \mathcal{C} \ \mid \ |\mean{I}_i| > 0\} \hskip 5pt {\forall } \ i \in [1,\mathop{dim}].
\end{align}
where: $i$: cell index, $I$: source buffer radiant intensity\vskip 5pt
The dual buffer result $\mathcal{C'}$ denotes the set of all transformed cells after the propagation, i.e., after applying the operator $\mathop{propagate}$ to each cell. This result is generated once in each iteration and gets pushed back and forth between the 2 buffers. Note that, $\mathop{dim}=width\cdot height$, which equals the tensor field resolution in this case. At last, we define a threshold for the minimum overall distribution error for the total radiant flux taken into account for a convergence criterion, initiating a stop sequence when falling below that threshold $\epsilon$ (, e.g., $0.5$) :\\
\paragraph{Criterion:}
\begin{align*}
	\Delta \Phi_{total} &= \sum_{c_i\in\mathcal{C}}\lvert\Delta I_i(\omega)\rvert = \sum_{c_i\in\mathcal{C}} \int_0^{2\pi} \lvert(I'_i(\omega)-I_i(\omega))\rvert \mathop{d\omega}  \overset{!}{<} \epsilon\\
\end{align*}
where $\phi_{total}$: total radiant flux, $I'$: target buffer radiant intensity\vskip 5pt
Note that we take the absolute value of the differences to prevent from mutual compensation.

Until now, we aimed to simulate the propagation of light in empty space (vacuum) providing us with a physically-motivated but very simplified base-approach. We aim to modulate the transmission of the intensities on top of the base approach with transmission profiles obtained from the tensor fields through principal component analysis, which yields ellipsoid glyph equations in the following.

%In diesem und den nachfolgenden Abschnitten werden die Beitr?ge der
%Arbeit motiviert, formal sauber (oft mathatisch, sprich mit
%Definitionen etc.) beschrieben, und bei Bedarf mithilfe von Beispielen
%verdeutlicht. Die Beschreibungen in diesem Kapitel sind meist
%unabh?ngig von einer konkreten Realisierung und Daten; diese werden im
%nachfolgenden Kapitel detailliert.

\section{Transmission Profiles and Weighting Functions}
\label{sec:trans}
Note that the following considerations are only given for the tensor profile operating mode with tensor influence incorporated. To obtain a kind of footprint of the tensor field, we set up ellipse equations for each glyph representation, obtained from applying PCA on each cell (tensor) in the grid. Concurrently, we extend our abstract mathematical operator \textit{propagate}, previously referenced in sect \ref{sec:scheme}, by a normalization strategy necessary for windowing weighting profiles, as it is the case for our elliptical transmission profiles, which we will capture in the following. At this point, we presume the precomputation of singular value decomposition and linear combination weights. We map the singular values (corresponding to eigenvalues $\lambda$ here for simplicity) in decreasing order to the ellipses radii (half-axes) $a(\Delta x)$ and $b(\Delta y)$ as follows:

Ellipse Equation:
\begin{align}
	r(\omega) = \frac{ab}{\sqrt{a^{2}\sin^{2}(\omega-\varphi)+b^{2}\cos^{2}(\omega-\varphi)}} =\frac{\lambda_1\lambda_2}{\sqrt{\lambda_1^2\sin^2(\omega-\varphi)+\lambda_2^2\cos^2(\omega-\varphi)}}.
\end{align}
That is, the singular values form the half-axes of the PCA ellipsoid, which yields us a symbolic definition for our transmission (transfer) functions as weighting profiles which we can now evaluate as a parsing and accumulating step. We compute the offset angle $\varphi$ w.r.t. the x-axis by exploiting the $\mathop{atan2}$-function with 4-quadrant evaluation for the first singular vector (ordered decreasingly in value), since we scale the $x$-axis with the first singular value. As a preliminary preparation, this profile is precomputed (sampled) for all discrete angular steps for each cell (, e.g., $360$) and stored for reuitilization as it is the case for the previously mentioned linear combination weights as well. The weighting scheme which we apply is depicted in Fig. \ref{weighting}.
\begin{figure}[!t]
  \centering
  {
    \includegraphics[width=0.5\textwidth]
    {img/polarplot.pdf}
  }
  \caption{red: Intensity Profile, cyan: Transmission Profile (glyph Ellipsoid Eq.)}
  \label{weighting}
\end{figure}
We interpret the polar functions as discrete, sampled 1D vectors with a uniform angular resolution. That means, we can use element-wise multiplication for the 1D-vectors in red (Intensity Profile) and cyan (Transmission Profile) to obtain a windowed version of the intensity profile and perform the following integration step to accumulate the total radiant power (flux) as follows. Note that we model angular dependent radiant intensity ${I}(\omega) [\frac{W}{\mathop{rad}}]$ as an underlying physical entity, which is well suited for describing point light sources. As a side note, the physical quantity perceived as brightness by human eyes, is the radiance (luminance), which is yielded by relating radiant intensity to a finite sensitive area ($[\frac{W}{\mathop{rad}}/m^2]$). The primary definition for our total accumulated incident radiant flux (again presuming that the cone is given in local coordinates - i.e. centered around the origin) is then denoted as:

\begin{align}
	\Phi_{\alpha} &= \varepsilon_\alpha(\int_{-\pi/4}^{-\alpha/2}{T}(\omega){I}(\omega)\mathop{d\omega}+\int_{\alpha/2}^{\pi/4}{T}(\omega){I}(\omega)\mathop{d\omega})+\int_{-\alpha/2}^{\alpha/2}{T}(\omega){I}(\omega)\mathop{d\omega}\\
	\Phi_{\beta} &= \varepsilon_\beta\int_{-\beta}^{\beta}{T}(\omega){I}(\omega)\mathop{d\omega}
\end{align}

This permits us to scale an oriented and clipped cosine-lobe $\cos_+$ with the total radiant flux w. respect to its own total energy when integrated likewise. A discrete cosine lobe never reaches its theoretical sum of $2$ exactly in the interval $[{-\frac{\pi}{2}},{\frac{\pi}{2}}]$. Therefore it needs to be computed to account for sampling (discretization) errors and normalized by following relation (, e.g., $\sum_{\cos} \approx 1.9949204635834517$).
\begin{align}
\sum_{\cos_+} &= \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\cos_+(\omega)\mathop{d\omega} \approx 2,\\
	\cos_k(\omega) &= \frac{\Phi_t}{\sum_{\cos_+}}cos_+(\omega-k\cdot\frac{\pi}{4})\hskip 20pt w. \ k\in[0,7].
\end{align}
This specific profile is injected in form of a placement to the corresponding neighbor index $k$. It yields a cosine lobe, which has an overall energy of $\Phi_t$, if the point light is interpreted with a Lambert-emitter (cf. \cite{lindlein}) characteristic (as part or area element of a diffusely reflecting surface) since the intensity profile matches a cosine in this case. This emitter characteristic is the case for global illumination diffuse indirect illumination (non-specular). It is used in similar style by Dachsbacher et al. for propagating intensities through LPVs (light propagation volumes).
But caution, we should not generate any exceeding energy right? Now each tensor ellipse equation has a mean transmission rate $[\%]$, which turns out to be a measure or synonym for the total rate of the outgoing transmitted radiant intensity $\int_0^{2\pi}{T}(\omega)\mathop{d\omega}$. Since energy can not be generated from transmission profiles (only transmitting, not generating), we have to normalize every tensor ellipse equation with the highest mean transmission rate in the field, which is then transmitting the whole outgoing intensity $\int_0^{2\pi}{I}(\omega)\mathop{d\omega}$ in total. This restricts the tensor ellipse mean to a maximum of $100\%$ for the tensor with the highest tensor magnitude. We will use the self-made definition of mean tensor magnitude $\mean{r}_{max}$ obtained from the PCA here for simplicity. That is, we take the principal component ellipsoid and calculate its analytical mean, which is then used for mean normalization to $\mean{r}_{max}=1.0$. Hence the tensor ellipse equations are normalized through the following step:
\begin{align}
	r_n(\omega) = \frac{1}{\mean{r}_{max}}r(\omega).
\end{align}
Notice, that this implements the concept of absorption implicitly, because tensors with a lower mean $\mean{r}$ than $100\%$ absorb energy wheras absorption equals ${\mu} = {1}-{\mean{r}}$ here. Now that we have coped with the problem of overall transmission, we are left with another problem that occurs with the weighting functions as they can still yield values above $1.0$. Consider following examples as $T(\omega)I(\omega)$ pairs (obviously normalized to $1.0$): $1.5\cdot 1.5=2.25$ and $0.5\cdot 0.5=0.25$, which disproves us that a mean $\bar{r}=1.0$ necessarily effects a mean transmission intensity $I_t$ of $1.0<1.25=\frac{2.5}{2}$. This would potentially generate intensity out of nothing which leads to crucially unstable behaviour, since we employ uncontrolled sources of energy in this case. Well, but only in case energy conservation principles were not respected.  To construct a solution, we ask for the source of the excessive energy. If the tensor mean complies with the constraint $\mean{r}\leq 1.0$, then the energy must lack in some other place (since the overall transmission fulfills the energy conservation constraint in that case). The only problem is that a normalization condition in usual style for the transmission profile ${T}(\omega)$ does not necessarily imply the same normalization for the term ${T}(\omega){I}(\omega)$. However, a proper normalization can provide corrective support, wich reveals another subsequent normalization constraint. The normalization factor is defined in 3 subsequent steps:
\begin{enumerate}
	\item Normalization of $TI$ to $\overline{TI}=1.0$
	\item Subsequent scaling w. mean intensity $\mean{I}$ for energy conservation principles
	\item Subsequent scaling w. mean tensor magnitude $\mean{T}$ for absorption principles
\end{enumerate}

Hence the transmitted outgoing intensity for a single direction is defined as:

\begin{align}
    {I}_t(\omega) = 
\begin{cases}
   n_f\varepsilon_\alpha{T}(\omega){I}(\omega),& \text{if } k\bmod 2 = 0\\
     n_f\varepsilon_\beta{T}(\omega){I}(\omega),              & \text{otherwise}
\end{cases}
\end{align}

This transforms our equation for the transmitted radiant flux into:
\begin{align}
	\Phi_{\alpha,t,i} &= \int_\alpha{I}_{t,i}(\omega)\mathop{d\omega} = n_f\Phi_{\alpha,i},\\
	\Phi_{\beta,t,i} &= \int_\beta{I}_{t,i}(\omega)\mathop{d\omega} = n_f\Phi_{\beta,i}.
\end{align}
Which is now respecting energy conservation principles very conveniently, as all we need is a set of means and we are capable of normalizing the weighted intensity profile transmitted. This collapses our abstract mathematical operator $\mathop{propagate}$, which is applied to each cell, into the following steps:
\begin{enumerate}
	\item Computation of normalization factor $n_f$
	\item Direction (component)-wise weighting to account for shared part in diagonal cones
	\item Integration (accumulation) of total radiant flux weighted with \underline{normalized} transmission profiles inside the angular neighbor range (read-access)
	\item Scaling of cosine lobe corresponding to neighbor direction and subsequent placing (injection) to corresponding neighbor (write-access)

\end{enumerate}
Remember that this operator needs to applied once to each cell to form a whole propagation cycle propagating intensities from source to target buffer. This approach ensures that no energy will be generated and therefore implements the 	physical law of energy conservation or Kirchhoff's current law if you want to think of a particle's diffusion process. Consequently, this allows us to assume that all sources of power (energy) are user-defined. In fact the approach has proven to be robust w.r.t. energy losses for several iterations as depicted in chap. \ref{chap:eval} in the stochastic error graph.

\section{Physical Model}
\label{sec:model}
To constitute a physical interpretation for the transmission of the tensor field, we imagine a melting crystal with  fibrous structures aligning with the major eigenvectors of the tensor field. This is what we call the footprint of the tensor field. In that sense of imagination, we would virtually imprint every tensor field into a crystalline structure as a preliminary initialization step. For an every day example, this can be observed in, e.g., the gemstone tiger's eyes' cat's eye effect, where crystal fibers align with the crystal axes respectively and effect a brightened line which never changes position or orientation when rotating the crystal. These oriented crysal fibers are then representing the biased propagation directions of light inside of the crystal lattice. The propability density function $P(\omega)$, for a photon being scattered in a particular direction $\omega$ \ref{ryder}, is then described as:
\begin{align}
 	P(\omega) &= \frac{T(\omega)}{\int_{k\pi}^{(k+1)\pi}T(\omega)\mathop{d\omega}}\\
 	w. \int_0^{\pi} P(\omega) &= 1.
\end{align}
This propability density function (PDF) indicates, whether there is directed anisotropy or spherical isotropy w.r.t the neighboring $180$ ($\pi \mathop{rad}$) degrees and how likely the spontaneuos scattering of a photon in this particular direction will happen. Each tensor is then a unique footprint which can be precomputed for convenience and described as polar profile just as our previously mentioned intensity profiles. We apply these polar profiles in a sense of transmission profiles as discussed in sect. \ref{sec:trans}.

\section{Light Propagation Scheme Behaviour}
To evaluate the plausibility of the light distribution results obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.8\textwidth]{img/isotropic.png}
    \label{a)}
    \caption*{light src in top center}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.8\textwidth]{img/isotropic-all-center.png}
    \label{b)}
    \caption*{light src in center}
  \end{minipage}
\caption{final light propagation distributions for ``isotropy''-testfield}
\label{isotropic}
\end{figure}
One can observe here, how light is propagated with no directional bias leading to an absolutely isotropic distribution in case  b), which is symmetric as expected. This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in sect. \ref{sec:att} and \ref{sect:conserve}. We can also depict the final distribution as a scalar field when using the total energy in each cell as scalar:

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/iso_scalar.png}
    \label{a)}
    \caption*{light src in center}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/iso_sat.png}
    \label{b)}
    \caption*{light src in center (rescaled)}
  \end{minipage}
\caption{final light distributions for ``isotropy''-test field}
\label{rings-tests}
\end{figure}
Note that we skipped the top center example, since we consider it a trivial add-on example. When we observe these scalar heat map plots, we first notice the red dots, which always represent light sources (since they are captured the most energy) in these cases. Specifically, the heat maps are obtained from the same light distributions we plot as polar profiles and should hence be interpreted concurrently. Remark the propagation front indicating a circular wave (circular isolines in a)).
In this next example, we use the tensor field ``rings'' as input data for the approach and observe how the intensities are channeled in a circular orbit around the central origin. The anisotropy measures decide whether the transmission profile is more narrow (radical) or more wide (liberal):
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/rings-two-special1.png}
    \label{a)}
    \caption*{antisymmtric light src's}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/rings-two-special2.png}
    \label{b)}
    \caption*{symmetric light src's}
  \end{minipage}
\caption{final light propagation distributions for ``rings''-test field}
\label{rings-tests}
\end{figure}
We consider Fig. \ref{rings-tests} to regard the behavior for circular tensor field lines. As expected, the intensities get directed in circular orbits by the tensor ellipsoid transmission profiles. Note the symmetry for both cases depicted. Again, we plot the final light distributions as heat maps:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/cos_parallel.png}
    \label{a)}
    \caption*{antisymmtric light src's}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/cos_anti.png}
    \label{b)}
    \caption*{symmetric light src's}
  \end{minipage}
\caption{final light propagation distributions for ``spiral''-test field}
\label{spiral-tests}
\end{figure}
It is noticeable, that the intensities concentrate preferably on the circular tensor field lines, which proves the approach's controlling ability expressing as direction.
The results for our tensor field ``spiral'' can be interpreted in analogous fashion as the previously discussed ``rings''-testfield. The influence incorporated by the tensor field is easily noticeable in this case as well:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/spiral-two-dense.png}
    \label{a)}
    \caption*{dense light src's}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/spiral-two-wide.png}
    \label{b)}
    \caption*{symmetric light src's}
  \end{minipage}
\caption{final light propagation distributions for ``spiral''-test field}
\label{spiral-tests}
\end{figure}
For the ``spiral'' tensor field we observe the spiral shape literally absorbing the intensities into its spiral arms, which becomes even more clear when we again observe the heat map representation:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/spiral-temp2.png}
    \label{a)}
    \caption*{dense light src's}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.9\textwidth]{img/spiral_full.png}
    \label{b)}
    \caption*{four light src's}
  \end{minipage}
\caption{final light propagation distributions for ``spiral''-test field}
\label{spiral-tests}
\end{figure}

The star example reveals another interesting feature of the approach: if the intensity impulsively hits an isotropic region, it gets directed according to its intensity profile (neglecting the transmission profile) which implements the principle of inertia for directed light propagation present in vacuum (light never changes its direction):
\begin{figure}[!t]
\centering
  \begin{minipage}{0.31\textwidth}
    \includegraphics[width=\textwidth]{img/star-0,0.png}
    \label{a)}
    \caption*{single light src}
  \end{minipage}
  \begin{minipage}{0.31\textwidth}
    \includegraphics[width=\textwidth]{img/star-two-wide.png}
    \label{b)}
    \caption*{symmetric light src's}
  \end{minipage}
   \begin{minipage}{0.31\textwidth}
    \includegraphics[width=\textwidth]{img/star-two-dense.png}
    \label{c)}
    \caption*{dual light src's}
  \end{minipage}
\caption{final light propagation distributions for ``star'' test field}
\label{spiral-tests}
\end{figure}
Remark, that the diagonal intensities are attenuated more strongly by default, which introduces a bias into the light paths which are chosen along the axes with higher propability, effectively. This gets even clearer, when we plot the heat map depiction again:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.31\textwidth}
    \includegraphics[width=\textwidth]{img/star_wide.png}
    \label{a)}
    \caption*{symmetric light src's}
  \end{minipage}
  \begin{minipage}{0.31\textwidth}
    \includegraphics[width=\textwidth]{img/star_close.png}
    \label{b)}
    \caption*{dual light src's}
  \end{minipage}
\caption{final light propagation distributions for ``star'' test field}
\label{spiral-tests}
\end{figure}
Note that we skipped the single light source example, since we consider it a trivial add-on example. We can regard the intensities condensed inside of a narrow beam of light.
Note, that a logarithmic scale needs to be incorporated to produce results on a large scale in a sense of higher spatial resolution (e.g. $101\times 101$), since the propagation attenuation follows the distance to the origin inversely proportional.

\section{Light Transport Gradient (LTG)}
\label{sec:ltg}
\subsection{Evolution}
Now, that we are able to propagate intensity in grid, directed by the eigenvectors of the tensor field, we aim to do something more elaborate to segment ROIs (regions of interest) characterised by divergence/convergence of tensor field lines and drastic changes in anisotropy. For this purpose, we imagine kind of a local and spatial gradient, as we place light sources in two close-by positions and eventually in two close-by directions and analyze the resulting final light distribution. We consider the light distributions and their difference image for the \enquote{isotropic} and the \enquote{drain} test field to have an idea, for which kind or types of cases/configurations we obtain inherently different light distributions, in particular.  We choose to use these test cases, because they constitute absolutely pure examples par excellence for a FTLE computation exposing no (\enquote{isotropic}), or on the other hand raw (\enquote{drain}) tensor field lines diverging. We will start off with the \enquote{isotropic} field from here to first analyze the offset response for no divergence or more specifically, no anisotropy at all. Fig. \ref{analysis1} depicts final light distributions (cf. sect. \ref{sec:ltg} Fig. \ref{rings-tests}) for placing the light source one step to the left and one to the right from the central axis (spatial offset). We can observe the light cones cancel out each other in the center of the difference image. That is, per default, light distributions cancel each other out for two close-by light source positions or directions (or more specifically: for continous overlaps in their distributions). In this case, for the \enquote{isotropic} test field, we have no tensor field lines diverging leading to a small overall difference for light sources in finite distance.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/scalar_iso_left.png}
    \label{a)}
    \caption*{a) final light distribution for left spatial offset}
  \end{minipage} $-$
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/scalar_iso_right.png}
    \label{b)}
    \caption*{b) final light distribution for right spatial offset}
  \end{minipage} $=$
   \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/scalar_iso_diff.png}
    \label{b)}
    \caption*{difference image for a) and b)}
  \end{minipage}
  \caption{``isotropic'' test field}
\label{analysis1}
\end{figure}

The complementing idea for the angular domain would be: placing the light source at the same position but pointing in differing directions, which is depicted in Fig \ref{analysis2}.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/scalar_iso_leftangle.png}
    \label{a)}
    \caption*{a) final light distribution for left angular offset}
  \end{minipage} $-$
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/scalar_iso_rightangle.png}
    \label{b)}
    \caption*{b) final light distribution for right angular offset}
  \end{minipage} $=$
   \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/scalar_iso_anglediff.png}
    \label{b)}
    \caption*{difference image for a) and b)}
  \end{minipage}
  \caption{``isotropic'' test field}
\label{analysis2}
\end{figure}
For the angular offset, we can again observe the light distributions cancelling each other out (per default) in the center (middle) of the difference image. For now, we can take home from these examples, that light distributions are cancelling each other out within their overlap in general, in case no anisotropy was introduced into the test tensor field. We again analyze the drain test field (cf. Fig. \ref{drain} in sect. \ref{chap:grundlagen}) for the behavior of the resulting light distributions in scalar field representation. This following example illustrated in Fig. \ref{evolution1} now introduces raw tensor field lines diverging w.r.t. the $y$-axis.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_left.png}
    \label{a)}
    \caption*{a) final light distribution for left spatial offset}
  \end{minipage} $-$
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_right.png}
    \label{b)}
    \caption*{b) final light distribution for right spatial offset}
  \end{minipage} $=$
   \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_diff.png}
    \label{b)}
    \caption*{difference image for a) and b)}
  \end{minipage}
  \caption{``drain'' test field}
\label{evolution1}
\end{figure}
We can see the light distributions diverge following tensor field lines, which ?n turn form these curves depicted in Fig. \ref{evolution1}. The separation of path (tensor field) lines and thus of light distributions is maximized at the center location of the field. We can now examine the light paths being dragged into long curves, which prevents the light distributions from overlapping at all. Hence, this test case will lead to larger differences and thus to an increased response of the FTLE field, which we will define in this section. 
%Imagine accumulating each pixel in the images to obtain the cumulative total energy in the (difference) grid, which could now be compared for each light source position and direction in terms of proper min-max scaling.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_leftangle.png}
    \label{a)}
    \caption*{a) final light distribution for left angular offset}
  \end{minipage} $-$
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_rightangle.png}
    \label{b)}
    \caption*{b) final light distribution for right angular offset}
  \end{minipage} $=$
   \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_anglediff.png}
    \label{b)}
    \caption*{difference image for a) and b)}
  \end{minipage}
  \caption{``drain'' test field}
\label{evolution2}
\end{figure}
For the angle, we can observe again that the difference is again increased, but does not 
Imagine accumulating each pixel in the images to obtain the cumulative total energy in the (difference) grid, which could now be compared for each light source position and direction in terms of proper min-max scaling. Now imagine we would accumulate these differences and assign each position a gradient magnitude in form of a scalar value with the summed difference of the final light distributions in the neighborhood. What would we eventually be yielded with. Well, we would obtain another scalar field considered to be the gradient image of the light distributions. We could very well extend this process for every light source direction discretized (sampled) and obtain a $3D$ volume representing the topological structure of the tensor field for every direction. All we need, is a formal definition for a central differences gradient approximation for light distributions given in polar profiles. The main difference to FTLE in vector field visualization is, that we use whole distributions of trajectories and hence need to use a starting direction for the distribution of trajectories, instead of just moving with the flow, as a tensor introduces a whole distribution of directions and does not only depend on the inital position, but also on the direction. This corresponding to the volume emerging from the FTLE plane image, as we introduce a new dimension $\omega$.

Remark that we denote the final buffer result as $\mathbf{b}=\mathcal{C}_{final}$ in the following.
\subsection{Definition}
 As tensor field lines are bidirectional, convergence and divergence can not be reasonably discerned in that case. We aim to extract ridges from a FTLE-like field to grasp the key regions separating anisotropy behavior (, i.e., changes in direction/tensor magnitude) in the field segmented as height ridges in the field, which are in turn LCS (Lagrangian Coherent Structures) well known for their rich substance in physical meaning. They often indicate transport barriers in flow fields. The nature and location of these structures and their interaction with each other is subject to the study of both turbulent and laminar mixing. They can even be observed in nature for, e.g., vortices in flow fields. For this purpose, we will use our, previously defined, $\mathop{propagate}$-operator, i.e., the propagation scheme from sect. \ref{sec:scheme} and \ref{sec:trans}, on every possible light source location $\mathbf{x}$ and direction $\omega$ in the grid to obtain the final set of buffers $\mathcal{B} = \{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_j\}$ of impulse responses of the algorithm. That is, we initialize the grid with every possible light source configuration and compute a whole final light distribution for each step:
\begin{align}
	\mathcal{B} = \mathop{propagate}\{c_i\in \mathcal{C} \ \mid \ |\mean{I}_i| > 0\} \ {\forall } \ i \in [1,\mathop{dim}] \ {\forall } \ j \in [1,\mathop{dim}\cdot \mathop{steps}].
	\label{ftle}
\end{align}
Whereas $j$ is the index of light source position and direction representing the initialization in each iteration and $\mathop{steps}$ is the count of the discrete angular steps. The initialization is done by placing a light src (delta-impulse) with $\mean{I}_i = 1.0$ at the exact location of index $j$ realizing a light source at location $(x\mid y)$ in direction $\omega$. Eq. \ref{ftle} now denotes the set of buffers $\mathcal{B}$ containing light distributions propagated from, i.e., computed for every possible direction and location. Now, that we have measured impulse responses from every location in our long $1D$ sample vector, we can incorporate finite (central) differences to obtain the gradient of the set of final buffers with final light distributions as elements $\mathcal{C}_{final}=\mathbf{b}\in\mathcal{B}$. Note, that we interpret a finite set $\mathcal{C}_{final}$ as a vector of cells $\mathbf{b}$ with concurrent indices here:
\begin{align}
	\nabla \mathbf{b}_{x,y,\omega} =
	\begin{pmatrix}
	\frac{\partial \mathbf{b}_{x,y,\omega}}{\partial x} \\ \frac{\partial \mathbf{b}_{x,y,\omega}}{\partial y} \\ \frac{\partial \mathbf{b}_{x,y,\omega}}{\partial \omega}
	\end{pmatrix} \approx
	\frac{1}{2}
	\begin{pmatrix}
	\sum_{\mathbf{c}}\lvert \mathbf{b}_{x+1,y,\omega} - \mathbf{b}_{x-1,y,\omega}\rvert \\ \sum_{\mathbf{c}}\lvert \mathbf{b}_{x,y+1,\omega} - \mathbf{b}_{x,y-1,\omega}\rvert \\ \sum_{\mathbf{c}}\lvert \mathbf{b}_{x,y,\omega+\pi/6} - \mathbf{b}_{x,y,\omega-\pi/6}\rvert
	\end{pmatrix} =
	\frac{1}{2}
	\begin{pmatrix}
	\sum_{\mathbf{c}\in \mathbf{b_x}}\lvert \mathbf{b_x}\rvert \\ \sum_{\mathbf{c}\in \mathbf{b_y}}\lvert\mathbf{b_{y}}\rvert \\ \sum_{\mathbf{c}\in \mathbf{b_{\omega}}}\lvert \mathbf{b_{\omega}}\rvert
	\end{pmatrix}.
\end{align}
 We choose to use a constant offset of $\frac{\pi}{6}\mathop{rad}$ for the angle because we want to prevent from generating results with content implicitly depending on the angular resolution and also since we need to exceed $\beta$ to have a uniform and consistent placing of cosine lobes (inside diagonal cones) concerning neighbor indices. Remark that the results do not depend on the spatial resolution implicitly, because increasement acts as a domain extension for and does effectively not increase the sampling rate of the propagation scheme. Also note that this step is very expensive considering runtime-costs since it follows the following runtime observation. We have $\mathop{width}\cdot \mathop{height}\cdot \mathop{steps}$ numbers in the grid and on top $\mathop{width}\cdot \mathop{height}\cdot \mathop{steps}$ buffers in total, which need to be processed. Also we remark the convergence criterion, which introduces another heuristic factor $\mathop{width}$, because the straight line distance to the field edges is a measure for eventual compliance, and increases with $d\sim width$ for any location in the grid. We set $\mathop{height}=\mathop{width}$ here for simplicity, which yields:
\begin{align}
	T(n)=\mathcal{O}(\mathop{width}\mathop{width}\mathop{steps}\mathop{width}\mathop{width}\mathop{steps}\mathop{width}) = \mathcal{O}(\mathop{width^5}\mathop{steps^2}),
\end{align}
which is the runtime behavior in order notation considering memory consumption and iteration step count. It is noticeable that the runtime increases fast with increasing resolution, which makes parallelization algorithms a crucial tool necessary to compute the results in reasonable time with additional computing power provided by a computer cluster. Next, we compute the Euclidean norm of the light transport gradient yielding a scalar field, which can be visualized directly or via extracted ridges in a visualization framework like ParaView from VTK format:
\begin{align}
	\lvert\nabla \mathbf{b}_{x,y,\omega}\rvert = \sqrt{(\frac{\partial \mathbf{b}_{x,y,\omega}}{\partial x})^2 + (\frac{\partial \mathbf{b}_{x,y,\omega}}{\partial y})^2 + (\frac{\partial \mathbf{b}_{x,y,\omega}}{\partial \omega})^2}.
\end{align}
Note that the resulting scalar field is $(n+1)$-dimensional ($x,y$ and direction $\omega\mapsto  z$) and needs to be flattened by averaging or projection in 3D for proper visualization in case the approach will be adopted in $3D$. We note, that our invented technique stands out compared to the already existent FTLE on tensor fields invented by Hlawatsch et al. \cite{hlawatsch}, as we use whole distributions of directions instead of following the major eigenvector strictly. These directional distributions allow for a complete analysis of the tensor field accounting for the incoming direction, as well.

\subsection{Functionality}
Remark, that the $z$-axis represents the $\omega$-axis for each test and example and that each data range is rescaled to its interval for proper color coding in heat maps of the resulting $3D$ volume. We will try to show the functionality of the method in a few simple examples. At first, we will have a null example to show the offset response of the approach induced by total spherical anisotropy and no absorption in the field. We consider the example \enquote{isotropic}, concerned to have no influence on the light propagation, as it models homogeneous identity matrices inducing absolutely isotropic transmission (every incident distribution is transmitted as is with no transmission weighting incorporated). This corresponds to the \enquote{non - tensor field} mode of the propagation scheme having no influence of any tensor field incorporated. In this configuration, we could eventually visualize light transport paths in a $2D$ topological scene.

\begin{figure}[t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[height=\textwidth]{img/isotropic2.png}
    \label{a)}
    \caption*{a) Glyphs}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[height=\textwidth]{img/iso_ftle.png}
    \label{b)}
    \caption*{b) LTG($0^\circ$) for a)}
  \end{minipage}
  \caption{\enquote{isotropic} test field}
\label{ftle_iso}
\end{figure}

When we look at the FTLE field, on the top right in Fig. \ref{ftle_iso}, we determine that we obtain a more or less uniform reponse (, if we would neglect the edges) just as expected. Because the divergence of TFLs is distributed uniformly in this test field, we get approximately the same response for any position, but yet introducing a systematic bias against the light source direction. This is explained by the fact, that the energy will leave the grid sooner on light source positions at the edges for light source directions pointing toward the edges leading to inherently decreased differences and hence gradient magnitudes (responses). Note, that this also causes a shift of the FTLE field to the left, as the light source points to the right ($0^\circ$ slice). As noticeable, no height ridges are detected in this case as expected, since the test field does not pose any specific regions exhibiting diverging tensor field lines. We consider this result the absolute offset response, whereas energies are redistributed and separated more strictly for introducing local features into the grid (field), leading to greater differences, which are in turn emphasized by local regions of interest displayed in deep red at their corresponding positions. Imagine this process comparable with ink absorbed by dry paper. As we are also interested in the whole ensemble of directions, we can use a volume or surface representation for the $n+1=3$-dimensional grid which results from our FTLE computation. The volume grid is what we call the final footprint of the tensor field which is necessarily derived from the initial glyph footprint. From it, we can observe the behavior of the tensor field for any direction. That is, the slices we observe are stacked along the $\omega\mapsto z$-axis in the following. As a subsequent step, we segment height ridges in $3D$ as thresholded volumes from the whole volume grid.

\begin{figure}[t]
\centering
  \begin{minipage}{0.4\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{img/iso_volume.png}
    \label{a)}
    \caption*{a) Volume}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.8\textwidth]{img/iso_isovolume.png}
    \label{b)}
    \caption*{b) IsoVolume Filter for a)}
  \end{minipage}
  \caption{\enquote{isotropic} test field}
\label{ftle_iso}
\end{figure}
To verify the concept of absorption implicitly, we incorporate absorption into test field \enquote{absorb}, whereas all cells are absorbing energy except the vertical center line, which is assumed to produce a clear ridge. 
\begin{figure}[t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/absorb.png}
    \label{a)}
    \caption*{a) Glyphs}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.9\textwidth]{img/absorb_ftle.png}
    \label{b)}
    \caption*{b) LTG($90^\circ$) for a)}
  \end{minipage}
  \caption{\enquote{absorb} test field}
\label{ftle_absorb}
\end{figure}

We observe the clear straight ridge in the example in Fig. \ref{ftle_absorb}. As expected, the ridge is located on the vertivcal center line. Note, that we (min-max) normalize each transmission profile with the highest mean transmission rate (tensor magnitude) in the field. From this test case, we can conclude, that regions with a comparably small tensor magnitude would not obtain any FTLE field response, which we will later observe for the real datasets \enquote{brain} and \enquote{heart} exhibiting such areas. We will now discuss synthetic examples designed to evaluate the performance of the FTLE. As a little warmup, we will have an experiment exhibting bare, raw tensor field lines diverging w.r.t. the $y$-axis:

\begin{figure}[t]
\centering
  \begin{minipage}{0.3\textwidth}
\centering
    \includegraphics[width=\textwidth]{img/drain(alt)-TFL.png}
    \label{a)}
    \caption*{a) TFLs and glyphs}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_drain_alt.png}
    \label{b)}
    \caption*{b) LTG($90^\circ$) for a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/drain_alt_gradient.png}
    \label{b)}
    \caption*{c) gradient for b)}
  \end{minipage}
  \caption{``drain'' test field}
\label{ftle_base}
\end{figure}
What we can observe here, is a feature called a ``ridge'' occuring as a high value line segment (red vertical track) which is already quite obvious for human vision. A gradient filter could be applied to extract and segment edges from the image as a preprocessing step for computer vision, which would look like the image in Fig. \ref{ftle_base} c).

These ridges represent ROIs with underlying Lagrangian Coherent Structures, which we already discussed in sect. \ref{sect:motiv}. It could for example represent a stress field inside of a mechanical tool with a central axis with maximum stress condensation. 
\begin{figure}[ht]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
  \includegraphics[width=0.9\textwidth]{img/drain_top.PNG}\vskip 10pt
    \includegraphics[width=0.9\textwidth]{img/drain_alt-profile.PNG}
    \caption*{Cross section on LTG volume}	
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/drain_alt_isovolume.PNG}
     \caption*{IsoVolume Filter}
    \label{b)}
  \end{minipage}
  \caption{LTG volume representation: \enquote{drain} test field}
\label{drain_contour}
\end{figure}
For the ``drain'' test field we consider the central channel which produces a clear line in the cross section. The light transport gradient responds most inside of the channel region and on the top T-arms independently of the stimulus direction. For the sake of completeness, we show that the approach's functionality also for diagonal directions in Fig. \ref{diag-ftle}.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/diagDrain.png}
    \label{a)}
    \caption*{a) TFLs and glyphs}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/diagDrain-ftle.png}
    \label{b)}
    \caption*{b) LTG($225^\circ$) for a)}
  \end{minipage}
  \caption{``diagonal drain'' test field}
\label{diag-ftle}
\end{figure}
Regarding the ridge for the \enquote{drain} test field, we note that it is a little smoothed compared to the axis-aligned ridges, as they obtain a larger propagation attenuation (because of the larger distance) leading to a stronger separation of trajectories in general and hence to a smoothed FTLE response. Also we note that the divergence of the tensor field lines is reduced compared to the classic \enquote{drain} example of the FTLE.

\begin{figure}[ht]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/inverse-TFL.png}
    \label{a)}
    \caption*{a) TFLs and glyphs }
  \end{minipage}
  \begin{minipage}[!t]{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/inverse-FTLE-raw.png}
    \label{b)}
    \caption*{b) LTG($90^\circ$) for a)}
  \end{minipage}
  \caption{``inverse''-test field}
\label{inverse-ftle}
\end{figure}

We also evaluated the approach for the ``inverse'' test field, depicted in Fig. \ref{inverse-ftle}, whereas we can observe height ridges for the $90^\circ$ slice in any position with diverging tensor field lines w.r.t. the light source direction. This is certainly not the case for the upper (missing) ridge whereas tensor field lines converge with respect to the UP ($90^\circ$) direction. If we determine the ridges pointing into the diverging direction, this proves that the method is capable of detecting ridges from parallel up to orthogonal to the light source direction, but not in antiparallel directions, which exhibit converging tensor field lines as bidirectional counterpart. Luckily, we measured the gradient for every possible sampled light source direction which complements this information into another slice (and the resulting volume of stacked slices).

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/inverse_isovolume.png}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/inverse_surface.png}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/inverse_contour.png}
  \end{minipage}
    \caption{IsoVolume Filter on LTG volume representation: ``inverse'' test field}
\label{inverse_contour}
\end{figure}
Note that the $z$-axis represents the light source direction angle $\omega$ in this case. The contour filter reveals the star-shaped structure of the tensor field which rotates along the angle axis. This is the effect we already captured for the single slice: For each axis direction we can detect ridges from parallel up to orthogonal to our predefined stimulus (light source) direction, but not in antiparallel direction. This causes the star shape to miss one branch or ray which is in turn the antiparallel direction missing for each light source direction, emphasizing the impression of rotation. We can also try to segment height ridges represented as surfaces from the resulting volume grids as depicted in Fig \ref{inverse_contour}. The orthogonal ridges in this example reach out to the edges because they are gaining a bias by never offsetting the light source in a transmission direction (ellipsoid orientation) for orthogonals. This turns out to be a general feature of the method, as we will later verify in Fig. \ref{res_ftle_volume} and \ref{star-ftle}.

%\begin{figure}[!t]
%\centering
%  \begin{minipage}{0.4\textwidth}
%    \includegraphics[height=\textwidth]{img/inverse_surface.png}
%    \label{a)}
%  \end{minipage}
%  \begin{minipage}{0.4\textwidth}
%    \includegraphics[height=\textwidth]{img/inverse-wireframe.PNG}
%    \label{b)}
%  \end{minipage}
%    \caption{LTG stacked slice surface representations: ``inverse'' test field}
%\label{inverse_surface}
%\end{figure}

For the ``inverse'' test field, we can observe a 3 part ``rotor'' indicating the ridges which are in parallel up to orthogonal to the current light source (slice) direction. This rotor can be obtained by applying a contour filter, but it can also be visualized by applying a simple thresholding on the surface representation of the volume, as depicted in Fig. \ref{inverse_surface}. Subsequently, we will analyze the \enquote{rings} test field to show its performance on circular ridges.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/tensorfieldlines.png}
    \label{a)}
    \caption*{a) TFLs and glyphs }
  \end{minipage}
  \begin{minipage}[!t]{0.4\textwidth}
  \centering
    \includegraphics[width=0.9\textwidth]{img/rings_ftle.png}
    \label{b)}
    \caption*{b) LTG($0^\circ$) for a)}
  \end{minipage}
  \caption{``rings''-test field}
\label{rings-ftle}
\end{figure}
What we can observe in Fig. \ref{rings-ftle}, is called an inside-outside test producing the ridge on the circumference of the circle with maximized radius fitting into the minimum bounding box. If one would place a light source on the circumference of this circle offsetting it spatially and in the angular domain, one would obtain one incomplete final light distribution (outside) and one complete final light distribution (inside), as this is the maximum radius at which the intensities can follow a circle. This hence leads to large differential light distributions and thus to a higher LTG magnitude. ROIs pointing outside the field respond the same for varying light source directions and positions yielding the blue areas in the corners. Regions introducing linear anisotropy respond uniformly inside of the circle in contrary to the center point, which is again not responding because of circular overlaps of the offset light distributions computed for the light transport gradient. The contour filter reveals these two edge locations in the field, concerning tensor field line divergence.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/rings_contour.png}
    \caption*{Contour}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/rings_isovolume.png}
    \caption*{IsoVolume}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/rings_volume.png}
    \caption*{Volume}
  \end{minipage}
    \caption{LTG volume representations: ``rings'' test field}
\label{rings_contour}
\end{figure}
The volumes, we examine here in Fig. \ref{rings_contour}, show that the field is symmetric, as we do not notice a change along the $\omega\mapsto z$-axis except for the natural shift of the field against the light source direction, which we discussed already in this subsection. Along the $z$-axis, the light source is rotated in all directions which causes the shift to form a full cycle from bottom to top ($[0^\circ,360^\circ]$).
At last, we will analyze the bow test field, which is now considered as a more complicated example. We again plot the FTLE field in color coding representation in Fig. \ref{bow-ftle}.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=\textwidth]{img/bow-TFL.png}
    \label{a)}
    \caption*{a) TFLs and glyphs }
  \end{minipage}
  \begin{minipage}[!t]{0.4\textwidth}
  \centering
    \includegraphics[width=\textwidth]{img/bow_ftle.png}
    \label{b)}
    \caption*{b) LTG($0^\circ$) for a)}
  \end{minipage}
  \caption{``bow''-test field}
\label{bow-ftle}
\end{figure}
One may observe here, that the ridges located against the direction of the light source are again visible more clearly for the $0^\circ$ slice (directed bias). These ridges represent regions revealing most diverging tensor field lines in the original field. The ridges on the left side are detected because they are in the range up to orthogonal to the light source direction (if we consider them pointing into the diverging direction).

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=\textwidth]{img/bow_contour.png}
    \caption*{Contour}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
  \centering
    \includegraphics[width=\textwidth]{img/bow_volume.png}
    \caption*{IsoVolume}
  \end{minipage}
    \caption{LTG volume representations: ``bow'' test field}
\label{bow_contour}
\end{figure}
Considering the volumes depicted in Fig. \ref{bow_contour} one may notice, that we obtain a riffled structure reminding of ground plates. On half the volume height ($180^\circ$) the structure is turned by $180^\circ$ as expected as this light source configuration lights up the mirror image to the one on the surface. The bow is literally shaped from the volume representation if one applies a threshold.

If we relate the resulting FTLE volume to the work of Hlawatsch et al. \cite{hlawatsch}, we find that their FTLE field is contained in our volume in form of a manifold, since they use the same basic idea, but consider only a single direction (of the major eigenvector), instead of taking all directions into account, as we do here. The resulting manifold included, is constrained to a height field (not self-overlapping or self-intersecting), because each position can only represent one individual direction, as there is only one direction for the major eigenvector at each position to be covered (, since it is obviously a definite, distinct measure). This direction emerges as a single scalar height (FTLE) value on each position corresponding to a height (direction) slice in our volume. That is, each position in their $2D$ scalar field is assigned a height in our $3D$ scalar field. We cover their results as a subset of our volume, which could be, e.g., a spiral plane in case all directions are given in circular arrangement in the underlying tensor field (cf. \enquote{rings} test field). If we would vary the angle of the major eigenvector with a uniform distribution along the tensor field's width, we would obtain a primitive diagonal plane w.r.t. the $xy$-plane reaching out in $z$-direction as contained subset. Therefore, we claim and consider our invented technique to be more generalized (to all directions) resulting in a topological $n+1$-D volume.
\chapter{Implementation}
\label{chap:imp}
In this chapter, we will describe the implementation process from top to bottom following runtime chronology. At first, we will load a configuration file setting angular resolution in (discrete) steps, which are spaced uniformly over the interval $[0,2\pi]$. The threshold for the convergence criterion was empirically determined and set in the range $[0.5,2.0]$. Since it is an energetic threshold, it is only dependent on the light source magnitude and not on any type of resolutions. An iteration limit is set to $\mathop{limit}=10\mathop{width}$ (e.g. for $\mathop{width}=10\cdot 100=1000$)  to prevent the approach from eventual oscillation behavior for intensities and hence overpropagation (propagating for too many timesteps). The spatial resolution is inferred by the tensor field reader method from the tensor field (see Propagation below). The light source profile can be defined for the \enquote{non-FTLE} mode where the user sets position and polar profile of single or multiple light sources as a symbolic term: ,e.g. , \texttt{clip(cos(theta-pi/2))} for a Lambert-emitter pointing upwards. After the configuration read process is finished, we allocate memory for certain buffers holding glyph and intensity information with a dimension of $\mathop{dim}=\mathop{width}\mathop{height}\mathop{steps}$. This $1D$ buffer is interpreted like this:

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.5\textwidth]
    {img/grid.pdf}
  }
  \caption{Indices and Dimensions in Grid}
  \label{indices}
\end{figure}\vskip 3pt
Fig. \ref{indices} depicts the \textit{deltaIndex}-map for computing the destination index for the scaled cosine lobes for cell $c$. The size of each cell is $\mathit{size}=\mathit{steps}$ in a $1D$ buffer arrangement, respectively.
We then initialize energetic sums and thresholds and run a while loop until the convergence criterion is met. We compute the difference $\Delta\Phi_{total}$ of the energetic sum in this iteration with the one from the last. If it falls short of this threshold, we initiate a stop sequence and leave the loop. To compute the FTLE, we use multiple stimuli in parallel (,e.g., Dirac-pulses within the scope of this work), and just compute the differences in final light distributions for two close-by light source positions and directions to accumulate them in a total energy sum (capacity), which in turn forms one component of the light transport gradient. A Euclidean norm yields a scalar field from the gradient at each position.

\begin{algorithm}
\caption{Propagate Light Distribution}\label{PLD}
\begin{algorithmic}[1]
\Procedure{propagateDist}{i,j,$\omega$}\Comment{Input position $j,i$ and direction $\omega$}
\State \Call{fill}{\textit{bufferA}, $0.0$} \Comment{Reset and Initialize}
\State \Call{fill}{\textit{bufferB}, $0.0$}
\State $\mathit{sumMem} \gets 0.0$
\State $\mathit{finished} \gets \mathit{false}$
\State $\mathit{index} \gets (j\cdot\mathit{width}+i)$ \Comment{compute $1D$ (cell) index}
\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{write light src }
\While{$\Delta\Phi_{total} < \mathit{\epsilon }$ } \Comment{while convergence criterion not met..}
\State $\mathit{sumA} \gets 0.0$ \Comment{reset sum}
\State \Call{propagate}{\null} \Comment{propagate \textit{bufferA} (src) in \textit{bufferB} (tar)}
\State $\mathit{sumA} \gets$\Call{sum}{\textit{bufferB}}  \Comment{sum up energies}
\State $\Delta\Phi_{total} \gets \lvert\mathit{sumA}-\mathit{sumMem}\rvert $\Comment{compute difference to prev. iter.}
\State $\mathit{sumMem} \gets \mathit{sumA}$\Comment{save sum for next iter.}
\State \Call{swap}{\textit{bufferA}, \textit{bufferB}} \Comment{swap buffers for restart}
\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{re-write light src}
\State \Call{fill}{\textit{bufferB}, $0.0$}
\If {$\mathit{ctr}>\mathit{limit}$}\Comment{stop on iteration limit}
\State break
\EndIf
\EndWhile\label{euclidendwhile} \Comment{final light distribution stored in bufferA..}
\State return \textit{bufferA}
\EndProcedure
\end{algorithmic}
\end{algorithm}
Note that we set the light source intensity to constant value $\mathit{steps}$ to require a mean intensity of $\mu_I=1.0$ comparable to the one of a unit circle independently on the angular resolution.

When we consider the LTG method, we already stated its runtime behavior ($\mathcal{O}(\mathop{width^5}\mathop{steps^2})$), which makes parallelization libraries and hardware to compute the results in reasonable time. First, we tried to implement a GPU-based parallelization for the propagation scheme, which turned out not to work as expected, since the relative runtime gain was too low make a big difference. At this point, we choose to use OpenMP as we figured it is a much more suitable architecture employed here, which allows to perform the computation on a multiprocessor architecture provided by the Heidelberg University allowing to compute $64$ hardware threads at maximum. As we need fat fed, instruction-rich threads to have a high performance gain, we choose to parallelize only the outer loop of the LTG method (propagating final light distributions for each light source direction). Remark, that each thread requires its own memory, as an independent memory block is needed to perform the propagation in an individual propagator object of class propagator, in order to perform the computation in parallel.

\paragraph{Propagation}
The method performing the light propagation, was implemented in a separate class called propagator, which we can initialize once for each programming thread (, which we can in turn provide with its own buffer memory). As an initialization we fill the buffers with zeros to have a blank buffer initially to fill up with meaningful intensity profiles. The transmission profiles are generated by loading the tensor field in memory, performing PCA on each of them and exploiting the thin $\mathbf{U}$-matrix for the first (greater) singular vector. The resulting principal component ellipsoid equation is precomputed, sampled and stored for reuse once for each cell.  We again sample cosines into each direction, average out their integrals (sums) and normalize each of them with the mean to have the integral energy scaled to $1$ (identity element for multiplication, as it is the case for \enquote{scaling the cosine lobes} here). Thus we ensure, that the transmitted energy is approximately $1.0=100\%$ for each lobe and account for angular sampling errors.  Now we need to precompute and store the linear combinations used for the \enquote{shared part} in diagonal cones. When hitting border cells, we propagate the intensities only to the present neighbor directions still located on the grid. Remark, that we can not just ignore the edges of the grid because it would lead to large trivial intensity differences on grid borders for the FTLE-similar approach LTG.

\begin{algorithm}
\caption{Propagate BufferA to BufferB}\label{PB}
\begin{algorithmic}[1]
\Procedure{propagate}{}
 \For{each cell $c$ in bufferA}
\State $\mu_{I_c}\gets mean(I_c(\omega))$ \Comment{Compute Mean Intensity}
\If {$\mu_{I_c}=0.0$}\Comment{break on trivial null sample}
\State break
\EndIf
\State $\mu_{T_c}\gets mean(T_c(\omega))$ \Comment{Comp. Mean Transmission}
\State $\mu_{\mathit{T_cI_c}}\gets mean(T_c(\omega)\cdot I_c(\omega))$ \Comment{Comp. Mean Transmitted Intensity}
\State $n_c \gets \frac{\mu_{T_c}\mu_{I_c}}{\mu_{T_cI_c}}$\Comment{Comp. normalization factor for cur. cell}
 \For{each direction $k$ in $[0,7]$}\Comment{For all neighbors, do..}
 	\State $\gamma \gets k\frac{\pi}{4}$\Comment{Offset to compute cone center angle}
 	\If {$k\bmod 2 = 0$}\Comment{if even (face) neighbor..}
	\State $\mathit{energy} = \int_{\gamma-\pi/4}^{\gamma+\pi/4}n_c\epsilon_\alpha T_c(\omega)I_c(\omega)\mathop{d\omega}$
	\Else\Comment{if odd (diagonal) neighbor..}
	\State $\mathit{energy} = \int_{\gamma-\beta}^{\gamma+\beta}n_c\epsilon_\beta T_c(\omega)I_c(\omega)\mathop{d\omega}$
	\EndIf
	\State $\mathit{index} \gets c+\mathit{deltaIndex}(k)$\Comment{assign neighbor destination index}
	\State $\mathit{bufferB(index,\omega)}\gets \mathit{energy}\cdot cos_k(\omega)+\mathit{bufferB(index,\omega)}$\\
	\Comment{Scale cosine lobe with summed energy and accumulate}\\
	\Comment{$y=a*x+y\rightarrow$ axpy-OP}
 \EndFor
 \EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

\paragraph{Languages and Libraries}
At first, we needed to generate convenient test cases, which we design using Mathematicas VectorPlot, used to visualize vector fields. To output a pre-defined test case, we need to define a little code, which we do using python scripts which do output VTK appended or ASCII format files. To implement the method we used C++'s STL for every dynamic memory (buffer), SFML to plot polar profiles, Eigen to perform matrix decompositions and VTK to in- and output VTK format files, which can in turn be visualized and explored interactively with the visualization framework ParaView. The user functions defined in the configuration file are read-in by a string parser (muParser Library), which is used just once for sampling the polar profile and storing the precomputed result. This causes a convenient input mechanism for stimuli, in case they should be predefined.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Results and Evaluation}
\label{chap:eval}
\section{Propagation Attenuation}
\label{sec:att}
It can be proven that the approach follows propagation attenuation principles and follows the inverse (square in $3D$) law ($\sim\frac{1}{r}$) as observed in the graph in Fig. \ref{att}. For that, we plot the relative intensity $i_{rel}(r_{rel})$ (in relation to the intensity at position (radius) $1$ to account for the inverse square law, which holds for relative distances. We evaluate the relative intensity by forming a ratio $\frac{I(r)}{I_1}$, that relates the step intensity to the src intensity at position 1.

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/inverse_law.pdf}
  }
  \caption{Propagation Attenuation}
  \label{att}
\end{figure}\vskip 3pt

\section{Behaviour for Total Anisotropy}
\label{sect:total}
We also wanted to check for the limitations of the approach by having total linear anisotropy (the case whereas one eigenvalue $\lambda_1>0$ and the other $\lambda_2=0$) incorporated. For this we set up bidirectional Dirac-pulses as transmission profiles to have the whole energy transmitted in just one, discrete direction.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/tensorfieldlines.png}
    \caption*{TFLs and glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/total_anisotropy1.png}
    \caption*{iteration $50$}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/total_anisotropy2.png}
    \caption*{iteration $103$}
    \label{c)}
  \end{minipage}
\caption{intensity propagation in grid for total anisotropy}
\label{anisotropy}
\end{figure}
What we can observe here, is kind of a sampling drift as the approach uses 8 discrete directions to propagate averaged intensity whereas the circle uses much more (namely infinite if presumed as perfect circle). Hence, the energy can never follow a perfect circle, but at best an octagon. This means, that there is a drift of at least two cells per diagonal (one overshoot per enter and escape each) as observable in Fig. \ref{anisotropy}. Nonetheless, the energies are transmitted in circular orbits around the circle center.	
\section{Energy Conservation Test}
\label{sect:conserve}
The energy conservation test was done in similar manner as thepropagation attenuation test in sect. \ref{sec:att}. We measure the impulse response of the algorithm for 80 iterations by summing the total energy in the grid after each iteration and comparing it to the initial energy sum (e.g. $\int_0^{2\pi}1\mathop{d\omega}=2\pi$) in iteration 0 in absolute values. This energetic error analysis yields the following graph depicted in Fig. \ref{energy}:

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/sum_difference.pdf}
  }
  \caption{Sum Error: src intensities: red: 1.0, blue 2.0, green: 5.0}
  \label{energy}
\end{figure}

The approach remains stable until iteration 40, which is at the time (point) when the numbers are too small and get rounded down under the double machine epsilon. You can see heavy fluctuations because the numbers get rounded upwards likewise. But what is lost, is lost and can not be regained. The error is thus accumulating gradually and with faster slope for higher source intensities, since floting point precision, i.e., rounding errors operate more heavily on large numbers since it is a constant relative loss $\sim\frac{1}{r}$ leading to a greater absolute $\Delta E$.

\section{Light Transport Gradient (LTG)}
The light transport gradient approach, which is closely related to the FTLE approach in vector fields, is evaluated for some more synthetic and real datasets in this section. We will start off with the comparison of our FTLE-based approach to the one invented by Hlawatsch et al. \cite{hlawatsch}, which we reimplemented for verification purposes.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=0.9\textwidth]{img/ftle_drain_alt.png}
    \caption*{LTG($0^\circ$)}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.9\textwidth]{img/drain_ftle_org_real.png}
     \caption*{FTLE: Hlawatsch's approach}
    \label{b)}
  \end{minipage}
\caption{\enquote{drain} test field}
\label{drain_ftle_comp}
\end{figure}
We can observe the comparison between our invented FTLE-like approach and the one of Hlawatsch et al. in the depiction in Fig. \ref{drain_ftle_comp}. Note, that a sharper but weaker ridge is detected on the right side of the image, which shines similar to a cone shape from the top. We interpret the missing spot in the top center as an artifact stemming from the way we handle our light propagation, as we allow whole distributions of trajectories to be propagated (with a scalar attribute interpretable as propability). This causes a sampling drift, we already discussed in sect. \ref{sect:total}, leading to an energy distribution ending up outside of the field, for this position. The directive effect turns out to be not that strict and delayed compared to the classic tensor field line FTLE. The crucial difference is that the compared work of Hlawatsch et al. obtains the FTLE field from the gradient $\nabla\mathbf{\upvarphi}(\mathbf{x})$ of the flow map, which is in turn obtained from sampling of trajectories following tensor field lines.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.29\textwidth}
  \includegraphics[height=\textwidth]{img/star.png}
    \caption*{\enquote{Star} test field}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.29\textwidth}
    \includegraphics[height=\textwidth]{img/star_ftle.PNG}
     \caption*{LTG slice($90^\circ$)}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.35\textwidth}
  \centering
    \includegraphics[height=\textwidth]{img/star_volume_alt.PNG}
     \caption*{IsoVolume on LTG volume}
    \label{b)}
  \end{minipage}
  \caption{LTG on \enquote{star} test field}
\label{star-ftle}
\end{figure}
When we look at the example depicted in Fig. \ref{star-ftle}, we can see the FTLE approach LTG responding most in the center of the field, which is the knot point of tensor field lines diverging in any direction. Remark, that in the middle image there is a light V-shaped ridge on the left on blue background which is outshined by the center spot because of its orders of larger magnitude. The volume illustrated on the right even shows a spiked structure for the volume representing all directions at once. The approach detects more ridges in diagonal directions stochastically as we can infer from the Gaussian-shaped plates in the thresholded volume, evenly spaced in $90^\circ$ shifts with $45^\circ$ (quarter volume) offset. These can be interpreted nicely, since the diagonal direction is never offset into its transmission direction leading to a trivial (none) difference for axis-parallel directions for the offset direction orthogonal to the light source direction. Hence, the differences will be factor $2$ at worst for the diagonals, since we measure differences in both offset directions, in case ridges in both axis-parallel and -diagonal directions are present. Accordingly, we observe Gaussian-plates in the volume instead of sharp, contrasted ridges for the \enquote{star} example. But wouldn't we expect a ridges only in parallel directions? Well, in this case, we could imagine not only tensor field lines diverge or converge, but also they could cross each other unaffectedly (just as with light paths), which could be the case when shifting away from the diagonal directions. For better imagination, offset the light source vertically or horizontally from the diagonals and observe the light paths hitting the axes (, which in fact do cross each other).

%A possible correction would be, to normalize each slice with a heuristic factor $\frac{1}{1+(\sqrt{2}-1)\lvert\cos2\omega-\frac{\pi}{4}\rvert}$ to get a uniform response over all directions and account for the longer diagonal run lengths (SLDs).

 We will now capture the star dataset revealing tensor field lines diverging in any direction from the center.
Now, we will evaluate a classic textbook example called the \enquote{double gyre} \cite{shawn}, which is a quadruple gyre in our case for formatting reasons, yet yielding the same conceptual results. We compare the glyph representation to the LTG magnitude for a single slice ($0^\circ$) again:
\begin{figure}[!t]
\centering
  \begin{minipage}[t]{0.3\textwidth}
  \includegraphics[height=\textwidth]{img/gyre.png}
    \caption*{a) Glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}[t]{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle.PNG}
     \caption*{b) LTG}
    \label{b)}
  \end{minipage}
  \begin{minipage}[t]{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle_org.PNG}
     \caption*{c) FTLE: Hlawatsch's approach}
    \label{b)}
  \end{minipage}
  \caption{Gyre test field}
\label{gyre-ftle}
\end{figure}
We can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t. the $x$-axis ($0^\circ$). Note that even regions in antiparallel directions ridgres are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging. Again, it is noticeable, that the results correspond to each other to a certain degree. This implies, that we obtained reasonable results generated by our method, but the classical tensor field line FTLE is more strict concerning its directive abilities and congruently does not do things by halves (like we do with our partitioned light distributions). Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure. We again depict volume representations for this example:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
  \includegraphics[height=\textwidth]{img/gyre_volume2.png}
    \caption*{a) IsoVolume}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_gauss_contour.PNG}
     \caption*{b) Contour}
    \label{b)}
  \end{minipage}
   \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_volume4.PNG}
     \caption*{c) GaussContour}
    \label{b)}
  \end{minipage}
  \caption{\enquote{Gyre} test field}
\label{gyre_volume}
\end{figure}
When we consider the volumes for the \enquote{gyre} test field, we can remark the holes, which represent the gyres in the centers of the volume for a low and a high threshold, whereby we extract (filter out) the previously observed ridges in $3D$ for all directions $\omega$ simultaneously.

Now, that the approach has passed the textbook examples, we push it forward to be applied in real (measured) data to show its performance in discerning and segmenting ROIs from a human brain. The approach will detect LCS in the brain's tissue from MRI-measured diffusion tensors of a male subject at the age of 27. We compare ParaView's glyph representation to the FTLE method LTG for the test field \enquote{brain}:
\begin{figure}[!t]
\centering
  \begin{minipage}[t]{0.3\textwidth}
  \includegraphics[height=\textwidth]{img/brain_cyl_glyphs.PNG}
    \caption*{a) glyph representation}
    \label{a)}
  \end{minipage}
  \begin{minipage}[t]{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/brain_ftle.PNG}
     \caption*{b) LTG}
    \label{b)}
  \end{minipage}
  \begin{minipage}[t]{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/brain_ftle_org.PNG}
     \caption*{c) FTLE: Hlawatsch's approach}
    \label{b)}
  \end{minipage}
  \caption{\enquote{Brain} test field}
\label{brain-ftle}
\end{figure}
Regarding Fig. \ref{gyre_volume}, we find that the approach detects the boundaries of concavities (lat.: Gyri) in the brain, forming the lobus cerebri (commonly known a lobes). The invented method could thus, e.g., aid to compute $3D$ models of a subject in medical sciences. It emphasizes regions with increased change in flow-rate (throughput), because in this areas the anisotropy and/or tensor magnitude is increased, preventing the intensities from being absorbed, hence yielding to a high FTLE-response. The regions highlighted in red, are caused by high tensor magnitudes, as we can infer from the glyph representation revealing no linear anisotropy at all. Note that we do not show the volume representation in this case, since all directions (slices) look approximately the same because the tensor field is only introducing changes in tensor magnitude not in anisotropy as inferred from the glyph representation. The regions with a tensor magnitude higher than zero restrict the light propagation to certain areas with the condition, that not all of the energy is absorbed by sorrounding cells in the local neighborhood. Again comparing the results to the classical FTLE (Hlawatsch's approach) on tensor fields in Fig. \ref{brain_ftle_comp}, we notice that the FTLE response is chaotic, since this MRI scan is subject to inherent MRI noise filling the background regions in the image. This noise produces random trajectories, leading to an increased gradient of the flow map and thus to a high FTLE response, which reveals trajectories running in random directions in background regions. Only in foreground regions, the trajectories are directed in definite manner, which is why the flow map's gradient is minimized along the brain structure (foreground, depicted in blue in Fig \ref{brain_ftle_comp} c)). Yet, our approach uses absorption to decrease noise present in the tensor field and reasonable min-max scaling to conduct $100\%$ of the intensity for the cell with the highest tensor magnitude. Hence, we do not yield a noisy representation but a FG-BG segmented image, with red highlights in positions, where the tensor magnitude is maximized (and thus absorption is minimized). Remark, that the resulting volume does not inherently depend on the direction, because the test field does not introduce any strong linear anisotropy, as visible from the glyph representation in Fig. \ref{brain_ftle_comp} a). If no linear anisotropy was incorporated, the FTLE response will also be independent on the direction to the greatest extent. The slices we obtain as our volume are, with reasonable explanation, more or less identical comparing them against each other for certain directions.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=0.9\textwidth]{img/brainDwnsmpl.png}
    \caption*{a) Glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.9\textwidth]{img/brainDwnsmpl-TFL.PNG}
     \caption*{b) Hlawatsch's app.: TFLs}
    \label{b)}
  \end{minipage}
\caption{\enquote{brain} test field}
\label{brain_ftle_comp}
\end{figure}
We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig. \ref{brain_ftle_comp}. We propose the following threshold correction to prevent noisy FTLE results:

We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig. \ref{brain_ftle_comp}. We propose the following threshold correction to prevent noisy FTLE results:

\begin{align*}
r_i(\omega) = 
\begin{cases}
   r_i(\omega) = r_i(\omega),& \text{if } \mean{r}_i > \theta,\\
     r_i(\omega),              & \text{otherwise}.
\end{cases}
\end{align*}

This normalization correction will ensure, that any mean tensor magnitude below threshold $\theta$ will be neglected for normalization and hence will not be amplified. The free parameter should be defined with heuristics concerning the signal-to-noise ratio. This now states a local normalization, which was not feasible with the noisy background before introducing this change, since a local normalization would have amplified noisy signals likewise. Now, we are capable of producing a tensor magnitude weighted FTLE mode with global normalization, which we used up to this point and another FTLE mode indicating changes in anisotropy separately while neglecting any tensor magnitudes for local normalization. This change will now yield a volume (now depending on the direction) for the local normalization mode, which we will show later on. We could also claim, that this correction would be necessary for Hlawatsch's approach as well, as their result is susceptive to noise for the \enquote{brain} test field in Fig. \ref{brain_ftle_comp} b). We first show the functionality of the correction on their approach in Fig. \ref{brain_ftle_noise}.

\begin{figure}[!t]
\centering
  \begin{minipage}[t]{0.4\textwidth}
  \includegraphics[width=0.9\textwidth]{img_new/brain-TFL.png}
    \caption*{a) Glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}[t]{0.4\textwidth}
    \includegraphics[width=0.9\textwidth]{img/brain_ftle_surf_noisereduction.PNG}
     \caption*{b) H.'s app.: TFLs (noise correction)}
    \label{b)}
  \end{minipage}
\caption{\enquote{brain} test field}
\label{brain_ftle_noise}
\end{figure}
Again observing the results for the tensor field lines, we now remark their proper correlation in BG regions (cf. straight lines in the background in Fig. \ref{brain_ftle_noise} a)). This yields a gradient of the flow map exactly equal to zero, which avoids any noisy FTLE responses, as visible in Fig. \ref{brain_ftle_noise} b) for Hlawatsch's approach.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
  \includegraphics[width=\textwidth]{img_new/brain_cross-section.PNG}
    \caption*{a) tensor magnitude}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img_new/brain_volume.PNG}
     \caption*{b) LTG}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img_new/brain_volume2.PNG}
     \caption*{c) FTLE H.'s approach}
    \label{b)}
  \end{minipage}
  \caption{\enquote{heart} test field}
\label{heart-ftle}
\end{figure}

The volumes depicted in Fig. \ref{heart-ftle} now show variation in $z$-direction, as they are now visualizing pure FTLE fields with no weighting of the tensor magnitude incorporated.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
  \includegraphics[height=\textwidth]{img/heart_org.PNG}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/heart.PNG}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/heart_ftle_org.PNG}
    \label{b)}
  \end{minipage}
  \caption{\enquote{Brain} test field: volumes}
\label{heart-ftle}
\end{figure}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
  \includegraphics[width=\textwidth]{img/heart_org.PNG}
    \caption*{a) tensor magnitude}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/heart.PNG}
     \caption*{b) LTG}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/heart_ftle_org.PNG}
     \caption*{c) Hlawatsch's approach}
    \label{b)}
  \end{minipage}
  \caption{\enquote{heart} test field}
\label{heart-ftle}
\end{figure}
After applying the LTG approach on the heart dataset, we observe the resulting tensor magnitude weighted FTLE-field, which is again not changing w.r.t. direction, which is why we do not show the volumes. The approach highlights the regions with diverging tensor field lines resulting from linear anisotropy or increased tensor magnitude, emerging as ridges. These ridges could now represent blood vessels and such. The red spot emphasized in the left of the image could, e.g., represent the highest diffusion velocity in tissue. When considering the FTLE for the heart dataset depicted in Fig. \ref{heart-ftle-comp}, we find that the trajectories exhibit strong correlations amongst them. This results from the low noise rate inherently present in this dataset. The inital value for our angle is given by $0$ degree, which is why the trajectories follow this angle strictly in case of a zero matrix (corresponding to a noise free cell in background regions). In this case, we obtain the same results as the approach from Hlawatsch et al., whereas our global min-max normalization (due to noise) is becoming a redundant step for noiseless input data. A proof of this elaboration can be found in Fig. \ref{heart-ftle-comp}.

\begin{figure}[!t]
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/heartDwnsmpl.png}
    \caption*{Glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/heartDwnsmpl-TFL.PNG}
     \caption*{Hlawatsch's approach: TFLs}
    \label{b)}
  \end{minipage}
\caption{\enquote{heart} test field}
\label{heart-ftle-comp}
\end{figure}
Again considering the trajectories for the heart test field, we note that they all correlate in background regions, which causes the gradient of the flow map to equal zero. This occurs, since we do not have any MRI noise incorporated into the background of the field. Consequently, we obtain results corresponding to Hlawatsch's approach, which is in general susceptive to noise, but not in case the noise was removed or is not present at all initially. From observing Hlawatsch's approach detecting the heart edges, we can infer that the flow is also congruent inside of the heart, inducing a gradient response only to the edges. In our case, we have an offset response incorporated which happens even for total spherical isotropy, wherefore we will obtain a response even if the flow does not change locally (presumed it is not absorbed by low tensor magnitudes). This in turn stems from the fact, that we use a energetic flood-fill difference, which inherently introduces an offset response.
\begin{figure}[!t]
 \begin{minipage}{0.3\textwidth}
 \includegraphics[width=\textwidth]{img_new/heart_volume.PNG}
   \label{a)}
 \end{minipage}
 \begin{minipage}{0.3\textwidth}
   \includegraphics[width=\textwidth]{img_new/heart_volume3.PNG}
   \label{b)}
 \end{minipage}
 \begin{minipage}{0.3\textwidth}
   \includegraphics[width=\textwidth]{img_new/heart_volume5.PNG}
   \label{b)}
 \end{minipage}
 \caption{\enquote{heart} test field: LTG volumes}
\label{heart-ftle}
\end{figure}
Fig. \ref{heart-ftle} depicts the volumes for the heart test field, we can observe a 3D ridge running in spiral along the $z$-direction.

\paragraph{Documentation}\vskip 1pt
\textit{Spatial resolution pyramid}
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
  \includegraphics[height=\textwidth]{img/inverse29.png}
    \caption*{$\mathop{res}=29\times29\times16$}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth} 
    \includegraphics[height=\textwidth]{img/inverse51.PNG}
     \caption*{$\mathop{res}=51\times51\times16$}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
  \centering
    \includegraphics[height=\textwidth]{img/inverse101.PNG}
     \caption*{$\mathop{res}=101\times101\times16$}
    \label{b)}
  \end{minipage}
  \caption{Spatial resolution pyramid of \enquote{inverse} test field - slice}
\label{res_ftle}
\end{figure}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
 % \centering
  \includegraphics[height=\textwidth]{img/inverse29_volume.png}
    \caption*{$\mathop{res}=29\times29\times16$}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
  %\centering
    \includegraphics[height=\textwidth]{img/inverse51_volume.PNG}
     \caption*{$\mathop{res}=51\times51\times16$}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
 % \centering
    \includegraphics[height=\textwidth]{img/inverse101_volume.PNG}
     \caption*{$\mathop{res}=101\times101\times16$}
    \label{b)}
  \end{minipage}
  \caption{Spatial resolution pyramid of \enquote{inverse} test field - volume}
\label{res_ftle_volume}
\end{figure}
When observing these (decomposed) spatial resolution pyramids shown, we can observe the slice image becoming more and more continous with increasing resolution to the right. The right image on the top, shows a result fulfilling the Nyquist-Shannon sampling theorem, as it appears with no jaggies or pixelation and thus successfully deceiving the human mind in being continous. The volumes on the other hand, exhibit a slight difference, which is due to the revealing of new discretized directions implicitly in the tensor field, as shifts in angle become more gradual and continuously for a more dense sampling of the same implicitly content-based field. This effects a more precise behavior, which now indicates the orthogonal ridges from each of the $2$ directions in parallel to the light source direction (axis-parallel) on each side of the volume. They are evenly spaced in a shift of half the volume height which represents the angle $180^\circ$ (2 antiparallel directions in axis directions). The blue span (hole), we perceive in the middle of the cross-section, is representing the antiparallel direction, which in turn is missing, since we only can detect divergence of tensor field lines in axis-parallel up to -orthogonal directions (if we consider them pointing into the diverging direction). Note that we already discussed this feature in section \ref{chap:meth}.
Also remark, that the ridges get sharper with increasing spatial test field (input data) resolution because the implicitly defined content of the tensor field is more densely sampled at a higher resolution allowing for more continous changes in directions of the major eigenvectors (glyph orientations).
\textit{Angular resolution pyramid}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
  \includegraphics[height=\textwidth]{img/slice16.png}
    \caption*{$\mathop{res}=101\times101\times16$}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
    \includegraphics[height=\textwidth]{img/slice32.PNG}
     \caption*{$\mathop{res}=101\times101\times32$}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
  \centering
    \includegraphics[height=\textwidth]{img/slice64.PNG}
     \caption*{$\mathop{res}=101\times101\times64$}
    \label{b)}
  \end{minipage}
  \caption{Angular resolution pyramid of \enquote{inverse} test field - slice}
\label{resA_ftle}
\end{figure}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
 	\centering
  \includegraphics[height=\textwidth]{img/vol16.png}
    \caption*{$\mathop{res}=101\times101\times16$}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
  \centering
    \includegraphics[height=\textwidth]{img/vol32.PNG}
     \caption*{$\mathop{res}=101\times101\times32$}
    \label{b)}
  \end{minipage}
  \begin{minipage}{0.3\textwidth}
  \centering
    \includegraphics[height=\textwidth]{img/vol64.PNG}
     \caption*{$\mathop{res}=101\times101\times64$}
    \label{b)}
  \end{minipage}
  \caption{Angular resolution pyramid of \enquote{inverse} test field - volume}
\label{resA_ftle_volume}
\end{figure}
Observing these angular resolution pyramids, one may again notice, that the results can be considered to be independent on the angular resolution, as well. The only change incorporated is that the ridges get more and more sharp with increasing angular resolution, which is to be expected, since the cutoffs of the transmission profiles (transfer functions) are resolved in more detail, too, which allows the light distributions to follow the tensor field lines more strictly. This turns out to be the counterpart to the spatial resolution.

Please note that all values in the test table where measured on following system:\\
\texttt{4 x 16-Core Gold 6142M 2.60GHz, 3072 GB RAM}.
\begin{table}
\small
\centering
\caption{Test Table LTG:}
\begin{tabularx}{\columnwidth}{X|X|X|X}
\hline\hline
test name - Fig. \# & resolution ($x\times y\times\omega$) & runtime $[s]$ & memory consumption\\
\toprule
\enquote{drain} - \ref{ftle_base} & $101\times101\times64=652864$ & $179878$ & $8\times652864\times 8\mathop{byte} +(101\times101 + 19\times64)\times8\mathop{byte}\approx47.02\mathop{MB}$\\ 
\hline 
\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times64=652864$ & $68787$ & $8\times652864\times 8\mathop{byte} +(101\times101 + 19\times64)\times8\mathop{byte}\approx47.02\mathop{MB}$\\
\hline 
 &  &  & \\ 
\hline 
\enquote{isotropic} - \ref{bow-ftle} & $51\times51\times3283232$ & $1357$ & $8\times83232\times 8\mathop{byte} +(51\times51 + 19\times32)\times8\mathop{byte}\approx5.998\mathop{MB}$\\ 
\hline
\enquote{rings} - \ref{bow-ftle} & $51\times51\times32=83232$ & $5761$ & $8\times83232\times 8\mathop{byte} +(51\times51 + 19\times32)\times8\mathop{byte}\approx5.998\mathop{MB}$\\ 
\hline 
\enquote{bow} - \ref{bow-ftle} & $51\times51\times32=83232$ & $6128$ & $8\times83232\times 8\mathop{byte} +(51\times51 + 19\times32)\times8\mathop{byte}\approx5.998\mathop{MB}$\\ 
\hline 
\enquote{star} - \ref{star-ftle} & $51\times51\times32=83232$ & $5815$ & $8\times83232\times 8\mathop{byte} +(51\times51 + 19\times32)\times8\mathop{byte}\approx5.998\mathop{MB}$\\ 
\hline 
\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times16=163216$ & $12308$ & $8\times163216\times 8\mathop{byte} +(101\times101 + 19\times16)\times8\mathop{byte}\approx11.754\mathop{MB}$\\ 
\hline 
\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times32=326432$ & $24571$ & $8\times326432\times 8\mathop{byte} +(101\times101 + 19\times32)\times8\mathop{byte}\approx23.5\mathop{MB}$\\ 
\hline
\enquote{brain} - \ref{brain-ftle} & $128\times128\times64=1048576$ & $2427$ & $8\times1048576\times 8\mathop{byte} +(128\times128 + 19\times64)\times8\mathop{byte}\approx75.5\mathop{MB}$\\
\hline
\enquote{heart} - \ref{heart-ftle} & $256\times256\times32=2097152$ & $117199$ & $8\times2097152\times 8\mathop{byte} +(256\times256 + 19\times32)\times8\mathop{byte}\approx151.0\mathop{MB}$\\
\hline 
\end{tabularx}
   \vskip 1pt
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusion and Future Work}
\label{chap:concl}
In this work, we have discussed the advantages and purposes of tensor field visualization in practical applications. Furthermore, we have presented a presented a technique to extract meaningful structures from tensor fields, which can, as a secondary purpose, also be used to visualize meaningful structures in $2D$ virtual geometrical scenes. To put it into context, we discussed its relations to vector field visualization techniques and other tensor field and light transport visualization methods. Also, we captured global illumination methods, since the proposed visualization method is derived from a global light transport method propagating intensities inside of a given Cartesian grid, consistent of cells defined by polar profiles (distributions), respectively. The intensities are redistributed stochastically by transmission profiles derived from the tensors in the field, used as windowing (weighting) functions. Based on this propagation algorithm, we can compute a FTLE-like field on the tensor field, as we obtain a final light distribution for each light source configuration (position and direction), which is considered as an energy-based flow map. In analogy to the computation of the FTLE field for vector fields, we analyze the gradient of the resulting flow map. The incongruity on the other hand, is that we need an additional inital direction on top of the inital position, because the tensors also dependent on the incoming direction, not as vectors only on their location $x,y$. As a result, we obtain a $3D$ scalar field, which can be visualized as volume (e.g. heatmap representation) to extract and segment ridges, i.e., key regions separating areas of correlating directional distributions. These ridges are in turn called Lagrangian coherent structures (LCS), which can be found in nature in , e.g., vortices in flow fields or moraines in glaciers. We designed comprehensive synthetic test tensor fields to evaluate our approach. Once we were convinced over its principal functionality, we applied the approach to some real data examples, namely DT-MRI scans of human brain and heart. Also we designed test runs to verify the physical plausibility of the results of the propagation scheme. The maps or fields we obtain, can make a statement about active regions in tensor fields, which do reveal the greatest tensor magnitude and anisotropy in combination. The approach is positioned in interesting relation to the works of Zirr et al. \cite{zirr}, which perform their FTPD (FTLE-like) computation by random sampling of trajectories to compute light transport separatrices. They do not incorporate any transmission weighting functions, but instead blocker (obstacle) or reflecting geometry, which causes a FTLE response as well, as edges or mirrors cause light paths to cross each other chaotically in similar manner. Furthermore, the work of Hlawatsch et al. \cite{hlawatsch} obtains an FTLE field on tensor fields in similar style, as they use a coordinate map to sample their trajectories.

A possible application would be mechanical component design, as we could segment ROIs exhibiting increased loads or in a greater scope even machine component design, which are produced in series. The invented technique could also aid geologist in analyzing stress in cross-sections of the earth to be able to predict earthquakes sufficiently soon enough. In general one could measure vortices in directional distributions and hence tensor fields with a topological background (e.g. extracting separatrices of manifolds). One of the most relevant problems in visualization is incorporating uncertainty \cite{johnson}. Measurement, simulation or systematic (model) errors, boundary conditions or turbulences may introduce local uncertainty, which turns into global uncertainty through moving with the flow. This introduces propability density functions (PDFs) instead of certain single trajectories into the flow field. A real world example for uncertain vector fields is the measurement of blood flow and tissue velocity by phase contrast MRI, which yields an uncertain vector field due to the inherent noise in MRI scans. This uncertain vector field could be eventually visualized with our proposed technique. We just need to put it into the right context. The analogy to certain vector fields would be, that we could interpret our principal component ellipsoid as a propability density function and even replace it with such. The trajectories would be represented as propability distributions, which themselves are in turn propagated, just as we propagate our intensity profiles (distributions). As an outlook, we could apply our invented method to uncertain vector fields, which could be sampled in transmission profiles in discrete polar coordinates, just as we do in our work with the polar ellipsoids. Future works might include the previously mentioned application of the approach to uncertain vector fields or evaluating its applicability to real data in segmenting meaningful structures from engineering and medical data. Also we could imagine to implement a GPU parallelized version of the method to be able to propagate the intensities more efficiently, which would decrease the runtime in general. We can conclude, that tensor field visualization is becoming more and more important in scientific domains and that we invented a useful new tool in this context, which is worth of being investigated more precisely to reveal its potential in detecting meaningful information (key regions) and abstracting from redundancies in tensor fields.

%Hier werden noch einmal die wichtigsten Ergebnisse und Erkenntnisse
%der Arbeit zusammengefasst (nicht einfach eine Wiederholung des
%Aufbaus der vorherigen Kapitel!), welche neuen Konzepte, Methoden und
%Werkzeuge Neues entwickelt wurden, welche Probleme nun (effizienter)
%gel?st werden k?nnen, und es wird ein Ausblick auf weiterf?hrende
%Arbeiten gegeben (z.B.~was Sie machen w?rden, wenn Sie noch 6 Monate
%mehr Zeit h?tten).

% References (Literaturverzeichnis):
% a) Style (with abbreviations: use alpha):
% see
% https://de.wikibooks.org/wiki/LaTeX-W%C3%B6rterbuch:_bibliographystyle
% for the different formats and styles

\bibliographystyle{ieeetr} %acm - as capital letter ALTERNATIVE
% b) The File:
\bibliography{references}

\end{document}

### 2019-07-06 thesis-template/msc-thesis.tex
update

@@
   \begin{minipage}{0.4\textwidth}
+   \centering
   \includegraphics[width=0.9\textwidth]{img/brainDwnsmpl.png}
@@
   \begin{minipage}{0.4\textwidth}
+   \centering
     \includegraphics[width=0.9\textwidth]{img/brainDwnsmpl-TFL.PNG}
@@
 \begin{cases}
-   r_i(\omega) = r_i(\omega),& \text{if } \mean{r}_i > \theta,\\
+    \frac{1}{\mean{r}_i}r_i(\omega),& \text{if } \mean{r}_i > \theta,\\
      r_i(\omega),              & \text{otherwise}.
@@
   \begin{minipage}[t]{0.4\textwidth}
+   \centering
   \includegraphics[width=0.9\textwidth]{img_new/brain-TFL.png}
@@
   \begin{minipage}[t]{0.4\textwidth}
+   \centering
     \includegraphics[width=0.9\textwidth]{img/brain_ftle_surf_noisereduction.PNG}
@@
   \begin{minipage}{0.3\textwidth}
-  \includegraphics[width=\textwidth]{img_new/brain_cross-section.PNG}
+  \includegraphics[height=\textwidth]{img_new/brain_cross-section.PNG}
     \caption*{a) tensor magnitude}
@@
   \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img_new/brain_volume.PNG}
+    \includegraphics[height=\textwidth]{img_new/brain_volume.PNG}
      \caption*{b) LTG}
@@
   \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img_new/brain_volume2.PNG}
+    \includegraphics[height=\textwidth]{img_new/brain_volume2.PNG}
      \caption*{c) FTLE H.'s approach}
@@
   \begin{minipage}{0.3\textwidth}
+  \centering
   \includegraphics[height=\textwidth]{img/heart_org.PNG}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
+    \caption*{a) tensor magnitude}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
     \includegraphics[height=\textwidth]{img/heart.PNG}
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
+     \caption*{b) LTG}
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
     \includegraphics[height=\textwidth]{img/heart_ftle_org.PNG}
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Brain} test field: volumes}
-\label{heart-ftle}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-  \includegraphics[width=\textwidth]{img/heart_org.PNG}
-    \caption*{a) tensor magnitude}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img/heart.PNG}
-     \caption*{b) LTG}
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img/heart_ftle_org.PNG}
      \caption*{c) Hlawatsch's approach}
@@
 \begin{figure}[!t]
-  \begin{minipage}{0.4\textwidth}
+\centering
+  \begin{minipage}{0.4\textwidth}
+   \centering
   \includegraphics[width=\textwidth]{img/heartDwnsmpl.png}
@@
   \begin{minipage}{0.4\textwidth}
+   \centering
     \includegraphics[width=\textwidth]{img/heartDwnsmpl-TFL.PNG}
@@
 This in turn stems from the fact, that we use a energetic flood-fill difference, which inherently introduces an offset response.
-\begin{figure}[!t]
- \begin{minipage}{0.3\textwidth}
- \includegraphics[width=\textwidth]{img_new/heart_volume.PNG}
-   \label{a)}
- \end{minipage}
- \begin{minipage}{0.3\textwidth}
-   \includegraphics[width=\textwidth]{img_new/heart_volume3.PNG}
-   \label{b)}
- \end{minipage}
- \begin{minipage}{0.3\textwidth}
-   \includegraphics[width=\textwidth]{img_new/heart_volume5.PNG}
+We can again apply a threshold for the normalization to obtain a result non-dependent on the noise to visualize raw changes in anisotropy and thus to obtain a variation in $z$-direction, ie., along the light source direction.
+\begin{figure}[!t]
+\centering
+ \begin{minipage}{0.3\textwidth}
+ \centering
+ \includegraphics[height=\textwidth]{img_new/heart_volume.PNG}
+   \label{a)}
+ \end{minipage}
+ \begin{minipage}{0.3\textwidth}
+ \centering
+   \includegraphics[height=\textwidth]{img_new/heart_volume3.PNG}
+   \label{b)}
+ \end{minipage}
+ \begin{minipage}{0.3\textwidth}
+ \centering
+   \includegraphics[height=\textwidth]{img_new/heart_volume5.PNG}
    \label{b)}
@@
 \end{figure}
-Fig. \ref{heart-ftle} depicts the volumes for the heart test field, we can observe a 3D ridge running in spiral along the $z$-direction.
+Fig.
+\ref{heart-ftle} depicts the volumes for the heart test field, whereas we can observe a 3D ridge running in spiral along the $z$-direction.
+The FTLE responses are shifted w.r.t.
+varying directions along the $z$-axis.
+
+We also introduced random numbers in to the our tensor field generation process.
+Therefore, we define a uniform grid with low resolution ($9\times9$) and interpolate bilinearly to obtain a tensor field with higher resolution exhibiting continous changes in anisotropy and tensor magnitude.
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+   \centering
+  \includegraphics[width=\textwidth]{img_new/random-global.png}
+    \caption*{Global normalization}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+   \centering
+    \includegraphics[width=\textwidth]{img_new/random-local.PNG}
+     \caption*{Local normalization}
+    \label{b)}
+  \end{minipage}
+\caption{\enquote{Random} test field}
+\label{random}
+\end{figure}
+Observing the generated random test field both with global and local normalization in Fig.
+\ref{random}a)\& b) respectively, we remark that the propability for small numbers to have a high ratio in combination is increased, as linear anisotropy occurs mainly in BG regions (low tensor magnitudes).
+In case we use global normalization, we will again obtain a FTLE field independent on the direction, since tensor magnitudes outshine the changes in anisotropy in this test case.
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+   \centering
+  \includegraphics[width=\textwidth]{img_new/random_global.png}
+    \caption*{Global normalization}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+   \centering
+    \includegraphics[width=\textwidth]{img_new/random_local_slice0.PNG}
+     \caption*{Local normalization}
+    \label{b)}
+  \end{minipage}
+\caption{\enquote{Random} test field}
+\label{random}
+\end{figure}
+In case we use local normalization, we could interpret the contributions revealing low tensor magnitudes as a result from white noise, which shows the inherent problem with local normalization.
 

### 2019-07-06 thesis-template/msc-thesis.tex
update

@@
 Student ID: & 3481802\\
-Supervisors: & Prof. Filip Sadlo, Dr. Susanne Kr?mker\\
+Supervisors: & Prof. Filip Sadlo, Dr. Susanne Krömker\\
 Submission date: & 07.07.2019
@@
     \includegraphics[height=\textwidth]{img/star_ftle.PNG}
-     \caption*{LTG slice($90^\circ$)}
+     \caption*{LTG slice($0^\circ$)}
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/gyre_ftle.PNG}
-     \caption*{b) LTG}
+     \caption*{b) LTG($0^\circ$)}
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/brain_ftle.PNG}
-     \caption*{b) LTG}
+     \caption*{b) LTG($0^\circ$)}
     \label{b)}
@@
   \centering
-  \includegraphics[height=\textwidth]{img/heart_org.PNG}
-    \caption*{a) tensor magnitude}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
   \includegraphics[height=\textwidth]{img/heart.PNG}
-     \caption*{b) LTG}
+    \caption*{a) LTG($0^\circ$): global}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[height=\textwidth]{img_new/heart_slice0.PNG}
+     \caption*{b) LTG($0^\circ$): local}
     \label{b)}
@@
 \ref{random}a)\& b) respectively, we remark that the propability for small numbers to have a high ratio in combination is increased, as linear anisotropy occurs mainly in BG regions (low tensor magnitudes).
-In case we use global normalization, we will again obtain a FTLE field independent on the direction, since tensor magnitudes outshine the changes in anisotropy in this test case.
+In case we use global normalization in Fig.
+\ref{random_FTLE}a), we will again obtain a FTLE field independent on the direction, since tensor magnitudes outshine the changes in anisotropy in this test case.
+The LTG approach works as expected, as we see the tensor magnitude weighted into the left image, yielding a high response for the tensor magnitude weighted FTLE response.
 \begin{figure}[!t]
@@
   \includegraphics[width=\textwidth]{img_new/random_global.png}
-    \caption*{Global normalization}
+    \caption*{LTG: Global normalization}
     \label{a)}
@@
     \includegraphics[width=\textwidth]{img_new/random_local_slice0.PNG}
-     \caption*{Local normalization}
+     \caption*{LTG: Local normalization}
     \label{b)}
@@
 \caption{\enquote{Random} test field}
-\label{random}
-\end{figure}
-In case we use local normalization, we could interpret the contributions revealing low tensor magnitudes as a result from white noise, which shows the inherent problem with local normalization.
+\label{random_FTLE}
+\end{figure}
+In case we use local normalization in Fig.
+\ref{random_FTLE}b), we could interpret the contributions revealing low tensor magnitudes as a result from white noise, which shows the inherent problem with local normalization.
+If we would just locally normalize the whole tensor field with the mean tensor magnitude of each cell, we would enhance noisy contributions, which turn out to be in general more propable to be anisotropic than FG data.
+Keeping this consideration in mind, we observe the approach detecting raw changes in anisotropy.
+Note, that this test field does not introduce clearly defined ridges with tensor field lines diverging w.r.t.
+to a middle axis.
+We just introduce rapid random variations in anisotropy along the spatial domain, which is why the FTLE response is independent on the direction in this case, as well.
+This example emphasizes, that ridges can also be yielded from rapid changes in anisotropy and are hence in this case, not directed and thus non directionally dependent.
+Also we take home from this example, that the thresholding is in fact necessary to form a reasonable local normalization as a posteriori verification.
 
@@
 \hline\hline
-test name - Fig. \# & resolution ($x\times y\times\omega$) & runtime $[s]$ & memory consumption\\
+test name - Fig. \# & resolution ($x\times y\times\omega$) & runtime $[s]$ & memory consumption (dynamic)\\
 \toprule
@@
 \hline 
- &  &  & \\ 
+\enquote{random} &  &  & \\ 
 \hline 

### 2019-07-07 thesis-template/msc-thesis.tex
update

@@
 
-\textbf{\large Master-Arbeit} % Bachelor-Arbeit
+\textbf{\large Master's Thesis} % Bachelor-Arbeit
 
@@
 \emph{dt.:}
-Ich versichere, dass ich diese Master-Arbeit selbstst?ndig verfasst und nur die angegebenen
-Quellen und Hilfsmittel verwendet habe und die Grunds?tze und
-,Empfehlungen ``Verantwortung in der Wissenschaft'' der Universit?t Heidelberg beachtet wurden. 
+Ich versichere, dass ich diese Master-Arbeit selbstständig verfasst und nur die angegebenen
+Quellen und Hilfsmittel verwendet habe und die Grundsätze und
+,Empfehlungen ``Verantwortung in der Wissenschaft'' der Universität Heidelberg beachtet wurden. 
 \bigskip
@@
 Tensorfelder werden meistens in Verbindung mit mechanischen Spanunngsverteilungen in 2D/3D-Gittern gebracht (vgl. Cauchy-Stress Tensor), haben aber auch andere praktische Bedeutungen in der Physik.
-Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren, die aber zus?tzlich auch verwendet werden kann um die Lichtausbreitung in einer topologischen Szene zu beschreiben.
-Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema f?r kartesische Gitter ab, dass die Prinzipien der Ausbreitungsd?mpfung und Energieerhaltung beachtet und Licht-verteilung/en f?r gegebene Lichtquellenrichtung/en und -positionen bis zur Konvergenz approximieren kann.
+Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren, die aber zusätzlich auch verwendet werden kann um die Lichtausbreitung in einer topologischen Szene zu beschreiben.
+Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema für kartesische Gitter ab, dass die Prinzipien der Ausbreitungsdämpfung und Energieerhaltung beachtet und Licht-verteilung/en für gegebene Lichtquellenrichtung/en und -positionen bis zur Konvergenz approximieren kann.
 Als unterliegendes Modell werden die Transmissionsprofile innerhalb diesem Gitter als Kristallfiberstruktur angenommen (vgl.
 Edelsteine: Katzenaugeneffekt beim Tigerauge).
-Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiema?en des unterliegenden Tensorfelds zu modellieren.
-Daf?r erstellen wir per Hauptachsentransformation (PCA) ein Eigensystem f?r jede Zelle des Tensorfelds und modellieren eine zugeh?rige Ellipsengleichung als (Transfer-/Wichtungs-) Transmissionsprofil.
-Folglich kann die resultierende Lichtverteilung als 2D Skalarfeld oder Polarplots f?r Richtungsinformation visualisiert/repr?sentiert werden.
-Wir messen Impulsantworten des Tensorfelds mit Delta-Pulsen an jeder Position und in jede Richtung um eine uniform gesamplete Map der globalen Lichtverteilungen zu erhalten.
+Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiemaßen des unterliegenden Tensorfelds zu modellieren.
+Dafür erstellen wir per Hauptachsentransformation (PCA) ein Eigensystem für jede Zelle des Tensorfelds und modellieren eine zugehörige Ellipsengleichung als (Transfer-/Wichtungs-) Transmissionsprofil.
+Folglich kann die resultierende Lichtverteilung als 2D Skalarfeld oder Polarplot für Richtungsinformation visualisiert/repräsentiert werden.
+Wir messen Impulsantworten des Tensorfelds mit Delta-Pulsen an jeder Position und in jede Richtung um eine uniform gesamplte Map der globalen Lichtverteilungen zu erhalten.
 Diese Map wird als ein globales Energieflussfeld betrachtet.
 Wir lassen uns von der Technik FTLE aus dem Bereich Particle Tracing der Vector Field Visualization inspirieren, indem wir den Gradient des Flussfelds der resultierenden Lichtverteilungen analysieren und visualisieren.
-?hnlich wie FTLE in Vektorfeldern Gr?te/K?mme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, anziehende/absto?ende/sattelnde Tensorfeldlinien die h?ufig Schl?sselstrukturen in der Natur repr?sentieren, in Tensorfeldern.
-Wir f?hren diese Gr??e als Global Illumination Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtflusses, zu visualisieren.
+Ähnlich wie FTLE in Vektorfeldern Gräte/Kämme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, anziehende/abstoßende/sattelnde Tensorfeldlinien die häufig Schlüsselstrukturen in der Natur repräsentieren, in Tensorfeldern.
+Wir führen diese Größe als Global Illumination Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtfluss, zu visualisieren.
 \newpage
@@
 tensor field representations.
-Tensor fields, are most commonly associated with stress distributions (cf.
-Cauchy-stress tensor) in 2D/3D grids, but also have some other meanings in practice in physics.
-By means of global illumination techniques, we motivate a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in a topologically defined scene geometry.
-As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence.
-The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (, e.g., gemstones: tiger's eye's cat's eye effect).
-The consequent task is to model anisotropic fiber structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
-For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer) function for the transmitted light profiles.
+Tensor fields, are commonly associated with stress distributions (cf.
+Cauchy-stress tensor) in 2D/3D grids, but also have other meanings in various fields.
+By means of global illumination techniques, we motivate a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in anisotropic media.
+As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles, and is able to approximate light distributions for given light source position(s) and direction(s).
+The transmission profiles within this grid are considered as \enquote{crystal fiber structures} as an underlying physical model (e.g.: gemstones: tiger's eye's cat's eye effect).
+The consequent task is to interpret anisotropic fiber structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
+For this, we derive an eigenbasis by principal component analysis for every cell of the tensor field and form an ellipsoid equation as transmission (conductivity) function for the transmitted light profiles.
 Hence, the resulting light distribution can be visualized via 2D scalar field color coding or polar plots for directional information.
-On this basis, we measure impulse responses of the tensor field with Dirac-pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions.
+On this basis, we measure impulse responses of the tensor field with Dirac pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions.
 This map is considered as a global illumination energy flow field.
 We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field, formed by the resulting light distributions.
-Similar to how the FTLE detects ridges/LCS (Lagrangian Coherent Structures) in vector fields (flow fields), our approach captures these structures likewise, namely regions revealing attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields.
-We denote this entity as light transport gradient (LTG) which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global light transport in the grid, directed by imprints in crystelline fiber structures modeled by the tensor field's ellipsoid glyhps.
-We evaluate our approaches for plausibility and comparability within an extensive test campaign and demonstrate its performance on several examples of both synthetic and real datasets.
+Similar to how the FTLE detects ridges/Lagrangian coherent structures in vector fields, our approach captures these structures likewise, namely regions revealing attracting/repelling/saddling tensor field lines which frequently represent key structures in nature, in tensor fields.
+We denote this entity as the light transport gradient (LTG), which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global light transport in the grid, directed by imprints in crystelline fiber structures modeled by the tensor field.
+We evaluate our approaches for plausibility and comparability within an extensive evaluation and demonstrate its performance on several examples of both synthetic and real datasets.
 \newpage
@@
 
-In this work, we will discuss the advantages of tensor field analysis and motivate a new method for tensor field visualization through a coherent light transport visualization.
-At first, we will shortly state the problem setting, objectives and give a quick overview about the structure of this work.
+In this work, we discuss the advantages of tensor field analysis and motivate a new method for tensor field visualization through a coherent light transport concept.
+We shortly state the problem setting, objectives and give a quick overview about the structure of this work.
 The main contributions of this work include:
 
-\paragraph{Contributions of this Work}
 \begin{itemize}
-	\item light transport model (propagation scheme) following basic but crucial physical principles for tensor field visualization using tensors as transmission profiles,
+	\item a light transport model (propagation scheme) following basic but crucial physical principles for tensor field visualization interpreting tensors as light transmission properties,
 	\item visualization of the propagation results in $2D$ polar profiles and scalar fields,
@@
 \ref{ToS}.
-While most techniques focus on symmetric tensor field visualization (like glyphs and tensor field lines), some other recent works adress the problem of asymmetric tensor field visualization.
+While most techniques focus on symmetric tensor field visualization (like glyphs and tensor field lines), some other recent works address the problem of asymmetric tensor field visualization.
 The issue with asymmetric (non-symmetric) or therein included antisymmetric (skew-symmetric) tensors is, that they do not always yield real eigenvalues, which is needed to determine an eigenbasis to set up glyphs, hyperstreamlines or tensor field lines (classically).
@@
 We imagine a flow field, i.e., a vector field which in turn represents the effective amount or concentration shifted into a particular direction at each point in space.
-It could induced by e.g.
+It could be induced by e.g.
 concentration or height gradients.
@@
 DOMs were recently revisited by Fattal \cite{fattal}.
-This work adresses artifacts stemming from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}.
+This work addresses artifacts stemming from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}.
 These methods are considered to be too slow to \enquote{run at interactive framerates} by Dachsbacher et al.
@@
  {
-    \includegraphics[width=0.8\textwidth]
+    \includegraphics[width=0.7\textwidth]
     {img/geogebra-export_large.pdf}
@@
   }
-  \caption{red: Intensity Profile, cyan: Transmission Profile (glyph Ellipsoid Eq.)}
+  \caption{yellow: Intensity Profile, cyan: Transmission Profile (glyph Ellipsoid Eq.)}
   \label{weighting}
@@
 \end{figure}
-We can see the light distributions diverge following tensor field lines, which ?n turn form these curves depicted in Fig.
+We can see the light distributions diverge following tensor field lines, which in turn form these curves depicted in Fig.
 \ref{evolution1}.
@@
 \end{figure}
-For the angle, we can observe again that the difference is again increased, but does not 
+For the angle difference in Fig.
+\ref{evolution2}, we can observe that the difference is again increased, because it does not overlap leading to a maximized difference.
 Imagine accumulating each pixel in the images to obtain the cumulative total energy in the (difference) grid, which could now be compared for each light source position and direction in terms of proper min-max scaling.
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \caption{Propagate Light Distribution}\label{PLD}
@@
 When we consider the LTG method, we already stated its runtime behavior ($\mathcal{O}(\mathop{width^5}\mathop{steps^2})$), which makes parallelization libraries and hardware to compute the results in reasonable time.
-First, we tried to implement a GPU-based parallelization for the propagation scheme, which turned out not to work as expected, since the relative runtime gain was too low make a big difference.
+First, we tried to implement a GPU-based parallelization for the propagation scheme with nVidia's Thrust library, which turned out not to work as expected, since the relative runtime gain was too low to make a big difference and we struggled to find a suitable implementation mapping the propagation process to atomic OPs (simple vector operations), which are the only operations allowed for GPU-parallelization.
 At this point, we choose to use OpenMP as we figured it is a much more suitable architecture employed here, which allows to perform the computation on a multiprocessor architecture provided by the Heidelberg University allowing to compute $64$ hardware threads at maximum.
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 \caption{Propagate BufferA to BufferB}\label{PB}
@@
 At first, we needed to generate convenient test cases, which we design using Mathematicas VectorPlot, used to visualize vector fields.
-To output a pre-defined test case, we need to define a little code, which we do using python scripts which do output VTK appended or ASCII format files.
+To output a predefined test case, we need to define a little code, for which we used python scripts which generate VTK appended or ASCII format files.
 To implement the method we used C++'s STL for every dynamic memory (buffer), SFML to plot polar profiles, Eigen to perform matrix decompositions and VTK to in- and output VTK format files, which can in turn be visualized and explored interactively with the visualization framework ParaView.
@@
 \label{sect:conserve}
-The energy conservation test was done in similar manner as thepropagation attenuation test in sect.
+The energy conservation test was done in similar manner as the propagation attenuation test in sect.
 \ref{sec:att}.
@@
 When we look at the example depicted in Fig.
-\ref{star-ftle}, we can see the FTLE approach LTG responding most in the center of the field, which is the knot point of tensor field lines diverging in any direction.
+\ref{star-ftle}, we can see the FTLE approach LTG responding most in the center of the field, which is the nodal point of tensor field lines diverging in any direction.
 Remark, that in the middle image there is a light V-shaped ridge on the left on blue background which is outshined by the center spot because of its orders of larger magnitude.
@@
   \begin{minipage}{0.3\textwidth}
+  \centering
   \includegraphics[height=\textwidth]{img_new/brain_cross-section.PNG}
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img_new/brain_volume.PNG}
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img_new/brain_volume2.PNG}
@@
 This example emphasizes, that ridges can also be yielded from rapid changes in anisotropy and are hence in this case, not directed and thus non directionally dependent.
-Also we take home from this example, that the thresholding is in fact necessary to form a reasonable local normalization as a posteriori verification.
+As a posteriori verification, we take home from this example, that the thresholding is in fact necessary to form a reasonable local normalization, because otherwise we would emphasize linear anisotropy resulting from noise.
 

### 2019-07-07 thesis-template/dummy.tex
update

--- 2019-06-25 thesis-template/dummy.tex
@@
 \usepackage{caption}
-
+\usepackage{subfig}
+%\usepackage{subfig}
+\newsubfloat{figure}
 \newcommand{\source}[1]{\vspace{-3pt} \caption*{ Source: {#1}} }
@@
 \centering
-  \begin{minipage}{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_drain_alt.png}
-    \label{grada)}
-    \caption*{a) LTG($90^\circ$) }
+  \begin{minipage}{0.3\textwidth}
+  \centering
+  \includegraphics[width=\textwidth]{img_new/brain_surf.PNG}
+     \label{a)}
+      a)
   \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/drain_alt_gradient.png}
-    \label{gradb)}
-    \caption*{b) gradient for a)}
+  \begin{minipage}{0.3\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{img_new/brain_volume2.PNG}
+    \label{b)}
+    b)
   \end{minipage}
-  \caption{``drain'' test field}
-  \label{grad)}
+  \begin{minipage}{0.3\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{img_new/brain_cross-section.PNG}
+    c)
+    \label{b)}
+  \end{minipage}
+  \caption{\enquote{brain} test field: a) LTG($0^\circ$), b) LTG volume, c) cross-section}
+\label{heart-ftle}
 \end{figure}
 
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+  \centering
+  \includegraphics[height=\textwidth]{img/inverse29.png}\label{1a}
+    a)
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[height=\textwidth]{img/inverse51.PNG}\label{2a}
+	b)
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[height=\textwidth]{img/inverse101.PNG}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{ \protect a) some text 
+  \protect b) some other text}
+\label{res_ftle}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Alternative: put content in separate files
+% Check the difference between including these files using \input{filename} and \include{filename} and see which one you like better
+%\chapter{Einleitung}\label{intro}
+%\input{introduction}
+%
+%\chapter{Voraussetzungen}\label{bg}
+%\input{background}
+
 \end{document}

### 2019-07-07 thesis-template/msc-thesis.tex
update

@@
 \end{titlepage}
+\newpage
+\thispagestyle{plain} % empty
+\mbox{}
 
@@
 tensor field representations.
-Tensor fields, are commonly associated with stress distributions (cf.
-Cauchy-stress tensor) in 2D/3D grids, but also have other meanings in various fields.
-By means of global illumination techniques, we motivate a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in anisotropic media.
+Tensor fields, are commonly associated with stress distributions (cf.~Cauchy-stress tensor) in 2D/3D grids, but also have many other meanings in various fields.
+We present the concept of applying techniques from the field of global illumination to tensor field analysis.
+Thereby, we interpret the tensors as a conductivity property for light transport.
 As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles, and is able to approximate light distributions for given light source position(s) and direction(s).
@@
 \begin{itemize}
-	\item a light transport model (propagation scheme) following basic but crucial physical principles for tensor field visualization interpreting tensors as light transmission properties,
+	\item a light transport model (propagation scheme) following basic but crucial physical principles,
 	\item visualization of the propagation results in $2D$ polar profiles and scalar fields,
-	\item a FTLE-related approach called light transport gradient (LTG) for visualizing key structures (ROIs), namely LCS in $2D$ second-order tensor fields, and
-	\item application of our approach to real data involving brain and heart datasets.
+	\item application of this model for tensor field visualization interpreting tensors as light transmission properties,
+	\item a FTLE (Finite-time Lyapunov exponents)-related approach called light transport gradient (LTG) for visualizing key structures, namely LCS (Lagrangian coherent structures) in $2D$ second-order tensor fields, and
+	\item application of our approach to both synthetic and real data involving brain and heart datasets.
 \end{itemize}
@@
 Tensors occur most commonly in medical, scientific and engineering applications, making them a frequently used mathematical entity in science.
-In general, a tensor is a $o$-D generalization of scalar, vector and matrix and appears, e.g., as Jacobian matrix (spatial gradient) in vector (flow) field visualization or as stress tensor in solid and fluid continuum mechanics.
-Basically, tensors describe compressions, tensions, rotations and volume changes in both solid and fluid material, as illustrated in Fig.
-\ref{ToS}.
+In general, a tensor is a order-$o$ generalization of scalar, vector and matrix and appears, e.g., as the Jacobian matrix (gradient in vector fields), or as stress tensor in solid and fluid continuum mechanics.
+Basically, tensors describe compressions, tensions, rotations and volume changes in both solid and fluid material, as illustrated in Fig.~\ref{ToS}.
 While most techniques focus on symmetric tensor field visualization (like glyphs and tensor field lines), some other recent works address the problem of asymmetric tensor field visualization.
-The issue with asymmetric (non-symmetric) or therein included antisymmetric (skew-symmetric) tensors is, that they do not always yield real eigenvalues, which is needed to determine an eigenbasis to set up glyphs, hyperstreamlines or tensor field lines (classically).
+The issue with asymmetric (non-symmetric) or therein included antisymmetric (skew-symmetric) tensors is, that they do not always yield real eigenvalues, which is needed to determine a real eigenbasis to set up glyphs, hyperstreamlines or tensor field lines.
 Nevertheless, there has been extensive work in this domain as well.
 Classically, the issue with asymmetric tensor fields is, that they do not yield real eigenvalues in general, which is needed to set up glyph representations.
-In this work, we will use the $\mathbf{U}$-matrix obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA principal ellipsoid representing the tensor transformation as suggested by Moler \cite{moler} to state another alternative of processing asymmetric tensors.
-2D tensor field visualization is useful for cases where only two-dimensional data is available, such as satellite remote sensing, particle imagery velocimetry (PIV) and earthquake data \cite{laramee}.
+In this work, we will use the matrix decomposition obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA principal ellipsoid representing the tensor transformation as suggested by Moler \cite{moler} to state another alternative of processing asymmetric tensors.
+2D tensor field visualization is useful for cases where only two-dimensional data is available, such as satellite remote sensing, particle imagery velocimetry and earthquake data \cite{laramee}.
 In their field of work, geologists and medical scientists are struggling with complicated number arrays which seem almost impossible to interpret in their raw, unprocessed form.
 Nevertheless, comprehensive tensor field visualization is aiding scientists in their everyday lifes understanding complex data loads, such as tensor fields.
-To cope with the arising problem of tensor field visualization, we can only try to invent new methods, continuously evaluating them for suitability for certain domains of science, to provide our techniques at best, while considering them for the right applications.\\
+To cope with the arising problem of tensor field visualization, one needs to advance and extend respective techniques.
 
@@
   \captionsetup{font={footnotesize,bf,it}}
-  \caption*{src: \url{https://upload.wikimedia.org/wikipedia/commons/e/eb/DIFFERENT_TYPES_OF_STRESS.png}}	
+  \caption*{source: \url{https://upload.wikimedia.org/wikipedia/commons/e/eb/DIFFERENT_TYPES_OF_STRESS.png}}	
   \label{ToS}
@@
   \captionsetup{font={footnotesize,bf,it}}
-  \caption*{src: \url{http://www2.cs.uh.edu/~chengu/Teaching/Fall2018/Lecs/Lec17.pdf}}	
+  \caption*{source: \url{http://www2.cs.uh.edu/~chengu/Teaching/Fall2018/Lecs/Lec17.pdf}}	
   \label{streamlines}
@@
 This implies, e.g., clustering, structuring or projection techniques.
-When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor - medical resonance imaging) \cite{hlawatsch}, when measuring water molecule diffusion characteristics within tissue or in simulated (or measured) stress tensor fields from solid and fluid mechanics.
-In these named domains of science the problem arises, that tensor fields can be generated (simulated) or measured with physical/mathematical background, but not be regarded intuitively.
+When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor - medical resonance imaging) \cite{hlawatsch}, when measuring water molecule diffusion characteristics within tissue or in simulated or measured stress tensor fields from solid and fluid mechanics.
+In these named domains of science the problem arises, that tensor fields can be generated or measured with physical/mathematical background, but not be regarded intuitively.
 Necessity begets ingenuity as always, which lets visualization methods and systems emerge in this context.
-Concurrently, the problem setting of this work encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids, commonly consisting of second-order tensors (cf.
-Cauchy stress tensor).
+Concurrently, the problem setting of this work encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids, commonly consisting of second-order tensors (cf.~Cauchy stress tensor).
 In general, these grids span a spatial domain with sample positions indicated by vertices or cell centers in the dual grid, depending on the mode of interpretation.
@@
 Congruently, the most basic objective of this work is to design a simple light propagation scheme following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation, for tensor field visualization.
-Of course, the propagation scheme will need some kind of stimulus which can be defined through symbolic functions (, e.g., $r(\omega)=1$) by the user which we sample (discretize) as a primary step.
+Of course, the propagation scheme will need some kind of stimulus which can be defined through symbolic functions (e.g.: $r(\omega)=1$) by the user which we sample (discretize) as a primary step.
 This way, the approach will operate as a light transport visualization technique for light distributions given in polar coordinates, provided the user set an initial profile into the grid.
@@
 These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature, such as moraines in glaciers or vortices in flow fields.
-Commonly, in vector field visualization, LCS separate regions of different flow behavior, but in our case they separate regions/subdomains of varying directional distributions (manifesting in diverging/converging tensor field lines $\rightarrow$ compressions, tensions, roations or varying tensor magnitude $\rightarrow$ volume changes) represented by tensors, as we use our tensor-controlled light transport visualization (propagation scheme) as a basis or framework to compute the FTLE field.
+Commonly, in vector field visualization, LCS separate regions of different flow behavior, but in our case they separate regions/subdomains of varying directional distributions (manifesting in diverging/converging tensor field lines leading to compressions, tensions, roations or varying tensor magnitude causing volume changes) represented by tensors, as we use our tensor-controlled light transport visualization (propagation scheme) as a basis or framework to compute the FTLE field.
 Consequently, we propose another method to compute an FTLE-like field on tensor fields.
@@
 The greater goal of this work is therefore the encoding of a tensor field into certain abstracted representations and the interpretation of meaningful results evaluating the proposed techniques for possible uses.
-The designed approach could ,e.g.
-, be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations.
-A classical example for tensor field visualization is plotting hyperstreamlines following major/minor eigenvectors, as depicted in Fig.
-\ref{streamlines}, which features a degenerate point (purple), a wedge (red) and a trisector (green).
+The designed approach could, e.g., be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations.
+A classical example for tensor field visualization is plotting hyperstreamlines following major/minor eigenvectors, as depicted in Fig.~\ref{streamlines}, which features a degenerate point (purple), a wedge (red) and a trisector (green).
 
@@
 For the rest of this work, we will give a quick overview on the structure and outline.
-Next, in chap.
-\ref{chap:grundlagen}, we will capture important fundamentals like tensors, polar coordinates, PCA in the context of singular value decomposition (generalization of eigenvalue decomposition), FTLE and GT datasets.
+Next, in Chap.~\ref{chap:grundlagen}, we will capture important fundamentals like tensors, polar coordinates, PCA in the context of singular value decomposition (generalization of eigenvalue decomposition), FTLE and GT (ground truth) datasets.
 Also we will discuss the most relevant related work in the domain of tensor field visualization and also in the domain of global illumination (indirect illumination rendering).
-Subsequently, in chap.
-\ref{chap:main}, we will propose the main contributions and methods of this work and document the results and evaluation in chap.
-\ref{chap:eval}.
-At last, a compact summary, conclusion and outlook will be given in chap.
-\ref{chap:concl}.\\
+Subsequently, in Chap.~\ref{chap:main}, we will propose the main contributions and methods of this work and document the results and evaluation in Chap.~\ref{chap:eval}.
+At last, a compact summary, conclusion and outlook will be given in Chap.~\ref{chap:concl}.\\
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Fundamentals}
-\label{chap:grundlagen}
-
-In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and we will capture tensor fields, polar coordinates, PCA, FTLE, GT datasets and related work in that sense.
-
-%Die ersten paar Abschnitte in diesem Kapitel f?hren in die Grundlagen
-%zur Arbeit ein. Das k?nnen beispielsweise Grundlagen zu Netzwerken
-%oder zur Informationsextraktion sein.
-
-\section{Tensor Fields}
-\label{sec:tensors}
-Now what is a tensor? A tensor is a $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
-\begin{table}[!b]
-\centering
-\caption{Tensor Shapes}
- \begin{tabular}{c|c|c|c|c|c|c}
-\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
-\hline 
-\textbf{shape} & scalar & vector & matrix & ``3D matrix'' & ... & $\mathop{o}$D matrix\\ 
-\end{tabular}
-\label{tensor}
- \end{table} 
-In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ in space (the dimension of the row and column vectors), which is equal to the rank of the tensor for full rank (in other words the number of column elements n for a $n{\times}n$-matrix).
-A tensor is considered non-degenerate if all eigenvalues $\lambda_i>\lambda_{i+1}$ manifesting in significant anisotropy.
-Tensors follow certain transformation rules which are defined for covariant/contravariant tensors, which indicate the change characteristics under a basis change.
-Contravariant tensors do not change under a change of basis, whereas covariant tensors are not invariant w.r.t.
-basis changes.
-Tensors are most commonly obtained and defined from mathematical transformation equations.
-Hence, any $oD$ number array could be a tensor, but the definition holds only if the transformation rules apply to these $oD$ number arrays (in a sense of mathematical/formal interpretation).
-This is mostly true for component indexed $oD$ number arrays following matrix multiplication and transformation rules found in physics and math manifesting in calculation rules like matrix multiplication.
-Thus, a tensor in general is a $o$D number array arranged in meaningful (transformation) order, to put it simple.
-In fact, tensors themselves are interpreted as transformations which reflect an incoming direction vector (measuring, e.g., the stress in that direction) to an outgoing (stress) vector.
-To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities are not simple numbers or vectors.
-They occur in form of a tensors, and thus as entities or quantities in physics.
-That is, at each point (typically a Euclidean space or manifold) in space (indexed by the grid) there is a whole distribution of directions, e.g., stresses (elastic and viscous), which needs to be characterized and represented by a single tensor each.
-Collectively, defined at each point in space, these tensors form a tensor field.
-\paragraph{stress tensor}
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.6\textwidth]
-    {img/stress_tensor.png}
-  }
-  \caption{Cauchy stress Tensor}
-  \captionsetup{font={footnotesize,bf,it}}
-  \caption*{src: \url{https://en.wikipedia.org/wiki/Cauchy_stress_tensor\#/media/File:Components_stress_tensor_cartesian.svg}}
-  \label{cauchy}
-\end{figure}
-
-The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig.
-\ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$, arranged in row-major order.
-It appears in similar form as viscous stress tensor in fluid mechanics:
-\begin{align*}
-\boldsymbol{\sigma}
- = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
-\boldsymbol{T}^{(\mathbf{e}_2)} \\
-\boldsymbol{T}^{(\mathbf{e}_3)} \\
-\end{matrix}}\right] =
-\left[{\begin{matrix}
-\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
-\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
-\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
-\end{matrix}}\right]  = \left[{\begin{matrix}
-\sigma _x & \tau _{xy} & \tau _{xz} \\
-\tau _{yx} & \sigma _y & \tau _{yz} \\
-\tau _{zx} & \tau _{zy} & \sigma _z \\
-\end{matrix}}\right] =
-\begin{bmatrix}
-\sigma _{11} & \sigma _{12} & \sigma _{13} \\
-\sigma _{21} & \sigma _{22} & \sigma _{23} \\
-\sigma _{31} & \sigma _{32} & \sigma _{33} 
-\end{bmatrix}.
-\end{align*}
-These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in plane normal directions ${x,y,z}$ ($3^2=9$ individual numbers).
-Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space.
-The stress tensor itself can be interpreted as a transformation, which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$.
-The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent (kind of a Euclidean distance of the stress components).
-The tensor itself is a complicated representation and needs to be projected onto its principal axes to lower its dimensionality, as it is the case for principal component analysis.
-It is possible to find $2$ principal stress directions in $2D$ through PCA (see sect.
-\ref{sec:pca} below), for which there are no shear stresses existent in equilibrium.
-These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour.
-The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''.
-So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig.
-\ref{hat} in sect.
-\ref{sec:pca}.
-Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle.
-It is therefore somehow representing the distortion of an absolutely isotropic object (unit circle) by whatever the underlying affine tensor transformation effects (e.g.
-stretching, shearing, mirroring, scaling, rotating).
-Let us consider it as some lower dimensional, more intuitive representation of the tensor which we were left with in the first place.
-To put it all together, a tensor is a mathematical transformation which is encoded inside of an abstract $oD$ number array, which comes in need, whenever we have a whole distribution of directions manifesting itself inside of a tensor (matrix) transformation.
-
-\section{Polar Coordinates}
-\label{sec:polar}
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.8\textwidth]
-    {img/polarkoordinaten.png}
-  }
-  \caption{polar space}
-  \captionsetup{font={footnotesize,bf,it}}
-   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:Ebene_polarkoordinaten.PNG}}
-  \label{polar}
-\end{figure}\vskip 3pt
-Mathematically speaking, a polar space is a system, which maps each angle $\omega$ a magnitude $r(\omega)$ through a symbolic function.
-The conversion formula from polar (given a radius $r$ and angle $\omega$) to Cartesian coordinates is depicted in Fig.
-\ref{polar} and is given by:
-\begin{align*}
-	x &= r \cos \omega, \\
- 	y &= r \sin \omega.
-\end{align*}
-Note, that we switched angle $\phi$ to $\omega$ for reasonable discrimination to the radiant intensity. The inverse direction (given coordinates $x,y$) is a bit more complicated:
-\begin{align*}
-	r &= \sqrt{x^2 + y^2} \\
-\omega &= \operatorname{atan2}(y, x)
-\end{align*}
-The polar domain is a circular axis version of the domain $[0..2\pi]$, with negative magnitudes in $r$ shifted by 180 degrees (point reflected) as this is the formal behaviour defined and required by the conversion formulas.
-
-\section{Principal Component Analysis}
-\label{sec:pca}
-Note, that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following.
-Also keep in mind, that we interpret a matrix or tensor as affine transformation.
-For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent.
-Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted.
-The anisotropy can be represented by, e.g., ellipsoid glyphs.
-
-\begin{figure}[!t]
-\hskip 22pt
-  \begin{minipage}[!t]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/HAT.png}
-   \caption{2D principal component ellipsoid\/ glyph (eigenframe)}
-   \label{hat}
-   \captionsetup{font={footnotesize,bf,it}}
-   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:HAT-ellipse0.svg}}
-  \end{minipage} \hskip 40pt
-  %\raggedright
-  \begin{minipage}[!t]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/glyphs.png}
-    \caption{3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
-    \label{3Dglyphs}
-    \captionsetup{font={footnotesize,bf,it}}
-   \caption*{src: \url{https://slideplayer.com/slide/5291764/}}
-  \end{minipage}
-\end{figure}
-
-Principal Component Analysis is an algorithm, which is able to capture the principal components (directions) in stochastic data.
-But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling.
-Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices.
-For (skew-) symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf.
-\cite{kieburg}).
-Note, that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases (cf.
-\cite{moler}).
-Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling (classically: $\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$).
-We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive).
-Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig.
-\ref{HAT} for any kind of transformation matrix.
-The singular value decomposition is defined as follows:
-\begin{align*}
-	\mathbf{A} = \mathbf{U} \boldsymbol{\Sigma} \mathbf{V}
-\end{align*}
-where
-\begin{itemize}
-	\item $\mathbf{U}$ is an $m ? m$ unitary matrix
-	\item $\boldsymbol{\Sigma}$ is a diagonal matrix {$m ? n$} with positive real numbers on the diagonal
-	\item $\mathbf{V}$ is an {$n ? n$} unitary matrix, $\mathbf{V}^*$ is its conjugate transpose
-\end{itemize}
-
-In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues (\cite{mroz}) ordered decreasingly by absolute value (corresponding to singular value order).
-The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses.
-We do this decomposition analysis once for each matrix and then store the precomputed results in a grid.
-The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step, which makes it a basic and central concept within the scope of this work.
-Within this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition.
-This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor.
-We will then use global-illumination techniques on top to generate a light transport flow map, which permits us somehow to forward the intensities which are given as predefined initialization values in form of symbolic functions defined by the user.
-This kind of light transport propagation scheme will also allow us to compute an FTLE-like field on the tensor fields.
-
-\section{Finite-Time Lyapunov Exponent}
-\label{sec:ftle}
-The Finite-Time Lyapunov Exponent is a measure, developed to analyze time-dependent dynamical systems for their separation abilities concerning massless tracer particles.
-We imagine a flow field, i.e., a vector field which in turn represents the effective amount or concentration shifted into a particular direction at each point in space.
-It could be induced by e.g.
-concentration or height gradients.
-This flow field is considered a time-dependent dynamical system, since it usually holds a time-component $t$ and thus maps each initial position an aged (altered) position: $\mathbf{\upvarphi}(\mathbf{x}) : \mathbf{x}(t_0) \longmapsto \mathbf{x}(t_0+T)$.
-We can now imagine placing two close by particles in distance $\delta$ into a flow field and run a time dependent simulation.
-Remark that this corresponds to the Lagrangian view in vector fields, which travels with the particles (instead of taking the position of a resting observer - Eulerian view):
-
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.6\textwidth]
-    {img/ftle_sep.pdf}
-  }
-  \caption{FTLE: separation of massless tracer particles}
-  \captionsetup{font={footnotesize,bf,it}}
-  \caption*{src: Skript Prof. Sadlo, Scientific Visualization SS2017, Heidelberg University}
-  \label{ftle_sep}
-\end{figure}
-
-We observe particle trajectories diverging w.r.t. the central axis ending up in distance $\Delta$. The FTLE measure for this single example (point in space) is then defined as:
-
-\begin{align*}
-	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
-\end{align*}
-
-Alternatively, an FTLE can also be computed from a flow map's gradient, which is more convenient to compute in general:
-\begin{align*}
-	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \lvert\lvert\nabla\mathbf{\upvarphi(x)}\rvert\rvert _2,
-\end{align*}
-whereas $\lvert\lvert A\rvert\rvert _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$.
-This will yield a map, which responds most where path lines diverge w.r.t.
-to a central axis.
-We consider an example which detects FTLE ridges inside of the cross-section of a double gyre.
-Note, that this example considers single trajectories diverging implicitly:
-\begin{figure}[!t]
-\hskip 22pt
-  \begin{minipage}[!t]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/diverge.png}
-   \caption{diverging streamlines for double gyre}
-   \label{diverge}
-   \captionsetup{font={footnotesize,bf,it}}
-   \caption*{src: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dg_3traj.png}}
-  \end{minipage} \hskip 40pt
-  %\raggedright
-  \begin{minipage}[!t]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_gyre.png}
-    \caption{FTLE field for double gyre}
-    \label{ftle_gyre}
-    \captionsetup{font={footnotesize,bf,it}}
-   \caption*{src: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dle_incT.png}}
-  \end{minipage}
-\end{figure}
-Fig \ref{diverge} and \ref{ftle_gyre} depict a scenario of diverging streamlines for a double gyre snapshot in time (yielding a time-independent dynamical system).
-Fig.
-\ref{ftle_gyre} features a high ridge on the margins (borders) and in particular at the center of the domain.
-This region of interest (ROI) is considered to indicate the maximum relative local divergence of tensor field lines, by the approach.
-Ridges also represent Lagrangian Coherent Structures, which are found in nature in vortices of correlating flow behavior.
-We could imagine to introduce this process for a whole distribution of trajectories, which would implement an FTLE for uncertain vector fields in a sense, that whole propability distributions of trajectories could be heeded.
-An overview about existing visualization techniques for uncertain data is given by Griete \& Schuhmann \cite{griete}.
-We require that the trajectories can cross each other unaffectedly, i.e., the superposition principle is supposed to be respected for the magnitudes.
-But behold, is this behavior not somehow familiar?
-If we elaborately think this through, we can come up with the idea that this is the behavior of light propagation in vacuum.
-Light paths are physically allowed to cross each other with no influence exerted on each other at all.
-Consequently, we choose to model the distributions of trajectories as light intensity profiles.
-That is, we interpret the polar profiles as a propability density function (PDF) for particle trajectories heading into this direction.
-To put it into context, our main contribution method LTG uses the FTLE as an inspiration or template to design a similar approach for tensor field visualization.
-The main difference to the template FTLE is that we follow eigenvectors instead of flow vectors and that we use whole distributions of trajectories, as a tensor field suggests whole distributions of directions, accordingly.
-
-\section{GT Datasets / Methods}
-\label{sec:gt}
-\paragraph{Glyphs and Tensor Field Lines}
-%\begin{figure}[!t]
-%  \centering
-% {
-%    \includegraphics[width=0.4\textwidth]
-%    {img/tensorfieldlines.png}
-%  }
-%  \caption{tensor field lines: bold lines; ellipsoids: ellipsoid glyphs}
-%  \label{scheme}
-%\end{figure}
-Glyphs are derived as a basis for the weighting profiles of the tensors in the following propagation scheme.
-They were chosen because it is an easy-to-use (understand) concept, which can also be interpreted nicely.
-In fact, these representations are used for the principal component analysis of direction vectors for the tensor field line extrapolation (integration) and also as transmission profiles for the light propagation scheme in a subsequent step.
-We can again emphasize glyphs as a crucial, central concept within this work.
-Tensor field lines (the type of tensor streamline analog that does not introduce artificial intertia) were implemented for ground truth (GT) data generation purposes.
-For example, when designing a FTLE-like field computation technique, we will need to know when tensor field lines diverge/converge.
-Examples for glyphs (ellipsoids) and tensor field lines (thin lines) are depicted in Fig.
-\ref{test-fields}.
-
-\paragraph{Test Data (Generation)}
-We used python's numpy library to generate predefined, synthetic test cases and caught two real data example from \url{TensorVis.org} and \url{https://www.cs.purdue.edu/homes/cs530/projects/project5.html}, which needed to be pre-processed through slicing and subsequent cropping to obtain $2D$ samples ($2\times2$) matrices).
-These test fields can be grouped into symmetric and asymmetric ones:\\
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/corner-TFL.png}
-    \caption*{``corner''}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/spiral-TFL.png}
-    \caption*{``spiral''}
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/drain_old.png}
-    \caption*{``V''}
-    \label{b)}
-  \end{minipage}
-\caption{Named test fields}
-\label{test-fields}
-\end{figure}
-
-\textbf{Asymmetric Tensor Fields}
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/drain(alt).png}
-    \caption*{a) glyphs}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/drain(alt)-TFL.png}
-    \caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{``drain'' test field}
-\label{drain}
-\end{figure}
-
-The ``drain''-Testfield, depicted in Fig.
-\ref{drain} is generated to have a kind of proof-of-concept (POC) for the FTLE-related approach.
-It poses raw, bare diverging tensor field lines without any other feature modulated on top.
-Thus, it can be expected to induce a high FTLE response, since it is an interpretable ground truth test and is considered to be a basic trigger/stimulus for the approach with no side effects introduced.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/rings.png}
-    \caption*{a) glyphs}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-   \centering
-    \includegraphics[width=0.8\textwidth]{img/tensorfieldlines.png}
-	\caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{``rings'' test field}
-\label{rings}
-\end{figure}
-The ``rings'' test field, depicted in Fig. \ref{rings} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/spiral.png}
-    \caption*{a) glyphs}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/spiral-TFL.png}
-    \caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{``spiral''-test field}
-\label{spiral}
-\end{figure}
-The ``spiral'' test field, depicted in Fig. \ref{spiral} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/inverse.png}
-    \caption*{a) glyphs}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/inverse-TFL.png}
-    \caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{``inverse'' test field}
-\label{inverse}
-\end{figure}
-With the \enquote{inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/gyre.png}
-    \caption*{a) glyphs}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/gyre-TFL.png}
-    \caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{a) ``gyre''-test field, b) tensor field lines for a)}
-\label{gyre}
-\end{figure}
-The test field ``gyre'', depicted in Fig \ref{gyre}, constitutes a sustained turn over the grid representing the redirection or forwarding along domain corners.
-
-\textbf{Symmteric Tensor Fields}
-We also aim to use a real data example of the $3D$ DTI-MRI diffusion tensor of the brain for a male subject at the age of 27 and another one featuring a human heart.
-As we visualize 2D symmetric second order tensor fields, we need to slice the dataset and crop the $2{\times}2$ matrices omitting the $z$-component of the original $3{\times}3$ matrices.
-Fig.
-\ref{brainglyphs} depicts the glyph representation of the test dataset for slice index $15$.
-To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{brain} and \enquote{heart} scans, which can be observed in Fig.
-\ref{real1} and \ref{real2}.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/brain_org.png}
-    \label{a)}
-    \caption*{tensor magnitude}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/brain_cyl_glyphs.png}
-    \label{b)}
-    \caption*{cyclinder glyphs}
-  \end{minipage}
-\caption{Real data examples: \enquote{brain} dataset}
-\label{real1}
-\end{figure}
-Observing the \enquote{brain} dataset in Fig \ref{real1}, one may notice that it introduces a lot of variances to tensor magnitude and linear anisotropy.
-We choose this DTI-MRI scan as an example-par-excellence for real application data given in a reasonable resolution.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/heart_org.png}
-    \label{a)}
-    \caption*{tensor magnitude}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-      \includegraphics[width=0.8\textwidth]{img/heart_cyl_glyphs.png}
-    \label{b)}
-    \caption*{cyclinder glyphs}
-  \end{minipage}
-\caption{Real data examples \enquote{heart} dataset}
-\label{real2}
-\end{figure}
-When we look at this DT-MRI scan of a human heart we can notice the elongated shape, which we consider to represent the aorta and a part of the rest of the heart.
-If we consider the glyph representation, we again infer, that we incorporate lots of anisotropic regions with varying tensor magnitudes here.
 
@@
 
-\paragraph{Global Illumination Methods}
+\section{Global Illumination Methods}
 Since we use a specific category of global illumination methods as a basis, we decide to only discuss lattice based methods which work, corresponding to their name, on a grid.
@@
 The radiative transfer equation is discretized spatially and in the angular domain.
-The radiance distribution (PDF) is stored in a grid and light is distributed and exchanged between neighboring volume elements which reduces the overall effort to local and independent computations.
+The radiance distribution is stored in a grid and light is distributed and exchanged between neighboring volume elements which reduces the overall effort to local and independent computations.
 Furthermore, there has been more recent work on describing the light propagation as a diffusion process within a simple photon transport model based on the Lattice-Boltzmann method \cite{geist}.
@@
 
-\paragraph{Symmetric Tensor Field Visualization}
+\section{Symmetric Tensor Field Visualization}
 Zheng \& and Pang \cite{pang&zheng} et al.
@@
 \\
-\paragraph{Asymmetric Tensor Field Visualization}
+\section{Asymmetric Tensor Field Visualization}
 Zheng and Pang \cite{pang&zheng} proposed the concept of dual eigenvectors and Zhang et al.
@@
 
+\chapter{Fundamentals}
+\label{chap:grundlagen}
+
+In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and we will capture tensor fields, polar coordinates, PCA, FTLE, GT datasets and related work in that sense.
+
+%Die ersten paar Abschnitte in diesem Kapitel f?hren in die Grundlagen
+%zur Arbeit ein. Das k?nnen beispielsweise Grundlagen zu Netzwerken
+%oder zur Informationsextraktion sein.
+
+\section{Tensor Fields}
+\label{sec:tensors}
+Now what is a tensor? A tensor is a $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
+\begin{table}[!b]
+\centering
+\caption{Tensor Shapes}
+ \begin{tabular}{c|c|c|c|c|c|c}
+\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
+\hline 
+\textbf{shape} & scalar & vector & matrix & ``3D matrix'' & ... & order-$o$ matrix\\ 
+\end{tabular}
+\label{tensor}
+ \end{table} 
+In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ in space (the dimension of the row and column vectors), which is equal to the rank of the tensor for full rank (in other words the number of column elements $n$ for a $n{\times}n$-matrix).
+A tensor is considered non-degenerate if all eigenvalues $\lambda_i>\lambda_{i+1}$ manifesting in significant anisotropy.
+Tensors follow certain transformation rules which are defined for covariant/contravariant tensors, which indicate the change characteristics under a basis change.
+Contravariant tensors do not change under a change of basis, whereas covariant tensors are not invariant w.r.t.
+basis changes.
+Tensors are most commonly obtained and defined from mathematical transformation equations.
+Hence, any order-$o$ number array could be a tensor, but the definition holds only if the transformation rules apply to these $o$-th order number arrays (in a sense of mathematical/formal interpretation).
+This is mostly true for component indexed order-$o$ number arrays following matrix multiplication and transformation rules found in physics and math manifesting in calculation rules like matrix multiplication.
+Thus, a tensor in general is a order-$o$ number array arranged in meaningful (transformation) order, to put it simple.
+In fact, tensors themselves are interpreted as transformations which reflect an incoming direction vector (measuring, e.g., the stress in that direction) to an outgoing (stress) vector.
+To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities are not simple numbers or vectors.
+They occur in form of a tensors, and thus as entities or quantities in physics.
+That is, at each point (typically a Euclidean space or manifold) in space (indexed by the grid) there is a whole distribution of directions, e.g., stresses (elastic and viscous), which needs to be characterized and represented by a single tensor each.
+Collectively, defined at each point in space, these tensors form a tensor field.
+\paragraph{stress tensor}
+\begin{figure}[!t]
+  \centering
+ {
+    \includegraphics[width=0.6\textwidth]
+    {img/stress_tensor.png}
+  }
+  \caption{Cauchy stress Tensor}
+  \captionsetup{font={footnotesize,bf,it}}
+  \caption*{source: \url{https://en.wikipedia.org/wiki/Cauchy_stress_tensor\#/media/File:Components_stress_tensor_cartesian.svg}}
+  \label{cauchy}
+\end{figure}
+
+The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig.~\ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$, arranged in row-major order.
+It appears in similar form as viscous stress tensor in fluid mechanics:
+\begin{align*}
+\boldsymbol{\sigma}
+ = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
+\boldsymbol{T}^{(\mathbf{e}_2)} \\
+\boldsymbol{T}^{(\mathbf{e}_3)} \\
+\end{matrix}}\right] =
+\left[{\begin{matrix}
+\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
+\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
+\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
+\end{matrix}}\right]  = \left[{\begin{matrix}
+\sigma _x & \tau _{xy} & \tau _{xz} \\
+\tau _{yx} & \sigma _y & \tau _{yz} \\
+\tau _{zx} & \tau _{zy} & \sigma _z \\
+\end{matrix}}\right] =
+\begin{bmatrix}
+\sigma _{11} & \sigma _{12} & \sigma _{13} \\
+\sigma _{21} & \sigma _{22} & \sigma _{23} \\
+\sigma _{31} & \sigma _{32} & \sigma _{33} 
+\end{bmatrix}.
+\end{align*}
+These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in plane normal directions ${x,y,z}$ ($3^2=9$ individual numbers).
+Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space.
+The stress tensor itself can be interpreted as a transformation, which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$.
+The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent (kind of a Euclidean distance of the stress components).
+The tensor itself is a complicated representation and needs to be projected onto its principal axes to lower its dimensionality, as it is the case for principal component analysis.
+It is possible to find $2$ principal stress directions in $2D$ through PCA (see Sect.~\ref{sec:pca} below), for which there are no shear stresses existent in equilibrium.
+These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour.
+The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''.
+So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig.~\ref{hat} in Sect.~\ref{sec:pca}.
+Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle.
+It is therefore somehow representing the distortion of an absolutely isotropic object (unit circle) by whatever the underlying affine tensor transformation effects (e.g.: stretching, shearing, mirroring, scaling, rotating).
+Let us consider it as some lower dimensional, more intuitive representation of the tensor which we were left with in the first place.
+To put it all together, a tensor is a mathematical transformation which is encoded inside of an abstract $oD$ number array, which comes in need, whenever we have a whole distribution of directions manifesting itself inside of a tensor (matrix) transformation.
+
+\section{Polar Coordinates}
+\label{sec:polar}
+\begin{figure}[!t]
+  \centering
+ {
+    \includegraphics[width=0.8\textwidth]
+    {img/polarkoordinaten.png}
+  }
+  \caption{polar space}
+  \captionsetup{font={footnotesize,bf,it}}
+   \caption*{source: \url{https://de.wikipedia.org/wiki/Datei:Ebene_polarkoordinaten.PNG}}
+  \label{polar}
+\end{figure}\vskip 3pt
+Mathematically speaking, a polar space is a system, which maps each angle $\omega$ a magnitude $r(\omega)$ through a symbolic function.
+The conversion formula from polar (given a radius $r$ and angle $\omega$) to Cartesian coordinates is depicted in Fig.~\ref{polar} and is given by:
+\begin{align*}
+	x &= r \cos \omega, \\
+ 	y &= r \sin \omega.
+\end{align*}
+Note, that we switched angle $\phi$ to $\omega$ for reasonable discrimination to the radiant intensity. The inverse direction (given coordinates $x,y$) is a bit more complicated:
+\begin{align*}
+	r &= \sqrt{x^2 + y^2}, \\
+\omega &= \operatorname{atan2}(y, x).
+\end{align*}
+The polar domain is a circular axis version of the domain $[0..2\pi]$, with negative magnitudes in $r$ shifted by 180 degrees (point reflected) as this is the formal behaviour defined and required by the conversion formulas.
+
+\section{Principal Component Analysis}
+\label{sec:pca}
+Note, that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following.
+Also keep in mind, that we interpret a matrix or tensor as affine transformation.
+For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent.
+Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted.
+The anisotropy can be represented by, e.g., ellipsoid glyphs.
+
+\begin{figure}[!t]
+\hskip 22pt
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/HAT.png}
+   \caption{2D principal component ellipsoid\/ glyph (eigenframe)}
+   \label{hat}
+   \captionsetup{font={footnotesize,bf,it}}
+   \caption*{source: \url{https://de.wikipedia.org/wiki/Datei:HAT-ellipse0.svg}}
+  \end{minipage} \hskip 40pt
+  %\raggedright
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/glyphs.png}
+    \caption{3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
+    \label{3Dglyphs}
+    \captionsetup{font={footnotesize,bf,it}}
+   \caption*{source: \url{https://slideplayer.com/slide/5291764/}}
+  \end{minipage}
+\end{figure}
+
+Principal Component Analysis is an algorithm, which is able to capture the principal components (directions) in stochastic data.
+But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling.
+Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices.
+For (skew-) symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf.~\cite{kieburg}).
+Note, that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases (cf.~\cite{moler}).
+Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling (classically: $\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$).
+We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive).
+Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig.~\ref{HAT} for any kind of transformation matrix.
+The singular value decomposition is defined as follows:
+\begin{align*}
+	\mathbf{A} = \mathbf{U} \boldsymbol{\Sigma} \mathbf{V}
+\end{align*}
+where
+\begin{itemize}
+	\item $\mathbf{U}$ is an $m\times m$ unitary matrix
+	\item $\boldsymbol{\Sigma}$ is a diagonal matrix {$m \times n$} with positive real numbers on the diagonal
+	\item $\mathbf{V}$ is an {$n\times n$} unitary matrix, $\mathbf{V}^*$ is its conjugate transpose
+\end{itemize}
+
+In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues (\cite{mroz}) ordered decreasingly by absolute value (corresponding to singular value order).
+The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses.
+We do this decomposition analysis once for each matrix and then store the precomputed results in a grid.
+The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step, which makes it a basic and central concept within the scope of this work.
+Within this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition.
+This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor.
+We will then use global-illumination techniques on top to generate a light transport flow map, which permits us somehow to forward the intensities which are given as predefined initialization values in form of symbolic functions defined by the user.
+This kind of light transport propagation scheme will also allow us to compute an FTLE-like field on the tensor fields.
+
+\section{Finite-Time Lyapunov Exponent}
+\label{sec:ftle}
+The Finite-Time Lyapunov Exponent is a measure, developed to analyze time-dependent dynamical systems for their separation abilities concerning massless tracer particles.
+We imagine a flow field, i.e., a vector field which in turn represents the effective amount or concentration shifted into a particular direction at each point in space.
+It could be induced by e.g.
+concentration or height gradients.
+This flow field is considered a time-dependent dynamical system, since it usually holds a time-component $t$ and thus maps each initial position an aged (altered) position: $\mathbf{\upvarphi}(\mathbf{x}) : \mathbf{x}(t_0) \longmapsto \mathbf{x}(t_0+T)$.
+We can now imagine placing two close by particles in distance $\delta$ into a flow field and run a time dependent simulation.
+Remark that this corresponds to the Lagrangian view in vector fields, which travels with the particles (instead of taking the position of a resting observer - Eulerian view):
+
+\begin{figure}[!t]
+  \centering
+ {
+    \includegraphics[width=0.6\textwidth]
+    {img/ftle_sep.pdf}
+  }
+  \caption{FTLE: separation of massless tracer particles}
+  \captionsetup{font={footnotesize,bf,it}}
+  \caption*{source: Skript Prof. Sadlo, Scientific Visualization SS2017, Heidelberg University}
+  \label{ftle_sep}
+\end{figure}
+
+We observe particle trajectories diverging w.r.t. the central axis ending up in distance $\Delta$. The FTLE measure for this single example (point in space) is then defined as:
+
+\begin{align*}
+	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
+\end{align*}
+
+Alternatively, an FTLE can also be computed from a flow map's gradient, which is more convenient to compute in general:
+\begin{align*}
+	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{\upvarphi(x)}\| _2,
+\end{align*}
+whereas $\| A\| _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$.
+This will yield a map, which responds most where path lines diverge w.r.t.
+to a central axis.
+We consider an example which detects FTLE ridges inside of the cross-section of a double gyre.
+Note, that this example considers single trajectories diverging implicitly:
+\begin{figure}[!t]
+\hskip 22pt
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/diverge.png}
+   \caption{diverging streamlines for double gyre}
+   \label{diverge}
+   \captionsetup{font={footnotesize,bf,it}}
+   \caption*{source: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dg_3traj.png}}
+  \end{minipage} \hskip 40pt
+  %\raggedright
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/ftle_gyre.png}
+    \caption{FTLE field for double gyre}
+    \label{ftle_gyre}
+    \captionsetup{font={footnotesize,bf,it}}
+   \caption*{source: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dle_incT.png}}
+  \end{minipage}
+\end{figure}
+Fig \ref{diverge} and \ref{ftle_gyre} depict a scenario of diverging streamlines for a double gyre snapshot in time (yielding a time-independent dynamical system).
+Fig.~\ref{ftle_gyre} features a high ridge on the margins (borders) and in particular at the center of the domain.
+This region of interest (ROI) is considered to indicate the maximum relative local divergence of tensor field lines, by the approach.
+Ridges also represent Lagrangian Coherent Structures, which are found in nature in vortices of correlating flow behavior.
+We could imagine to introduce this process for a whole distribution of trajectories, which would implement an FTLE for uncertain vector fields in a sense, that whole propability distributions of trajectories could be heeded.
+An overview about existing visualization techniques for uncertain data is given by Griete \& Schuhmann \cite{griete}.
+We require that the trajectories can cross each other unaffectedly, i.e., the superposition principle is supposed to be respected for the magnitudes.
+But behold, is this behavior not somehow familiar?
+If we elaborately think this through, we can come up with the idea that this is the behavior of light propagation in vacuum.
+Light paths are physically allowed to cross each other with no influence exerted on each other at all.
+Consequently, we choose to model the distributions of trajectories as light intensity profiles.
+That is, we interpret the polar profiles as a propability density function (PDF) for particle trajectories heading into this direction.
+To put it into context, our main contribution method LTG uses the FTLE as an inspiration or template to design a similar approach for tensor field visualization.
+The main difference to the template FTLE is that we follow eigenvectors instead of flow vectors and that we use whole distributions of trajectories, as a tensor field suggests whole distributions of directions, accordingly.
+
+\section{GT Datasets / Methods}
+\label{sec:gt}
+\subsection{Glyphs and Tensor Field Lines}
+%\begin{figure}[!t]
+%  \centering
+% {
+%    \includegraphics[width=0.4\textwidth]
+%    {img/tensorfieldlines.png}
+%  }
+%  \caption{tensor field lines: bold lines; ellipsoids: ellipsoid glyphs}
+%  \label{scheme}
+%\end{figure}
+Glyphs are derived as a basis for the weighting profiles of the tensors in the following propagation scheme.
+They were chosen because it is an easy-to-use (understand) concept, which can also be interpreted nicely.
+In fact, these representations are used for the principal component analysis of direction vectors for the tensor field line extrapolation (integration) and also as transmission profiles for the light propagation scheme in a subsequent step.
+We can again emphasize glyphs as a crucial, central concept within this work.
+Tensor field lines (the type of tensor streamline analog that does not introduce artificial intertia) were implemented for ground truth (GT) data generation purposes.
+For example, when designing a FTLE-like field computation technique, we will need to know when tensor field lines diverge/converge.
+Examples for glyphs (ellipsoids) and tensor field lines (thin lines) are depicted in Fig.~\ref{test-fields}.
+
+\subsection{Test Data (Generation)}
+We used python's numpy library to generate predefined, synthetic test cases and caught two real data example from \url{TensorVis.org} and \url{https://www.cs.purdue.edu/homes/cs530/projects/project5.html}, which needed to be pre-processed through slicing and subsequent cropping to obtain $2D$ samples ($2\times2$) matrices).
+These test fields can be grouped into symmetric and asymmetric ones:\\
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+    \includegraphics[height=\textwidth]{img/corner-TFL.png}
+    \caption*{``corner''}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+    \includegraphics[height=\textwidth]{img/spiral-TFL.png}
+    \caption*{``spiral''}
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+    \includegraphics[height=\textwidth]{img/drain_old.png}
+    \caption*{``V''}
+    \label{b)}
+  \end{minipage}
+\caption{Named test fields}
+\label{test-fields}
+\end{figure}
+
+\subsubsection{Tensor Fields}
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=0.8\textwidth]{img/drain(alt).png}
+    \caption*{a) glyphs}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.8\textwidth]{img/drain(alt)-TFL.png}
+    \caption*{b) tensor field lines for a)}
+    \label{b)}
+  \end{minipage}
+\caption{``drain'' test field}
+\label{drain}
+\end{figure}
+
+The ``drain''-Testfield, depicted in Fig.~\ref{drain} is generated to have a kind of proof-of-concept (POC) for the FTLE-related approach.
+It poses raw, bare diverging tensor field lines without any other feature modulated on top.
+Thus, it can be expected to induce a high FTLE response, since it is an interpretable ground truth test and is considered to be a basic trigger/stimulus for the approach with no side effects introduced.
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=0.8\textwidth]{img/rings.png}
+    \caption*{a) glyphs}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+   \centering
+    \includegraphics[width=0.8\textwidth]{img/tensorfieldlines.png}
+	\caption*{b) tensor field lines for a)}
+    \label{b)}
+  \end{minipage}
+\caption{``rings'' test field}
+\label{rings}
+\end{figure}
+The ``rings'' test field, depicted in Fig.~\ref{rings} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.8\textwidth]{img/spiral.png}
+    \caption*{a) glyphs}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.8\textwidth]{img/spiral-TFL.png}
+    \caption*{b) tensor field lines for a)}
+    \label{b)}
+  \end{minipage}
+\caption{``spiral''-test field}
+\label{spiral}
+\end{figure}
+The ``spiral'' test field, depicted in Fig.~\ref{spiral} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.8\textwidth]{img/inverse.png}
+    \caption*{a) glyphs}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.8\textwidth]{img/inverse-TFL.png}
+    \caption*{b) tensor field lines for a)}
+    \label{b)}
+  \end{minipage}
+\caption{``inverse'' test field}
+\label{inverse}
+\end{figure}
+With the \enquote{inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.8\textwidth]{img/gyre.png}
+    \caption*{a) glyphs}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.8\textwidth]{img/gyre-TFL.png}
+    \caption*{b) tensor field lines for a)}
+    \label{b)}
+  \end{minipage}
+\caption{a) ``gyre''-test field, b) tensor field lines for a)}
+\label{gyre}
+\end{figure}
+The test field ``gyre'', depicted in Fig \ref{gyre}, constitutes a sustained turn over the grid representing the redirection or forwarding along domain corners.
+
+\subsubsection{Symmteric Tensor Fields}
+We also aim to use a real data example of the $3D$ DTI-MRI diffusion tensor of the brain for a male subject at the age of 27 and another one featuring a human heart.
+As we visualize 2D symmetric second order tensor fields, we need to slice the dataset and crop the $2{\times}2$ matrices omitting the $z$-component of the original $3{\times}3$ matrices.
+Fig.~\ref{brainglyphs} depicts the glyph representation of the test dataset for slice index $15$.
+To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{brain} and \enquote{heart} scans, which can be observed in Figs.~\ref{real1} and \ref{real2}.
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=0.8\textwidth]{img/brain_org.png}
+    \label{a)}
+    \caption*{tensor magnitude}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=0.8\textwidth]{img/brain_cyl_glyphs.png}
+    \label{b)}
+    \caption*{cyclinder glyphs}
+  \end{minipage}
+\caption{Real data examples: \enquote{brain} dataset}
+\label{real1}
+\end{figure}
+Observing the \enquote{brain} dataset in Figs.
+\ref{real1}, one may notice that it introduces a lot of variances to tensor magnitude and linear anisotropy.
+We choose this DTI-MRI scan as an example-par-excellence for real application data given in a reasonable resolution.
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.8\textwidth]{img/heart_org.png}
+    \label{a)}
+    \caption*{tensor magnitude}
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \centering
+      \includegraphics[width=0.8\textwidth]{img/heart_cyl_glyphs.png}
+    \label{b)}
+    \caption*{cyclinder glyphs}
+  \end{minipage}
+\caption{Real data examples \enquote{heart} dataset}
+\label{real2}
+\end{figure}
+When we look at this DT-MRI scan of a human heart we can notice the elongated shape, which we consider to represent the aorta and a part of the rest of the heart.
+If we consider the glyph representation, we again infer, that we incorporate lots of anisotropic regions with varying tensor magnitudes here.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-This chapter will summarize the main contributions of this work studied and state targeted aims and goals as a first step in sect.
-\ref{sec:motiv}.
-Then we will have a short introduction to the implemented global illumination scheme, derived and simplified from a global illumination approach, invented for the Crytek GmbH in the context of global indirect illumination rendering, in sect.
-\ref{sec:scheme}.
+This chapter will summarize the main contributions of this work studied and state targeted aims and goals as a first step in Sect.~\ref{sec:motiv}.
+Then we will have a short introduction to the implemented global illumination scheme, derived and simplified from a global illumination approach, invented for the Crytek GmbH in the context of global indirect illumination rendering, in Sect.~\ref{sec:scheme}.
 Last, we will state the effectively implemented tensor field visualization methods: light transport visualization (directed propagation scheme yielding $2D$ scalar fields) in sects.
-\ref{sec:scheme} \& \ref{sec:trans} and light transport gradient (LTG: yielding $3D$ scalar fields) in sect.
-\ref{sec:ltg}, which are proposed as the invented tensor visualization techniques and hence as conceptual contributions to the domain of tensor field visualization.
+\ref{sec:scheme} \& \ref{sec:trans} and light transport gradient (LTG: yielding $3D$ scalar fields) in Sect.~\ref{sec:ltg}, which are proposed as the invented tensor visualization techniques and hence as conceptual contributions to the domain of tensor field visualization.
 
@@
 We then generate transmission profiles, which we sample in a preliminary step, from the eigensystems (principal component ellipsoids) of the tensors for each cell in the domain (field).
-This step is done, to provide an orientation for a lattice of crystal fibre structures (cf.
-optical fibres) to redirect the intensities in analogy to the anisotropy characteristics of the tensor field within a glyph representation, obtained from principal component analysis (PCA).
+This step is done, to provide an orientation for a lattice of crystal fibre structures (cf.~optical fibres) to redirect the intensities in analogy to the anisotropy characteristics of the tensor field within a glyph representation, obtained from principal component analysis (PCA).
 Our approach should be implicitly adaptive to any kind of 2D dataset or slices of 3D datasets.
-Thus, we require it to be compatible to handle any kind of matrix (symm./anti-symm.) and any kind of resolution of the field (fully-adaptive).
+Thus, we require it to be compatible to handle any kind of matrix (symmetric/anti-symmetric) and any kind of resolution of the field (fully-adaptive).
 At last, our ambition is to segment key locations (LCS/ridges) in the field with tensor field lines (hyperstreamlines) converging/diverging, which is necessarily the same (for bidirectional eigenvectors/singular vectors of tensors).
@@
 
-\section{Light Transport (Propagation) Scheme}
+\section{Light Propagation Scheme}
 \label{sec:scheme}
@@
 This polar profiles are first sampled into a vector with elements ranging over the interval $[0,2\pi]$.
-Polar profiles for different light sources are shown in Fig.
-\ref{polar}.
+Polar profiles for different light sources are shown in Fig.~\ref{polar}.
 \begin{figure}[!t]
@@
   \captionsetup{font={footnotesize,bf,it}}
-  \caption*{src: \url{http://www.optik.uni-erlangen.de/odem/vorlesung/ws1213/Photometrie_Geometrische_Optik.pdf}}
+  \caption*{source: \url{http://www.optik.uni-erlangen.de/odem/vorlesung/ws1213/Photometrie_Geometrische_Optik.pdf}}
   \label{polar}
@@
 But this would omit every propagation attenuation principles (any principles at all specifically), if one takes a closer look:
-\paragraph{Accumulation stage}
+\subsection{Accumulation stage}
 Each cone type (yellow and green) is an individual neighbor-dependent angular area.
@@
 
-\paragraph{Injection stage}
+\subsection{Injection stage}
  We place a cosine lobe scaled with the integrated (accumulated) energies in each of the 8 cone directions (angular domains), whereas outgoing integrated and summed energies are weighted by the linear combination weights (factors) in overlapping (green) areas to split up the contributions accordingly.
-The contribution areas are depicted in Fig.
-\ref{scheme}, the cosine lobes can be observed in Fig.
-\ref{iter} respectively.
+The contribution areas are depicted in Fig.~\ref{scheme}, the cosine lobes can be observed in Fig.~\ref{iter} respectively.
 This manner of propagation is inspired by the propagation scheme of Dachsbacher et al.
@@
 We measure the convergence by setting up an overall distribution error w.r.t to the previous iteration.
-Note that even though we propagate the energy directly through the diagonal cones (which forms a square initially), the 8 degrees of freedom introduce a circular (spherical) wave front after few iterations (for greater resolutions) as depicted in Fig.
-\ref{iter}.
+Note that even though we propagate the energy directly through the diagonal cones (which forms a square initially), the 8 degrees of freedom introduce a circular (spherical) wave front after few iterations (for greater resolutions) as depicted in Fig.~\ref{iter}.
 This occurs, because of a faster propagation attenuation which counteracts the larger angle of diagonal cones.
-There is a redistribution of intensities in the long term resulting in a, since the shared part effects a uniformly spaced isotropic propagation profile resulting in a circular torus-shaped distribution of intensities, as also observed in Fig.
-\ref{iter}.
+There is a redistribution of intensities in the long term resulting in a, since the shared part effects a uniformly spaced isotropic propagation profile resulting in a circular torus-shaped distribution of intensities, as also observed in Fig.~\ref{iter}.
 Also note, that this approach satisfies energy conservation and propagation attenuation principles and applies to the principles of light propagation in vacuum (by default).
 We infer, that we can assume to follow physical models of spherical waves.
-The verification of these principal requirements and specifications can be observed in the evaluation in chap.
-\ref{chap:eval}.
+The verification of these principal requirements and specifications can be observed in the evaluation in Chap.~\ref{chap:eval}.
 We first show the approach for a few iterations with a manual stop criterion to have a kind of imagination for the propagation process:
@@
 The propagation definition denotes the transformation from one grid (A-buffer) into another (B-buffer) through proper application of the \textit{propagate}-operator to each cell.
-We will extend this operator with a normalization correction in sect.
-\ref{sec:trans} in form of a precise mathematical formulation respectively.
+We will extend this operator with a normalization correction in Sect.~\ref{sec:trans} in form of a precise mathematical formulation respectively.
 For now, just imagine the approach in raw operating mode, with no tensor influence incorporated.
@@
 As a preliminary preparation, this profile is precomputed (sampled) for all discrete angular steps for each cell (, e.g., $360$) and stored for reuitilization as it is the case for the previously mentioned linear combination weights as well.
-The weighting scheme which we apply is depicted in Fig.
-\ref{weighting}.
+The weighting scheme which we apply is depicted in Fig.~\ref{weighting}.
 \begin{figure}[!t]
@@
   }
-  \caption{yellow: Intensity Profile, cyan: Transmission Profile (glyph Ellipsoid Eq.)}
+  \caption{Red: Intensity Profile, cyan: Transmission Profile (glyph Ellipsoid Eq.)}
   \label{weighting}
@@
 
-This permits us to scale an oriented and clipped cosine-lobe $\cos_+$ with the total radiant flux w.
-respect to its own total energy when integrated likewise.
+This permits us to scale an oriented and clipped cosine-lobe $\cos_+$ with the total radiant flux with respect to its own total energy when integrated likewise.
 A discrete cosine lobe never reaches its theoretical sum of $2$ exactly in the interval $[{-\frac{\pi}{2}},{\frac{\pi}{2}}]$.
@@
 \sum_{\cos_+} &= \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\cos_+(\omega)\mathop{d\omega} \approx 2,\\
-	\cos_k(\omega) &= \frac{\Phi_t}{\sum_{\cos_+}}cos_+(\omega-k\cdot\frac{\pi}{4})\hskip 20pt w. \ k\in[0,7].
+	\cos_k(\omega) &= \frac{\Phi_t}{\sum_{\cos_+}}cos_+(\omega-k\cdot\frac{\pi}{4})\hskip 20pt \mathit{with} \ k\in[0,7].
 \end{align}
 This specific profile is injected in form of a placement to the corresponding neighbor index $k$.
-It yields a cosine lobe, which has an overall energy of $\Phi_t$, if the point light is interpreted with a Lambert-emitter (cf.
-\cite{lindlein}) characteristic (as part or area element of a diffusely reflecting surface) since the intensity profile matches a cosine in this case.
+It yields a cosine lobe, which has an overall energy of $\Phi_t$, if the point light is interpreted with a Lambert-emitter (cf.~\cite{lindlein}) characteristic (as part or area element of a diffusely reflecting surface) since the intensity profile matches a cosine in this case.
 This emitter characteristic is the case for global illumination diffuse indirect illumination (non-specular).
@@
 However, a proper normalization can provide corrective support, wich reveals another subsequent normalization constraint.
-The normalization factor is defined in 3 subsequent steps:
+The normalization factor is defined in three subsequent steps:
 \begin{enumerate}
 	\item Normalization of $TI$ to $\overline{TI}=1.0$
-	\item Subsequent scaling w. mean intensity $\mean{I}$ for energy conservation principles
-	\item Subsequent scaling w. mean tensor magnitude $\mean{T}$ for absorption principles
+	\item Subsequent scaling with mean intensity $\mean{I}$ for energy conservation principles
+	\item Subsequent scaling with mean tensor magnitude $\mean{T}$ for absorption principles
 \end{enumerate}
@@
 In fact the approach has proven to be robust w.r.t.
-energy losses for several iterations as depicted in chap.
-\ref{chap:eval} in the stochastic error graph.
+energy losses for several iterations as depicted in Chap.~\ref{chap:eval} in the stochastic error graph.
 
@@
  	P(\omega) &= \frac{T(\omega)}{\int_{k\pi}^{(k+1)\pi}T(\omega)\mathop{d\omega}}\\
- 	w. \int_0^{\pi} P(\omega) &= 1.
+ 	\mathit{with} \int_0^{\pi} P(\omega) &= 1.
 \end{align}
@@
 Each tensor is then a unique footprint which can be precomputed for convenience and described as polar profile just as our previously mentioned intensity profiles.
-We apply these polar profiles in a sense of transmission profiles as discussed in sect.
-\ref{sec:trans}.
-
-\section{Light Propagation Scheme Behaviour}
+We apply these polar profiles in a sense of transmission profiles as discussed in Sect.~\ref{sec:trans}.
+
+\section{Light Propagation Scheme}
 To evaluate the plausibility of the light distribution results obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates:
@@
     \label{a)}
-    \caption*{light src in top center}
+    \caption*{light source in top center}
   \end{minipage}
@@
     \label{b)}
-    \caption*{light src in center}
+    \caption*{light source in center}
   \end{minipage}
@@
 One can observe here, how light is propagated with no directional bias leading to an absolutely isotropic distribution in case  b), which is symmetric as expected.
-This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in sect.
-\ref{sec:att} and \ref{sect:conserve}.
+This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in Sect.~\ref{sec:att} and \ref{sect:conserve}.
 We can also depict the final distribution as a scalar field when using the total energy in each cell as scalar:
@@
     \label{a)}
-    \caption*{light src in center}
+    \caption*{light source in center}
   \end{minipage}
@@
     \label{b)}
-    \caption*{light src in center (rescaled)}
+    \caption*{light source in center (rescaled)}
   \end{minipage}
@@
     \label{a)}
-    \caption*{antisymmtric light src's}
+    \caption*{antisymmtric light sources}
   \end{minipage}
@@
     \label{b)}
-    \caption*{symmetric light src's}
+    \caption*{symmetric light sources}
   \end{minipage}
@@
 \end{figure}
-We consider Fig.
-\ref{rings-tests} to regard the behavior for circular tensor field lines.
+We consider Fig.~\ref{rings-tests} to regard the behavior for circular tensor field lines.
 As expected, the intensities get directed in circular orbits by the tensor ellipsoid transmission profiles.
@@
     \label{a)}
-    \caption*{dense light src's}
+    \caption*{dense light sources}
   \end{minipage}
@@
     \label{b)}
-    \caption*{four light src's}
+    \caption*{four light sources}
   \end{minipage}
@@
     \label{a)}
-    \caption*{single light src}
+    \caption*{single light source}
   \end{minipage}
@@
     \label{c)}
-    \caption*{dual light src's}
+    \caption*{dual light sources}
   \end{minipage}
@@
     \label{a)}
-    \caption*{symmetric light src's}
+    \caption*{symmetric light sources}
   \end{minipage}
@@
     \label{b)}
-    \caption*{dual light src's}
+    \caption*{dual light sources}
   \end{minipage}
@@
 Note that we skipped the single light source example, since we consider it a trivial add-on example. We can regard the intensities condensed inside of a narrow beam of light.
-Note, that a logarithmic scale needs to be incorporated to produce results on a large scale in a sense of higher spatial resolution (e.g.
-$101\times 101$), since the propagation attenuation follows the distance to the origin inversely proportional.
-
-\section{Light Transport Gradient (LTG)}
+Note, that a logarithmic scale needs to be incorporated to produce results on a large scale in a sense of higher spatial resolution (e.g.: $101\times 101$), since the propagation attenuation follows the distance to the origin inversely proportional.
+
+\section{Light Transport Gradient}
+In this section, we will motivate the main contribution on tensor field analysis in this work.
+The approach called Light Transport Gradient (LTG) states a FTLE-related approach performed on tensor fields, but employing whole distributions of trajectories interpreted as light transmission profiles.
+Since we incorporate whole distributions of trajectories, we do not strictly follow the major eigenvectors like Hlawatsch's et al.
+\cite{hlawatsch} do with their FTLE on single trajectories following tensor field lines.
+Instead, we follow the directional distribution defined by the tensor field's principal ellipsoids for different input angles, which yields a FTLE volume, congruently depending on the direction, in general.
+In place of a 2D FTLE field concerning only the direction of the major eigenvector in each cell, we obtain all the directionally dependent FTLEs unrolled into slices and hence transform the FTLE field into a function FTLE$(\omega\mapsto z)$.
+The resulting volume contains a manifold which is not self-overlapping and is a continous plane for a continous tensor field.
 \label{sec:ltg}
 \subsection{Evolution}
-Now, that we are able to propagate intensity in grid, directed by the eigenvectors of the tensor field, we aim to do something more elaborate to segment ROIs (regions of interest) characterised by divergence/convergence of tensor field lines and drastic changes in anisotropy.
+Now, that we are able to propagate intensity in grid, directed by the eigenvectors of the tensor field, we aim to do something more elaborate to segment ROIs characterised by divergence/convergence of tensor field lines and drastic changes in anisotropy.
 For this purpose, we imagine kind of a local and spatial gradient, as we place light sources in two close-by positions and eventually in two close-by directions and analyze the resulting final light distribution.
@@
 We will start off with the \enquote{isotropic} field from here to first analyze the offset response for no divergence or more specifically, no anisotropy at all.
-Fig.
-\ref{analysis1} depicts final light distributions (cf.
-sect.
-\ref{sec:ltg} Fig.
-\ref{rings-tests}) for placing the light source one step to the left and one to the right from the central axis (spatial offset).
+Fig.~\ref{analysis1} depicts final light distributions (cf.~Sect.~\ref{sec:ltg} Fig.~\ref{rings-tests}) for placing the light source one step to the left and one to the right from the central axis (spatial offset).
 We can observe the light cones cancel out each other in the center of the difference image.
@@
 For now, we can take home from these examples, that light distributions are cancelling each other out within their overlap in general, in case no anisotropy was introduced into the test tensor field.
-We again analyze the drain test field (cf.
-Fig.
-\ref{drain} in sect.
-\ref{chap:grundlagen}) for the behavior of the resulting light distributions in scalar field representation.
-This following example illustrated in Fig.
-\ref{evolution1} now introduces raw tensor field lines diverging w.r.t.
+We again analyze the drain test field (cf.~Fig.~\ref{drain} in Sect.~\ref{chap:grundlagen}) for the behavior of the resulting light distributions in scalar field representation.
+This following example illustrated in Fig.~\ref{evolution1} now introduces raw tensor field lines diverging w.r.t.
 the $y$-axis.
@@
 \end{figure}
-We can see the light distributions diverge following tensor field lines, which in turn form these curves depicted in Fig.
-\ref{evolution1}.
+We can see the light distributions diverge following tensor field lines, which in turn form these curves depicted in Fig.~\ref{evolution1}.
 The separation of path (tensor field) lines and thus of light distributions is maximized at the center location of the field.
@@
 \end{figure}
-For the angle difference in Fig.
-\ref{evolution2}, we can observe that the difference is again increased, because it does not overlap leading to a maximized difference.
+For the angle difference in Fig.~\ref{evolution2}, we can observe that the difference is again increased, because it does not overlap leading to a maximized difference.
 Imagine accumulating each pixel in the images to obtain the cumulative total energy in the (difference) grid, which could now be compared for each light source position and direction in terms of proper min-max scaling.
@@
 This corresponding to the volume emerging from the FTLE plane image, as we introduce a new dimension $\omega$.
-
 Remark that we denote the final buffer result as $\mathbf{b}=\mathcal{C}_{final}$ in the following.
@@
 They can even be observed in nature for, e.g., vortices in flow fields.
-For this purpose, we will use our, previously defined, $\mathop{propagate}$-operator, i.e., the propagation scheme from sect.
-\ref{sec:scheme} and \ref{sec:trans}, on every possible light source location $\mathbf{x}$ and direction $\omega$ in the grid to obtain the final set of buffers $\mathcal{B} = \{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_j\}$ of impulse responses of the algorithm.
+For this purpose, we will use our, previously defined, $\mathop{propagate}$-operator, i.e., the propagation scheme from Sect.~\ref{sec:scheme} and \ref{sec:trans}, on every possible light source location $\mathbf{x}$ and direction $\omega$ in the grid to obtain the final set of buffers $\mathcal{B} = \{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_j\}$ of impulse responses of the algorithm.
 That is, we initialize the grid with every possible light source configuration and compute a whole final light distribution for each step:
@@
 Whereas $j$ is the index of light source position and direction representing the initialization in each iteration and $\mathop{steps}$ is the count of the discrete angular steps.
-The initialization is done by placing a light src (delta-impulse) with $\mean{I}_i = 1.0$ at the exact location of index $j$ realizing a light source at location $(x\mid y)$ in direction $\omega$.
+The initialization is done by placing a light source (delta-impulse) with $\mean{I}_i = 1.0$ at the exact location of index $j$ realizing a light source at location $(x\mid y)$ in direction $\omega$.
 Eq.
@@
 
-When we look at the FTLE field, on the top right in Fig.
-\ref{ftle_iso}, we determine that we obtain a more or less uniform reponse (, if we would neglect the edges) just as expected.
+When we look at the FTLE field, on the top right in Fig.~\ref{ftle_iso}, we determine that we obtain a more or less uniform reponse (, if we would neglect the edges) just as expected.
 Because the divergence of TFLs is distributed uniformly in this test field, we get approximately the same response for any position, but yet introducing a systematic bias against the light source direction.
@@
 
-We observe the clear straight ridge in the example in Fig.
-\ref{ftle_absorb}.
+We observe the clear straight ridge in the example in Fig.~\ref{ftle_absorb}.
 As expected, the ridge is located on the vertivcal center line.
@@
 What we can observe here, is a feature called a ``ridge'' occuring as a high value line segment (red vertical track) which is already quite obvious for human vision.
-A gradient filter could be applied to extract and segment edges from the image as a preprocessing step for computer vision, which would look like the image in Fig.
-\ref{ftle_base} c).
-
-These ridges represent ROIs with underlying Lagrangian Coherent Structures, which we already discussed in sect.
-\ref{sect:motiv}.
+A gradient filter could be applied to extract and segment edges from the image as a preprocessing step for computer vision, which would look like the image in Fig.~\ref{ftle_base} c).
+
+These ridges represent ROIs with underlying Lagrangian Coherent Structures, which we already discussed in Sect.~\ref{sect:motiv}.
 It could for example represent a stress field inside of a mechanical tool with a central axis with maximum stress condensation.
@@
 The light transport gradient responds most inside of the channel region and on the top T-arms independently of the stimulus direction.
-For the sake of completeness, we show that the approach's functionality also for diagonal directions in Fig.
-\ref{diag-ftle}.
+For the sake of completeness, we show that the approach's functionality also for diagonal directions in Fig.~\ref{diag-ftle}.
 
@@
 
-We also evaluated the approach for the ``inverse'' test field, depicted in Fig.
-\ref{inverse-ftle}, whereas we can observe height ridges for the $90^\circ$ slice in any position with diverging tensor field lines w.r.t.
+We also evaluated the approach for the ``inverse'' test field, depicted in Fig.~\ref{inverse-ftle}, whereas we can observe height ridges for the $90^\circ$ slice in any position with diverging tensor field lines w.r.t.
 the light source direction.
@@
 The orthogonal ridges in this example reach out to the edges because they are gaining a bias by never offsetting the light source in a transmission direction (ellipsoid orientation) for orthogonals.
-This turns out to be a general feature of the method, as we will later verify in Fig.
+This turns out to be a general feature of the method, as we will later verify in Figs.
 \ref{res_ftle_volume} and \ref{star-ftle}.
@@
 For the ``inverse'' test field, we can observe a 3 part ``rotor'' indicating the ridges which are in parallel up to orthogonal to the current light source (slice) direction.
-This rotor can be obtained by applying a contour filter, but it can also be visualized by applying a simple thresholding on the surface representation of the volume, as depicted in Fig.
-\ref{inverse_surface}.
+This rotor can be obtained by applying a contour filter, but it can also be visualized by applying a simple thresholding on the surface representation of the volume, as depicted in Fig.~\ref{inverse_surface}.
 Subsequently, we will analyze the \enquote{rings} test field to show its performance on circular ridges.
@@
 \end{figure}
-What we can observe in Fig.
-\ref{rings-ftle}, is called an inside-outside test producing the ridge on the circumference of the circle with maximized radius fitting into the minimum bounding box.
+What we can observe in Fig.~\ref{rings-ftle}, is called an inside-outside test producing the ridge on the circumference of the circle with maximized radius fitting into the minimum bounding box.
 If one would place a light source on the circumference of this circle offsetting it spatially and in the angular domain, one would obtain one incomplete final light distribution (outside) and one complete final light distribution (inside), as this is the maximum radius at which the intensities can follow a circle.
@@
 \end{figure}
-The volumes, we examine here in Fig.
-\ref{rings_contour}, show that the field is symmetric, as we do not notice a change along the $\omega\mapsto z$-axis except for the natural shift of the field against the light source direction, which we discussed already in this subsection.
+The volumes, we examine here in Fig.~\ref{rings_contour}, show that the field is symmetric, as we do not notice a change along the $\omega\mapsto z$-axis except for the natural shift of the field against the light source direction, which we discussed already in this subsection.
 Along the $z$-axis, the light source is rotated in all directions which causes the shift to form a full cycle from bottom to top ($[0^\circ,360^\circ]$).
-At last, we will analyze the bow test field, which is now considered as a more complicated example. We again plot the FTLE field in color coding representation in Fig. \ref{bow-ftle}.
+At last, we will analyze the bow test field, which is now considered as a more complicated example. We again plot the FTLE field in color coding representation in Fig.~\ref{bow-ftle}.
 
@@
 \end{figure}
-Considering the volumes depicted in Fig.
-\ref{bow_contour} one may notice, that we obtain a riffled structure reminding of ground plates.
+Considering the volumes depicted in Fig.~\ref{bow_contour} one may notice, that we obtain a riffled structure reminding of ground plates.
 On half the volume height ($180^\circ$) the structure is turned by $180^\circ$ as expected as this light source configuration lights up the mirror image to the one on the surface.
@@
 That is, each position in their $2D$ scalar field is assigned a height in our $3D$ scalar field.
-We cover their results as a subset of our volume, which could be, e.g., a spiral plane in case all directions are given in circular arrangement in the underlying tensor field (cf.
-\enquote{rings} test field).
+We cover their results as a subset of our volume, which could be, e.g., a spiral plane in case all directions are given in circular arrangement in the underlying tensor field (cf.~\enquote{rings} test field).
 If we would vary the angle of the major eigenvector with a uniform distribution along the tensor field's width, we would obtain a primitive diagonal plane w.r.t.
@@
 Therefore, we claim and consider our invented technique to be more generalized (to all directions) resulting in a topological $n+1$-D volume.
-\chapter{Implementation}
+
+\chapter{Algorithm}
 \label{chap:imp}
@@
 Since it is an energetic threshold, it is only dependent on the light source magnitude and not on any type of resolutions.
-An iteration limit is set to $\mathop{limit}=10\mathop{width}$ (e.g.
-for $\mathop{width}=10\cdot 100=1000$)  to prevent the approach from eventual oscillation behavior for intensities and hence overpropagation (propagating for too many timesteps).
+An iteration limit is set to $\mathop{limit}=10\mathop{width}$ (e.g.: for $\mathop{width}=10\cdot 100=1000$)  to prevent the approach from eventual oscillation behavior for intensities and hence overpropagation (propagating for too many timesteps).
 The spatial resolution is inferred by the tensor field reader method from the tensor field (see Propagation below).
-The light source profile can be defined for the \enquote{non-FTLE} mode where the user sets position and polar profile of single or multiple light sources as a symbolic term: ,e.g.
-, \texttt{clip(cos(theta-pi/2))} for a Lambert-emitter pointing upwards.
+The light source profile can be defined for the \enquote{non-FTLE} mode where the user sets position and polar profile of single or multiple light sources as a symbolic term: e.g.
+\texttt{clip(cos(theta-pi/2))} for a Lambert-emitter pointing upwards.
 After the configuration read process is finished, we allocate memory for certain buffers holding glyph and intensity information with a dimension of $\mathop{dim}=\mathop{width}\mathop{height}\mathop{steps}$.
@@
   }
-  \caption{Indices and Dimensions in Grid}
+  \caption{Indices and dimensions in grid}
   \label{indices}
-\end{figure}\vskip 3pt
-Fig.
-\ref{indices} depicts the \textit{deltaIndex}-map for computing the destination index for the scaled cosine lobes for cell $c$.
+\end{figure}
+Fig.~\ref{indices} depicts the \textit{deltaIndex}-map for computing the destination index for the scaled cosine lobes for cell $c$.
 The size of each cell is $\mathit{size}=\mathit{steps}$ in a $1D$ buffer arrangement, respectively.
-We then initialize energetic sums and thresholds and run a while loop until the convergence criterion is met.
-We compute the difference $\Delta\Phi_{total}$ of the energetic sum in this iteration with the one from the last.
+
+\section{Stimuli and Convergence}
+We then initialize energetic sums concerning the convergence and thresholds and run a while loop until the convergence criterion is met.
+We compute the difference $\Delta\Phi_{total}$ of the energetic sum in this iteration with the one from the previous iteration.
 If it falls short of this threshold, we initiate a stop sequence and leave the loop.
-To compute the FTLE, we use multiple stimuli in parallel (,e.g., Dirac-pulses within the scope of this work), and just compute the differences in final light distributions for two close-by light source positions and directions to accumulate them in a total energy sum (capacity), which in turn forms one component of the light transport gradient.
+To compute the FTLE, we use multiple stimuli in parallel (e.g.: Dirac pulses within the scope of this work), and just compute the differences in final light distributions for two close-by light source positions and directions to accumulate them in a total energy sum (capacity), which in turn forms one component of the light transport gradient.
 A Euclidean norm yields a scalar field from the gradient at each position.
@@
 \State $\mathit{index} \gets (j\cdot\mathit{width}+i)$ \Comment{compute $1D$ (cell) index}
-\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{write light src }
+\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{write light source }
 \While{$\Delta\Phi_{total} < \mathit{\epsilon }$ } \Comment{while convergence criterion not met..}
 \State $\mathit{sumA} \gets 0.0$ \Comment{reset sum}
-\State \Call{propagate}{\null} \Comment{propagate \textit{bufferA} (src) in \textit{bufferB} (tar)}
+\State \Call{propagate}{\null} \Comment{propagate \textit{bufferA} (source) in \textit{bufferB} (tar)}
 \State $\mathit{sumA} \gets$\Call{sum}{\textit{bufferB}}  \Comment{sum up energies}
@@
 \State \Call{swap}{\textit{bufferA}, \textit{bufferB}} \Comment{swap buffers for restart}
-\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{re-write light src}
+\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{re-write light source}
 \State \Call{fill}{\textit{bufferB}, $0.0$}
@@
 
-\paragraph{Propagation}
+\section{Propagation}
 The method performing the light propagation, was implemented in a separate class called propagator, which we can initialize once for each programming thread (, which we can in turn provide with its own buffer memory).
 As an initialization we fill the buffers with zeros to have a blank buffer initially to fill up with meaningful intensity profiles.
-The transmission profiles are generated by loading the tensor field in memory, performing PCA on each of them and exploiting the thin $\mathbf{U}$-matrix for the first (greater) singular vector.
+The transmission profiles are generated by loading the tensor field in memory, performing PCA on each of them and exploiting the thin $\mathbf{U}$ matrix for the first (greater) singular vector.
 The resulting principal component ellipsoid equation is precomputed, sampled and stored for reuse once for each cell.
@@
 
-\paragraph{Languages and Libraries}
+\section{Languages and Libraries}
 At first, we needed to generate convenient test cases, which we design using Mathematicas VectorPlot, used to visualize vector fields.
@@
 \label{sec:att}
-It can be proven that the approach follows propagation attenuation principles and follows the inverse (square in $3D$) law ($\sim\frac{1}{r}$) as observed in the graph in Fig.
-\ref{att}.
+It can be proven that the approach follows propagation attenuation principles and follows the inverse (square in $3D$) law ($\sim\frac{1}{r}$) as observed in the graph in Fig.~\ref{att}.
 For that, we plot the relative intensity $i_{rel}(r_{rel})$ (in relation to the intensity at position (radius) $1$ to account for the inverse square law, which holds for relative distances.
-We evaluate the relative intensity by forming a ratio $\frac{I(r)}{I_1}$, that relates the step intensity to the src intensity at position 1.
+We evaluate the relative intensity by forming a ratio $\frac{I(r)}{I_1}$, that relates the step intensity to the source intensity at distance $1$.
 
@@
   }
-  \caption{Propagation Attenuation}
+  \caption{Propagation attenuation}
   \label{att}
-\end{figure}\vskip 3pt
-
-\section{Behaviour for Total Anisotropy}
+\end{figure}
+
+\section{Behavior for Total Anisotropy}
 \label{sect:total}
@@
 Hence, the energy can never follow a perfect circle, but at best an octagon.
-This means, that there is a drift of at least two cells per diagonal (one overshoot per enter and escape each) as observable in Fig.
-\ref{anisotropy}.
+This means, that there is a drift of at least two cells per diagonal (one overshoot per enter and escape each) as observable in Fig.~\ref{anisotropy}.
 Nonetheless, the energies are transmitted in circular orbits around the circle center.
@@
 \label{sect:conserve}
-The energy conservation test was done in similar manner as the propagation attenuation test in sect.
-\ref{sec:att}.
-We measure the impulse response of the algorithm for 80 iterations by summing the total energy in the grid after each iteration and comparing it to the initial energy sum (e.g.
-$\int_0^{2\pi}1\mathop{d\omega}=2\pi$) in iteration 0 in absolute values.
-This energetic error analysis yields the following graph depicted in Fig.
-\ref{energy}:
+The energy conservation test was done in similar manner as the propagation attenuation test in Sect.~\ref{sec:att}.
+We measure the impulse response of the algorithm for 80 iterations by summing the total energy in the grid after each iteration and comparing it to the initial energy sum (e.g.: $\int_0^{2\pi}1\mathop{d\omega}=2\pi$) in iteration 0 in absolute values.
+This energetic error analysis yields the following graph depicted in Fig.~\ref{energy}:
 
@@
   }
-  \caption{Sum Error: src intensities: red: 1.0, blue 2.0, green: 5.0}
+  \caption{Sum Error: source intensities: red: 1.0, blue 2.0, green: 5.0}
   \label{energy}
@@
 \centering
-  \begin{minipage}{0.4\textwidth}
-  \includegraphics[width=0.9\textwidth]{img/ftle_drain_alt.png}
-    \caption*{LTG($0^\circ$)}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \includegraphics[width=0.9\textwidth]{img/drain_ftle_org_real.png}
-     \caption*{FTLE: Hlawatsch's approach}
+  \begin{minipage}[t]{0.3\textwidth}
+  \includegraphics[height=\textwidth]{img/drain(alt)-TFL.png}
+    \caption*{a) Glyphs}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}[t]{0.3\textwidth}
+    \includegraphics[height=\textwidth]{img/ftle_drain_alt.PNG}
+     \caption*{b) LTG($0^\circ$)}
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}[t]{0.3\textwidth}
+    \includegraphics[height=\textwidth]{img/drain_ftle_org_real.PNG}
+     \caption*{c) FTLE: Hlawatsch's approach}
     \label{b)}
@@
 \end{figure}
+
 We can observe the comparison between our invented FTLE-like approach and the one of Hlawatsch et al.
-in the depiction in Fig.
-\ref{drain_ftle_comp}.
+in the depiction in Fig.~\ref{drain_ftle_comp}.
 Note, that a sharper but weaker ridge is detected on the right side of the image, which shines similar to a cone shape from the top.
 We interpret the missing spot in the top center as an artifact stemming from the way we handle our light propagation, as we allow whole distributions of trajectories to be propagated (with a scalar attribute interpretable as propability).
-This causes a sampling drift, we already discussed in sect.
-\ref{sect:total}, leading to an energy distribution ending up outside of the field, for this position.
+This causes a sampling drift, we already discussed in Sect.~\ref{sect:total}, leading to an energy distribution ending up outside of the field, for this position.
 The directive effect turns out to be not that strict and delayed compared to the classic tensor field line FTLE.
@@
 \end{figure}
-When we look at the example depicted in Fig.
-\ref{star-ftle}, we can see the FTLE approach LTG responding most in the center of the field, which is the nodal point of tensor field lines diverging in any direction.
+When we look at the example depicted in Fig.~\ref{star-ftle}, we can see the FTLE approach LTG responding most in the center of the field, which is the nodal point of tensor field lines diverging in any direction.
 Remark, that in the middle image there is a light V-shaped ridge on the left on blue background which is outshined by the center spot because of its orders of larger magnitude.
@@
 \end{figure}
-We can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t.
+In Fig.~\ref{gyre-ftle, we can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t.
 the $x$-axis ($0^\circ$).
-Note that even regions in antiparallel directions ridgres are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging.
-Again, it is noticeable, that the results correspond to each other to a certain degree.
-This implies, that we obtained reasonable results generated by our method, but the classical tensor field line FTLE is more strict concerning its directive abilities and congruently does not do things by halves (like we do with our partitioned light distributions).
-Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure.
+Note that even regions in antiparallel directions ridges are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging.
+Again, it is noticeable, that the result of Hlawatsch correspond to our result to a certain degree.
+It is noticeable, that Hlawatsch's approach does not consider the initial direction and hence will detect ridges also in antiparallel directions (gradient of the flow map).
+This yields symmetric ridges w.r.t.
+central axes, whereas we detect ridges only up to orthogonally to the initial light source direction as visible in Fig.~\ref{gyre-ftle}b)\& c), respectively.
+In the center location of each gyre, we interpolate the sorrounding ring with an isotropic cell in the middle, leading to chaotic trajectories leaving the gyre and thus leading to an increased gradient of the flow map for Hlawatsch's result.
+This causes the FTLE field to respond in the center of the gyres in Fig.~\ref{gyre-ftle}c).
+We interpret this behavior within the scope of interpolated trajectories, which are in general more likely to escape the gyre, since their angle is interpolated (by PCA), which is susceptive to rapid changes and hence leads to unstable behavior.
+What Hlawatsch et al.
+detect, could be called local ridges (in terms rather short), what our method detects could rather be called global ridges (in terms rather long).
+This implies, that we obtained reasonable results generated by our method, but the classical tensor field line FTLE is susceptive to instability concerning its directive abilities and does not do things by halves (like we do with our partitioned light distributions), which causes sharper ridges in general.
+When we  Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure.
 We again depict volume representations for this example:
@@
 \end{figure}
-Regarding Fig.
-\ref{gyre_volume}, we find that the approach detects the boundaries of concavities (lat.: Gyri) in the brain, forming the lobus cerebri (commonly known a lobes).
+Regarding Fig.~\ref{gyre_volume}, we find that the approach detects the boundaries of concavities (lat.: Gyri) in the brain, forming the lobus cerebri (commonly known a lobes).
 The invented method could thus, e.g., aid to compute $3D$ models of a subject in medical sciences.
@@
 The regions with a tensor magnitude higher than zero restrict the light propagation to certain areas with the condition, that not all of the energy is absorbed by sorrounding cells in the local neighborhood.
-Again comparing the results to the classical FTLE (Hlawatsch's approach) on tensor fields in Fig.
-\ref{brain_ftle_comp}, we notice that the FTLE response is chaotic, since this MRI scan is subject to inherent MRI noise filling the background regions in the image.
+Again comparing the results to the classical FTLE (Hlawatsch's approach) on tensor fields in Fig.~\ref{brain_ftle_comp}, we notice that the FTLE response is chaotic, since this MRI scan is subject to inherent MRI noise filling the background regions in the image.
 This noise produces random trajectories, leading to an increased gradient of the flow map and thus to a high FTLE response, which reveals trajectories running in random directions in background regions.
@@
 Hence, we do not yield a noisy representation but a FG-BG segmented image, with red highlights in positions, where the tensor magnitude is maximized (and thus absorption is minimized).
-Remark, that the resulting volume does not inherently depend on the direction, because the test field does not introduce any strong linear anisotropy, as visible from the glyph representation in Fig.
-\ref{brain_ftle_comp} a).
+Remark, that the resulting volume does not inherently depend on the direction, because the test field does not introduce any strong linear anisotropy, as visible from the glyph representation in Fig.~\ref{brain_ftle_comp} a).
 If no linear anisotropy was incorporated, the FTLE response will also be independent on the direction to the greatest extent.
@@
 \end{figure}
-We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig.
-\ref{brain_ftle_comp}.
+We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig.~\ref{brain_ftle_comp}.
 We propose the following threshold correction to prevent noisy FTLE results:
 
-We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig.
-\ref{brain_ftle_comp}.
+We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig.~\ref{brain_ftle_comp}.
 We propose the following threshold correction to prevent noisy FTLE results:
@@
 This change will now yield a volume (now depending on the direction) for the local normalization mode, which we will show later on.
-We could also claim, that this correction would be necessary for Hlawatsch's approach as well, as their result is susceptive to noise for the \enquote{brain} test field in Fig.
-\ref{brain_ftle_comp} b).
-We first show the functionality of the correction on their approach in Fig.
-\ref{brain_ftle_noise}.
+We could also claim, that this correction would be necessary for Hlawatsch's approach as well, as their result is susceptive to noise for the \enquote{brain} test field in Fig.~\ref{brain_ftle_comp} b).
+We first show the functionality of the correction on their approach in Fig.~\ref{brain_ftle_noise}.
 
@@
 \end{figure}
-Again observing the results for the tensor field lines, we now remark their proper correlation in BG regions (cf.
-straight lines in the background in Fig.
-\ref{brain_ftle_noise} a)).
-This yields a gradient of the flow map exactly equal to zero, which avoids any noisy FTLE responses, as visible in Fig.
-\ref{brain_ftle_noise} b) for Hlawatsch's approach.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-  \centering
-  \includegraphics[height=\textwidth]{img_new/brain_cross-section.PNG}
-    \caption*{a) tensor magnitude}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[height=\textwidth]{img_new/brain_volume.PNG}
-     \caption*{b) LTG}
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[height=\textwidth]{img_new/brain_volume2.PNG}
-     \caption*{c) FTLE H.'s approach}
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{heart} test field}
+Again observing the results for the tensor field lines, we now remark their proper correlation in BG regions (cf.~straight lines in the background in Fig.~\ref{brain_ftle_noise} a)).
+This yields a gradient of the flow map exactly equal to zero, which avoids any noisy FTLE responses, as visible in Fig.~\ref{brain_ftle_noise} b) for Hlawatsch's approach.
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+  \centering
+  \includegraphics[width=\textwidth]{img_new/brain_surf.PNG}
+     \label{a)}
+      a)
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{img_new/brain_volume2.PNG}
+    \label{b)}
+    b)
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{img_new/brain_cross-section.PNG}
+    c)
+    \label{b)}
+  \end{minipage}
+  \caption{\enquote{brain} test field LTG: a) LTG($0^\circ$), b) LTG volume, c) cross-section}
 \label{heart-ftle}
@@
 
-The volumes depicted in Fig. \ref{heart-ftle} now show variation in $z$-direction, as they are now visualizing pure FTLE fields with no weighting of the tensor magnitude incorporated.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-  \centering
-  \includegraphics[height=\textwidth]{img/heart.PNG}
-    \caption*{a) LTG($0^\circ$): global}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img_new/heart_slice0.PNG}
-     \caption*{b) LTG($0^\circ$): local}
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img/heart_ftle_org.PNG}
-     \caption*{c) Hlawatsch's approach}
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{heart} test field}
+The volumes depicted in Fig.~\ref{heart-ftle} now show variation in $z$-direction, as they are now visualizing pure FTLE fields with no weighting of the tensor magnitude incorporated.
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+  \centering
+  \includegraphics[width=\textwidth]{img/heart.PNG}
+	a)
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{img_new/heart_slice0.PNG}
+	b)
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{img/heart_ftle_org.PNG}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{\enquote{heart} test field LTG: a) global, b) local, c) Hlawatsch's approach}
 \label{heart-ftle}
 \end{figure}
-After applying the LTG approach on the heart dataset, we observe the resulting tensor magnitude weighted FTLE-field, which is again not changing w.r.t.
-direction, which is why we do not show the volumes.
+After applying the LTG approach on the heart dataset, we observe the resulting tensor magnitude weighted FTLE-field (global) and the locally normalized (local) in Figs.
+\ref{heart-ftle}a)\& b).
+Note that the FTLE response for the globally normalized tensor field is again not changing w.r.t.
+direction, which is why we do not show the volumes for the global noramlization.
 The approach highlights the regions with diverging tensor field lines resulting from linear anisotropy or increased tensor magnitude, emerging as ridges.
+In Fig.~\ref{heart-ftle}a) we observe the tensor magnitude weighted mode for an inital global normalization, whereas anisotropy characteristics get outshined by tensor magnitude influences.
+On the other hand, in Fig.~\ref{heart-ftle}b) we remark the typical edge artifacts for FG regions known from the classical FTLE.
+Also, we notice the approach detecting changes in anisotropy in lack of any directed ridges, which reveal diverging tensor field lines w.r.t.
+a central axis.
+The results seem to correspond to the ones of Hlawatsch et al.
+certainly only in parts of the domain, because we unrolled all the FTLE responses distributed over all discrete directions, which is why some of the FTLE ridges are missing.
+Note that these ridges are not that sharp compared to Hlawatsch's approach, as we use whole distributions of trajectories which should be interpreted in a more continous way (changing w.r.t.
+direction).
 These ridges could now represent blood vessels and such.
 The red spot emphasized in the left of the image could, e.g., represent the highest diffusion velocity in tissue.
-When considering the FTLE for the heart dataset depicted in Fig.
-\ref{heart-ftle-comp}, we find that the trajectories exhibit strong correlations amongst them.
+When considering the FTLE for the heart dataset depicted in Fig.~\ref{heart-ftle-comp}, we find that the trajectories exhibit strong correlations amongst them.
 This results from the low noise rate inherently present in this dataset.
@@
 In this case, we obtain the same results as the approach from Hlawatsch et al., whereas our global min-max normalization (due to noise) is becoming a redundant step for noiseless input data.
-A proof of this elaboration can be found in Fig.
-\ref{heart-ftle-comp}.
+A proof of this elaboration can be found in Fig.~\ref{heart-ftle-comp}.
 
@@
   \includegraphics[width=\textwidth]{img/heartDwnsmpl.png}
-    \caption*{Glyphs}
+	a)
     \label{a)}
@@
     \includegraphics[width=\textwidth]{img/heartDwnsmpl-TFL.PNG}
-     \caption*{Hlawatsch's approach: TFLs}
-    \label{b)}
-  \end{minipage}
-\caption{\enquote{heart} test field}
+	b)
+    \label{b)}
+  \end{minipage}
+\caption{\enquote{heart} test field: a) glyphs, b) TFLs}
 \label{heart-ftle-comp}
@@
 \end{figure}
-Fig.
-\ref{heart-ftle} depicts the volumes for the heart test field, whereas we can observe a 3D ridge running in spiral along the $z$-direction.
+Fig.~\ref{heart-ftle} depicts the volumes for the heart test field, whereas we can observe a 3D ridge running in spiral along the $z$-direction.
 The FTLE responses are shifted w.r.t.
@@
   \end{minipage}
-\caption{\enquote{Random} test field}
+\caption{\enquote{Random} test field: a) global, b) local normalization}
 \label{random}
 \end{figure}
-Observing the generated random test field both with global and local normalization in Fig.
-\ref{random}a)\& b) respectively, we remark that the propability for small numbers to have a high ratio in combination is increased, as linear anisotropy occurs mainly in BG regions (low tensor magnitudes).
-In case we use global normalization in Fig.
-\ref{random_FTLE}a), we will again obtain a FTLE field independent on the direction, since tensor magnitudes outshine the changes in anisotropy in this test case.
+Observing the generated random test field both with global and local normalization in Fig.~\ref{random}a)\& b) respectively, we remark that the propability for small numbers to have a high ratio in combination is increased, as linear anisotropy occurs mainly in BG regions (low tensor magnitudes).
+In case we use global normalization in Fig.~\ref{random_FTLE}a), we will again obtain a FTLE field independent on the direction, since tensor magnitudes outshine the changes in anisotropy in this test case.
 The LTG approach works as expected, as we see the tensor magnitude weighted into the left image, yielding a high response for the tensor magnitude weighted FTLE response.
@@
   \includegraphics[width=\textwidth]{img_new/random_global.png}
-    \caption*{LTG: Global normalization}
+b)
     \label{a)}
@@
     \includegraphics[width=\textwidth]{img_new/random_local_slice0.PNG}
-     \caption*{LTG: Local normalization}
-    \label{b)}
-  \end{minipage}
-\caption{\enquote{Random} test field}
+	a)
+    \label{b)}
+  \end{minipage}
+\caption{\enquote{Random} test field LTG: a) global, b) local normalization}
 \label{random_FTLE}
 \end{figure}
-In case we use local normalization in Fig.
-\ref{random_FTLE}b), we could interpret the contributions revealing low tensor magnitudes as a result from white noise, which shows the inherent problem with local normalization.
+In case we use local normalization in Fig.~\ref{random_FTLE}b), we could interpret the contributions revealing low tensor magnitudes as a result from white noise, which shows the inherent problem with local normalization.
 If we would just locally normalize the whole tensor field with the mean tensor magnitude of each cell, we would enhance noisy contributions, which turn out to be in general more propable to be anisotropic than FG data.
@@
   \includegraphics[height=\textwidth]{img/inverse29.png}
-    \caption*{$\mathop{res}=29\times29\times16$}
+    a) $\mathop{res}=29\times29\times16$
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/inverse51.PNG}
-     \caption*{$\mathop{res}=51\times51\times16$}
+	b) $\mathop{res}=51\times51\times16$
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/inverse101.PNG}
-     \caption*{$\mathop{res}=101\times101\times16$}
+	c) $\mathop{res}=101\times101\times16$
     \label{b)}
@@
   \includegraphics[height=\textwidth]{img/inverse29_volume.png}
-    \caption*{$\mathop{res}=29\times29\times16$}
+	a) $\mathop{res}=29\times29\times16$
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/inverse51_volume.PNG}
-     \caption*{$\mathop{res}=51\times51\times16$}
+	b) $\mathop{res}=51\times51\times16$
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/inverse101_volume.PNG}
-     \caption*{$\mathop{res}=101\times101\times16$}
+	c) $\mathop{res}=101\times101\times16$
     \label{b)}
@@
   \includegraphics[height=\textwidth]{img/slice16.png}
-    \caption*{$\mathop{res}=101\times101\times16$}
+	a) $\mathop{res}=101\times101\times16$
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/slice32.PNG}
-     \caption*{$\mathop{res}=101\times101\times32$}
+	b) $\mathop{res}=101\times101\times32$
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/slice64.PNG}
-     \caption*{$\mathop{res}=101\times101\times64$}
+	c) $\mathop{res}=101\times101\times64$
     \label{b)}
@@
   \includegraphics[height=\textwidth]{img/vol16.png}
-    \caption*{$\mathop{res}=101\times101\times16$}
+	a) $\mathop{res}=101\times101\times16$
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/vol32.PNG}
-     \caption*{$\mathop{res}=101\times101\times32$}
+	b) $\mathop{res}=101\times101\times32$
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/vol64.PNG}
-     \caption*{$\mathop{res}=101\times101\times64$}
+	c) $\mathop{res}=101\times101\times64$
     \label{b)}
@@
 Please note that all values in the test table where measured on following system:\\
-\texttt{4 x 16-Core Gold 6142M 2.60GHz, 3072 GB RAM}.
+\texttt{4 x 16-Core Gold 6142M 2.60GHz, 3072 GB RAM}.\\
+The (dynamic) memory consumption was calculated with following formula:
+\begin{align*}
+	\mathop{mem}=8\mathop{dim}\mathop{steps}\cdot 8\mathop{byte}+(dim+19\cdot\mathop{steps})\cdot 8\mathop{byte}
+\end{align*}
 \begin{table}
@@
 \hline\hline
-test name - Fig. \# & resolution ($x\times y\times\omega$) & runtime $[s]$ & memory consumption (dynamic)\\
+test name - Fig.~\# & resolution ($x\times y\times\omega$) & runtime $[s]$ & memory consumption (dynamic)$\approx$\\
 \toprule
-\enquote{drain} - \ref{ftle_base} & $101\times101\times64=652864$ & $179878$ & $8\times652864\times 8\mathop{byte} +(101\times101 + 19\times64)\times8\mathop{byte}\approx47.02\mathop{MB}$\\ 
+\enquote{drain} - \ref{ftle_base} & $101\times101\times64=652864$ & $179878$ & $47.02\mathop{MB}$\\ 
 \hline 
-\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times64=652864$ & $68787$ & $8\times652864\times 8\mathop{byte} +(101\times101 + 19\times64)\times8\mathop{byte}\approx47.02\mathop{MB}$\\
+\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times64=652864$ & $68787$ & $47.02\mathop{MB}$\\
 \hline 
-\enquote{random} &  &  & \\ 
+\enquote{random-local} & $101\times101\times32=326432$ & $59896$ & $23.5\mathop{MB}$\\ 
 \hline
-\enquote{isotropic} - \ref{bow-ftle} & $51\times51\times3283232$ & $1357$ & $8\times83232\times 8\mathop{byte} +(51\times51 + 19\times32)\times8\mathop{byte}\approx5.998\mathop{MB}$\\ 
+\enquote{random-global} & $101\times101\times32=326432$ & $757$ & $23.5\mathop{MB}$\\ 
 \hline 
-\enquote{rings} - \ref{bow-ftle} & $51\times51\times32=83232$ & $5761$ & $8\times83232\times 8\mathop{byte} +(51\times51 + 19\times32)\times8\mathop{byte}\approx5.998\mathop{MB}$\\ 
+\enquote{isotropic} - \ref{bow-ftle} & $51\times51\times3283232$ & $1357$ & $5.998\mathop{MB}$\\ 
 \hline
-\enquote{bow} - \ref{bow-ftle} & $51\times51\times32=83232$ & $6128$ & $8\times83232\times 8\mathop{byte} +(51\times51 + 19\times32)\times8\mathop{byte}\approx5.998\mathop{MB}$\\ 
+\enquote{rings} - \ref{bow-ftle} & $51\times51\times32=83232$ & $5761$ & $5.998\mathop{MB}$\\ 
 \hline 
-\enquote{star} - \ref{star-ftle} & $51\times51\times32=83232$ & $5815$ & $8\times83232\times 8\mathop{byte} +(51\times51 + 19\times32)\times8\mathop{byte}\approx5.998\mathop{MB}$\\ 
+\enquote{bow} - \ref{bow-ftle} & $51\times51\times32=83232$ & $6128$ & $5.998\mathop{MB}$\\ 
 \hline 
-\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times16=163216$ & $12308$ & $8\times163216\times 8\mathop{byte} +(101\times101 + 19\times16)\times8\mathop{byte}\approx11.754\mathop{MB}$\\ 
+\enquote{star} - \ref{star-ftle} & $51\times51\times32=83232$ & $5815$ & $5.998\mathop{MB}$\\ 
 \hline 
-\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times32=326432$ & $24571$ & $8\times326432\times 8\mathop{byte} +(101\times101 + 19\times32)\times8\mathop{byte}\approx23.5\mathop{MB}$\\ 
+\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times16=163216$ & $12308$ & $11.754\mathop{MB}$\\ 
 \hline 
-\enquote{brain} - \ref{brain-ftle} & $128\times128\times64=1048576$ & $2427$ & $8\times1048576\times 8\mathop{byte} +(128\times128 + 19\times64)\times8\mathop{byte}\approx75.5\mathop{MB}$\\
+\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times32=326432$ & $24571$ & $23.5\mathop{MB}$\\ 
 \hline
-\enquote{heart} - \ref{heart-ftle} & $256\times256\times32=2097152$ & $117199$ & $8\times2097152\times 8\mathop{byte} +(256\times256 + 19\times32)\times8\mathop{byte}\approx151.0\mathop{MB}$\\
+\enquote{brain} - \ref{brain-ftle} & $128\times128\times64=1048576$ & $2427$ & $75.5\mathop{MB}$\\
+\hline
+\enquote{heart} - \ref{heart-ftle} & $256\times256\times32=2097152$ & $117199$ & $151.0\mathop{MB}$\\
 \hline 

### 2019-07-07 thesis-template/dummy.tex
update

@@
 \centering
-  \begin{minipage}{0.3\textwidth}
+  \begin{minipage}[t]{0.4\textwidth}
    \centering
-  \includegraphics[width=\textwidth]{img_new/brain_surf.PNG}
+  \includegraphics[height=0.9\textwidth]{img_new/brain-TFL.png}
+	a) TFLs: noise correction
     \label{a)}
-      a)
   \end{minipage}
-  \begin{minipage}{0.3\textwidth}
+  \begin{minipage}[t]{0.4\textwidth}
    \centering
-    \includegraphics[width=\textwidth]{img_new/brain_volume2.PNG}
-    \label{b)}
-    b)
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img_new/brain_cross-section.PNG}
-    c)
+    \includegraphics[width=0.9\textwidth]{img/brain_ftle_surf_noisereduction.PNG}
+	b) Hlawatsch's approach: noise correction
     \label{b)}
   \end{minipage}
-  \caption{\enquote{brain} test field: a) LTG($0^\circ$), b) LTG volume, c) cross-section}
-\label{heart-ftle}
+\caption{\enquote{Brain} test field}
+\label{brain_ftle_noise}
 \end{figure}
-
 \begin{figure}[!t]

### 2019-07-07 thesis-template/msc-thesis.tex
update

@@
 \end{titlepage}
-\newpage
-\thispagestyle{plain} % empty
-\mbox{}
+\newpage\null\thispagestyle{empty}\newpage
 
@@
 Ich versichere, dass ich diese Master-Arbeit selbstständig verfasst und nur die angegebenen
-Quellen und Hilfsmittel verwendet habe und die Grundsätze und
-,Empfehlungen ``Verantwortung in der Wissenschaft'' der Universität Heidelberg beachtet wurden. 
+Quellen und Hilfsmittel verwendet habe und die Grundsätze und Empfehlungen ``Verantwortung in der Wissenschaft'' der Universität Heidelberg beachtet wurden. 
 \bigskip
@@
 \emph{eng.:}
-I hereby assure, that I drafted this thesis independently and only used the sources and materials labeled as references and that the conventions/principles and recommendations ``Verantwortung in der Wissenschaft'' of the Heidelberg University have been regarded.
+I hereby assure, that I drafted this thesis independently and only used the sources and materials labeled as references and that the conventions, principles and recommendations ``Verantwortung in der Wissenschaft'' of the Heidelberg University have been regarded.
 \vspace*{50pt}
@@
 Also we will discuss the most relevant related work in the domain of tensor field visualization and also in the domain of global illumination (indirect illumination rendering).
-Subsequently, in Chap.~\ref{chap:main}, we will propose the main contributions and methods of this work and document the results and evaluation in Chap.~\ref{chap:eval}.
+Subsequently, in Chap.~\ref{chap:meth}, we will propose the main contributions and methods of this work and document the results and evaluation in Chap.~\ref{chap:eval}.
 At last, a compact summary, conclusion and outlook will be given in Chap.~\ref{chap:concl}.\\
@@
   \label{polar}
-\end{figure}\vskip 3pt
+\end{figure}
 Mathematically speaking, a polar space is a system, which maps each angle $\omega$ a magnitude $r(\omega)$ through a symbolic function.
@@
    \caption{2D principal component ellipsoid\/ glyph (eigenframe)}
-   \label{hat}
+   \label{HAT}
    \captionsetup{font={footnotesize,bf,it}}
@@
     \includegraphics[width=\textwidth]{img/glyphs.png}
-    \caption{3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
+    \caption{3D ellipsoid glyphs (l:linear, p:planar, s:spherical)}
     \label{3Dglyphs}
@@
     \includegraphics[width=\textwidth]{img/diverge.png}
-   \caption{diverging streamlines for double gyre}
+	a)
    \label{diverge}
    \captionsetup{font={footnotesize,bf,it}}
-   \caption*{source: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dg_3traj.png}}
-  \end{minipage} \hskip 40pt
+  \end{minipage} \hskip 30pt
   %\raggedright
@@
     \includegraphics[width=\textwidth]{img/ftle_gyre.png}
-    \caption{FTLE field for double gyre}
-    \label{ftle_gyre}
+	b)
     \captionsetup{font={footnotesize,bf,it}}
@@
   \end{minipage}
-\end{figure}
-Fig \ref{diverge} and \ref{ftle_gyre} depict a scenario of diverging streamlines for a double gyre snapshot in time (yielding a time-independent dynamical system).
-Fig.~\ref{ftle_gyre} features a high ridge on the margins (borders) and in particular at the center of the domain.
+  \caption{double gyre: a) diverging streamlines, b) FTLE field}
+  \label{ftle_gyre}
+  \caption*{source: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/}}
+\end{figure}
+Fig.~\ref{ftle_gyre}a) depicts a scenario of diverging streamlines for a double gyre snapshot in time (yielding a time-independent dynamical system).
+Fig.~\ref{ftle_gyre}b) features a high ridge on the margins (borders) and in particular at the center of the domain.
 This region of interest (ROI) is considered to indicate the maximum relative local divergence of tensor field lines, by the approach.
@@
 We could imagine to introduce this process for a whole distribution of trajectories, which would implement an FTLE for uncertain vector fields in a sense, that whole propability distributions of trajectories could be heeded.
-An overview about existing visualization techniques for uncertain data is given by Griete \& Schuhmann \cite{griete}.
+An overview about existing visualization techniques for uncertain data is given by Griete \& Schuhmann \cite{griethe}.
 We require that the trajectories can cross each other unaffectedly, i.e., the superposition principle is supposed to be respected for the magnitudes.
@@
 \subsection{Test Data (Generation)}
-We used python's numpy library to generate predefined, synthetic test cases and caught two real data example from \url{TensorVis.org} and \url{https://www.cs.purdue.edu/homes/cs530/projects/project5.html}, which needed to be pre-processed through slicing and subsequent cropping to obtain $2D$ samples ($2\times2$) matrices).
+We used python's numpy library to generate predefined, synthetic test cases and caught two real data example from \url{TensorVis.org} and \url{https://www.cs.purdue.edu/homes/cs530/projects/project5.html}, which needed to be pre-processed through slicing and subsequent cropping to obtain 2D samples ($2\times2$) matrices).
 These test fields can be grouped into symmetric and asymmetric ones:\\
@@
   \end{minipage}
-\caption{Named test fields}
+\caption{Several test fields}
 \label{test-fields}
@@
 
-\subsubsection{Tensor Fields}
+\subsubsection{Asymmetric Tensor Fields}
 
@@
     \includegraphics[width=0.8\textwidth]{img/drain(alt).png}
-    \caption*{a) glyphs}
+	a)
     \label{a)}
@@
     \includegraphics[width=0.8\textwidth]{img/drain(alt)-TFL.png}
-    \caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{``drain'' test field}
+	b)
+    \label{b)}
+  \end{minipage}
+\caption{``drain'' test field: a) glyphs, b) tensor field lines for a)}
 \label{drain}
@@
     \includegraphics[width=0.8\textwidth]{img/rings.png}
-    \caption*{a) glyphs}
+	a)
     \label{a)}
@@
     \includegraphics[width=0.8\textwidth]{img/tensorfieldlines.png}
-	\caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{``rings'' test field}
+	b)
+    \label{b)}
+  \end{minipage}
+\caption{``rings'' test field: a) glyphs, b) tensor field lines for a)}
 \label{rings}
@@
     \includegraphics[width=0.8\textwidth]{img/spiral.png}
-    \caption*{a) glyphs}
+	a)
     \label{a)}
@@
     \includegraphics[width=0.8\textwidth]{img/spiral-TFL.png}
-    \caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{``spiral''-test field}
+	b)
+    \label{b)}
+  \end{minipage}
+\caption{``spiral''-test field: a) glyphs, b) tensor field lines for a)}
 \label{spiral}
@@
     \includegraphics[width=0.8\textwidth]{img/inverse.png}
-    \caption*{a) glyphs}
+	a)
     \label{a)}
@@
     \includegraphics[width=0.8\textwidth]{img/inverse-TFL.png}
-    \caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{``inverse'' test field}
+	b)
+    \label{b)}
+  \end{minipage}
+\caption{``inverse'' test field: a) glyphs, b) tensor field lines for a)}
 \label{inverse}
@@
     \includegraphics[width=0.8\textwidth]{img/gyre.png}
-    \caption*{a) glyphs}
+	a)
     \label{a)}
@@
     \includegraphics[width=0.8\textwidth]{img/gyre-TFL.png}
-    \caption*{b) tensor field lines for a)}
-    \label{b)}
-  \end{minipage}
-\caption{a) ``gyre''-test field, b) tensor field lines for a)}
+	b)
+    \label{b)}
+  \end{minipage}
+\caption{``gyre'' test field: a) glyphs, b) tensor field lines for a)}
 \label{gyre}
@@
 
-\subsubsection{Symmteric Tensor Fields}
-We also aim to use a real data example of the $3D$ DTI-MRI diffusion tensor of the brain for a male subject at the age of 27 and another one featuring a human heart.
+\subsubsection{Symmetric Tensor Fields}
+We also aim to use a real data example of the 3D DTI-MRI diffusion tensor of the brain for a male subject at the age of 27 and another one featuring a human heart.
 As we visualize 2D symmetric second order tensor fields, we need to slice the dataset and crop the $2{\times}2$ matrices omitting the $z$-component of the original $3{\times}3$ matrices.
-Fig.~\ref{brainglyphs} depicts the glyph representation of the test dataset for slice index $15$.
+Fig.~\ref{real1} depicts the glyph representation of the test dataset for slice index $15$.
 To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{brain} and \enquote{heart} scans, which can be observed in Figs.~\ref{real1} and \ref{real2}.
@@
     \label{a)}
-    \caption*{tensor magnitude}
+	a)
   \end{minipage}
@@
     \label{b)}
-    \caption*{cyclinder glyphs}
-  \end{minipage}
-\caption{Real data examples: \enquote{brain} dataset}
+    b)
+  \end{minipage}
+\caption{\enquote{Brain} dataset: a) tensor magnitude, b) cyclinder glyphs}
 \label{real1}
@@
     \label{b)}
-    \caption*{cyclinder glyphs}
-  \end{minipage}
-\caption{Real data examples \enquote{heart} dataset}
+	b)
+  \end{minipage}
+\caption{\enquote{Heart} dataset: a) tensor magnitude, b) cyclinder glyphs}
 \label{real2}
@@
   }
-  \caption{Red: Intensity Profile, cyan: Transmission Profile (glyph Ellipsoid Eq.)}
+  \caption{Red: Intensity Profile, cyan: Transmission Profile (glyph ellipsoid eq.)}
   \label{weighting}
@@
 These oriented crysal fibers are then representing the biased propagation directions of light inside of the crystal lattice.
-The propability density function $P(\omega)$, for a photon being scattered in a particular direction $\omega$ \ref{ryder}, is then described as:
+The propability density function $P(\omega)$, for a photon being scattered in a particular direction $\omega$ \cite{ryder}, is then described as:
 \begin{align}
@@
 Each tensor is then a unique footprint which can be precomputed for convenience and described as polar profile just as our previously mentioned intensity profiles.
-We apply these polar profiles in a sense of transmission profiles as discussed in Sect.~\ref{sec:trans}.
+We apply these polar profiles in a sense of transmission profiles as discussed in Sect.~\ref{sec:polar}.
 
@@
     \label{a)}
-    \caption*{light source in top center}
+	a) light source in top center
   \end{minipage}
@@
     \label{b)}
-    \caption*{light source in center}
+	b) light source in center
   \end{minipage}
@@
 One can observe here, how light is propagated with no directional bias leading to an absolutely isotropic distribution in case  b), which is symmetric as expected.
-This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in Sect.~\ref{sec:att} and \ref{sect:conserve}.
+This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in Sects.~\ref{sec:att} and \ref{sect:conserve}.
 We can also depict the final distribution as a scalar field when using the total energy in each cell as scalar:
@@
     \label{a)}
-    \caption*{light source in center}
+	a) light source in center
   \end{minipage}
@@
     \label{b)}
-    \caption*{light source in center (rescaled)}
+	b) light source in center (rescaled)
   \end{minipage}
@@
     \label{a)}
-    \caption*{antisymmtric light sources}
+	a) antisymmtric light sources
   \end{minipage}
@@
     \label{b)}
-    \caption*{symmetric light sources}
+	b) symmetric light sources
   \end{minipage}
@@
     \label{a)}
-    \caption*{dense light sources}
+	a) dense light sources
   \end{minipage}
@@
     \label{a)}
-    \caption*{dense light sources}
+	a) two light sources
   \end{minipage}
@@
     \label{b)}
-    \caption*{four light sources}
+	b) four light sources
   \end{minipage}
@@
     \label{a)}
-    \caption*{single light source}
+	a) single light source
   \end{minipage}
@@
     \label{c)}
-    \caption*{dual light sources}
+	c) dual light sources
   \end{minipage}
@@
     \label{a)}
-    \caption*{symmetric light sources}
+	a) symmetric light sources
   \end{minipage}
@@
     \label{b)}
-    \caption*{dual light sources}
+	b) dual light sources
   \end{minipage}
@@
     \label{a)}
-    \caption*{a) final light distribution for left spatial offset}
+	a) final light distribution for left spatial offset
   \end{minipage} $-$
@@
     \label{b)}
-    \caption*{b) final light distribution for right spatial offset}
+	b) final light distribution for right spatial offset
   \end{minipage} $=$
@@
     \label{b)}
-    \caption*{difference image for a) and b)}
+	difference image for a) and b)
   \end{minipage}
@@
     \label{a)}
-    \caption*{a) final light distribution for left angular offset}
+	a) final light distribution for left angular offset
   \end{minipage} $-$
@@
     \label{b)}
-    \caption*{b) final light distribution for right angular offset}
+	b) final light distribution for right angular offset
   \end{minipage} $=$
@@
     \label{b)}
-    \caption*{difference image for a) and b)}
-  \end{minipage}
-  \caption{``drain'' test field}
+	difference image for a) and b)
+  \end{minipage}
+  \caption{``drain'' test field: }
 \label{evolution2}
@@
     \label{a)}
-    \caption*{a) Glyphs}
+	a)
   \end{minipage}
@@
     \label{b)}
-    \caption*{b) LTG($0^\circ$) for a)}
-  \end{minipage}
-  \caption{\enquote{isotropic} test field}
+	b)
+  \end{minipage}
+  \caption{\enquote{isotropic} test field: a) glyphs, b) LTG($0^\circ$) for a)}
 \label{ftle_iso}
@@
     \label{a)}
-    \caption*{a) Volume}
+	a)
   \end{minipage}
@@
     \label{b)}
-    \caption*{b) IsoVolume Filter for a)}
-  \end{minipage}
-  \caption{\enquote{isotropic} test field}
+	b)
+  \end{minipage}
+  \caption{\enquote{isotropic} test field: a) volume, b) IsoVolume for a)}
 \label{ftle_iso}
@@
     \label{b)}
-    \caption*{b) LTG($90^\circ$) for a)}
-  \end{minipage}
-  \caption{\enquote{absorb} test field}
+	b)
+  \end{minipage}
+  \caption{\enquote{absorb} test field: a) glyphs, b) LTG($90^\circ$) for a)}
 \label{ftle_absorb}
@@
     \label{a)}
-    \caption*{a) TFLs and glyphs}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
+	a)
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
     \includegraphics[width=\textwidth]{img/ftle_drain_alt.png}
     \label{b)}
-    \caption*{b) LTG($90^\circ$) for a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
+	b)
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
     \includegraphics[width=\textwidth]{img/drain_alt_gradient.png}
     \label{b)}
-    \caption*{c) gradient for b)}
-  \end{minipage}
-  \caption{``drain'' test field}
+	c)
+  \end{minipage}
+  \caption{``drain'' test field: a) TFLs and glyphs, b) LTG($90^\circ$) for a), c) gradient for b)}
 \label{ftle_base}
@@
     \includegraphics[width=0.9\textwidth]{img/drain_alt-profile.PNG}
-    \caption*{Cross section on LTG volume}	
+	a)
     \label{a)}
@@
     \includegraphics[width=0.9\textwidth]{img/drain_alt_isovolume.PNG}
-     \caption*{IsoVolume Filter}
-    \label{b)}
-  \end{minipage}
-  \caption{LTG volume representation: \enquote{drain} test field}
+	b)
+    \label{b)}
+  \end{minipage}
+  \caption{LTG volumes: \enquote{drain} test field a) cross-section, b) IsoVolume}
 \label{drain_contour}
@@
     \label{a)}
-    \caption*{a) TFLs and glyphs}
+	a)
   \end{minipage}
@@
     \label{b)}
-    \caption*{b) LTG($225^\circ$) for a)}
-  \end{minipage}
-  \caption{``diagonal drain'' test field}
+	b)
+  \end{minipage}
+  \caption{``diagonal drain'' test field: a) TFLs and glyphs, b) LTG($225^\circ$) for a)}
 \label{diag-ftle}
@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
 \centering
@@
     \label{a)}
-    \caption*{a) TFLs and glyphs }
+	a)
   \end{minipage}
@@
     \label{b)}
-    \caption*{b) LTG($90^\circ$) for a)}
-  \end{minipage}
-  \caption{``inverse''-test field}
+	b)
+  \end{minipage}
+  \caption{``Inverse''-test field: a) TFLs and glyphs, b) LTG($90^\circ$) for a)}
 \label{inverse-ftle}
@@
     \includegraphics[height=\textwidth]{img/inverse_isovolume.png}
+    a)
   \end{minipage}
@@
     \includegraphics[height=\textwidth]{img/inverse_surface.png}
+    b)
   \end{minipage}
@@
     \includegraphics[height=\textwidth]{img/inverse_contour.png}
-  \end{minipage}
-    \caption{IsoVolume Filter on LTG volume representation: ``inverse'' test field}
-\label{inverse_contour}
+    c)
+  \end{minipage}
+    \caption{LTG volumes: ``inverse'' test field: a) IsoVolume, b) Surface, c) Contour}
+\label{inverse_volume}
 \end{figure}
@@
 For the ``inverse'' test field, we can observe a 3 part ``rotor'' indicating the ridges which are in parallel up to orthogonal to the current light source (slice) direction.
-This rotor can be obtained by applying a contour filter, but it can also be visualized by applying a simple thresholding on the surface representation of the volume, as depicted in Fig.~\ref{inverse_surface}.
+This rotor can be obtained by applying a contour filter, but it can also be visualized by applying a simple thresholding on the surface representation of the volume, as depicted in Fig.~\ref{inverse_volume}.
 Subsequently, we will analyze the \enquote{rings} test field to show its performance on circular ridges.
@@
     \label{b)}
-    \caption*{b) LTG($0^\circ$) for a)}
-  \end{minipage}
-  \caption{``rings''-test field}
+	b)
+  \end{minipage}
+  \caption{``Rings''-test field: a) TFLs and glyphs, b) LTG($0^\circ$) for a)}
 \label{rings-ftle}
@@
     \includegraphics[height=\textwidth]{img/rings_contour.png}
-    \caption*{Contour}
+	a)
   \end{minipage}
@@
     \includegraphics[height=\textwidth]{img/rings_isovolume.png}
-    \caption*{IsoVolume}
+	b)
   \end{minipage}
@@
     \includegraphics[height=\textwidth]{img/rings_volume.png}
-    \caption*{Volume}
-  \end{minipage}
-    \caption{LTG volume representations: ``rings'' test field}
+	c)
+  \end{minipage}
+    \caption{LTG volumes: ``rings'' test field: a) Contour, b) IsoVolume, c) Volume}
 \label{rings_contour}
@@
     \label{b)}
-    \caption*{b) LTG($0^\circ$) for a)}
-  \end{minipage}
-  \caption{``bow''-test field}
+	b)
+  \end{minipage}
+  \caption{``bow''-test field: a) TFLs and glyphs, b) LTG($0^\circ$) for a)}
 \label{bow-ftle}
@@
     \includegraphics[width=\textwidth]{img/bow_contour.png}
-    \caption*{Contour}
+	a)
   \end{minipage}
@@
     \includegraphics[width=\textwidth]{img/bow_volume.png}
-    \caption*{IsoVolume}
-  \end{minipage}
-    \caption{LTG volume representations: ``bow'' test field}
+	b)
+  \end{minipage}
+    \caption{LTG volumes: ``bow'' test field : a) Contour, b) IsoVolume}
 \label{bow_contour}
@@
   \begin{minipage}{0.3\textwidth}
+  \centering
     \includegraphics[width=\textwidth]{img/tensorfieldlines.png}
-    \caption*{TFLs and glyphs}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
+	a) TFLs and glyphs
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
     \includegraphics[width=\textwidth]{img/total_anisotropy1.png}
-    \caption*{iteration $50$}
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
+    b) iteration $50$
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
     \includegraphics[width=\textwidth]{img/total_anisotropy2.png}
-    \caption*{iteration $103$}
+	c) iteration $103$
     \label{c)}
@@
   \includegraphics[height=\textwidth]{img/drain(alt)-TFL.png}
-    \caption*{a) Glyphs}
+	a)
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/ftle_drain_alt.PNG}
-     \caption*{b) LTG($0^\circ$)}
+	b)
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/drain_ftle_org_real.PNG}
-     \caption*{c) FTLE: Hlawatsch's approach}
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{drain} test field}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{\enquote{Drain} test field: a) Glyphs, b) LTG($0^\circ$), c) FTLE: Hlawatsch's approach}
 \label{drain_ftle_comp}
@@
   \includegraphics[height=\textwidth]{img/star.png}
-    \caption*{\enquote{Star} test field}
+	a)
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/star_ftle.PNG}
-     \caption*{LTG slice($0^\circ$)}
+	b)
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/star_volume_alt.PNG}
-     \caption*{IsoVolume on LTG volume}
-    \label{b)}
-  \end{minipage}
-  \caption{LTG on \enquote{star} test field}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{\enquote{Star} test field: a) field, b) LTG slice($0^\circ$), c) IsoVolume on LTG volume}
 \label{star-ftle}
@@
   \includegraphics[height=\textwidth]{img/gyre.png}
-    \caption*{a) Glyphs}
+	a)
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/gyre_ftle.PNG}
-     \caption*{b) LTG($0^\circ$)}
+	b)
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/gyre_ftle_org.PNG}
-     \caption*{c) FTLE: Hlawatsch's approach}
-    \label{b)}
-  \end{minipage}
-  \caption{Gyre test field}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{\enquote{Gyre} test field: a) Glyphs, b) LTG($0^\circ$), c) FTLE: Hlawatsch's approach}
 \label{gyre-ftle}
 \end{figure}
-In Fig.~\ref{gyre-ftle, we can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t.
+In Fig.~\ref{gyre-ftle}, we can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t.
 the $x$-axis ($0^\circ$).
@@
 This implies, that we obtained reasonable results generated by our method, but the classical tensor field line FTLE is susceptive to instability concerning its directive abilities and does not do things by halves (like we do with our partitioned light distributions), which causes sharper ridges in general.
-When we  Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure.
+Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure.
 We again depict volume representations for this example:
@@
   \includegraphics[height=\textwidth]{img/gyre_volume2.png}
-    \caption*{a) IsoVolume}
+	a)
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/gyre_gauss_contour.PNG}
-     \caption*{b) Contour}
+	b)
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/gyre_volume4.PNG}
-     \caption*{c) GaussContour}
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Gyre} test field}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{\enquote{Gyre} test field LTG volumes: a) IsoVolume, b) Contour, c) GaussContour}
 \label{gyre_volume}
@@
   \end{minipage}
-\caption{\enquote{brain} test field}
+\caption{\enquote{Brain} test field}
 \label{brain_ftle_comp}
@@
   \includegraphics[width=0.9\textwidth]{img_new/brain-TFL.png}
-    \caption*{a) Glyphs}
+	a) TFLs: noise correction
     \label{a)}
@@
     \includegraphics[width=0.9\textwidth]{img/brain_ftle_surf_noisereduction.PNG}
-     \caption*{b) H.'s app.: TFLs (noise correction)}
-    \label{b)}
-  \end{minipage}
-\caption{\enquote{brain} test field}
+	b) Hlawatsch's approach: noise correction
+    \label{b)}
+  \end{minipage}
+\caption{\enquote{Brain} test field}
 \label{brain_ftle_noise}
@@
   \end{minipage}
-  \caption{\enquote{brain} test field LTG: a) LTG($0^\circ$), b) LTG volume, c) cross-section}
+  \caption{\enquote{Brain} test field LTG: a) LTG($0^\circ$), b) LTG volume, c) cross-section}
 \label{heart-ftle}
@@
   \end{minipage}
-  \caption{\enquote{heart} test field LTG: a) global, b) local, c) Hlawatsch's approach}
+  \caption{\enquote{Heart} test field: a) LTG global, b) LTG local, c) Hlawatsch's approach}
 \label{heart-ftle}
@@
   \end{minipage}
-\caption{\enquote{heart} test field: a) glyphs, b) TFLs}
+\caption{\enquote{Heart} test field: a) glyphs, b) TFLs}
 \label{heart-ftle-comp}
@@
    \label{a)}
- \end{minipage}
- \begin{minipage}{0.3\textwidth}
- \centering
-   \includegraphics[height=\textwidth]{img_new/heart_volume3.PNG}
+ \end{minipage}\hskip 12pt
+ \begin{minipage}{0.3\textwidth}
+ \centering
+   \includegraphics[width=\textwidth]{img_new/heart_volume3.PNG}
    \label{b)}
@@
  \end{minipage}
- \caption{\enquote{heart} test field: LTG volumes}
+ \caption{\enquote{Heart} test field: LTG volumes}
 \label{heart-ftle}
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
   \includegraphics[height=\textwidth]{img/inverse29.png}
@@
   \begin{minipage}{0.3\textwidth} 
+    \centering
     \includegraphics[height=\textwidth]{img/inverse51.PNG}
@@
   \begin{minipage}{0.3\textwidth}
- % \centering
+	\centering
   \includegraphics[height=\textwidth]{img/inverse29_volume.png}
-	a) $\mathop{res}=29\times29\times16$
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  %\centering
+	a)~$\mathop{res}=29\times29\times16$
+    \label{a)}
+  \end{minipage}%\hskip 17pt
+  \begin{minipage}{0.3\textwidth}
+	\centering
     \includegraphics[height=\textwidth]{img/inverse51_volume.PNG}
-	b) $\mathop{res}=51\times51\times16$
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
- % \centering
+	b)~$\mathop{res}=51\times51\times16$
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+	\centering
     \includegraphics[height=\textwidth]{img/inverse101_volume.PNG}
-	c) $\mathop{res}=101\times101\times16$
+	c)~$\mathop{res}=101\times101\times16$
     \label{b)}
@@
   \begin{minipage}{0.3\textwidth}
+  \centering
   \includegraphics[height=\textwidth]{img/slice16.png}
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img/slice32.PNG}
@@
   \includegraphics[height=\textwidth]{img/vol16.png}
-	a) $\mathop{res}=101\times101\times16$
+	a)~$\mathop{res}=101\times101\times16$
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/vol32.PNG}
-	b) $\mathop{res}=101\times101\times32$
+	b)~$\mathop{res}=101\times101\times32$
     \label{b)}
@@
     \includegraphics[height=\textwidth]{img/vol64.PNG}
-	c) $\mathop{res}=101\times101\times64$
+	c)~$\mathop{res}=101\times101\times64$
     \label{b)}
@@
 \hline 
-\enquote{random-local} & $101\times101\times32=326432$ & $59896$ & $23.5\mathop{MB}$\\ 
+\enquote{random-local} - \ref{random_FTLE} & $101\times101\times32=326432$ & $59896$ & $23.5\mathop{MB}$\\ 
 \hline
-\enquote{random-global} & $101\times101\times32=326432$ & $757$ & $23.5\mathop{MB}$\\ 
+\enquote{random-global} - \ref{random_FTLE} & $101\times101\times32=326432$ & $757$ & $23.5\mathop{MB}$\\ 
 \hline 

### 2019-07-07 TensorFields/gyre.py

@@ -1,4 +1,4 @@
-import numpy as np
+import numpy
 import math as m
 import os
 
@@ -46,7 +46,7 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101
+length = 29
 s = 4
 #deg = -45
 

### 2019-07-07 thesis-template/msc-thesis.tex
update

@@
 
-Tensorfelder werden meistens in Verbindung mit mechanischen Spanunngsverteilungen in 2D/3D-Gittern gebracht (vgl. Cauchy-Stress Tensor), haben aber auch andere praktische Bedeutungen in der Physik.
-Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren, die aber zusätzlich auch verwendet werden kann um die Lichtausbreitung in einer topologischen Szene zu beschreiben.
-Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema für kartesische Gitter ab, dass die Prinzipien der Ausbreitungsdämpfung und Energieerhaltung beachtet und Licht-verteilung/en für gegebene Lichtquellenrichtung/en und -positionen bis zur Konvergenz approximieren kann.
+Tensorfelder werden häufig in Verbindung mit mechanischen Spanunngsverteilungen in 2D/3D-Gittern gebracht (vgl. Cauchy-Stress Tensor), haben aber auch andere praktische Bedeutungen in der Physik.
+Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren, die aber alternativ auch verwendet werden kann um die Lichtausbreitung in anisotropen Medien zu beschreiben.
+Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema für kartesische Gitter ab, welches die Prinzipien der Ausbreitungsdämpfung und Energieerhaltung beachtet und Lichtverteilung/en für gegebene Lichtquellenrichtung/en und -positionen approximieren kann.
 Als unterliegendes Modell werden die Transmissionsprofile innerhalb diesem Gitter als Kristallfiberstruktur angenommen (vgl.
 Edelsteine: Katzenaugeneffekt beim Tigerauge).
-Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiemaßen des unterliegenden Tensorfelds zu modellieren.
-Dafür erstellen wir per Hauptachsentransformation (PCA) ein Eigensystem für jede Zelle des Tensorfelds und modellieren eine zugehörige Ellipsengleichung als (Transfer-/Wichtungs-) Transmissionsprofil.
-Folglich kann die resultierende Lichtverteilung als 2D Skalarfeld oder Polarplot für Richtungsinformation visualisiert/repräsentiert werden.
-Wir messen Impulsantworten des Tensorfelds mit Delta-Pulsen an jeder Position und in jede Richtung um eine uniform gesamplte Map der globalen Lichtverteilungen zu erhalten.
-Diese Map wird als ein globales Energieflussfeld betrachtet.
-Wir lassen uns von der Technik FTLE aus dem Bereich Particle Tracing der Vector Field Visualization inspirieren, indem wir den Gradient des Flussfelds der resultierenden Lichtverteilungen analysieren und visualisieren.
-Ähnlich wie FTLE in Vektorfeldern Gräte/Kämme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, anziehende/abstoßende/sattelnde Tensorfeldlinien die häufig Schlüsselstrukturen in der Natur repräsentieren, in Tensorfeldern.
-Wir führen diese Größe als Global Illumination Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtfluss, zu visualisieren.
+Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiemaßen des unterliegenden Tensorfelds zu interpretieren.
+Dafür erstellen wir per Hauptachsentransformation ein Eigensystem für jede Zelle des Tensorfelds und modellieren eine zugehörige Ellipsengleichung als (Wichtungs-) Transmissionsprofil.
+Dabei werden die physikalischen Eigenschaften des Tensorfelds (Orientierung/Anisotropie/Amplitude) als Lichtleitfähigkeit interpretiert und die Lichtverteilung berechnet.
+Folglich kann die resultierende Lichtverteilung als 2D Skalarfeld oder Polarplot für zusätzliche Richtungsinformation repräsentiert und visualisiert werden.
+Wir messen Impulsantworten des Lichtausbreitungs-Prozesses moduliert mit den anisotropen Charakteristika des Tensorfelds, indem Delta-Pulse an jeder Position und in jede Richtung als Stimuli verwendet werden.
+Wir lassen uns von der Methode FTLE aus dem Bereich Particle Tracing der Vector Field Visualization inspirieren, indem wir den Gradient des Flussfelds der resultierenden Lichtverteilungen analysieren und visualisieren.
+Ähnlich wie FTLE in Vektorfeldern Gräte/Kämme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, nämlich anziehende/abstoßende/sattelnde Tensorfeldlinien die häufig Schlüsselstrukturen in der Natur repräsentieren, in Tensorfeldern.
+Wir führen diese Größe als Light Transport Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtfluss, zu visualisieren.
 \newpage
@@
 tensor field representations.
-Tensor fields, are commonly associated with stress distributions (cf.~Cauchy-stress tensor) in 2D/3D grids, but also have many other meanings in various fields.
+Tensor fields, are commonly associated with stress distributions in 2D/3D grids, but also have many other meanings in various fields.
+By means of global illumination techniques, we motivate a new method to visualize tensor fields,
+which is also capable of visualizing the light transport (propagation) in anisotropic media.
 We present the concept of applying techniques from the field of global illumination to tensor field analysis.
@@
 Similar to how the FTLE detects ridges/Lagrangian coherent structures in vector fields, our approach captures these structures likewise, namely regions revealing attracting/repelling/saddling tensor field lines which frequently represent key structures in nature, in tensor fields.
-We denote this entity as the light transport gradient (LTG), which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global light transport in the grid, directed by imprints in crystelline fiber structures modeled by the tensor field.
+We denote this entity as the light transport gradient (LTG), which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global light transport in the grid.
 We evaluate our approaches for plausibility and comparability within an extensive evaluation and demonstrate its performance on several examples of both synthetic and real datasets.
@@
 Classically, the issue with asymmetric tensor fields is, that they do not yield real eigenvalues in general, which is needed to set up glyph representations.
-In this work, we will use the matrix decomposition obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA principal ellipsoid representing the tensor transformation as suggested by Moler \cite{moler} to state another alternative of processing asymmetric tensors.
+In this work, we will use the matrix decomposition obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA (principal component analyis) principal ellipsoid representing the tensor transformation as suggested by Moler \cite{moler} to state another alternative of processing asymmetric tensors.
 2D tensor field visualization is useful for cases where only two-dimensional data is available, such as satellite remote sensing, particle imagery velocimetry and earthquake data \cite{laramee}.
@@
 This implies, e.g., clustering, structuring or projection techniques.
-When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor - medical resonance imaging) \cite{hlawatsch}, when measuring water molecule diffusion characteristics within tissue or in simulated or measured stress tensor fields from solid and fluid mechanics.
+When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor - medical resonance imaging) \cite{hlawatsch}, when measuring water molecule diffusion characteristics within tissue, or furthermore in simulated or measured stress tensor fields from solid and fluid mechanics.
 In these named domains of science the problem arises, that tensor fields can be generated or measured with physical/mathematical background, but not be regarded intuitively.
@@
 A Cartesian grid, which we will utilize in this context, is a grid normalized to a certain unit length, as a special case of a regular grid.
-That is, each cell is formed as a unit cube in relation to the normalizing unit length.
+That is, each cell is formed as a unit cube (square) in relation to the normalizing unit length.
 We aim to use global illumination methods to model light transport, directed by tensor fields, within such type of grids, which we consider as light propagation volumes (LPVs) \cite{dachsbacher}.
-Congruently, the most basic objective of this work is to design a simple light propagation scheme following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation, for tensor field visualization.
-Of course, the propagation scheme will need some kind of stimulus which can be defined through symbolic functions (e.g.: $r(\omega)=1$) by the user which we sample (discretize) as a primary step.
+Congruently, the most basic objective of this work is to design a simple light propagation scheme following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation, eventually to visualize tensor fields.
+Of course, the propagation scheme will need some kind of stimulus which can be defined through symbolic functions (e.g., $r(\omega)=1$) by the user which we sample (discretize) as a primary step.
 This way, the approach will operate as a light transport visualization technique for light distributions given in polar coordinates, provided the user set an initial profile into the grid.
-Furthermore, this propagation scheme should either be used to visualize the light transport paths, defined by the anisotropy characteristics of tensor fields and tensor magnitudes of a given tensor field in a $2D$ grid.
+Furthermore, this propagation scheme should either be used to visualize the light transport paths, defined by the anisotropy characteristics of tensor fields and tensor magnitudes of a given tensor field in a 2D grid.
 Or, as a secondary purpose, be used to visualize the light transport in a defined, given geometrical scene in a $2D$ grid.
-However, the tensor field needs to be preprocessed somehow to generate a footprint map of transmission profiles defined by the tensor field defined by the tensor field's anisotropy characteristics, which we will handle through matrix decompositions.
-For this purpose, we will let the transmission profiles act on our simple light propagation scheme to have a kind of tensor controlled directive effect on the light transport.
+However, the tensor field needs to be preprocessed somehow to generate a footprint map of transmission profiles defined by the tensor field's anisotropy characteristics, which we will handle through PCA yielding a principal component ellipsoid.
+For this purpose, we will let the transmission profiles act on our simple light propagation scheme to have a kind of  directive effect on the light transport controlled by the tensor's anisotropy characteristics.
 The transmission profiles should modulate the outgoing intensities in a sense of redirection or more appropriately, redistribution.
-This will yield an approach, capable of computing approximated final light distributions for a given light source configuration inside of a $2D$ lattice.
-As this process requires a lot of user-interaction and is not to be concerned as generalizing to all possible scenarios.
-In the sense of automization and optimzation, another consequent aim is to implement an approach borrowed from vector-field visualization to reveal structures, which are hard to grasp by ordinary methods.
+This will yield an approach, capable of computing approximated final light distributions for a given light source configuration inside of a 2D lattice.
+As this process requires a lot of user-interaction and is not to be concerned as generalizing to all possible scenarios, we aim to do something more elaborate on the basis of the light propagation scheme.
+We place light sources in form of Dirac pulses at any position and pointing in any direction represented in the grid, which represents a flow map, which we will analyze for its gradient to reveal the topolgical structure of the tensor field.
+In the sense of automization and optimization, another consequent aim is to implement an approach borrowed from vector-field visualization to reveal structures, which are hard to grasp by ordinary methods.
 We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges representing LCS (Lagrangian Coherent Structures).
-These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature, such as moraines in glaciers or vortices in flow fields.
-Commonly, in vector field visualization, LCS separate regions of different flow behavior, but in our case they separate regions/subdomains of varying directional distributions (manifesting in diverging/converging tensor field lines leading to compressions, tensions, roations or varying tensor magnitude causing volume changes) represented by tensors, as we use our tensor-controlled light transport visualization (propagation scheme) as a basis or framework to compute the FTLE field.
+These structures can be classified (identified) as repelling, attracting and saddling (i.e., forming a saddle) local features, which frequently represent key structures in nature, such as moraines in glaciers or vortices in flow fields.
+Commonly, in vector field visualization, LCS separate regions of different flow behavior, but in our case they separate regions, or in other words subdomains, of varying directional distributions (manifesting in diverging/converging tensor field lines leading to compressions, tensions, roations or varying tensor magnitude causing volume changes) represented by tensors, as we use our tensor-controlled light transport visualization (propagation scheme) as a basis or framework to compute the FTLE field.
 Consequently, we propose another method to compute an FTLE-like field on tensor fields.
-The FTLE field responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions and since we effectively measure the finite-time Lyapunov exponent (FTLE) with both forward and backward integration time (stimulus directions, in our case), we will detect both cases.
+The FTLE field responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other conceptually, because of bidirectional stress directions.
+Since we effectively measure the finite-time Lyapunov exponent (FTLE) with both forward and backward integration time (stimulus directions, in our case), we will detect both cases dependent on our initial light source direction.
 Both invented methods, will give new insights into the topology and anisotropy of tensor fields and provide new tools for the structural analysis.
@@
 The designed approach could, e.g., be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations.
-A classical example for tensor field visualization is plotting hyperstreamlines following major/minor eigenvectors, as depicted in Fig.~\ref{streamlines}, which features a degenerate point (purple), a wedge (red) and a trisector (green).
+A classical example for tensor field visualization is plotting hyperstreamlines following major/minor eigenvectors, as depicted in Fig.~\ref{streamlines}, which features a structurally unstable point (purple), a wedge (red) and a trisector (green).
 
@@
 Next, in Chap.~\ref{chap:grundlagen}, we will capture important fundamentals like tensors, polar coordinates, PCA in the context of singular value decomposition (generalization of eigenvalue decomposition), FTLE and GT (ground truth) datasets.
-Also we will discuss the most relevant related work in the domain of tensor field visualization and also in the domain of global illumination (indirect illumination rendering).
+Furthermore, we will discuss the most relevant related work in the domain of tensor field visualization and also in the domain of global illumination (indirect illumination rendering), since we use these methods as template models, as well.
 Subsequently, in Chap.~\ref{chap:meth}, we will propose the main contributions and methods of this work and document the results and evaluation in Chap.~\ref{chap:eval}.
@@
 There has been extensive recent work on symmetric tensors but in comparison little on asymmetric ones.
-Since we are not able to cover the whole scope in this work, we will focus on the most relevant works on tensor field visualization.
+Since we are not able to cover the whole scope in this work, we will focus on the most relevant works on global illumination and tensor field visualization.
 We will talk about the publications from Hlawatsch et al.
@@
 Thus, they aim on producing a most realistic image, what in our case turns out to be a most informative image in correspondence.
-The most early work in this field is propably the Discrete Ordinates Method (DOM) by Chandrasekhar \cite{Chandrasekhar}, which is used specifically for computing radiative transfer in participating media (volume rendering).
+The most early work in this field is propably the Discrete Ordinates Method (DOM) by Chandrasekhar \cite{Chandrasekhar}, which is used specifically for computing radiative transfer in participating media (for volume rendering).
 The radiative transfer equation is discretized spatially and in the angular domain.
-The radiance distribution is stored in a grid and light is distributed and exchanged between neighboring volume elements which reduces the overall effort to local and independent computations.
-Furthermore, there has been more recent work on describing the light propagation as a diffusion process within a simple photon transport model based on the Lattice-Boltzmann method \cite{geist}.
+The radiance distributions are represented in a grid and light is distributed and exchanged between neighboring volume elements which reduces the overall effort to local and independent computations.
+Furthermore, there has been more recent work on describing the light propagation as a diffusion process within a simple photon transport model, based on the Lattice-Boltzmann method \cite{geist}.
 DOMs were recently revisited by Fattal \cite{fattal}.
-This work addresses artifacts stemming from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}.
+This work addresses artifacts emerging from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}.
 These methods are considered to be too slow to \enquote{run at interactive framerates} by Dachsbacher et al.
@@
 Zheng \& and Pang \cite{pang&zheng} et al.
-proposed a texture-based visualization approach called Hyper LIC, which extends the concept of LIC to symmetric tensor fields by using an anisotropic 2D-filter kernel oriented along the major/minor eigenvectors.
-The concept of following the major/minor eigenvector along tensor field lines was invented by Vilanova et al.
+proposed a texture-based visualization approach called HyperLIC, which extends the concept of LIC to symmetric tensor fields by using an anisotropic 2D-filter kernel oriented along the major or minor eigenvectors.
+The concept of following the major/minor eigenvector along tensor field lines was first invented by Vilanova et al.
 \cite{vilanova}.
-Tensorlines \cite{weinstein} introduce a kind of artificial inertia (running average), making them resistent to noise.
+Tensorlines \cite{weinstein} introduce a kind of artificial inertia (running average), making them more resistent to noise.
 The notion of tensor field topology (critical, degenerate points, separatrices) and the concept of hyperstreamlines was first introduced by Delmarcelle and Hesselink \cite{delmarcelle&hesselink}.
@@
 Feng et al.
-\cite{feng} used Voronoi Tesselation for placing the glyphs.
+\cite{feng} used Voronoi tessellation for placing the glyphs.
 Superquadric glyphs have been proposed by Schultz and Kindlmann \cite{schultz&kindlmann} for general (non-positive definite) tensors.
 De Leeuw and van Wijk \cite{deleeuw&vanwijk} visualize the partial derivative gradient, the Jacobian of the tensor field, to sense the local properties of the field.
-A diffusion tensor can also be visualized by box \cite{makris}, ellipsoid \cite{pierpaoli}, composite \cite{westin2} or superquadric \cite{kindlmann} glyphs.
+A symmetric tensor can also be visualized by box \cite{makris}, ellipsoid \cite{pierpaoli}, composite \cite{westin2} or superquadric \cite{kindlmann} glyphs.
 There is also a set of scalar measures available in symmetric tensor field visualization such as: mean diffusivity, fractional anisotropy \cite{basser} and anisotropy coefficients \cite{westin}.
@@
 \cite{hlawatsch} obtain a coordinate (flow) map from tensor field lines following major/minor eigenvectors (fibre trajectories) using the maximum eigenvalue of the Cauchy-Green deformation tensor to compute the FTLE-field in homogeneous regions.
-In a similar manner, we aim to use light transport techniques to obtain a final light distribution (flow map) through successive light propagation cycles directed (in form of transmission profiles obtained from ellipsoid glyphs) by the tensor field.
+In a similar manner, we aim to use Global Illumination methods propagating light, directed by tensor fields, to obtain a final light distribution (flow map) through successive light propagation cycles.
+The light distributions are redirected by the tensor field employing transmission profiles obtained from ellipsoid glyphs.
 DT-MRI diffusion tensor imaging can help to reveal the functional \cite{denis} or topological \cite{wakana} structure of the brain and muscles by analysing celebral and muscle tissue tractography, which yields a tensor field representing the anisotropy characteristcs of the brain's white matter.
@@
 \cite{xavier} use invariant manifolds to extract topological features from tensor fields.
-Other feature extraction-based approaches extract general features as: edges \cite{hancock}, tensor shape\/ orientation \cite{gordon}, interfaces \cite{donnel}, creases \cite{tricochet}.
+Other feature extraction-based approaches extract general features as: edges \cite{hancock}, tensor shape\/ orientation \cite{gordon}, interfaces \cite{donnel} and creases \cite{tricochet}.
 \\
@@
 \cite{zirr}, since it allows a similar measure to be computed through defining a geometrical scene and setting the transmission (defined by the tensor field) to an isotropic and constant value of $100\%$.
-In this operating mode, the approach operates as a light transport visualization technique neglecting any tensor field transmission profiles and is thus capable of displaying bounding separatrices of a light transport scenario perturbated by obstacles \cite{zirr}.
+In this operating mode, the approach operates as a light transport visualization technique neglecting any tensor field transmission profiles and is thus capable of displaying bounding separatrices of a light transport scenario perturbated by obstacles and mirrors \cite{zirr}.
 %Typischerweise im letzten Abschnitt dieses Kapitels wird dann auf
@@
 
-In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and we will capture tensor fields, polar coordinates, PCA, FTLE, GT datasets and related work in that sense.
+In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and we will capture tensor fields, polar coordinates, PCA, FTLE and GT (ground truth) datasets in that sense.
 
@@
 \label{sec:tensors}
-Now what is a tensor? A tensor is a $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
+Now what is a tensor? A tensor is an $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
 \begin{table}[!b]
@@
  \end{table} 
-In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ in space (the dimension of the row and column vectors), which is equal to the rank of the tensor for full rank (in other words the number of column elements $n$ for a $n{\times}n$-matrix).
+In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ in space (the dimension of the row and column vectors), which is equal to the rank of the tensor for full rank.
+Or in other words the number of column elements $n$ for a $n{\times}n$-matrix.
 A tensor is considered non-degenerate if all eigenvalues $\lambda_i>\lambda_{i+1}$ manifesting in significant anisotropy.
-Tensors follow certain transformation rules which are defined for covariant/contravariant tensors, which indicate the change characteristics under a basis change.
+Tensors follow certain transformation rules which are defined for covariant and contravariant tensors, which indicate the change characteristics under a basis change.
 Contravariant tensors do not change under a change of basis, whereas covariant tensors are not invariant w.r.t.
@@
 In fact, tensors themselves are interpreted as transformations which reflect an incoming direction vector (measuring, e.g., the stress in that direction) to an outgoing (stress) vector.
-To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities are not simple numbers or vectors.
+To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features, as these entities are not simple numbers or vectors.
 They occur in form of a tensors, and thus as entities or quantities in physics.
@@
 Let us consider it as some lower dimensional, more intuitive representation of the tensor which we were left with in the first place.
-To put it all together, a tensor is a mathematical transformation which is encoded inside of an abstract $oD$ number array, which comes in need, whenever we have a whole distribution of directions manifesting itself inside of a tensor (matrix) transformation.
+To put it all together, a tensor is a mathematical transformation which is encoded inside of an abstract order-$o$ number array, which comes in need, whenever we have a whole distribution of directions manifesting itself inside of a tensor (matrix) transformation.
 

### 2019-07-08 thesis-template/dummy.tex
update

@@
 \centering
-  \begin{minipage}[t]{0.4\textwidth}
+  \begin{minipage}{0.4\textwidth}
     \centering
-  \includegraphics[height=0.9\textwidth]{img_new/brain-TFL.png}
-	a) TFLs: noise correction
+    \includegraphics[width=0.9\textwidth]{img/spiral-two-dense.png}
     \label{a)}
+	a)~two light sources
   \end{minipage}
-  \begin{minipage}[t]{0.4\textwidth}
+  \begin{minipage}{0.4\textwidth}
     \centering
-    \includegraphics[width=0.9\textwidth]{img/brain_ftle_surf_noisereduction.PNG}
-	b) Hlawatsch's approach: noise correction
+    \includegraphics[width=0.9\textwidth]{img/spiral-two-wide.png}
     \label{b)}
+	b)~symmetric light sources
   \end{minipage}
-\caption{\enquote{Brain} test field}
-\label{brain_ftle_noise}
-\end{figure}
-\begin{figure}[!t]
+  \begin{minipage}{0.4\textwidth}
     \centering
-  \begin{minipage}{0.3\textwidth}
+    \includegraphics[width=0.9\textwidth]{img/spiral-temp2.png}
+    \label{a)}
+	c)~two light sources
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
     \centering
-  \includegraphics[height=\textwidth]{img/inverse29.png}\label{1a}
-    a)
-    \label{a)}
+    \includegraphics[width=0.9\textwidth]{img/spiral_full.png}
+    \label{b)}
+	d)~four light sources
   \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img/inverse51.PNG}\label{2a}
-	b)
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img/inverse101.PNG}
-	c)
-    \label{b)}
-  \end{minipage}
-  \caption{ \protect a) some text 
-  \protect b) some other text}
-\label{res_ftle}
+\caption{final light propagation distributions for ``spiral''-test field}
+\label{spiral-tests}
 \end{figure}

### 2019-07-08 thesis-template/msc-thesis.tex
update

@@
 These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in plane normal directions ${x,y,z}$ ($3^2=9$ individual numbers).
-Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space.
+Thus, the stress tensor poses a full valid representation of the stress distribution at any point in space.
 The stress tensor itself can be interpreted as a transformation, which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$.
 The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent (kind of a Euclidean distance of the stress components).
-The tensor itself is a complicated representation and needs to be projected onto its principal axes to lower its dimensionality, as it is the case for principal component analysis.
-It is possible to find $2$ principal stress directions in $2D$ through PCA (see Sect.~\ref{sec:pca} below), for which there are no shear stresses existent in equilibrium.
+The tensor itself is a complicated representation and needs to be projected onto its principal axes to lower its dimensionality, as it could be done with, e.g., principal component analysis.
+It is possible to find $2$ principal stress directions in 2D through PCA (see Sect.~\ref{sec:pca} below), for which there are no shear stresses existent in equilibrium.
 These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour.
 The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''.
-So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig.~\ref{hat} in Sect.~\ref{sec:pca}.
+So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig.~\ref{HAT} in Sect.~\ref{sec:pca}.
 Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle.
@@
 To put it all together, a tensor is a mathematical transformation which is encoded inside of an abstract order-$o$ number array, which comes in need, whenever we have a whole distribution of directions manifesting itself inside of a tensor (matrix) transformation.
-
+\newpage
 \section{Polar Coordinates}
@@
 The polar domain is a circular axis version of the domain $[0..2\pi]$, with negative magnitudes in $r$ shifted by 180 degrees (point reflected) as this is the formal behaviour defined and required by the conversion formulas.
+Remark that we exclude negative magnitudes, in lack of physical interpretation for negative energies.
 
@@
 Note, that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following.
-Also keep in mind, that we interpret a matrix or tensor as affine transformation.
+Also keep in mind, that we interpret a matrix or tensor as an affine transformation.
 For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent.
@@
 
-Principal Component Analysis is an algorithm, which is able to capture the principal components (directions) in stochastic data.
+Principal component analysis is an algorithm, which is able to capture the principal components (directions) in stochastic data.
 But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling.
-Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices.
+Now, principal component analysis can be done by eigenvalue or singular value decomposition, whereas the former yields complex eigenvalues for asymmetric matrices.
 For (skew-) symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf.~\cite{kieburg}).
@@
 We will then use global-illumination techniques on top to generate a light transport flow map, which permits us somehow to forward the intensities which are given as predefined initialization values in form of symbolic functions defined by the user.
-This kind of light transport propagation scheme will also allow us to compute an FTLE-like field on the tensor fields.
+This type of light transport propagation scheme will also allow us to compute an FTLE-like field on the tensor fields.
 
@@
 We can now imagine placing two close by particles in distance $\delta$ into a flow field and run a time dependent simulation.
-Remark that this corresponds to the Lagrangian view in vector fields, which travels with the particles (instead of taking the position of a resting observer - Eulerian view):
+Remark that this corresponds to the Lagrangian view in vector fields, which travels with the particles (instead of taking the position of a resting observer called Eulerian view).
 
@@
 
-We observe particle trajectories diverging w.r.t. the central axis ending up in distance $\Delta$. The FTLE measure for this single example (point in space) is then defined as:
+In Fig. \ref{ftle_sep}, we observe particle trajectories diverging w.r.t. the central axis ending up in distance $\Delta$. The FTLE measure for this single example (point in space) is then defined as:
 
 \begin{align*}
-	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
+	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}.
 \end{align*}
@@
 \begin{align*}
-	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{\upvarphi(x)}\| _2,
+	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{\upvarphi(x)}\|_2.
 \end{align*}
@@
 Fig.~\ref{ftle_gyre}b) features a high ridge on the margins (borders) and in particular at the center of the domain.
-This region of interest (ROI) is considered to indicate the maximum relative local divergence of tensor field lines, by the approach.
-Ridges also represent Lagrangian Coherent Structures, which are found in nature in vortices of correlating flow behavior.
-We could imagine to introduce this process for a whole distribution of trajectories, which would implement an FTLE for uncertain vector fields in a sense, that whole propability distributions of trajectories could be heeded.
+The FTLE approach considers this region of interest (ROI) to indicate the maximum relative local divergence of tensor field lines, by the approach.
+Ridges also represent Lagrangian coherent structures, which are found in nature in vortices of correlating flow behavior.
+We could imagine to introduce this process for a whole distribution of trajectories, which would also implement a FTLE for uncertain vector fields in a sense, that whole propability distributions of trajectories could be heeded.
 An overview about existing visualization techniques for uncertain data is given by Griete \& Schuhmann \cite{griethe}.
@@
 Consequently, we choose to model the distributions of trajectories as light intensity profiles.
-That is, we interpret the polar profiles as a propability density function (PDF) for particle trajectories heading into this direction.
+That is, we interpret the polar profiles as a propability density function (PDF) for transmitting particle trajectories heading into a certain direction.
 To put it into context, our main contribution method LTG uses the FTLE as an inspiration or template to design a similar approach for tensor field visualization.
-The main difference to the template FTLE is that we follow eigenvectors instead of flow vectors and that we use whole distributions of trajectories, as a tensor field suggests whole distributions of directions, accordingly.
+The main difference to the template FTLE is that we follow eigenvectors instead of flow vectors and that we use whole distributions (i.e., light distributions specifically) of trajectories, as a tensor field suggests whole distributions of directions, accordingly.
+This interpretation will be the connecting key to incorporate tensor fields into our light propagation scheme.
 
@@
 Glyphs are derived as a basis for the weighting profiles of the tensors in the following propagation scheme.
-They were chosen because it is an easy-to-use (understand) concept, which can also be interpreted nicely.
-In fact, these representations are used for the principal component analysis of direction vectors for the tensor field line extrapolation (integration) and also as transmission profiles for the light propagation scheme in a subsequent step.
+They were chosen because it is an easy-to-use (and understand) concept, which can also be interpreted nicely.
+Tensor field lines (the type of tensor streamline analog that does not introduce artificial intertia) were implemented for ground truth (GT) data generation purposes.
 We can again emphasize glyphs as a crucial, central concept within this work.
-Tensor field lines (the type of tensor streamline analog that does not introduce artificial intertia) were implemented for ground truth (GT) data generation purposes.
-For example, when designing a FTLE-like field computation technique, we will need to know when tensor field lines diverge/converge.
+In fact, these representations are used for the principal component analysis of direction vectors for the tensor field line extrapolation (reimplementing Hlawatsch's \cite{hlawatsch} approach) and also as transmission profiles for our invented light propagation scheme in a subsequent step.
+For example, when designing a FTLE-like field computation technique, we will need to know when tensor field lines diverge/converge as ground truth information.
 Examples for glyphs (ellipsoids) and tensor field lines (thin lines) are depicted in Fig.~\ref{test-fields}.
@@
 We used python's numpy library to generate predefined, synthetic test cases and caught two real data example from \url{TensorVis.org} and \url{https://www.cs.purdue.edu/homes/cs530/projects/project5.html}, which needed to be pre-processed through slicing and subsequent cropping to obtain 2D samples ($2\times2$) matrices).
+Later on, we will find that we need a noise correction, to process real data.
 These test fields can be grouped into symmetric and asymmetric ones:\\
 
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
+\subsubsection{Asymmetric Tensor Fields}
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img/corner-TFL.png}
-    \caption*{``corner''}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
+	a)``corner''
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img/spiral-TFL.png}
-    \caption*{``spiral''}
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
+	b)``spiral''
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img/drain_old.png}
-    \caption*{``V''}
+	c)``V''
     \label{b)}
@@
 
-\subsubsection{Asymmetric Tensor Fields}
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/drain(alt).png}
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+      \centering
+    \includegraphics[height=\textwidth]{img/drain(alt)-TFL.png}
 	a)
@@
   \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/drain(alt)-TFL.png}
+  \begin{minipage}{0.3\textwidth}
+      \centering
+    \includegraphics[height=\textwidth]{img/tensorfieldlines.png}
 	b)
@@
   \end{minipage}
-\caption{``drain'' test field: a) glyphs, b) tensor field lines for a)}
-\label{drain}
-\end{figure}
-
-The ``drain''-Testfield, depicted in Fig.~\ref{drain} is generated to have a kind of proof-of-concept (POC) for the FTLE-related approach.
+  \begin{minipage}{0.3\textwidth}
+      \centering
+    \includegraphics[height=\textwidth]{img/inverse-TFL.png}
+	c)
+    \label{b)}
+  \end{minipage}
+\caption{Several test fields: a)``corner'', ``spiral'', ``V''}
+\label{test_fields2}
+\end{figure}
+
+The ``drain''-Testfield, depicted in Fig.~\ref{test_fields2}a) is generated to have a kind of proof-of-concept for the FTLE-related approach.
 It poses raw, bare diverging tensor field lines without any other feature modulated on top.
 Thus, it can be expected to induce a high FTLE response, since it is an interpretable ground truth test and is considered to be a basic trigger/stimulus for the approach with no side effects introduced.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/rings.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-   \centering
-    \includegraphics[width=0.8\textwidth]{img/tensorfieldlines.png}
-	b)
-    \label{b)}
-  \end{minipage}
-\caption{``rings'' test field: a) glyphs, b) tensor field lines for a)}
-\label{rings}
-\end{figure}
-The ``rings'' test field, depicted in Fig.~\ref{rings} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/spiral.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/spiral-TFL.png}
-	b)
-    \label{b)}
-  \end{minipage}
-\caption{``spiral''-test field: a) glyphs, b) tensor field lines for a)}
-\label{spiral}
-\end{figure}
-The ``spiral'' test field, depicted in Fig.~\ref{spiral} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/inverse.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/inverse-TFL.png}
-	b)
-    \label{b)}
-  \end{minipage}
-\caption{``inverse'' test field: a) glyphs, b) tensor field lines for a)}
-\label{inverse}
-\end{figure}
+The ``rings'' test field, depicted in Fig.~\ref{test_fields2} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
+The ``spiral'' test field, depicted in Fig.~\ref{test_fields2} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
 With the \enquote{inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.
@@
 \end{figure}
-The test field ``gyre'', depicted in Fig \ref{gyre}, constitutes a sustained turn over the grid representing the redirection or forwarding along domain corners.
+The test field ``gyre'', depicted in Fig \ref{gyre}, constitutes a classical textbook example for the FTLE and is considered for verifcation purposes.
 
@@
 Fig.~\ref{real1} depicts the glyph representation of the test dataset for slice index $15$.
-To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{brain} and \enquote{heart} scans, which can be observed in Figs.~\ref{real1} and \ref{real2}.
+To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{Brain} and \enquote{Heart} scans, which can be observed in Figs.~\ref{real1} and \ref{real2}.
 \begin{figure}[!t]
@@
 \end{figure}
-When we look at this DT-MRI scan of a human heart we can notice the elongated shape, which we consider to represent the aorta and a part of the rest of the heart.
+When we look at this DT-MRI scan of a human heart, we can notice the elongated shape, which we consider to represent the aorta and a part of the rest of the heart.
 If we consider the glyph representation, we again infer, that we incorporate lots of anisotropic regions with varying tensor magnitudes here.
@@
 \label{sec:spec} 
-Our ambition is, to motivate a simple and efficient global illumination propagation scheme, able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t.
+Our ambition is to motivate a simple and efficient global illumination propagation scheme, able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t.
 energy conservation and propagation attenuation principles.
 We then generate transmission profiles, which we sample in a preliminary step, from the eigensystems (principal component ellipsoids) of the tensors for each cell in the domain (field).
-This step is done, to provide an orientation for a lattice of crystal fibre structures (cf.~optical fibres) to redirect the intensities in analogy to the anisotropy characteristics of the tensor field within a glyph representation, obtained from principal component analysis (PCA).
+This step is done, to provide an orientation for a lattice of crystal fibre structures (cf.~optical fibres) to redirect the intensities in the grid in analogy to the anisotropy characteristics of the underlying tensor field within a glyph representation, obtained from PCA.
 Our approach should be implicitly adaptive to any kind of 2D dataset or slices of 3D datasets.
@@
 For the sake of interpretability, we exclude negative values (energies) from the scope of our calculations which we express through mathematical clip functions, limiting negative values to $0.0$, which we denote by subscript $+$ in the following.
-We will describe the emission charactersitics of point lights exchanging energies in the lattice by parsing such polar functions.
-Our grid consists of $\mathop{dim}=\mathop{width}\cdot \mathop{height}$ virtual point lights (VPLs), which are in turn light source with (theoretically) no finite extend represented through polar functions.
+We will describe the emission characteristics of point lights exchanging energies in the lattice by parsing such polar functions.
+Our grid consists of $\mathop{dim}=\mathop{width}\cdot \mathop{height}$ virtual point lights (VPLs), which are in turn light sources with (theoretically) no finite extend represented through polar functions.
 As a start-up thinker for the propagation scheme, we will think of a grid cell in 2D as a center pixel within an 8-neighborhood with a previously injected point light (imagine a spherical emitter: $r(\omega)=1$ here for simplicity).
@@
 Naively, one would consider to just transmit the intensity in the direction its pointing into.
-But this would omit every propagation attenuation principles (any principles at all specifically), if one takes a closer look:
+But this would omit every propagation attenuation principles (any principles at all specifically), if one takes a closer look.
 \subsection{Accumulation stage}
 Each cone type (yellow and green) is an individual neighbor-dependent angular area.
-The diagonals take in a greater angular domain while they are located at greater distance and thus need to be attenuated more strongly (consider an isotropic stimulus).
+The diagonals take in a greater angular domain, while they are located at greater distance and thus need to be attenuated more strongly (consider an isotropic stimulus).
 We aim for distributing contributions for overlapping cones to transmit attenuated (because of the larger factor $\sqrt{2}$ distance) intensity in form of a shared part to the diagonals in a single step (which also prevents from zigzagging diagonal intensities in case we would transmit diagonal intensities around the corner).
@@
 \end{align}
-
+This permits us to scale an oriented and clipped cosine-lobe $\cos_+$ with the total radiant flux with respect to its own total energy when integrated likewise.
+A discrete cosine lobe never reaches its theoretical sum of $2$ exactly in the interval $[{-\frac{\pi}{2}},{\frac{\pi}{2}}]$.
+Therefore it needs to be computed to account for sampling (discretization) errors and normalized by following relation (e.g., $\sum_{\cos} \approx 1.9949204635834517$).
+\begin{align}
+\sum_{\cos_+} &= \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\cos_+(\omega)\mathop{d\omega} \approx 2,\\
+	\cos_k(\omega) &= \frac{\Phi_t}{\sum_{\cos_+}}cos_+(\omega-k\cdot\frac{\pi}{4})\hskip 20pt \mathit{with} \ k\in[0,7].
+\end{align}
 \subsection{Injection stage}
  We place a cosine lobe scaled with the integrated (accumulated) energies in each of the 8 cone directions (angular domains), whereas outgoing integrated and summed energies are weighted by the linear combination weights (factors) in overlapping (green) areas to split up the contributions accordingly.
-The contribution areas are depicted in Fig.~\ref{scheme}, the cosine lobes can be observed in Fig.~\ref{iter} respectively.
+The contribution areas are depicted in Fig.~\ref{scheme}, the cosine lobes can be observed in Fig.~\ref{iter}, respectively.
 This manner of propagation is inspired by the propagation scheme of Dachsbacher et al.
@@
 Basically, we sum up the weighted intensities corresponding to the respective neighbor index and scale an oriented cosine lobe (in direction of the neighbor) with the total flux $\Phi_t$ accumulated inside the neighbor range to transmit all of the parsed intensity for each neighbor distributed into a cosine distribution.
+Remark, that a cosine distribution physically corresponds to Lambert emitter charactersitics and hence to diffusely reflecting surfaces.
 We repeat this process once for each neighbor and each cell respectively.
@@
 The implementation is done in a dual buffer approach which pushes the energies back and forth (buffer A to buffer B) for each propagation cyvle until a state of convergence is reached, which is the case when the energy is spreaded throughout the grid and enters a stationary state of equilibrium, characterized by equal out (at grid borders) to in (at light source positions) energy-flow with no more nettings going on inside the domain.
-We measure the convergence by setting up an overall distribution error w.r.t to the previous iteration.
-Note that even though we propagate the energy directly through the diagonal cones (which forms a square initially), the 8 degrees of freedom introduce a circular (spherical) wave front after few iterations (for greater resolutions) as depicted in Fig.~\ref{iter}.
+We measure the convergence by setting up an overall distribution error w.r.t.
+the previous iteration.
+Note, that even though we propagate the energy directly through the diagonal cones (which forms a square initially), the 8 degrees of freedom introduce a circular (spherical) wave front after few iterations (for greater resolutions), as depicted in Fig.~\ref{iter}.
 This occurs, because of a faster propagation attenuation which counteracts the larger angle of diagonal cones.
-There is a redistribution of intensities in the long term resulting in a, since the shared part effects a uniformly spaced isotropic propagation profile resulting in a circular torus-shaped distribution of intensities, as also observed in Fig.~\ref{iter}.
+There is a redistribution of intensities in the long term, since the shared part effects a uniformly spaced isotropic propagation profile resulting in a circular torus-shaped distribution of intensities, as observed in Fig.~\ref{iter}.
 Also note, that this approach satisfies energy conservation and propagation attenuation principles and applies to the principles of light propagation in vacuum (by default).
@@
   \label{iter}
-\end{figure}\vskip 3pt
-Fig \ref{iter} depicts the named iteration indices $0,1,2,3,7,11$ (cycles) for the propagation scheme for a circular pulse stimulus which gets turned off after 1 iteration to demonstrate raw energy spreading in the field (impulse response of the algorithm).
+\end{figure}
+Fig \ref{iter} depicts the iteration indices $0,1,2,3,7,11$ (cycles) for the propagation scheme for a circular pulse stimulus which gets turned off after 1 iteration to demonstrate raw energy spreading in the field (impulse response of the propagation scheme).
 In step 1 one can observe the placing of cosine lobes scaled with the respective accumulated and weighted energy for the summed area in each of the 8 neighbor directions.
@@
 
-As already mentioned, a dual-buffer approach is chosen here, as we can efficiently propagate from one into the other and just swap pointers to 
-start all over (after reinitialization) until convergence.
+As already mentioned, a dual-buffer approach is chosen here, as we can efficiently propagate from one into the other and just swap pointers to start all over (after reinitialization) until convergence.
 Another reason is it yields a reasonable source-target or read-write segmentation.
-We introduce a new operator called \textit{propagate}, permitting us to pre-define the propagation in a symbolic manner.
+We introduce a new operator called \textit{propagate}, permitting us to predefine the propagation in a symbolic manner.
 The propagation definition denotes the transformation from one grid (A-buffer) into another (B-buffer) through proper application of the \textit{propagate}-operator to each cell.
-We will extend this operator with a normalization correction in Sect.~\ref{sec:trans} in form of a precise mathematical formulation respectively.
+We will extend this operator with a normalization correction in Sect.~\ref{sec:trans} in form of a precise mathematical formulation, respectively.
 For now, just imagine the approach in raw operating mode, with no tensor influence incorporated.
@@
 For now, we want to define our operator $\mathop{propagate}$: Each Grid or Buffer is basically a set of cells defined as $\mathcal{C} = \{\mathbf{c}_1, \mathbf{c}_2, ..., \mathbf{c}_i\}$.
-The dual buffer result $\mathcal{C'}$ is denoted symbolically as follows for all cells with a mean intensity greater than zero (those which do not contain trivial null samples):
+The dual buffer result $\mathcal{C'}$ is denoted symbolically as follows for all cells with a mean intensity greater than zero (those which do not contain trivial zero samples):
 \begin{align}
-	\mathcal{C'} = \mathop{propagate}\{c_i\in \mathcal{C} \ \mid \ |\mean{I}_i| > 0\} \hskip 5pt {\forall } \ i \in [1,\mathop{dim}].
+	\mathcal{C'} = \mathop{propagate}\{c_i\in \mathcal{C} \ \mid \ |\mean{I}_i| > 0\} \hskip 5pt {\forall } \ i \in [1,\mathop{dim}],
 \end{align}
-where: $i$: cell index, $I$: source buffer radiant intensity\vskip 5pt
+whereas: $i$: cell index, $I$: source buffer radiant intensity.\vskip 5pt
 The dual buffer result $\mathcal{C'}$ denotes the set of all transformed cells after the propagation, i.e., after applying the operator $\mathop{propagate}$ to each cell.
-This result is generated once in each iteration and gets pushed back and forth between the 2 buffers.
+This result is generated once in each iteration and conceptually gets pushed back and forth between the two buffers.
 Note that, $\mathop{dim}=width\cdot height$, which equals the tensor field resolution in this case.
@@
 \begin{align*}
-	\Delta \Phi_{total} &= \sum_{c_i\in\mathcal{C}}\lvert\Delta I_i(\omega)\rvert = \sum_{c_i\in\mathcal{C}} \int_0^{2\pi} \lvert(I'_i(\omega)-I_i(\omega))\rvert \mathop{d\omega}  \overset{!}{<} \epsilon\\
+	\Delta \Phi_{total} &= \sum_{c_i\in\mathcal{C}}\lvert\Delta I_i(\omega)\rvert = \sum_{c_i\in\mathcal{C}} \int_0^{2\pi} \lvert(I'_i(\omega)-I_i(\omega))\rvert \mathop{d\omega}  \overset{!}{<} \epsilon.\\
 \end{align*}
@@
 Note that we take the absolute value of the differences to prevent from mutual compensation.
-
 Until now, we aimed to simulate the propagation of light in empty space (vacuum) providing us with a physically-motivated but very simplified base-approach.
@@
 \label{sec:trans}
-Note that the following considerations are only given for the tensor profile operating mode with tensor influence incorporated.
+Note that the following considerations are only given for the tensor profile operating mode with tensor influence incorporated and now the tensor properties are modeled on top.
 To obtain a kind of footprint of the tensor field, we set up ellipse equations for each glyph representation, obtained from applying PCA on each cell (tensor) in the grid.
-Concurrently, we extend our abstract mathematical operator \textit{propagate}, previously referenced in sect \ref{sec:scheme}, by a normalization strategy necessary for windowing weighting profiles, as it is the case for our elliptical transmission profiles, which we will capture in the following.
-At this point, we presume the precomputation of singular value decomposition and linear combination weights.
+Concurrently, we extend our abstract mathematical operator \textit{propagate}, previously defined in Sect.
+\ref{sec:scheme}, by a normalization strategy necessary for windowing weighting profiles, as it is the case for our elliptical transmission profiles, which we will capture in the following.
+At this point, we presume the precomputation of singular value decomposition (dependent of each cell) and linear combination weights (independent of cells).
 We map the singular values (corresponding to eigenvalues $\lambda$ here for simplicity) in decreasing order to the ellipses radii (half-axes) $a(\Delta x)$ and $b(\Delta y)$ as follows:
@@
 the x-axis by exploiting the $\mathop{atan2}$-function with 4-quadrant evaluation for the first singular vector (ordered decreasingly in value), since we scale the $x$-axis with the first singular value.
-As a preliminary preparation, this profile is precomputed (sampled) for all discrete angular steps for each cell (, e.g., $360$) and stored for reuitilization as it is the case for the previously mentioned linear combination weights as well.
-The weighting scheme which we apply is depicted in Fig.~\ref{weighting}.
+As a preliminary preparation, this profile is precomputed (sampled) for all discrete angular steps for each cell (e.g., $360$) and stored for reuitilization, as it is the case for the previously mentioned linear combination weights, as well.
+The weighting scheme, which we apply is depicted in Fig.~\ref{weighting}.
 \begin{figure}[!t]
@@
 \begin{align}
-	\Phi_{\alpha} &= \varepsilon_\alpha(\int_{-\pi/4}^{-\alpha/2}{T}(\omega){I}(\omega)\mathop{d\omega}+\int_{\alpha/2}^{\pi/4}{T}(\omega){I}(\omega)\mathop{d\omega})+\int_{-\alpha/2}^{\alpha/2}{T}(\omega){I}(\omega)\mathop{d\omega}\\
-	\Phi_{\beta} &= \varepsilon_\beta\int_{-\beta}^{\beta}{T}(\omega){I}(\omega)\mathop{d\omega}
+	\Phi_{\alpha} &= \varepsilon_\alpha(\int_{-\pi/4}^{-\alpha/2}{T}(\omega){I}(\omega)\mathop{d\omega}+\int_{\alpha/2}^{\pi/4}{T}(\omega){I}(\omega)\mathop{d\omega})+\int_{-\alpha/2}^{\alpha/2}{T}(\omega){I}(\omega)\mathop{d\omega},\\
+	\Phi_{\beta} &= \varepsilon_\beta\int_{-\beta}^{\beta}{T}(\omega){I}(\omega)\mathop{d\omega}.
 \end{align}
 
-This permits us to scale an oriented and clipped cosine-lobe $\cos_+$ with the total radiant flux with respect to its own total energy when integrated likewise.
-A discrete cosine lobe never reaches its theoretical sum of $2$ exactly in the interval $[{-\frac{\pi}{2}},{\frac{\pi}{2}}]$.
-Therefore it needs to be computed to account for sampling (discretization) errors and normalized by following relation (, e.g., $\sum_{\cos} \approx 1.9949204635834517$).
-\begin{align}
-\sum_{\cos_+} &= \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\cos_+(\omega)\mathop{d\omega} \approx 2,\\
-	\cos_k(\omega) &= \frac{\Phi_t}{\sum_{\cos_+}}cos_+(\omega-k\cdot\frac{\pi}{4})\hskip 20pt \mathit{with} \ k\in[0,7].
-\end{align}
 This specific profile is injected in form of a placement to the corresponding neighbor index $k$.
@@
 But caution, we should not generate any exceeding energy right?
-Now each tensor ellipse equation has a mean transmission rate $[\%]$, which turns out to be a measure or synonym for the total rate of the outgoing transmitted radiant intensity $\int_0^{2\pi}{T}(\omega)\mathop{d\omega}$.
+Each tensor ellipse equation features a mean transmission rate $[\%]$, which turns out to be a measure or synonym for the total rate of the outgoing transmitted radiant intensity $\int_0^{2\pi}{T}(\omega)\mathop{d\omega}$ $[\%]$.
 Since energy can not be generated from transmission profiles (only transmitting, not generating), we have to normalize every tensor ellipse equation with the highest mean transmission rate in the field, which is then transmitting the whole outgoing intensity $\int_0^{2\pi}{I}(\omega)\mathop{d\omega}$ in total.
@@
 \begin{align}
-	r_n(\omega) = \frac{1}{\mean{r}_{max}}r(\omega).
+	r_i(\omega) = \frac{1}{\mean{r}_{max}}r(\omega).
 \end{align}
-Notice, that this implements the concept of absorption implicitly, because tensors with a lower mean $\mean{r}$ than $100\%$ absorb energy wheras absorption equals ${\mu} = {1}-{\mean{r}}$ here.
-Now that we have coped with the problem of overall transmission, we are left with another problem that occurs with the weighting functions as they can still yield values above $1.0$.
-Consider following examples as $T(\omega)I(\omega)$ pairs (obviously normalized to $1.0$): $1.5\cdot 1.5=2.25$ and $0.5\cdot 0.5=0.25$, which disproves us that a mean $\bar{r}=1.0$ necessarily effects a mean transmission intensity $I_t$ of $1.0<1.25=\frac{2.5}{2}$.
-This would potentially generate intensity out of nothing which leads to crucially unstable behaviour, since we employ uncontrolled sources of energy in this case.
+Notice, that this implements the concept of absorption implicitly, because tensors with a lower mean $\mean{r}_i$ than $100\%$ absorb energy wheras absorption equals ${\mu} = {1}-{\mean{r}}$ here.
+Now, that we have coped with the problem of overall transmission, we are left with another problem that occurs with the weighting functions as they can still yield values above $1.0$.
+Remark, that the next example could be interpreted congruently to Fig.\ref{weighting}.
+Consider following examples as $T(\omega)I(\omega)$ pairs (obviously normalized to $\mean{r}=1.0$): $1.5\cdot 1.5=2.25$ and $0.5\cdot 0.5=0.25$, which already disproves us that a mean transmission rate $\mean{r}=1.0$ necessarily effects a mean transmission intensity $I_t$ of $1.0<1.25=\frac{2.5}{2}$ and thus to an overall transmitted intensity of $100\%$.
+This would potentially generate intensity out of nothing, which leads to crucially unstable behaviour, since we would employ uncontrolled sources of energy in this case.
 Well, but only in case energy conservation principles were not respected.
@@
 \begin{enumerate}
-	\item Normalization of $TI$ to $\overline{TI}=1.0$
+	\item Normalization of $T(\omega)I(\omega)$ to $\overline{TI}=1.0$
 	\item Subsequent scaling with mean intensity $\mean{I}$ for energy conservation principles
@@
    n_f\varepsilon_\alpha{T}(\omega){I}(\omega),& \text{if } k\bmod 2 = 0\\
-     n_f\varepsilon_\beta{T}(\omega){I}(\omega),              & \text{otherwise}
+     n_f\varepsilon_\beta{T}(\omega){I}(\omega),              & \text{otherwise}.
 \end{cases}
@@
 \begin{align}
-	\Phi_{\alpha,t,i} &= \int_\alpha{I}_{t,i}(\omega)\mathop{d\omega} = n_f\Phi_{\alpha,i},\\
-	\Phi_{\beta,t,i} &= \int_\beta{I}_{t,i}(\omega)\mathop{d\omega} = n_f\Phi_{\beta,i}.
+ \Phi_{t,i} = 
+\begin{cases}
+   \int_\alpha{I}_{t,i}(\omega)\mathop{d\omega} = n_f\Phi_{\alpha,i},& \text{if } k\bmod 2 = 0\\
+     \int_\beta{I}_{t,i}(\omega)\mathop{d\omega} = n_f\Phi_{\beta,i},              & \text{otherwise}.
+\end{cases}
 \end{align}
@@
 \end{enumerate}
-Remember that this operator needs to applied once to each cell to form a whole propagation cycle propagating intensities from source to target buffer.
-This approach ensures that no energy will be generated and therefore implements the 	physical law of energy conservation or Kirchhoff's current law if you want to think of a particle's diffusion process.
+Remember, that this operator needs to be applied once to each cell to form a whole propagation cycle propagating intensities from source to target buffer.
+This approach ensures, that no energy will be generated and therefore implements the physical law of energy conservation or Kirchhoff's current law if you want to think of a electronic network.
 Consequently, this allows us to assume that all sources of power (energy) are user-defined.
-In fact the approach has proven to be robust w.r.t.
+In fact the approach, has proven to be robust w.r.t.
 energy losses for several iterations as depicted in Chap.~\ref{chap:eval} in the stochastic error graph.
@@
 In that sense of imagination, we would virtually imprint every tensor field into a crystalline structure as a preliminary initialization step.
-For an every day example, this can be observed in, e.g., the gemstone tiger's eyes' cat's eye effect, where crystal fibers align with the crystal axes respectively and effect a brightened line which never changes position or orientation when rotating the crystal.
+For an every day example, this can be observed in, e.g., the gemstone tiger's eye's cat's eye effect, where crystal fibers align with the crystal axes respectively and effect a brightened line which never changes position or orientation when rotating the crystal.
 These oriented crysal fibers are then representing the biased propagation directions of light inside of the crystal lattice.
-The propability density function $P(\omega)$, for a photon being scattered in a particular direction $\omega$ \cite{ryder}, is then described as:
+The propability density function $P(\omega)$, for a photon being scattered (redistributed) in a particular direction $\omega$ \cite{ryder}, is then described as:
 \begin{align}
@@
 Each tensor is then a unique footprint which can be precomputed for convenience and described as polar profile just as our previously mentioned intensity profiles.
-We apply these polar profiles in a sense of transmission profiles as discussed in Sect.~\ref{sec:polar}.
-
-\section{Light Propagation Scheme}
-To evaluate the plausibility of the light distribution results obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates:
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \includegraphics[width=0.8\textwidth]{img/isotropic.png}
-    \label{a)}
-	a) light source in top center
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
+We apply these polar profiles in a sense of transmission profiles as previously discussed in Sect.~\ref{sec:polar}.
+
+\section{Light Distributions}
+To evaluate the plausibility of the light distribution results, obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates:
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=0.8\textwidth]{img/iso_sat.png}
+    \label{a)}
+	a)~light source in top center
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+  \centering
     \includegraphics[width=0.8\textwidth]{img/isotropic-all-center.png}
     \label{b)}
-	b) light source in center
+	b)~light source in center
   \end{minipage}
@@
 \end{figure}
-One can observe here, how light is propagated with no directional bias leading to an absolutely isotropic distribution in case  b), which is symmetric as expected.
+One can observe in Fig.~\ref{isotropic}, how light is propagated with no directional bias leading to an absolutely isotropic distribution in case  b), which is symmetric as expected.
 This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in Sects.~\ref{sec:att} and \ref{sect:conserve}.
@@
 
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/iso_scalar.png}
-    \label{a)}
-	a) light source in center
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/iso_sat.png}
-    \label{b)}
-	b) light source in center (rescaled)
-  \end{minipage}
-\caption{final light distributions for ``isotropy''-test field}
-\label{rings-tests}
-\end{figure}
-Note that we skipped the top center example, since we consider it a trivial add-on example.
-When we observe these scalar heat map plots, we first notice the red dots, which always represent light sources (since they are captured the most energy) in these cases.
+When we observe these polar and scalar heat map plots in Fig.~\ref{isotropic}, we first notice the red dots, which always represent light sources (since they are captured the most energy) in these cases.
 Specifically, the heat maps are obtained from the same light distributions we plot as polar profiles and should hence be interpreted concurrently.
@@
     \label{a)}
-	a) antisymmtric light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/rings-two-special2.png}
-    \label{b)}
-	b) symmetric light sources
+	a)~antisymmetric light sources
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.9\textwidth]{img/cos_parallel.png}
+    \label{b)}
+	b)~symmetric light sources
   \end{minipage}
@@
 \end{figure}
+
 We consider Fig.~\ref{rings-tests} to regard the behavior for circular tensor field lines.
@@
 Note the symmetry for both cases depicted.
-Again, we plot the final light distributions as heat maps:
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/cos_parallel.png}
-    \label{a)}
-	a) antisymmtric light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/cos_anti.png}
-    \label{b)}
-	b) symmetric light sources
+Again, we plot the final light distributions as heat maps: It is noticeable in Fig.
+\ref{rings-tests}, that the intensities concentrate preferably on the circular tensor field lines, which proves the approach's controlling ability expressing as direction.
+ The results for our tensor field ``spiral'' can be interpreted in analogous fashion as the previously discussed ``rings''-testfield in Fig.
+\ref{spiral-tests}.
+The influence incorporated by the tensor field is easily noticeable in this case as well.
+
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.9\textwidth]{img/spiral-two-dense.png}
+    \label{a)}
+	a)~two light sources
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.9\textwidth]{img/spiral-two-wide.png}
+    \label{b)}
+	b)~symmetric light sources
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.9\textwidth]{img/spiral-temp2.png}
+    \label{a)}
+	c)~two light sources
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.9\textwidth]{img/spiral_full.png}
+    \label{b)}
+	d)~four light sources
   \end{minipage}
@@
 \end{figure}
-It is noticeable, that the intensities concentrate preferably on the circular tensor field lines, which proves the approach's controlling ability expressing as direction.
-The results for our tensor field ``spiral'' can be interpreted in analogous fashion as the previously discussed ``rings''-testfield.
-The influence incorporated by the tensor field is easily noticeable in this case as well:
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral-two-dense.png}
-    \label{a)}
-	a) dense light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral-two-wide.png}
-    \label{b)}
-	b) symmetric light sources
-  \end{minipage}
-\caption{final light propagation distributions for ``spiral''-test field}
-\label{spiral-tests}
-\end{figure}
-For the ``spiral'' tensor field we observe the spiral shape literally absorbing the intensities into its spiral arms, which becomes even more clear when we again observe the heat map representation:
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral-temp2.png}
-    \label{a)}
-	a) two light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral_full.png}
-    \label{b)}
-	b) four light sources
-  \end{minipage}
-\caption{final light propagation distributions for ``spiral''-test field}
-\label{spiral-tests}
-\end{figure}
-
-The star example reveals another interesting feature of the approach: if the intensity impulsively hits an isotropic region, it gets directed according to its intensity profile (neglecting the transmission profile) which implements the principle of inertia for directed light propagation present in vacuum (light never changes its direction):
+In Fig.
+\ref{spiral-tests}a)\& b) we observe the spiral shape literally absorbing the intensities into its spiral arms, which becomes even more clear when we again observe the heat map representation in Fig.
+\ref{spiral-tests}c)\& d).
+
+The star example reveals another interesting feature of the approach in Fig.
+\ref{star-tests1} if the intensity impulsively hits an isotropic region, it gets directed according to its intensity profile (neglecting the transmission profile) which implements the principle of inertia for directed light propagation present in vacuum (light never changes its direction).
+Considering Fig.
+\ref{spiral-tests} we remark, that the diagonal intensities are attenuated more strongly by default, which introduces a bias into the light paths which are chosen along the axes with higher propability, effectively.
+This gets even clearer, when we plot the heat map depiction in Fig.
+\ref{star-tests2} again.
 \begin{figure}[!t]
@@
   \begin{minipage}{0.31\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/star-0,0.png}
     \label{a)}
-	a) single light source
+	a)~single light source
   \end{minipage}
   \begin{minipage}{0.31\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/star-two-wide.png}
     \label{b)}
-	b) symmetric light sources
+	b)~symmetric light sources
   \end{minipage}
@@
     \label{c)}
-	c) dual light sources
+	c)~dual light sources
   \end{minipage}
 \caption{final light propagation distributions for ``star'' test field}
-\label{spiral-tests}
-\end{figure}
-Remark, that the diagonal intensities are attenuated more strongly by default, which introduces a bias into the light paths which are chosen along the axes with higher propability, effectively.
-This gets even clearer, when we plot the heat map depiction again:
+\label{star-tests1}
+\end{figure}
+
 \begin{figure}[!t]
@@
   \begin{minipage}{0.31\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/star_wide.png}
     \label{a)}
-	a) symmetric light sources
+	a)~symmetric light sources
   \end{minipage}
   \begin{minipage}{0.31\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/star_close.png}
     \label{b)}
-	b) dual light sources
+	b)~dual light sources
   \end{minipage}
 \caption{final light propagation distributions for ``star'' test field}
-\label{spiral-tests}
+\label{star-tests2}
 \end{figure}
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/scalar_iso_left.png}
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/scalar_iso_right.png}
@@
    \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/scalar_iso_diff.png}
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/scalar_iso_leftangle.png}
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/scalar_iso_rightangle.png}

### 2019-07-08 thesis-template/msc-thesis.tex
update

@@
 Also note, that this approach satisfies energy conservation and propagation attenuation principles and applies to the principles of light propagation in vacuum (by default).
-We infer, that we can assume to follow physical models of spherical waves.
+We infer, that we can assume to follow physical models of circular waves.
 The verification of these principal requirements and specifications can be observed in the evaluation in Chap.~\ref{chap:eval}.
@@
 \section{Light Distributions}
-To evaluate the plausibility of the light distribution results, obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates:
-\begin{figure}[!t]
-\centering
+To evaluate the plausibility of the light distribution results, obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates.
+As an alternative representation, we plot the energy distribution summed over all directions for each cell, which yields a scalar field in Fig.
+\ref{isotropic} b).
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=0.8\textwidth]{img/isotropic-all-center.png}
+    \label{a)}
+	a)~light source in top center
+  \end{minipage}
   \begin{minipage}{0.4\textwidth}
@@
     \includegraphics[width=0.8\textwidth]{img/iso_sat.png}
-    \label{a)}
-	a)~light source in top center
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/isotropic-all-center.png}
     \label{b)}
@@
 \end{figure}
-One can observe in Fig.~\ref{isotropic}, how light is propagated with no directional bias leading to an absolutely isotropic distribution in case  b), which is symmetric as expected.
+One can observe in Fig.~\ref{isotropic}, how light is propagated with no directional bias leading to an absolutely isotropic distribution, which is symmetric as expected.
 This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in Sects.~\ref{sec:att} and \ref{sect:conserve}.
-We can also depict the final distribution as a scalar field when using the total energy in each cell as scalar:
-
-When we observe these polar and scalar heat map plots in Fig.~\ref{isotropic}, we first notice the red dots, which always represent light sources (since they are captured the most energy) in these cases.
-Specifically, the heat maps are obtained from the same light distributions we plot as polar profiles and should hence be interpreted concurrently.
+When we observe these polar and scalar heat map plots, we first notice the red dots, which always represent light sources (since they cover the most energy) in these cases.
+Specifically, the heat maps are obtained from the same light distributions, which we plot as polar profiles and should hence be interpreted concurrently.
 Remark the propagation front indicating a circular wave (circular isolines in a)).
-In this next example, we use the tensor field ``rings'' as input data for the approach and observe how the intensities are channeled in a circular orbit around the central origin.
+In this next example in Fig.
+\ref{rings-tests}, we use the tensor field ``rings'' as input data for the approach and observe how the intensities are channeled in a circular orbit around the central origin.
 The anisotropy measures decide whether the transmission profile is more narrow (radical) or more wide (liberal):
@@
 As expected, the intensities get directed in circular orbits by the tensor ellipsoid transmission profiles.
-Note the symmetry for both cases depicted.
+Note the symmetry for both light source directions.
 Again, we plot the final light distributions as heat maps: It is noticeable in Fig.
-\ref{rings-tests}, that the intensities concentrate preferably on the circular tensor field lines, which proves the approach's controlling ability expressing as direction.
- The results for our tensor field ``spiral'' can be interpreted in analogous fashion as the previously discussed ``rings''-testfield in Fig.
+\ref{rings-tests}, that the intensities concentrate preferably on the circular tensor field lines, which proves the approach's controlling ability expressed through redirection.
+The results for our tensor field ``spiral'' can be interpreted in analogous fashion, similar to the previously discussed ``rings''-testfield in Fig.
 \ref{spiral-tests}.
@@
 In Fig.
-\ref{spiral-tests}a)\& b) we observe the spiral shape literally absorbing the intensities into its spiral arms, which becomes even more clear when we again observe the heat map representation in Fig.
-\ref{spiral-tests}c)\& d).
+\ref{spiral-tests}a),b) we observe the spiral shape literally absorbing the intensities into its spiral arms, which becomes even more clear, when we again observe the heat map representation in Fig.
+\ref{spiral-tests}c),d).
 
 The star example reveals another interesting feature of the approach in Fig.
-\ref{star-tests1} if the intensity impulsively hits an isotropic region, it gets directed according to its intensity profile (neglecting the transmission profile) which implements the principle of inertia for directed light propagation present in vacuum (light never changes its direction).
-Considering Fig.
-\ref{spiral-tests} we remark, that the diagonal intensities are attenuated more strongly by default, which introduces a bias into the light paths which are chosen along the axes with higher propability, effectively.
-This gets even clearer, when we plot the heat map depiction in Fig.
-\ref{star-tests2} again.
+\ref{star-tests1}.
+When the intensity impulsively hits an isotropic region, it gets directed according to its intensity profile (neglecting the transmission profile), which implements the principle of inertia for directed light propagation present in vacuum (light never changes its direction).
 \begin{figure}[!t]
@@
 \end{figure}
-
+Considering the star test field in Fig.
+\ref{star-tests1}, we remark, that the diagonal intensities are attenuated more strongly by default, which introduces a bias into the light paths which are chosen along the axes with higher propability, effectively.
+This gets even clearer, when we plot the heat map depiction in Fig.
+\ref{star-tests2} again.
 \begin{figure}[!t]
@@
 \end{figure}
-Note that we skipped the single light source example, since we consider it a trivial add-on example. We can regard the intensities condensed inside of a narrow beam of light.
-Note, that a logarithmic scale needs to be incorporated to produce results on a large scale in a sense of higher spatial resolution (e.g.: $101\times 101$), since the propagation attenuation follows the distance to the origin inversely proportional.
+Note that we skipped the single light source example, since we consider it a trivial add-on example.
+We can regard the intensities condensed inside of a narrow beam of light.
+Remark, that a logarithmic scale needs to be incorporated to produce results on a large scale in a sense of higher spatial resolution (e.g.: $101\times 101$), since the propagation attenuation follows the distance to the origin inversely proportional ($\sim\frac{1}{r}$) for circular waves.
 
 \section{Light Transport Gradient}
-In this section, we will motivate the main contribution on tensor field analysis in this work.
+In this section, we will motivate the main contribution to tensor field analysis in this work.
 The approach called Light Transport Gradient (LTG) states a FTLE-related approach performed on tensor fields, but employing whole distributions of trajectories interpreted as light transmission profiles.
@@
 In place of a 2D FTLE field concerning only the direction of the major eigenvector in each cell, we obtain all the directionally dependent FTLEs unrolled into slices and hence transform the FTLE field into a function FTLE$(\omega\mapsto z)$.
-The resulting volume contains a manifold which is not self-overlapping and is a continous plane for a continous tensor field.
+The resulting volume contains Hlawatsch's result in form of a manifold which is not self-overlapping and is a continous plane for a continous tensor field.
 \label{sec:ltg}
 \subsection{Evolution}
-Now, that we are able to propagate intensity in grid, directed by the eigenvectors of the tensor field, we aim to do something more elaborate to segment ROIs characterised by divergence/convergence of tensor field lines and drastic changes in anisotropy.
-For this purpose, we imagine kind of a local and spatial gradient, as we place light sources in two close-by positions and eventually in two close-by directions and analyze the resulting final light distribution.
-We consider the light distributions and their difference image for the \enquote{isotropic} and the \enquote{drain} test field to have an idea, for which kind or types of cases/configurations we obtain inherently different light distributions, in particular.
+Now, that we are able to propagate intensity in the grid, directed by the eigenvectors of the tensor field, we aim to do something more elaborate to segment ROIs characterized by divergence/convergence of tensor field lines and drastic changes in anisotropy.
+For this purpose, we imagine kind of a local and spatial gradient, as we place light sources in two close-by positions and eventually also in two close-by directions and analyze the resulting final light distribution.
+We consider the light distributions and their difference image for the \enquote{isotropic} and the \enquote{drain} test field to have an idea, for which kind or types of cases, i.e.
+initial configurations we obtain inherently different light distributions, in particular.
 We choose to use these test cases, because they constitute absolutely pure examples par excellence for a FTLE computation exposing no (\enquote{isotropic}), or on the other hand raw (\enquote{drain}) tensor field lines diverging.
@@
 That is, per default, light distributions cancel each other out for two close-by light source positions or directions (or more specifically: for continous overlaps in their distributions).
-In this case, for the \enquote{isotropic} test field, we have no tensor field lines diverging leading to a small overall difference for light sources in finite distance.
+In this case, for the \enquote{isotropic} test field, we have no tensor field lines diverging leading to a small overlap and thus to a decreased overall difference for light sources in finite distance.
 
@@
     \label{a)}
-	a) final light distribution for left spatial offset
+	a)~final light distribution for left spatial offset
   \end{minipage} $-$
@@
     \label{b)}
-	b) final light distribution for right spatial offset
+	b)~final light distribution for right spatial offset
   \end{minipage} $=$
@@
     \label{b)}
-	difference image for a) and b)
+	c)~difference image for a) and b)
   \end{minipage}
@@
     \label{a)}
-	a) final light distribution for left angular offset
+	a)~final light distribution for left angular offset
   \end{minipage} $-$
@@
     \label{b)}
-	b) final light distribution for right angular offset
+	b)~final light distribution for right angular offset
   \end{minipage} $=$
    \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/scalar_iso_anglediff.png}
     \label{b)}
-	difference image for a) and b)
+	c)~difference image for a) and b)
   \end{minipage}
@@
 \end{figure}
-For the angular offset, we can again observe the light distributions cancelling each other out (per default) in the center (middle) of the difference image.
+For the angular offset, we can again observe the light distributions cancel each other out (per default) in the center (middle) of the difference image.
 For now, we can take home from these examples, that light distributions are cancelling each other out within their overlap in general, in case no anisotropy was introduced into the test tensor field.
-We again analyze the drain test field (cf.~Fig.~\ref{drain} in Sect.~\ref{chap:grundlagen}) for the behavior of the resulting light distributions in scalar field representation.
+We again analyze the drain test field (cf.~Fig.~\ref{test_fields2}a) in Sect.~\ref{chap:grundlagen}) for the behavior of the resulting light distributions in scalar field representation.
 This following example illustrated in Fig.~\ref{evolution1} now introduces raw tensor field lines diverging w.r.t.
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/ftle_left.png}
     \label{a)}
-	a) final light distribution for left spatial offset
+	a)~final light distribution for left spatial offset
   \end{minipage} $-$
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/ftle_right.png}
     \label{b)}
-	b) final light distribution for right spatial offset
+	b)~final light distribution for right spatial offset
   \end{minipage} $=$
    \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[width=\textwidth]{img/ftle_diff.png}
     \label{b)}
-	difference image for a) and b)
+	c)~difference image for a) and b)
   \end{minipage}
@@
 \end{figure}
-We can see the light distributions diverge following tensor field lines, which in turn form these curves depicted in Fig.~\ref{evolution1}.
+We can remark the light distributions diverge following tensor field lines, which in turn form these curves depicted in Fig.~\ref{evolution1}.
 The separation of path (tensor field) lines and thus of light distributions is maximized at the center location of the field.
@@
   \begin{minipage}{0.3\textwidth}
+\centering
     \includegraphics[width=\textwidth]{img/ftle_leftangle.png}
     \label{a)}
-	a) final light distribution for left angular offset
+	a)~final light distribution for left angular offset
   \end{minipage} $-$
   \begin{minipage}{0.3\textwidth}
+\centering
     \includegraphics[width=\textwidth]{img/ftle_rightangle.png}
     \label{b)}
-	b) final light distribution for right angular offset
+	b)~final light distribution for right angular offset
   \end{minipage} $=$
    \begin{minipage}{0.3\textwidth}
+\centering
     \includegraphics[width=\textwidth]{img/ftle_anglediff.png}
     \label{b)}
-	difference image for a) and b)
+	c)~difference image for a) and b)
   \end{minipage}
@@
 Imagine accumulating each pixel in the images to obtain the cumulative total energy in the (difference) grid, which could now be compared for each light source position and direction in terms of proper min-max scaling.
-Now imagine we would accumulate these differences and assign each position a gradient magnitude in form of a scalar value with the summed difference of the final light distributions in the neighborhood.
-What would we eventually be yielded with.
-Well, we would obtain another scalar field considered to be the gradient image of the light distributions.
-We could very well extend this process for every light source direction discretized (sampled) and obtain a $3D$ volume representing the topological structure of the tensor field for every direction.
+Now imagine we would accumulate these differences and assign each position a gradient magnitude with components in form of a scalar value with the summed difference of the final light distributions in the neighborhood.
+What would we finally be yielded with?
+Well, we would obtain another scalar field considered to be the gradient image of the resulting light distributions.
+We could very well extend this process for every light source direction discretized (sampled) and obtain a 3D volume representing the topological structure of the tensor field for every direction.
 All we need, is a formal definition for a central differences gradient approximation for light distributions given in polar profiles.
-The main difference to FTLE in vector field visualization is, that we use whole distributions of trajectories and hence need to use a starting direction for the distribution of trajectories, instead of just moving with the flow, as a tensor introduces a whole distribution of directions and does not only depend on the inital position, but also on the direction.
+The main difference to FTLE in vector field visualization is, that we use whole distributions of trajectories and hence use an initial direction for the distribution of trajectories, instead of just moving with the flow, as a tensor introduces a whole distribution of directions and does not only depend on the inital position, but also on the direction.
 This corresponding to the volume emerging from the FTLE plane image, as we introduce a new dimension $\omega$.
-Remark that we denote the final buffer result as $\mathbf{b}=\mathcal{C}_{final}$ in the following.
+Remark that we denote the final buffer result concerning the propagatipon scheme as $\mathbf{b}=\mathcal{C}_{final}$ in the following.
 \subsection{Definition}
  As tensor field lines are bidirectional, convergence and divergence can not be reasonably discerned in that case.
-We aim to extract ridges from a FTLE-like field to grasp the key regions separating anisotropy behavior (, i.e., changes in direction/tensor magnitude) in the field segmented as height ridges in the field, which are in turn LCS (Lagrangian Coherent Structures) well known for their rich substance in physical meaning.
+We aim to extract ridges from a FTLE-like field to grasp the key regions separating anisotropy behavior (i.e., changes in direction/tensor magnitude) segmented as height ridges in the field, which are in turn LCS (Lagrangian Coherent Structures), well known for their rich substance in physical meaning.
 They often indicate transport barriers in flow fields.
@@
 \end{align}
- We choose to use a constant offset of $\frac{\pi}{6}\mathop{rad}$ for the angle because we want to prevent from generating results with content implicitly depending on the angular resolution and also since we need to exceed $\beta$ to have a uniform and consistent placing of cosine lobes (inside diagonal cones) concerning neighbor indices.
+ We choose to use a constant offset of $\frac{\pi}{6}\mathop{rad}$ for the angle because we want to prevent from generating results, with content implicitly depending on the angular resolution and also since we need to exceed $\beta$ to have a uniform and consistent placing of cosine lobes (inside diagonal cones) concerning neighbor indices.
 Remark that the results do not depend on the spatial resolution implicitly, because increasement acts as a domain extension for and does effectively not increase the sampling rate of the propagation scheme.
-Also note that this step is very expensive considering runtime-costs since it follows the following runtime observation.
+Also note, that this step is very expensive considering runtime-costs since it follows the following runtime observation.
 We have $\mathop{width}\cdot \mathop{height}\cdot \mathop{steps}$ numbers in the grid and on top $\mathop{width}\cdot \mathop{height}\cdot \mathop{steps}$ buffers in total, which need to be processed.
@@
 We note, that our invented technique stands out compared to the already existent FTLE on tensor fields invented by Hlawatsch et al.
-\cite{hlawatsch}, as we use whole distributions of directions instead of following the major eigenvector strictly.
+\cite{hlawatsch}, as we use whole distributions of directions instead of following the major eigenvectors strictly.
 These directional distributions allow for a complete analysis of the tensor field accounting for the incoming direction, as well.
@@
 We will try to show the functionality of the method in a few simple examples.
-At first, we will have a null example to show the offset response of the approach induced by total spherical anisotropy and no absorption in the field.
-We consider the example \enquote{isotropic}, concerned to have no influence on the light propagation, as it models homogeneous identity matrices inducing absolutely isotropic transmission (every incident distribution is transmitted as is with no transmission weighting incorporated).
+At first, we will have a null example to show the offset response of the approach, induced by total spherical anisotropy and no absorption in the field.
+We consider the example \enquote{Isotropic}, concerned to have no influence on the light propagation, as it models homogeneous identity matrices inducing absolutely isotropic transmission (every incident distribution is transmitted as is with no transmission weighting incorporated).
 This corresponds to the \enquote{non - tensor field} mode of the propagation scheme having no influence of any tensor field incorporated.
@@
 
-When we look at the FTLE field, on the top right in Fig.~\ref{ftle_iso}, we determine that we obtain a more or less uniform reponse (, if we would neglect the edges) just as expected.
+When we look at the FTLE field in Fig.~\ref{ftle_iso}b), we determine that we obtain a more or less uniform reponse, if we would neglect the edges, just as expected.
 Because the divergence of TFLs is distributed uniformly in this test field, we get approximately the same response for any position, but yet introducing a systematic bias against the light source direction.
@@
 As noticeable, no height ridges are detected in this case as expected, since the test field does not pose any specific regions exhibiting diverging tensor field lines.
-We consider this result the absolute offset response, whereas energies are redistributed and separated more strictly for introducing local features into the grid (field), leading to greater differences, which are in turn emphasized by local regions of interest displayed in deep red at their corresponding positions.
-Imagine this process comparable with ink absorbed by dry paper.
-As we are also interested in the whole ensemble of directions, we can use a volume or surface representation for the $n+1=3$-dimensional grid which results from our FTLE computation.
+We consider this result the absolute offset response, whereas energies are redistributed and separated more strictly for introducing local features into the grid (field), leading to greater differences, which are in turn emphasized by local ROIs, displayed in deep red at their corresponding positions.
+Imagine this process comparable with ink being absorbed by dry paper.
+As we are also interested in the whole ensemble of directions, we can use a volume or surface representation for the $n+1=$3-dimensional grid which results from our FTLE computation.
 The volume grid is what we call the final footprint of the tensor field which is necessarily derived from the initial glyph footprint.
@@
 We observe the clear straight ridge in the example in Fig.~\ref{ftle_absorb}.
-As expected, the ridge is located on the vertivcal center line.
+As expected, the ridge is located on the vertical center line.
 Note, that we (min-max) normalize each transmission profile with the highest mean transmission rate (tensor magnitude) in the field.
@@
     \label{a)}
-	a)
+	a)~TFLs and glyphs
   \end{minipage}
@@
     \label{b)}
-	b)
+	b)~LTG($90^\circ$) for a)
   \end{minipage}
@@
     \label{b)}
-	c)
-  \end{minipage}
-  \caption{``drain'' test field: a) TFLs and glyphs, b) LTG($90^\circ$) for a), c) gradient for b)}
+	c)~Gradient for b)
+  \end{minipage}
+  \caption{``drain'' test field}
 \label{ftle_base}
@@
 A gradient filter could be applied to extract and segment edges from the image as a preprocessing step for computer vision, which would look like the image in Fig.~\ref{ftle_base} c).
-
-These ridges represent ROIs with underlying Lagrangian Coherent Structures, which we already discussed in Sect.~\ref{sect:motiv}.
+These ridges represent ROIs with underlying Lagrangian Coherent Structures, which we already discussed in Sect.~\ref{sec:motiv}.
 It could for example represent a stress field inside of a mechanical tool with a central axis with maximum stress condensation.
@@
 \end{figure}
-Regarding the ridge for the \enquote{drain} test field, we note that it is a little smoothed compared to the axis-aligned ridges, as they obtain a larger propagation attenuation (because of the larger distance) leading to a stronger separation of trajectories in general and hence to a smoothed FTLE response.
+Regarding the ridge for the \enquote{diagonal drain} test field, we note that it is a little smoothed compared to the axis-aligned ridges, as they obtain a larger propagation attenuation (because of the larger distance) leading to a stronger separation of trajectories in general and hence to a smoothed FTLE response.
 Also we note that the divergence of the tensor field lines is reduced compared to the classic \enquote{drain} example of the FTLE.
@@
 Note that the $z$-axis represents the light source direction angle $\omega$ in this case.
-The contour filter reveals the star-shaped structure of the tensor field which rotates along the angle axis.
+Observing the volumes in Fig.
+\ref{inverse_volume}, the contour filter reveals the star-shaped structure of the tensor field which rotates along the angle axis.
 This is the effect we already captured for the single slice: For each axis direction we can detect ridges from parallel up to orthogonal to our predefined stimulus (light source) direction, but not in antiparallel direction.
@@
 We can also try to segment height ridges represented as surfaces from the resulting volume grids as depicted in Fig \ref{inverse_contour}.
-The orthogonal ridges in this example reach out to the edges because they are gaining a bias by never offsetting the light source in a transmission direction (ellipsoid orientation) for orthogonals.
+The orthogonal ridges in this example reach out to the edges because they are gaining a bias, through never offsetting the light source in a transmission direction (ellipsoid orientation) for orthogonals.
 This turns out to be a general feature of the method, as we will later verify in Figs.
@@
   \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/rings_isovolume.png}
+    \includegraphics[height=\textwidth]{img_new/rings_vertridges.png}
 	b)
@@
   \end{minipage}
-    \caption{LTG volumes: ``rings'' test field: a) Contour, b) IsoVolume, c) Volume}
+    \caption{LTG volumes: ``rings'' test field: a) Contour, b) VCGRidge, c) Volume}
 \label{rings_contour}

### 2019-07-08 thesis-template/dummy.tex
update

@@
 \usepackage{subfig}
+\usepackage{url}
 %\usepackage{subfig}
@@
 \begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral-two-dense.png}
-    \label{a)}
-	a)~two light sources
+\hskip 22pt
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/diverge.png}
+	a)
+   \label{diverge}
+   \captionsetup{font={footnotesize,bf,it}}
+  \end{minipage} \hskip 30pt
+  %\raggedright
+  \begin{minipage}[!t]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{img/ftle_gyre.png}
+	b)
+    \captionsetup{font={footnotesize,bf,it}}
   \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral-two-wide.png}
-    \label{b)}
-	b)~symmetric light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral-temp2.png}
-    \label{a)}
-	c)~two light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral_full.png}
-    \label{b)}
-	d)~four light sources
-  \end{minipage}
-\caption{final light propagation distributions for ``spiral''-test field}
-\label{spiral-tests}
+  \caption{double gyre: a) diverging streamlines, b) FTLE field}
+  \label{ftle_gyre}
+  \caption*{source: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/}}
 \end{figure}

### 2019-07-08 thesis-template/msc-thesis.tex
update

@@
 	\item a light transport model (propagation scheme) following basic but crucial physical principles,
-	\item visualization of the propagation results in $2D$ polar profiles and scalar fields,
+	\item visualization of the propagation results in 2D polar profiles and scalar fields,
 	\item application of this model for tensor field visualization interpreting tensors as light transmission properties,
-	\item a FTLE (Finite-time Lyapunov exponents)-related approach called light transport gradient (LTG) for visualizing key structures, namely LCS (Lagrangian coherent structures) in $2D$ second-order tensor fields, and
+	\item a FTLE (Finite-time Lyapunov exponents)-related approach called light transport gradient (LTG) for visualizing key structures, namely LCS (Lagrangian coherent structures) in 2D second-order tensor fields, and
 	\item application of our approach to both synthetic and real data involving brain and heart datasets.
@@
 Necessity begets ingenuity as always, which lets visualization methods and systems emerge in this context.
-Concurrently, the problem setting of this work encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids, commonly consisting of second-order tensors (cf.~Cauchy stress tensor).
+Concurrently, the problem setting of this work encomprises the visualization of tensor fields, i.e.~, mostly 2D/3D-grids, commonly consisting of second-order tensors (cf.~Cauchy stress tensor).
 In general, these grids span a spatial domain with sample positions indicated by vertices or cell centers in the dual grid, depending on the mode of interpretation.
@@
 Furthermore, this propagation scheme should either be used to visualize the light transport paths, defined by the anisotropy characteristics of tensor fields and tensor magnitudes of a given tensor field in a 2D grid.
-Or, as a secondary purpose, be used to visualize the light transport in a defined, given geometrical scene in a $2D$ grid.
+Or, as a secondary purpose, be used to visualize the light transport in a defined, given geometrical scene in a 2D grid.
 However, the tensor field needs to be preprocessed somehow to generate a footprint map of transmission profiles defined by the tensor field's anisotropy characteristics, which we will handle through PCA yielding a principal component ellipsoid.
@@
 We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges representing LCS (Lagrangian Coherent Structures).
-These structures can be classified (identified) as repelling, attracting and saddling (i.e., forming a saddle) local features, which frequently represent key structures in nature, such as moraines in glaciers or vortices in flow fields.
+These structures can be classified (identified) as repelling, attracting and saddling (i.e.~, forming a saddle) local features, which frequently represent key structures in nature, such as moraines in glaciers or vortices in flow fields.
 Commonly, in vector field visualization, LCS separate regions of different flow behavior, but in our case they separate regions, or in other words subdomains, of varying directional distributions (manifesting in diverging/converging tensor field lines leading to compressions, tensions, roations or varying tensor magnitude causing volume changes) represented by tensors, as we use our tensor-controlled light transport visualization (propagation scheme) as a basis or framework to compute the FTLE field.
@@
 Since we are not able to cover the whole scope in this work, we will focus on the most relevant works on global illumination and tensor field visualization.
-We will talk about the publications from Hlawatsch et al.
-\cite{hlawatsch} and Zirr et al.
-\cite{zirr} in particular, since they constitute thematically close related works.
+We will talk about the publications from Hlawatsch et al.~\cite{hlawatsch} and Zirr et al.~\cite{zirr} in particular, since they constitute thematically close related works.
 But first, we will cover related works on global illumination methods, as these form a basis or prerequisite for our work.
@@
 This work addresses artifacts emerging from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}.
-These methods are considered to be too slow to \enquote{run at interactive framerates} by Dachsbacher et al.
-\cite{dachsbacher}, though \enquote{efficient and potentially highly parallel}.
+These methods are considered to be too slow to \enquote{run at interactive framerates} by Dachsbacher et al.~\cite{dachsbacher}, though \enquote{efficient and potentially highly parallel}.
 Our work is inspired by Lattice-Boltzmann methods (volume grid radiosity) and the lattice-based discrete ordinate methods concerning the light propagation scheme.
@@
 \section{Symmetric Tensor Field Visualization}
-Zheng \& and Pang \cite{pang&zheng} et al.
-proposed a texture-based visualization approach called HyperLIC, which extends the concept of LIC to symmetric tensor fields by using an anisotropic 2D-filter kernel oriented along the major or minor eigenvectors.
-The concept of following the major/minor eigenvector along tensor field lines was first invented by Vilanova et al.
-\cite{vilanova}.
+Zheng \& and Pang \cite{pang&zheng} et al.~proposed a texture-based visualization approach called HyperLIC, which extends the concept of LIC to symmetric tensor fields by using an anisotropic 2D-filter kernel oriented along the major or minor eigenvectors.
+The concept of following the major/minor eigenvector along tensor field lines was first invented by Vilanova et al.~\cite{vilanova}.
 Tensorlines \cite{weinstein} introduce a kind of artificial inertia (running average), making them more resistent to noise.
 The notion of tensor field topology (critical, degenerate points, separatrices) and the concept of hyperstreamlines was first introduced by Delmarcelle and Hesselink \cite{delmarcelle&hesselink}.
-The topology in $3D$ tensor fields was first analyzed by Zheng et al.
-\cite{zheng}.
-The placement of tensor field lines as hyperstreamlines for glyph packing was recently improved by Spencer et al.
-\cite{spencer}.
-A good overview of seeding strategies concerning glyphs is given by McLoughlin et al.
-\cite{mcloughlin}.
-Feng et al.
-\cite{feng} used Voronoi tessellation for placing the glyphs.
+The topology in 3D tensor fields was first analyzed by Zheng et al.~\cite{zheng}.
+The placement of tensor field lines as hyperstreamlines for glyph packing was recently improved by Spencer et al.~\cite{spencer}.
+A good overview of seeding strategies concerning glyphs is given by McLoughlin et al.~\cite{mcloughlin}.
+Feng et al.~\cite{feng} used Voronoi tessellation for placing the glyphs.
 Superquadric glyphs have been proposed by Schultz and Kindlmann \cite{schultz&kindlmann} for general (non-positive definite) tensors.
@@
 These coefficients can also be visualized by direct volume rendering \cite{kindlmann2}.
-Hlawatsch et al.
-\cite{hlawatsch} obtain a coordinate (flow) map from tensor field lines following major/minor eigenvectors (fibre trajectories) using the maximum eigenvalue of the Cauchy-Green deformation tensor to compute the FTLE-field in homogeneous regions.
+Hlawatsch et al.~\cite{hlawatsch} obtain a coordinate (flow) map from tensor field lines following major/minor eigenvectors (fibre trajectories) using the maximum eigenvalue of the Cauchy-Green deformation tensor to compute the FTLE-field in homogeneous regions.
 In a similar manner, we aim to use Global Illumination methods propagating light, directed by tensor fields, to obtain a final light distribution (flow map) through successive light propagation cycles.
@@
 An overview about DT-MRI visualization techniques is given by \cite{vilanova}.
-Tricoche et al.
-\cite{xavier} use invariant manifolds to extract topological features from tensor fields.
+Tricoche et al.~\cite{xavier} use invariant manifolds to extract topological features from tensor fields.
 Other feature extraction-based approaches extract general features as: edges \cite{hancock}, tensor shape\/ orientation \cite{gordon}, interfaces \cite{donnel} and creases \cite{tricochet}.
@@
 \section{Asymmetric Tensor Field Visualization}
-Zheng and Pang \cite{pang&zheng} proposed the concept of dual eigenvectors and Zhang et al.
-\cite{zhang} extended it to pseudo-eigenvectors and introduced the eigenvector/eigenvalue manifolds to visualize eigenvectors in the complex domain.
-Laramee et al.
-\cite{laramee} focused on the efficient implementation and visualization of these structures and provided an interactive visualization system for asymmetric tensors applicable in fluid and solid dynamics.
+Zheng and Pang \cite{pang&zheng} proposed the concept of dual eigenvectors and Zhang et al.~\cite{zhang} extended it to pseudo-eigenvectors and introduced the eigenvector/eigenvalue manifolds to visualize eigenvectors in the complex domain.
+Laramee et al.~\cite{laramee} focused on the efficient implementation and visualization of these structures and provided an interactive visualization system for asymmetric tensors applicable in fluid and solid dynamics.
 The concept of tensor magnitude has been introduced by \cite{laramee} for means of physical interpretation.
-They also proposed an efficient glyph and hyperstreamline hybrid approach, which made dynamic interaction in real-time in $2D$ tensor fields feasible.
-Palacios et al.
-\cite{palacios} extracted isosurfaces of tensor magnitude, mode and isotropy index.
+They also proposed an efficient glyph and hyperstreamline hybrid approach, which made dynamic interaction in real-time in 2D tensor fields feasible.
+Palacios et al.~\cite{palacios} extracted isosurfaces of tensor magnitude, mode and isotropy index.
 
@@
 LCS have been defined as time-dependent analog of separatrices \cite{haller} which are concerned to be robust under noisy conditions \cite{haller2}, which we expect from our technique as well by implication.
-Hlawatsch et al.
-\cite{hlawatsch} obtain similar results (a FTLE-like field on tensor fields), but obtain them by sampling of fiber trajectories instead of global illumination light distributions, as it is the case for this work.
+Hlawatsch et al.~\cite{hlawatsch} obtain similar results (a FTLE-like field on tensor fields), but obtain them by sampling of fiber trajectories instead of global illumination light distributions, as it is the case for this work.
 Our work shares the same basic idea, deriving and computing a FTLE-like field for tensor field visualization generalizing FTLE from vectors to tensors, yet using a light propagation scheme and hence using whole distributions of trajectories incorporating polar profiles.
@@
 We will detect LCS with an FTLE-derived method considering differences in final light transport distributions through computing the gradient of the resulting flow maps.
-The FTLE-related approach named ``light transport gradient'', to be designed, also states a generalization of the light transport visualization method FTPD proposed by Zirr et al.
-\cite{zirr}, since it allows a similar measure to be computed through defining a geometrical scene and setting the transmission (defined by the tensor field) to an isotropic and constant value of $100\%$.
+The FTLE-related approach named ``light transport gradient'', to be designed, also states a generalization of the light transport visualization method FTPD proposed by Zirr et al.~\cite{zirr}, since it allows a similar measure to be computed through defining a geometrical scene and setting the transmission (defined by the tensor field) to an isotropic and constant value of $100\%$.
 In this operating mode, the approach operates as a light transport visualization technique neglecting any tensor field transmission profiles and is thus capable of displaying bounding separatrices of a light transport scenario perturbated by obstacles and mirrors \cite{zirr}.
@@
 Tensors follow certain transformation rules which are defined for covariant and contravariant tensors, which indicate the change characteristics under a basis change.
-Contravariant tensors do not change under a change of basis, whereas covariant tensors are not invariant w.r.t.
-basis changes.
+Contravariant tensors do not change under a change of basis, whereas covariant tensors are not invariant w.r.t.~basis changes.
 Tensors are most commonly obtained and defined from mathematical transformation equations.
@@
 The Finite-Time Lyapunov Exponent is a measure, developed to analyze time-dependent dynamical systems for their separation abilities concerning massless tracer particles.
-We imagine a flow field, i.e., a vector field which in turn represents the effective amount or concentration shifted into a particular direction at each point in space.
+We imagine a flow field, i.e.~, a vector field which in turn represents the effective amount or concentration shifted into a particular direction at each point in space.
 It could be induced by e.g.
@@
 
-In Fig. \ref{ftle_sep}, we observe particle trajectories diverging w.r.t. the central axis ending up in distance $\Delta$. The FTLE measure for this single example (point in space) is then defined as:
+In Fig.~\ref{ftle_sep}, we observe particle trajectories diverging w.r.t.~the central axis ending up in distance $\Delta$. The FTLE measure for this single example (point in space) is then defined as:
 
@@
 whereas $\| A\| _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$.
-This will yield a map, which responds most where path lines diverge w.r.t.
-to a central axis.
+This will yield a map, which responds most where path lines diverge w.r.t.~to a central axis.
 We consider an example which detects FTLE ridges inside of the cross-section of a double gyre.
@@
 An overview about existing visualization techniques for uncertain data is given by Griete \& Schuhmann \cite{griethe}.
-We require that the trajectories can cross each other unaffectedly, i.e., the superposition principle is supposed to be respected for the magnitudes.
+We require that the trajectories can cross each other unaffectedly, i.e.~, the superposition principle is supposed to be respected for the magnitudes.
 But behold, is this behavior not somehow familiar?
@@
 To put it into context, our main contribution method LTG uses the FTLE as an inspiration or template to design a similar approach for tensor field visualization.
-The main difference to the template FTLE is that we follow eigenvectors instead of flow vectors and that we use whole distributions (i.e., light distributions specifically) of trajectories, as a tensor field suggests whole distributions of directions, accordingly.
+The main difference to the template FTLE is that we follow eigenvectors instead of flow vectors and that we use whole distributions (i.e.~, light distributions specifically) of trajectories, as a tensor field suggests whole distributions of directions, accordingly.
 This interpretation will be the connecting key to incorporate tensor fields into our light propagation scheme.
@@
     \includegraphics[height=\textwidth]{img/corner-TFL.png}
-	a)``corner''
+	a)``Corner''
     \label{a)}
@@
     \includegraphics[height=\textwidth]{img/spiral-TFL.png}
-	b)``spiral''
+	b)``Spiral''
     \label{b)}
@@
   \end{minipage}
-\caption{Several test fields: a)``corner'', ``spiral'', ``V''}
+\caption{Several test fields: a)``Drain'', ``Rings'', ``Inverse''}
 \label{test_fields2}
@@
 
-The ``drain''-Testfield, depicted in Fig.~\ref{test_fields2}a) is generated to have a kind of proof-of-concept for the FTLE-related approach.
+The ``Drain''-Testfield, depicted in Fig.~\ref{test_fields2}a) is generated to have a kind of proof-of-concept for the FTLE-related approach.
 It poses raw, bare diverging tensor field lines without any other feature modulated on top.
@@
 The ``rings'' test field, depicted in Fig.~\ref{test_fields2} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
-The ``spiral'' test field, depicted in Fig.~\ref{test_fields2} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
-With the \enquote{inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.
+The ``Spiral'' test field, depicted in Fig.~\ref{test_fields2} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
+With the \enquote{Inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.
 
@@
   \end{minipage}
-\caption{``gyre'' test field: a) glyphs, b) tensor field lines for a)}
+\caption{``Gyre'' test field: a) glyphs, b) tensor field lines for a)}
 \label{gyre}
 \end{figure}
-The test field ``gyre'', depicted in Fig \ref{gyre}, constitutes a classical textbook example for the FTLE and is considered for verifcation purposes.
+The test field ``Gyre'', depicted in Fig \ref{gyre}, constitutes a classical textbook example for the FTLE and is considered for verifcation purposes.
 
@@
 \end{figure}
-Observing the \enquote{brain} dataset in Figs.
+Observing the \enquote{Brain} dataset in Figs.
 \ref{real1}, one may notice that it introduces a lot of variances to tensor magnitude and linear anisotropy.
@@
 Then we will have a short introduction to the implemented global illumination scheme, derived and simplified from a global illumination approach, invented for the Crytek GmbH in the context of global indirect illumination rendering, in Sect.~\ref{sec:scheme}.
-Last, we will state the effectively implemented tensor field visualization methods: light transport visualization (directed propagation scheme yielding $2D$ scalar fields) in sects.
-\ref{sec:scheme} \& \ref{sec:trans} and light transport gradient (LTG: yielding $3D$ scalar fields) in Sect.~\ref{sec:ltg}, which are proposed as the invented tensor visualization techniques and hence as conceptual contributions to the domain of tensor field visualization.
+Last, we will state the effectively implemented tensor field visualization methods: light transport visualization (directed propagation scheme yielding 2D scalar fields) in sects.
+\ref{sec:scheme} \& \ref{sec:trans} and light transport gradient (LTG: yielding 3D scalar fields) in Sect.~\ref{sec:ltg}, which are proposed as the invented tensor visualization techniques and hence as conceptual contributions to the domain of tensor field visualization.
 
@@
 \label{sec:spec} 
-Our ambition is to motivate a simple and efficient global illumination propagation scheme, able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t.
-energy conservation and propagation attenuation principles.
+Our ambition is to motivate a simple and efficient global illumination propagation scheme, able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t.~energy conservation and propagation attenuation principles.
 We then generate transmission profiles, which we sample in a preliminary step, from the eigensystems (principal component ellipsoids) of the tensors for each cell in the domain (field).
@@
 They are circular axis versions of the domain $[0..2\pi]$, with negative magnitudes point reflected by 180 degrees.
-Thus, they are functions with an angular domain making them directionally dependent, which is a well suited representation for point lights (the type of light source with theoretically no finite extend), as we use them in a style of Dachsbacher et al.
-\cite{dachsbacher}.
+Thus, they are functions with an angular domain making them directionally dependent, which is a well suited representation for point lights (the type of light source with theoretically no finite extend), as we use them in a style of Dachsbacher et al.~\cite{dachsbacher}.
 For the sake of interpretability, we exclude negative values (energies) from the scope of our calculations which we express through mathematical clip functions, limiting negative values to $0.0$, which we denote by subscript $+$ in the following.
@@
 The face neighbors take in the angular domain of $d_1=\alpha+2\beta$ (yellow cone), whereas the diagonal neighbors (green cones) span a domain of $d_2=2\beta$.
-The linear combination weights for the overlapping (green) parts ($\beta$) are obtained from the relative overlapping angular area of the angular span $\beta$ w.r.t.
-diagonal cones ($2\beta$) and yellow face cones ($\alpha+2\beta = 90^\circ$) (yellow cone is overlapped by green cones at full angle).
+The linear combination weights for the overlapping (green) parts ($\beta$) are obtained from the relative overlapping angular area of the angular span $\beta$ w.r.t.~diagonal cones ($2\beta$) and yellow face cones ($\alpha+2\beta = 90^\circ$) (yellow cone is overlapped by green cones at full angle).
 The linear combination weights are subsequently normalized to $\sum = 1.0$ (normalization condition):
@@
 Remark, that we can precompute these linear combination weights and reutilize them in each iteration.
-We define the energetic contributions (presuming the cone is given in local coordinates - i.e.
-centered around the origin) as follows:
+We define the energetic contributions (presuming the cone is given in local coordinates - i.e.~centered around the origin) as follows:
 \begin{align}
@@
 The contribution areas are depicted in Fig.~\ref{scheme}, the cosine lobes can be observed in Fig.~\ref{iter}, respectively.
-This manner of propagation is inspired by the propagation scheme of Dachsbacher et al.
-\cite{dachsbacher} (see the paper on how to ``project the flux into a point light'').
+This manner of propagation is inspired by the propagation scheme of Dachsbacher et al.~\cite{dachsbacher} (see the paper on how to ``project the flux into a point light'').
 Basically, we sum up the weighted intensities corresponding to the respective neighbor index and scale an oriented cosine lobe (in direction of the neighbor) with the total flux $\Phi_t$ accumulated inside the neighbor range to transmit all of the parsed intensity for each neighbor distributed into a cosine distribution.
@@
 The implementation is done in a dual buffer approach which pushes the energies back and forth (buffer A to buffer B) for each propagation cyvle until a state of convergence is reached, which is the case when the energy is spreaded throughout the grid and enters a stationary state of equilibrium, characterized by equal out (at grid borders) to in (at light source positions) energy-flow with no more nettings going on inside the domain.
-We measure the convergence by setting up an overall distribution error w.r.t.
-the previous iteration.
+We measure the convergence by setting up an overall distribution error w.r.t.~the previous iteration.
 Note, that even though we propagate the energy directly through the diagonal cones (which forms a square initially), the 8 degrees of freedom introduce a circular (spherical) wave front after few iterations (for greater resolutions), as depicted in Fig.~\ref{iter}.
@@
 whereas: $i$: cell index, $I$: source buffer radiant intensity.\vskip 5pt
-The dual buffer result $\mathcal{C'}$ denotes the set of all transformed cells after the propagation, i.e., after applying the operator $\mathop{propagate}$ to each cell.
+The dual buffer result $\mathcal{C'}$ denotes the set of all transformed cells after the propagation, i.e.~, after applying the operator $\mathop{propagate}$ to each cell.
 This result is generated once in each iteration and conceptually gets pushed back and forth between the two buffers.
@@
 That is, the singular values form the half-axes of the PCA ellipsoid, which yields us a symbolic definition for our transmission (transfer) functions as weighting profiles which we can now evaluate as a parsing and accumulating step.
-We compute the offset angle $\varphi$ w.r.t.
-the x-axis by exploiting the $\mathop{atan2}$-function with 4-quadrant evaluation for the first singular vector (ordered decreasingly in value), since we scale the $x$-axis with the first singular value.
+We compute the offset angle $\varphi$ w.r.t.~the x-axis by exploiting the $\mathop{atan2}$-function with 4-quadrant evaluation for the first singular vector (ordered decreasingly in value), since we scale the $x$-axis with the first singular value.
 As a preliminary preparation, this profile is precomputed (sampled) for all discrete angular steps for each cell (e.g., $360$) and stored for reuitilization, as it is the case for the previously mentioned linear combination weights, as well.
@@
 As a side note, the physical quantity perceived as brightness by human eyes, is the radiance (luminance), which is yielded by relating radiant intensity to a finite sensitive area ($[\frac{W}{\mathop{rad}}/m^2]$).
-The primary definition for our total accumulated incident radiant flux (again presuming that the cone is given in local coordinates - i.e.
-centered around the origin) is then denoted as:
+The primary definition for our total accumulated incident radiant flux (again presuming that the cone is given in local coordinates - i.e.~centered around the origin) is then denoted as:
 
@@
 This emitter characteristic is the case for global illumination diffuse indirect illumination (non-specular).
-It is used in similar style by Dachsbacher et al.
-for propagating intensities through LPVs (light propagation volumes).
+It is used in similar style by Dachsbacher et al.~for propagating intensities through LPVs (light propagation volumes).
 But caution, we should not generate any exceeding energy right?
@@
 Consequently, this allows us to assume that all sources of power (energy) are user-defined.
-In fact the approach, has proven to be robust w.r.t.
-energy losses for several iterations as depicted in Chap.~\ref{chap:eval} in the stochastic error graph.
+In fact the approach, has proven to be robust w.r.t.~energy losses for several iterations as depicted in Chap.~\ref{chap:eval} in the stochastic error graph.
 
@@
 To evaluate the plausibility of the light distribution results, obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates.
-As an alternative representation, we plot the energy distribution summed over all directions for each cell, which yields a scalar field in Fig.
-\ref{isotropic} b).
+As an alternative representation, we plot the energy distribution summed over all directions for each cell, which yields a scalar field in Fig.~\ref{isotropic} b).
 \begin{figure}[!t]
@@
   \end{minipage}
-\caption{final light propagation distributions for ``isotropy''-testfield}
+\caption{final light propagation distributions for ``Isotropy''-testfield}
 \label{isotropic}
@@
 Remark the propagation front indicating a circular wave (circular isolines in a)).
-In this next example in Fig.
-\ref{rings-tests}, we use the tensor field ``rings'' as input data for the approach and observe how the intensities are channeled in a circular orbit around the central origin.
+In this next example in Fig.~\ref{rings-tests}, we use the tensor field ``Rings'' as input data for the approach and observe how the intensities are channeled in a circular orbit around the central origin.
 The anisotropy measures decide whether the transmission profile is more narrow (radical) or more wide (liberal):
@@
   \end{minipage}
-\caption{final light propagation distributions for ``rings''-test field}
+\caption{final light propagation distributions for ``Rings''-test field}
 \label{rings-tests}
@@
 Note the symmetry for both light source directions.
-Again, we plot the final light distributions as heat maps: It is noticeable in Fig.
-\ref{rings-tests}, that the intensities concentrate preferably on the circular tensor field lines, which proves the approach's controlling ability expressed through redirection.
-The results for our tensor field ``spiral'' can be interpreted in analogous fashion, similar to the previously discussed ``rings''-testfield in Fig.
-\ref{spiral-tests}.
+Again, we plot the final light distributions as heat maps: It is noticeable in Fig.~\ref{rings-tests}, that the intensities concentrate preferably on the circular tensor field lines, which proves the approach's controlling ability expressed through redirection.
+The results for our tensor field ``Spiral'' can be interpreted in analogous fashion, similar to the previously discussed ``Rings''-testfield in Fig.~\ref{spiral-tests}.
 The influence incorporated by the tensor field is easily noticeable in this case as well.
@@
   \end{minipage}
-\caption{final light propagation distributions for ``spiral''-test field}
+\caption{final light propagation distributions for ``Spiral''-test field}
 \label{spiral-tests}
 \end{figure}
-In Fig.
-\ref{spiral-tests}a),b) we observe the spiral shape literally absorbing the intensities into its spiral arms, which becomes even more clear, when we again observe the heat map representation in Fig.
-\ref{spiral-tests}c),d).
-
-The star example reveals another interesting feature of the approach in Fig.
-\ref{star-tests1}.
+In Fig.~\ref{spiral-tests}a),b) we observe the spiral shape literally absorbing the intensities into its spiral arms, which becomes even more clear, when we again observe the heat map representation in Fig.~\ref{spiral-tests}c),d).
+
+The star example reveals another interesting feature of the approach in Fig.~\ref{star-tests1}.
 When the intensity impulsively hits an isotropic region, it gets directed according to its intensity profile (neglecting the transmission profile), which implements the principle of inertia for directed light propagation present in vacuum (light never changes its direction).
@@
   \end{minipage}
-\caption{final light propagation distributions for ``star'' test field}
+\caption{final light propagation distributions for ``Star'' test field}
 \label{star-tests1}
 \end{figure}
-Considering the star test field in Fig.
-\ref{star-tests1}, we remark, that the diagonal intensities are attenuated more strongly by default, which introduces a bias into the light paths which are chosen along the axes with higher propability, effectively.
-This gets even clearer, when we plot the heat map depiction in Fig.
-\ref{star-tests2} again.
+Considering the star test field in Fig.~\ref{star-tests1}, we remark, that the diagonal intensities are attenuated more strongly by default, which introduces a bias into the light paths which are chosen along the axes with higher propability, effectively.
+This gets even clearer, when we plot the heat map depiction in Fig.~\ref{star-tests2} again.
 \begin{figure}[!t]
@@
   \end{minipage}
-\caption{final light propagation distributions for ``star'' test field}
+\caption{final light propagation distributions for ``Star'' test field}
 \label{star-tests2}
@@
 The approach called Light Transport Gradient (LTG) states a FTLE-related approach performed on tensor fields, but employing whole distributions of trajectories interpreted as light transmission profiles.
-Since we incorporate whole distributions of trajectories, we do not strictly follow the major eigenvectors like Hlawatsch's et al.
-\cite{hlawatsch} do with their FTLE on single trajectories following tensor field lines.
+Since we incorporate whole distributions of trajectories, we do not strictly follow the major eigenvectors like Hlawatsch's et al.~\cite{hlawatsch} do with their FTLE on single trajectories following tensor field lines.
 Instead, we follow the directional distribution defined by the tensor field's principal ellipsoids for different input angles, which yields a FTLE volume, congruently depending on the direction, in general.
@@
 For this purpose, we imagine kind of a local and spatial gradient, as we place light sources in two close-by positions and eventually also in two close-by directions and analyze the resulting final light distribution.
-We consider the light distributions and their difference image for the \enquote{isotropic} and the \enquote{drain} test field to have an idea, for which kind or types of cases, i.e.
-initial configurations we obtain inherently different light distributions, in particular.
-We choose to use these test cases, because they constitute absolutely pure examples par excellence for a FTLE computation exposing no (\enquote{isotropic}), or on the other hand raw (\enquote{drain}) tensor field lines diverging.
-We will start off with the \enquote{isotropic} field from here to first analyze the offset response for no divergence or more specifically, no anisotropy at all.
+We consider the light distributions and their difference image for the \enquote{Isotropic} and the \enquote{Drain} test field to have an idea, for which kind or types of cases, i.e.~initial configurations we obtain inherently different light distributions, in particular.
+We choose to use these test cases, because they constitute absolutely pure examples par excellence for a FTLE computation exposing no (\enquote{Isotropic}), or on the other hand raw (\enquote{Drain}) tensor field lines diverging.
+We will start off with the \enquote{Isotropic} field from here to first analyze the offset response for no divergence or more specifically, no anisotropy at all.
 Fig.~\ref{analysis1} depicts final light distributions (cf.~Sect.~\ref{sec:ltg} Fig.~\ref{rings-tests}) for placing the light source one step to the left and one to the right from the central axis (spatial offset).
@@
 That is, per default, light distributions cancel each other out for two close-by light source positions or directions (or more specifically: for continous overlaps in their distributions).
-In this case, for the \enquote{isotropic} test field, we have no tensor field lines diverging leading to a small overlap and thus to a decreased overall difference for light sources in finite distance.
+In this case, for the \enquote{Isotropic} test field, we have no tensor field lines diverging leading to a small overlap and thus to a decreased overall difference for light sources in finite distance.
 
@@
   \end{minipage}
-  \caption{``isotropic'' test field}
+  \caption{``Isotropic'' test field}
 \label{analysis1}
@@
   \end{minipage}
-  \caption{``isotropic'' test field}
+  \caption{``Isotropic'' test field}
 \label{analysis2}
@@
 We again analyze the drain test field (cf.~Fig.~\ref{test_fields2}a) in Sect.~\ref{chap:grundlagen}) for the behavior of the resulting light distributions in scalar field representation.
-This following example illustrated in Fig.~\ref{evolution1} now introduces raw tensor field lines diverging w.r.t.
-the $y$-axis.
+This following example illustrated in Fig.~\ref{evolution1} now introduces raw tensor field lines diverging w.r.t.~the $y$-axis.
 
@@
   \end{minipage}
-  \caption{``drain'' test field}
+  \caption{``Drain'' test field}
 \label{evolution1}
@@
   \end{minipage}
-  \caption{``drain'' test field: }
+  \caption{``Drain'' test field: }
 \label{evolution2}
@@
  As tensor field lines are bidirectional, convergence and divergence can not be reasonably discerned in that case.
-We aim to extract ridges from a FTLE-like field to grasp the key regions separating anisotropy behavior (i.e., changes in direction/tensor magnitude) segmented as height ridges in the field, which are in turn LCS (Lagrangian Coherent Structures), well known for their rich substance in physical meaning.
+We aim to extract ridges from a FTLE-like field to grasp the key regions separating anisotropy behavior (i.e.~, changes in direction/tensor magnitude) segmented as height ridges in the field, which are in turn LCS (Lagrangian Coherent Structures), well known for their rich substance in physical meaning.
 They often indicate transport barriers in flow fields.
@@
 They can even be observed in nature for, e.g., vortices in flow fields.
-For this purpose, we will use our, previously defined, $\mathop{propagate}$-operator, i.e., the propagation scheme from Sect.~\ref{sec:scheme} and \ref{sec:trans}, on every possible light source location $\mathbf{x}$ and direction $\omega$ in the grid to obtain the final set of buffers $\mathcal{B} = \{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_j\}$ of impulse responses of the algorithm.
+For this purpose, we will use our, previously defined, $\mathop{propagate}$-operator, i.e.~, the propagation scheme from Sect.~\ref{sec:scheme} and \ref{sec:trans}, on every possible light source location $\mathbf{x}$ and direction $\omega$ in the grid to obtain the final set of buffers $\mathcal{B} = \{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_j\}$ of impulse responses of the algorithm.
 That is, we initialize the grid with every possible light source configuration and compute a whole final light distribution for each step:
@@
 Eq.
-\ref{ftle} now denotes the set of buffers $\mathcal{B}$ containing light distributions propagated from, i.e., computed for every possible direction and location.
+\ref{ftle} now denotes the set of buffers $\mathcal{B}$ containing light distributions propagated from, i.e.~, computed for every possible direction and location.
 Now, that we have measured impulse responses from every location in our long $1D$ sample vector, we can incorporate finite (central) differences to obtain the gradient of the set of final buffers with final light distributions as elements $\mathcal{C}_{final}=\mathbf{b}\in\mathcal{B}$.
@@
 \end{align}
-Note that the resulting scalar field is $(n+1)$-dimensional ($x,y$ and direction $\omega\mapsto  z$) and needs to be flattened by averaging or projection in 3D for proper visualization in case the approach will be adopted in $3D$.
-We note, that our invented technique stands out compared to the already existent FTLE on tensor fields invented by Hlawatsch et al.
-\cite{hlawatsch}, as we use whole distributions of directions instead of following the major eigenvectors strictly.
+Note that the resulting scalar field is $(n+1)$-dimensional ($x,y$ and direction $\omega\mapsto  z$) and needs to be flattened by averaging or projection in 3D for proper visualization in case the approach will be adopted in 3D.
+We note, that our invented technique stands out compared to the already existent FTLE on tensor fields invented by Hlawatsch et al.~\cite{hlawatsch}, as we use whole distributions of directions instead of following the major eigenvectors strictly.
 These directional distributions allow for a complete analysis of the tensor field accounting for the incoming direction, as well.
@@
 \subsection{Functionality}
-Remark, that the $z$-axis represents the $\omega$-axis for each test and example and that each data range is rescaled to its interval for proper color coding in heat maps of the resulting $3D$ volume.
+Remark, that the $z$-axis represents the $\omega$-axis for each test and example and that each data range is rescaled to its interval for proper color coding in heat maps of the resulting 3D volume.
 We will try to show the functionality of the method in a few simple examples.
@@
 This corresponds to the \enquote{non - tensor field} mode of the propagation scheme having no influence of any tensor field incorporated.
-In this configuration, we could eventually visualize light transport paths in a $2D$ topological scene.
+In this configuration, we could eventually visualize light transport paths in a 2D topological scene.
 
@@
   \end{minipage}
-  \caption{\enquote{isotropic} test field: a) glyphs, b) LTG($0^\circ$) for a)}
+  \caption{\enquote{Isotropic} test field: a) glyphs, b) LTG($0^\circ$) for a)}
 \label{ftle_iso}
@@
 That is, the slices we observe are stacked along the $\omega\mapsto z$-axis in the following.
-As a subsequent step, we segment height ridges in $3D$ as thresholded volumes from the whole volume grid.
+As a subsequent step, we segment height ridges in 3D as thresholded volumes from the whole volume grid.
 
@@
   \begin{minipage}{0.4\textwidth}
+   \centering
     \includegraphics[width=0.8\textwidth]{img/iso_isovolume.png}
@@
   \end{minipage}
-  \caption{\enquote{isotropic} test field: a) volume, b) IsoVolume for a)}
+  \caption{\enquote{Isotropic} test field: a) volume, b) IsoVolume for a)}
 \label{ftle_iso}
 \end{figure}
-To verify the concept of absorption implicitly, we incorporate absorption into test field \enquote{absorb}, whereas all cells are absorbing energy except the vertical center line, which is assumed to produce a clear ridge.
+To verify the concept of absorption implicitly, we incorporate absorption into test field \enquote{Absorb}, whereas all cells are absorbing energy except the vertical center line, which is assumed to produce a clear ridge.
 \begin{figure}[t]
@@
   \begin{minipage}{0.4\textwidth}
+   \centering
     \includegraphics[width=0.9\textwidth]{img/absorb_ftle.png}
@@
   \end{minipage}
-  \caption{\enquote{absorb} test field: a) glyphs, b) LTG($90^\circ$) for a)}
+  \caption{\enquote{Absorb} test field: a) glyphs, b) LTG($90^\circ$) for a)}
 \label{ftle_absorb}
@@
 Note, that we (min-max) normalize each transmission profile with the highest mean transmission rate (tensor magnitude) in the field.
-From this test case, we can conclude, that regions with a comparably small tensor magnitude would not obtain any FTLE field response, which we will later observe for the real datasets \enquote{brain} and \enquote{heart} exhibiting such areas.
+From this test case, we can conclude, that regions with a comparably small tensor magnitude would not obtain any FTLE field response, which we will later observe for the real datasets \enquote{Brain} and \enquote{Heart} exhibiting such areas.
 We will now discuss synthetic examples designed to evaluate the performance of the FTLE.
-As a little warmup, we will have an experiment exhibting bare, raw tensor field lines diverging w.r.t.
-the $y$-axis:
+As a little warmup, we will have an experiment exhibting bare, raw tensor field lines diverging w.r.t.~the $y$-axis:
 
@@
   \end{minipage}
-  \caption{``drain'' test field}
+  \caption{``Drain'' test field}
 \label{ftle_base}
 \end{figure}
-What we can observe here, is a feature called a ``ridge'' occuring as a high value line segment (red vertical track) which is already quite obvious for human vision.
+What we can observe here, is a feature called a ridge occuring as a high value line segment (red vertical track) which is already quite obvious for human vision.
 A gradient filter could be applied to extract and segment edges from the image as a preprocessing step for computer vision, which would look like the image in Fig.~\ref{ftle_base} c).
@@
 It could for example represent a stress field inside of a mechanical tool with a central axis with maximum stress condensation.
-\begin{figure}[ht]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
   \includegraphics[width=0.9\textwidth]{img/drain_top.PNG}\vskip 10pt
@@
   \begin{minipage}{0.4\textwidth}
-  \centering
     \includegraphics[width=0.9\textwidth]{img/drain_alt_isovolume.PNG}
@@
   \end{minipage}
-  \caption{LTG volumes: \enquote{drain} test field a) cross-section, b) IsoVolume}
+  \caption{LTG volumes: \enquote{Drain} test field a) cross-section, b) IsoVolume}
 \label{drain_contour}
 \end{figure}
-For the ``drain'' test field we consider the central channel which produces a clear line in the cross section.
+For the ``Drain'' test field we consider the central channel which produces a clear line in the cross section.
 The light transport gradient responds most inside of the channel region and on the top T-arms independently of the stimulus direction.
@@
   \end{minipage}
-  \caption{``diagonal drain'' test field: a) TFLs and glyphs, b) LTG($225^\circ$) for a)}
+  \caption{``Diagonal Drain'' test field: a) TFLs and glyphs, b) LTG($225^\circ$) for a)}
 \label{diag-ftle}
 \end{figure}
-Regarding the ridge for the \enquote{diagonal drain} test field, we note that it is a little smoothed compared to the axis-aligned ridges, as they obtain a larger propagation attenuation (because of the larger distance) leading to a stronger separation of trajectories in general and hence to a smoothed FTLE response.
-Also we note that the divergence of the tensor field lines is reduced compared to the classic \enquote{drain} example of the FTLE.
+Regarding the ridge for the \enquote{Diagonal Drain} test field, we note that it is a little smoothed compared to the axis-aligned ridges, as they obtain a larger propagation attenuation (because of the larger distance) leading to a stronger separation of trajectories in general and hence to a smoothed FTLE response.
+Also we note that the divergence of the tensor field lines is reduced compared to the classic \enquote{Drain} example of the FTLE.
 
@@
 
-We also evaluated the approach for the ``inverse'' test field, depicted in Fig.~\ref{inverse-ftle}, whereas we can observe height ridges for the $90^\circ$ slice in any position with diverging tensor field lines w.r.t.
-the light source direction.
+We also evaluated the approach for the ``Inverse'' test field, depicted in Fig.~\ref{inverse-ftle}, whereas we can observe height ridges for the $90^\circ$ slice in any position with diverging tensor field lines w.r.t.~the light source direction.
 This is certainly not the case for the upper (missing) ridge whereas tensor field lines converge with respect to the UP ($90^\circ$) direction.
@@
   \end{minipage}
-    \caption{LTG volumes: ``inverse'' test field: a) IsoVolume, b) Surface, c) Contour}
+    \caption{LTG volumes: ``Inverse'' test field: a) IsoVolume, b) Surface, c) Contour}
 \label{inverse_volume}
@@
 Note that the $z$-axis represents the light source direction angle $\omega$ in this case.
-Observing the volumes in Fig.
-\ref{inverse_volume}, the contour filter reveals the star-shaped structure of the tensor field which rotates along the angle axis.
+Observing the volumes in Fig.~\ref{inverse_volume}, the contour filter reveals the rotor-shaped structure of the tensor field which rotates along the angle axis.
 This is the effect we already captured for the single slice: For each axis direction we can detect ridges from parallel up to orthogonal to our predefined stimulus (light source) direction, but not in antiparallel direction.
 This causes the star shape to miss one branch or ray which is in turn the antiparallel direction missing for each light source direction, emphasizing the impression of rotation.
-We can also try to segment height ridges represented as surfaces from the resulting volume grids as depicted in Fig \ref{inverse_contour}.
+We can also try to segment height ridges represented as surfaces from the resulting volume grids as depicted in Fig \ref{inverse_volume}.
 The orthogonal ridges in this example reach out to the edges because they are gaining a bias, through never offsetting the light source in a transmission direction (ellipsoid orientation) for orthogonals.
 This turns out to be a general feature of the method, as we will later verify in Figs.
-\ref{res_ftle_volume} and \ref{star-ftle}.
+\ref{res_ftle_volume} and \ref{star-ftle} in Chap.
+\ref{chap:eval}.
 
@@
 
-For the ``inverse'' test field, we can observe a 3 part ``rotor'' indicating the ridges which are in parallel up to orthogonal to the current light source (slice) direction.
-This rotor can be obtained by applying a contour filter, but it can also be visualized by applying a simple thresholding on the surface representation of the volume, as depicted in Fig.~\ref{inverse_volume}.
-Subsequently, we will analyze the \enquote{rings} test field to show its performance on circular ridges.
+Subsequently, we will analyze the \enquote{Rings} test field to show its performance on circular ridges.
 
@@
   \end{minipage}
-    \caption{LTG volumes: ``rings'' test field: a) Contour, b) VCGRidge, c) Volume}
+    \caption{LTG volumes: ``Rings'' test field: a) Contour, b) VCGRidge, c) Volume}
 \label{rings_contour}
@@
   \end{minipage}
-  \caption{``bow''-test field: a) TFLs and glyphs, b) LTG($0^\circ$) for a)}
+  \caption{``Bow''-test field: a) TFLs and glyphs, b) LTG($0^\circ$) for a)}
 \label{bow-ftle}
@@
   \end{minipage}
-    \caption{LTG volumes: ``bow'' test field : a) Contour, b) IsoVolume}
+    \caption{LTG volumes: ``Bow'' test field : a) Contour, b) IsoVolume}
 \label{bow_contour}
@@
 
-If we relate the resulting FTLE volume to the work of Hlawatsch et al.
-\cite{hlawatsch}, we find that their FTLE field is contained in our volume in form of a manifold, since they use the same basic idea, but consider only a single direction (of the major eigenvector), instead of taking all directions into account, as we do here.
-The resulting manifold included, is constrained to a height field (not self-overlapping or self-intersecting), because each position can only represent one individual direction, as there is only one direction for the major eigenvector at each position to be covered (, since it is obviously a definite, distinct measure).
+If we relate the resulting FTLE volume to the work of Hlawatsch et al.~\cite{hlawatsch}, we find that their FTLE field is contained in our volume in form of a manifold, since they use the same basic idea, but consider only a single direction (of the major eigenvector), instead of taking all directions into account, as we do here.
+The resulting manifold included, is constrained to a height field (not self-overlapping or self-intersecting), because each position can only represent one individual direction, as there is only one direction for the major eigenvector at each position to be covered, since it is obviously a definite, distinct measure.
 This direction emerges as a single scalar height (FTLE) value on each position corresponding to a height (direction) slice in our volume.
-That is, each position in their $2D$ scalar field is assigned a height in our $3D$ scalar field.
-We cover their results as a subset of our volume, which could be, e.g., a spiral plane in case all directions are given in circular arrangement in the underlying tensor field (cf.~\enquote{rings} test field).
-If we would vary the angle of the major eigenvector with a uniform distribution along the tensor field's width, we would obtain a primitive diagonal plane w.r.t.
-the $xy$-plane reaching out in $z$-direction as contained subset.
-Therefore, we claim and consider our invented technique to be more generalized (to all directions) resulting in a topological $n+1$-D volume.
+That is, each position in their 2D scalar field is assigned a height in our 3D scalar field.
+We cover their results as a subset of our volume, which could be, e.g., a spiral plane in case all directions are given in circular arrangement in the underlying tensor field (cf.~\enquote{Rings} test field).
+If we would vary the angle of the major eigenvector with a uniform distribution along the tensor field's width, we would obtain a primitive diagonal plane w.r.t.~the $xy$-plane reaching out in $z$-direction as contained subset.
+Therefore, we claim and consider our invented technique to be more generalized (to all directions) resulting in a topological $(n+1)$-D volume.
 
@@
 The threshold for the convergence criterion was empirically determined and set in the range $[0.5,2.0]$.
-Since it is an energetic threshold, it is only dependent on the light source magnitude and not on any type of resolutions.
+Since it is an energetic threshold, it is only dependent on the light source magnitude and not on any type of resolution.
 An iteration limit is set to $\mathop{limit}=10\mathop{width}$ (e.g.: for $\mathop{width}=10\cdot 100=1000$) to prevent the approach from eventual oscillation behavior for intensities and hence overpropagation (propagating for too many timesteps).
-The spatial resolution is inferred by the tensor field reader method from the tensor field (see Propagation below).
-The light source profile can be defined for the \enquote{non-FTLE} mode where the user sets position and polar profile of single or multiple light sources as a symbolic term: e.g.
-\texttt{clip(cos(theta-pi/2))} for a Lambert-emitter pointing upwards.
+The spatial resolution is inferred by the tensor field reader from the tensor field.
+The light source profile can be defined for the \enquote{non-FTLE} mode where the user sets position and polar profile of single or multiple light sources as a symbolic term: e.g., \texttt{clip(cos(theta-pi/2))} for a Lambert-emitter pointing upwards.
 After the configuration read process is finished, we allocate memory for certain buffers holding glyph and intensity information with a dimension of $\mathop{dim}=\mathop{width}\mathop{height}\mathop{steps}$.
-This $1D$ buffer is interpreted like this:
+This 1D buffer is interpreted corresponding to Fig.~\ref{indices}.
 
@@
 If it falls short of this threshold, we initiate a stop sequence and leave the loop.
-To compute the FTLE, we use multiple stimuli in parallel (e.g.: Dirac pulses within the scope of this work), and just compute the differences in final light distributions for two close-by light source positions and directions to accumulate them in a total energy sum (capacity), which in turn forms one component of the light transport gradient.
+To compute the FTLE, we use multiple stimuli in parallel (e.g., Dirac pulses within the scope of this work), and just compute the differences in final light distributions for two close-by light source positions and directions to accumulate them in a total energy sum (capacity), which in turn forms one component of the light transport gradient.
 A Euclidean norm yields a scalar field from the gradient at each position.
@@
 Note that we set the light source intensity to constant value $\mathit{steps}$ to require a mean intensity of $\mu_I=1.0$ comparable to the one of a unit circle independently on the angular resolution.
-
 When we consider the LTG method, we already stated its runtime behavior ($\mathcal{O}(\mathop{width^5}\mathop{steps^2})$), which makes parallelization libraries and hardware to compute the results in reasonable time.
-First, we tried to implement a GPU-based parallelization for the propagation scheme with nVidia's Thrust library, which turned out not to work as expected, since the relative runtime gain was too low to make a big difference and we struggled to find a suitable implementation mapping the propagation process to atomic OPs (simple vector operations), which are the only operations allowed for GPU-parallelization.
-At this point, we choose to use OpenMP as we figured it is a much more suitable architecture employed here, which allows to perform the computation on a multiprocessor architecture provided by the Heidelberg University allowing to compute $64$ hardware threads at maximum.
+First, we tried to implement a GPU-based parallelization for the propagation scheme with nVidia's Thrust library, which turned out not to work as expected, since the relative runtime gain was too low to make a big difference and we struggled to find a suitable implementation mapping the propagation process to atomic OPs (simple vector operations), which are the only operations allowed for GPU-based parallelization.
+At this point, we choose to use OpenMP for CPU-based parallelization, as we figured it is a much more suitable architecture employed here, which allows to perform the computation on a multiprocessor architecture provided by the Heidelberg University allowing to compute $64$ hardware threads at maximum.
 As we need fat fed, instruction-rich threads to have a high performance gain, we choose to parallelize only the outer loop of the LTG method (propagating final light distributions for each light source direction).
@@
 The method performing the light propagation, was implemented in a separate class called propagator, which we can initialize once for each programming thread (, which we can in turn provide with its own buffer memory).
-As an initialization we fill the buffers with zeros to have a blank buffer initially to fill up with meaningful intensity profiles.
+As an initialization step, we fill the buffers with zeros to have a blank buffer initially, effectively to fill up with propagated intensity profiles.
 The transmission profiles are generated by loading the tensor field in memory, performing PCA on each of them and exploiting the thin $\mathbf{U}$ matrix for the first (greater) singular vector.
 The resulting principal component ellipsoid equation is precomputed, sampled and stored for reuse once for each cell.
-We again sample cosines into each direction, average out their integrals (sums) and normalize each of them with the mean to have the integral energy scaled to $1$ (identity element for multiplication, as it is the case for \enquote{scaling the cosine lobes} here).
+We again sample cosines into each direction, average out their integrals (sums) and normalize each of them with the mean to have the integral energy scaled to $1.0$ (identity element for multiplication, as it is the case for \enquote{scaling the cosine lobes} here).
 Thus we ensure, that the transmitted energy is approximately $1.0=100\%$ for each lobe and account for angular sampling errors.
-Now we need to precompute and store the linear combinations used for the \enquote{shared part} in diagonal cones.
-When hitting border cells, we propagate the intensities only to the present neighbor directions still located on the grid.
-Remark, that we can not just ignore the edges of the grid because it would lead to large trivial intensity differences on grid borders for the FTLE-similar approach LTG.
+Now, we need to precompute and store the linear combinations used for the \enquote{shared part} in diagonal cones.
+When hitting border cells, we propagate the intensities only to the present neighbor directions, still located on the grid.
+Remark, that we can not just ignore the edges of the grid, because it would lead to large trivial intensity differences on grid borders for the FTLE-similar approach LTG.
 
@@
 At first, we needed to generate convenient test cases, which we design using Mathematicas VectorPlot, used to visualize vector fields.
-To output a predefined test case, we need to define a little code, for which we used python scripts which generate VTK appended or ASCII format files.
-To implement the method we used C++'s STL for every dynamic memory (buffer), SFML to plot polar profiles, Eigen to perform matrix decompositions and VTK to in- and output VTK format files, which can in turn be visualized and explored interactively with the visualization framework ParaView.
-The user functions defined in the configuration file are read-in by a string parser (muParser Library), which is used just once for sampling the polar profile and storing the precomputed result.
-This causes a convenient input mechanism for stimuli, in case they should be predefined.
+To output a predefined test case, we used python scripts which generate VTK appended or ASCII format files.
+To implement the propagator method we used C++'s STL for every dynamic memory (buffer), SFML to plot polar profiles, Eigen to perform matrix decompositions and VTK to in- and output VTK format files, which can in turn be visualized and explored interactively with the visualization framework ParaView.
+The user functions are defined in the configuration file and are read-in by a string parser (muParser Library), which is used just once for sampling the polar profile and then storing the precomputed result in memory.
+This provides a convenient input mechanism for stimuli, in case they should be predefined.
 
@@
 \label{sec:att}
-It can be proven that the approach follows propagation attenuation principles and follows the inverse (square in $3D$) law ($\sim\frac{1}{r}$) as observed in the graph in Fig.~\ref{att}.
-For that, we plot the relative intensity $i_{rel}(r_{rel})$ (in relation to the intensity at position (radius) $1$ to account for the inverse square law, which holds for relative distances.
+It can be proven that, the approach follows propagation attenuation principles and follows the inverse (square in 3D) law ($\sim\frac{1}{r}$) found for circular waves, as observed in the graph in Fig.~\ref{att}.
+For that, we plot the relative intensity $i_{\mathop{rel}}(r_{\mathop{rel}})$ (in relation to the intensity at distance (radius) $1$ to account for the inverse square law, which holds for relative distances.
 We evaluate the relative intensity by forming a ratio $\frac{I(r)}{I_1}$, that relates the step intensity to the source intensity at distance $1$.
@@
 \end{figure}
+Considering the graph in Fig \ref{att}, we remark the reasonably fitted $\frac{1}{r}$ curve with measures we output into a text file.
+We can now state, that the approach implements the concept of propagation attenuation in 2D.
 
@@
 \label{sect:total}
-We also wanted to check for the limitations of the approach by having total linear anisotropy (the case whereas one eigenvalue $\lambda_1>0$ and the other $\lambda_2=0$) incorporated.
-For this we set up bidirectional Dirac-pulses as transmission profiles to have the whole energy transmitted in just one, discrete direction.
+We also wanted to check for the limitations of the approach by having total linear anisotropy incorporated.
+This is always the case, when one eigenvalue $\lambda_1>0$ and the other $\lambda_2=0$.
+For this we set up bidirectional Dirac pulses as transmission profiles to have the whole energy transmitted in just one, discrete direction.
 \begin{figure}[!t]
@@
 \label{sect:conserve}
-The energy conservation test was done in similar manner as the propagation attenuation test in Sect.~\ref{sec:att}.
+The energy conservation test was done in similar manner, as the propagation attenuation test in Sect.~\ref{sec:att}.
 We measure the impulse response of the algorithm for 80 iterations by summing the total energy in the grid after each iteration and comparing it to the initial energy sum (e.g.: $\int_0^{2\pi}1\mathop{d\omega}=2\pi$) in iteration 0 in absolute values.
@@
 But what is lost, is lost and can not be regained.
-The error is thus accumulating gradually and with faster slope for higher source intensities, since floting point precision, i.e., rounding errors operate more heavily on large numbers since it is a constant relative loss $\sim\frac{1}{r}$ leading to a greater absolute $\Delta E$.
+The error is thus accumulating gradually and with faster slope for higher source intensities, since floting point precision, i.e.~, rounding errors operate more heavily on large numbers since it is a constant relative loss $\sim\frac{1}{r}$ leading to a greater absolute energetic difference $\Delta E$.
 
@@
 The light transport gradient approach, which is closely related to the FTLE approach in vector fields, is evaluated for some more synthetic and real datasets in this section.
-We will start off with the comparison of our FTLE-based approach to the one invented by Hlawatsch et al.
-\cite{hlawatsch}, which we reimplemented for verification purposes.
+We will start off with the comparison of our FTLE-based approach to the one invented by Hlawatsch et al.~\cite{hlawatsch}, which we reimplemented for verification purposes.
 
@@
   \begin{minipage}[t]{0.3\textwidth}
+    \centering
   \includegraphics[height=\textwidth]{img/drain(alt)-TFL.png}
@@
   \end{minipage}
+    \centering
   \begin{minipage}[t]{0.3\textwidth}
@@
 
-We can observe the comparison between our invented FTLE-like approach and the one of Hlawatsch et al.
-in the depiction in Fig.~\ref{drain_ftle_comp}.
-Note, that a sharper but weaker ridge is detected on the right side of the image, which shines similar to a cone shape from the top.
+We can observe the comparison between our invented FTLE-like approach and the one of Hlawatsch et al.~in the depiction in Fig.~\ref{drain_ftle_comp}.
+Note, that a sharper but weaker ridge is detected in c), which shines similar to a cone shape from the top.
 We interpret the missing spot in the top center as an artifact stemming from the way we handle our light propagation, as we allow whole distributions of trajectories to be propagated (with a scalar attribute interpretable as propability).
@@
 The directive effect turns out to be not that strict and delayed compared to the classic tensor field line FTLE.
-The crucial difference is that the compared work of Hlawatsch et al.
-obtains the FTLE field from the gradient $\nabla\mathbf{\upvarphi}(\mathbf{x})$ of the flow map, which is in turn obtained from sampling of trajectories following tensor field lines.
-
+The crucial difference is, that the compared work of Hlawatsch et al.~obtains the FTLE field from the gradient $\nabla\mathbf{\upvarphi}(\mathbf{x})$ of the flow map, which is in turn obtained from sampling of trajectories following tensor field lines.
+We will now capture the star dataset revealing tensor field lines diverging in any direction from the center.
 \begin{figure}[!t]
@@
   \begin{minipage}{0.29\textwidth}
+    \centering
   \includegraphics[height=\textwidth]{img/star.png}
@@
   \begin{minipage}{0.29\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img/star_ftle.PNG}
@@
   \begin{minipage}{0.35\textwidth}
+    \centering
   \centering
@@
 \end{figure}
-When we look at the example depicted in Fig.~\ref{star-ftle}, we can see the FTLE approach LTG responding most in the center of the field, which is the nodal point of tensor field lines diverging in any direction.
+When we look at the example depicted in Fig.~\ref{star-ftle}, we can see the approach LTG responding most in the center of the field, which is the nodal point of tensor field lines diverging in any direction.
 Remark, that in the middle image there is a light V-shaped ridge on the left on blue background which is outshined by the center spot because of its orders of larger magnitude.
-The volume illustrated on the right even shows a spiked structure for the volume representing all directions at once.
-The approach detects more ridges in diagonal directions stochastically as we can infer from the Gaussian-shaped plates in the thresholded volume, evenly spaced in $90^\circ$ shifts with $45^\circ$ (quarter volume) offset.
+The volume illustrated in c) even shows a spiked structure for the volume representing all directions at once.
+The approach detects more ridges in diagonal directions stochastically, as we can infer from the Gaussian-shaped plates in the thresholded volume, evenly spaced in $90^\circ$ shifts with $45^\circ$ (quarter volume) offset.
 These can be interpreted nicely, since the diagonal direction is never offset into its transmission direction leading to a trivial (none) difference for axis-parallel directions for the offset direction orthogonal to the light source direction.
 Hence, the differences will be factor $2$ at worst for the diagonals, since we measure differences in both offset directions, in case ridges in both axis-parallel and -diagonal directions are present.
-Accordingly, we observe Gaussian-plates in the volume instead of sharp, contrasted ridges for the \enquote{star} example.
+Accordingly, we observe Gaussian-plates in the volume instead of sharp, contrasted ridges for the \enquote{Star} example.
 But wouldn't we expect a ridges only in parallel directions?
@@
 For better imagination, offset the light source vertically or horizontally from the diagonals and observe the light paths hitting the axes (, which in fact do cross each other).
+This is a very special example, in particular, which is commonly not well generalizable to other cases.
 
@@
 
- We will now capture the star dataset revealing tensor field lines diverging in any direction from the center.
-Now, we will evaluate a classic textbook example called the \enquote{double gyre} \cite{shawn}, which is a quadruple gyre in our case for formatting reasons, yet yielding the same conceptual results.
+Now, we will evaluate a classic textbook example called the \enquote{Double Gyre} \cite{shawn}, which is a quadruple gyre in our case for formatting reasons, yet yielding the same conceptual results.
 We compare the glyph representation to the LTG magnitude for a single slice ($0^\circ$) again:
@@
   \begin{minipage}[t]{0.3\textwidth}
+    \centering
   \includegraphics[height=\textwidth]{img/gyre.png}
@@
   \begin{minipage}[t]{0.3\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img/gyre_ftle.PNG}
@@
   \begin{minipage}[t]{0.3\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img/gyre_ftle_org.PNG}
@@
   \end{minipage}
-  \caption{\enquote{Gyre} test field: a) Glyphs, b) LTG($0^\circ$), c) FTLE: Hlawatsch's approach}
+  \caption{\enquote{Gyre} test field: a) glyphs, b) LTG($0^\circ$), c) FTLE: Hlawatsch's approach}
 \label{gyre-ftle}
 \end{figure}
-In Fig.~\ref{gyre-ftle}, we can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t.
-the $x$-axis ($0^\circ$).
-Note that even regions in antiparallel directions ridges are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging.
+In Fig.~\ref{gyre-ftle}, we can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t.~the $x$-axis ($0^\circ$).
+Note, that even ridges in antiparallel directions are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging.
 Again, it is noticeable, that the result of Hlawatsch correspond to our result to a certain degree.
 It is noticeable, that Hlawatsch's approach does not consider the initial direction and hence will detect ridges also in antiparallel directions (gradient of the flow map).
-This yields symmetric ridges w.r.t.
-central axes, whereas we detect ridges only up to orthogonally to the initial light source direction as visible in Fig.~\ref{gyre-ftle}b)\& c), respectively.
-In the center location of each gyre, we interpolate the sorrounding ring with an isotropic cell in the middle, leading to chaotic trajectories leaving the gyre and thus leading to an increased gradient of the flow map for Hlawatsch's result.
-This causes the FTLE field to respond in the center of the gyres in Fig.~\ref{gyre-ftle}c).
-We interpret this behavior within the scope of interpolated trajectories, which are in general more likely to escape the gyre, since their angle is interpolated (by PCA), which is susceptive to rapid changes and hence leads to unstable behavior.
-What Hlawatsch et al.
-detect, could be called local ridges (in terms rather short), what our method detects could rather be called global ridges (in terms rather long).
+This yields symmetric ridges w.r.t.~central axes, whereas we detect ridges only up to orthogonally to the initial light source direction, as visible in Fig.~\ref{gyre-ftle}b),c), respectively.
+In the center location of each gyre, we interpolate the sorrounding ring with an isotropic cell in the middle, leading to chaotic trajectories and thus leading to an increased gradient of the flow map for Hlawatsch's result in c).
+Whereas we detect this position as redudant because of circular overlaps of light distributions, they incorporate trajectories not influencing each other at all.
+This causes the FTLE field to respond in the center of the gyres in Fig.~\ref{gyre-ftle}c), but not for the LTG in b).
+We interpret this behavior within the scope of interpolated trajectories, which are in general more likely to produce chaotic trajectories, since their angle is interpolated (by PCA), which is susceptive to rapid changes and hence leads to unstable behavior on discontinuities.
+What Hlawatsch et al.~detect, could be called local ridges (in terms rather short), what our method detects could rather be called global ridges (in terms rather long).
 This implies, that we obtained reasonable results generated by our method, but the classical tensor field line FTLE is susceptive to instability concerning its directive abilities and does not do things by halves (like we do with our partitioned light distributions), which causes sharper ridges in general.
-Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure.
+Also remember, that the resulting ridges are not to be concerned sharp ones for the LTG, since we measure the separation of whole light distributions, which should be considered a continous response or measure.
 We again depict volume representations for this example:
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
   \includegraphics[height=\textwidth]{img/gyre_volume2.png}
@@
   \begin{minipage}{0.3\textwidth}
+    \centering
     \includegraphics[height=\textwidth]{img/gyre_gauss_contour.PNG}
@@
    \begin{minipage}{0.3\textwidth}
+     \centering
     \includegraphics[height=\textwidth]{img/gyre_volume4.PNG}
@@
 \end{figure}
-When we consider the volumes for the \enquote{gyre} test field, we can remark the holes, which represent the gyres in the centers of the volume for a low and a high threshold, whereby we extract (filter out) the previously observed ridges in $3D$ for all directions $\omega$ simultaneously.
+When we consider the volumes for the \enquote{Gyre} test field, we can remark the holes, which represent the gyres in the centers of the volume for a low and a high threshold, whereby we extract (filter out) the previously observed ridges in 3D for all directions $\omega$ simultaneously.
 
@@
 The approach will detect LCS in the brain's tissue from MRI-measured diffusion tensors of a male subject at the age of 27.
-We compare ParaView's glyph representation to the FTLE method LTG for the test field \enquote{brain}:
+We compare ParaView's glyph representation to the FTLE method LTG for the test field \enquote{Brain}:
 \begin{figure}[!t]
@@
   \begin{minipage}[t]{0.3\textwidth}
+  \centering
   \includegraphics[height=\textwidth]{img/brain_cyl_glyphs.PNG}
-    \caption*{a) glyph representation}
+	a)
     \label{a)}
@@
   \begin{minipage}[t]{0.3\textwidth}
+  \centering
     \includegraphics[height=\textwidth]{img/brain_ftle.PNG}
-     \caption*{b) LTG($0^\circ$)}
+	b)
     \label{b)}
@@
   \begin{minipage}[t]{0.3\textwidth}
+  \centering
     \includegraphics[height=\textwidth]{img/brain_ftle_org.PNG}
-     \caption*{c) FTLE: Hlawatsch's approach}
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Brain} test field}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{\enquote{Brain} test field: a)Glyphs, b) LTG($0^\circ$), c) FTLE: Hlawatsch's approach}
 \label{brain-ftle}
@@
 Regarding Fig.~\ref{gyre_volume}, we find that the approach detects the boundaries of concavities (lat.: Gyri) in the brain, forming the lobus cerebri (commonly known a lobes).
-The invented method could thus, e.g., aid to compute $3D$ models of a subject in medical sciences.
-It emphasizes regions with increased change in flow-rate (throughput), because in this areas the anisotropy and/or tensor magnitude is increased, preventing the intensities from being absorbed, hence yielding to a high FTLE-response.
-The regions highlighted in red, are caused by high tensor magnitudes, as we can infer from the glyph representation revealing no linear anisotropy at all.
-Note that we do not show the volume representation in this case, since all directions (slices) look approximately the same because the tensor field is only introducing changes in tensor magnitude not in anisotropy as inferred from the glyph representation.
+The invented method could thus, e.g., aid to compute 3D models of a subject in medical sciences.
+It emphasizes regions with increased change in flow-rate (throughput), because in this areas the anisotropy and/or tensor magnitude is increased, preventing the intensities from being absorbed, hence leading to a high FTLE-response.
+The regions highlighted in red, are caused by high tensor magnitudes, as we can infer from the glyph representation in Fig.~\ref{brain_ftle_comp}a) revealing no linear anisotropy at all.
+Note, that we do not show the volume representation in this case, since all directions (slices) look approximately the same because the tensor field is only introducing changes in tensor magnitude not in anisotropy as inferred from the glyph representation.
 The regions with a tensor magnitude higher than zero restrict the light propagation to certain areas with the condition, that not all of the energy is absorbed by sorrounding cells in the local neighborhood.
 Again comparing the results to the classical FTLE (Hlawatsch's approach) on tensor fields in Fig.~\ref{brain_ftle_comp}, we notice that the FTLE response is chaotic, since this MRI scan is subject to inherent MRI noise filling the background regions in the image.
-This noise produces random trajectories, leading to an increased gradient of the flow map and thus to a high FTLE response, which reveals trajectories running in random directions in background regions.
+This noise produces random trajectories visible in Fig.
+\ref{brain_ftle_comp}b), leading to an increased gradient of the flow map and thus to a high FTLE response, which reveals trajectories running in random directions in background regions.
 Only in foreground regions, the trajectories are directed in definite manner, which is why the flow map's gradient is minimized along the brain structure (foreground, depicted in blue in Fig \ref{brain_ftle_comp} c)).
@@
   \includegraphics[width=0.9\textwidth]{img/brainDwnsmpl.png}
-    \caption*{a) Glyphs}
+	a)~Glyphs
     \label{a)}
@@
     \includegraphics[width=0.9\textwidth]{img/brainDwnsmpl-TFL.PNG}
-     \caption*{b) Hlawatsch's app.: TFLs}
+	b)~TFLs
     \label{b)}
@@
 \end{figure}
-We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig.~\ref{brain_ftle_comp}.
-We propose the following threshold correction to prevent noisy FTLE results:
-
-We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig.~\ref{brain_ftle_comp}.
+We note, that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig.~\ref{brain_ftle_comp}.
 We propose the following threshold correction to prevent noisy FTLE results:
@@
 This now states a local normalization, which was not feasible with the noisy background before introducing this change, since a local normalization would have amplified noisy signals likewise.
-Now, we are capable of producing a tensor magnitude weighted FTLE mode with global normalization, which we used up to this point and another FTLE mode indicating changes in anisotropy separately while neglecting any tensor magnitudes for local normalization.
+Now, we are capable of producing a tensor magnitude weighted FTLE mode with global normalization, which we used up to this point and another FTLE mode indicating changes in anisotropy separately, while neglecting any tensor magnitudes for local normalization.
 This change will now yield a volume (now depending on the direction) for the local normalization mode, which we will show later on.
-We could also claim, that this correction would be necessary for Hlawatsch's approach as well, as their result is susceptive to noise for the \enquote{brain} test field in Fig.~\ref{brain_ftle_comp} b).
+We could also claim, that this correction would be necessary for Hlawatsch's approach as well, as their result is susceptive to noise for the \enquote{Brain} test field in Fig.~\ref{brain_ftle_comp} b).
 We first show the functionality of the correction on their approach in Fig.~\ref{brain_ftle_noise}.
@@
   \includegraphics[width=0.9\textwidth]{img_new/brain-TFL.png}
-	a) TFLs: noise correction
+	a)
     \label{a)}
@@
     \includegraphics[width=0.9\textwidth]{img/brain_ftle_surf_noisereduction.PNG}
-	b) Hlawatsch's approach: noise correction
-    \label{b)}
-  \end{minipage}
-\caption{\enquote{Brain} test field}
+	b)
+    \label{b)}
+  \end{minipage}
+\caption{\enquote{Brain} test field: a) TFLs: noise correction, b) Hlawatsch's approach: noise correction}
 \label{brain_ftle_noise}
@@
   \caption{\enquote{Brain} test field LTG: a) LTG($0^\circ$), b) LTG volume, c) cross-section}
-\label{heart-ftle}
-\end{figure}
-
-The volumes depicted in Fig.~\ref{heart-ftle} now show variation in $z$-direction, as they are now visualizing pure FTLE fields with no weighting of the tensor magnitude incorporated.
+\label{brain-volumes}
+\end{figure}
+
+The volumes depicted in Fig.~\ref{brain-volumes} now show variation in $z$-direction, as they are now visualizing pure FTLE fields with no weighting of the tensor magnitude incorporated.
 
@@
 \end{figure}
-After applying the LTG approach on the heart dataset, we observe the resulting tensor magnitude weighted FTLE-field (global) and the locally normalized (local) in Figs.
-\ref{heart-ftle}a)\& b).
-Note that the FTLE response for the globally normalized tensor field is again not changing w.r.t.
-direction, which is why we do not show the volumes for the global noramlization.
+After applying the LTG approach on the heart dataset, we observe the resulting tensor magnitude weighted FTLE-field (global) and the locally normalized (local) field in Figs.
+\ref{heart-ftle}a),b).
+Note, that the FTLE response for the globally normalized tensor field is again not changing w.r.t.~direction, which is why we do not show the volumes for the global normalization.
 The approach highlights the regions with diverging tensor field lines resulting from linear anisotropy or increased tensor magnitude, emerging as ridges.
 In Fig.~\ref{heart-ftle}a) we observe the tensor magnitude weighted mode for an inital global normalization, whereas anisotropy characteristics get outshined by tensor magnitude influences.
-On the other hand, in Fig.~\ref{heart-ftle}b) we remark the typical edge artifacts for FG regions known from the classical FTLE.
-Also, we notice the approach detecting changes in anisotropy in lack of any directed ridges, which reveal diverging tensor field lines w.r.t.
-a central axis.
-The results seem to correspond to the ones of Hlawatsch et al.
-certainly only in parts of the domain, because we unrolled all the FTLE responses distributed over all discrete directions, which is why some of the FTLE ridges are missing.
-Note that these ridges are not that sharp compared to Hlawatsch's approach, as we use whole distributions of trajectories which should be interpreted in a more continous way (changing w.r.t.
-direction).
+On the other hand, in Fig.~\ref{heart-ftle}b) we remark the typical edge artifacts for FG regions known from the classical FTLE for our local normalization mode.
+Also, we notice the approach detecting changes in anisotropy in lack of any directed ridges, which reveal diverging tensor field lines w.r.t.~a central axis.
+The results seem to correspond to the ones of Hlawatsch et al.~certainly only in parts of the domain, because we unrolled all the FTLE responses distributed over all discrete directions, which is why some of the FTLE ridges are missing.
+Note that these ridges are not that sharp compared to Hlawatsch's approach, as we use whole distributions of trajectories, which should be interpreted in a more continous way (changing w.r.t.~direction).
 These ridges could now represent blood vessels and such.
@@
 This occurs, since we do not have any MRI noise incorporated into the background of the field.
-Consequently, we obtain results corresponding to Hlawatsch's approach, which is in general susceptive to noise, but not in case the noise was removed or is not present at all initially.
+Consequently, we obtain results corresponding to Hlawatsch's approach, which is in general susceptive to noise, but not in case the noise was removed or is not present at all, initially.
 From observing Hlawatsch's approach detecting the heart edges, we can infer that the flow is also congruent inside of the heart, inducing a gradient response only to the edges.
-In our case, we have an offset response incorporated which happens even for total spherical isotropy, wherefore we will obtain a response even if the flow does not change locally (presumed it is not absorbed by low tensor magnitudes).
-This in turn stems from the fact, that we use a energetic flood-fill difference, which inherently introduces an offset response.
-We can again apply a threshold for the normalization to obtain a result non-dependent on the noise to visualize raw changes in anisotropy and thus to obtain a variation in $z$-direction, ie., along the light source direction.
+In our case, we have an offset response incorporated, which happens even for total spherical isotropy, wherefore we will obtain a response even if the flow does not change locally (presumed it is not absorbed by low tensor magnitudes).
+This in turn stems from the fact, that we use an energetic flood-fill difference, which inherently introduces an offset response.
+We can again apply a threshold for the normalization to obtain a result non-dependent on the noise to visualize raw changes in anisotropy and thus to obtain a variation in $z$-direction, i.e., along the dimension of inital light source directions.
 \begin{figure}[!t]
@@
  \caption{\enquote{Heart} test field: LTG volumes}
-\label{heart-ftle}
-\end{figure}
-Fig.~\ref{heart-ftle} depicts the volumes for the heart test field, whereas we can observe a 3D ridge running in spiral along the $z$-direction.
-The FTLE responses are shifted w.r.t.
-varying directions along the $z$-axis.
-
-We also introduced random numbers in to the our tensor field generation process.
+\label{heart-ftle-volumes}
+\end{figure}
+Fig.~\ref{heart-ftle-volumes} depicts the volumes for the heart test field, whereas we can observe a 3D ridge running in spiral along the $z$-direction.
+The FTLE responses are shifted w.r.t.~varying directions along the $z$-axis.
+
+We also introduced random numbers to our tensor field generation process.
 Therefore, we define a uniform grid with low resolution ($9\times9$) and interpolate bilinearly to obtain a tensor field with higher resolution exhibiting continous changes in anisotropy and tensor magnitude.
@@
   \includegraphics[width=\textwidth]{img_new/random-global.png}
-    \caption*{Global normalization}
+	a)
     \label{a)}
@@
     \includegraphics[width=\textwidth]{img_new/random-local.PNG}
-     \caption*{Local normalization}
+	b)
     \label{b)}
@@
 \end{figure}
-Observing the generated random test field both with global and local normalization in Fig.~\ref{random}a)\& b) respectively, we remark that the propability for small numbers to have a high ratio in combination is increased, as linear anisotropy occurs mainly in BG regions (low tensor magnitudes).
+Observing the generated random test field both with global and local normalization in Fig.~\ref{random}a), b) respectively, we remark that the propability for small numbers to have a high ratio in combination is increased, as linear anisotropy occurs mainly in BG regions (low tensor magnitudes).
 In case we use global normalization in Fig.~\ref{random_FTLE}a), we will again obtain a FTLE field independent on the direction, since tensor magnitudes outshine the changes in anisotropy in this test case.
@@
   \includegraphics[width=\textwidth]{img_new/random_global.png}
-b)
+	a)
     \label{a)}
@@
     \includegraphics[width=\textwidth]{img_new/random_local_slice0.PNG}
-	a)
+	b)
     \label{b)}
@@
 Keeping this consideration in mind, we observe the approach detecting raw changes in anisotropy.
-Note, that this test field does not introduce clearly defined ridges with tensor field lines diverging w.r.t.
-to a middle axis.
+Note, that this test field does not introduce clearly defined ridges with tensor field lines diverging w.r.t.~to a middle axis.
 We just introduce rapid random variations in anisotropy along the spatial domain, which is why the FTLE response is independent on the direction in this case, as well.
@@
   \end{minipage}
-  \caption{Spatial resolution pyramid of \enquote{inverse} test field - slice}
+  \caption{Spatial resolution pyramid of \enquote{Inverse} test field - slice}
 \label{res_ftle}
@@
   \end{minipage}
-  \caption{Spatial resolution pyramid of \enquote{inverse} test field - volume}
+  \caption{Spatial resolution pyramid of \enquote{Inverse} test field - volume}
 \label{res_ftle_volume}
@@
 The right image on the top, shows a result fulfilling the Nyquist-Shannon sampling theorem, as it appears with no jaggies or pixelation and thus successfully deceiving the human mind in being continous.
-The volumes on the other hand, exhibit a slight difference, which is due to the revealing of new discretized directions implicitly in the tensor field, as shifts in angle become more gradual and continuously for a more dense sampling of the same implicitly content-based field.
+The volumes on the other hand, exhibit a slight difference, which is due to the revealing of new discretized directions implicitly in the tensor field, as shifts in angle become more gradual and continuously for a more dense sampling of the same implicitly defined content-based field.
 This effects a more precise behavior, which now indicates the orthogonal ridges from each of the $2$ directions in parallel to the light source direction (axis-parallel) on each side of the volume.
 They are evenly spaced in a shift of half the volume height which represents the angle $180^\circ$ (2 antiparallel directions in axis directions).
-The blue span (hole), we perceive in the middle of the cross-section, is representing the antiparallel direction, which in turn is missing, since we only can detect divergence of tensor field lines in axis-parallel up to -orthogonal directions (if we consider them pointing into the diverging direction).
-Note that we already discussed this feature in section \ref{chap:meth}.
-Also remark, that the ridges get sharper with increasing spatial test field (input data) resolution because the implicitly defined content of the tensor field is more densely sampled at a higher resolution allowing for more continous changes in directions of the major eigenvectors (glyph orientations).
+The blue span (hole), we perceive in the middle of the cross-section, is representing the antiparallel direction, which in turn is missing, since we only can detect divergence of tensor field lines in axis-parallel up to -orthogonal directions (if we consider the light source pointing into the diverging direction).
+Note, that we already discussed this feature in Sect.~\ref{chap:meth}.
+Also remark, that the ridges get sharper with increasing spatial test field (input data) resolution, because the implicitly defined content of the tensor field is sampled more densely at a higher resolution allowing for more continous changes in directions of the major eigenvectors (glyph orientations).
 \textit{Angular resolution pyramid}
@@
   \end{minipage}
-  \caption{Angular resolution pyramid of \enquote{inverse} test field - slice}
+  \caption{Angular resolution pyramid of \enquote{Inverse} test field - slice}
 \label{resA_ftle}
@@
   \end{minipage}
-  \caption{Angular resolution pyramid of \enquote{inverse} test field - volume}
+  \caption{Angular resolution pyramid of \enquote{Inverse} test field - volume}
 \label{resA_ftle_volume}
@@
 \toprule
-\enquote{drain} - \ref{ftle_base} & $101\times101\times64=652864$ & $179878$ & $47.02\mathop{MB}$\\ 
+\enquote{Drain} - \ref{ftle_base} & $101\times101\times64=652864$ & $179878$ & $47.02\mathop{MB}$\\ 
 \hline 
-\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times64=652864$ & $68787$ & $47.02\mathop{MB}$\\
+\enquote{Inverse} - \ref{inverse-ftle} & $101\times101\times64=652864$ & $68787$ & $47.02\mathop{MB}$\\
 \hline 
-\enquote{random-local} - \ref{random_FTLE} & $101\times101\times32=326432$ & $59896$ & $23.5\mathop{MB}$\\ 
+\enquote{Random-local} - \ref{random_FTLE} & $101\times101\times32=326432$ & $59896$ & $23.5\mathop{MB}$\\ 
 \hline
-\enquote{random-global} - \ref{random_FTLE} & $101\times101\times32=326432$ & $757$ & $23.5\mathop{MB}$\\ 
+\enquote{Random-global} - \ref{random_FTLE} & $101\times101\times32=326432$ & $757$ & $23.5\mathop{MB}$\\ 
 \hline 
-\enquote{isotropic} - \ref{bow-ftle} & $51\times51\times3283232$ & $1357$ & $5.998\mathop{MB}$\\ 
+\enquote{Isotropic} - \ref{bow-ftle} & $51\times51\times3283232$ & $1357$ & $5.998\mathop{MB}$\\ 
 \hline
-\enquote{rings} - \ref{bow-ftle} & $51\times51\times32=83232$ & $5761$ & $5.998\mathop{MB}$\\ 
+\enquote{Rings} - \ref{bow-ftle} & $51\times51\times32=83232$ & $5761$ & $5.998\mathop{MB}$\\ 
 \hline 
-\enquote{bow} - \ref{bow-ftle} & $51\times51\times32=83232$ & $6128$ & $5.998\mathop{MB}$\\ 
+\enquote{Bow} - \ref{bow-ftle} & $51\times51\times32=83232$ & $6128$ & $5.998\mathop{MB}$\\ 
 \hline 
-\enquote{star} - \ref{star-ftle} & $51\times51\times32=83232$ & $5815$ & $5.998\mathop{MB}$\\ 
+\enquote{Star} - \ref{star-ftle} & $51\times51\times32=83232$ & $5815$ & $5.998\mathop{MB}$\\ 
 \hline 
-\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times16=163216$ & $12308$ & $11.754\mathop{MB}$\\ 
+\enquote{Inverse} - \ref{inverse-ftle} & $101\times101\times16=163216$ & $12308$ & $11.754\mathop{MB}$\\ 
 \hline 
-\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times32=326432$ & $24571$ & $23.5\mathop{MB}$\\ 
+\enquote{Inverse} - \ref{inverse-ftle} & $101\times101\times32=326432$ & $24571$ & $23.5\mathop{MB}$\\ 
 \hline
-\enquote{brain} - \ref{brain-ftle} & $128\times128\times64=1048576$ & $2427$ & $75.5\mathop{MB}$\\
+\enquote{Brain} - \ref{brain-ftle} & $128\times128\times64=1048576$ & $2427$ & $75.5\mathop{MB}$\\
 \hline
-\enquote{heart} - \ref{heart-ftle} & $256\times256\times32=2097152$ & $117199$ & $151.0\mathop{MB}$\\
+\enquote{Heart} - \ref{heart-ftle} & $256\times256\times32=2097152$ & $117199$ & $151.0\mathop{MB}$\\
 \hline 
@@
 In this work, we have discussed the advantages and purposes of tensor field visualization in practical applications.
-Furthermore, we have presented a presented a technique to extract meaningful structures from tensor fields, which can, as a secondary purpose, also be used to visualize meaningful structures in $2D$ virtual geometrical scenes.
+Furthermore, we have presented a presented a technique to extract meaningful structures from tensor fields, which can, as a secondary purpose, also be used to visualize meaningful structures in 2D virtual geometrical scenes.
 To put it into context, we discussed its relations to vector field visualization techniques and other tensor field and light transport visualization methods.
 Also, we captured global illumination methods, since the proposed visualization method is derived from a global light transport method propagating intensities inside of a given Cartesian grid, consistent of cells defined by polar profiles (distributions), respectively.
-The intensities are redistributed stochastically by transmission profiles derived from the tensors in the field, used as windowing (weighting) functions.
+The intensities are redistributed stochastically by transmission profiles, derived from the tensors in the field, used as windowing (weighting) functions.
 Based on this propagation algorithm, we can compute a FTLE-like field on the tensor field, as we obtain a final light distribution for each light source configuration (position and direction), which is considered as an energy-based flow map.
 In analogy to the computation of the FTLE field for vector fields, we analyze the gradient of the resulting flow map.
-The incongruity on the other hand, is that we need an additional inital direction on top of the inital position, because the tensors also dependent on the incoming direction, not as vectors only on their location $x,y$.
-As a result, we obtain a $3D$ scalar field, which can be visualized as volume (e.g.
+The incongruity on the other hand, is that we need an additional initial direction on top of the inital position, because the tensors also depend on the incoming direction, not as vectors only on their location $x,y$.
+As a result, we obtain a 3D scalar field, which can be visualized as volume (e.g.
 heatmap representation) to extract and segment ridges, i.e., key regions separating areas of correlating directional distributions.
-These ridges are in turn called Lagrangian coherent structures (LCS), which can be found in nature in , e.g., vortices in flow fields or moraines in glaciers.
+These ridges are in turn called Lagrangian coherent structures (LCS), which can be found in nature in, e.g., vortices in flow fields or moraines in glaciers.
 We designed comprehensive synthetic test tensor fields to evaluate our approach.
 Once we were convinced over its principal functionality, we applied the approach to some real data examples, namely DT-MRI scans of human brain and heart.
-Also we designed test runs to verify the physical plausibility of the results of the propagation scheme.
-The maps or fields we obtain, can make a statement about active regions in tensor fields, which do reveal the greatest tensor magnitude and anisotropy in combination.
-The approach is positioned in interesting relation to the works of Zirr et al.
-\cite{zirr}, which perform their FTPD (FTLE-like) computation by random sampling of trajectories to compute light transport separatrices.
+Also, we designed test runs to verify the physical plausibility of the results of the propagation scheme.
+The maps or fields we obtain, can make a statement about active regions in tensor fields, which do reveal the greatest tensor magnitude and anisotropy in combination or alternatively separately.
+The approach is positioned in interesting relation to the works of Zirr et al.~\cite{zirr}, which perform their FTPD (FTLE-like) computation by random sampling of trajectories to compute light transport separatrices.
 They do not incorporate any transmission weighting functions, but instead blocker (obstacle) or reflecting geometry, which causes a FTLE response as well, as edges or mirrors cause light paths to cross each other chaotically in similar manner.
-Furthermore, the work of Hlawatsch et al.
-\cite{hlawatsch} obtains an FTLE field on tensor fields in similar style, as they use a coordinate map to sample their trajectories.
+Furthermore, the work of Hlawatsch et al.~\cite{hlawatsch} obtains an FTLE field on tensor fields in similar style, as they use a coordinate map to sample their trajectories.
 
 A possible application would be mechanical component design, as we could segment ROIs exhibiting increased loads or in a greater scope even machine component design, which are produced in series.
-The invented technique could also aid geologist in analyzing stress in cross-sections of the earth to be able to predict earthquakes sufficiently soon enough.
-In general one could measure vortices in directional distributions and hence tensor fields with a topological background (e.g.
+The invented technique could also aid geologists in analyzing stress in cross-sections of the earth to be able to predict earthquakes sufficiently soon enough.
+In general, one could also measure vortices in directional distributions and hence tensor fields with a topological background (e.g.
 extracting separatrices of manifolds).
@@
 As an outlook, we could apply our invented method to uncertain vector fields, which could be sampled in transmission profiles in discrete polar coordinates, just as we do in our work with the polar ellipsoids.
-Future works might include the previously mentioned application of the approach to uncertain vector fields or evaluating its applicability to real data in segmenting meaningful structures from engineering and medical data.
+Future works might include the previously mentioned application of the approach to uncertain vector fields, or evaluating its applicability to real data in segmenting meaningful structures from engineering and medical data.
 Also we could imagine to implement a GPU parallelized version of the method to be able to propagate the intensities more efficiently, which would decrease the runtime in general.

### 2019-07-18 beamer2thesis/dummy.tex
update

--- 2019-07-08 thesis-template/msc-thesis.tex
@@
-	
-% This template was initially provided by Dulip Withanage.
-% Modifications for the database systems research group
-% were made by Conny Junghans,  Jannik Strötgen and Michael Gertz
-
-\documentclass[%enabledeprecatedfontcommands,
-     12pt,         % font size
-     a4paper,      % paper format
-     BCOR10mm,     % binding correction
-     DIV14,        % stripe size for margin calculation
-%     liststotoc,   % table listing in toc
-%     bibtotoc,     % bibliography in toc
-%     idxtotoc,     % index in toc
-%     parskip       % paragraph skip instad of paragraph indent
-     ]{scrreprt}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% PACKAGES:
-
-% fix bitmap fonts
-\usepackage{lmodern}
+\documentclass{scrartcl}
 \usepackage{tabularx}
 \usepackage{booktabs}
-% Use German :
-\usepackage[ngerman,english]{babel}
-\usepackage{amsmath}
-\usepackage{upgreek}
-\usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-% Input and font encoding
-\usepackage[latin1]{inputenc}
-\usepackage[T1]{fontenc}
-% Index-generation
-\usepackage{makeidx}
-% Einbinden von URLs:
-\usepackage{url}
-% Special \LaTex symbols (e.g. \BibTeX):
-%\usepackage{doc}
+\usepackage{csquotes}
 % Include Graphic-files:
 \usepackage{graphicx}
-\usepackage{csquotes}
-% Include doc++ generated tex-files:
-%\usepackage{docxx}
-% Include PDF links
-%\usepackage[pdftex, bookmarks=true]{hyperref}
+\usepackage{caption}
+\usepackage{subfig}
+\usepackage{url}
+%\usepackage{subfig}
+\newsubfloat{figure}
+\newcommand{\source}[1]{\vspace{-3pt} \caption*{ Source: {#1}} }
 
-% Fuer anderthalbzeiligen Textsatz
-\usepackage{setspace}
-
-\usepackage{caption}
-
-% hyperrefs in the documents
-\usepackage[bookmarks=true,colorlinks,pdfpagelabels,pdfstartview = FitH,bookmarksopen = true,bookmarksnumbered = true,linkcolor = black,plainpages = false,hypertexnames = false,citecolor = black,urlcolor=black]{hyperref}
-%\usepackage{hyperref}
-\usepackage{letterspace}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% OTHER SETTINGS:
-
-% Pagestyle:
-\pagestyle{headings}
-
-% Choose language
-\newcommand{\setlang}[1]{\selectlanguage{#1}\nonfrenchspacing}
-\newcommand*\mean[1]{\bar{#1}}
-\newcommand{\source}[1]{\vspace{-3pt} \caption*{ Source: {#1}} }
-\setlength{\emergencystretch}{3em}
 \begin{document}
 
-% TITLE:
-\pagenumbering{roman} 
-\begin{titlepage}
-
-\vspace*{1cm}
-\begin{center}
-\vspace*{3cm}
-\textbf{ 
-\Large Heidelberg University\\
-\smallskip
-\Large Institute of Computer Science\\
-\smallskip
-\Large Visual Computing Group (VCG)\\
-\smallskip
-}
-
-\vspace{3cm}
-
-\textbf{\large Master's Thesis} % Bachelor-Arbeit
-
-\vspace{0.5\baselineskip}
-{\huge
-\textbf{Light Transport Techniques for Tensor Field Visualization}
-}
-\end{center}
-
-\vfill
-
-{\large
-\begin{tabular}[l]{ll}
-Sebastian Bek\\
-Student ID: & 3481802\\
-Supervisors: & Prof. Filip Sadlo, Dr. Susanne Krömker\\
-Submission date: & 07.07.2019
-\end{tabular}
-}
-
-\end{titlepage}
-\newpage\null\thispagestyle{empty}\newpage
-
-\onehalfspacing
-
-\thispagestyle{empty}
-
-\vspace*{100pt}
-\noindent
-\emph{dt.:}
-Ich versichere, dass ich diese Master-Arbeit selbstständig verfasst und nur die angegebenen
-Quellen und Hilfsmittel verwendet habe und die Grundsätze und Empfehlungen ``Verantwortung in der Wissenschaft'' der Universität Heidelberg beachtet wurden. 
-\bigskip
-\\
-\emph{eng.:}
-I hereby assure, that I drafted this thesis independently and only used the sources and materials labeled as references and that the conventions, principles and recommendations ``Verantwortung in der Wissenschaft'' of the Heidelberg University have been regarded.
-\vspace*{50pt}
-\noindent
-
-\underline{\phantom{mmmmmmmmmmmmmmmmmmmm}}
-
-\medskip
-\noindent 
-Abgabedatum / Due Date: 07.07.2019
-\newpage
-
-% Add a brief summary of your topic and contributions (Zusammenfassung) in German *and* in English:
-\chapter*{Zusammenfassung}
-%\input{zusammenfassung}
-
-Tensorfelder werden häufig in Verbindung mit mechanischen Spanunngsverteilungen in 2D/3D-Gittern gebracht (vgl. Cauchy-Stress Tensor), haben aber auch andere praktische Bedeutungen in der Physik.
-Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren, die aber alternativ auch verwendet werden kann um die Lichtausbreitung in anisotropen Medien zu beschreiben.
-Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema für kartesische Gitter ab, welches die Prinzipien der Ausbreitungsdämpfung und Energieerhaltung beachtet und Lichtverteilung/en für gegebene Lichtquellenrichtung/en und -positionen approximieren kann.
-Als unterliegendes Modell werden die Transmissionsprofile innerhalb diesem Gitter als Kristallfiberstruktur angenommen (vgl.
-Edelsteine: Katzenaugeneffekt beim Tigerauge).
-Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiemaßen des unterliegenden Tensorfelds zu interpretieren.
-Dafür erstellen wir per Hauptachsentransformation ein Eigensystem für jede Zelle des Tensorfelds und modellieren eine zugehörige Ellipsengleichung als (Wichtungs-) Transmissionsprofil.
-Dabei werden die physikalischen Eigenschaften des Tensorfelds (Orientierung/Anisotropie/Amplitude) als Lichtleitfähigkeit interpretiert und die Lichtverteilung berechnet.
-Folglich kann die resultierende Lichtverteilung als 2D Skalarfeld oder Polarplot für zusätzliche Richtungsinformation repräsentiert und visualisiert werden.
-Wir messen Impulsantworten des Lichtausbreitungs-Prozesses moduliert mit den anisotropen Charakteristika des Tensorfelds, indem Delta-Pulse an jeder Position und in jede Richtung als Stimuli verwendet werden.
-Wir lassen uns von der Methode FTLE aus dem Bereich Particle Tracing der Vector Field Visualization inspirieren, indem wir den Gradient des Flussfelds der resultierenden Lichtverteilungen analysieren und visualisieren.
-Ähnlich wie FTLE in Vektorfeldern Gräte/Kämme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, nämlich anziehende/abstoßende/sattelnde Tensorfeldlinien die häufig Schlüsselstrukturen in der Natur repräsentieren, in Tensorfeldern.
-Wir führen diese Größe als Light Transport Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtfluss, zu visualisieren.
-\newpage
-
-\chapter*{Abstract}
-%\input{abstract}
-Many problems in science and engineering require
-tensor field representations.
-Tensor fields, are commonly associated with stress distributions in 2D/3D grids, but also have many other meanings in various fields.
-By means of global illumination techniques, we motivate a new method to visualize tensor fields,
-which is also capable of visualizing the light transport (propagation) in anisotropic media.
-We present the concept of applying techniques from the field of global illumination to tensor field analysis.
-Thereby, we interpret the tensors as a conductivity property for light transport.
-As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles, and is able to approximate light distributions for given light source position(s) and direction(s).
-The transmission profiles within this grid are considered as \enquote{crystal fiber structures} as an underlying physical model (e.g.: gemstones: tiger's eye's cat's eye effect).
-The consequent task is to interpret anisotropic fiber structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
-For this, we derive an eigenbasis by principal component analysis for every cell of the tensor field and form an ellipsoid equation as transmission (conductivity) function for the transmitted light profiles.
-Hence, the resulting light distribution can be visualized via 2D scalar field color coding or polar plots for directional information.
-On this basis, we measure impulse responses of the tensor field with Dirac pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions.
-This map is considered as a global illumination energy flow field.
-We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field, formed by the resulting light distributions.
-Similar to how the FTLE detects ridges/Lagrangian coherent structures in vector fields, our approach captures these structures likewise, namely regions revealing attracting/repelling/saddling tensor field lines which frequently represent key structures in nature, in tensor fields.
-We denote this entity as the light transport gradient (LTG), which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global light transport in the grid.
-We evaluate our approaches for plausibility and comparability within an extensive evaluation and demonstrate its performance on several examples of both synthetic and real datasets.
-\newpage
-
-% MAIN PART:
-% Table of contents (Inhaltsverzeichnis)
-\tableofcontents
-\cleardoublepage
-\pagenumbering{arabic}
-
-% List of figures (Abbildungsverzeichnis):
-%\listoffigures
-% List of tables (Tabellenverzeichnis):
-%\listoftables
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Here, the actual content of your thesis begins
-% You can either put all the text here or use individual files to store the chapters of your thesis.
-% Below are templates for both alternatives.
-
-\chapter{Introduction}\label{intro}
-
-In this work, we discuss the advantages of tensor field analysis and motivate a new method for tensor field visualization through a coherent light transport concept.
-We shortly state the problem setting, objectives and give a quick overview about the structure of this work.
-The main contributions of this work include:
-
+\begin{table}[h]
+\scriptsize
+\centering
+\caption{Challenges in generative modeling:}
+\begin{tabularx}{\textwidth}{X|X|X}
+Global Illumination Methods & Symmetric Tensor Field Visualization & Asymmetric Tensor Field Visualization\\
+\toprule
 \begin{itemize}
-	\item a light transport model (propagation scheme) following basic but crucial physical principles,
-	\item visualization of the propagation results in $2D$ polar profiles and scalar fields,
-	\item application of this model for tensor field visualization interpreting tensors as light transmission properties,
-	\item a FTLE (Finite-time Lyapunov exponents)-related approach called light transport gradient (LTG) for visualizing key structures, namely LCS (Lagrangian coherent structures) in $2D$ second-order tensor fields, and
-	\item application of our approach to both synthetic and real data involving brain and heart datasets.
+	\item
 \end{itemize}
-%Dieses Kapitel gibt einen ?berblick ?ber die Arbeit. Gerade der
-%Abschnitt zur Motivation soll allgemein verst?ndlich geschrieben
-%werden. Die Einleitung sollte auch wichtige Referenzen enthalten.
-
-\section{Motivation}
-\label{sec:motiv}
-Tensor field visualization is gaining importance as a relevant tool for the analysis of fluid and solid mechanics, where tensor representations are crucially needed.
-This is always the case, when a directional distribution is needed for each single point in space.
-Tensors occur most commonly in medical, scientific and engineering applications, making them a frequently used mathematical entity in science.
-In general, a tensor is a order-$o$ generalization of scalar, vector and matrix and appears, e.g., as the Jacobian matrix (gradient in vector fields), or as stress tensor in solid and fluid continuum mechanics.
-Basically, tensors describe compressions, tensions, rotations and volume changes in both solid and fluid material, as illustrated in Fig.~\ref{ToS}.
-While most techniques focus on symmetric tensor field visualization (like glyphs and tensor field lines), some other recent works address the problem of asymmetric tensor field visualization.
-The issue with asymmetric (non-symmetric) or therein included antisymmetric (skew-symmetric) tensors is, that they do not always yield real eigenvalues, which is needed to determine a real eigenbasis to set up glyphs, hyperstreamlines or tensor field lines.
-Nevertheless, there has been extensive work in this domain as well.
-Classically, the issue with asymmetric tensor fields is, that they do not yield real eigenvalues in general, which is needed to set up glyph representations.
-In this work, we will use the matrix decomposition obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA (principal component analyis) principal ellipsoid representing the tensor transformation as suggested by Moler \cite{moler} to state another alternative of processing asymmetric tensors.
-2D tensor field visualization is useful for cases where only two-dimensional data is available, such as satellite remote sensing, particle imagery velocimetry and earthquake data \cite{laramee}.
-In their field of work, geologists and medical scientists are struggling with complicated number arrays which seem almost impossible to interpret in their raw, unprocessed form.
-Nevertheless, comprehensive tensor field visualization is aiding scientists in their everyday lifes understanding complex data loads, such as tensor fields.
-To cope with the arising problem of tensor field visualization, one needs to advance and extend respective techniques.
-
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.6\textwidth]
-    {img/DIFFERENT_TYPES_OF_STRESS.png}
-  }
-  \caption{Different types of stress}
-  \captionsetup{font={footnotesize,bf,it}}
-  \caption*{source: \url{https://upload.wikimedia.org/wikipedia/commons/e/eb/DIFFERENT_TYPES_OF_STRESS.png}}	
-  \label{ToS}
-\end{figure}
-
-\section{Objectives}
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.5\textwidth]
-    {img/hyperstreamlines.png}
-  }
-  \caption{Hyperstreamlines in tensor field visualization}
-  \captionsetup{font={footnotesize,bf,it}}
-  \caption*{source: \url{http://www2.cs.uh.edu/~chengu/Teaching/Fall2018/Lecs/Lec17.pdf}}	
-  \label{streamlines}
-\end{figure}
-The main purpose of visualization in general, is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation.
-This implies, e.g., clustering, structuring or projection techniques.
-When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor - medical resonance imaging) \cite{hlawatsch}, when measuring water molecule diffusion characteristics within tissue, or furthermore in simulated or measured stress tensor fields from solid and fluid mechanics.
-In these named domains of science the problem arises, that tensor fields can be generated or measured with physical/mathematical background, but not be regarded intuitively.
-Necessity begets ingenuity as always, which lets visualization methods and systems emerge in this context.
-Concurrently, the problem setting of this work encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids, commonly consisting of second-order tensors (cf.~Cauchy stress tensor).
-In general, these grids span a spatial domain with sample positions indicated by vertices or cell centers in the dual grid, depending on the mode of interpretation.
-Tensor fields feature a sampling grid, consistent of tensors, which is spaced uniformly in the commonly used regular grid.
-A Cartesian grid, which we will utilize in this context, is a grid normalized to a certain unit length, as a special case of a regular grid.
-That is, each cell is formed as a unit cube (square) in relation to the normalizing unit length.
-We aim to use global illumination methods to model light transport, directed by tensor fields, within such type of grids, which we consider as light propagation volumes (LPVs) \cite{dachsbacher}.
-Congruently, the most basic objective of this work is to design a simple light propagation scheme following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation, eventually to visualize tensor fields.
-Of course, the propagation scheme will need some kind of stimulus which can be defined through symbolic functions (e.g., $r(\omega)=1$) by the user which we sample (discretize) as a primary step.
-This way, the approach will operate as a light transport visualization technique for light distributions given in polar coordinates, provided the user set an initial profile into the grid.
-Furthermore, this propagation scheme should either be used to visualize the light transport paths, defined by the anisotropy characteristics of tensor fields and tensor magnitudes of a given tensor field in a 2D grid.
-Or, as a secondary purpose, be used to visualize the light transport in a defined, given geometrical scene in a $2D$ grid.
-However, the tensor field needs to be preprocessed somehow to generate a footprint map of transmission profiles defined by the tensor field's anisotropy characteristics, which we will handle through PCA yielding a principal component ellipsoid.
-For this purpose, we will let the transmission profiles act on our simple light propagation scheme to have a kind of  directive effect on the light transport controlled by the tensor's anisotropy characteristics.
-The transmission profiles should modulate the outgoing intensities in a sense of redirection or more appropriately, redistribution.
-This will yield an approach, capable of computing approximated final light distributions for a given light source configuration inside of a 2D lattice.
-As this process requires a lot of user-interaction and is not to be concerned as generalizing to all possible scenarios, we aim to do something more elaborate on the basis of the light propagation scheme.
-We place light sources in form of Dirac pulses at any position and pointing in any direction represented in the grid, which represents a flow map, which we will analyze for its gradient to reveal the topolgical structure of the tensor field.
-In the sense of automization and optimization, another consequent aim is to implement an approach borrowed from vector-field visualization to reveal structures, which are hard to grasp by ordinary methods.
-We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges representing LCS (Lagrangian Coherent Structures).
-These structures can be classified (identified) as repelling, attracting and saddling (i.e., forming a saddle) local features, which frequently represent key structures in nature, such as moraines in glaciers or vortices in flow fields.
-Commonly, in vector field visualization, LCS separate regions of different flow behavior, but in our case they separate regions, or in other words subdomains, of varying directional distributions (manifesting in diverging/converging tensor field lines leading to compressions, tensions, roations or varying tensor magnitude causing volume changes) represented by tensors, as we use our tensor-controlled light transport visualization (propagation scheme) as a basis or framework to compute the FTLE field.
-Consequently, we propose another method to compute an FTLE-like field on tensor fields.
-The FTLE field responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other conceptually, because of bidirectional stress directions.
-Since we effectively measure the finite-time Lyapunov exponent (FTLE) with both forward and backward integration time (stimulus directions, in our case), we will detect both cases dependent on our initial light source direction.
-Both invented methods, will give new insights into the topology and anisotropy of tensor fields and provide new tools for the structural analysis.
-The greater goal of this work is therefore the encoding of a tensor field into certain abstracted representations and the interpretation of meaningful results evaluating the proposed techniques for possible uses.
-The designed approach could, e.g., be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations.
-A classical example for tensor field visualization is plotting hyperstreamlines following major/minor eigenvectors, as depicted in Fig.~\ref{streamlines}, which features a structurally unstable point (purple), a wedge (red) and a trisector (green).
-
-%In diesem Abschnitt sollen neben den Herausforderungen und der
-%Problemstellung insbesondere die Ziele der Arbeit beschrieben werden. 
-\section{Structure of this Work}
-
-For the rest of this work, we will give a quick overview on the structure and outline.
-Next, in Chap.~\ref{chap:grundlagen}, we will capture important fundamentals like tensors, polar coordinates, PCA in the context of singular value decomposition (generalization of eigenvalue decomposition), FTLE and GT (ground truth) datasets.
-Furthermore, we will discuss the most relevant related work in the domain of tensor field visualization and also in the domain of global illumination (indirect illumination rendering), since we use these methods as template models, as well.
-Subsequently, in Chap.~\ref{chap:meth}, we will propose the main contributions and methods of this work and document the results and evaluation in Chap.~\ref{chap:eval}.
-At last, a compact summary, conclusion and outlook will be given in Chap.~\ref{chap:concl}.\\
-
-%Dieser Abschnitt wird meist recht kurz gehalten und beschreibt im
-%Prinzip nur den Aufbau des Rests der Arbeit. Zum Beispiel: In Kapitel
-%\ref{chap:grundlagen} geben wir einen ?berblick ?ber die  Grundlagen
-%zu der Arbeit sowie ?ber verwandte Arbeiten. In Kapitel
-%\ref{chap:main} stellen wir dann \ldots vor. \ldots etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Related Work}
-\label{chap:related}
-
-There has been extensive recent work on symmetric tensors but in comparison little on asymmetric ones.
-Since we are not able to cover the whole scope in this work, we will focus on the most relevant works on global illumination and tensor field visualization.
-We will talk about the publications from Hlawatsch et al.
-\cite{hlawatsch} and Zirr et al.
-\cite{zirr} in particular, since they constitute thematically close related works.
-But first, we will cover related works on global illumination methods, as these form a basis or prerequisite for our work.
-Please note, that they do not focus on tensor field visualization at all.
-Remember, that we will inventively employ global illumination methods for tensor field visualization as an innovative step and that we want to cover previous works on our base approach, as well.
-
-\section{Global Illumination Methods}
-Since we use a specific category of global illumination methods as a basis, we decide to only discuss lattice based methods which work, corresponding to their name, on a grid.
-Remark, that Global Illumination methods classically focus on realistic indirect lighting.
-Thus, they aim on producing a most realistic image, what in our case turns out to be a most informative image in correspondence.
-The most early work in this field is propably the Discrete Ordinates Method (DOM) by Chandrasekhar \cite{Chandrasekhar}, which is used specifically for computing radiative transfer in participating media (for volume rendering).
-The radiative transfer equation is discretized spatially and in the angular domain.
-The radiance distributions are represented in a grid and light is distributed and exchanged between neighboring volume elements which reduces the overall effort to local and independent computations.
-Furthermore, there has been more recent work on describing the light propagation as a diffusion process within a simple photon transport model, based on the Lattice-Boltzmann method \cite{geist}.
-DOMs were recently revisited by Fattal \cite{fattal}.
-This work addresses artifacts emerging from spatial discretization, which are called \enquote{ray effects}, and from successive interpolation called \enquote{light smearing}.
-These methods are considered to be too slow to \enquote{run at interactive framerates} by Dachsbacher et al.
-\cite{dachsbacher}, though \enquote{efficient and potentially highly parallel}.
-Our work is inspired by Lattice-Boltzmann methods (volume grid radiosity) and the lattice-based discrete ordinate methods concerning the light propagation scheme.
-
-\section{Symmetric Tensor Field Visualization}
-Zheng \& and Pang \cite{pang&zheng} et al.
-proposed a texture-based visualization approach called HyperLIC, which extends the concept of LIC to symmetric tensor fields by using an anisotropic 2D-filter kernel oriented along the major or minor eigenvectors.
-The concept of following the major/minor eigenvector along tensor field lines was first invented by Vilanova et al.
-\cite{vilanova}.
-Tensorlines \cite{weinstein} introduce a kind of artificial inertia (running average), making them more resistent to noise.
-The notion of tensor field topology (critical, degenerate points, separatrices) and the concept of hyperstreamlines was first introduced by Delmarcelle and Hesselink \cite{delmarcelle&hesselink}.
-The topology in $3D$ tensor fields was first analyzed by Zheng et al.
-\cite{zheng}.
-The placement of tensor field lines as hyperstreamlines for glyph packing was recently improved by Spencer et al.
-\cite{spencer}.
-A good overview of seeding strategies concerning glyphs is given by McLoughlin et al.
-\cite{mcloughlin}.
-Feng et al.
-\cite{feng} used Voronoi tessellation for placing the glyphs.
-Superquadric glyphs have been proposed by Schultz and Kindlmann \cite{schultz&kindlmann} for general (non-positive definite) tensors.
-De Leeuw and van Wijk \cite{deleeuw&vanwijk} visualize the partial derivative gradient, the Jacobian of the tensor field, to sense the local properties of the field.
-A symmetric tensor can also be visualized by box \cite{makris}, ellipsoid \cite{pierpaoli}, composite \cite{westin2} or superquadric \cite{kindlmann} glyphs.
-There is also a set of scalar measures available in symmetric tensor field visualization such as: mean diffusivity, fractional anisotropy \cite{basser} and anisotropy coefficients \cite{westin}.
-These coefficients can also be visualized by direct volume rendering \cite{kindlmann2}.
-Hlawatsch et al.
-\cite{hlawatsch} obtain a coordinate (flow) map from tensor field lines following major/minor eigenvectors (fibre trajectories) using the maximum eigenvalue of the Cauchy-Green deformation tensor to compute the FTLE-field in homogeneous regions.
-In a similar manner, we aim to use Global Illumination methods propagating light, directed by tensor fields, to obtain a final light distribution (flow map) through successive light propagation cycles.
-The light distributions are redirected by the tensor field employing transmission profiles obtained from ellipsoid glyphs.
-DT-MRI diffusion tensor imaging can help to reveal the functional \cite{denis} or topological \cite{wakana} structure of the brain and muscles by analysing celebral and muscle tissue tractography, which yields a tensor field representing the anisotropy characteristcs of the brain's white matter.
-An overview about DT-MRI visualization techniques is given by \cite{vilanova}.
-Tricoche et al.
-\cite{xavier} use invariant manifolds to extract topological features from tensor fields.
-Other feature extraction-based approaches extract general features as: edges \cite{hancock}, tensor shape\/ orientation \cite{gordon}, interfaces \cite{donnel} and creases \cite{tricochet}.
-\\
-\section{Asymmetric Tensor Field Visualization}
-Zheng and Pang \cite{pang&zheng} proposed the concept of dual eigenvectors and Zhang et al.
-\cite{zhang} extended it to pseudo-eigenvectors and introduced the eigenvector/eigenvalue manifolds to visualize eigenvectors in the complex domain.
-Laramee et al.
-\cite{laramee} focused on the efficient implementation and visualization of these structures and provided an interactive visualization system for asymmetric tensors applicable in fluid and solid dynamics.
-The concept of tensor magnitude has been introduced by \cite{laramee} for means of physical interpretation.
-They also proposed an efficient glyph and hyperstreamline hybrid approach, which made dynamic interaction in real-time in $2D$ tensor fields feasible.
-Palacios et al.
-\cite{palacios} extracted isosurfaces of tensor magnitude, mode and isotropy index.
-
-Last, we will discuss our work in relation to other publications.
-While these previously mentioned works use classical approaches like tensor field lines and glyphs, we will also introduce the concept of Lagrangian coherent structures into our framework contribution to tensor field analysis.
-LCS have been defined as time-dependent analog of separatrices \cite{haller} which are concerned to be robust under noisy conditions \cite{haller2}, which we expect from our technique as well by implication.
-Hlawatsch et al.
-\cite{hlawatsch} obtain similar results (a FTLE-like field on tensor fields), but obtain them by sampling of fiber trajectories instead of global illumination light distributions, as it is the case for this work.
-Our work shares the same basic idea, deriving and computing a FTLE-like field for tensor field visualization generalizing FTLE from vectors to tensors, yet using a light propagation scheme and hence using whole distributions of trajectories incorporating polar profiles.
-Concerning the tensor field visualization method (LTG), our approach is based on FTLE, and principal component analysis (PCA).
-We will detect LCS with an FTLE-derived method considering differences in final light transport distributions through computing the gradient of the resulting flow maps.
-The FTLE-related approach named ``light transport gradient'', to be designed, also states a generalization of the light transport visualization method FTPD proposed by Zirr et al.
-\cite{zirr}, since it allows a similar measure to be computed through defining a geometrical scene and setting the transmission (defined by the tensor field) to an isotropic and constant value of $100\%$.
-In this operating mode, the approach operates as a light transport visualization technique neglecting any tensor field transmission profiles and is thus capable of displaying bounding separatrices of a light transport scenario perturbated by obstacles and mirrors \cite{zirr}.
-%Typischerweise im letzten Abschnitt dieses Kapitels wird dann auf
-%verwandte Arbeiten eingegangen. Entsprechende Arbeiten sind geeignet
-%zu zitieren. Beispiel: Die wurde erstmalig in den Arbeiten von Spitz
-%und Gertz \cite{Spitz2016a} gezeigt \ldots Details dazu werden in
-%dem Buch von Newman zu Netzwerken \cite{Newman2010} erl?utert \ldots.
+&
+\begin{itemize}
+	\item
+\end{itemize}
+& 
+\begin{itemize}
+	\item
+\end{itemize}
+\hline 
+\end{tabularx}
+   \vskip 1pt
+\end{table}
 
@@
 
-\chapter{Fundamentals}
-\label{chap:grundlagen}
-
-In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and we will capture tensor fields, polar coordinates, PCA, FTLE and GT (ground truth) datasets in that sense.
-
-%Die ersten paar Abschnitte in diesem Kapitel f?hren in die Grundlagen
-%zur Arbeit ein. Das k?nnen beispielsweise Grundlagen zu Netzwerken
-%oder zur Informationsextraktion sein.
-
-\section{Tensor Fields}
-\label{sec:tensors}
-Now what is a tensor? A tensor is an $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
-\begin{table}[!b]
-\centering
-\caption{Tensor Shapes}
- \begin{tabular}{c|c|c|c|c|c|c}
-\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
-\hline 
-\textbf{shape} & scalar & vector & matrix & ``3D matrix'' & ... & order-$o$ matrix\\ 
-\end{tabular}
-\label{tensor}
- \end{table} 
-In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ in space (the dimension of the row and column vectors), which is equal to the rank of the tensor for full rank.
-Or in other words the number of column elements $n$ for a $n{\times}n$-matrix.
-A tensor is considered non-degenerate if all eigenvalues $\lambda_i>\lambda_{i+1}$ manifesting in significant anisotropy.
-Tensors follow certain transformation rules which are defined for covariant and contravariant tensors, which indicate the change characteristics under a basis change.
-Contravariant tensors do not change under a change of basis, whereas covariant tensors are not invariant w.r.t.
-basis changes.
-Tensors are most commonly obtained and defined from mathematical transformation equations.
-Hence, any order-$o$ number array could be a tensor, but the definition holds only if the transformation rules apply to these $o$-th order number arrays (in a sense of mathematical/formal interpretation).
-This is mostly true for component indexed order-$o$ number arrays following matrix multiplication and transformation rules found in physics and math manifesting in calculation rules like matrix multiplication.
-Thus, a tensor in general is a order-$o$ number array arranged in meaningful (transformation) order, to put it simple.
-In fact, tensors themselves are interpreted as transformations which reflect an incoming direction vector (measuring, e.g., the stress in that direction) to an outgoing (stress) vector.
-To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features, as these entities are not simple numbers or vectors.
-They occur in form of a tensors, and thus as entities or quantities in physics.
-That is, at each point (typically a Euclidean space or manifold) in space (indexed by the grid) there is a whole distribution of directions, e.g., stresses (elastic and viscous), which needs to be characterized and represented by a single tensor each.
-Collectively, defined at each point in space, these tensors form a tensor field.
-\paragraph{stress tensor}
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.6\textwidth]
-    {img/stress_tensor.png}
-  }
-  \caption{Cauchy stress Tensor}
-  \captionsetup{font={footnotesize,bf,it}}
-  \caption*{source: \url{https://en.wikipedia.org/wiki/Cauchy_stress_tensor\#/media/File:Components_stress_tensor_cartesian.svg}}
-  \label{cauchy}
-\end{figure}
-
-The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig.~\ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$, arranged in row-major order.
-It appears in similar form as viscous stress tensor in fluid mechanics:
-\begin{align*}
-\boldsymbol{\sigma}
- = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
-\boldsymbol{T}^{(\mathbf{e}_2)} \\
-\boldsymbol{T}^{(\mathbf{e}_3)} \\
-\end{matrix}}\right] =
-\left[{\begin{matrix}
-\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
-\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
-\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
-\end{matrix}}\right]  = \left[{\begin{matrix}
-\sigma _x & \tau _{xy} & \tau _{xz} \\
-\tau _{yx} & \sigma _y & \tau _{yz} \\
-\tau _{zx} & \tau _{zy} & \sigma _z \\
-\end{matrix}}\right] =
-\begin{bmatrix}
-\sigma _{11} & \sigma _{12} & \sigma _{13} \\
-\sigma _{21} & \sigma _{22} & \sigma _{23} \\
-\sigma _{31} & \sigma _{32} & \sigma _{33} 
-\end{bmatrix}.
-\end{align*}
-These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in plane normal directions ${x,y,z}$ ($3^2=9$ individual numbers).
-Thus, the stress tensor poses a full valid representation of the stress distribution at any point in space.
-The stress tensor itself can be interpreted as a transformation, which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$.
-The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent (kind of a Euclidean distance of the stress components).
-The tensor itself is a complicated representation and needs to be projected onto its principal axes to lower its dimensionality, as it could be done with, e.g., principal component analysis.
-It is possible to find $2$ principal stress directions in 2D through PCA (see Sect.~\ref{sec:pca} below), for which there are no shear stresses existent in equilibrium.
-These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour.
-The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''.
-So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig.~\ref{HAT} in Sect.~\ref{sec:pca}.
-Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle.
-It is therefore somehow representing the distortion of an absolutely isotropic object (unit circle) by whatever the underlying affine tensor transformation effects (e.g.: stretching, shearing, mirroring, scaling, rotating).
-Let us consider it as some lower dimensional, more intuitive representation of the tensor which we were left with in the first place.
-To put it all together, a tensor is a mathematical transformation which is encoded inside of an abstract order-$o$ number array, which comes in need, whenever we have a whole distribution of directions manifesting itself inside of a tensor (matrix) transformation.
-\newpage
-\section{Polar Coordinates}
-\label{sec:polar}
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.8\textwidth]
-    {img/polarkoordinaten.png}
-  }
-  \caption{polar space}
-  \captionsetup{font={footnotesize,bf,it}}
-   \caption*{source: \url{https://de.wikipedia.org/wiki/Datei:Ebene_polarkoordinaten.PNG}}
-  \label{polar}
-\end{figure}
-Mathematically speaking, a polar space is a system, which maps each angle $\omega$ a magnitude $r(\omega)$ through a symbolic function.
-The conversion formula from polar (given a radius $r$ and angle $\omega$) to Cartesian coordinates is depicted in Fig.~\ref{polar} and is given by:
-\begin{align*}
-	x &= r \cos \omega, \\
- 	y &= r \sin \omega.
-\end{align*}
-Note, that we switched angle $\phi$ to $\omega$ for reasonable discrimination to the radiant intensity. The inverse direction (given coordinates $x,y$) is a bit more complicated:
-\begin{align*}
-	r &= \sqrt{x^2 + y^2}, \\
-\omega &= \operatorname{atan2}(y, x).
-\end{align*}
-The polar domain is a circular axis version of the domain $[0..2\pi]$, with negative magnitudes in $r$ shifted by 180 degrees (point reflected) as this is the formal behaviour defined and required by the conversion formulas.
-Remark that we exclude negative magnitudes, in lack of physical interpretation for negative energies.
-
-\section{Principal Component Analysis}
-\label{sec:pca}
-Note, that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following.
-Also keep in mind, that we interpret a matrix or tensor as an affine transformation.
-For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent.
-Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted.
-The anisotropy can be represented by, e.g., ellipsoid glyphs.
-
-\begin{figure}[!t]
-\hskip 22pt
-  \begin{minipage}[!t]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/HAT.png}
-   \caption{2D principal component ellipsoid\/ glyph (eigenframe)}
-   \label{HAT}
-   \captionsetup{font={footnotesize,bf,it}}
-   \caption*{source: \url{https://de.wikipedia.org/wiki/Datei:HAT-ellipse0.svg}}
-  \end{minipage} \hskip 40pt
-  %\raggedright
-  \begin{minipage}[!t]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/glyphs.png}
-    \caption{3D ellipsoid glyphs (l:linear, p:planar, s:spherical)}
-    \label{3Dglyphs}
-    \captionsetup{font={footnotesize,bf,it}}
-   \caption*{source: \url{https://slideplayer.com/slide/5291764/}}
-  \end{minipage}
-\end{figure}
-
-Principal component analysis is an algorithm, which is able to capture the principal components (directions) in stochastic data.
-But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling.
-Now, principal component analysis can be done by eigenvalue or singular value decomposition, whereas the former yields complex eigenvalues for asymmetric matrices.
-For (skew-) symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf.~\cite{kieburg}).
-Note, that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases (cf.~\cite{moler}).
-Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling (classically: $\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$).
-We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive).
-Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig.~\ref{HAT} for any kind of transformation matrix.
-The singular value decomposition is defined as follows:
-\begin{align*}
-	\mathbf{A} = \mathbf{U} \boldsymbol{\Sigma} \mathbf{V}
-\end{align*}
-where
-\begin{itemize}
-	\item $\mathbf{U}$ is an $m\times m$ unitary matrix
-	\item $\boldsymbol{\Sigma}$ is a diagonal matrix {$m \times n$} with positive real numbers on the diagonal
-	\item $\mathbf{V}$ is an {$n\times n$} unitary matrix, $\mathbf{V}^*$ is its conjugate transpose
-\end{itemize}
-
-In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues (\cite{mroz}) ordered decreasingly by absolute value (corresponding to singular value order).
-The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses.
-We do this decomposition analysis once for each matrix and then store the precomputed results in a grid.
-The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step, which makes it a basic and central concept within the scope of this work.
-Within this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition.
-This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor.
-We will then use global-illumination techniques on top to generate a light transport flow map, which permits us somehow to forward the intensities which are given as predefined initialization values in form of symbolic functions defined by the user.
-This type of light transport propagation scheme will also allow us to compute an FTLE-like field on the tensor fields.
-
-\section{Finite-Time Lyapunov Exponent}
-\label{sec:ftle}
-The Finite-Time Lyapunov Exponent is a measure, developed to analyze time-dependent dynamical systems for their separation abilities concerning massless tracer particles.
-We imagine a flow field, i.e., a vector field which in turn represents the effective amount or concentration shifted into a particular direction at each point in space.
-It could be induced by e.g.
-concentration or height gradients.
-This flow field is considered a time-dependent dynamical system, since it usually holds a time-component $t$ and thus maps each initial position an aged (altered) position: $\mathbf{\upvarphi}(\mathbf{x}) : \mathbf{x}(t_0) \longmapsto \mathbf{x}(t_0+T)$.
-We can now imagine placing two close by particles in distance $\delta$ into a flow field and run a time dependent simulation.
-Remark that this corresponds to the Lagrangian view in vector fields, which travels with the particles (instead of taking the position of a resting observer called Eulerian view).
-
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.6\textwidth]
-    {img/ftle_sep.pdf}
-  }
-  \caption{FTLE: separation of massless tracer particles}
-  \captionsetup{font={footnotesize,bf,it}}
-  \caption*{source: Skript Prof. Sadlo, Scientific Visualization SS2017, Heidelberg University}
-  \label{ftle_sep}
-\end{figure}
-
-In Fig. \ref{ftle_sep}, we observe particle trajectories diverging w.r.t. the central axis ending up in distance $\Delta$. The FTLE measure for this single example (point in space) is then defined as:
-
-\begin{align*}
-	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}.
-\end{align*}
-
-Alternatively, an FTLE can also be computed from a flow map's gradient, which is more convenient to compute in general:
-\begin{align*}
-	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{\upvarphi(x)}\|_2.
-\end{align*}
-whereas $\| A\| _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$.
-This will yield a map, which responds most where path lines diverge w.r.t.
-to a central axis.
-We consider an example which detects FTLE ridges inside of the cross-section of a double gyre.
-Note, that this example considers single trajectories diverging implicitly:
-\begin{figure}[!t]
-\hskip 22pt
-  \begin{minipage}[!t]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/diverge.png}
-	a)
-   \label{diverge}
-   \captionsetup{font={footnotesize,bf,it}}
-  \end{minipage} \hskip 30pt
-  %\raggedright
-  \begin{minipage}[!t]{0.4\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_gyre.png}
-	b)
-    \captionsetup{font={footnotesize,bf,it}}
-   \caption*{source: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/dle_incT.png}}
-  \end{minipage}
-  \caption{double gyre: a) diverging streamlines, b) FTLE field}
-  \label{ftle_gyre}
-  \caption*{source: \url{https://shaddenlab.berkeley.edu/uploads/LCS-tutorial/}}
-\end{figure}
-Fig.~\ref{ftle_gyre}a) depicts a scenario of diverging streamlines for a double gyre snapshot in time (yielding a time-independent dynamical system).
-Fig.~\ref{ftle_gyre}b) features a high ridge on the margins (borders) and in particular at the center of the domain.
-The FTLE approach considers this region of interest (ROI) to indicate the maximum relative local divergence of tensor field lines, by the approach.
-Ridges also represent Lagrangian coherent structures, which are found in nature in vortices of correlating flow behavior.
-We could imagine to introduce this process for a whole distribution of trajectories, which would also implement a FTLE for uncertain vector fields in a sense, that whole propability distributions of trajectories could be heeded.
-An overview about existing visualization techniques for uncertain data is given by Griete \& Schuhmann \cite{griethe}.
-We require that the trajectories can cross each other unaffectedly, i.e., the superposition principle is supposed to be respected for the magnitudes.
-But behold, is this behavior not somehow familiar?
-If we elaborately think this through, we can come up with the idea that this is the behavior of light propagation in vacuum.
-Light paths are physically allowed to cross each other with no influence exerted on each other at all.
-Consequently, we choose to model the distributions of trajectories as light intensity profiles.
-That is, we interpret the polar profiles as a propability density function (PDF) for transmitting particle trajectories heading into a certain direction.
-To put it into context, our main contribution method LTG uses the FTLE as an inspiration or template to design a similar approach for tensor field visualization.
-The main difference to the template FTLE is that we follow eigenvectors instead of flow vectors and that we use whole distributions (i.e., light distributions specifically) of trajectories, as a tensor field suggests whole distributions of directions, accordingly.
-This interpretation will be the connecting key to incorporate tensor fields into our light propagation scheme.
-
-\section{GT Datasets / Methods}
-\label{sec:gt}
-\subsection{Glyphs and Tensor Field Lines}
-%\begin{figure}[!t]
-%  \centering
-% {
-%    \includegraphics[width=0.4\textwidth]
-%    {img/tensorfieldlines.png}
-%  }
-%  \caption{tensor field lines: bold lines; ellipsoids: ellipsoid glyphs}
-%  \label{scheme}
-%\end{figure}
-Glyphs are derived as a basis for the weighting profiles of the tensors in the following propagation scheme.
-They were chosen because it is an easy-to-use (and understand) concept, which can also be interpreted nicely.
-Tensor field lines (the type of tensor streamline analog that does not introduce artificial intertia) were implemented for ground truth (GT) data generation purposes.
-We can again emphasize glyphs as a crucial, central concept within this work.
-In fact, these representations are used for the principal component analysis of direction vectors for the tensor field line extrapolation (reimplementing Hlawatsch's \cite{hlawatsch} approach) and also as transmission profiles for our invented light propagation scheme in a subsequent step.
-For example, when designing a FTLE-like field computation technique, we will need to know when tensor field lines diverge/converge as ground truth information.
-Examples for glyphs (ellipsoids) and tensor field lines (thin lines) are depicted in Fig.~\ref{test-fields}.
-
-\subsection{Test Data (Generation)}
-We used python's numpy library to generate predefined, synthetic test cases and caught two real data example from \url{TensorVis.org} and \url{https://www.cs.purdue.edu/homes/cs530/projects/project5.html}, which needed to be pre-processed through slicing and subsequent cropping to obtain 2D samples ($2\times2$) matrices).
-Later on, we will find that we need a noise correction, to process real data.
-These test fields can be grouped into symmetric and asymmetric ones:\\
-
-\subsubsection{Asymmetric Tensor Fields}
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[height=\textwidth]{img/corner-TFL.png}
-	a)``corner''
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[height=\textwidth]{img/spiral-TFL.png}
-	b)``spiral''
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[height=\textwidth]{img/drain_old.png}
-	c)``V''
-    \label{b)}
-  \end{minipage}
-\caption{Several test fields}
-\label{test-fields}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-      \centering
-    \includegraphics[height=\textwidth]{img/drain(alt)-TFL.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-      \centering
-    \includegraphics[height=\textwidth]{img/tensorfieldlines.png}
-	b)
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-      \centering
-    \includegraphics[height=\textwidth]{img/inverse-TFL.png}
-	c)
-    \label{b)}
-  \end{minipage}
-\caption{Several test fields: a)``corner'', ``spiral'', ``V''}
-\label{test_fields2}
-\end{figure}
-
-The ``drain''-Testfield, depicted in Fig.~\ref{test_fields2}a) is generated to have a kind of proof-of-concept for the FTLE-related approach.
-It poses raw, bare diverging tensor field lines without any other feature modulated on top.
-Thus, it can be expected to induce a high FTLE response, since it is an interpretable ground truth test and is considered to be a basic trigger/stimulus for the approach with no side effects introduced.
-The ``rings'' test field, depicted in Fig.~\ref{test_fields2} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
-The ``spiral'' test field, depicted in Fig.~\ref{test_fields2} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
-With the \enquote{inverse} test field we want to demonstrate the directive separation of intensities into four quadrants.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/gyre.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/gyre-TFL.png}
-	b)
-    \label{b)}
-  \end{minipage}
-\caption{``gyre'' test field: a) glyphs, b) tensor field lines for a)}
-\label{gyre}
-\end{figure}
-The test field ``gyre'', depicted in Fig \ref{gyre}, constitutes a classical textbook example for the FTLE and is considered for verifcation purposes.
-
-\subsubsection{Symmetric Tensor Fields}
-We also aim to use a real data example of the 3D DTI-MRI diffusion tensor of the brain for a male subject at the age of 27 and another one featuring a human heart.
-As we visualize 2D symmetric second order tensor fields, we need to slice the dataset and crop the $2{\times}2$ matrices omitting the $z$-component of the original $3{\times}3$ matrices.
-Fig.~\ref{real1} depicts the glyph representation of the test dataset for slice index $15$.
-To have some real examples involved, we apply the invented technique to both DTI-MRI \enquote{Brain} and \enquote{Heart} scans, which can be observed in Figs.~\ref{real1} and \ref{real2}.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/brain_org.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/brain_cyl_glyphs.png}
-    \label{b)}
-    b)
-  \end{minipage}
-\caption{\enquote{Brain} dataset: a) tensor magnitude, b) cyclinder glyphs}
-\label{real1}
-\end{figure}
-Observing the \enquote{brain} dataset in Figs.
-\ref{real1}, one may notice that it introduces a lot of variances to tensor magnitude and linear anisotropy.
-We choose this DTI-MRI scan as an example-par-excellence for real application data given in a reasonable resolution.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/heart_org.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-      \includegraphics[width=0.8\textwidth]{img/heart_cyl_glyphs.png}
-    \label{b)}
-	b)
-  \end{minipage}
-\caption{\enquote{Heart} dataset: a) tensor magnitude, b) cyclinder glyphs}
-\label{real2}
-\end{figure}
-When we look at this DT-MRI scan of a human heart, we can notice the elongated shape, which we consider to represent the aorta and a part of the rest of the heart.
-If we consider the glyph representation, we again infer, that we incorporate lots of anisotropic regions with varying tensor magnitudes here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-
-\chapter{Method}
-\label{chap:meth}
-
-This chapter will summarize the main contributions of this work studied and state targeted aims and goals as a first step in Sect.~\ref{sec:motiv}.
-Then we will have a short introduction to the implemented global illumination scheme, derived and simplified from a global illumination approach, invented for the Crytek GmbH in the context of global indirect illumination rendering, in Sect.~\ref{sec:scheme}.
-Last, we will state the effectively implemented tensor field visualization methods: light transport visualization (directed propagation scheme yielding $2D$ scalar fields) in sects.
-\ref{sec:scheme} \& \ref{sec:trans} and light transport gradient (LTG: yielding $3D$ scalar fields) in Sect.~\ref{sec:ltg}, which are proposed as the invented tensor visualization techniques and hence as conceptual contributions to the domain of tensor field visualization.
-
-%Dieses Kapitel stellt meist den Hauptteil der Arbeit dar. Vor dem
-%ersten Abschnitt sollte ein kurzer ?berblick (ein paar wenige S?tze
-%mit Verweise auf nachfolgende Abschnitte) gegeben werden. Beispiel: Im
-%nachfolgenden Abschnitt \ref{sec:overview} wir ein ?berblick ?ber die Anforderungen an
-%das Modell gegeben.
-
-\section{Motivation}
-\label{sec:spec} 
-Our ambition is to motivate a simple and efficient global illumination propagation scheme, able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t.
-energy conservation and propagation attenuation principles.
-We then generate transmission profiles, which we sample in a preliminary step, from the eigensystems (principal component ellipsoids) of the tensors for each cell in the domain (field).
-This step is done, to provide an orientation for a lattice of crystal fibre structures (cf.~optical fibres) to redirect the intensities in the grid in analogy to the anisotropy characteristics of the underlying tensor field within a glyph representation, obtained from PCA.
-Our approach should be implicitly adaptive to any kind of 2D dataset or slices of 3D datasets.
-Thus, we require it to be compatible to handle any kind of matrix (symmetric/anti-symmetric) and any kind of resolution of the field (fully-adaptive).
-At last, our ambition is to segment key locations (LCS/ridges) in the field with tensor field lines (hyperstreamlines) converging/diverging, which is necessarily the same (for bidirectional eigenvectors/singular vectors of tensors).
-Thus, we aim to generate a new FTLE-related method for tensor field visualization by applying a light transport gradient to the resulting light distributions for every possible position and direction of the light source in the grid (lattice) generated by stimulation with Dirac-pulses.
-Our approach should efficiently handle (relatively) large datasets with reasonable resolutions, which turns out to be possible with additional computing power empowered by a cluster.
-%Knapp zwei Seiten, in dem die Anforderungen, die Zielsetzung und die
-%Methoden ?berblicksartig beschrieben werden. Hier sollte die
-%Beschreibung ``technischer'' bzw.~``formaler'' sein als in der
-%Einleitung, da der Leser nun mit den Grundlagen und verwandten
-%Arbeiten vertraut ist.
-
-\section{Light Propagation Scheme}
-\label{sec:scheme}
-As already discussed, we will design a new method for visualizing tensor fields, which is based on the physical nature of light propagation.
-For this purpose, we motivate a light transport propagation scheme designed to efficiently propagate initial intensity profiles stored in polar coordinates given as symbolic impulses (stimuli) by the user defined in the configuration file.
-This polar profiles are first sampled into a vector with elements ranging over the interval $[0,2\pi]$.
-Polar profiles for different light sources are shown in Fig.~\ref{polar}.
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.8\textwidth]
-    {img/strahler.png}
-  }
-  \caption{Polar profiles for different types of emitters ($m=0$: spherical emitter, $m=1$: Lambert emitter, $m=2$: cone emitter}
-  \captionsetup{font={footnotesize,bf,it}}
-  \caption*{source: \url{http://www.optik.uni-erlangen.de/odem/vorlesung/ws1213/Photometrie_Geometrische_Optik.pdf}}
-  \label{polar}
-\end{figure}\vskip 3pt
-As a reminder, polar profiles map each angle $\omega$ a magnitude $r$ forming a function $r(\omega)$.
-They are circular axis versions of the domain $[0..2\pi]$, with negative magnitudes point reflected by 180 degrees.
-Thus, they are functions with an angular domain making them directionally dependent, which is a well suited representation for point lights (the type of light source with theoretically no finite extend), as we use them in a style of Dachsbacher et al.
-\cite{dachsbacher}.
-For the sake of interpretability, we exclude negative values (energies) from the scope of our calculations which we express through mathematical clip functions, limiting negative values to $0.0$, which we denote by subscript $+$ in the following.
-We will describe the emission characteristics of point lights exchanging energies in the lattice by parsing such polar functions.
-Our grid consists of $\mathop{dim}=\mathop{width}\cdot \mathop{height}$ virtual point lights (VPLs), which are in turn light sources with (theoretically) no finite extend represented through polar functions.
-As a start-up thinker for the propagation scheme, we will think of a grid cell in 2D as a center pixel within an 8-neighborhood with a previously injected point light (imagine a spherical emitter: $r(\omega)=1$ here for simplicity).
-Inside of this cell, we imagine a circular polar profile of radiant intensities for every angle $\omega$ given in $[\frac{W}{\mathop{rad}}]$.
-Remember, that all considerations from now on are given concerning a single grid cell respectively.
-Each cell has 8 (4 faces/edges and 4 diagonals) unique neighbors, which are adjacent point lights in the grid (initialized to $I(\omega)=0$).
-We consider one face neighbor (top) to shortly explain the steps of the scheme.
-The diagonals and the others are obtained from derivation and symmetry.
-Naively, one would consider to just transmit the intensity in the direction its pointing into.
-But this would omit every propagation attenuation principles (any principles at all specifically), if one takes a closer look.
-\subsection{Accumulation stage}
-Each cone type (yellow and green) is an individual neighbor-dependent angular area.
-The diagonals take in a greater angular domain, while they are located at greater distance and thus need to be attenuated more strongly (consider an isotropic stimulus).
-We aim for distributing contributions for overlapping cones to transmit attenuated (because of the larger factor $\sqrt{2}$ distance) intensity in form of a shared part to the diagonals in a single step (which also prevents from zigzagging diagonal intensities in case we would transmit diagonal intensities around the corner).
-Therefore, the intensity gathered in polar coordinates inside of angular ranges marked in green, needs to be partitioned between 2 of the 8 cone index neighbors with following considerations:
-
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.7\textwidth]
-    {img/geogebra-export_large.pdf}
-  }
-  \caption{Propagation Scheme}
-  \label{scheme}
-\end{figure}\vskip 3pt
-
-The face neighbors take in the angular domain of $d_1=\alpha+2\beta$ (yellow cone), whereas the diagonal neighbors (green cones) span a domain of $d_2=2\beta$.
-The linear combination weights for the overlapping (green) parts ($\beta$) are obtained from the relative overlapping angular area of the angular span $\beta$ w.r.t.
-diagonal cones ($2\beta$) and yellow face cones ($\alpha+2\beta = 90^\circ$) (yellow cone is overlapped by green cones at full angle).
-The linear combination weights are subsequently normalized to $\sum = 1.0$ (normalization condition):
-\begin{align}
-	\varepsilon_\alpha &= \frac{\frac{\beta}{\alpha+2\beta}}{\frac{\beta}{\alpha+2\beta} + \frac{\beta}{2\beta}} \approx 0.362291\\
-	\varepsilon_\beta &= \frac{\frac{\beta}{2\beta}}{\frac{\beta}{\alpha+2\beta} + \frac{\beta}{2\beta}} \approx 0.63771.
-\end{align}
-
-Remark, that we can precompute these linear combination weights and reutilize them in each iteration.
-We define the energetic contributions (presuming the cone is given in local coordinates - i.e.
-centered around the origin) as follows:
-\begin{align}
-	\Phi_{\alpha} &= \varepsilon_\alpha\int_\alpha {I}(\omega)\mathop{d\omega} =\varepsilon_\alpha(\int_{-\pi/4}^{-\alpha/2}{I}(\omega)\mathop{d\omega}+\int_{\alpha/2}^{\pi/4}{I}(\omega)\mathop{d\omega})+\int_{-\alpha/2}^{\alpha/2}{I}(\omega)\mathop{d\omega}\\
-	\Phi_{\beta} &= \varepsilon_\beta\int_\beta {I}(\omega)\mathop{d\omega} = \varepsilon_\beta\int_{-\beta}^{\beta}{I}(\omega)\mathop{d\omega}
-\end{align}
-This permits us to scale an oriented and clipped cosine-lobe $\cos_+$ with the total radiant flux with respect to its own total energy when integrated likewise.
-A discrete cosine lobe never reaches its theoretical sum of $2$ exactly in the interval $[{-\frac{\pi}{2}},{\frac{\pi}{2}}]$.
-Therefore it needs to be computed to account for sampling (discretization) errors and normalized by following relation (e.g., $\sum_{\cos} \approx 1.9949204635834517$).
-\begin{align}
-\sum_{\cos_+} &= \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\cos_+(\omega)\mathop{d\omega} \approx 2,\\
-	\cos_k(\omega) &= \frac{\Phi_t}{\sum_{\cos_+}}cos_+(\omega-k\cdot\frac{\pi}{4})\hskip 20pt \mathit{with} \ k\in[0,7].
-\end{align}
-\subsection{Injection stage}
- We place a cosine lobe scaled with the integrated (accumulated) energies in each of the 8 cone directions (angular domains), whereas outgoing integrated and summed energies are weighted by the linear combination weights (factors) in overlapping (green) areas to split up the contributions accordingly.
-The contribution areas are depicted in Fig.~\ref{scheme}, the cosine lobes can be observed in Fig.~\ref{iter}, respectively.
-This manner of propagation is inspired by the propagation scheme of Dachsbacher et al.
-\cite{dachsbacher} (see the paper on how to ``project the flux into a point light'').
-Basically, we sum up the weighted intensities corresponding to the respective neighbor index and scale an oriented cosine lobe (in direction of the neighbor) with the total flux $\Phi_t$ accumulated inside the neighbor range to transmit all of the parsed intensity for each neighbor distributed into a cosine distribution.
-Remark, that a cosine distribution physically corresponds to Lambert emitter charactersitics and hence to diffusely reflecting surfaces.
-We repeat this process once for each neighbor and each cell respectively.
-When we are done with the whole grid, we just recursively use the result and keep on running propagation cycles.
-The implementation is done in a dual buffer approach which pushes the energies back and forth (buffer A to buffer B) for each propagation cyvle until a state of convergence is reached, which is the case when the energy is spreaded throughout the grid and enters a stationary state of equilibrium, characterized by equal out (at grid borders) to in (at light source positions) energy-flow with no more nettings going on inside the domain.
-We measure the convergence by setting up an overall distribution error w.r.t.
-the previous iteration.
-Note, that even though we propagate the energy directly through the diagonal cones (which forms a square initially), the 8 degrees of freedom introduce a circular (spherical) wave front after few iterations (for greater resolutions), as depicted in Fig.~\ref{iter}.
-This occurs, because of a faster propagation attenuation which counteracts the larger angle of diagonal cones.
-There is a redistribution of intensities in the long term, since the shared part effects a uniformly spaced isotropic propagation profile resulting in a circular torus-shaped distribution of intensities, as observed in Fig.~\ref{iter}.
-Also note, that this approach satisfies energy conservation and propagation attenuation principles and applies to the principles of light propagation in vacuum (by default).
-We infer, that we can assume to follow physical models of spherical waves.
-The verification of these principal requirements and specifications can be observed in the evaluation in Chap.~\ref{chap:eval}.
-We first show the approach for a few iterations with a manual stop criterion to have a kind of imagination for the propagation process:
-
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=1.0\textwidth]
-    {img/steps_gallery_format_arrow_alt.png}
-  }
-  \caption{Impulse response for isotropic stimulus (few iteration snaphots: $t$)}
-  \label{iter}
-\end{figure}
-Fig \ref{iter} depicts the iteration indices $0,1,2,3,7,11$ (cycles) for the propagation scheme for a circular pulse stimulus which gets turned off after 1 iteration to demonstrate raw energy spreading in the field (impulse response of the propagation scheme).
-In step 1 one can observe the placing of cosine lobes scaled with the respective accumulated and weighted energy for the summed area in each of the 8 neighbor directions.
-In the subsequent steps, remark the evolution from an 8-neighborhood in step 1 to a circular shaped torus.
-
-As already mentioned, a dual-buffer approach is chosen here, as we can efficiently propagate from one into the other and just swap pointers to start all over (after reinitialization) until convergence.
-Another reason is it yields a reasonable source-target or read-write segmentation.
-We introduce a new operator called \textit{propagate}, permitting us to predefine the propagation in a symbolic manner.
-The propagation definition denotes the transformation from one grid (A-buffer) into another (B-buffer) through proper application of the \textit{propagate}-operator to each cell.
-We will extend this operator with a normalization correction in Sect.~\ref{sec:trans} in form of a precise mathematical formulation, respectively.
-For now, just imagine the approach in raw operating mode, with no tensor influence incorporated.
-That is, evaluating and accumulating the polar profiles, weighting with the linear combination weights and scaling and placing of a cosine lobe to the respective neighbor.
-For now, we want to define our operator $\mathop{propagate}$: Each Grid or Buffer is basically a set of cells defined as $\mathcal{C} = \{\mathbf{c}_1, \mathbf{c}_2, ..., \mathbf{c}_i\}$.
-The dual buffer result $\mathcal{C'}$ is denoted symbolically as follows for all cells with a mean intensity greater than zero (those which do not contain trivial zero samples):
-\begin{align}
-	\mathcal{C'} = \mathop{propagate}\{c_i\in \mathcal{C} \ \mid \ |\mean{I}_i| > 0\} \hskip 5pt {\forall } \ i \in [1,\mathop{dim}],
-\end{align}
-whereas: $i$: cell index, $I$: source buffer radiant intensity.\vskip 5pt
-The dual buffer result $\mathcal{C'}$ denotes the set of all transformed cells after the propagation, i.e., after applying the operator $\mathop{propagate}$ to each cell.
-This result is generated once in each iteration and conceptually gets pushed back and forth between the two buffers.
-Note that, $\mathop{dim}=width\cdot height$, which equals the tensor field resolution in this case.
-At last, we define a threshold for the minimum overall distribution error for the total radiant flux taken into account for a convergence criterion, initiating a stop sequence when falling below that threshold $\epsilon$ (, e.g., $0.5$) :\\
-\paragraph{Criterion:}
-\begin{align*}
-	\Delta \Phi_{total} &= \sum_{c_i\in\mathcal{C}}\lvert\Delta I_i(\omega)\rvert = \sum_{c_i\in\mathcal{C}} \int_0^{2\pi} \lvert(I'_i(\omega)-I_i(\omega))\rvert \mathop{d\omega}  \overset{!}{<} \epsilon.\\
-\end{align*}
-where $\phi_{total}$: total radiant flux, $I'$: target buffer radiant intensity\vskip 5pt
-Note that we take the absolute value of the differences to prevent from mutual compensation.
-Until now, we aimed to simulate the propagation of light in empty space (vacuum) providing us with a physically-motivated but very simplified base-approach.
-We aim to modulate the transmission of the intensities on top of the base approach with transmission profiles obtained from the tensor fields through principal component analysis, which yields ellipsoid glyph equations in the following.
-
-%In diesem und den nachfolgenden Abschnitten werden die Beitr?ge der
-%Arbeit motiviert, formal sauber (oft mathatisch, sprich mit
-%Definitionen etc.) beschrieben, und bei Bedarf mithilfe von Beispielen
-%verdeutlicht. Die Beschreibungen in diesem Kapitel sind meist
-%unabh?ngig von einer konkreten Realisierung und Daten; diese werden im
-%nachfolgenden Kapitel detailliert.
-
-\section{Transmission Profiles and Weighting Functions}
-\label{sec:trans}
-Note that the following considerations are only given for the tensor profile operating mode with tensor influence incorporated and now the tensor properties are modeled on top.
-To obtain a kind of footprint of the tensor field, we set up ellipse equations for each glyph representation, obtained from applying PCA on each cell (tensor) in the grid.
-Concurrently, we extend our abstract mathematical operator \textit{propagate}, previously defined in Sect.
-\ref{sec:scheme}, by a normalization strategy necessary for windowing weighting profiles, as it is the case for our elliptical transmission profiles, which we will capture in the following.
-At this point, we presume the precomputation of singular value decomposition (dependent of each cell) and linear combination weights (independent of cells).
-We map the singular values (corresponding to eigenvalues $\lambda$ here for simplicity) in decreasing order to the ellipses radii (half-axes) $a(\Delta x)$ and $b(\Delta y)$ as follows:
-
-Ellipse Equation:
-\begin{align}
-	r(\omega) = \frac{ab}{\sqrt{a^{2}\sin^{2}(\omega-\varphi)+b^{2}\cos^{2}(\omega-\varphi)}} =\frac{\lambda_1\lambda_2}{\sqrt{\lambda_1^2\sin^2(\omega-\varphi)+\lambda_2^2\cos^2(\omega-\varphi)}}.
-\end{align}
-That is, the singular values form the half-axes of the PCA ellipsoid, which yields us a symbolic definition for our transmission (transfer) functions as weighting profiles which we can now evaluate as a parsing and accumulating step.
-We compute the offset angle $\varphi$ w.r.t.
-the x-axis by exploiting the $\mathop{atan2}$-function with 4-quadrant evaluation for the first singular vector (ordered decreasingly in value), since we scale the $x$-axis with the first singular value.
-As a preliminary preparation, this profile is precomputed (sampled) for all discrete angular steps for each cell (e.g., $360$) and stored for reuitilization, as it is the case for the previously mentioned linear combination weights, as well.
-The weighting scheme, which we apply is depicted in Fig.~\ref{weighting}.
-\begin{figure}[!t]
-  \centering
-  {
-    \includegraphics[width=0.5\textwidth]
-    {img/polarplot.pdf}
-  }
-  \caption{Red: Intensity Profile, cyan: Transmission Profile (glyph ellipsoid eq.)}
-  \label{weighting}
-\end{figure}
-We interpret the polar functions as discrete, sampled 1D vectors with a uniform angular resolution.
-That means, we can use element-wise multiplication for the 1D-vectors in red (Intensity Profile) and cyan (Transmission Profile) to obtain a windowed version of the intensity profile and perform the following integration step to accumulate the total radiant power (flux) as follows.
-Note that we model angular dependent radiant intensity ${I}(\omega) [\frac{W}{\mathop{rad}}]$ as an underlying physical entity, which is well suited for describing point light sources.
-As a side note, the physical quantity perceived as brightness by human eyes, is the radiance (luminance), which is yielded by relating radiant intensity to a finite sensitive area ($[\frac{W}{\mathop{rad}}/m^2]$).
-The primary definition for our total accumulated incident radiant flux (again presuming that the cone is given in local coordinates - i.e.
-centered around the origin) is then denoted as:
-
-\begin{align}
-	\Phi_{\alpha} &= \varepsilon_\alpha(\int_{-\pi/4}^{-\alpha/2}{T}(\omega){I}(\omega)\mathop{d\omega}+\int_{\alpha/2}^{\pi/4}{T}(\omega){I}(\omega)\mathop{d\omega})+\int_{-\alpha/2}^{\alpha/2}{T}(\omega){I}(\omega)\mathop{d\omega},\\
-	\Phi_{\beta} &= \varepsilon_\beta\int_{-\beta}^{\beta}{T}(\omega){I}(\omega)\mathop{d\omega}.
-\end{align}
-
-This specific profile is injected in form of a placement to the corresponding neighbor index $k$.
-It yields a cosine lobe, which has an overall energy of $\Phi_t$, if the point light is interpreted with a Lambert-emitter (cf.~\cite{lindlein}) characteristic (as part or area element of a diffusely reflecting surface) since the intensity profile matches a cosine in this case.
-This emitter characteristic is the case for global illumination diffuse indirect illumination (non-specular).
-It is used in similar style by Dachsbacher et al.
-for propagating intensities through LPVs (light propagation volumes).
-But caution, we should not generate any exceeding energy right?
-Each tensor ellipse equation features a mean transmission rate $[\%]$, which turns out to be a measure or synonym for the total rate of the outgoing transmitted radiant intensity $\int_0^{2\pi}{T}(\omega)\mathop{d\omega}$ $[\%]$.
-Since energy can not be generated from transmission profiles (only transmitting, not generating), we have to normalize every tensor ellipse equation with the highest mean transmission rate in the field, which is then transmitting the whole outgoing intensity $\int_0^{2\pi}{I}(\omega)\mathop{d\omega}$ in total.
-This restricts the tensor ellipse mean to a maximum of $100\%$ for the tensor with the highest tensor magnitude.
-We will use the self-made definition of mean tensor magnitude $\mean{r}_{max}$ obtained from the PCA here for simplicity.
-That is, we take the principal component ellipsoid and calculate its analytical mean, which is then used for mean normalization to $\mean{r}_{max}=1.0$.
-Hence the tensor ellipse equations are normalized through the following step:
-\begin{align}
-	r_i(\omega) = \frac{1}{\mean{r}_{max}}r(\omega).
-\end{align}
-Notice, that this implements the concept of absorption implicitly, because tensors with a lower mean $\mean{r}_i$ than $100\%$ absorb energy wheras absorption equals ${\mu} = {1}-{\mean{r}}$ here.
-Now, that we have coped with the problem of overall transmission, we are left with another problem that occurs with the weighting functions as they can still yield values above $1.0$.
-Remark, that the next example could be interpreted congruently to Fig.\ref{weighting}.
-Consider following examples as $T(\omega)I(\omega)$ pairs (obviously normalized to $\mean{r}=1.0$): $1.5\cdot 1.5=2.25$ and $0.5\cdot 0.5=0.25$, which already disproves us that a mean transmission rate $\mean{r}=1.0$ necessarily effects a mean transmission intensity $I_t$ of $1.0<1.25=\frac{2.5}{2}$ and thus to an overall transmitted intensity of $100\%$.
-This would potentially generate intensity out of nothing, which leads to crucially unstable behaviour, since we would employ uncontrolled sources of energy in this case.
-Well, but only in case energy conservation principles were not respected.
-To construct a solution, we ask for the source of the excessive energy.
-If the tensor mean complies with the constraint $\mean{r}\leq 1.0$, then the energy must lack in some other place (since the overall transmission fulfills the energy conservation constraint in that case).
-The only problem is that a normalization condition in usual style for the transmission profile ${T}(\omega)$ does not necessarily imply the same normalization for the term ${T}(\omega){I}(\omega)$.
-However, a proper normalization can provide corrective support, wich reveals another subsequent normalization constraint.
-The normalization factor is defined in three subsequent steps:
-\begin{enumerate}
-	\item Normalization of $T(\omega)I(\omega)$ to $\overline{TI}=1.0$
-	\item Subsequent scaling with mean intensity $\mean{I}$ for energy conservation principles
-	\item Subsequent scaling with mean tensor magnitude $\mean{T}$ for absorption principles
-\end{enumerate}
-
-Hence the transmitted outgoing intensity for a single direction is defined as:
-
-\begin{align}
-    {I}_t(\omega) = 
-\begin{cases}
-   n_f\varepsilon_\alpha{T}(\omega){I}(\omega),& \text{if } k\bmod 2 = 0\\
-     n_f\varepsilon_\beta{T}(\omega){I}(\omega),              & \text{otherwise}.
-\end{cases}
-\end{align}
-
-This transforms our equation for the transmitted radiant flux into:
-\begin{align}
- \Phi_{t,i} = 
-\begin{cases}
-   \int_\alpha{I}_{t,i}(\omega)\mathop{d\omega} = n_f\Phi_{\alpha,i},& \text{if } k\bmod 2 = 0\\
-     \int_\beta{I}_{t,i}(\omega)\mathop{d\omega} = n_f\Phi_{\beta,i},              & \text{otherwise}.
-\end{cases}
-\end{align}
-Which is now respecting energy conservation principles very conveniently, as all we need is a set of means and we are capable of normalizing the weighted intensity profile transmitted.
-This collapses our abstract mathematical operator $\mathop{propagate}$, which is applied to each cell, into the following steps:
-\begin{enumerate}
-	\item Computation of normalization factor $n_f$
-	\item Direction (component)-wise weighting to account for shared part in diagonal cones
-	\item Integration (accumulation) of total radiant flux weighted with \underline{normalized} transmission profiles inside the angular neighbor range (read-access)
-	\item Scaling of cosine lobe corresponding to neighbor direction and subsequent placing (injection) to corresponding neighbor (write-access)
-
-\end{enumerate}
-Remember, that this operator needs to be applied once to each cell to form a whole propagation cycle propagating intensities from source to target buffer.
-This approach ensures, that no energy will be generated and therefore implements the physical law of energy conservation or Kirchhoff's current law if you want to think of a electronic network.
-Consequently, this allows us to assume that all sources of power (energy) are user-defined.
-In fact the approach, has proven to be robust w.r.t.
-energy losses for several iterations as depicted in Chap.~\ref{chap:eval} in the stochastic error graph.
-
-\section{Physical Model}
-\label{sec:model}
-To constitute a physical interpretation for the transmission of the tensor field, we imagine a melting crystal with  fibrous structures aligning with the major eigenvectors of the tensor field.
-This is what we call the footprint of the tensor field.
-In that sense of imagination, we would virtually imprint every tensor field into a crystalline structure as a preliminary initialization step.
-For an every day example, this can be observed in, e.g., the gemstone tiger's eye's cat's eye effect, where crystal fibers align with the crystal axes respectively and effect a brightened line which never changes position or orientation when rotating the crystal.
-These oriented crysal fibers are then representing the biased propagation directions of light inside of the crystal lattice.
-The propability density function $P(\omega)$, for a photon being scattered (redistributed) in a particular direction $\omega$ \cite{ryder}, is then described as:
-\begin{align}
- 	P(\omega) &= \frac{T(\omega)}{\int_{k\pi}^{(k+1)\pi}T(\omega)\mathop{d\omega}}\\
- 	\mathit{with} \int_0^{\pi} P(\omega) &= 1.
-\end{align}
-This propability density function (PDF) indicates, whether there is directed anisotropy or spherical isotropy w.r.t the neighboring $180$ ($\pi \mathop{rad}$) degrees and how likely the spontaneuos scattering of a photon in this particular direction will happen.
-Each tensor is then a unique footprint which can be precomputed for convenience and described as polar profile just as our previously mentioned intensity profiles.
-We apply these polar profiles in a sense of transmission profiles as previously discussed in Sect.~\ref{sec:polar}.
-
-\section{Light Distributions}
-To evaluate the plausibility of the light distribution results, obtained from the light propagation scheme, we plot the transmission profiles overlaid with the resulting intensity profiles for several synthetic test fields, whereas red: tensor glyphs (transmission profiles) and blue: intensity profiles are both depicted in polar coordinates:
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/iso_sat.png}
-    \label{a)}
-	a)~light source in top center
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.8\textwidth]{img/isotropic-all-center.png}
-    \label{b)}
-	b)~light source in center
-  \end{minipage}
-\caption{final light propagation distributions for ``isotropy''-testfield}
-\label{isotropic}
-\end{figure}
-One can observe in Fig.~\ref{isotropic}, how light is propagated with no directional bias leading to an absolutely isotropic distribution in case  b), which is symmetric as expected.
-This result is considered to be the absolute ground truth example for the light propagation scheme, which is evaluated separately in Sects.~\ref{sec:att} and \ref{sect:conserve}.
-We can also depict the final distribution as a scalar field when using the total energy in each cell as scalar:
-
-When we observe these polar and scalar heat map plots in Fig.~\ref{isotropic}, we first notice the red dots, which always represent light sources (since they are captured the most energy) in these cases.
-Specifically, the heat maps are obtained from the same light distributions we plot as polar profiles and should hence be interpreted concurrently.
-Remark the propagation front indicating a circular wave (circular isolines in a)).
-In this next example, we use the tensor field ``rings'' as input data for the approach and observe how the intensities are channeled in a circular orbit around the central origin.
-The anisotropy measures decide whether the transmission profile is more narrow (radical) or more wide (liberal):
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/rings-two-special1.png}
-    \label{a)}
-	a)~antisymmetric light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/cos_parallel.png}
-    \label{b)}
-	b)~symmetric light sources
-  \end{minipage}
-\caption{final light propagation distributions for ``rings''-test field}
-\label{rings-tests}
-\end{figure}
-
-We consider Fig.~\ref{rings-tests} to regard the behavior for circular tensor field lines.
-As expected, the intensities get directed in circular orbits by the tensor ellipsoid transmission profiles.
-Note the symmetry for both cases depicted.
-Again, we plot the final light distributions as heat maps: It is noticeable in Fig.
-\ref{rings-tests}, that the intensities concentrate preferably on the circular tensor field lines, which proves the approach's controlling ability expressing as direction.
- The results for our tensor field ``spiral'' can be interpreted in analogous fashion as the previously discussed ``rings''-testfield in Fig.
-\ref{spiral-tests}.
-The influence incorporated by the tensor field is easily noticeable in this case as well.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral-two-dense.png}
-    \label{a)}
-	a)~two light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral-two-wide.png}
-    \label{b)}
-	b)~symmetric light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral-temp2.png}
-    \label{a)}
-	c)~two light sources
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.9\textwidth]{img/spiral_full.png}
-    \label{b)}
-	d)~four light sources
-  \end{minipage}
-\caption{final light propagation distributions for ``spiral''-test field}
-\label{spiral-tests}
-\end{figure}
-In Fig.
-\ref{spiral-tests}a)\& b) we observe the spiral shape literally absorbing the intensities into its spiral arms, which becomes even more clear when we again observe the heat map representation in Fig.
-\ref{spiral-tests}c)\& d).
-
-The star example reveals another interesting feature of the approach in Fig.
-\ref{star-tests1} if the intensity impulsively hits an isotropic region, it gets directed according to its intensity profile (neglecting the transmission profile) which implements the principle of inertia for directed light propagation present in vacuum (light never changes its direction).
-Considering Fig.
-\ref{spiral-tests} we remark, that the diagonal intensities are attenuated more strongly by default, which introduces a bias into the light paths which are chosen along the axes with higher propability, effectively.
-This gets even clearer, when we plot the heat map depiction in Fig.
-\ref{star-tests2} again.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.31\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img/star-0,0.png}
-    \label{a)}
-	a)~single light source
-  \end{minipage}
-  \begin{minipage}{0.31\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img/star-two-wide.png}
-    \label{b)}
-	b)~symmetric light sources
-  \end{minipage}
-   \begin{minipage}{0.31\textwidth}
-    \includegraphics[width=\textwidth]{img/star-two-dense.png}
-    \label{c)}
-	c)~dual light sources
-  \end{minipage}
-\caption{final light propagation distributions for ``star'' test field}
-\label{star-tests1}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.31\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img/star_wide.png}
-    \label{a)}
-	a)~symmetric light sources
-  \end{minipage}
-  \begin{minipage}{0.31\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img/star_close.png}
-    \label{b)}
-	b)~dual light sources
-  \end{minipage}
-\caption{final light propagation distributions for ``star'' test field}
-\label{star-tests2}
-\end{figure}
-Note that we skipped the single light source example, since we consider it a trivial add-on example. We can regard the intensities condensed inside of a narrow beam of light.
-Note, that a logarithmic scale needs to be incorporated to produce results on a large scale in a sense of higher spatial resolution (e.g.: $101\times 101$), since the propagation attenuation follows the distance to the origin inversely proportional.
-
-\section{Light Transport Gradient}
-In this section, we will motivate the main contribution on tensor field analysis in this work.
-The approach called Light Transport Gradient (LTG) states a FTLE-related approach performed on tensor fields, but employing whole distributions of trajectories interpreted as light transmission profiles.
-Since we incorporate whole distributions of trajectories, we do not strictly follow the major eigenvectors like Hlawatsch's et al.
-\cite{hlawatsch} do with their FTLE on single trajectories following tensor field lines.
-Instead, we follow the directional distribution defined by the tensor field's principal ellipsoids for different input angles, which yields a FTLE volume, congruently depending on the direction, in general.
-In place of a 2D FTLE field concerning only the direction of the major eigenvector in each cell, we obtain all the directionally dependent FTLEs unrolled into slices and hence transform the FTLE field into a function FTLE$(\omega\mapsto z)$.
-The resulting volume contains a manifold which is not self-overlapping and is a continous plane for a continous tensor field.
-\label{sec:ltg}
-\subsection{Evolution}
-Now, that we are able to propagate intensity in grid, directed by the eigenvectors of the tensor field, we aim to do something more elaborate to segment ROIs characterised by divergence/convergence of tensor field lines and drastic changes in anisotropy.
-For this purpose, we imagine kind of a local and spatial gradient, as we place light sources in two close-by positions and eventually in two close-by directions and analyze the resulting final light distribution.
-We consider the light distributions and their difference image for the \enquote{isotropic} and the \enquote{drain} test field to have an idea, for which kind or types of cases/configurations we obtain inherently different light distributions, in particular.
-We choose to use these test cases, because they constitute absolutely pure examples par excellence for a FTLE computation exposing no (\enquote{isotropic}), or on the other hand raw (\enquote{drain}) tensor field lines diverging.
-We will start off with the \enquote{isotropic} field from here to first analyze the offset response for no divergence or more specifically, no anisotropy at all.
-Fig.~\ref{analysis1} depicts final light distributions (cf.~Sect.~\ref{sec:ltg} Fig.~\ref{rings-tests}) for placing the light source one step to the left and one to the right from the central axis (spatial offset).
-We can observe the light cones cancel out each other in the center of the difference image.
-That is, per default, light distributions cancel each other out for two close-by light source positions or directions (or more specifically: for continous overlaps in their distributions).
-In this case, for the \enquote{isotropic} test field, we have no tensor field lines diverging leading to a small overall difference for light sources in finite distance.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img/scalar_iso_left.png}
-    \label{a)}
-	a) final light distribution for left spatial offset
-  \end{minipage} $-$
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img/scalar_iso_right.png}
-    \label{b)}
-	b) final light distribution for right spatial offset
-  \end{minipage} $=$
-   \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img/scalar_iso_diff.png}
-    \label{b)}
-	difference image for a) and b)
-  \end{minipage}
-  \caption{``isotropic'' test field}
-\label{analysis1}
-\end{figure}
-
-The complementing idea for the angular domain would be: placing the light source at the same position but pointing in differing directions, which is depicted in Fig \ref{analysis2}.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img/scalar_iso_leftangle.png}
-    \label{a)}
-	a) final light distribution for left angular offset
-  \end{minipage} $-$
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img/scalar_iso_rightangle.png}
-    \label{b)}
-	b) final light distribution for right angular offset
-  \end{minipage} $=$
-   \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img/scalar_iso_anglediff.png}
-    \label{b)}
-	difference image for a) and b)
-  \end{minipage}
-  \caption{``isotropic'' test field}
-\label{analysis2}
-\end{figure}
-For the angular offset, we can again observe the light distributions cancelling each other out (per default) in the center (middle) of the difference image.
-For now, we can take home from these examples, that light distributions are cancelling each other out within their overlap in general, in case no anisotropy was introduced into the test tensor field.
-We again analyze the drain test field (cf.~Fig.~\ref{drain} in Sect.~\ref{chap:grundlagen}) for the behavior of the resulting light distributions in scalar field representation.
-This following example illustrated in Fig.~\ref{evolution1} now introduces raw tensor field lines diverging w.r.t.
-the $y$-axis.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_left.png}
-    \label{a)}
-	a) final light distribution for left spatial offset
-  \end{minipage} $-$
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_right.png}
-    \label{b)}
-	b) final light distribution for right spatial offset
-  \end{minipage} $=$
-   \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_diff.png}
-    \label{b)}
-	difference image for a) and b)
-  \end{minipage}
-  \caption{``drain'' test field}
-\label{evolution1}
-\end{figure}
-We can see the light distributions diverge following tensor field lines, which in turn form these curves depicted in Fig.~\ref{evolution1}.
-The separation of path (tensor field) lines and thus of light distributions is maximized at the center location of the field.
-We can now examine the light paths being dragged into long curves, which prevents the light distributions from overlapping at all.
-Hence, this test case will lead to larger differences and thus to an increased response of the FTLE field, which we will define in this section.
-%Imagine accumulating each pixel in the images to obtain the cumulative total energy in the (difference) grid, which could now be compared for each light source position and direction in terms of proper min-max scaling.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_leftangle.png}
-    \label{a)}
-	a) final light distribution for left angular offset
-  \end{minipage} $-$
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_rightangle.png}
-    \label{b)}
-	b) final light distribution for right angular offset
-  \end{minipage} $=$
-   \begin{minipage}{0.3\textwidth}
-    \includegraphics[width=\textwidth]{img/ftle_anglediff.png}
-    \label{b)}
-	difference image for a) and b)
-  \end{minipage}
-  \caption{``drain'' test field: }
-\label{evolution2}
-\end{figure}
-For the angle difference in Fig.~\ref{evolution2}, we can observe that the difference is again increased, because it does not overlap leading to a maximized difference.
-Imagine accumulating each pixel in the images to obtain the cumulative total energy in the (difference) grid, which could now be compared for each light source position and direction in terms of proper min-max scaling.
-Now imagine we would accumulate these differences and assign each position a gradient magnitude in form of a scalar value with the summed difference of the final light distributions in the neighborhood.
-What would we eventually be yielded with.
-Well, we would obtain another scalar field considered to be the gradient image of the light distributions.
-We could very well extend this process for every light source direction discretized (sampled) and obtain a $3D$ volume representing the topological structure of the tensor field for every direction.
-All we need, is a formal definition for a central differences gradient approximation for light distributions given in polar profiles.
-The main difference to FTLE in vector field visualization is, that we use whole distributions of trajectories and hence need to use a starting direction for the distribution of trajectories, instead of just moving with the flow, as a tensor introduces a whole distribution of directions and does not only depend on the inital position, but also on the direction.
-This corresponding to the volume emerging from the FTLE plane image, as we introduce a new dimension $\omega$.
-Remark that we denote the final buffer result as $\mathbf{b}=\mathcal{C}_{final}$ in the following.
-\subsection{Definition}
- As tensor field lines are bidirectional, convergence and divergence can not be reasonably discerned in that case.
-We aim to extract ridges from a FTLE-like field to grasp the key regions separating anisotropy behavior (, i.e., changes in direction/tensor magnitude) in the field segmented as height ridges in the field, which are in turn LCS (Lagrangian Coherent Structures) well known for their rich substance in physical meaning.
-They often indicate transport barriers in flow fields.
-The nature and location of these structures and their interaction with each other is subject to the study of both turbulent and laminar mixing.
-They can even be observed in nature for, e.g., vortices in flow fields.
-For this purpose, we will use our, previously defined, $\mathop{propagate}$-operator, i.e., the propagation scheme from Sect.~\ref{sec:scheme} and \ref{sec:trans}, on every possible light source location $\mathbf{x}$ and direction $\omega$ in the grid to obtain the final set of buffers $\mathcal{B} = \{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_j\}$ of impulse responses of the algorithm.
-That is, we initialize the grid with every possible light source configuration and compute a whole final light distribution for each step:
-\begin{align}
-	\mathcal{B} = \mathop{propagate}\{c_i\in \mathcal{C} \ \mid \ |\mean{I}_i| > 0\} \ {\forall } \ i \in [1,\mathop{dim}] \ {\forall } \ j \in [1,\mathop{dim}\cdot \mathop{steps}].
-	\label{ftle}
-\end{align}
-Whereas $j$ is the index of light source position and direction representing the initialization in each iteration and $\mathop{steps}$ is the count of the discrete angular steps.
-The initialization is done by placing a light source (delta-impulse) with $\mean{I}_i = 1.0$ at the exact location of index $j$ realizing a light source at location $(x\mid y)$ in direction $\omega$.
-Eq.
-\ref{ftle} now denotes the set of buffers $\mathcal{B}$ containing light distributions propagated from, i.e., computed for every possible direction and location.
-Now, that we have measured impulse responses from every location in our long $1D$ sample vector, we can incorporate finite (central) differences to obtain the gradient of the set of final buffers with final light distributions as elements $\mathcal{C}_{final}=\mathbf{b}\in\mathcal{B}$.
-Note, that we interpret a finite set $\mathcal{C}_{final}$ as a vector of cells $\mathbf{b}$ with concurrent indices here:
-\begin{align}
-	\nabla \mathbf{b}_{x,y,\omega} =
-	\begin{pmatrix}
-	\frac{\partial \mathbf{b}_{x,y,\omega}}{\partial x} \\ \frac{\partial \mathbf{b}_{x,y,\omega}}{\partial y} \\ \frac{\partial \mathbf{b}_{x,y,\omega}}{\partial \omega}
-	\end{pmatrix} \approx
-	\frac{1}{2}
-	\begin{pmatrix}
-	\sum_{\mathbf{c}}\lvert \mathbf{b}_{x+1,y,\omega} - \mathbf{b}_{x-1,y,\omega}\rvert \\ \sum_{\mathbf{c}}\lvert \mathbf{b}_{x,y+1,\omega} - \mathbf{b}_{x,y-1,\omega}\rvert \\ \sum_{\mathbf{c}}\lvert \mathbf{b}_{x,y,\omega+\pi/6} - \mathbf{b}_{x,y,\omega-\pi/6}\rvert
-	\end{pmatrix} =
-	\frac{1}{2}
-	\begin{pmatrix}
-	\sum_{\mathbf{c}\in \mathbf{b_x}}\lvert \mathbf{b_x}\rvert \\ \sum_{\mathbf{c}\in \mathbf{b_y}}\lvert\mathbf{b_{y}}\rvert \\ \sum_{\mathbf{c}\in \mathbf{b_{\omega}}}\lvert \mathbf{b_{\omega}}\rvert
-	\end{pmatrix}.
-\end{align}
- We choose to use a constant offset of $\frac{\pi}{6}\mathop{rad}$ for the angle because we want to prevent from generating results with content implicitly depending on the angular resolution and also since we need to exceed $\beta$ to have a uniform and consistent placing of cosine lobes (inside diagonal cones) concerning neighbor indices.
-Remark that the results do not depend on the spatial resolution implicitly, because increasement acts as a domain extension for and does effectively not increase the sampling rate of the propagation scheme.
-Also note that this step is very expensive considering runtime-costs since it follows the following runtime observation.
-We have $\mathop{width}\cdot \mathop{height}\cdot \mathop{steps}$ numbers in the grid and on top $\mathop{width}\cdot \mathop{height}\cdot \mathop{steps}$ buffers in total, which need to be processed.
-Also we remark the convergence criterion, which introduces another heuristic factor $\mathop{width}$, because the straight line distance to the field edges is a measure for eventual compliance, and increases with $d\sim width$ for any location in the grid.
-We set $\mathop{height}=\mathop{width}$ here for simplicity, which yields:
-\begin{align}
-	T(n)=\mathcal{O}(\mathop{width}\mathop{width}\mathop{steps}\mathop{width}\mathop{width}\mathop{steps}\mathop{width}) = \mathcal{O}(\mathop{width^5}\mathop{steps^2}),
-\end{align}
-which is the runtime behavior in order notation considering memory consumption and iteration step count.
-It is noticeable that the runtime increases fast with increasing resolution, which makes parallelization algorithms a crucial tool necessary to compute the results in reasonable time with additional computing power provided by a computer cluster.
-Next, we compute the Euclidean norm of the light transport gradient yielding a scalar field, which can be visualized directly or via extracted ridges in a visualization framework like ParaView from VTK format:
-\begin{align}
-	\lvert\nabla \mathbf{b}_{x,y,\omega}\rvert = \sqrt{(\frac{\partial \mathbf{b}_{x,y,\omega}}{\partial x})^2 + (\frac{\partial \mathbf{b}_{x,y,\omega}}{\partial y})^2 + (\frac{\partial \mathbf{b}_{x,y,\omega}}{\partial \omega})^2}.
-\end{align}
-Note that the resulting scalar field is $(n+1)$-dimensional ($x,y$ and direction $\omega\mapsto  z$) and needs to be flattened by averaging or projection in 3D for proper visualization in case the approach will be adopted in $3D$.
-We note, that our invented technique stands out compared to the already existent FTLE on tensor fields invented by Hlawatsch et al.
-\cite{hlawatsch}, as we use whole distributions of directions instead of following the major eigenvector strictly.
-These directional distributions allow for a complete analysis of the tensor field accounting for the incoming direction, as well.
-
-\subsection{Functionality}
-Remark, that the $z$-axis represents the $\omega$-axis for each test and example and that each data range is rescaled to its interval for proper color coding in heat maps of the resulting $3D$ volume.
-We will try to show the functionality of the method in a few simple examples.
-At first, we will have a null example to show the offset response of the approach induced by total spherical anisotropy and no absorption in the field.
-We consider the example \enquote{isotropic}, concerned to have no influence on the light propagation, as it models homogeneous identity matrices inducing absolutely isotropic transmission (every incident distribution is transmitted as is with no transmission weighting incorporated).
-This corresponds to the \enquote{non - tensor field} mode of the propagation scheme having no influence of any tensor field incorporated.
-In this configuration, we could eventually visualize light transport paths in a $2D$ topological scene.
-
-\begin{figure}[t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img/isotropic2.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[height=\textwidth]{img/iso_ftle.png}
-    \label{b)}
-	b)
-  \end{minipage}
-  \caption{\enquote{isotropic} test field: a) glyphs, b) LTG($0^\circ$) for a)}
-\label{ftle_iso}
-\end{figure}
-
-When we look at the FTLE field, on the top right in Fig.~\ref{ftle_iso}, we determine that we obtain a more or less uniform reponse (, if we would neglect the edges) just as expected.
-Because the divergence of TFLs is distributed uniformly in this test field, we get approximately the same response for any position, but yet introducing a systematic bias against the light source direction.
-This is explained by the fact, that the energy will leave the grid sooner on light source positions at the edges for light source directions pointing toward the edges leading to inherently decreased differences and hence gradient magnitudes (responses).
-Note, that this also causes a shift of the FTLE field to the left, as the light source points to the right ($0^\circ$ slice).
-As noticeable, no height ridges are detected in this case as expected, since the test field does not pose any specific regions exhibiting diverging tensor field lines.
-We consider this result the absolute offset response, whereas energies are redistributed and separated more strictly for introducing local features into the grid (field), leading to greater differences, which are in turn emphasized by local regions of interest displayed in deep red at their corresponding positions.
-Imagine this process comparable with ink absorbed by dry paper.
-As we are also interested in the whole ensemble of directions, we can use a volume or surface representation for the $n+1=3$-dimensional grid which results from our FTLE computation.
-The volume grid is what we call the final footprint of the tensor field which is necessarily derived from the initial glyph footprint.
-From it, we can observe the behavior of the tensor field for any direction.
-That is, the slices we observe are stacked along the $\omega\mapsto z$-axis in the following.
-As a subsequent step, we segment height ridges in $3D$ as thresholded volumes from the whole volume grid.
-
-\begin{figure}[t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.8\textwidth]{img/iso_volume.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \includegraphics[width=0.8\textwidth]{img/iso_isovolume.png}
-    \label{b)}
-	b)
-  \end{minipage}
-  \caption{\enquote{isotropic} test field: a) volume, b) IsoVolume for a)}
-\label{ftle_iso}
-\end{figure}
-To verify the concept of absorption implicitly, we incorporate absorption into test field \enquote{absorb}, whereas all cells are absorbing energy except the vertical center line, which is assumed to produce a clear ridge.
-\begin{figure}[t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{img/absorb.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-    \includegraphics[width=0.9\textwidth]{img/absorb_ftle.png}
-    \label{b)}
-	b)
-  \end{minipage}
-  \caption{\enquote{absorb} test field: a) glyphs, b) LTG($90^\circ$) for a)}
-\label{ftle_absorb}
-\end{figure}
-
-We observe the clear straight ridge in the example in Fig.~\ref{ftle_absorb}.
-As expected, the ridge is located on the vertivcal center line.
-Note, that we (min-max) normalize each transmission profile with the highest mean transmission rate (tensor magnitude) in the field.
-From this test case, we can conclude, that regions with a comparably small tensor magnitude would not obtain any FTLE field response, which we will later observe for the real datasets \enquote{brain} and \enquote{heart} exhibiting such areas.
-We will now discuss synthetic examples designed to evaluate the performance of the FTLE.
-As a little warmup, we will have an experiment exhibting bare, raw tensor field lines diverging w.r.t.
-the $y$-axis:
-
-\begin{figure}[t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-	\centering
-    \includegraphics[width=\textwidth]{img/drain(alt)-TFL.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/ftle_drain_alt.png}
-    \label{b)}
-	b)
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/drain_alt_gradient.png}
-    \label{b)}
-	c)
-  \end{minipage}
-  \caption{``drain'' test field: a) TFLs and glyphs, b) LTG($90^\circ$) for a), c) gradient for b)}
-\label{ftle_base}
-\end{figure}
-What we can observe here, is a feature called a ``ridge'' occuring as a high value line segment (red vertical track) which is already quite obvious for human vision.
-A gradient filter could be applied to extract and segment edges from the image as a preprocessing step for computer vision, which would look like the image in Fig.~\ref{ftle_base} c).
-
-These ridges represent ROIs with underlying Lagrangian Coherent Structures, which we already discussed in Sect.~\ref{sect:motiv}.
-It could for example represent a stress field inside of a mechanical tool with a central axis with maximum stress condensation.
-\begin{figure}[ht]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-  \includegraphics[width=0.9\textwidth]{img/drain_top.PNG}\vskip 10pt
-    \includegraphics[width=0.9\textwidth]{img/drain_alt-profile.PNG}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{img/drain_alt_isovolume.PNG}
-	b)
-    \label{b)}
-  \end{minipage}
-  \caption{LTG volumes: \enquote{drain} test field a) cross-section, b) IsoVolume}
-\label{drain_contour}
-\end{figure}
-For the ``drain'' test field we consider the central channel which produces a clear line in the cross section.
-The light transport gradient responds most inside of the channel region and on the top T-arms independently of the stimulus direction.
-For the sake of completeness, we show that the approach's functionality also for diagonal directions in Fig.~\ref{diag-ftle}.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{img/diagDrain.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{img/diagDrain-ftle.png}
-    \label{b)}
-	b)
-  \end{minipage}
-  \caption{``diagonal drain'' test field: a) TFLs and glyphs, b) LTG($225^\circ$) for a)}
-\label{diag-ftle}
-\end{figure}
-Regarding the ridge for the \enquote{drain} test field, we note that it is a little smoothed compared to the axis-aligned ridges, as they obtain a larger propagation attenuation (because of the larger distance) leading to a stronger separation of trajectories in general and hence to a smoothed FTLE response.
-Also we note that the divergence of the tensor field lines is reduced compared to the classic \enquote{drain} example of the FTLE.
-
-\begin{figure}[h]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{img/inverse-TFL.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}[!t]{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{img/inverse-FTLE-raw.png}
-    \label{b)}
-	b)
-  \end{minipage}
-  \caption{``Inverse''-test field: a) TFLs and glyphs, b) LTG($90^\circ$) for a)}
-\label{inverse-ftle}
-\end{figure}
-
-We also evaluated the approach for the ``inverse'' test field, depicted in Fig.~\ref{inverse-ftle}, whereas we can observe height ridges for the $90^\circ$ slice in any position with diverging tensor field lines w.r.t.
-the light source direction.
-This is certainly not the case for the upper (missing) ridge whereas tensor field lines converge with respect to the UP ($90^\circ$) direction.
-If we determine the ridges pointing into the diverging direction, this proves that the method is capable of detecting ridges from parallel up to orthogonal to the light source direction, but not in antiparallel directions, which exhibit converging tensor field lines as bidirectional counterpart.
-Luckily, we measured the gradient for every possible sampled light source direction which complements this information into another slice (and the resulting volume of stacked slices).
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/inverse_isovolume.png}
-    a)
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/inverse_surface.png}
-    b)
-  \end{minipage}
-   \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/inverse_contour.png}
-    c)
-  \end{minipage}
-    \caption{LTG volumes: ``inverse'' test field: a) IsoVolume, b) Surface, c) Contour}
-\label{inverse_volume}
-\end{figure}
-Note that the $z$-axis represents the light source direction angle $\omega$ in this case.
-The contour filter reveals the star-shaped structure of the tensor field which rotates along the angle axis.
-This is the effect we already captured for the single slice: For each axis direction we can detect ridges from parallel up to orthogonal to our predefined stimulus (light source) direction, but not in antiparallel direction.
-This causes the star shape to miss one branch or ray which is in turn the antiparallel direction missing for each light source direction, emphasizing the impression of rotation.
-We can also try to segment height ridges represented as surfaces from the resulting volume grids as depicted in Fig \ref{inverse_contour}.
-The orthogonal ridges in this example reach out to the edges because they are gaining a bias by never offsetting the light source in a transmission direction (ellipsoid orientation) for orthogonals.
-This turns out to be a general feature of the method, as we will later verify in Figs.
-\ref{res_ftle_volume} and \ref{star-ftle}.
-
-%\begin{figure}[!t]
-%\centering
-%  \begin{minipage}{0.4\textwidth}
-%    \includegraphics[height=\textwidth]{img/inverse_surface.png}
-%    \label{a)}
-%  \end{minipage}
-%  \begin{minipage}{0.4\textwidth}
-%    \includegraphics[height=\textwidth]{img/inverse-wireframe.PNG}
-%    \label{b)}
-%  \end{minipage}
-%    \caption{LTG stacked slice surface representations: ``inverse'' test field}
-%\label{inverse_surface}
-%\end{figure}
-
-For the ``inverse'' test field, we can observe a 3 part ``rotor'' indicating the ridges which are in parallel up to orthogonal to the current light source (slice) direction.
-This rotor can be obtained by applying a contour filter, but it can also be visualized by applying a simple thresholding on the surface representation of the volume, as depicted in Fig.~\ref{inverse_volume}.
-Subsequently, we will analyze the \enquote{rings} test field to show its performance on circular ridges.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{img/tensorfieldlines.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}[!t]{0.4\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{img/rings_ftle.png}
-    \label{b)}
-	b)
-  \end{minipage}
-  \caption{``Rings''-test field: a) TFLs and glyphs, b) LTG($0^\circ$) for a)}
-\label{rings-ftle}
-\end{figure}
-What we can observe in Fig.~\ref{rings-ftle}, is called an inside-outside test producing the ridge on the circumference of the circle with maximized radius fitting into the minimum bounding box.
-If one would place a light source on the circumference of this circle offsetting it spatially and in the angular domain, one would obtain one incomplete final light distribution (outside) and one complete final light distribution (inside), as this is the maximum radius at which the intensities can follow a circle.
-This hence leads to large differential light distributions and thus to a higher LTG magnitude.
-ROIs pointing outside the field respond the same for varying light source directions and positions yielding the blue areas in the corners.
-Regions introducing linear anisotropy respond uniformly inside of the circle in contrary to the center point, which is again not responding because of circular overlaps of the offset light distributions computed for the light transport gradient.
-The contour filter reveals these two edge locations in the field, concerning tensor field line divergence.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/rings_contour.png}
-	a)
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/rings_isovolume.png}
-	b)
-  \end{minipage}
-   \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/rings_volume.png}
-	c)
-  \end{minipage}
-    \caption{LTG volumes: ``rings'' test field: a) Contour, b) IsoVolume, c) Volume}
-\label{rings_contour}
-\end{figure}
-The volumes, we examine here in Fig.~\ref{rings_contour}, show that the field is symmetric, as we do not notice a change along the $\omega\mapsto z$-axis except for the natural shift of the field against the light source direction, which we discussed already in this subsection.
-Along the $z$-axis, the light source is rotated in all directions which causes the shift to form a full cycle from bottom to top ($[0^\circ,360^\circ]$).
-At last, we will analyze the bow test field, which is now considered as a more complicated example. We again plot the FTLE field in color coding representation in Fig.~\ref{bow-ftle}.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/bow-TFL.png}
-    \label{a)}
-	a)
-  \end{minipage}
-  \begin{minipage}[!t]{0.4\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/bow_ftle.png}
-    \label{b)}
-	b)
-  \end{minipage}
-  \caption{``bow''-test field: a) TFLs and glyphs, b) LTG($0^\circ$) for a)}
-\label{bow-ftle}
-\end{figure}
-One may observe here, that the ridges located against the direction of the light source are again visible more clearly for the $0^\circ$ slice (directed bias).
-These ridges represent regions revealing most diverging tensor field lines in the original field.
-The ridges on the left side are detected because they are in the range up to orthogonal to the light source direction (if we consider them pointing into the diverging direction).
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/bow_contour.png}
-	a)
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/bow_volume.png}
-	b)
-  \end{minipage}
-    \caption{LTG volumes: ``bow'' test field : a) Contour, b) IsoVolume}
-\label{bow_contour}
-\end{figure}
-Considering the volumes depicted in Fig.~\ref{bow_contour} one may notice, that we obtain a riffled structure reminding of ground plates.
-On half the volume height ($180^\circ$) the structure is turned by $180^\circ$ as expected as this light source configuration lights up the mirror image to the one on the surface.
-The bow is literally shaped from the volume representation if one applies a threshold.
-
-If we relate the resulting FTLE volume to the work of Hlawatsch et al.
-\cite{hlawatsch}, we find that their FTLE field is contained in our volume in form of a manifold, since they use the same basic idea, but consider only a single direction (of the major eigenvector), instead of taking all directions into account, as we do here.
-The resulting manifold included, is constrained to a height field (not self-overlapping or self-intersecting), because each position can only represent one individual direction, as there is only one direction for the major eigenvector at each position to be covered (, since it is obviously a definite, distinct measure).
-This direction emerges as a single scalar height (FTLE) value on each position corresponding to a height (direction) slice in our volume.
-That is, each position in their $2D$ scalar field is assigned a height in our $3D$ scalar field.
-We cover their results as a subset of our volume, which could be, e.g., a spiral plane in case all directions are given in circular arrangement in the underlying tensor field (cf.~\enquote{rings} test field).
-If we would vary the angle of the major eigenvector with a uniform distribution along the tensor field's width, we would obtain a primitive diagonal plane w.r.t.
-the $xy$-plane reaching out in $z$-direction as contained subset.
-Therefore, we claim and consider our invented technique to be more generalized (to all directions) resulting in a topological $n+1$-D volume.
-
-\chapter{Algorithm}
-\label{chap:imp}
-In this chapter, we will describe the implementation process from top to bottom following runtime chronology.
-At first, we will load a configuration file setting angular resolution in (discrete) steps, which are spaced uniformly over the interval $[0,2\pi]$.
-The threshold for the convergence criterion was empirically determined and set in the range $[0.5,2.0]$.
-Since it is an energetic threshold, it is only dependent on the light source magnitude and not on any type of resolutions.
-An iteration limit is set to $\mathop{limit}=10\mathop{width}$ (e.g.: for $\mathop{width}=10\cdot 100=1000$)  to prevent the approach from eventual oscillation behavior for intensities and hence overpropagation (propagating for too many timesteps).
-The spatial resolution is inferred by the tensor field reader method from the tensor field (see Propagation below).
-The light source profile can be defined for the \enquote{non-FTLE} mode where the user sets position and polar profile of single or multiple light sources as a symbolic term: e.g.
-\texttt{clip(cos(theta-pi/2))} for a Lambert-emitter pointing upwards.
-After the configuration read process is finished, we allocate memory for certain buffers holding glyph and intensity information with a dimension of $\mathop{dim}=\mathop{width}\mathop{height}\mathop{steps}$.
-This $1D$ buffer is interpreted like this:
-
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.5\textwidth]
-    {img/grid.pdf}
-  }
-  \caption{Indices and dimensions in grid}
-  \label{indices}
-\end{figure}
-Fig.~\ref{indices} depicts the \textit{deltaIndex}-map for computing the destination index for the scaled cosine lobes for cell $c$.
-The size of each cell is $\mathit{size}=\mathit{steps}$ in a $1D$ buffer arrangement, respectively.
-
-\section{Stimuli and Convergence}
-We then initialize energetic sums concerning the convergence and thresholds and run a while loop until the convergence criterion is met.
-We compute the difference $\Delta\Phi_{total}$ of the energetic sum in this iteration with the one from the previous iteration.
-If it falls short of this threshold, we initiate a stop sequence and leave the loop.
-To compute the FTLE, we use multiple stimuli in parallel (e.g.: Dirac pulses within the scope of this work), and just compute the differences in final light distributions for two close-by light source positions and directions to accumulate them in a total energy sum (capacity), which in turn forms one component of the light transport gradient.
-A Euclidean norm yields a scalar field from the gradient at each position.
-
-\begin{algorithm}[t]
-\caption{Propagate Light Distribution}\label{PLD}
-\begin{algorithmic}[1]
-\Procedure{propagateDist}{i,j,$\omega$}\Comment{Input position $j,i$ and direction $\omega$}
-\State \Call{fill}{\textit{bufferA}, $0.0$} \Comment{Reset and Initialize}
-\State \Call{fill}{\textit{bufferB}, $0.0$}
-\State $\mathit{sumMem} \gets 0.0$
-\State $\mathit{finished} \gets \mathit{false}$
-\State $\mathit{index} \gets (j\cdot\mathit{width}+i)$ \Comment{compute $1D$ (cell) index}
-\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{write light source }
-\While{$\Delta\Phi_{total} < \mathit{\epsilon }$ } \Comment{while convergence criterion not met..}
-\State $\mathit{sumA} \gets 0.0$ \Comment{reset sum}
-\State \Call{propagate}{\null} \Comment{propagate \textit{bufferA} (source) in \textit{bufferB} (tar)}
-\State $\mathit{sumA} \gets$\Call{sum}{\textit{bufferB}}  \Comment{sum up energies}
-\State $\Delta\Phi_{total} \gets \lvert\mathit{sumA}-\mathit{sumMem}\rvert $\Comment{compute difference to prev. iter.}
-\State $\mathit{sumMem} \gets \mathit{sumA}$\Comment{save sum for next iter.}
-\State \Call{swap}{\textit{bufferA}, \textit{bufferB}} \Comment{swap buffers for restart}
-\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{re-write light source}
-\State \Call{fill}{\textit{bufferB}, $0.0$}
-\If {$\mathit{ctr}>\mathit{limit}$}\Comment{stop on iteration limit}
-\State break
-\EndIf
-\EndWhile\label{euclidendwhile} \Comment{final light distribution stored in bufferA..}
-\State return \textit{bufferA}
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-Note that we set the light source intensity to constant value $\mathit{steps}$ to require a mean intensity of $\mu_I=1.0$ comparable to the one of a unit circle independently on the angular resolution.
-
-When we consider the LTG method, we already stated its runtime behavior ($\mathcal{O}(\mathop{width^5}\mathop{steps^2})$), which makes parallelization libraries and hardware to compute the results in reasonable time.
-First, we tried to implement a GPU-based parallelization for the propagation scheme with nVidia's Thrust library, which turned out not to work as expected, since the relative runtime gain was too low to make a big difference and we struggled to find a suitable implementation mapping the propagation process to atomic OPs (simple vector operations), which are the only operations allowed for GPU-parallelization.
-At this point, we choose to use OpenMP as we figured it is a much more suitable architecture employed here, which allows to perform the computation on a multiprocessor architecture provided by the Heidelberg University allowing to compute $64$ hardware threads at maximum.
-As we need fat fed, instruction-rich threads to have a high performance gain, we choose to parallelize only the outer loop of the LTG method (propagating final light distributions for each light source direction).
-Remark, that each thread requires its own memory, as an independent memory block is needed to perform the propagation in an individual propagator object of class propagator, in order to perform the computation in parallel.
-
-\section{Propagation}
-The method performing the light propagation, was implemented in a separate class called propagator, which we can initialize once for each programming thread (, which we can in turn provide with its own buffer memory).
-As an initialization we fill the buffers with zeros to have a blank buffer initially to fill up with meaningful intensity profiles.
-The transmission profiles are generated by loading the tensor field in memory, performing PCA on each of them and exploiting the thin $\mathbf{U}$ matrix for the first (greater) singular vector.
-The resulting principal component ellipsoid equation is precomputed, sampled and stored for reuse once for each cell.
-We again sample cosines into each direction, average out their integrals (sums) and normalize each of them with the mean to have the integral energy scaled to $1$ (identity element for multiplication, as it is the case for \enquote{scaling the cosine lobes} here).
-Thus we ensure, that the transmitted energy is approximately $1.0=100\%$ for each lobe and account for angular sampling errors.
-Now we need to precompute and store the linear combinations used for the \enquote{shared part} in diagonal cones.
-When hitting border cells, we propagate the intensities only to the present neighbor directions still located on the grid.
-Remark, that we can not just ignore the edges of the grid because it would lead to large trivial intensity differences on grid borders for the FTLE-similar approach LTG.
-
-\begin{algorithm}[t]
-\caption{Propagate BufferA to BufferB}\label{PB}
-\begin{algorithmic}[1]
-\Procedure{propagate}{}
- \For{each cell $c$ in bufferA}
-\State $\mu_{I_c}\gets mean(I_c(\omega))$ \Comment{Compute Mean Intensity}
-\If {$\mu_{I_c}=0.0$}\Comment{break on trivial null sample}
-\State break
-\EndIf
-\State $\mu_{T_c}\gets mean(T_c(\omega))$ \Comment{Comp. Mean Transmission}
-\State $\mu_{\mathit{T_cI_c}}\gets mean(T_c(\omega)\cdot I_c(\omega))$ \Comment{Comp. Mean Transmitted Intensity}
-\State $n_c \gets \frac{\mu_{T_c}\mu_{I_c}}{\mu_{T_cI_c}}$\Comment{Comp. normalization factor for cur. cell}
- \For{each direction $k$ in $[0,7]$}\Comment{For all neighbors, do..}
- 	\State $\gamma \gets k\frac{\pi}{4}$\Comment{Offset to compute cone center angle}
- 	\If {$k\bmod 2 = 0$}\Comment{if even (face) neighbor..}
-	\State $\mathit{energy} = \int_{\gamma-\pi/4}^{\gamma+\pi/4}n_c\epsilon_\alpha T_c(\omega)I_c(\omega)\mathop{d\omega}$
-	\Else\Comment{if odd (diagonal) neighbor..}
-	\State $\mathit{energy} = \int_{\gamma-\beta}^{\gamma+\beta}n_c\epsilon_\beta T_c(\omega)I_c(\omega)\mathop{d\omega}$
-	\EndIf
-	\State $\mathit{index} \gets c+\mathit{deltaIndex}(k)$\Comment{assign neighbor destination index}
-	\State $\mathit{bufferB(index,\omega)}\gets \mathit{energy}\cdot cos_k(\omega)+\mathit{bufferB(index,\omega)}$\\
-	\Comment{Scale cosine lobe with summed energy and accumulate}\\
-	\Comment{$y=a*x+y\rightarrow$ axpy-OP}
- \EndFor
- \EndFor
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\section{Languages and Libraries}
-At first, we needed to generate convenient test cases, which we design using Mathematicas VectorPlot, used to visualize vector fields.
-To output a predefined test case, we need to define a little code, for which we used python scripts which generate VTK appended or ASCII format files.
-To implement the method we used C++'s STL for every dynamic memory (buffer), SFML to plot polar profiles, Eigen to perform matrix decompositions and VTK to in- and output VTK format files, which can in turn be visualized and explored interactively with the visualization framework ParaView.
-The user functions defined in the configuration file are read-in by a string parser (muParser Library), which is used just once for sampling the polar profile and storing the precomputed result.
-This causes a convenient input mechanism for stimuli, in case they should be predefined.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Results and Evaluation}
-\label{chap:eval}
-\section{Propagation Attenuation}
-\label{sec:att}
-It can be proven that the approach follows propagation attenuation principles and follows the inverse (square in $3D$) law ($\sim\frac{1}{r}$) as observed in the graph in Fig.~\ref{att}.
-For that, we plot the relative intensity $i_{rel}(r_{rel})$ (in relation to the intensity at position (radius) $1$ to account for the inverse square law, which holds for relative distances.
-We evaluate the relative intensity by forming a ratio $\frac{I(r)}{I_1}$, that relates the step intensity to the source intensity at distance $1$.
-
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.6\textwidth]
-    {img/inverse_law.pdf}
-  }
-  \caption{Propagation attenuation}
-  \label{att}
-\end{figure}
-
-\section{Behavior for Total Anisotropy}
-\label{sect:total}
-We also wanted to check for the limitations of the approach by having total linear anisotropy (the case whereas one eigenvalue $\lambda_1>0$ and the other $\lambda_2=0$) incorporated.
-For this we set up bidirectional Dirac-pulses as transmission profiles to have the whole energy transmitted in just one, discrete direction.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/tensorfieldlines.png}
-	a) TFLs and glyphs
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/total_anisotropy1.png}
-    b) iteration $50$
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/total_anisotropy2.png}
-	c) iteration $103$
-    \label{c)}
-  \end{minipage}
-\caption{intensity propagation in grid for total anisotropy}
-\label{anisotropy}
-\end{figure}
-What we can observe here, is kind of a sampling drift as the approach uses 8 discrete directions to propagate averaged intensity whereas the circle uses much more (namely infinite if presumed as perfect circle).
-Hence, the energy can never follow a perfect circle, but at best an octagon.
-This means, that there is a drift of at least two cells per diagonal (one overshoot per enter and escape each) as observable in Fig.~\ref{anisotropy}.
-Nonetheless, the energies are transmitted in circular orbits around the circle center.
-\section{Energy Conservation Test}
-\label{sect:conserve}
-The energy conservation test was done in similar manner as the propagation attenuation test in Sect.~\ref{sec:att}.
-We measure the impulse response of the algorithm for 80 iterations by summing the total energy in the grid after each iteration and comparing it to the initial energy sum (e.g.: $\int_0^{2\pi}1\mathop{d\omega}=2\pi$) in iteration 0 in absolute values.
-This energetic error analysis yields the following graph depicted in Fig.~\ref{energy}:
-
-\begin{figure}[!t]
-  \centering
- {
-    \includegraphics[width=0.6\textwidth]
-    {img/sum_difference.pdf}
-  }
-  \caption{Sum Error: source intensities: red: 1.0, blue 2.0, green: 5.0}
-  \label{energy}
-\end{figure}
-
-The approach remains stable until iteration 40, which is at the time (point) when the numbers are too small and get rounded down under the double machine epsilon.
-You can see heavy fluctuations because the numbers get rounded upwards likewise.
-But what is lost, is lost and can not be regained.
-The error is thus accumulating gradually and with faster slope for higher source intensities, since floting point precision, i.e., rounding errors operate more heavily on large numbers since it is a constant relative loss $\sim\frac{1}{r}$ leading to a greater absolute $\Delta E$.
-
-\section{Light Transport Gradient (LTG)}
-The light transport gradient approach, which is closely related to the FTLE approach in vector fields, is evaluated for some more synthetic and real datasets in this section.
-We will start off with the comparison of our FTLE-based approach to the one invented by Hlawatsch et al.
-\cite{hlawatsch}, which we reimplemented for verification purposes.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}[t]{0.3\textwidth}
-  \includegraphics[height=\textwidth]{img/drain(alt)-TFL.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}[t]{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/ftle_drain_alt.PNG}
-	b)
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}[t]{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/drain_ftle_org_real.PNG}
-	c)
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Drain} test field: a) Glyphs, b) LTG($0^\circ$), c) FTLE: Hlawatsch's approach}
-\label{drain_ftle_comp}
-\end{figure}
-
-We can observe the comparison between our invented FTLE-like approach and the one of Hlawatsch et al.
-in the depiction in Fig.~\ref{drain_ftle_comp}.
-Note, that a sharper but weaker ridge is detected on the right side of the image, which shines similar to a cone shape from the top.
-We interpret the missing spot in the top center as an artifact stemming from the way we handle our light propagation, as we allow whole distributions of trajectories to be propagated (with a scalar attribute interpretable as propability).
-This causes a sampling drift, we already discussed in Sect.~\ref{sect:total}, leading to an energy distribution ending up outside of the field, for this position.
-The directive effect turns out to be not that strict and delayed compared to the classic tensor field line FTLE.
-The crucial difference is that the compared work of Hlawatsch et al.
-obtains the FTLE field from the gradient $\nabla\mathbf{\upvarphi}(\mathbf{x})$ of the flow map, which is in turn obtained from sampling of trajectories following tensor field lines.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.29\textwidth}
-  \includegraphics[height=\textwidth]{img/star.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.29\textwidth}
-    \includegraphics[height=\textwidth]{img/star_ftle.PNG}
-	b)
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.35\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img/star_volume_alt.PNG}
-	c)
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Star} test field: a) field, b) LTG slice($0^\circ$), c) IsoVolume on LTG volume}
-\label{star-ftle}
-\end{figure}
-When we look at the example depicted in Fig.~\ref{star-ftle}, we can see the FTLE approach LTG responding most in the center of the field, which is the nodal point of tensor field lines diverging in any direction.
-Remark, that in the middle image there is a light V-shaped ridge on the left on blue background which is outshined by the center spot because of its orders of larger magnitude.
-The volume illustrated on the right even shows a spiked structure for the volume representing all directions at once.
-The approach detects more ridges in diagonal directions stochastically as we can infer from the Gaussian-shaped plates in the thresholded volume, evenly spaced in $90^\circ$ shifts with $45^\circ$ (quarter volume) offset.
-These can be interpreted nicely, since the diagonal direction is never offset into its transmission direction leading to a trivial (none) difference for axis-parallel directions for the offset direction orthogonal to the light source direction.
-Hence, the differences will be factor $2$ at worst for the diagonals, since we measure differences in both offset directions, in case ridges in both axis-parallel and -diagonal directions are present.
-Accordingly, we observe Gaussian-plates in the volume instead of sharp, contrasted ridges for the \enquote{star} example.
-But wouldn't we expect a ridges only in parallel directions?
-Well, in this case, we could imagine not only tensor field lines diverge or converge, but also they could cross each other unaffectedly (just as with light paths), which could be the case when shifting away from the diagonal directions.
-For better imagination, offset the light source vertically or horizontally from the diagonals and observe the light paths hitting the axes (, which in fact do cross each other).
-
-%A possible correction would be, to normalize each slice with a heuristic factor $\frac{1}{1+(\sqrt{2}-1)\lvert\cos2\omega-\frac{\pi}{4}\rvert}$ to get a uniform response over all directions and account for the longer diagonal run lengths (SLDs).
-
- We will now capture the star dataset revealing tensor field lines diverging in any direction from the center.
-Now, we will evaluate a classic textbook example called the \enquote{double gyre} \cite{shawn}, which is a quadruple gyre in our case for formatting reasons, yet yielding the same conceptual results.
-We compare the glyph representation to the LTG magnitude for a single slice ($0^\circ$) again:
-\begin{figure}[!t]
-\centering
-  \begin{minipage}[t]{0.3\textwidth}
-  \includegraphics[height=\textwidth]{img/gyre.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}[t]{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/gyre_ftle.PNG}
-	b)
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}[t]{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/gyre_ftle_org.PNG}
-	c)
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Gyre} test field: a) Glyphs, b) LTG($0^\circ$), c) FTLE: Hlawatsch's approach}
-\label{gyre-ftle}
-\end{figure}
-In Fig.~\ref{gyre-ftle}, we can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t.
-the $x$-axis ($0^\circ$).
-Note that even regions in antiparallel directions ridges are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging.
-Again, it is noticeable, that the result of Hlawatsch correspond to our result to a certain degree.
-It is noticeable, that Hlawatsch's approach does not consider the initial direction and hence will detect ridges also in antiparallel directions (gradient of the flow map).
-This yields symmetric ridges w.r.t.
-central axes, whereas we detect ridges only up to orthogonally to the initial light source direction as visible in Fig.~\ref{gyre-ftle}b)\& c), respectively.
-In the center location of each gyre, we interpolate the sorrounding ring with an isotropic cell in the middle, leading to chaotic trajectories leaving the gyre and thus leading to an increased gradient of the flow map for Hlawatsch's result.
-This causes the FTLE field to respond in the center of the gyres in Fig.~\ref{gyre-ftle}c).
-We interpret this behavior within the scope of interpolated trajectories, which are in general more likely to escape the gyre, since their angle is interpolated (by PCA), which is susceptive to rapid changes and hence leads to unstable behavior.
-What Hlawatsch et al.
-detect, could be called local ridges (in terms rather short), what our method detects could rather be called global ridges (in terms rather long).
-This implies, that we obtained reasonable results generated by our method, but the classical tensor field line FTLE is susceptive to instability concerning its directive abilities and does not do things by halves (like we do with our partitioned light distributions), which causes sharper ridges in general.
-Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure.
-We again depict volume representations for this example:
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-  \includegraphics[height=\textwidth]{img/gyre_volume2.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/gyre_gauss_contour.PNG}
-	b)
-    \label{b)}
-  \end{minipage}
-   \begin{minipage}{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/gyre_volume4.PNG}
-	c)
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Gyre} test field LTG volumes: a) IsoVolume, b) Contour, c) GaussContour}
-\label{gyre_volume}
-\end{figure}
-When we consider the volumes for the \enquote{gyre} test field, we can remark the holes, which represent the gyres in the centers of the volume for a low and a high threshold, whereby we extract (filter out) the previously observed ridges in $3D$ for all directions $\omega$ simultaneously.
-
-Now, that the approach has passed the textbook examples, we push it forward to be applied in real (measured) data to show its performance in discerning and segmenting ROIs from a human brain.
-The approach will detect LCS in the brain's tissue from MRI-measured diffusion tensors of a male subject at the age of 27.
-We compare ParaView's glyph representation to the FTLE method LTG for the test field \enquote{brain}:
-\begin{figure}[!t]
-\centering
-  \begin{minipage}[t]{0.3\textwidth}
-  \includegraphics[height=\textwidth]{img/brain_cyl_glyphs.PNG}
-    \caption*{a) glyph representation}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}[t]{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/brain_ftle.PNG}
-     \caption*{b) LTG($0^\circ$)}
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}[t]{0.3\textwidth}
-    \includegraphics[height=\textwidth]{img/brain_ftle_org.PNG}
-     \caption*{c) FTLE: Hlawatsch's approach}
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Brain} test field}
-\label{brain-ftle}
-\end{figure}
-Regarding Fig.~\ref{gyre_volume}, we find that the approach detects the boundaries of concavities (lat.: Gyri) in the brain, forming the lobus cerebri (commonly known a lobes).
-The invented method could thus, e.g., aid to compute $3D$ models of a subject in medical sciences.
-It emphasizes regions with increased change in flow-rate (throughput), because in this areas the anisotropy and/or tensor magnitude is increased, preventing the intensities from being absorbed, hence yielding to a high FTLE-response.
-The regions highlighted in red, are caused by high tensor magnitudes, as we can infer from the glyph representation revealing no linear anisotropy at all.
-Note that we do not show the volume representation in this case, since all directions (slices) look approximately the same because the tensor field is only introducing changes in tensor magnitude not in anisotropy as inferred from the glyph representation.
-The regions with a tensor magnitude higher than zero restrict the light propagation to certain areas with the condition, that not all of the energy is absorbed by sorrounding cells in the local neighborhood.
-Again comparing the results to the classical FTLE (Hlawatsch's approach) on tensor fields in Fig.~\ref{brain_ftle_comp}, we notice that the FTLE response is chaotic, since this MRI scan is subject to inherent MRI noise filling the background regions in the image.
-This noise produces random trajectories, leading to an increased gradient of the flow map and thus to a high FTLE response, which reveals trajectories running in random directions in background regions.
-Only in foreground regions, the trajectories are directed in definite manner, which is why the flow map's gradient is minimized along the brain structure (foreground, depicted in blue in Fig \ref{brain_ftle_comp} c)).
-Yet, our approach uses absorption to decrease noise present in the tensor field and reasonable min-max scaling to conduct $100\%$ of the intensity for the cell with the highest tensor magnitude.
-Hence, we do not yield a noisy representation but a FG-BG segmented image, with red highlights in positions, where the tensor magnitude is maximized (and thus absorption is minimized).
-Remark, that the resulting volume does not inherently depend on the direction, because the test field does not introduce any strong linear anisotropy, as visible from the glyph representation in Fig.~\ref{brain_ftle_comp} a).
-If no linear anisotropy was incorporated, the FTLE response will also be independent on the direction to the greatest extent.
-The slices we obtain as our volume are, with reasonable explanation, more or less identical comparing them against each other for certain directions.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-   \centering
-  \includegraphics[width=0.9\textwidth]{img/brainDwnsmpl.png}
-    \caption*{a) Glyphs}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-   \centering
-    \includegraphics[width=0.9\textwidth]{img/brainDwnsmpl-TFL.PNG}
-     \caption*{b) Hlawatsch's app.: TFLs}
-    \label{b)}
-  \end{minipage}
-\caption{\enquote{Brain} test field}
-\label{brain_ftle_comp}
-\end{figure}
-We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig.~\ref{brain_ftle_comp}.
-We propose the following threshold correction to prevent noisy FTLE results:
-
-We note that this real data example incorporates noise into the BG regions as observable from the chaotic trajectories in Fig.~\ref{brain_ftle_comp}.
-We propose the following threshold correction to prevent noisy FTLE results:
-
-\begin{align*}
-r_i(\omega) = 
-\begin{cases}
-    \frac{1}{\mean{r}_i}r_i(\omega),& \text{if } \mean{r}_i > \theta,\\
-     r_i(\omega),              & \text{otherwise}.
-\end{cases}
-\end{align*}
-
-This normalization correction will ensure, that any mean tensor magnitude below threshold $\theta$ will be neglected for normalization and hence will not be amplified.
-The free parameter should be defined with heuristics concerning the signal-to-noise ratio.
-This now states a local normalization, which was not feasible with the noisy background before introducing this change, since a local normalization would have amplified noisy signals likewise.
-Now, we are capable of producing a tensor magnitude weighted FTLE mode with global normalization, which we used up to this point and another FTLE mode indicating changes in anisotropy separately while neglecting any tensor magnitudes for local normalization.
-This change will now yield a volume (now depending on the direction) for the local normalization mode, which we will show later on.
-We could also claim, that this correction would be necessary for Hlawatsch's approach as well, as their result is susceptive to noise for the \enquote{brain} test field in Fig.~\ref{brain_ftle_comp} b).
-We first show the functionality of the correction on their approach in Fig.~\ref{brain_ftle_noise}.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}[t]{0.4\textwidth}
-   \centering
-  \includegraphics[width=0.9\textwidth]{img_new/brain-TFL.png}
-	a) TFLs: noise correction
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}[t]{0.4\textwidth}
-   \centering
-    \includegraphics[width=0.9\textwidth]{img/brain_ftle_surf_noisereduction.PNG}
-	b) Hlawatsch's approach: noise correction
-    \label{b)}
-  \end{minipage}
-\caption{\enquote{Brain} test field}
-\label{brain_ftle_noise}
-\end{figure}
-Again observing the results for the tensor field lines, we now remark their proper correlation in BG regions (cf.~straight lines in the background in Fig.~\ref{brain_ftle_noise} a)).
-This yields a gradient of the flow map exactly equal to zero, which avoids any noisy FTLE responses, as visible in Fig.~\ref{brain_ftle_noise} b) for Hlawatsch's approach.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-  \centering
-  \includegraphics[width=\textwidth]{img_new/brain_surf.PNG}
-     \label{a)}
-      a)
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img_new/brain_volume2.PNG}
-    \label{b)}
-    b)
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{img_new/brain_cross-section.PNG}
-    c)
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Brain} test field LTG: a) LTG($0^\circ$), b) LTG volume, c) cross-section}
-\label{heart-ftle}
-\end{figure}
-
-The volumes depicted in Fig.~\ref{heart-ftle} now show variation in $z$-direction, as they are now visualizing pure FTLE fields with no weighting of the tensor magnitude incorporated.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-  \centering
-  \includegraphics[width=\textwidth]{img/heart.PNG}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img_new/heart_slice0.PNG}
-	b)
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[width=\textwidth]{img/heart_ftle_org.PNG}
-	c)
-    \label{b)}
-  \end{minipage}
-  \caption{\enquote{Heart} test field: a) LTG global, b) LTG local, c) Hlawatsch's approach}
-\label{heart-ftle}
-\end{figure}
-After applying the LTG approach on the heart dataset, we observe the resulting tensor magnitude weighted FTLE-field (global) and the locally normalized (local) in Figs.
-\ref{heart-ftle}a)\& b).
-Note that the FTLE response for the globally normalized tensor field is again not changing w.r.t.
-direction, which is why we do not show the volumes for the global noramlization.
-The approach highlights the regions with diverging tensor field lines resulting from linear anisotropy or increased tensor magnitude, emerging as ridges.
-In Fig.~\ref{heart-ftle}a) we observe the tensor magnitude weighted mode for an inital global normalization, whereas anisotropy characteristics get outshined by tensor magnitude influences.
-On the other hand, in Fig.~\ref{heart-ftle}b) we remark the typical edge artifacts for FG regions known from the classical FTLE.
-Also, we notice the approach detecting changes in anisotropy in lack of any directed ridges, which reveal diverging tensor field lines w.r.t.
-a central axis.
-The results seem to correspond to the ones of Hlawatsch et al.
-certainly only in parts of the domain, because we unrolled all the FTLE responses distributed over all discrete directions, which is why some of the FTLE ridges are missing.
-Note that these ridges are not that sharp compared to Hlawatsch's approach, as we use whole distributions of trajectories which should be interpreted in a more continous way (changing w.r.t.
-direction).
-These ridges could now represent blood vessels and such.
-The red spot emphasized in the left of the image could, e.g., represent the highest diffusion velocity in tissue.
-When considering the FTLE for the heart dataset depicted in Fig.~\ref{heart-ftle-comp}, we find that the trajectories exhibit strong correlations amongst them.
-This results from the low noise rate inherently present in this dataset.
-The inital value for our angle is given by $0$ degree, which is why the trajectories follow this angle strictly in case of a zero matrix (corresponding to a noise free cell in background regions).
-In this case, we obtain the same results as the approach from Hlawatsch et al., whereas our global min-max normalization (due to noise) is becoming a redundant step for noiseless input data.
-A proof of this elaboration can be found in Fig.~\ref{heart-ftle-comp}.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-   \centering
-  \includegraphics[width=\textwidth]{img/heartDwnsmpl.png}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-   \centering
-    \includegraphics[width=\textwidth]{img/heartDwnsmpl-TFL.PNG}
-	b)
-    \label{b)}
-  \end{minipage}
-\caption{\enquote{Heart} test field: a) glyphs, b) TFLs}
-\label{heart-ftle-comp}
-\end{figure}
-Again considering the trajectories for the heart test field, we note that they all correlate in background regions, which causes the gradient of the flow map to equal zero.
-This occurs, since we do not have any MRI noise incorporated into the background of the field.
-Consequently, we obtain results corresponding to Hlawatsch's approach, which is in general susceptive to noise, but not in case the noise was removed or is not present at all initially.
-From observing Hlawatsch's approach detecting the heart edges, we can infer that the flow is also congruent inside of the heart, inducing a gradient response only to the edges.
-In our case, we have an offset response incorporated which happens even for total spherical isotropy, wherefore we will obtain a response even if the flow does not change locally (presumed it is not absorbed by low tensor magnitudes).
-This in turn stems from the fact, that we use a energetic flood-fill difference, which inherently introduces an offset response.
-We can again apply a threshold for the normalization to obtain a result non-dependent on the noise to visualize raw changes in anisotropy and thus to obtain a variation in $z$-direction, ie., along the light source direction.
-\begin{figure}[!t]
-\centering
- \begin{minipage}{0.3\textwidth}
- \centering
- \includegraphics[height=\textwidth]{img_new/heart_volume.PNG}
-   \label{a)}
- \end{minipage}\hskip 12pt
- \begin{minipage}{0.3\textwidth}
- \centering
-   \includegraphics[width=\textwidth]{img_new/heart_volume3.PNG}
-   \label{b)}
- \end{minipage}
- \begin{minipage}{0.3\textwidth}
- \centering
-   \includegraphics[height=\textwidth]{img_new/heart_volume5.PNG}
-   \label{b)}
- \end{minipage}
- \caption{\enquote{Heart} test field: LTG volumes}
-\label{heart-ftle}
-\end{figure}
-Fig.~\ref{heart-ftle} depicts the volumes for the heart test field, whereas we can observe a 3D ridge running in spiral along the $z$-direction.
-The FTLE responses are shifted w.r.t.
-varying directions along the $z$-axis.
-
-We also introduced random numbers in to the our tensor field generation process.
-Therefore, we define a uniform grid with low resolution ($9\times9$) and interpolate bilinearly to obtain a tensor field with higher resolution exhibiting continous changes in anisotropy and tensor magnitude.
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-   \centering
-  \includegraphics[width=\textwidth]{img_new/random-global.png}
-    \caption*{Global normalization}
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-   \centering
-    \includegraphics[width=\textwidth]{img_new/random-local.PNG}
-     \caption*{Local normalization}
-    \label{b)}
-  \end{minipage}
-\caption{\enquote{Random} test field: a) global, b) local normalization}
-\label{random}
-\end{figure}
-Observing the generated random test field both with global and local normalization in Fig.~\ref{random}a)\& b) respectively, we remark that the propability for small numbers to have a high ratio in combination is increased, as linear anisotropy occurs mainly in BG regions (low tensor magnitudes).
-In case we use global normalization in Fig.~\ref{random_FTLE}a), we will again obtain a FTLE field independent on the direction, since tensor magnitudes outshine the changes in anisotropy in this test case.
-The LTG approach works as expected, as we see the tensor magnitude weighted into the left image, yielding a high response for the tensor magnitude weighted FTLE response.
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.4\textwidth}
-   \centering
-  \includegraphics[width=\textwidth]{img_new/random_global.png}
-b)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.4\textwidth}
-   \centering
-    \includegraphics[width=\textwidth]{img_new/random_local_slice0.PNG}
-	a)
-    \label{b)}
-  \end{minipage}
-\caption{\enquote{Random} test field LTG: a) global, b) local normalization}
-\label{random_FTLE}
-\end{figure}
-In case we use local normalization in Fig.~\ref{random_FTLE}b), we could interpret the contributions revealing low tensor magnitudes as a result from white noise, which shows the inherent problem with local normalization.
-If we would just locally normalize the whole tensor field with the mean tensor magnitude of each cell, we would enhance noisy contributions, which turn out to be in general more propable to be anisotropic than FG data.
-Keeping this consideration in mind, we observe the approach detecting raw changes in anisotropy.
-Note, that this test field does not introduce clearly defined ridges with tensor field lines diverging w.r.t.
-to a middle axis.
-We just introduce rapid random variations in anisotropy along the spatial domain, which is why the FTLE response is independent on the direction in this case, as well.
-This example emphasizes, that ridges can also be yielded from rapid changes in anisotropy and are hence in this case, not directed and thus non directionally dependent.
-As a posteriori verification, we take home from this example, that the thresholding is in fact necessary to form a reasonable local normalization, because otherwise we would emphasize linear anisotropy resulting from noise.
-
-\paragraph{Documentation}\vskip 1pt
-\textit{Spatial resolution pyramid}
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-    \centering
-  \includegraphics[height=\textwidth]{img/inverse29.png}
-    a) $\mathop{res}=29\times29\times16$
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth} 
-    \centering
-    \includegraphics[height=\textwidth]{img/inverse51.PNG}
-	b) $\mathop{res}=51\times51\times16$
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img/inverse101.PNG}
-	c) $\mathop{res}=101\times101\times16$
-    \label{b)}
-  \end{minipage}
-  \caption{Spatial resolution pyramid of \enquote{inverse} test field - slice}
-\label{res_ftle}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-	\centering
-  \includegraphics[height=\textwidth]{img/inverse29_volume.png}
-	a)~$\mathop{res}=29\times29\times16$
-    \label{a)}
-  \end{minipage}%\hskip 17pt
-  \begin{minipage}{0.3\textwidth}
-	\centering
-    \includegraphics[height=\textwidth]{img/inverse51_volume.PNG}
-	b)~$\mathop{res}=51\times51\times16$
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-	\centering
-    \includegraphics[height=\textwidth]{img/inverse101_volume.PNG}
-	c)~$\mathop{res}=101\times101\times16$
-    \label{b)}
-  \end{minipage}
-  \caption{Spatial resolution pyramid of \enquote{inverse} test field - volume}
-\label{res_ftle_volume}
-\end{figure}
-When observing these (decomposed) spatial resolution pyramids shown, we can observe the slice image becoming more and more continous with increasing resolution to the right.
-The right image on the top, shows a result fulfilling the Nyquist-Shannon sampling theorem, as it appears with no jaggies or pixelation and thus successfully deceiving the human mind in being continous.
-The volumes on the other hand, exhibit a slight difference, which is due to the revealing of new discretized directions implicitly in the tensor field, as shifts in angle become more gradual and continuously for a more dense sampling of the same implicitly content-based field.
-This effects a more precise behavior, which now indicates the orthogonal ridges from each of the $2$ directions in parallel to the light source direction (axis-parallel) on each side of the volume.
-They are evenly spaced in a shift of half the volume height which represents the angle $180^\circ$ (2 antiparallel directions in axis directions).
-The blue span (hole), we perceive in the middle of the cross-section, is representing the antiparallel direction, which in turn is missing, since we only can detect divergence of tensor field lines in axis-parallel up to -orthogonal directions (if we consider them pointing into the diverging direction).
-Note that we already discussed this feature in section \ref{chap:meth}.
-Also remark, that the ridges get sharper with increasing spatial test field (input data) resolution because the implicitly defined content of the tensor field is more densely sampled at a higher resolution allowing for more continous changes in directions of the major eigenvectors (glyph orientations).
-\textit{Angular resolution pyramid}
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
-  \centering
-  \includegraphics[height=\textwidth]{img/slice16.png}
-	a) $\mathop{res}=101\times101\times16$
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-    \centering
-    \includegraphics[height=\textwidth]{img/slice32.PNG}
-	b) $\mathop{res}=101\times101\times32$
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img/slice64.PNG}
-	c) $\mathop{res}=101\times101\times64$
-    \label{b)}
-  \end{minipage}
-  \caption{Angular resolution pyramid of \enquote{inverse} test field - slice}
-\label{resA_ftle}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-  \begin{minipage}{0.3\textwidth}
- 	\centering
-  \includegraphics[height=\textwidth]{img/vol16.png}
-	a)~$\mathop{res}=101\times101\times16$
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img/vol32.PNG}
-	b)~$\mathop{res}=101\times101\times32$
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[height=\textwidth]{img/vol64.PNG}
-	c)~$\mathop{res}=101\times101\times64$
-    \label{b)}
-  \end{minipage}
-  \caption{Angular resolution pyramid of \enquote{inverse} test field - volume}
-\label{resA_ftle_volume}
-\end{figure}
-Observing these angular resolution pyramids, one may again notice, that the results can be considered to be independent on the angular resolution, as well.
-The only change incorporated is that the ridges get more and more sharp with increasing angular resolution, which is to be expected, since the cutoffs of the transmission profiles (transfer functions) are resolved in more detail, too, which allows the light distributions to follow the tensor field lines more strictly.
-This turns out to be the counterpart to the spatial resolution.
-
-Please note that all values in the test table where measured on following system:\\
-\texttt{4 x 16-Core Gold 6142M 2.60GHz, 3072 GB RAM}.\\
-The (dynamic) memory consumption was calculated with following formula:
-\begin{align*}
-	\mathop{mem}=8\mathop{dim}\mathop{steps}\cdot 8\mathop{byte}+(dim+19\cdot\mathop{steps})\cdot 8\mathop{byte}
-\end{align*}
-\begin{table}
-\small
-\centering
-\caption{Test Table LTG:}
-\begin{tabularx}{\columnwidth}{X|X|X|X}
-\hline\hline
-test name - Fig.~\# & resolution ($x\times y\times\omega$) & runtime $[s]$ & memory consumption (dynamic)$\approx$\\
-\toprule
-\enquote{drain} - \ref{ftle_base} & $101\times101\times64=652864$ & $179878$ & $47.02\mathop{MB}$\\ 
-\hline 
-\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times64=652864$ & $68787$ & $47.02\mathop{MB}$\\
-\hline 
-\enquote{random-local} - \ref{random_FTLE} & $101\times101\times32=326432$ & $59896$ & $23.5\mathop{MB}$\\ 
-\hline
-\enquote{random-global} - \ref{random_FTLE} & $101\times101\times32=326432$ & $757$ & $23.5\mathop{MB}$\\ 
-\hline 
-\enquote{isotropic} - \ref{bow-ftle} & $51\times51\times3283232$ & $1357$ & $5.998\mathop{MB}$\\ 
-\hline
-\enquote{rings} - \ref{bow-ftle} & $51\times51\times32=83232$ & $5761$ & $5.998\mathop{MB}$\\ 
-\hline 
-\enquote{bow} - \ref{bow-ftle} & $51\times51\times32=83232$ & $6128$ & $5.998\mathop{MB}$\\ 
-\hline 
-\enquote{star} - \ref{star-ftle} & $51\times51\times32=83232$ & $5815$ & $5.998\mathop{MB}$\\ 
-\hline 
-\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times16=163216$ & $12308$ & $11.754\mathop{MB}$\\ 
-\hline 
-\enquote{inverse} - \ref{inverse-ftle} & $101\times101\times32=326432$ & $24571$ & $23.5\mathop{MB}$\\ 
-\hline
-\enquote{brain} - \ref{brain-ftle} & $128\times128\times64=1048576$ & $2427$ & $75.5\mathop{MB}$\\
-\hline
-\enquote{heart} - \ref{heart-ftle} & $256\times256\times32=2097152$ & $117199$ & $151.0\mathop{MB}$\\
-\hline 
-\end{tabularx}
-   \vskip 1pt
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Conclusion and Future Work}
-\label{chap:concl}
-In this work, we have discussed the advantages and purposes of tensor field visualization in practical applications.
-Furthermore, we have presented a presented a technique to extract meaningful structures from tensor fields, which can, as a secondary purpose, also be used to visualize meaningful structures in $2D$ virtual geometrical scenes.
-To put it into context, we discussed its relations to vector field visualization techniques and other tensor field and light transport visualization methods.
-Also, we captured global illumination methods, since the proposed visualization method is derived from a global light transport method propagating intensities inside of a given Cartesian grid, consistent of cells defined by polar profiles (distributions), respectively.
-The intensities are redistributed stochastically by transmission profiles derived from the tensors in the field, used as windowing (weighting) functions.
-Based on this propagation algorithm, we can compute a FTLE-like field on the tensor field, as we obtain a final light distribution for each light source configuration (position and direction), which is considered as an energy-based flow map.
-In analogy to the computation of the FTLE field for vector fields, we analyze the gradient of the resulting flow map.
-The incongruity on the other hand, is that we need an additional inital direction on top of the inital position, because the tensors also dependent on the incoming direction, not as vectors only on their location $x,y$.
-As a result, we obtain a $3D$ scalar field, which can be visualized as volume (e.g.
-heatmap representation) to extract and segment ridges, i.e., key regions separating areas of correlating directional distributions.
-These ridges are in turn called Lagrangian coherent structures (LCS), which can be found in nature in , e.g., vortices in flow fields or moraines in glaciers.
-We designed comprehensive synthetic test tensor fields to evaluate our approach.
-Once we were convinced over its principal functionality, we applied the approach to some real data examples, namely DT-MRI scans of human brain and heart.
-Also we designed test runs to verify the physical plausibility of the results of the propagation scheme.
-The maps or fields we obtain, can make a statement about active regions in tensor fields, which do reveal the greatest tensor magnitude and anisotropy in combination.
-The approach is positioned in interesting relation to the works of Zirr et al.
-\cite{zirr}, which perform their FTPD (FTLE-like) computation by random sampling of trajectories to compute light transport separatrices.
-They do not incorporate any transmission weighting functions, but instead blocker (obstacle) or reflecting geometry, which causes a FTLE response as well, as edges or mirrors cause light paths to cross each other chaotically in similar manner.
-Furthermore, the work of Hlawatsch et al.
-\cite{hlawatsch} obtains an FTLE field on tensor fields in similar style, as they use a coordinate map to sample their trajectories.
-
-A possible application would be mechanical component design, as we could segment ROIs exhibiting increased loads or in a greater scope even machine component design, which are produced in series.
-The invented technique could also aid geologist in analyzing stress in cross-sections of the earth to be able to predict earthquakes sufficiently soon enough.
-In general one could measure vortices in directional distributions and hence tensor fields with a topological background (e.g.
-extracting separatrices of manifolds).
-One of the most relevant problems in visualization is incorporating uncertainty \cite{johnson}.
-Measurement, simulation or systematic (model) errors, boundary conditions or turbulences may introduce local uncertainty, which turns into global uncertainty through moving with the flow.
-This introduces propability density functions (PDFs) instead of certain single trajectories into the flow field.
-A real world example for uncertain vector fields is the measurement of blood flow and tissue velocity by phase contrast MRI, which yields an uncertain vector field due to the inherent noise in MRI scans.
-This uncertain vector field could be eventually visualized with our proposed technique.
-We just need to put it into the right context.
-The analogy to certain vector fields would be, that we could interpret our principal component ellipsoid as a propability density function and even replace it with such.
-The trajectories would be represented as propability distributions, which themselves are in turn propagated, just as we propagate our intensity profiles (distributions).
-As an outlook, we could apply our invented method to uncertain vector fields, which could be sampled in transmission profiles in discrete polar coordinates, just as we do in our work with the polar ellipsoids.
-Future works might include the previously mentioned application of the approach to uncertain vector fields or evaluating its applicability to real data in segmenting meaningful structures from engineering and medical data.
-Also we could imagine to implement a GPU parallelized version of the method to be able to propagate the intensities more efficiently, which would decrease the runtime in general.
-We can conclude, that tensor field visualization is becoming more and more important in scientific domains and that we invented a useful new tool in this context, which is worth of being investigated more precisely to reveal its potential in detecting meaningful information (key regions) and abstracting from redundancies in tensor fields.
-
-%Hier werden noch einmal die wichtigsten Ergebnisse und Erkenntnisse
-%der Arbeit zusammengefasst (nicht einfach eine Wiederholung des
-%Aufbaus der vorherigen Kapitel!), welche neuen Konzepte, Methoden und
-%Werkzeuge Neues entwickelt wurden, welche Probleme nun (effizienter)
-%gel?st werden k?nnen, und es wird ein Ausblick auf weiterf?hrende
-%Arbeiten gegeben (z.B.~was Sie machen w?rden, wenn Sie noch 6 Monate
-%mehr Zeit h?tten).
-
-% References (Literaturverzeichnis):
-% a) Style (with abbreviations: use alpha):
-% see
-% https://de.wikibooks.org/wiki/LaTeX-W%C3%B6rterbuch:_bibliographystyle
-% for the different formats and styles
-
-\bibliographystyle{ieeetr} %acm - as capital letter ALTERNATIVE
-% b) The File:
-\bibliography{references}
-
 \end{document}

### 2019-07-19 TensorFields/gyre.py
update

@@ -1,4 +1,4 @@
-import numpy
+import numpy as np
 import math as m
 import os
 

### 2019-07-19 TensorFields/random.py

@@ -1,145 +1,0 @@
-
-import numpy as np
-import math as m
-import os
-
-def multiply(X, Y): ## yields the matrix product X*Y
-    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
-    return result
-
-def rotate(matrix, phi):
-    xx = m.cos(phi) # cosine of phi in radian
-    xy = -m.sin(phi)
-    yx = m.sin(phi)
-    yy = m.cos(phi)
-    rotation = np.array([[xx, xy], [yx, yy]])
-
-    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
-
-def mirrorX(): # mirroring at x-axis
-    mirror = np.array([[-1, 0], [0, 1]])
-    return mirror
-
-def mirrorY(matrix): # mirroring at y-axis
-    mirror = np.array([[1, 0], [0, -1]])
-    return mirror
-
-## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
-
-def scale(xx,yy):
-    scale = np.array([[xx, 0], [0, yy]])
-    return scale
-
-def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    det = np.linalg.det(matrix)
-    if det != 0:
-        c = 1/np.sqrt(det)#*np.pi)
-    else:
-        c = 0
-
-    scale = np.array([[c, 0], [0, c]])
-    return np.matmul(scale, matrix)
-
-def random(): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
-    sigmaX = np.asscalar(np.random.uniform(-1,1,1))
-    sigmaY = np.asscalar(np.random.uniform(-1,1,1))
-    sigmaXY = np.asscalar(np.random.uniform(-1,1,1))
-    print("sigmaX: " + str(sigmaX)) # print filelen of temp: should equal length*length*matrixHeight
-    randomM = np.array([[sigmaX, sigmaXY], [sigmaXY, sigmaY]])
-    return randomM
-
-
-# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
-
-# calculater of txt file len
-def file_len(fname):
-    with open(fname) as f:
-        for i, l in enumerate(f):
-            pass
[…]

### 2019-07-19 TensorFields/randomField.py

@@ -1,0 +1,145 @@
+
+import numpy as np
+import math as m
+import os
+
+def multiply(X, Y): ## yields the matrix product X*Y
+    result = [[sum(a * b for a, b in zip(X_row, Y_col)) for Y_col in zip(*Y)] for X_row in X]
+    return result
+
+def rotate(matrix, phi):
+    xx = m.cos(phi) # cosine of phi in radian
+    xy = -m.sin(phi)
+    yx = m.sin(phi)
+    yy = m.cos(phi)
+    rotation = np.array([[xx, xy], [yx, yy]])
+
+    return np.matmul(rotation, matrix) ## rotation applied from the left (as last step)
+
+def mirrorX(): # mirroring at x-axis
+    mirror = np.array([[-1, 0], [0, 1]])
+    return mirror
+
+def mirrorY(matrix): # mirroring at y-axis
+    mirror = np.array([[1, 0], [0, -1]])
+    return mirror
+
+## begin with scaling and mirroring (in arbitrary order) to correctly interpret your results!, rotations on top...
+
+def scale(xx,yy):
+    scale = np.array([[xx, 0], [0, yy]])
+    return scale
+
+def normalize(matrix): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    det = np.linalg.det(matrix)
+    if det != 0:
+        c = 1/np.sqrt(det)#*np.pi)
+    else:
+        c = 0
+
+    scale = np.array([[c, 0], [0, c]])
+    return np.matmul(scale, matrix)
+
+def random(): ## normalize uniformly to max(det(tensor_field))*Pi for real data to capture relative change (instead of directional)
+    sigmaX = np.asscalar(np.random.uniform(-1,1,1))
+    sigmaY = np.asscalar(np.random.uniform(-1,1,1))
+    sigmaXY = np.asscalar(np.random.uniform(-1,1,1))
+    print("sigmaX: " + str(sigmaX)) # print filelen of temp: should equal length*length*matrixHeight
+    randomM = np.array([[sigmaX, sigmaXY], [sigmaXY, sigmaY]])
+    return randomM
+
+
+# --> create parameter to control intensity of light src until saturation - inverse mode (black/white) (1-I): "Negativ"
+
+# calculater of txt file len
+def file_len(fname):
+    with open(fname) as f:
+        for i, l in enumerate(f):
+            pass
[…]

### 2019-07-23 beamer2thesis/dummy.tex
update

@@
-\documentclass{scrartcl}
+\documentclass{beamer}
 \usepackage{tabularx}
@@
 
-\begin{table}[h]
-\scriptsize
-\centering
-\caption{Challenges in generative modeling:}
-\begin{tabularx}{\textwidth}{X|X|X}
-Global Illumination Methods & Symmetric Tensor Field Visualization & Asymmetric Tensor Field Visualization\\
-\toprule
+\frame{
+\frametitle{{Introduction - Scalar Fields}}
+\begin{columns}
+\begin{column}{.5\textwidth}
 \begin{itemize}
-	\item
+	\item Scalar fields are visualized by heat maps (color codings) classically
+	\bigskip
+	\item Each position in space is mapped a scalar height value
+	\bigskip
+	\item Examples: temperature field, height field
 \end{itemize}
-&
+\end{column}
+\begin{column}{.5\textwidth}
+\begin{figure}[t]
+\includegraphics[width=0.7\textwidth]{Scalar_field.png} 
+\caption{Scalar field,  \textit{Source: \textcircled{1}}}
+\end{figure}
+\end{column}
+\end{columns}
+
+} % END OF FRAME
+
+\frame{
+\frametitle{{Introduction - Vector Fields}}
+\begin{columns}
+\begin{column}{.5\textwidth}
 \begin{itemize}
-	\item
+	\item Vector fields are visualized by a collection of arrows with a given magnitude and direction classically
+	\bigskip
+	\item Each position in space is mapped a scalar magnitude and an angle
+	\bigskip
+	\item Examples: flow field, magnetic field, gravitational field
 \end{itemize}
-& 
+\end{column}
+\begin{column}{.5\textwidth}
+\begin{figure}[t]
+\includegraphics[width=0.6\textwidth]{vector_field.pdf} a)
+\includegraphics[width=0.6\textwidth]{vector_field2.pdf} b)
+\caption{Vector fields: a) $\{x,y\}$, b) $\{-y,x\}$}
+\end{figure}
+
+\end{column}
+\end{columns}
+
+} % END OF FRAME
+
+\frame{
+\frametitle{{Introduction - Tensor Fields}}
+\begin{columns}
+\begin{column}{.5\textwidth}
 \begin{itemize}
-	\item
+	\item Tensor fields are commonly visualized by:
+	\begin{itemize}
+		\item Glyphs
+		\item Tensor field lines (TFLs) $\Rightarrow$ Hyperstreamlines
 	\end{itemize}
-\hline 
-\end{tabularx}
-   \vskip 1pt
-\end{table}
+	\bigskip
+	\item Each position in space is mapped a tensor describing a directional distribution
+	\bigskip
+	\item Scalar Measures: anisotropy index, tensor magnitude
+\end{itemize}
+\end{column}
+\begin{column}{.5\textwidth}
+\begin{figure}[t]
+\includegraphics[width=0.6\textwidth]{hyperstreamlines.png} a)
+\includegraphics[width=0.6\textwidth]{random-global.png} b)
+\caption{Tensor fields: a) Tensor field lines, b) Glyphs}
+\end{figure}
+
+\end{column}
+\end{columns}
+
+} % END OF FRAME
+
+\frame{
+\frametitle{{Motivation - Tensor Fields}}
+Applications:
+\begin{itemize}
+	\item Vector fields: to describe the directionally dependent spatial gradient called Jacobian-matrix,
+	\item Fluid and solid continuum mechanics: to describe a whole distribution of stresses
+	\item DT-MRI: diffusion tensor - magnetic resonance imaging: to describe the diffusion characteristics of water molecules within tissue
+\end{itemize}
+} % END OF FRAME
 

### 2019-07-23 TensorFields/drain.py

@@ -46,7 +46,7 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101
+length = 13
 matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
 
 # generate normalized tensors from linear transformations

### 2019-07-23 TensorFields/drain_alt.py

@@ -46,7 +46,7 @@ def file_len(fname):
     return i + 1
 
 identity = np.array([[1, 0],[0, 1]]) ## row-major ordering: row-by-row
-length = 101
+length = 13
 matrixArray = np.ndarray(shape=(length,length), dtype=np.ndarray) # initialize ndarray w 0
 
 # generate normalized tensors from linear transformations
@@ -57,7 +57,7 @@ with open('temp.txt', 'wb') as f:
             x = (length - 1 - j)/(length-1)*round(i-(length-1)/2) # --> center around midpoint
             matrixArray[j][i] = normalize(np.array([[1, 0], [0, 1]])) # initialization w. normed identity
             rad = m.atan2(y,x) # use vector atan2 to determine angle and quadrant of vector [x|y]
-            scaled = np.matmul(scale(10,1),identity) # use anisotropic scaling
+            scaled = np.matmul(scale(4,1),identity) # use anisotropic scaling
             rotated = rotate(scaled, rad)
             normalized = normalize(rotated) # ..->transforms
             matrixArray[j][i] = normalized # update matrixArray entry

### 2019-07-23 beamer2thesis/dummy.tex
update

@@
 \frametitle{{Motivation - Tensor Fields}}
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+      \centering
+    \includegraphics[height=0.7\textwidth]{random-global.png}
+	a)
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+      \centering
+    \includegraphics[height=0.7\textwidth]{random_global.png}
+	b)
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+      \centering
+    \includegraphics[height=0.7\textwidth]{random_local_slice0.png}
+	c)
+    \label{b)}
+  \end{minipage}
+\caption{Random field: a) field, b) tensor magnitude, c) FTLE}
+\label{test_fields2}
+\end{figure}
 Applications:
@@
 	\item DT-MRI: diffusion tensor - magnetic resonance imaging: to describe the diffusion characteristics of water molecules within tissue
+\end{itemize}
+} % END OF FRAME
+
+\frame{
+\frametitle{{Random Test Field}}
+\begin{columns}
+\begin{column}{.5\textwidth}
+\begin{itemize}
+	\item Tensor fields are commonly visualized by:
+	\begin{itemize}
+		\item Glyphs
+		\item Tensor field lines (TFLs) $\Rightarrow$ Hyperstreamlines
+	\end{itemize}
+	\bigskip
+	\item Each position in space is mapped a tensor describing a directional distribution
+	\bigskip
+	\item Scalar Measures: anisotropy index, tensor magnitude
+\end{itemize}
+\end{column}
+\begin{column}{.5\textwidth}
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{random-global.png}
+	a)
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{random-local.png}
+	b)
+  \end{minipage}
+\label{bow_contour}
+\end{figure}
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{random_global.png}
+	c)
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{random_local_slice0.png}
+	d)
+  \end{minipage}
+    \caption{Random test field : a) Global, b) Local normalization, c) Tensor mag. d) LTG (FTLE)}
+\label{bow_contour}
+\end{figure}
+\end{column}
+\end{columns}
+
+} % END OF FRAME
+
+\frame{
+\frametitle{{Random Test Field - Evaluation}}
+\begin{columns}
+\begin{column}{.5\textwidth}
+\begin{itemize}
+	\item numerics: limit number representation to a particular range
+	\item preposition for relatively high tensor mags.: $\lambda_1\approx\lambda_2\approx1.0$ (because ellipsoid area equals $A=\lambda_1\cdot\lambda_2\cdot\pi$)\\
+\end{itemize}
+\end{column}
+\begin{column}{.5\textwidth}
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{random-global.png}
+	a)
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{random-local.png}
+	b)
+  \end{minipage}
+\label{bow_contour}
+\end{figure}
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{random_global.png}
+	c)
+  \end{minipage}
+  \begin{minipage}{0.4\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{random_local_slice0.png}
+	d)
+  \end{minipage}
+    \caption{Random test field : a) Global, b) Local normalization, c) Tensor mag. d) LTG (FTLE)}
+\label{bow_contour}
+\end{figure}
+\end{column}
+\end{columns}
+
+%$\Rightarrow$ chances are about $50\%$ for relatively low tensor magnitudes (ellipsoid areas)
+} % END OF FRAME
+
+\frame{
+\frametitle{{Heart dataset - Evaluation}}
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+  \centering
+  \includegraphics[width=\textwidth]{heartDwnsmpl.PNG}
+	a)
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{heart_slice20.PNG}
+	b)
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{heart_slice20_rescale.PNG}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{Heart test field: a) Glyphs, b) LTG local, c) LTG local (rescale)}
+\label{heart-ftle}
+\end{figure}
+\begin{itemize}
+	\item light ridges enhanced by rescaling for the down-left direction
+	\item these regions exhibit most drastic changes in anisotropy and/or divergence for TFLs
+	\item edge artifacts: edges effect high gradients for the flow map
+\end{itemize}
+} % END OF FRAME
+
+\frame{
+\frametitle{{Heart dataset - Evaluation}}
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+  \centering
+  \includegraphics[width=\textwidth]{heart_vcgRidges4.PNG}
+	a)
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{heart_isoVolume3.PNG}
+	b)
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{heart_contour.PNG}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{Heart test field: a) Glyphs, b) LTG local, c) LTG local (rescale)}
+\label{heart-ftle}
+\end{figure}
+\begin{itemize}
+	\item volumes reveal blob-shaped features
+	\item these features in turn form 3D ridges representing LCS
 \end{itemize}

### 2019-07-24 beamer2thesis/dummy.tex
update

--- 2019-07-18 beamer2thesis/dummy.tex
@@
-\documentclass{scrartcl}
+\documentclass[aspectratio=169]{beamer}
+\usetheme[language=ngerman,
+titlepagelogo=logopolito,
+bullet=circle,
+pageofpages=of,
+titleline=true,
+color=blue
+]{TorinoTh}
+
+%\usepackage[ngerman]{babel}
+%\usepackage[utf8]{inputenc}
 \usepackage{tabularx}
 \usepackage{booktabs}
-\usepackage{csquotes}
-% Include Graphic-files:
+\usepackage{multicol}
+\usepackage{ulem}
+\usepackage{makecell}
+\usepackage{upgreek}
+\usepackage{movie15}
+%\usepackage{isodate}
+
+\addto{\captionsngerman}{%
+  \renewcommand*{\contentsname}{Contents}
+  \renewcommand*{\listfigurename}{Figures}
+  \renewcommand*{\listtablename}{Tables}
+  \renewcommand*{\figurename}{Fig.}	
+  \renewcommand*{\tablename}{Tab.}
+}
+\newcommand*\mean[1]{\bar{#1}}
+\newcommand{\tabitem}{~~\llap{\textbullet}~~}
+\newcommand\widebar[1]{\mathop{\overline{#1}}}
+
+\usepackage{color}
 \usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subfig}
-\usepackage{url}
-%\usepackage{subfig}
-\newsubfloat{figure}
+\usepackage{fancybox}
+
+\usepackage{beamerthemesplit}
+\usetheme[compress]{Heidelberg}
+\definecolor{unirot}{rgb}{0.4,0.4,0.3} % babyblue 0,0.58,1
+\usecolortheme[named=unirot]{structure}
+%\setbeamercolor{alerted text}{fg=red}
+\newcommand*\hilite[1]{\textcolor{red}{#1}}
+%\def\hilite<#1>{%
+  %\temporal<#1>{\color{black}}{\color{unirot}}%
+               %{\color{gray}}}
+
+\title[Light Transport Techniques for Tensor Field Visualization]{Light Transport Techniques for Tensor Field Visualization}
+\subtitle{Master's Thesis Presentation}
+\vspace*{2pt}
+\author[Sebastian Bek]{Sebastian Bek\vspace*{7pt}}
+\date{July 24th 2019}
+\institute[Uni HD]{
+Heidelberg University\\
+Visual Computing Group (VCG)\\
+Supervisors: Prof. Filip Sadlo, Dr. Susanne Krömker\\
+}
+
+%\color{unirot}{sebibek@gmail.com}
+%\newsubfloat{figure}
 \newcommand{\source}[1]{\vspace{-3pt} \caption*{ Source: {#1}} }
+\setlength{\belowcaptionskip}{-10pt}
+\setlength{\intextsep}{-30pt}
 
@@
 
-\begin{table}[h]
-\scriptsize
+%\frame[plain]{\titlepage}
+
+\frame{
+\frametitle{{Light Transport Gradient - Heart dataset}}
+\begin{figure}[!t]
 \centering
-\caption{Challenges in generative modeling:}
-\begin{tabularx}{\textwidth}{X|X|X}
-Global Illumination Methods & Symmetric Tensor Field Visualization & Asymmetric Tensor Field Visualization\\
-\toprule
+  \begin{minipage}{0.3\textwidth}
+   \centering
+  \includegraphics[width=0.7\textwidth]{heartDwnsmpl.png}
+	a)
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+   \centering
+    \includegraphics[width=0.7\textwidth]{heartDwnsmpl-TFL.PNG}
+	b)
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+   \centering
+    \includegraphics[width=0.7\textwidth]{heart_ftle_org.PNG}
+	c)
+    \label{b)}
+  \end{minipage}
+\caption{Heart test field: a) glyphs, b) TFLs, c) FTLE: H.'s approach}
+\label{heart-ftle-comp}
+\end{figure}
+
 \begin{itemize}
-	\item
+	\item No noise present in the Heart dataset $\Rightarrow$ TFLs reveal correlated trajectories
+	\item FTLE response on structure obtained in the style of Hlawatsch et al.
 \end{itemize}
-&
+} % END OF FRAME
+
+\frame{
+\frametitle{{Heart dataset - Evaluation}}
+\begin{figure}[!t]
+\centering
+  \begin{minipage}{0.3\textwidth}
+  \centering
+  \includegraphics[height=0.9\textwidth]{heartDwnsmpl.PNG}
+	a)
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[width=0.9\textwidth]{heart_ftle_org.PNG}
+	b)
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}{0.3\textwidth}
+  \centering
+    \includegraphics[width=0.9\textwidth]{heart_slice20_rescale.PNG}
+	c)
+    \label{b)}
+  \end{minipage}
+  \caption{Heart test field: a) Glyphs, b) LTG local, c) LTG local (rescale)}
+\label{heart-ftle}
+\end{figure}
 \begin{itemize}
-	\item
+	\item light ridges enhanced by rescaling for the down-left direction (red lines in c))
+	\item regions exhibit most drastic changes in anisotropy and/or divergence for TFLs
+	\item edge artifacts: edges effect high gradients for the flow map
 \end{itemize}
-& 
-\begin{itemize}
-	\item
-\end{itemize}
-\hline 
-\end{tabularx}
-   \vskip 1pt
-\end{table}
+} % END OF FRAME
+
+%Volume non-directionally dependent
 

### 2019-07-25 beamer2thesis/dummy.tex
update

@@
-\documentclass[aspectratio=169]{beamer}
+\documentclass[aspectratio=169,urlcolor=black]{beamer}
 \usetheme[language=ngerman,
@@
 pageofpages=of,
-titleline=true,
-color=blue
+color=blue,
+titleline=true
 ]{TorinoTh}
@@
 \usepackage{movie15}
+%\usepackage{xcolor}
 %\usepackage{isodate}
-
+\hypersetup{colorlinks,linkcolor=black,urlcolor=black}
 \addto{\captionsngerman}{%
@@
 %\newsubfloat{figure}
-\newcommand{\source}[1]{\vspace{-3pt} \caption*{ Source: {#1}} }
+\newcommand{\source}[1]{\hspace{-3pt} {\tiny \raisebox{-0.75in}{\rotatebox[origin=t]{90}{Source: \color{black}{#1}}}}}
 \setlength{\belowcaptionskip}{-10pt}
-\setlength{\intextsep}{-30pt}
+%\setlength{\intextsep}{-30pt}
 
@@
 %\frame[plain]{\titlepage}
+\frame{
+\frametitle{{Finite-Time Lyapunov exponents}}
 
-\frame{
-\frametitle{{Light Transport Gradient - Heart dataset}}
-\begin{figure}[!t]
+\begin{columns}
+\begin{column}{.5\textwidth}
+\begin{small}
+\begin{block}{\centering \textbf{Definition}}
+Local: 
+$
+	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
+$\\
+Global: 
+$
+	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{\upvarphi(x)}\|_2
+$\\
+{\scriptsize whereas $\| A\| _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$}
+\end{block}
+\end{small}
+\begin{itemize}
+	\item measure for separation abilities of LTV (time-variant) systems concerning massless tracer particles
+	\item Lagrangian view in vector fields: placing two close-by particles into and moving with the flow
+\end{itemize}
+
+\end{column}
+\begin{column}{.5\textwidth}
+\begin{figure}[t]
+\begin{minipage}{0.85\textwidth}
 \centering
-  \begin{minipage}{0.3\textwidth}
-   \centering
-  \includegraphics[width=0.7\textwidth]{heartDwnsmpl.png}
-	a)
-    \label{a)}
+\includegraphics[height=0.6\textwidth]{gyre_rot.pdf} a)
+\includegraphics[height=0.6\textwidth]{ftle_sep.pdf} b)
+\caption{a) Vector field, b) diverging pathlines}
 \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-   \centering
-    \includegraphics[width=0.7\textwidth]{heartDwnsmpl-TFL.PNG}
-	b)
-    \label{b)}
+\begin{minipage}{0.1\textwidth}
+\source{b) Skript Prof. Sadlo, SciVis SS2017}
 \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-   \centering
-    \includegraphics[width=0.7\textwidth]{heart_ftle_org.PNG}
-	c)
-    \label{b)}
-  \end{minipage}
-\caption{Heart test field: a) glyphs, b) TFLs, c) FTLE: H.'s approach}
-\label{heart-ftle-comp}
+
 \end{figure}
 
-\begin{itemize}
-	\item No noise present in the Heart dataset $\Rightarrow$ TFLs reveal correlated trajectories
-	\item FTLE response on structure obtained in the style of Hlawatsch et al.
-\end{itemize}
+\end{column}
+\end{columns}
 } % END OF FRAME
@@
 \frame{
-\frametitle{{Heart dataset - Evaluation}}
-\begin{figure}[!t]
+\frametitle{{Finite-Time Lyapunov exponents}}
+
+\begin{columns}
+\begin{column}{.5\textwidth}
+\begin{small}
+\begin{block}{\centering \textbf{Definition}}
+Local: 
+$
+	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
+$\\
+Global: 
+$
+	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{\upvarphi(x)}\|_2
+$\\
+{\scriptsize whereas $\| A\| _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$}
+\end{block}
+\end{small}
+\begin{itemize}
+	\item measure for separation abilities of LTV (time-variant) systems concerning massless tracer particles
+	\item Lagrangian view in vector fields: placing two close-by particles into and moving with the flow
+\end{itemize}
+
+\end{column}
+\begin{column}{.5\textwidth}
+\begin{figure}[t]
 \centering
-  \begin{minipage}{0.3\textwidth}
-  \centering
-  \includegraphics[height=0.9\textwidth]{heartDwnsmpl.PNG}
-	a)
-    \label{a)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{heart_ftle_org.PNG}
-	b)
-    \label{b)}
-  \end{minipage}
-  \begin{minipage}{0.3\textwidth}
-  \centering
-    \includegraphics[width=0.9\textwidth]{heart_slice20_rescale.PNG}
-	c)
-    \label{b)}
-  \end{minipage}
-  \caption{Heart test field: a) Glyphs, b) LTG local, c) LTG local (rescale)}
-\label{heart-ftle}
+\includegraphics[height=0.35\textwidth]{gyre.pdf}~a)
+\includegraphics[height=0.35\textwidth]{ftle.png}~b)
+\caption{a) Vector field, b) FTLE field}
+
 \end{figure}
-\begin{itemize}
-	\item light ridges enhanced by rescaling for the down-left direction (red lines in c))
-	\item regions exhibit most drastic changes in anisotropy and/or divergence for TFLs
-	\item edge artifacts: edges effect high gradients for the flow map
-\end{itemize}
+
+\end{column}
+\end{columns}
 } % END OF FRAME
 
+%glyphs are used to represent the eigensystem, i.e., the anisotropy characteristics and/or tensor magnitude
 %Volume non-directionally dependent

### 2019-07-28 beamer2thesis/dummy.tex
update

@@
 \usepackage{fancybox}
+\usepackage[singlespacing]{setspace}
 
@@
 \setlength{\belowcaptionskip}{-10pt}
-%\setlength{\intextsep}{-30pt}
+\setlength{\intextsep}{-20pt}
 
@@
 %\frame[plain]{\titlepage}
-\frame{
-\frametitle{{Finite-Time Lyapunov exponents}}
-
-\begin{columns}
-\begin{column}{.5\textwidth}
-\begin{small}
-\begin{block}{\centering \textbf{Definition}}
-Local: 
-$
-	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
-$\\
-Global: 
-$
-	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{\upvarphi(x)}\|_2
-$\\
-{\scriptsize whereas $\| A\| _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$}
-\end{block}
-\end{small}
-\begin{itemize}
-	\item measure for separation abilities of LTV (time-variant) systems concerning massless tracer particles
-	\item Lagrangian view in vector fields: placing two close-by particles into and moving with the flow
-\end{itemize}
-
-\end{column}
-\begin{column}{.5\textwidth}
-\begin{figure}[t]
-\begin{minipage}{0.85\textwidth}
-\centering
-\includegraphics[height=0.6\textwidth]{gyre_rot.pdf} a)
-\includegraphics[height=0.6\textwidth]{ftle_sep.pdf} b)
-\caption{a) Vector field, b) diverging pathlines}
-\end{minipage}
-\begin{minipage}{0.1\textwidth}
-\source{b) Skript Prof. Sadlo, SciVis SS2017}
-\end{minipage}
-
-\end{figure}
-
-\end{column}
-\end{columns}
-} % END OF FRAME
 
 \frame{
-\frametitle{{Finite-Time Lyapunov exponents}}
-
-\begin{columns}
-\begin{column}{.5\textwidth}
-\begin{small}
-\begin{block}{\centering \textbf{Definition}}
-Local: 
-$
-	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
-$\\
-Global: 
-$
-	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{\upvarphi(x)}\|_2
-$\\
-{\scriptsize whereas $\| A\| _2 = \sqrt{\lambda_{max}(A^TA)}$ is the spectral norm of matrix $A$}
+\frametitle{{Propagation Scheme - Physical Model}}
+\begin{block}{\centering \textbf{Crystal Fiber Structures}}
+\begin{itemize}
+	\item crystal lattices reveal gradients dependent on direction, which leads to anisotropic light transport inside the medium (birefringence)
+	\item the propability for redistribution of a photon in a particular direction is given by the phase function model yielded from Rayleigh scattering:
+	\begin{align*}
+ 	P(\omega) &= \frac{T(\omega)}{\int_{k\pi}^{(k+1)\pi}T(\omega)\mathop{d\omega}}\\
+ 	\mathit{with} \int_0^{\pi} P(\omega) &= 1.
+	\end{align*}
+	\item each tensor is then a unique footprint: $t_{i,j}\mapsto T(\omega)$
+\end{itemize}
 \end{block}
-\end{small}
-\begin{itemize}
-	\item measure for separation abilities of LTV (time-variant) systems concerning massless tracer particles
-	\item Lagrangian view in vector fields: placing two close-by particles into and moving with the flow
-\end{itemize}
-
-\end{column}
-\begin{column}{.5\textwidth}
-\begin{figure}[t]
-\centering
-\includegraphics[height=0.35\textwidth]{gyre.pdf}~a)
-\includegraphics[height=0.35\textwidth]{ftle.png}~b)
-\caption{a) Vector field, b) FTLE field}
-
-\end{figure}
-
-\end{column}
-\end{columns}
 } % END OF FRAME

### 2019-07-30 beamer2thesis/dummy.tex
update

@@
 \newcommand\widebar[1]{\mathop{\overline{#1}}}
+\usepackage{caption}
+\captionsetup{font=scriptsize}
 
@@
 \setlength{\intextsep}{-20pt}
-
+%\usepackage[style=verbose,backend=biber]{biblatex}
+%\addbibresource{references.bib}
+%\usepackage[multiple]{footmisc}
+%\usepackage[bottom]{footmisc}
+\renewcommand\thempfootnote{\arabic{mpfootnote}}
 \begin{document}
@@
 \frame{
-\frametitle{{Propagation Scheme - Physical Model}}
-\begin{block}{\centering \textbf{Crystal Fiber Structures}}
+\frametitle{{Eigenanalysis - Tensor Fields}}
+\begin{columns}
+\begin{column}{.6\textwidth}
+Direction, invariant under transformation $A$, sought:
+\begin{block}{\centering \textbf{Eigenvalue Problem}}
+\begin{align*}
+	A \cdot \boldsymbol{\epsilon} = \lambda \boldsymbol{\epsilon}
+\end{align*}
+with: $A$: square matrix, $\lambda$: eigenvalue $\mathbf{\epsilon}$: eigenvector
+\end{block}
 \begin{itemize}
-	\item crystal lattices reveal gradients dependent on direction, which leads to anisotropic light transport inside the medium (birefringence)
-	\item the propability for redistribution of a photon in a particular direction is given by the phase function model yielded from Rayleigh scattering:
-	\begin{align*}
- 	P(\omega) &= \frac{T(\omega)}{\int_{k\pi}^{(k+1)\pi}T(\omega)\mathop{d\omega}}\\
- 	\mathit{with} \int_0^{\pi} P(\omega) &= 1.
-	\end{align*}
-	\item each tensor is then a unique footprint: $t_{i,j}\mapsto T(\omega)$
+	\item eigensystems:\\ full set of eigenvalues and -vectors
+	\item we propose to use singular value systems (SVS) in analogy to Eigensystems\\
+	{\tiny Proof: "Glyphs for General Second-Order 2D and 3D Tensors", Gerrits et al., 2017}
 \end{itemize}
+\end{column}
+\begin{column}{.4\textwidth}
+\begin{figure}[t]
+\begin{minipage}{0.85\textwidth}
+\includegraphics[width=\textwidth]{HAT_edit2.png}
+\caption*{Eigensystem: characteristic ellipsoid}
+\end{minipage}
+\begin{minipage}{0.1\textwidth}
+\source{\url{https://slideplayer.com/slide/5291764}}
+\end{minipage}
+
+\end{figure}
+
+\end{column}
+\end{columns}
+} % END OF FRAME
+
+\frame{
+\frametitle{{Computation of Eigensystem - Tensor Fields}}
+
+\begin{columns}
+\begin{column}{.6\textwidth}
+\begin{small}
+\begin{block}{\centering \textbf{Matrix Decompositions}}
+1) Eigenvalue Decomposition: \hskip 46pt
+$
+	A = {R}{S}{R}^*
+$\\
+2) Singular Value Decomposition (SVD): 
+$
+	{A} = {U} {\Sigma} {V}^*
+$
 \end{block}
+\end{small}
+\begin{small}
+\begin{enumerate}
+	\item eigenvalues \hskip 14pt and eigenvectors
+	\item singular values and singular vectors (SVs)
+\end{enumerate}
+\begin{itemize}
+	\item SVs represent the axes of characteristic ellipsoid \footnote{Gerrits et al.
+''Glyphs for General Second-Order 2D and 3D Tensors", IEEE Transactions on Visualization and Computer Graphics, 2017.}
+	\item singular values ($s_i=\sqrt{\lvert\lambda_i\rvert}$)\footnote{M. Kieburg, “What is the Relation between Eigenvalues \& Singular
+Values?, 2016} are lengths of axes
+\end{itemize}
+\end{small}
+
+\end{column}
+\begin{column}{.4\textwidth}
+\begin{figure}[t]
+\begin{minipage}{0.9\textwidth}
+\includegraphics[width=\textwidth]{SVD.png}
+\caption*{Singular value system}
+\end{minipage}
+\begin{minipage}{0.05\textwidth}
+\source{\url{https://w.wiki/6WN}}
+\end{minipage}
+\end{figure}
+
+\end{column}
+\end{columns}
 } % END OF FRAME

### 2019-07-31 beamer2thesis/dummy.tex
update

--- 2019-07-28 beamer2thesis/dummy.tex
@@
 
+\renewcommand\thempfootnote{\arabic{mpfootnote}}
 \begin{document}
@@
 \frame{
-\frametitle{{Propagation Scheme - Physical Model}}
-\begin{block}{\centering \textbf{Crystal Fiber Structures}}
+\frametitle{{Tensor Fields - Tensor Field Lines}}
+
+\begin{columns}
+\begin{column}{.6\textwidth}
+\begin{small}
+\begin{block}{\centering \textbf{Procedure}}
+follow major/minor eigenvectors along pathlines
+\end{block}
+\end{small}
+\bigskip
+\begin{small}
 \begin{itemize}
-	\item crystal lattices reveal gradients dependent on direction, which leads to anisotropic light transport inside the medium (birefringence)
-	\item the propability for redistribution of a photon in a particular direction is given by the phase function model yielded from Rayleigh scattering:
-	\begin{align*}
- 	P(\omega) &= \frac{T(\omega)}{\int_{k\pi}^{(k+1)\pi}T(\omega)\mathop{d\omega}}\\
- 	\mathit{with} \int_0^{\pi} P(\omega) &= 1.
-	\end{align*}
-	\item each tensor is then a unique footprint: $t_{i,j}\mapsto T(\omega)$
+	\item ambiguity for nearly isotropic cases:\\ small changes in components effect large changes in tensor field line orientation (see Fig.)
+	\item therefore susceptive to noise in data
+	\item trajectories can cross each other\\
+	$\Rightarrow$ effects erroneous trajectories, hard to interpret
 \end{itemize}
-\end{block}
+\end{small}
+
+\end{column}
+\begin{column}{.4\textwidth}
+\centering
+\begin{figure}[t]
+\centering
+\includegraphics[height=0.8\textwidth]{nearly_isotropic.pdf}
+\caption*{Ambiguity}
+\end{figure}
+
+\end{column}
+\end{columns}
 } % END OF FRAME

### 2019-07-31 beamer2thesis/dummy.tex
update

@@
 
+\setlength{\skip\footins}{0.1cm}
+\setlength{\footnotesep}{0.1cm}
 \renewcommand\thempfootnote{\arabic{mpfootnote}}
@@
 \frame{
-\frametitle{{Tensor Fields - Tensor Field Lines}}
-
+\frametitle{{Propagation Scheme - Template}}
 \begin{columns}
 \begin{column}{.6\textwidth}
-\begin{small}
-\begin{block}{\centering \textbf{Procedure}}
-follow major/minor eigenvectors along pathlines
-\end{block}
-\end{small}
-\bigskip
-\begin{small}
+To interpret eigensystems of tensors as conductivity property for light transport, we will need a propagation scheme:\bigskip
 \begin{itemize}
-	\item ambiguity for nearly isotropic cases:\\ small changes in components effect large changes in tensor field line orientation (see Fig.)
-	\item therefore susceptive to noise in data
-	\item trajectories can cross each other\\
-	$\Rightarrow$ effects erroneous trajectories, hard to interpret
+	\item using light propagation volumes (LPV) similar to Dachsbacher et al. (cf. Fig.)\bigskip
+	\item exchanging itensities between neighboring cells: $\Phi_{t} = \int I(\omega)\mathop{d\omega}$\\
+	{\small with $\Phi_{t}$: radiant flux, $I(\omega)$: radiant intensity}
 \end{itemize}
-\end{small}
 
@@
 \begin{column}{.4\textwidth}
-\centering
 \begin{figure}[t]
 \centering
-\includegraphics[height=0.8\textwidth]{nearly_isotropic.pdf}
-\caption*{Ambiguity}
+\begin{minipage}{0.79\textwidth}
+\centering
+\includegraphics[width=0.7\textwidth]{LPV.pdf}
+\includegraphics[width=0.9\textwidth]{LPV2.pdf}
+\caption*{Light Propagation Volumes}
+\end{minipage}
+\begin{minipage}{0.15\textwidth}
+\source{\textit{A. Kaplanyan and C. Dachsbacher, ''Cascaded light}} 
+\hspace{-3pt} {\tiny \raisebox{-0.75in}{\rotatebox[origin=t]{90}{propagation volumes for realtime indirect illumination"}}}
+\end{minipage}
 \end{figure}

### 2019-07-31 beamer2thesis/dummy.tex
update

--- 2019-07-31 beamer2thesis/dummy.tex
@@
 \frame{
-\frametitle{{Tensor Fields - Tensor Field Lines}}
-
+\frametitle{{Random Test Field}}
 \begin{columns}
-\begin{column}{.6\textwidth}
-\begin{small}
-\begin{block}{\centering \textbf{Procedure}}
-follow major/minor eigenvectors along pathlines
-\end{block}
-\end{small}
-\bigskip
+\begin{column}{.5\textwidth}
 \begin{small}
 \begin{itemize}
-	\item ambiguity for nearly isotropic cases:\\ small changes in components effect large changes in tensor field line orientation (see Fig.)
-	\item therefore susceptive to noise in data
-	\item trajectories can cross each other\\
-	$\Rightarrow$ effects erroneous trajectories, hard to interpret
+	\item numerics: limit number representation to a particular range
+	\bigskip
+	\item characteristic ellipsoid area: $A=\sigma_1\cdot\sigma_2\cdot\pi$
+	$\Rightarrow$ preposition for high tensor magnitudes: $\sigma_1\approx\sigma_2\approx1.0$ (implicit spherical isotropy)\bigskip
+	\item local normalization is verified to emphasize regions with increased anisotropy
 \end{itemize}
 \end{small}
-
 \end{column}
-\begin{column}{.4\textwidth}
+\begin{column}{.5\textwidth}
+\begin{figure}[!t]
 \centering
-\begin{figure}[t]
-\centering
-\includegraphics[height=0.8\textwidth]{nearly_isotropic.pdf}
-\caption*{Ambiguity}
+    \includegraphics[width=0.33\textwidth]{random-global.png} a)
+    \includegraphics[width=0.33\textwidth]{random-local.png} b)
+    \includegraphics[width=0.33\textwidth]{random_global.png} c)
+    \includegraphics[width=0.33\textwidth]{random_local_slice0.png}	d)
+    \caption*{Random test field : a) Glyphs, b) Local normalization, c) Tensor magnitude d) LTG}
+\label{bow_contour}
 \end{figure}
-
+%\begin{figure}[!t]
+%\centering
+%    
+%    
+%\label{bow_contour}
+%\end{figure}
 \end{column}
 \end{columns}
+
 } % END OF FRAME

### 2019-08-01 beamer2thesis/dummy.tex
update

--- 2019-07-31 beamer2thesis/dummy.tex
@@
 \frame{
-\frametitle{{Tensor Fields - Tensor Field Lines}}
+\frametitle{{Propagation Scheme - Physical Models}}
+\begin{itemize}
+	\item phase function model from Rayleigh scattering:
+\begin{block}{\centering \textbf{Phase Functions\footnote[frame]{\url{http://www.met.reading.ac.uk/~jp902366/rad/phase_function_report.pdf}\bigskip}}}
+	\begin{align*}
+ 	P(\omega) &= \frac{T(\omega)}{\int_{k\pi}^{(k+1)\pi}T(\omega)\mathop{d\omega}}\\
+ 	\mathit{with} \int_0^{\pi} P(\omega) &= 1.
+	\end{align*}
+\end{block}
+	\item each tensor is then a unique footprint: $t_{i,j}\mapsto T(\omega)$
+\end{itemize}
 
-\begin{columns}
-\begin{column}{.6\textwidth}
-\begin{small}
-\begin{block}{\centering \textbf{Procedure}}
-follow major/minor eigenvectors along pathlines
-\end{block}
-\end{small}
-\bigskip
-\begin{small}
+} % END OF FRAME
+
+\frame{
+\frametitle{{Propagation Scheme - Physical Models}}
 \begin{itemize}
-	\item ambiguity for nearly isotropic cases:\\ small changes in components effect large changes in tensor field line orientation (see Fig.)
-	\item therefore susceptive to noise in data
-	\item trajectories can cross each other\\
-	$\Rightarrow$ effects erroneous trajectories, hard to interpret
+	\item crystal lattices reveal refractive indices dependent on direction, which lead to anisotropic light transport inside the medium (birefringence: e.g. quartz, ruby):
+\begin{block}{\centering \textbf{Indicatrix\footnote[frame]{\url{http://edafologia.ugr.es/optmine/intro/indicatw.htm}} - Refractive Ellipsoid}}
+	\begin{align*}
+	T(\omega) = \frac{ab}{\sqrt{a^{2}\sin^{2}(\omega)+b^{2}\cos^{2}(\omega)}} =\frac{n_1n_2}{\sqrt{n_1^2\sin^2(\omega)+n_2^2\cos^2(\omega)}}
+\end{align*}
+\end{block}\bigskip
+	\item each tensor is then a unique footprint: $t_{i,j}\mapsto T(\omega)$
 \end{itemize}
-\end{small}
 
-\end{column}
-\begin{column}{.4\textwidth}
-\centering
-\begin{figure}[t]
-\centering
-\includegraphics[height=0.8\textwidth]{nearly_isotropic.pdf}
-\caption*{Ambiguity}
-\end{figure}
-
-\end{column}
-\end{columns}
 } % END OF FRAME

### 2019-08-05 beamer2thesis/dummy.tex
update

--- 2019-07-31 beamer2thesis/dummy.tex
@@
 
-%\frame[plain]{\titlepage}
-
+\frame[plain]{\titlepage}
 \frame{
-\frametitle{{Propagation Scheme - Template}}
-\begin{columns}
-\begin{column}{.6\textwidth}
-To interpret eigensystems of tensors as conductivity property for light transport, we will need a propagation scheme:\bigskip
-\begin{itemize}
-	\item using light propagation volumes (LPV) similar to Dachsbacher et al. (cf. Fig.)\bigskip
-	\item exchanging itensities between neighboring cells: $\Phi_{t} = \int I(\omega)\mathop{d\omega}$\\
-	{\small with $\Phi_{t}$: radiant flux, $I(\omega)$: radiant intensity}
-\end{itemize}
-
-\end{column}
-\begin{column}{.4\textwidth}
+\frametitle{{Light Transport Gradient - Scheme}}
 \begin{figure}[t]
 \centering
-\begin{minipage}{0.79\textwidth}
+  \begin{minipage}{0.6\textwidth}
   \centering
-\includegraphics[width=0.7\textwidth]{LPV.pdf}
-\includegraphics[width=0.9\textwidth]{LPV2.pdf}
-\caption*{Light Propagation Volumes}
-\end{minipage}
-\begin{minipage}{0.15\textwidth}
-\source{\textit{A. Kaplanyan and C. Dachsbacher, ''Cascaded light}} 
-\hspace{-3pt} {\tiny \raisebox{-0.75in}{\rotatebox[origin=t]{90}{propagation volumes for realtime indirect illumination"}}}
+	\includegraphics[height=0.7\textwidth]{skizze/addins/grids_one_base_edit.pdf}
+	\caption*{Light Transport Gradient - Domain of Light Distributions}
 	\end{minipage}
@@
 
-\end{column}
-\end{columns}
+} % END OF FRAME
+
+\frame{
+\frametitle{{Light Transport Gradient - Scheme}}
+\begin{figure}[t]
+\centering
+  \begin{minipage}{0.6\textwidth}
+  \centering
+	\includegraphics[height=0.7\textwidth]{skizze/addins/grids_one_edit.pdf}
+	\caption*{Light Transport Gradient - Domain of Light Distributions}
+	\end{minipage}
+\end{figure}
+
+} % END OF FRAME
+
+\frame{
+\frametitle{{Light Transport Gradient - Scheme}}
+\begin{figure}[t]
+\centering
+  \begin{minipage}{0.5\textwidth}
+  \centering
+	\includegraphics[height=0.85\textwidth]{skizze/addins/grid_two_base.pdf}
+	\caption*{Light Transport Gradient - Neighborhood}
+	\end{minipage}
+\end{figure}
+
+} % END OF FRAME
+\frame{
+\frametitle{{Light Transport Gradient - Scheme}}
+\begin{figure}[t]
+\centering
+  \begin{minipage}{0.5\textwidth}
+  \centering
+	\includegraphics[height=0.85\textwidth]{skizze/addins/grids_two_edit.pdf}
+	\caption*{Light Transport Gradient - Neighborhood}
+	\end{minipage}
+\end{figure}
+
 } % END OF FRAME

### 2019-08-06 beamer2thesis/dummy.tex
update

@@
 \frame[plain]{\titlepage}
+
 \frame{
-\frametitle{{Light Transport Gradient - Scheme}}
+\frametitle{{Fundamentals - Finite-Time Lyapunov exponents}}
+
+\begin{columns}
+\begin{column}{.5\textwidth}
+\begin{small}
+\begin{block}{\centering \textbf{Definition}}
+Local: 
+$
+	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
+$\\
+Global: 
+$
+	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{u(x)}\|_2
+$\\
+{\scriptsize where $\| A\| _2 = \sqrt{\lambda_{\max}(A^TA)}$ is the spectral norm of matrix $A$}
+\end{block}
+\end{small}
+\begin{itemize}
+	\item measures gradient of flow map $\nabla\mathbf{u(x)}$
+	\item FTLE field responds largest where path lines diverge
+\end{itemize}
+
+\end{column}
+\begin{column}{.5\textwidth}
 \begin{figure}[t]
 \centering
-  \begin{minipage}{0.6\textwidth}
+\caption*{Double Gyre Field}\bigskip
+\begin{minipage}{\textwidth}
+{\scriptsize a)}
 \centering
-	\includegraphics[height=0.7\textwidth]{skizze/addins/grids_one_base_edit.pdf}
-	\caption*{Light Transport Gradient - Domain of Light Distributions}
+\includegraphics[height=0.25\textwidth]{diverge.png}
+\end{minipage}
+\begin{minipage}{\textwidth}
+{\scriptsize b)}
+\centering
+\includegraphics[height=0.25\textwidth]{ftle.png}
+\caption*{a) Vector field, b) FTLE field}
+\end{minipage}
+\begin{minipage}{0.05\textwidth}
+\centering
+\source{http://tiny.cc/2zbjaz}
 \end{minipage}
@@
 
+\end{column}
+\end{columns}
 } % END OF FRAME
 
-\frame{
-\frametitle{{Light Transport Gradient - Scheme}}
-\begin{figure}[t]
-\centering
-  \begin{minipage}{0.6\textwidth}
-  \centering
-	\includegraphics[height=0.7\textwidth]{skizze/addins/grids_one_edit.pdf}
-	\caption*{Light Transport Gradient - Domain of Light Distributions}
-	\end{minipage}
-\end{figure}
-
-} % END OF FRAME
-
-\frame{
-\frametitle{{Light Transport Gradient - Scheme}}
-\begin{figure}[t]
-\centering
-  \begin{minipage}{0.5\textwidth}
-  \centering
-	\includegraphics[height=0.85\textwidth]{skizze/addins/grid_two_base.pdf}
-	\caption*{Light Transport Gradient - Neighborhood}
-	\end{minipage}
-\end{figure}
-
-} % END OF FRAME
-\frame{
-\frametitle{{Light Transport Gradient - Scheme}}
-\begin{figure}[t]
-\centering
-  \begin{minipage}{0.5\textwidth}
-  \centering
-	\includegraphics[height=0.85\textwidth]{skizze/addins/grids_two_edit.pdf}
-	\caption*{Light Transport Gradient - Neighborhood}
-	\end{minipage}
-\end{figure}
-
-} % END OF FRAME
+%\frame{
+%\frametitle{{Light Transport Gradient - Scheme}}
+%\begin{figure}[t]
+%\centering
+%  \begin{minipage}{0.6\textwidth}
+%  \centering
+%	\includegraphics[height=0.7\textwidth]{skizze/addins/grids_one_base_edit.pdf}
+%	\caption*{Light Transport Gradient - Domain of Light Distributions}
+%	\end{minipage}
+%\end{figure}
+%
+%} % END OF FRAME
+%
+%\frame{
+%\frametitle{{Light Transport Gradient - Scheme}}
+%\begin{figure}[t]
+%\centering
+%  \begin{minipage}{0.6\textwidth}
+%  \centering
+%	\includegraphics[height=0.7\textwidth]{skizze/addins/grids_one_edit.pdf}
+%	\caption*{Light Transport Gradient - Domain of Light Distributions}
+%	\end{minipage}
+%\end{figure}
+%
+%} % END OF FRAME
+%
+%
+%\frame{
+%\frametitle{{Light Transport Gradient - Scheme}}
+%\begin{figure}[t]
+%\centering
+%  \begin{minipage}{0.5\textwidth}
+%  \centering
+%	\includegraphics[height=0.85\textwidth]{skizze/addins/grid_two_base.pdf}
+%	\caption*{Light Transport Gradient - Neighborhood}
+%	\end{minipage}
+%\end{figure}	
+%
+%} % END OF FRAME
+%\frame{
+%\frametitle{{Light Transport Gradient - Scheme}}
+%\begin{figure}[t]
+%\centering
+%  \begin{minipage}{0.5\textwidth}
+%  \centering
+%	\includegraphics[height=0.85\textwidth]{skizze/addins/grids_two_edit.pdf}
+%	\caption*{Light Transport Gradient - Neighborhood}
+%	\end{minipage}
+%\end{figure}
+%
+%} % END OF FRAME
 

### 2019-08-07 beamer2thesis/dummy.tex
update

@@
 \frame{
-\frametitle{{Fundamentals - Finite-Time Lyapunov exponents}}
-
+\frametitle{{Introduction - Tensor Fields}}
 \begin{columns}
 \begin{column}{.5\textwidth}
-\begin{small}
-\begin{block}{\centering \textbf{Definition}}
-Local: 
-$
-	\sigma = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
-$\\
-Global: 
-$
-	\sigma(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{u(x)}\|_2
-$\\
-{\scriptsize where $\| A\| _2 = \sqrt{\lambda_{\max}(A^TA)}$ is the spectral norm of matrix $A$}
-\end{block}
-\end{small}
 \begin{itemize}
-	\item measures gradient of flow map $\nabla\mathbf{u(x)}$
-	\item FTLE field responds largest where path lines diverge
+	\item tensor-valued function: $\mathop{t_{ij}\mathbf{(x)}\colon} \mathbf{x}\mapsto t$, with $\mathbf{x}\in\mathcal{R}^n$: location vector
+	\bigskip
+	\item tensor fields can be visualized by:
+	\begin{itemize}
+		\item a) glyphs
+		\item b) tensor field lines (TFLs)\\ $\Rightarrow$ streamlines for eigenvector fields
+		\item c) scalar measures:\\ anisotropy index, tensor magnitude
 	\end{itemize}
+	\bigskip
+	\item examples: stress tensor
 
+\end{itemize}
 \end{column}
@@
 \centering
-\caption*{Double Gyre Field}\bigskip
-\begin{minipage}{\textwidth}
-{\scriptsize a)}
+\caption*{\textit{Random tensor field}}\bigskip
+\begin{minipage}{0.49\textwidth}
 \centering
-\includegraphics[height=0.25\textwidth]{diverge.png}
+\includegraphics[height=\textwidth]{random-global.png} {\scriptsize a) Glyphs}
 \end{minipage}
-\begin{minipage}{\textwidth}
-{\scriptsize b)}
+\begin{minipage}{0.49\textwidth}
 \centering
-\includegraphics[height=0.25\textwidth]{ftle.png}
-\caption*{a) Vector field, b) FTLE field}
-\end{minipage}
-\begin{minipage}{0.05\textwidth}
-\centering
-\source{http://tiny.cc/2zbjaz}
+\includegraphics[height=\textwidth]{random-global-TFL.png} {\scriptsize b) Tensor field lines}
 \end{minipage}
@@
 \end{columns}
+
 } % END OF FRAME

### 2019-08-09 beamer2thesis/dummy.tex
update

@@
 \frame{
-\frametitle{{Introduction - Tensor Fields}}
+\frametitle{{Light Transport Gradient (LTG) - Isotropic field}}
 \begin{columns}
 \begin{column}{.5\textwidth}
+\begin{small}
 \begin{itemize}
-	\item tensor-valued function: $\mathop{t_{ij}\mathbf{(x)}\colon} \mathbf{x}\mapsto t$, with $\mathbf{x}\in\mathcal{R}^n$: location vector
+	\item Isotropic test field reveals offset response for the approach
 	\bigskip
-	\item tensor fields can be visualized by:
-	\begin{itemize}
-		\item a) glyphs
-		\item b) tensor field lines (TFLs)\\ $\Rightarrow$ streamlines for eigenvector fields
-		\item c) scalar measures:\\ anisotropy index, tensor magnitude
+	\item natural shift against the inital light source (slice) direction (left)\\
+%	$\Rightarrow$ initial light source configurations located close and pointing towards borders, are biased for low LTG (FTLE) responses
 \end{itemize}
-	\bigskip
-	\item examples: stress tensor
-
-\end{itemize}
+\end{small}
 \end{column}
@@
 \begin{figure}[t]
-\centering
-\caption*{\textit{Random tensor field}}\bigskip
-\begin{minipage}{0.49\textwidth}
-\centering
-\includegraphics[height=\textwidth]{random-global.png} {\scriptsize a) Glyphs}
+\caption*{\textit{Isotropic test field}}\bigskip
+\begin{minipage}{.49\textwidth}
+\includegraphics[height=\textwidth]{isotropic2.png} \caption*{a)~Glyphs}
 \end{minipage}
-\begin{minipage}{0.49\textwidth}
-\centering
-\includegraphics[height=\textwidth]{random-global-TFL.png} {\scriptsize b) Tensor field lines}
+\begin{minipage}{.49\textwidth}
+\includegraphics[height=\textwidth]{iso_ftle.png} \caption*{b)~LTG($0^\circ$)}
 \end{minipage}
+
 \end{figure}

### 2019-08-11 beamer2thesis/dummy.tex
update

@@
 \frame{
-\frametitle{{Light Transport Gradient (LTG) - Isotropic field}}
-\begin{columns}
-\begin{column}{.5\textwidth}
+\frametitle{{Related Work - Symmetric Tensor Field Visualization}}
+
 \begin{small}
 \begin{itemize}
-	\item Isotropic test field reveals offset response for the approach
-	\bigskip
-	\item natural shift against the inital light source (slice) direction (left)\\
-%	$\Rightarrow$ initial light source configurations located close and pointing towards borders, are biased for low LTG (FTLE) responses
+	\item glyphs\footnote[frame]{McLoughlin et al. ''Over two decades of integration-based, geometric flow visualization", 2010}: represent anisotropy with shape and orientation
+	\item tensor field lines\footnote[frame]{Vilanova et al. ''Dti visualization with streamsurfaces beaand evenly-spaced volume seeding", 2004} (TFLs): follow tensor field along major eigenvector
+	\item tensor field topology\footnote[frame]{Delmarcelle and Hesselink, “Visualizing second order tensor fields with hyperstreamlines”}: e.g. separatrices, degenerate points
+	\item tensorlines\footnote[frame]{Weinstein et al. ''Tensorlines: Advection-diffusion
+based propagation through diffusion tensor fields", 1999}: introduce artificial inertia to TFLs to increase stability
+	\item HyperLIC\footnote[frame]{Zheng and Pang ''HyperLIC", 2003}: use Line Integral Convolution from vector field visualization on TFLs
+	\item FTLE\footnote[frame]{Hlawatsch et al. ''Coherent structures of
+characteristic curves in symmetric second order tensor fields", 2011}: exploit the gradient of the flow map of TFLs to generate an FTLE field
+	\item scalar measures: fractional anisotropy\footnote[frame]{Basser ''Microstructural and physiological features of tissues elucidated by quantitative-diffusion-tensor MRI", 1996}, anisotropy coefficients\footnote[frame]{Westin et al.
+''Geometrical
+diffusion measures for MRI from tensor basis analysis", 2002}
+
 \end{itemize}
 \end{small}
-\end{column}
-\begin{column}{.5\textwidth}
-\begin{figure}[t]
-\caption*{\textit{Isotropic test field}}\bigskip
-\begin{minipage}{.49\textwidth}
-\includegraphics[height=\textwidth]{isotropic2.png} \caption*{a)~Glyphs}
-\end{minipage}
-\begin{minipage}{.49\textwidth}
-\includegraphics[height=\textwidth]{iso_ftle.png} \caption*{b)~LTG($0^\circ$)}
-\end{minipage}
-
-\end{figure}
-
-\end{column}
-\end{columns}
 

### 2019-08-11 beamer2thesis/dummy.tex
update

--- 2019-08-09 beamer2thesis/dummy.tex
@@
 \frame{
-\frametitle{{Light Transport Gradient (LTG) - Isotropic field}}
+\frametitle{{Fundamentals - Finite-Time Lyapunov Exponents (FTLE)}}
+
 \begin{columns}
@@
 \begin{small}
+\begin{block}{\centering \textbf{Definition}}
+Local: 
+$
+	\mathop{FTLE} = \frac{1}{\lvert T\rvert}\ln \frac{\Delta}{\delta}
+$ (cf. Fig. b))\\
+Global: 
+$
+	\mathop{FTLE}(\mathbf{x})=\frac{1}{\lvert T\rvert}\ln \|\nabla\mathbf{u(x)}\|_2
+$\\
+{\tiny where $\| A\| _2 = \sqrt{\lambda_{\max}(A^TA)}$ is the spectral norm of matrix $A$}
+\end{block}
+\end{small}
 \begin{itemize}
-	\item Isotropic test field reveals offset response for the approach
-	\bigskip
-	\item natural shift against the inital light source (slice) direction (left)\\
-%	$\Rightarrow$ initial light source configurations located close and pointing towards borders, are biased for low LTG (FTLE) responses
+	\item measure for separation abilities of time-variant systems
+	\item Lagrangian view in vector fields: placing particle, moving with flow
 \end{itemize}
-\end{small}
+
 \end{column}
@@
 \begin{figure}[t]
-\caption*{\textit{Isotropic test field}}\bigskip
-\begin{minipage}{.49\textwidth}
-\includegraphics[height=\textwidth]{isotropic2.png} \caption*{a)~Glyphs}
+\centering
+\begin{minipage}{0.85\textwidth}
+\centering
+\includegraphics[height=0.6\textwidth]{diverge_rot.png} {\scriptsize a)}
+\includegraphics[height=0.6\textwidth]{ftle_sep.pdf} {\scriptsize b)}
+\caption*{a) Vector field, b) Diverging pathlines}
 \end{minipage}
-\begin{minipage}{.49\textwidth}
-\includegraphics[height=\textwidth]{iso_ftle.png} \caption*{b)~LTG($0^\circ$)}
+\begin{minipage}{0.1\textwidth}
+\source{b) Skript Prof. Sadlo, SciVis SS2017}
 \end{minipage}
@@
 \end{columns}
-
 } % END OF FRAME

### 2019-08-12 beamer2thesis/dummy.tex
upddate

--- 2019-08-09 beamer2thesis/dummy.tex
@@
 \frame{
-\frametitle{{Light Transport Gradient (LTG) - Isotropic field}}
-\begin{columns}
-\begin{column}{.5\textwidth}
-\begin{small}
+\frametitle{{Transmission Profiles}}
+{\small To interpret singular value systems of tensors as conductivity property for light transport, we will need transfer functions (transmission profiles) for the propagation scheme:}
+\begin{block}{\centering \textbf{Ellipse Equation}}
+\begin{align*}
+	T(\omega) = \frac{ab}{\sqrt{a^{2}\sin^{2}(\omega-\varphi)+b^{2}\cos^{2}(\omega-\varphi)}} =\frac{\sigma_1\sigma_2}{\sqrt{\sigma_1^2\sin^2(\omega-\varphi)+\sigma_2^2\cos^2(\omega-\varphi)}}
+\end{align*}
+{\small with $\varphi=\mathop{atan2}(\sigma_{1,y},\sigma_{1,x})$ and $\sigma$: singular value, $a$: x-radius ellipse, $b$: y-radius ellipse}
+\end{block}
+$\Rightarrow$ transmission profiles $r(\omega)$ are defined as polar functions by the mapping of singular values to half-axes radii and the shift angle of the major singular vector $\varphi$
+
+} % END OF FRAME
+
+\frame{
+\frametitle{{Related Work - Asymmetric Tensor Field Visualization}}
+
 \begin{itemize}
-	\item Isotropic test field reveals offset response for the approach
+	\item dual eigenvectors\footnote[frame]{Zheng and Pang ''2d asymmetric tensor analysis", 2005}: use complex conjugate eigenvectors as co-visualization for the complex domain along with ordinary eigenvectors to represent the real domain
 	\bigskip
-	\item natural shift against the inital light source (slice) direction (left)\\
-%	$\Rightarrow$ initial light source configurations located close and pointing towards borders, are biased for low LTG (FTLE) responses
+	\item pseudo eigenvectors\footnote[frame]{Laramee et al. ''2d asymmetric tensor field topology", 2012}: extension for dual eigenvectors to a full set or graph
+	\bigskip
+	\item scalar measures: tensor magnitude\footnote[frame]{Lin et al.
+''Asymmetric tensor field visualization for surfaces", 2011}, tensor mode\footnote[frame]{Palacios et al.
+''Feature surfaces in symmetric tensor fields based on eigenvalue
+manifold", 2015}, isotropy index\footnote[frame]{see footnote 12}
 \end{itemize}
-\end{small}
-\end{column}
-\begin{column}{.5\textwidth}
-\begin{figure}[t]
-\caption*{\textit{Isotropic test field}}\bigskip
-\begin{minipage}{.49\textwidth}
-\includegraphics[height=\textwidth]{isotropic2.png} \caption*{a)~Glyphs}
-\end{minipage}
-\begin{minipage}{.49\textwidth}
-\includegraphics[height=\textwidth]{iso_ftle.png} \caption*{b)~LTG($0^\circ$)}
-\end{minipage}
-
-\end{figure}
-
-\end{column}
-\end{columns}
 

### 2019-08-13 beamer2thesis/dummy.tex
update

@@
 \frame{
-\frametitle{{Transmission Profiles}}
-{\small To interpret singular value systems of tensors as conductivity property for light transport, we will need transfer functions (transmission profiles) for the propagation scheme:}
-\begin{block}{\centering \textbf{Ellipse Equation}}
-\begin{align*}
-	T(\omega) = \frac{ab}{\sqrt{a^{2}\sin^{2}(\omega-\varphi)+b^{2}\cos^{2}(\omega-\varphi)}} =\frac{\sigma_1\sigma_2}{\sqrt{\sigma_1^2\sin^2(\omega-\varphi)+\sigma_2^2\cos^2(\omega-\varphi)}}
-\end{align*}
-{\small with $\varphi=\mathop{atan2}(\sigma_{1,y},\sigma_{1,x})$ and $\sigma$: singular value, $a$: x-radius ellipse, $b$: y-radius ellipse}
-\end{block}
-$\Rightarrow$ transmission profiles $r(\omega)$ are defined as polar functions by the mapping of singular values to half-axes radii and the shift angle of the major singular vector $\varphi$
-
+\frametitle{{Evaluation - Light Transport Gradient - Gyre Field}}
+\begin{figure}[!t]
+\centering
+\caption*{\textit{Gyre test field}}\bigskip
+  \begin{minipage}[t]{0.3\textwidth}
+    \centering
+  \includegraphics[height=0.7\textwidth]{img/gyre.png}
+	{\scriptsize a)~Glyphs}
+    \label{a)}
+  \end{minipage}
+  \begin{minipage}[t]{0.3\textwidth}
+    \centering
+    \includegraphics[height=0.7\textwidth]{gyre_ftle.PNG}
+	{\scriptsize b)~LTG($0^\circ$)}
+    \label{b)}
+  \end{minipage}
+  \begin{minipage}[t]{0.3\textwidth}
+    \centering
+    \includegraphics[height=0.7\textwidth]{gyre_ftle_org.PNG}
+	{\scriptsize c)~}{\tiny T-FTLE: Hlawatsch's approach}
+    \label{b)}
+  \end{minipage}
+\label{gyre-ftle}
+\end{figure}
+\bigskip
+\begin{itemize}
+	\item LTG is detecting ridges for diverging light distributions
+	\item decreased response in gyre centers for circular overlap in difference image
+\end{itemize}
 } % END OF FRAME
